### 2017-03-30 doc/RMD/data-comparison.Rmd
add makefile

---
title: "R Notebook"
output: html_notebook
---

This document serves to compare the data presented in Sydney's original XLSX 
file and the one generated by Amaradasa. He cleaned it up by inserting "unk" for
missing values. Because there was no file trail for this other than his own
comment, I will compare these two.

```{r setup, include = FALSE}
knitr::opts_knit$set(root.dir = "../../")
```

```{r}
library('tidyverse')
library('readxl')
```

Sydney's data is stored in excel format, so we have to use readxl to parse it.
First, we want to know what sheets exist.

```{r}
excel_sheets("../Analysis4 ForManu/A1_Copy of binned-genotypes_SE.xlsx")
```

We will be using the GenAlexBinned sheet. From my talks with Sydney, this sheet
contains the SSR data binned into the expected allelels. Note, this sequence of
data input and cleaning was done iteratively and what you are seeing is the 
final result.

There are always quirks with the data when it's in excel. Often, there are too
many rows, so we have to remove them with `slice()`.

Quirks specific to thse data:

 - GenAlEx format has an extra header region that should be ignored
 - The loci are formatted like so: `110-4(F)`, where the `(F)` part is not 
   informative
 - The header for the strata has extra information

```{r sydney_data}
syd <- read_excel("../Analysis4 ForManu/A1_Copy of binned-genotypes_SE.xlsx", 
                sheet = "GenAlexBinned", skip = 1) %>%
  select(-1) %>%                # removing first column, which is empty
  gather(locus, allele, -1) %>% # gather all loci into tidy columns
  mutate(locus = trimws(locus) %>% substr(1, nchar(.) - 3)) %>% # remove (F) designator
  spread(locus, allele) %>%     # spread data out with individual loci in columns
  separate(iso_st_mcg_org_loc_yr_hst_cult_rep, 
           c("Isolate", "Severity", "MCG", "State", "Source", "Year", "Host"), 
           sep = "_") %>%
  mutate_if(is.character, trimws) %>% # Trim whitespace forom all character columns
  mutate(Severity = as.numeric(Severity)) %>%
  slice(-n()) # remove last row
syd
```

### 2017-03-30 doc/RMD/data-comparison.Rmd
update analysis

@@
 This document serves to compare the data presented in Sydney's original XLSX 
-file and the one generated by Amaradasa. He cleaned it up by inserting "unk" for
+file and the one generated by Sajeewa. He cleaned it up by inserting "unk" for
 missing values. Because there was no file trail for this other than his own
@@
 ```{r setup, include = FALSE}
-knitr::opts_knit$set(root.dir = "../../")
+knitr::opts_knit$set(root.dir = PROJHOME)
 ```
@@
 library('readxl')
+library('assertr')
 ```
+
+## Assertations
+
+Of course, to make sure the data are cromulent, we want to ensure they live up to
+our standards of what has been reported. For this, we will use the *assertr*
+package, which lets us check our data in various ways. I'm taking the data
+presented in the paper thus far to create the assertions.
+
+```{r}
+size_ranges <- 
+"locus\trange\tn
+5-2	318-324	4
+6-2	483-495	3
+7-2	158-174	7
+8-3	244-270	7
+9-2	360-382	9
+12-2	214-222	5
+17-3	342-363	7
+20-3	280-282	2
+55-4	153-216	10
+110-4	370-386	5
+114-4	339-416	10
+"
+sr <- read_tsv(size_ranges) %>%
+  separate(range, c("lower", "upper"), sep = "-") # split range into lower and upper bounds
+sr
+# Making the assertation string.
+# To do this, we are using `sprintf()` to print the assert statement:
+#   assert(within_bounds(lower, upper), `locus`)
+# which means: "assert that the alleles (data) within this locus (variable) are
+# within the defined range". 
+# 
+# The `apply()` function takes a matrix or data frame and applies a given function over each
+# row (MARGIN = 1) or column (MARGIN = 2). In this case, we are specifying rows.
+# 
+# We then collapse all these statements with the pipe operator: %>%
+# Then we add the name of the operation to the beginning
+assrt_string <- apply(sr, MARGIN = 1, function(x){
+    sprintf("  assert(within_bounds(%s, %s), `%s`)", x[2], x[3], x[1]) 
+  }) %>% 
+  paste(collapse = " %>%\n") %>%
+  paste("assert_alleles_within_bounds <- .", ., sep = " %>%\n")
+#
+# now we can see what this statement looks like
+cat(assrt_string)
+# here, we evaluate it, making it available for us later on.
+eval(parse(text = assrt_string))
+
+YEARS <- c("2003", "2004", "2005", "2007", "2008", "2009", "2010", "2012")
+
+STATE <- c("AU", "TS", "FR", "BL", "MX", "NE", "NY", "MN", "MI", "OR", 
+           "WA", "CO", "WI", "ID", "CA", "ND")
+
+HOST <- c("Merlot", "Pinto", "redkid", "Beryl", "Bunsi", "37", "38", 
+          "11A", "cornell", "G122", "Orion", "PO7104", "PO7863", "WM31", 
+          "GH", "BO7104", "Black", "Vista", "SR06233", "BL", "Fuji", "unk", 
+          "Zorro", "PO7883", "Emerson", "Weihing", "Yellow")
+
+check_data_cromulence <- . %>%
+  chain_start() %>%
+  verify(nrow(.) == 366) %>%          # 366 isolates
+  verify(has_all_names(sr$locus)) %>% # Has all loci
+  assert_alleles_within_bounds %>%    # each locus has alleles within bounds
+  assert(in_set(YEARS), Year) %>%     # has all specified years
+  assert(in_set(STATE), State) %>%    # has all specified States (incl. countries)
+  assert(in_set(HOST), Host) %>%      # has all specified Hosts (incl. countries)
+  chain_end(error_fun = warn_report)
+```
+
+## Reading in Sydney's data
 
@@
   mutate(locus = trimws(locus) %>% substr(1, nchar(.) - 3)) %>% # remove (F) designator
+  mutate(allele = as.integer(allele)) %>% # force alleles to integers
   spread(locus, allele) %>%     # spread data out with individual loci in columns
@@
   mutate(Severity = as.numeric(Severity)) %>%
-  slice(-n()) # remove last row
+  mutate(Source = ifelse(Source == "", "unk", Source)) %>%
+  slice(-n()) %>% # remove last row
+  arrange(Isolate) %>%
+  check_data_cromulence
 syd
 ```
+
+## Reading in Sajeewa's data
+
+These data can be read in via `reader::read_csv()`
+
+```{r sajeewa_data}
+column_specification <- cols(
+  Individual = col_character(),
+  iso_st_mcg_org_loc_yr_hst = col_character(),
+  `5-2(F)` = col_integer(),
+  `6-2(F)` = col_integer(),
+  `7-2(F)` = col_integer(),
+  `8-3(H)` = col_integer(),
+  `9-2(F)` = col_integer(),
+  `12-2(H)` = col_integer(),
+  `17-3(H)` = col_integer(),
+  `20-3(F)` = col_integer(),
+  `55-4(F)` = col_integer(),
+  `110-4(H)` = col_integer(),
+  `114-4(H)` = col_integer()
+)
+saj <- read_csv("../Analysis4 ForManu/A2_Copy4 EUR_AUS_forManu.csv", skip = 2, col_types = column_specification) %>%
+  select(-1) %>%
+  gather(locus, allele, -1) %>% # gather all loci into tidy columns
+  mutate(locus = trimws(locus) %>% substr(1, nchar(.) - 3)) %>% # remove (F) designator
+  spread(locus, allele) %>%     # spread data out with individual loci in columns
+  separate(iso_st_mcg_org_loc_yr_hst, # Note: this corresponds closer to the initial data
+           c("Isolate", "Severity", "MCG", "State", "Source", "Year", "Host"), 
+           sep = "_") %>%
+  mutate_if(is.character, trimws) %>% # Trim whitespace forom all character columns
+  mutate(Severity = as.numeric(Severity)) %>%
+  arrange(Isolate) %>%
+  check_data_cromulence
+saj
+```
+
+## Comparison
+
+We've sorted each data set by Isolate, so that means that we should expect them
+to have the same data. The `setequal()` checks whether or not both data sets
+have the same rows (in any order)
+
+```{r, equal_sets}
+setequal(saj, syd)
+```
+
+Okay, something's not cromulent here. We'll have to manaully inspect these:
+
+```{r}
+syddiff <- setdiff(syd, saj)
+sajdiff <- setdiff(saj, syd)
+the_difference <- bind_rows(syd = syddiff, saj = sajdiff, .id = "source") %>% arrange(Isolate)
+head(the_difference, n = 20)
+```
+
+It appears that there are a couple of discrepancies:
+
+1. Isolate 805 is flagged for some reason in Sydney's data set
+2. Tasmania (TS, Sydney) has been changed to Australia (AU, Sajeewa)
+3. Belgium (BL, Sydney) has been changed to France (FR, Sajeewa)
+
+## Session Information
+
+```{r session_info}
+options(width = 100)
+devtools::session_info()
+```

### 2017-03-30 doc/RMD/data-comparison.Rmd
rerun with dplyr specified

@@
 ```{r, equal_sets}
-setequal(saj, syd)
+dplyr::setequal(saj, syd)
 ```
@@
 ```{r}
-syddiff <- setdiff(syd, saj)
-sajdiff <- setdiff(saj, syd)
+syddiff <- dplyr::setdiff(syd, saj)
+sajdiff <- dplyr::setdiff(saj, syd)
 the_difference <- bind_rows(syd = syddiff, saj = sajdiff, .id = "source") %>% arrange(Isolate)

### 2017-03-30 doc/RMD/data-comparison.Rmd
update conclusion

@@
 title: "R Notebook"
-output: html_notebook
+output: 
+  html_notebook:
+    toc: true
 ---
@@
 
-It appears that there are a couple of discrepancies:
+## Conclusion
+
+Both data sets failed the cromulence check in a common manner:
+
+```
+- Column 'Host' violates assertion 'in_set(HOST)' 2 times
+  index  value
+1   263 B07104
+2   317 PO7683
+```
+
+These are the values I grabbed from the manuscript that could possibly match:
+
+ - `PO7863`
+ - `PO7883`
+ - `BO7104` (The difference is that this is the letter "O" and above is the number "0")
+
+Sydney's data set had an extra discrepancy, but that's noted in the difference
+between data sets below:
 
@@
 3. Belgium (BL, Sydney) has been changed to France (FR, Sajeewa)
+4. Host for isolate 499 has been changed from ExRico (Sydney) to Bunsi (Sajeewa)
 

### 2017-03-30 doc/RMD/data-comparison.Rmd
update comparison

@@
 
+### Sydney's comments
+
+We have no clue why 805 is flagged. The changes to the metadata are correct.
+The change from BL to FR was because of confusion between the source of the 
+collectors vs. source of isolate.
+
+TS was grouped into AU because it was too small of a sample size to do anything.
+
+From here on out, we will use Sajeewa's data for analysis.
+
 ## Session Information

### 2017-03-30 doc/RMD/MLG-distribution.Rmd
add mlg-distribution and run make

---
title: "R Notebook"
output: 
  html_notebook:
    toc: true
---

```{r setup, include = FALSE}
knitr::opts_knit$set(root.dir = PROJHOME)
```

In this document, I will create a graph that shows the distribution of MLGs
across populations

```{r}
library('tidyverse')
library('assertr')
library('poppr')
library('igraph')
```

## Loading data and assertions

```{r read_data}
dat <- read.genalex("../Analysis4 ForManu/A2_Copy4 EUR_AUS_forManu.csv", ploidy = 1)
splitStrata(dat) <- ~Isolate/Severity/MCG/Region/Source/Year/Host
dat
```

The incoming strata includes both Severity and Isolate. Since these are not
necessary for delimiting the strata, we will place them in the "other" slot
after converting Severity to numeric. Placing this information in the "other"
slot ensures that these data will travel with the object.

```{r fix_strata}
dat_strata <- strata(dat) %>%
  mutate_all(as.character) %>%
  mutate(Severity = as.numeric(Severity))
strata(dat)     <- select(dat_strata, -Severity, -Isolate)
indNames(dat)   <- dat_strata$Isolate
other(dat)$meta <- select(dat_strata, Severity, Isolate)
```

```{r}
setPop(dat) <- ~Region
dat
```

## Crossing populations

We can use `mlg.crosspop()` to tabulte which MLGs cross populations.

```{r}
make_from_to <- function(poplist){
  x <- combn(poplist, 2)
  tibble::data_frame(from = x[1, ], to = x[2, ])
}
crosses <- mlg.crosspop(dat, df = TRUE, quiet = TRUE)

pop_graph <- crosses %>%
  group_by(MLG) %>% 
  summarize(Population = list(make_from_to(Population)), Size = sum(Count)) %>%
  unnest() %>%
  group_by(from, to) %>%
  summarize(weight = length(Size)) %>%
  ungroup()

gcross <- graph_from_data_frame(pop_graph, directed = FALSE)
traffic <- apply(gcross[], 1, sum)
# V(gcross)$weight <- table(pop(clonecorrect(dat, ~Region)))[V(gcross)]
colors <- viridis::magma(max(pop_graph$weight), end = 0.9)
set.seed(50)
gcross %>%
  plot(., vertex.size = table(pop(clonecorrect(dat, ~Region)))[names(V(gcross))],
       edge.width = E(.)$weight, 
       layout = layout_in_circle(.),
       edge.color = colors[E(.)$weight])

```

### 2017-03-31 doc/RMD/MLG-distribution.Rmd
update graphing

@@
 
-pop_graph <- crosses %>%
+pg <- 
+  crosses %>%
   group_by(MLG) %>% 
-  summarize(Population = list(make_from_to(Population)), Size = sum(Count)) %>%
-  unnest() %>%
+  summarize(Population = list(make_from_to(Population)), Size = length(Count))
+
+pop_graph <- pg %>%  unnest() %>%
   group_by(from, to) %>%
-  summarize(weight = length(Size)) %>%
-  ungroup()
+  mutate(weight = length(Size)) %>%
+  ungroup() %>%
+  select(from, to, weight, everything())
+
+unchoose <- function(x) ceiling(sqrt(x * 2))
 
 gcross           <- graph_from_data_frame(pop_graph, directed = FALSE)
-traffic <- apply(gcross[], 1, sum)
-# V(gcross)$weight <- table(pop(clonecorrect(dat, ~Region)))[V(gcross)]
-colors <- viridis::magma(max(pop_graph$weight), end = 0.9)
+E(gcross)$name   <- as.character(pop_graph$MLG)
+E(gcross)$weight <- unchoose(table(E(gcross)$name)[E(gcross)$name])
+colors <- viridis::magma(max(E(gcross)$weight) - 1, end = 0.8)
 set.seed(50)
@@
   plot(., vertex.size = table(pop(clonecorrect(dat, ~Region)))[names(V(gcross))],
-       edge.width = E(.)$weight, 
-       layout = layout_in_circle(.),
-       edge.color = colors[E(.)$weight])
+       edge.width = E(.)$weight - 1,
+       layout = layout_as_star(., center = "MI"),
+       # layout = layout_in_circle(.),
+       edge.color = colors[E(.)$weight - 1])
 
 ```
+
+```{r}
+
+pops <- crosses %>%
+  group_by(Population) %>%
+  summarize(MLG = list(as.character(MLG))) %>%
+  group_by(Population) %>%
+  mutate(interactions = list(pg$Population[pg$MLG %in% unlist(MLG)] %>% setNames(unlist(MLG)) %>% bind_rows(.id = "MLG") %>% distinct)) %>%
+  mutate(graph = map(interactions, ~select(.x, from, to, MLG) %>% graph_from_data_frame(directed = FALSE)))
+psize <- table(pop(clonecorrect(dat, ~Region)))
+par(mfrow = c(3, 4))
+
+apply(pops, 1, function(i){
+    plot(i$graph, 
+         layout = layout_as_star(i$graph, center = i$Population), 
+         main = i$Population, 
+         vertex.size = psize[names(V(i$graph))])
+  })
+par(mfrow = c(1, 1))
+do.call("union", pops$graph) %>% 
+  plot(., vertex.size = psize[names(V(.))],
+       layout = layout_as_star(., center = "MI"))
+```

### 2017-03-31 doc/RMD/MLG-distribution.Rmd
update friggen final graph

@@
 
-```{r}
+```{r, fig.width = 10, fig.height = 10}
 make_from_to <- function(poplist){
@@
 
-```{r}
+```{r, fig.width = 10, fig.height = 10}
 
@@
 par(mfrow = c(1, 1))
-do.call("union", pops$graph) %>% 
+
+igraph::union(pops$graph[[1]], pops$graph[-1]) %>% 
   plot(., vertex.size = psize[names(V(.))],

### 2017-03-31 doc/RMD/MLG-distribution.Rmd
Add cool new version of graph

@@
 colors <- viridis::magma(max(E(gcross)$weight) - 1, end = 0.8)
-set.seed(50)
+set.seed(500)
 gcross %>%
@@
 ```{r, fig.width = 10, fig.height = 10}
-
+psize <- table(pop(clonecorrect(dat, ~Region)))
 pops <- crosses %>%
+  mutate_if(is.factor, as.character) %>%
   group_by(Population) %>%
@@
   group_by(Population) %>%
-  mutate(interactions = list(pg$Population[pg$MLG %in% unlist(MLG)] %>% setNames(unlist(MLG)) %>% bind_rows(.id = "MLG") %>% distinct)) %>%
-  mutate(graph = map(interactions, ~select(.x, from, to, MLG) %>% graph_from_data_frame(directed = FALSE)))
-psize <- table(pop(clonecorrect(dat, ~Region)))
+  mutate(size = psize[unlist(Population)]) %>%
+  mutate(n_private = size - lengths(MLG)) %>%
+  mutate(interactions = list(pg$Population[pg$MLG %in% unlist(MLG)] %>% 
+                               setNames(unlist(MLG)) %>% 
+                               bind_rows(.id = "MLG") %>% 
+                               distinct() %>%
+                               filter(Population == from | Population == to))) %>%
+  mutate(graph = map(interactions, ~select(.x, from, to, MLG) %>% 
+                       rename(name = MLG) %>% 
+                       graph_from_data_frame(directed = FALSE)))
+pops
 par(mfrow = c(3, 4))
@@
 
-igraph::union(pops$graph[[1]], pops$graph[-1]) %>% 
-  plot(., vertex.size = psize[names(V(.))],
-       layout = layout_as_star(., center = "MI"))
+gnew <- igraph::union(pops$graph[[1]], pops$graph[-1])
+
+glayout <- layout_as_star(gnew, center = "MI")
+
+glayout_df <- glayout %>% 
+  data.frame %>% 
+  setNames(c("x", "y")) %>% 
+  as.list %>% 
+  c(Population = list(names(V(gnew)))) %>% 
+  dplyr::as_data_frame() %>%
+  inner_join(pops)
+V(gnew)$size <- glayout_df$size
+V(gnew)$color <- viridis::viridis(100)[round(100 * (1 - glayout_df$n_private/glayout_df$size))]
+gnew <- add_vertices(gnew, length(V(gnew)), size = glayout_df$n_private, color = "grey90")
+glayout2 <- rbind(glayout, glayout)
+plot(gnew, layout = glayout2)
+
 ```

### 2017-03-31 doc/RMD/MLG-distribution.Rmd
add ggraph

@@
 library('igraph')
+library("ggraph")
 ```
@@
 pops
+
+possible_edges <- expand.grid(from = pops$Population, to = pops$Population)
+possible_edges$Population <- possible_edges$from
+from_edges <- right_join(possible_edges, select(pops, Population, size)) %>% 
+  tbl_df %>% 
+  rename(from_size = size) %>%
+  select(-Population)
+to_edges <- right_join(mutate(possible_edges, Population = to), select(pops, Population, size)) %>% 
+  tbl_df %>% 
+  rename(to_size = size) %>%
+  select(-Population)
+
+possible_edges <- inner_join(from_edges, to_edges)
+
 par(mfrow = c(3, 4))
@@
 V(gnew)$size <- glayout_df$size
-V(gnew)$color <- viridis::viridis(100)[round(100 * (1 - glayout_df$n_private/glayout_df$size))]
+V(gnew)$weight <- (1 - glayout_df$n_private/glayout_df$size)
+V(gnew)$color <- viridis::viridis(100)[round(V(gnew)$weight * 100)]
 gnew <- add_vertices(gnew, length(V(gnew)), size = glayout_df$n_private, color = "grey90")
@@
 
+caps <- as_data_frame(gnew, what = "edges") %>% select(from, to) %>% 
+  inner_join(possible_edges) %>% tbl_df
+
+caps %>% mutate(from_size = scale(from_size, center = FALSE)/10) %>%
+  mutate(to_size = scale(to_size, center = FALSE)/10) -> spac
+
+glay <- create_layout(gnew, "manual", node.positions = as.data.frame(glayout2) %>% setNames(c("x", "y")))
+ggraph(glay) +
+  geom_node_circle(aes(r = scale(size, center = FALSE)/10, fill = size, alpha = 1 - weight)) +
+  geom_edge_fan() +
+  # geom_edge_fan2(aes(start_cap = circle(rep(spac$from_size, 2)/2, "mm"),
+  #                   end_cap = circle(rep(spac$to_size, 2)/2, "mm"))) +
+  geom_node_label(aes(label = name), repel = TRUE) +
+  viridis::scale_fill_viridis(option = "C") +
+  coord_fixed() +
+  theme_void() +
+  labs(list(
+    title = "Shared haplotypes across regions",
+    fill = "Number of\nGenotypes",
+    alpha = "Fraction of private genotpes",
+    caption = "Outer circle: Number of genotypes in the region\nInner Circle: Number of private genotypes in the region"
+  ))
 ```

### 2017-03-31 doc/RMD/MLG-distribution.Rmd
whoops. place circle in front of edge

@@
 ggraph(glay) +
-  geom_node_circle(aes(r = scale(size, center = FALSE)/10, fill = size, alpha = 1 - weight)) +
   geom_edge_fan() +
@@
   #                   end_cap = circle(rep(spac$to_size, 2)/2, "mm"))) +
+  geom_node_circle(aes(r = scale(size, center = FALSE)/10, fill = size, alpha = 1 - weight)) +
   geom_node_label(aes(label = name), repel = TRUE) +

### 2017-04-01 doc/RMD/MLG-distribution.Rmd
Rework MLG-graph

--- 2017-03-31 doc/RMD/MLG-distribution.Rmd
@@
 
-```{r, fig.width = 10, fig.height = 10}
-make_from_to <- function(poplist){
-  x <- combn(poplist, 2)
-  tibble::data_frame(from = x[1, ], to = x[2, ])
-}
-crosses <- mlg.crosspop(dat, df = TRUE, quiet = TRUE)
+I realized that it's possible to use an MLG table with matrix multiplication to
+get an adjency matrix.
 
-pg <- 
-  crosses %>%
-  group_by(MLG) %>% 
-  summarize(Population = list(make_from_to(Population)), Size = length(Count))
-
-pop_graph <- pg %>%  unnest() %>%
-  group_by(from, to) %>%
-  mutate(weight = length(Size)) %>%
-  ungroup() %>%
-  select(from, to, weight, everything())
-
-unchoose <- function(x) ceiling(sqrt(x * 2))
-
-gcross           <- graph_from_data_frame(pop_graph, directed = FALSE)
-E(gcross)$name   <- as.character(pop_graph$MLG)
-E(gcross)$weight <- unchoose(table(E(gcross)$name)[E(gcross)$name])
-colors <- viridis::magma(max(E(gcross)$weight) - 1, end = 0.8)
-set.seed(500)
-gcross %>%
-  plot(., vertex.size = table(pop(clonecorrect(dat, ~Region)))[names(V(gcross))],
-       edge.width = E(.)$weight - 1,
-       layout = layout_as_star(., center = "MI"),
-       # layout = layout_in_circle(.),
-       edge.color = colors[E(.)$weight - 1])
-
+```{r}
+datmlg  <- mlg.table(dat, plot = FALSE) > 0 # presence/absence of MLG
+crosses <- mlg.crosspop(dat, quiet = TRUE, df = TRUE) %>% tbl_df()
+adjmat  <- datmlg %*% t(datmlg) 
+cols    <- sort(colnames(adjmat))
+adjmat  <- adjmat[cols, cols]
+adjmat
+crosses
 ```
 
+Now that we have the adjacency matrix, we can use it to construct our graph:
+
 ```{r, fig.width = 10, fig.height = 10}
-psize <- table(pop(clonecorrect(dat, ~Region)))
-pops <- crosses %>%
-  mutate_if(is.factor, as.character) %>%
-  group_by(Population) %>%
-  summarize(MLG = list(as.character(MLG))) %>%
-  group_by(Population) %>%
-  mutate(size = psize[unlist(Population)]) %>%
-  mutate(n_private = size - lengths(MLG)) %>%
-  mutate(interactions = list(pg$Population[pg$MLG %in% unlist(MLG)] %>% 
-                               setNames(unlist(MLG)) %>% 
-                               bind_rows(.id = "MLG") %>% 
-                               distinct() %>%
-                               filter(Population == from | Population == to))) %>%
-  mutate(graph = map(interactions, ~select(.x, from, to, MLG) %>% 
-                       rename(name = MLG) %>% 
-                       graph_from_data_frame(directed = FALSE)))
-pops
+g           <- graph_from_adjacency_matrix(adjmat, mode = "undirected", diag = FALSE)
+V(g)$size   <- diag(adjmat)
+g           <- delete_vertices(g, degree(g) == 0)
+shared_mlg  <- (crosses %>% group_by(Population) %>% summarize(n = n()))$n
+V(g)$weight <- 1 - shared_mlg/V(g)$size
+el          <- as_adj_edge_list(g)
+el          <- el[lengths(el) > 0]
+popgraphs <- setNames(vector(mode = "list", length = length(el)), names(el))
+for (v in names(el)){
+  idx  <- el[[v]]
+  mlgs <- crosses %>%           # How to get all MLGs from a single population:
+    filter(Population == v) %>%         # Grab only the population e and then
+    select(MLG) %>%                     # remove everything but the MLGs to do an
+    inner_join(crosses, by = "MLG") %>% # inner join of the original list and then
+    filter(Population != v) %>%         # remove the query population to give
+    arrange(Population)                 # the neigboring populations in order.
+  MLGS <- as.character(mlgs$MLG)
+  E(g)[idx]$label  <- substr(MLGS, 5, nchar(MLGS))
+  E(g)[idx]$weight <- as.integer(table(MLGS)[MLGS]) # weight == n populations visited
+  popgraphs[[v]]   <- subgraph.edges(g, eids = idx)
+}
 par(mfrow = c(3, 4))
+for (i in names(popgraphs)){
+  pg <- popgraphs[[i]]
+  labs <- ifelse(E(pg)$weight > 1, E(pg)$label, NA)
+  labs <- ifelse(duplicated(labs), NA, labs)
+  plot(pg, 
+       main = i, 
+       layout = layout_as_star(pg, center = i), 
+       edge.width = E(pg)$weight,
+       edge.label = labs)
+}
+par(mfrow = c(1, 1))
+center_node <- degree(g) %>% which.max() %>% names()
 
-apply(pops, 1, function(i){
-    plot(i$graph, 
-         layout = layout_as_star(i$graph, center = i$Population), 
-         main = i$Population, 
-         vertex.size = psize[names(V(i$graph))])
-  })
-par(mfrow = c(1, 1))
+new_layout <- structure(c(1, 0.809016994374947, 0.309016994374947, -0.309016994374947, 
+-0.809016994374947, -1, -0.809016994374947, -0.309016994374948, 
+0, 0.309016994374947, 0.809016994374947, 0, 0.587785252292473, 
+0.951056516295154, 0.951056516295154, 0.587785252292473, 1.22464679914735e-16, 
+-0.587785252292473, -0.951056516295154, 0, -0.951056516295154, 
+-0.587785252292473), .Dim = c(11L, 2L), .Dimnames = list(c("MN", 
+"AU", "OR", "FR", "WA", "CA", "NE", "CO", "MI", "ND", "NY"), 
+    c("x", "y")))
+new_layout <- new_layout[V(g)$name, ]
+g <- add_vertices(g, length(V(g)), size = V(g)$size - shared_mlg, color = "grey90")
+plot(g, 
+     layout = rbind(new_layout, new_layout),
+     edge.width = E(g)$weight,
+     edge.label = NA)
 
-gnew <- igraph::union(pops$graph[[1]], pops$graph[-1])
-
-glayout <- layout_as_star(gnew, center = "MI")
-
-glayout_df <- glayout %>% 
-  data.frame %>% 
-  setNames(c("x", "y")) %>% 
-  as.list %>% 
-  c(Population = list(names(V(gnew)))) %>% 
-  dplyr::as_data_frame() %>%
-  inner_join(pops)
-V(gnew)$size <- glayout_df$size
-V(gnew)$color <- viridis::viridis(100)[round(100 * (1 - glayout_df$n_private/glayout_df$size))]
-gnew <- add_vertices(gnew, length(V(gnew)), size = glayout_df$n_private, color = "grey90")
-glayout2 <- rbind(glayout, glayout)
-plot(gnew, layout = glayout2)
-
+glay <- create_layout(g, "manual", node.positions = as.data.frame(rbind(new_layout, new_layout)))
+ggraph(glay) +
+  geom_edge_fan(aes(alpha = weight + 1)) +
+  geom_node_circle(aes(r = scale(size, center = FALSE)/10, fill = size, alpha = weight)) +
+  geom_node_label(aes(label = name), repel = TRUE) +
+  viridis::scale_fill_viridis(option = "C") +
+  coord_fixed() +
+  theme_void() +
+  labs(list(
+    title = "Shared haplotypes across regions",
+    fill = "Number of\nGenotypes",
+    alpha = "Fraction of\nprivate genotpes",
+    edge_alpha = "Populations\nper haploytpe",
+    caption = "Outer circle: Number of haplotypes in the region\nInner Circle: Number of private haplotypes in the region"
+  ))
 ```

### 2017-04-03 doc/RMD/MLG-distribution.Rmd
massive update to plots. These looks quite good.

@@
 library("ggraph")
+library('magrittr')
 ```
@@
 dat
+
+repeat_lengths <-
+  c(
+  `5-2` = 2.000000,
+  `5-3` = 0.500000,
+  `6-2` = 6.000000,
+  `7-2` = 2.000000,
+  `8-3` = 2.000000,
+  `9-2` = 2.000000,
+  `12-2` = 2.000000,
+  `17-3` = 3.000000,
+  `20-3` = 2.000000,
+  `36-4` = 0.500000,
+  `50-4` = 0.500000,
+  `55-4` = 4.000000,
+  `92-4` = 2.000000,
+  `106-4` = 4.000000,
+  `110-4` = 4.000000,
+  `114-4` = 4.000000
+  )
+repeat_lengths <- ifelse(repeat_lengths < 1, 4, repeat_lengths)
+```
+
+```{r excel_data}
+ex <- readxl::read_excel("../Analysis4 ForManu/A1_Copy of binned-genotypes_SE.xlsx", sheet = "GenAlex", skip = 1) %>%
+  select(-1) %>%                # removing first column, which is empty
+  gather(locus, allele, -1) %>% # gather all loci into tidy columns
+  mutate(locus = trimws(locus)) %>% # remove (F) designator
+  mutate(allele = as.integer(allele)) %>% # force alleles to integers
+  spread(locus, allele) %>%
+  slice(-n())
+ex <- ex[!names(ex) %in% locNames(dat)]
+
+# Function to select an adjacent allele. It will select the
+# next allele if the next allele is not missing and it's distance
+# is one away and the previous allele for the same conditions.
+# If none of the conditions are met, it will retain the allele.
+cromulent_allele <- Vectorize(function(lower, allele, higher){
+  if (!is.na(higher) && abs(allele - higher) == 1){
+    out <- higher
+  } else if (!is.na(lower) && abs(allele - lower) == 1){
+    out <- lower
+  } else {
+    out <- allele
+  }
+  out
+})
+ex
+exsummary <- ex %>% 
+  gather(locus, allele, -1) %>% # tidy the data
+  group_by(locus, allele) %>%   
+  summarize(n = n()) %>%        # summarize by count 
+  ungroup() %>%
+  group_by(locus) %>%           # group the loci, add the lower and upper alleles,
+  mutate(lower = lag(allele), higher = lead(allele)) %>% # and then create new_alleles
+  mutate(new_allele = ifelse(n < 3, cromulent_allele(lower, allele, higher), allele)) %>%
+  select(locus, new_allele, allele)
+exsummary
+corrected_loci <- ex %>% gather(locus, allele, -1) %>%
+  left_join(exsummary, by = c("locus", "allele")) %>%
+  mutate(allele = new_allele) %>%
+  select(-new_allele) %>%
+  spread(locus, allele)
+datdf <- genind2df(dat, usepop = FALSE) %>% bind_cols(corrected_loci[-1])
+dat  <- df2genind(datdf, ind.names = indNames(dat), strata = strata(dat), ploidy = 1) %>% as.genclone()
 ```
@@
 dat
+keeploci <- !locNames(dat) %in% colnames(corrected_loci)
+genotype_curve(dat, sample = 1000)
+locus_table(dat)
+# rl <- fix_replen(dat, repeat_lengths)
+# mlg.filter(dat, distance = bruvo.dist, replen = rl) <- .Machine$double.eps
+dat
+genotype_curve(dat[loc = keeploci, mlg.reset = TRUE])
+# dat <- dat[loc = keeploci, mlg.reset = TRUE]
+dat
 ```
@@
 ```{r}
+make_graph_list <- function(dat){ # dat is a genclone/snpclone object
   datmlg  <- mlg.table(dat, plot = FALSE) > 0 # presence/absence of MLG
@@
   adjmat  <- adjmat[cols, cols]
-adjmat
-crosses
-```
-
-Now that we have the adjacency matrix, we can use it to construct our graph:
-
-```{r, fig.width = 10, fig.height = 10}
+  # Creating Graph
   g           <- graph_from_adjacency_matrix(adjmat, mode = "undirected", diag = FALSE)
@@
   el          <- el[lengths(el) > 0]
-popgraphs <- setNames(vector(mode = "list", length = length(el)), names(el))
+  popgraphs <- setNames(vector(mode = "list", length = length(el) + 1), c(names(el), "total"))
   for (v in names(el)){
@@
   }
-par(mfrow = c(3, 4))
-for (i in names(popgraphs)){
-  pg <- popgraphs[[i]]
+  popgraphs[["total"]] <- g
+  popgraphs
+}
+
+plot_mlg_graph <- function(g, glayout = NULL){
+  if (is.null(glayout)){
+    glayout <- layout_nicely(g)
+  } else {
+    glayout <- glayout[V(g)$name, ]
+  }
+  shared_mlg  <- (1 - V(g)$weight) * V(g)$size
+  g           <- add_vertices(g, length(V(g)), size = V(g)$size - shared_mlg, color = "grey90")
+  g2          <- g
+  V(g2)$label <- ifelse(!is.na(V(g2)$name), sprintf("%s [(%d/%d)]", V(g2)$name, shared_mlg, V(g2)$size), NA)
+  glay        <- create_layout(g2, "manual", node.positions = as.data.frame(rbind(glayout, glayout)))
+  x_nudge     <- ifelse(abs(glay$x) == 1, -glay$x/10, glay$x/10)
+  ggraph(glay) +
+    geom_edge_fan(aes(alpha = weight + 1), width = 0.75) +
+    geom_node_circle(aes(r = drop(scale(size, center = FALSE))/10, fill = size)) +
+    geom_node_label(aes(label = label), repel = TRUE, parse = TRUE, label.size = 0.75, nudge_x = x_nudge) +
+    viridis::scale_fill_viridis(option = "C") +
+    scale_edge_alpha_continuous(range = c(0.25, 1), breaks = c(2:5)) +
+    coord_fixed() +
+    theme_void() +
+    theme(text = element_text(size = 14)) +
+    labs(list(
+      title = "Shared haplotypes across regions",
+      fill = "Number of\nGenotypes",
+      edge_alpha = "Populations\nper haploytpe",
+      caption = "Outer circle: Number of haplotypes in the region\nInner Circle: Number of private haplotypes in the region"
+    ))
+}
+
+plot_mlg_subgraph <- function(graphlist){
+  for (i in names(graphlist)){
+    pg   <- graphlist[[i]]
     labs <- ifelse(E(pg)$weight > 1, E(pg)$label, NA)
@@
   }
+}
+
+```
+
+Now that we have the adjacency matrix, we can use it to construct our graph:
+
+```{r, fig.width = 7, fig.height = 7}
+graph16loc <- make_graph_list(dat)
+graph11loc <- make_graph_list(dat[loc = keeploci, mlg.reset = TRUE])
+
+good_layout <- read.table(
+text = 
+"   x                            y
+AU  0.8090169943749470071737  5.877852522924730260812e-01
+CA -1.0000000000000000000000  1.224646799147350002426e-16
+CO -0.3090169943749480063744 -9.510565162951539752711e-01
+FR -0.3090169943749470071737  9.510565162951539752711e-01
+MI  0.0000000000000000000000  0.000000000000000000000e+00
+MN  1.0000000000000000000000  0.000000000000000000000e+00
+ND  0.3090169943749470071737 -9.510565162951539752711e-01
+NE -0.8090169943749470071737 -5.877852522924730260812e-01
+NY  0.8090169943749470071737 -5.877852522924730260812e-01
+OR  0.3090169943749470071737  9.510565162951539752711e-01
+WA -0.8090169943749470071737  5.877852522924730260812e-01"
+) %>% as.matrix()
+
+par(mfrow = c(3, 4))
+plot_mlg_subgraph(graph16loc[-length(graph16loc)])
+
+par(mfrow = c(3, 4))
+plot_mlg_subgraph(graph11loc[-length(graph11loc)])
+
 par(mfrow = c(1, 1))
-center_node <- degree(g) %>% which.max() %>% names()
+plot_mlg_graph(graph16loc$total, good_layout)
+plot_mlg_graph(graph11loc$total, good_layout)
+```
 
-new_layout <- structure(c(1, 0.809016994374947, 0.309016994374947, -0.309016994374947, 
--0.809016994374947, -1, -0.809016994374947, -0.309016994374948, 
-0, 0.309016994374947, 0.809016994374947, 0, 0.587785252292473, 
-0.951056516295154, 0.951056516295154, 0.587785252292473, 1.22464679914735e-16, 
--0.587785252292473, -0.951056516295154, 0, -0.951056516295154, 
--0.587785252292473), .Dim = c(11L, 2L), .Dimnames = list(c("MN", 
-"AU", "OR", "FR", "WA", "CA", "NE", "CO", "MI", "ND", "NY"), 
-    c("x", "y")))
-new_layout <- new_layout[V(g)$name, ]
-g <- add_vertices(g, length(V(g)), size = V(g)$size - shared_mlg, color = "grey90")
-plot(g, 
-     layout = rbind(new_layout, new_layout),
-     edge.width = E(g)$weight,
-     edge.label = NA)
-
-glay <- create_layout(g, "manual", node.positions = as.data.frame(rbind(new_layout, new_layout)))
-ggraph(glay) +
-  geom_edge_fan(aes(alpha = weight + 1)) +
-  geom_node_circle(aes(r = scale(size, center = FALSE)/10, fill = size, alpha = weight)) +
-  geom_node_label(aes(label = name), repel = TRUE) +
-  viridis::scale_fill_viridis(option = "C") +
-  coord_fixed() +
-  theme_void() +
-  labs(list(
-    title = "Shared haplotypes across regions",
-    fill = "Number of\nGenotypes",
-    alpha = "Fraction of\nprivate genotpes",
-    edge_alpha = "Populations\nper haploytpe",
-    caption = "Outer circle: Number of haplotypes in the region\nInner Circle: Number of private haplotypes in the region"
-  ))
+```{r, eval = FALSE}
+null_vertices <- length(V(g)) %>% seq(from = ./2 + 1, to = ., by = 1)
+sg <- g %>% 
+  delete_vertices(null_vertices) %>%
+  cluster_spinglass()
+modularity(sg)
+communities(sg)
 ```

### 2017-04-03 doc/RMD/MLG-distribution.Rmd
add section

@@
 
-```
-
-Now that we have the adjacency matrix, we can use it to construct our graph:
-
-```{r, fig.width = 7, fig.height = 7}
-graph16loc <- make_graph_list(dat)
-graph11loc <- make_graph_list(dat[loc = keeploci, mlg.reset = TRUE])
-
 good_layout <- read.table(
@@
 
+```
+
+## Graphs
+
+With the fuctions above, we can create and plot the graphs
+
+```{r, fig.width = 7, fig.height = 7}
+# Creating the graphs
+graph16loc <- make_graph_list(dat)
+graph11loc <- make_graph_list(dat[loc = keeploci, mlg.reset = TRUE])
+
+# Plotting the subgraphs
 par(mfrow = c(3, 4))
@@
 
+# Plotting the full graphs
 par(mfrow = c(1, 1))

### 2017-04-05 doc/RMD/MLG-distribution.Rmd
update analysis to weight by psex and include notes

--- 2017-04-03 doc/RMD/MLG-distribution.Rmd
@@
 I realized that it's possible to use an MLG table with matrix multiplication to
-get an adjency matrix.
+get an adjency matrix. Here I'm making a function to create a list of graphs.
+
+The vertices, representing populations, will have the following attributes:
+
+ - Size: equal to the number of MLGs within the population
+ - Weight: fraction of private MLGs
+
+The edges, representing multilocus genotypes are a bit trickier to handle since
+multiple edges can represent a single MLG passing through populations. They have
+the following attributes:
+
+ - Label: the MLG this edge belongs to. Note that multiple edges can share the
+          same label.
+ - Width: the number of populations this MLG passes through
+ - Weight: 1 - (probability of a second encounter by chance); 1 - psex
+
+The reason why I'm weighting these nodes as 1 - psex is for subsequent community
+analysis. The more likely you are to encounter a genotype by chance, the less
+influence that genotype should have on the connectivity.
 
 ```{r}
+rerange <- function(x){
+  minx <- min(x, na.rm = TRUE)
+  maxx <- max(x, na.rm = TRUE)
+  if (minx < 0)
+    x <- x + abs(minx)
+    maxx <- maxx + abs(minx)
+  if (maxx > 1)
+    x <- x/maxx
+  return(x)
+}
 make_graph_list <- function(dat){ # dat is a genclone/snpclone object
+  # w       <- pgen(dat, by_pop = FALSE) %>% rowSums() %>% exp() %>% setNames(paste0("MLG.", mll(dat)))
+  # w       <- w[unique(names(w))]
+  w       <- psex(dat, by_pop = FALSE, method = "multiple") %>% split(mll(dat))
+  names(w)<- paste0("MLG.", names(w))
   datmlg  <- mlg.table(dat, plot = FALSE) > 0 # presence/absence of MLG
@@
   shared_mlg  <- (crosses %>% group_by(Population) %>% summarize(n = n()))$n
-  V(g)$weight <- 1 - shared_mlg/V(g)$size
+  V(g)$weight <- 1 - shared_mlg/V(g)$size # fraction of private MLGs
   el          <- as_adj_edge_list(g)
@@
     E(g)[idx]$label  <- substr(MLGS, 5, nchar(MLGS))
-    E(g)[idx]$weight <- as.integer(table(MLGS)[MLGS]) # weight == n populations visited
+    E(g)[idx]$width  <- as.integer(table(MLGS)[MLGS]) # size == n populations visited
+    E(g)[idx]$weight <- 1 - map_dbl(w[MLGS], 2)       # weight == 1 - psex 2nd encounter
     popgraphs[[v]]   <- subgraph.edges(g, eids = idx)
@@
   x_nudge     <- ifelse(abs(glay$x) == 1, -glay$x/10, glay$x/10)
+  breaks      <- quantile(1 - E(g)$weight)
+  breaks      <- setNames(breaks, format(breaks, digits = 2))
   ggraph(glay) +
-    geom_edge_fan(aes(alpha = weight + 1), width = 0.75) +
+    geom_edge_fan(aes(alpha = 1 - weight, width = width + 1)) +
     geom_node_circle(aes(r = drop(scale(size, center = FALSE))/10, fill = size)) +
@@
     viridis::scale_fill_viridis(option = "C") +
-    scale_edge_alpha_continuous(range = c(0.25, 1), breaks = c(2:5)) +
+    scale_edge_alpha_continuous(range = c(1, 0.25), breaks = breaks) +
+    scale_edge_width_continuous(range = c(0.25, 1.25), breaks = c(2:5)) +
     coord_fixed() +
@@
       fill = "Number of\nGenotypes",
-      edge_alpha = "Populations\nper haploytpe",
+      edge_alpha = "Probability of\nsecond encounter",
+      edge_width = "Populations\nPer haplotype",
       caption = "Outer circle: Number of haplotypes in the region\nInner Circle: Number of private haplotypes in the region"
@@
     pg   <- graphlist[[i]]
-    labs <- ifelse(E(pg)$weight > 1, E(pg)$label, NA)
+    labs <- ifelse(E(pg)$width > 1, E(pg)$label, NA)
     labs <- ifelse(duplicated(labs), NA, labs)
@@
          layout = layout_as_star(pg, center = i), 
-         edge.width = E(pg)$weight,
+         edge.width = E(pg)$width,
+         edge.color = grey(rerange(1 - E(pg)$weight)), 
          edge.label = labs)
@@
 }
-
-```
-
-Now that we have the adjacency matrix, we can use it to construct our graph:
-
-```{r, fig.width = 7, fig.height = 7}
-graph16loc <- make_graph_list(dat)
-graph11loc <- make_graph_list(dat[loc = keeploci, mlg.reset = TRUE])
 
@@
 
+```
+
+## Graphs
+
+With the fuctions above, we can create and plot the graphs.
+
+```{r}
+# Creating the graphs
+graph16loc <- make_graph_list(dat)
+graph11loc <- make_graph_list(dat[loc = keeploci, mlg.reset = TRUE])
+```
+
+### Individual population subgraphs
+
+Each subgraph shows all of the connections for a single population.
+
+```{r, fig.width = 10, fig.height = 10}
+# Plotting the subgraphs
 par(mfrow = c(3, 4))
@@
 plot_mlg_subgraph(graph11loc[-length(graph11loc)])
-
-par(mfrow = c(1, 1))
-plot_mlg_graph(graph16loc$total, good_layout)
-plot_mlg_graph(graph11loc$total, good_layout)
 ```
 
-```{r, eval = FALSE}
-null_vertices <- length(V(g)) %>% seq(from = ./2 + 1, to = ., by = 1)
-sg <- g %>% 
-  delete_vertices(null_vertices) %>%
-  cluster_spinglass()
-modularity(sg)
-communities(sg)
+### Cross-regional graphs
+
+#### With 16 loci
+
+First, I'm going to show the results of a community analysis. I'm using the 
+igraph function `cluster_optimal()` to cluster the nodes.
+
+```{r, fig.width = 7, fig.height = 7}
+par(mfrow = c(1, 1))
+(g16o <- cluster_optimal(graph16loc$total))
+plot_mlg_graph(graph16loc$total, good_layout) + labs(list(subtitle = "(16 loci)"))
 ```
+
+What we see is that We are given three clusters showing a clustering of the
+plains states, the west coast, and Australia, France, and Minnesota. The last
+cluster is likely driven by the single genotype shared between these three
+regions that has a low probability of a second encounter.
+
+#### With 11 loci
+
+```{r, fig.width = 7, fig.height = 7}
+(g11o <- cluster_optimal(graph11loc$total))
+plot_mlg_graph(graph11loc$total, good_layout) + labs(list(subtitle = "(11 loci)"))
+```

### 2017-04-05 doc/RMD/MLG-distribution.Rmd
rerun with session information

@@
     viridis::scale_fill_viridis(option = "C") +
-    scale_edge_alpha_continuous(range = c(1, 0.25), breaks = breaks) +
+    scale_edge_alpha_continuous(range = c(1, 0.25), breaks = rev(breaks)) +
     scale_edge_width_continuous(range = c(0.25, 1.25), breaks = c(2:5)) +
@@
 ```
+
+## Session Information
+
+```{r}
+options(width = 100)
+devtools::session_info()
+```

### 2017-04-05 doc/RMD/MLG-distribution.Rmd
add quiet statement

@@
 keeploci <- !locNames(dat) %in% colnames(corrected_loci)
-genotype_curve(dat, sample = 1000)
+genotype_curve(dat, sample = 1000, quiet = TRUE)
 locus_table(dat)
@@
 dat
-genotype_curve(dat[loc = keeploci, mlg.reset = TRUE])
+genotype_curve(dat[loc = keeploci, mlg.reset = TRUE], quiet = TRUE)
 # dat <- dat[loc = keeploci, mlg.reset = TRUE]

### 2017-04-06 doc/RMD/MLG-distribution.Rmd
rearrange layout

@@
 
-plot_mlg_graph <- function(g, glayout = NULL){
+plot_mlg_graph <- function(g, glayout = NULL, label = TRUE){
   if (is.null(glayout)){
@@
   breaks      <- setNames(breaks, format(breaks, digits = 2))
-  ggraph(glay) +
+  outg <- ggraph(glay) +
     geom_edge_fan(aes(alpha = 1 - weight, width = width + 1)) +
-    geom_node_circle(aes(r = drop(scale(size, center = FALSE))/10, fill = size)) +
-    geom_node_label(aes(label = label), repel = TRUE, parse = TRUE, label.size = 0.75, nudge_x = x_nudge) +
+    geom_node_circle(aes(r = drop(scale(size, center = FALSE))/10, fill = size))
+  outg <- if (label) outg + geom_node_label(aes(label = label), repel = TRUE, parse = TRUE, label.size = 0.75, nudge_x = x_nudge) else outg
+  outg + 
     viridis::scale_fill_viridis(option = "C") +
@@
 
+alt_layout <- read.table(
+text = 
+"   x                            y
+AU  0.8090169943749470071737  5.877852522924730260812e-01
+OR -1.0000000000000000000000  1.224646799147350002426e-16
+CO -0.3090169943749480063744 -9.510565162951539752711e-01
+WA -0.3090169943749470071737  9.510565162951539752711e-01
+MI  0.0000000000000000000000  0.000000000000000000000e+00
+MN  1.0000000000000000000000  0.000000000000000000000e+00
+ND  0.3090169943749470071737 -9.510565162951539752711e-01
+NE -0.8090169943749470071737 -5.877852522924730260812e-01
+NY  0.8090169943749470071737 -5.877852522924730260812e-01
+FR  0.3090169943749470071737  9.510565162951539752711e-01
+CA -0.8090169943749470071737  5.877852522924730260812e-01"
+) %>% as.matrix()
+alt_layout <- alt_layout[rownames(good_layout), ]
 ```
@@
 (g16o <- cluster_optimal(graph16loc$total))
-plot_mlg_graph(graph16loc$total, good_layout) + labs(list(subtitle = "(16 loci)"))
+plot_mlg_graph(graph16loc$total, alt_layout) + labs(list(subtitle = "(16 loci)"))
 ```
@@
 (g11o <- cluster_optimal(graph11loc$total))
-plot_mlg_graph(graph11loc$total, good_layout) + labs(list(subtitle = "(11 loci)"))
+plot_mlg_graph(graph11loc$total, alt_layout) + labs(list(subtitle = "(11 loci)"))
 ```

### 2017-04-06 doc/RMD/MLG-distribution.Rmd
genotype -> haplotype
because that's what they are

@@
       title = "Shared haplotypes across regions",
-      fill = "Number of\nGenotypes",
+      fill = "Number of\nhaplotypes",
       edge_alpha = "Probability of\nsecond encounter",
-      edge_width = "Populations\nPer haplotype",
+      edge_width = "Populations\nper haplotype",
       caption = "Outer circle: Number of haplotypes in the region\nInner Circle: Number of private haplotypes in the region"

### 2017-04-06 doc/RMD/MLG-distribution.Rmd
rename populations

--- 2017-04-05 doc/RMD/MLG-distribution.Rmd
@@
 splitStrata(dat) <- ~Isolate/Severity/MCG/Region/Source/Year/Host
+pops <- levels(strata(dat)$Region)
+levels(strata(dat)$Region) <- case_when(pops == "FR" ~ "France", 
+                                        pops == "MX" ~ "Mexico", 
+                                        pops == "AU" ~ "Australia", 
+                                        TRUE ~ pops)
 dat
@@
   shared_mlg  <- (crosses %>% group_by(Population) %>% summarize(n = n()))$n
-  V(g)$weight <- 1 - shared_mlg/V(g)$size # fraction of private MLGs
+  V(g)$weight <- shared_mlg/V(g)$size # fraction of shared MLGs
   el          <- as_adj_edge_list(g)
@@
   }
-  shared_mlg  <- (1 - V(g)$weight) * V(g)$size
+  shared_mlg  <- V(g)$weight * V(g)$size
   g           <- add_vertices(g, length(V(g)), size = V(g)$size - shared_mlg, color = "grey90")
@@
 "   x                            y
-AU  0.8090169943749470071737  5.877852522924730260812e-01
+Australia  0.8090169943749470071737  5.877852522924730260812e-01
 CA -1.0000000000000000000000  1.224646799147350002426e-16
 CO -0.3090169943749480063744 -9.510565162951539752711e-01
-FR -0.3090169943749470071737  9.510565162951539752711e-01
+France -0.3090169943749470071737  9.510565162951539752711e-01
 MI  0.0000000000000000000000  0.000000000000000000000e+00
@@
 ) %>% as.matrix()
+
+alt_layout <- read.table(
+text = 
+"   x                            y
+Australia  0.8090169943749470071737  5.877852522924730260812e-01
+OR -1.0000000000000000000000  1.224646799147350002426e-16
+CO -0.3090169943749480063744 -9.510565162951539752711e-01
+WA -0.3090169943749470071737  9.510565162951539752711e-01
+MI  0.0000000000000000000000  0.000000000000000000000e+00
+MN  1.0000000000000000000000  0.000000000000000000000e+00
+ND  0.3090169943749470071737 -9.510565162951539752711e-01
+NE -0.8090169943749470071737 -5.877852522924730260812e-01
+NY  0.8090169943749470071737 -5.877852522924730260812e-01
+France  0.3090169943749470071737  9.510565162951539752711e-01
+CA -0.8090169943749470071737  5.877852522924730260812e-01"
+) %>% as.matrix()
+alt_layout <- alt_layout[rownames(good_layout), ]
 

### 2017-04-11 doc/RMD/data-comparison.Rmd
update data-comparison to filter data as well

@@
 "locus\trange\tn
-5-2	318-324	4
-6-2	483-495	3
-7-2	158-174	7
-8-3	244-270	7
-9-2	360-382	9
-12-2	214-222	5
-17-3	342-363	7
-20-3	280-282	2
-55-4	153-216	10
-110-4	370-386	5
-114-4	339-416	10
+5-2(F)	318-324	4
+6-2(F)	483-495	3
+7-2(F)	158-174	7
+8-3(H)	244-270	7
+9-2(F)	360-382	9
+12-2(H)	214-222	5
+17-3(H)	342-363	7
+20-3(F)	280-282	2
+55-4(F)	153-216	10
+110-4(H)	370-386	5
+114-4(H)	339-416	10
 "
@@
   gather(locus, allele, -1) %>% # gather all loci into tidy columns
-  mutate(locus = trimws(locus) %>% substr(1, nchar(.) - 3)) %>% # remove (F) designator
+  mutate(locus = trimws(locus)) %>% # remove whitespace in locus names
   mutate(allele = as.integer(allele)) %>% # force alleles to integers
@@
   gather(locus, allele, -1) %>% # gather all loci into tidy columns
-  mutate(locus = trimws(locus) %>% substr(1, nchar(.) - 3)) %>% # remove (F) designator
+  mutate(locus = trimws(locus)) %>% # remove any whitespace in the locus names
   spread(locus, allele) %>%     # spread data out with individual loci in columns
@@
 
+## Saving data as genclone object
+
+First we need to transform the data to something that
+
+```{r}
+dat <- saj %>% 
+  select(-(Isolate:Host)) %>% 
+  df2genind(ind.names = saj$Isolate, 
+            strata = select(saj, MCG:Host), # Filtering out severity and Isolate
+            ploidy = 1) %>%
+  as.genclone()
+nameStrata(dat)[2] <- "Region"
+pops <- levels(strata(dat)$Region)
+levels(strata(dat)$Region) <- case_when(pops == "FR" ~ "France", 
+                                        pops == "MX" ~ "Mexico", 
+                                        pops == "AU" ~ "Australia", 
+                                        TRUE ~ pops)
+dat
+locNames(dat)
+
+repeat_lengths <-
+  c(
+  `5-2(F)` = 2.000000,
+  `5-3(F)` = 0.500000,
+  `6-2(F)` = 6.000000,
+  `7-2(F)` = 2.000000,
+  `8-3(H)` = 2.000000,
+  `9-2(F)` = 2.000000,
+  `12-2(H)` = 2.000000,
+  `17-3(H)` = 3.000000,
+  `20-3(F)` = 2.000000,
+  `36-4(F)` = 0.500000,
+  `50-4(F)` = 0.500000,
+  `55-4(F)` = 4.000000,
+  `92-4(F)` = 2.000000,
+  `106-4(H)` = 4.000000,
+  `110-4(H)` = 4.000000,
+  `114-4(H)` = 4.000000
+  )
+repeat_lengths <- ifelse(repeat_lengths < 1, 4, repeat_lengths)
+```
+
+### Adding inconsistent loci
+
+Only 11 loci were manually curated by Sydney and Sajeewa. I'm taking a look at
+the other 5 loci because they may be informative. In order to do that, I'm going
+to have to clean them up by estimating the the true allele size.
+
+```{r read_excel}
+ex <- readxl::read_excel("../Analysis4 ForManu/A1_Copy of binned-genotypes_SE.xlsx", sheet = "GenAlex", skip = 1) %>%
+  select(-1) %>%                # removing first column, which is empty
+  gather(locus, allele, -1) %>% # gather all loci into tidy columns
+  mutate(locus = trimws(locus)) %>% # remove (F) designator
+  mutate(allele = as.integer(allele)) %>% # force alleles to integers
+  spread(locus, allele) %>%
+  slice(-n())
+ex <- ex[!names(ex) %in% locNames(dat)]
+
+# Function to select an adjacent allele. It will select the
+# next allele if the next allele is not missing and it's distance
+# is one away and the previous allele for the same conditions.
+# If none of the conditions are met, it will retain the allele.
+cromulent_allele <- Vectorize(function(lower, allele, higher){
+  if (!is.na(higher) && abs(allele - higher) == 1){
+    out <- higher
+  } else if (!is.na(lower) && abs(allele - lower) == 1){
+    out <- lower
+  } else {
+    out <- allele
+  }
+  out
+})
+ex
+exsummary <- ex %>% 
+  gather(locus, allele, -1) %>% # tidy the data
+  group_by(locus, allele) %>%   
+  summarize(n = n()) %>%        # summarize by count 
+  ungroup() %>%
+  group_by(locus) %>%           # group the loci, add the lower and upper alleles,
+  mutate(lower = lag(allele), higher = lead(allele)) %>% # and then create new_alleles
+  mutate(new_allele = ifelse(n < 3, cromulent_allele(lower, allele, higher), allele)) %>%
+  select(locus, new_allele, allele)
+exsummary
+```
+
+Now that we have our data set filtered (to a degree), we can merge the data with
+`dat` that we defined above.
+
+```{r merging_data}
+corrected_loci <- ex %>% gather(locus, allele, -1) %>%
+  left_join(exsummary, by = c("locus", "allele")) %>%
+  mutate(allele = new_allele) %>%
+  select(-new_allele) %>%
+  spread(locus, allele) %>%
+  separate(iso_st_mcg_org_loc_yr_hst_cult_rep, # Note: this corresponds closer to the initial data
+           c("Isolate", "Severity", "MCG", "State", "Source", "Year", "Host"), 
+           sep = "_") %>%
+  arrange(Isolate) %>%
+  select(-(Severity:Host))
+datdf <- genind2df(dat, usepop = FALSE) %>% 
+  rownames_to_column(var = "Isolate") %>% 
+  left_join(corrected_loci, by = "Isolate")
+stopifnot(identical(datdf$Isolate, indNames(dat)))
+dat   <- datdf %>% 
+  select(-Isolate) %>% 
+  df2genind(, ind.names = indNames(dat), strata = strata(dat), ploidy = 1) %>% 
+  as.genclone()
+```
+
+The original data includes both Severity and Isolate. Since these are not
+necessary for delimiting the strata, we will place them in the "other" slot
+after converting Severity to numeric. Placing this information in the "other"
+slot ensures that these data will travel with the object.
+
+```{r fix_strata}
+stopifnot(identical(indNames(dat), saj$Isolate))
+other(dat)$meta <- select(saj, Severity, Isolate)
+other(dat)$REPLEN <- fix_replen(dat, repeat_lengths)
+```
+
+```{r, save_data}
+save(dat, datdf, corrected_loci, file = "data/sclerotinia_16_loci.dat")
+```
+
 ## Session Information

### 2017-04-11 doc/RMD/MLG-distribution.Rmd
update MLG-distribution to save SVG files

@@
 library('tidyverse')
-library('assertr')
 library('poppr')
@@
 library("ggraph")
-library('magrittr')
 ```
 
-## Loading data and assertions
+## Loading data and setting strata
 
-```{r read_data}
-dat <- read.genalex("../Analysis4 ForManu/A2_Copy4 EUR_AUS_forManu.csv", ploidy = 1)
-splitStrata(dat) <- ~Isolate/Severity/MCG/Region/Source/Year/Host
-pops <- levels(strata(dat)$Region)
-levels(strata(dat)$Region) <- case_when(pops == "FR" ~ "France", 
-                                        pops == "MX" ~ "Mexico", 
-                                        pops == "AU" ~ "Australia", 
-                                        TRUE ~ pops)
+I'm loading the data and splitting it into the 11 locus and 16 locus versions
+
+```{r}
+load("data/sclerotinia_16_loci.dat")
+setPop(dat) <- ~Region
 dat
-
-repeat_lengths <-
-  c(
-  `5-2` = 2.000000,
-  `5-3` = 0.500000,
-  `6-2` = 6.000000,
-  `7-2` = 2.000000,
-  `8-3` = 2.000000,
-  `9-2` = 2.000000,
-  `12-2` = 2.000000,
-  `17-3` = 3.000000,
-  `20-3` = 2.000000,
-  `36-4` = 0.500000,
-  `50-4` = 0.500000,
-  `55-4` = 4.000000,
-  `92-4` = 2.000000,
-  `106-4` = 4.000000,
-  `110-4` = 4.000000,
-  `114-4` = 4.000000
-  )
-repeat_lengths <- ifelse(repeat_lengths < 1, 4, repeat_lengths)
+locNames(dat)
+other(dat)$REPLEN
+head(other(dat)$meta)
+keeploci <- !locNames(dat) %in% colnames(corrected_loci)
+dat11 <- dat[loc = keeploci, mlg.reset = TRUE] # reducing to 11 loci and recalculating mlgs
+dat11
 ```
 
-```{r excel_data}
-ex <- readxl::read_excel("../Analysis4 ForManu/A1_Copy of binned-genotypes_SE.xlsx", sheet = "GenAlex", skip = 1) %>%
-  select(-1) %>%                # removing first column, which is empty
-  gather(locus, allele, -1) %>% # gather all loci into tidy columns
-  mutate(locus = trimws(locus)) %>% # remove (F) designator
-  mutate(allele = as.integer(allele)) %>% # force alleles to integers
-  spread(locus, allele) %>%
-  slice(-n())
-ex <- ex[!names(ex) %in% locNames(dat)]
+Now we can inspect the genotype accumulation curves. Notice that the 16 loci
+version plateaus, but we know that these extra loci are inconsistent, so we 
+might not want to trust them.
 
-# Function to select an adjacent allele. It will select the
-# next allele if the next allele is not missing and it's distance
-# is one away and the previous allele for the same conditions.
-# If none of the conditions are met, it will retain the allele.
-cromulent_allele <- Vectorize(function(lower, allele, higher){
-  if (!is.na(higher) && abs(allele - higher) == 1){
-    out <- higher
-  } else if (!is.na(lower) && abs(allele - lower) == 1){
-    out <- lower
-  } else {
-    out <- allele
-  }
-  out
-})
-ex
-exsummary <- ex %>% 
-  gather(locus, allele, -1) %>% # tidy the data
-  group_by(locus, allele) %>%   
-  summarize(n = n()) %>%        # summarize by count 
-  ungroup() %>%
-  group_by(locus) %>%           # group the loci, add the lower and upper alleles,
-  mutate(lower = lag(allele), higher = lead(allele)) %>% # and then create new_alleles
-  mutate(new_allele = ifelse(n < 3, cromulent_allele(lower, allele, higher), allele)) %>%
-  select(locus, new_allele, allele)
-exsummary
-corrected_loci <- ex %>% gather(locus, allele, -1) %>%
-  left_join(exsummary, by = c("locus", "allele")) %>%
-  mutate(allele = new_allele) %>%
-  select(-new_allele) %>%
-  spread(locus, allele)
-datdf <- genind2df(dat, usepop = FALSE) %>% bind_cols(corrected_loci[-1])
-dat  <- df2genind(datdf, ind.names = indNames(dat), strata = strata(dat), ploidy = 1) %>% as.genclone()
+```{r genotype_curves, fig.width = 5, fig.height = 5}
+locus_table(dat)
+genotype_curve(dat, sample = 1000, quiet = TRUE)
+genotype_curve(dat11, quiet = TRUE)
 ```
 
-The incoming strata includes both Severity and Isolate. Since these are not
-necessary for delimiting the strata, we will place them in the "other" slot
-after converting Severity to numeric. Placing this information in the "other"
-slot ensures that these data will travel with the object.
-
-```{r fix_strata}
-dat_strata <- strata(dat) %>%
-  mutate_all(as.character) %>%
-  mutate(Severity = as.numeric(Severity))
-strata(dat)     <- select(dat_strata, -Severity, -Isolate)
-indNames(dat)   <- dat_strata$Isolate
-other(dat)$meta <- select(dat_strata, Severity, Isolate)
-```
+Creating contracted genotypes for use later
 
 ```{r}
-setPop(dat) <- ~Region
+mlg.filter(dat, distance = bruvo.dist, replen = other(dat)$REPLEN) <- .Machine$double.eps
 dat
-keeploci <- !locNames(dat) %in% colnames(corrected_loci)
-genotype_curve(dat, sample = 1000, quiet = TRUE)
-locus_table(dat)
-# rl <- fix_replen(dat, repeat_lengths)
-# mlg.filter(dat, distance = bruvo.dist, replen = rl) <- .Machine$double.eps
-dat
-genotype_curve(dat[loc = keeploci, mlg.reset = TRUE], quiet = TRUE)
-# dat <- dat[loc = keeploci, mlg.reset = TRUE]
-dat
+mlg.filter(dat11, distance = bruvo.dist, replen = other(dat11)$REPLEN) <- .Machine$double.eps
+dat11
+mll(dat) <- "original"
+mll(dat11) <- "original"
 ```
@@
 
-```{r, fig.width = 10, fig.height = 10}
+```{r subgraphs, fig.width = 10, fig.height = 10}
 # Plotting the subgraphs
@@
 
+I'm plotting two graphs here because it's useful to see if both of them are
+congruent in their community predictions.
+
 #### With 16 loci
@@
 
-```{r, fig.width = 7, fig.height = 7}
+```{r graph-16-loci, fig.width = 7, fig.height = 7}
 par(mfrow = c(1, 1))
@@
 plot_mlg_graph(graph16loc$total, alt_layout) + labs(list(subtitle = "(16 loci)"))
+ggsave(filename = "results/figures/publication/FigureS1Z.svg", width = 88, units = "mm", scale = 2)
 ```
@@
 
-```{r, fig.width = 7, fig.height = 7}
+```{r graph-11-loci, fig.width = 7, fig.height = 7}
 (g11o <- cluster_optimal(graph11loc$total))
 plot_mlg_graph(graph11loc$total, alt_layout) + labs(list(subtitle = "(11 loci)"))
+ggsave(filename = "results/figures/publication/Figure3Z.svg", width = 88, units = "mm", scale = 2)
 ```

### 2017-04-11 doc/RMD/MLG-distribution.Rmd
update SVG images to not be too dang tall

@@
 plot_mlg_graph(graph16loc$total, alt_layout) + labs(list(subtitle = "(16 loci)"))
-ggsave(filename = "results/figures/publication/FigureS1Z.svg", width = 88, units = "mm", scale = 2)
+ggsave(filename = "results/figures/publication/FigureS1Z.svg", width = 88, height = 88, units = "mm", scale = 2)
 ```
@@
 plot_mlg_graph(graph11loc$total, alt_layout) + labs(list(subtitle = "(11 loci)"))
-ggsave(filename = "results/figures/publication/Figure3Z.svg", width = 88, units = "mm", scale = 2)
+ggsave(filename = "results/figures/publication/Figure3Z.svg", width = 88, height = 88, units = "mm", scale = 2)
 ```

### 2017-04-11 doc/RMD/MLG-distribution.Rmd
update to save graphs
update name of data

@@
 ---
-title: "R Notebook"
+title: "Analyzing MLG distribution"
 output: 
@@
 ```{r}
-load("data/sclerotinia_16_loci.dat")
+load("data/sclerotinia_16_loci.rda")
 setPop(dat) <- ~Region
@@
 
+```{r, saving_data}
+save(graph16loc, graph11loc, alt_layout, file = "data/mlg-crosspop-graph.rda")
+```
+
 ## Session Information

### 2017-04-11 doc/RMD/RDA-analysis.Rmd
first stab at dbRDA
I still need to read up on this.

---
title: "dbRDA analysis"
output: 
  html_notebook:
    toc: true
---

```{r setup, include = FALSE}
knitr::opts_knit$set(root.dir = PROJHOME)
```

This document serves to re-assess Sajeewa's dbRDA analysis with the repeat
lengths that have been modified to avoid inconsistent allele calls.

Modification in this case means that a tiny amount has been added or subtracted
to the repeat length to ensure that the alleles are all unique after division.

Below is my attempt at reproduction of Sajeewa's analysis.

## Packages and Data

```{r, load_packages}
library('tidyverse')
library('poppr')
library('vegan')
```

```{r, load_data}
load("data/sclerotinia_16_loci.rda")
keeploci <- !locNames(dat) %in% colnames(corrected_loci)
setPop(dat) <- ~Host/Source/Region/Year
dat11 <- dat[loc = keeploci, mlg.reset = TRUE]
dat11cc <- clonecorrect(dat11, ~Host/Source/Region/Year, keep = 1:4)
dat16cc   <- clonecorrect(dat, ~Host/Source/Region/Year, keep = 1:4)
dat11cc
dat16cc
# environmental variables for dbRDA
stopifnot(identical(indNames(dat11cc), other(dat11cc)$meta$Isolate))
ENV11 <- bind_cols(select(other(dat11cc)$meta, Severity), 
                   select(strata(dat11cc), -MCG))
ENV11
stopifnot(identical(indNames(dat16cc), other(dat16cc)$meta$Isolate))
ENV16 <- bind_cols(select(other(dat16cc)$meta, Severity), 
                   select(strata(dat16cc), -MCG))
ENV16
```

## Calculate Bruvo's distance

```{r}
bruvo11cc <- bruvo.dist(dat11cc, replen = other(dat)$REPLEN)
bruvo16cc <- bruvo.dist(dat16cc, replen = other(dat)$REPLEN)
```

```{r}
commtab <- mlg.table(dat11cc, plot = FALSE)
cap11cc <- capscale(bruvo11cc ~ Region + Host + Year + Severity, 
                    data = ENV11, 
                    comm = , 
                    add = TRUE)

```

```{r}
capsum <- summary(cap11cc)
summary(summary(cap11cc))
##########Extracting 10 most important centroids for factor constrains 
###############It doesn't matter how you start the ggplot aesthetics. Can be DecendCentro CAP1 and CAP2 or site_centroids
############### CAP1 and CAP2. Same graph is resulted

Centroids <- data.frame(capsum$centroids) #no need to convert to a dataframe
LenCent <- sqrt((Centroids[,1]^2)*(Centroids[,2]^2))   ###  This calculates the arrow length (ie. strength of the eigenvector)
NewCentro <- cbind(Centroids, LenCent)
DecendCentro <- head(NewCentro[order(NewCentro[,7], decreasing = T),], n=23)   ###  this finds the n strongest eigenvectors by ordering them
#above "head" is correct

DecendCentro$cent_type <- rownames(head(DecendCentro, n=23))
DecendCentro <- head(DecendCentro, n = 10)
DecendCentro #constraint factor cetroids

## Testing plot in ggplot

ggplot(data= DecendCentro, mapping = aes(x = CAP1, y = CAP2))+
  geom_text(aes(label = cent_type))+ # label tells geom_text which text you want to plot
  ylim(-3.5,3.5)+ #NB note that this is a convenience wrapper and may cut data out of your plot
  xlim(-3,3.5) #important if you are calculating stats in the plot - these will be excluded

# sites/isolates
site_centroids <- data.frame(capsum$sites)
site_centroids$isolate_names <- rownames(site_centroids) # you can also put label = rownames() etc in the ggplot code. But sometimes you may want it in a column for later

##plotting both main centroids and isolates together
# I am plotting isolates firest and then envt centroids, alpha = transparancy
ggplot(DecendCentro, aes(x = CAP1, y = CAP2))+
  geom_point(data = site_centroids, colour = "orange", size =3.5, alpha =0.7)+
  geom_text(aes(label = cent_type))+
  ylim(-4,3.5)+
  xlim(-3.5,3.5)

#######strongest arrows: THESE ARE BIPLOT scores#############

arrws <- data.frame(capsum$biplot)
lenarrw <- sqrt(arrws[,1]^2*arrws[,2]^2)   ###  This calculates the arrow length (ie. strength of the eigenvector)
arrws <- cbind(arrws, lenarrw)
arrws <- head(arrws[order(arrws[,7], decreasing = T),], n=10)   ###  this finds the n strongest eigenvectors by ordering them
arrws

arrws$class <- (rownames(arrws)) #turning rownames into a variable
mult <- attributes(capsum)$const/2.5 # 97% sure this is the scaling for the arrows, pls let me know if I'm wrong
arrws$class
#arrws$class <- c("MX", "ND", "CA", "mtc", "NE", "MCG ")  # change this if you want to change how names are displayed
#arrws$class <- c("MX", "mtc", "NE", "ND", "MCG", "CA")
##locations abbreviations; see the original for full names

#####plotting centroids, isolates and biplot arrows together#########
##@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@########

arrws$class <- c("Mexico", "2005", "CA", "Vista", "CO", "Host: unknown", "OR", "Severity", "MN", "WA")
mult <- attributes(capsum)$const/3

(Cap_plot <- ggplot(DecendCentro, aes(x = CAP1, y = CAP2))+
   geom_text(aes(label = cent_type), colour = "blue")+
   geom_point(data = site_centroids, alpha=1/2, colour = "dark orange", size = 2.5, )+
   #coord_cartesian(x = c(-3.5, 3.5), y = c(-4.5, 2.5))+# For bruvo
   coord_cartesian(x = c(-3.5, 3.5), y = c(-2.5, 5))+
   geom_segment(data = arrws,
                aes(x = 0, xend = mult * CAP1, y = 0, yend = mult * CAP2),
                arrow = arrow(length = unit(0.3, "cm")), colour = "dark grey") + #grid is required for arrow to work.
   geom_text(data = arrws, color = "black",
             aes(x= (mult + mult/10) * CAP1, y = (mult + mult/10) * CAP2, label = arrws$class), #otherwise you could use hjust and vjust. I prefer this option
             size = 5, hjust = 0.5)+ theme_bw())
```

### 2017-04-13 doc/RMD/data-comparison.Rmd
update name of data
rerun with correct locus configuration
This ensures the MLGs are the same as the paper.

@@
 stopifnot(identical(datdf$Isolate, indNames(dat)))
+datdf <- datdf[c("Isolate", names(repeat_lengths))]
 dat   <- datdf %>% 
   select(-Isolate) %>% 
-  df2genind(, ind.names = indNames(dat), strata = strata(dat), ploidy = 1) %>% 
+  df2genind(ind.names = indNames(dat), strata = strata(dat), ploidy = 1) %>% 
   as.genclone()
@@
 
+```{r}
+setPop(dat) <- ~Region
+dat
+locNames(dat)
+other(dat)$REPLEN
+head(other(dat)$meta)
+keeploci <- !locNames(dat) %in% colnames(corrected_loci)
+dat11 <- dat[loc = keeploci, mlg.reset = TRUE] # reducing to 11 loci and recalculating mlgs
+dat11
+```
+
 ```{r, save_data}
-save(dat, datdf, corrected_loci, file = "data/sclerotinia_16_loci.dat")
+save(dat, dat11, datdf, keeploci, corrected_loci, file = "data/sclerotinia_16_loci.rda")
 ```

### 2017-04-14 doc/RMD/RDA-analysis.Rmd
update RDA analysis (I think?)

@@
 library('vegan')
+library('ggrepel')
 ```
@@
 load("data/sclerotinia_16_loci.rda")
-keeploci <- !locNames(dat) %in% colnames(corrected_loci)
 setPop(dat)   <- ~Host/Source/Region/Year
-dat11 <- dat[loc = keeploci, mlg.reset = TRUE]
+setPop(dat11) <- ~Host/Source/Region/Year
 dat11cc <- clonecorrect(dat11, ~Host/Source/Region/Year, keep = 1:4)
@@
 stopifnot(identical(indNames(dat11cc), other(dat11cc)$meta$Isolate))
-ENV11 <- bind_cols(select(other(dat11cc)$meta, Severity), 
-                   select(strata(dat11cc), -MCG))
+stopifnot(identical(indNames(dat16cc), other(dat16cc)$meta$Isolate))
+
+makeENV <- function(DAT, CC){
+  # Creating the data frame with severity
+  META   <- select(other(DAT)$meta, -Isolate)
+  STRATA <- select(strata(DAT), -MCG)
+  MLL    <- data.frame(MLG = mll(DAT))
+  sev    <- bind_cols(META, STRATA, MLL) %>%
+    group_by(Host, Source, Year, Region, MLG) %>% 
+    summarize(Severity = mean(Severity)) %>% # Get mean severity per MLG
+    ungroup()
+  # Ensuring the data is in the correct order
+  META   <- select(other(CC)$meta, -Isolate)
+  STRATA <- select(strata(CC), -MCG)
+  MLL    <- data.frame(MLG = mll(CC))
+  bind_cols(META, STRATA, MLL) %>% 
+    left_join(sev)
+}
+
+ENV11 <- makeENV(dat11, dat11cc)
 ENV11
-stopifnot(identical(indNames(dat16cc), other(dat16cc)$meta$Isolate))
-ENV16 <- bind_cols(select(other(dat16cc)$meta, Severity), 
-                   select(strata(dat16cc), -MCG))
+stopifnot(identical(ENV11$MLG, mll(dat11cc)))
+ENV11 <- select(ENV11, -MLG)
+ENV16 <- makeENV(dat, dat16cc)
 ENV16
+stopifnot(identical(ENV11$MLG, mll(dat11cc)))
+ENV16 <- select(ENV16, -MLG)
 ```
 
-## Calculate Bruvo's distance
+## Function to tie everything together
 
 ```{r}
-bruvo11cc <- bruvo.dist(dat11cc, replen = other(dat)$REPLEN)
-bruvo16cc <- bruvo.dist(dat16cc, replen = other(dat)$REPLEN)
+choose_dbrda <- function(bdist, ENV, CHOOSER = "ordistep", ...){
+  mod0  <- capscale(bdist ~ 1, data = ENV, add = TRUE)
+  mod1  <- capscale(bdist ~ ., data = ENV, add = TRUE)
+  CHOOSER <- match.fun(CHOOSER)
+  the_model <- CHOOSER(mod0, scope = formula(mod1), ...)
+  return(the_model)
+}
+arrowMul <- function(arrows, data, at = c(0, 0), fill = 0.75) {
+    u <- c(range(data[,1], range(data[,2])))
+    u <- u - rep(at, each = 2)
+    r <- c(range(arrows[, 1], na.rm = TRUE), range(arrows[, 2], na.rm = TRUE))
+    rev <- sign(diff(u))[-2]
+    if (rev[1] < 0)
+        u[1:2] <- u[2:1]
+    if (rev[2] < 0)
+        u[3:4] <- u[4:3]
+    u <- u/r
+    u <- u[is.finite(u) & u > 0]
+    fill * min(u)
+}
+plot_dbrda <- function(db){
+  dbsum     <- scores(db, display = c("cn", "bp", "sites"), scaling = "sites")
+  Centroids <- as.data.frame(dbsum$centroids)
+  Centroids <- rownames_to_column(Centroids, var = "cent_type")
+  Centroids <- mutate_(Centroids, .dots = list(Length = ~sqrt(CAP1^2 * CAP2^2)))
+  # Centroids
+  SampleCentroids <- rownames_to_column(data.frame(dbsum$sites), var = "isolate_names")
+
+  mul     <- arrowMul(dbsum$biplot[, 1:2], dbsum$sites) * 10
+  Arrows  <- data.frame(dbsum$biplot * mul)
+  Arrows  <- rownames_to_column(Arrows, var = "class")
+  Arrows  <- mutate_(Arrows, .dots = list(Length = ~sqrt(CAP1^2 * CAP2^2)))
+  Arrows  <- arrange(Arrows, Length)
+  Arrows  <- top_n(Arrows, 8)
+  ggplot(DecendCentro, aes(x = CAP1, y = CAP2))+
+   geom_point(data = site_centroids, alpha = 1/2, fill = "dark orange", color = "black", size = 2.5, pch = 21)+
+   coord_cartesian() +
+   geom_segment(aes(x = 0, xend = CAP1, 
+                    y = 0, yend = CAP2),
+                arrow = arrow(length = unit(0.3, "cm")), 
+                data = Arrows
+                ) + 
+   geom_label_repel(aes(x = CAP1, y = CAP2, label = class), data = Arrows)
+}
+```
+
+# Calculations
+
+```{r rda_calculation, cache = TRUE}
+dat11cc.bruvo <- dat11cc %>% bruvo.dist(replen = other(.)$REPLEN)
+cap11cc       <- choose_dbrda(dat11cc.bruvo, ENV = ENV11, CHOOSER = "ordistep")
+dat16cc.bruvo <- dat16cc %>% bruvo.dist(replen = other(.)$REPLEN)
+cap16cc       <- choose_dbrda(dat16cc.bruvo, ENV = ENV16, CHOOSER = "ordistep")
 ```
@@
 ```{r}
-commtab <- mlg.table(dat11cc, plot = FALSE)
-cap11cc <- capscale(bruvo11cc ~ Region + Host + Year + Severity, 
-                    data = ENV11, 
-                    comm = , 
-                    add = TRUE)
-
+plot_dbrda(cap11cc) + theme_bw() + theme(text = element_text(size = 14))
+plot_dbrda(cap16cc) + theme_bw() + theme(text = element_text(size = 14))
 ```
-
-```{r}
-capsum <- summary(cap11cc)
-summary(summary(cap11cc))
-##########Extracting 10 most important centroids for factor constrains 
-###############It doesn't matter how you start the ggplot aesthetics. Can be DecendCentro CAP1 and CAP2 or site_centroids
-############### CAP1 and CAP2. Same graph is resulted
-
-Centroids <- data.frame(capsum$centroids) #no need to convert to a dataframe
-LenCent <- sqrt((Centroids[,1]^2)*(Centroids[,2]^2))   ###  This calculates the arrow length (ie. strength of the eigenvector)
-NewCentro <- cbind(Centroids, LenCent)
-DecendCentro <- head(NewCentro[order(NewCentro[,7], decreasing = T),], n=23)   ###  this finds the n strongest eigenvectors by ordering them
-#above "head" is correct
-
-DecendCentro$cent_type <- rownames(head(DecendCentro, n=23))
-DecendCentro <- head(DecendCentro, n = 10)
-DecendCentro #constraint factor cetroids
-
-## Testing plot in ggplot
-
-ggplot(data= DecendCentro, mapping = aes(x = CAP1, y = CAP2))+
-  geom_text(aes(label = cent_type))+ # label tells geom_text which text you want to plot
-  ylim(-3.5,3.5)+ #NB note that this is a convenience wrapper and may cut data out of your plot
-  xlim(-3,3.5) #important if you are calculating stats in the plot - these will be excluded
-
-# sites/isolates
-site_centroids <- data.frame(capsum$sites)
-site_centroids$isolate_names <- rownames(site_centroids) # you can also put label = rownames() etc in the ggplot code. But sometimes you may want it in a column for later
-
-##plotting both main centroids and isolates together
-# I am plotting isolates firest and then envt centroids, alpha = transparancy
-ggplot(DecendCentro, aes(x = CAP1, y = CAP2))+
-  geom_point(data = site_centroids, colour = "orange", size =3.5, alpha =0.7)+
-  geom_text(aes(label = cent_type))+
-  ylim(-4,3.5)+
-  xlim(-3.5,3.5)
-
-#######strongest arrows: THESE ARE BIPLOT scores#############
-
-arrws <- data.frame(capsum$biplot)
-lenarrw <- sqrt(arrws[,1]^2*arrws[,2]^2)   ###  This calculates the arrow length (ie. strength of the eigenvector)
-arrws <- cbind(arrws, lenarrw)
-arrws <- head(arrws[order(arrws[,7], decreasing = T),], n=10)   ###  this finds the n strongest eigenvectors by ordering them
-arrws
-
-arrws$class <- (rownames(arrws)) #turning rownames into a variable
-mult <- attributes(capsum)$const/2.5 # 97% sure this is the scaling for the arrows, pls let me know if I'm wrong
-arrws$class
-#arrws$class <- c("MX", "ND", "CA", "mtc", "NE", "MCG ")  # change this if you want to change how names are displayed
-#arrws$class <- c("MX", "mtc", "NE", "ND", "MCG", "CA")
-##locations abbreviations; see the original for full names
-
-#####plotting centroids, isolates and biplot arrows together#########
-##@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@@########
-
-arrws$class <- c("Mexico", "2005", "CA", "Vista", "CO", "Host: unknown", "OR", "Severity", "MN", "WA")
-mult <- attributes(capsum)$const/3
-
-(Cap_plot <- ggplot(DecendCentro, aes(x = CAP1, y = CAP2))+
-   geom_text(aes(label = cent_type), colour = "blue")+
-   geom_point(data = site_centroids, alpha=1/2, colour = "dark orange", size = 2.5, )+
-   #coord_cartesian(x = c(-3.5, 3.5), y = c(-4.5, 2.5))+# For bruvo
-   coord_cartesian(x = c(-3.5, 3.5), y = c(-2.5, 5))+
-   geom_segment(data = arrws,
-                aes(x = 0, xend = mult * CAP1, y = 0, yend = mult * CAP2),
-                arrow = arrow(length = unit(0.3, "cm")), colour = "dark grey") + #grid is required for arrow to work.
-   geom_text(data = arrws, color = "black",
-             aes(x= (mult + mult/10) * CAP1, y = (mult + mult/10) * CAP2, label = arrws$class), #otherwise you could use hjust and vjust. I prefer this option
-             size = 5, hjust = 0.5)+ theme_bw())
-```

### 2017-04-17 doc/RMD/RDA-analysis.Rmd
Add documentation for the code

@@
 
+So, here Sajeewa clone-corrected the data according to the combination of
+Host, Source (aka Field), Region (aka State/Country), and Year.
+
 ```{r, load_data}
@@
 dat16cc
-# environmental variables for dbRDA
+# Asserting that nothing messed up with the metadata.
 stopifnot(identical(indNames(dat11cc), other(dat11cc)$meta$Isolate))
@@
 
+# function to gather Environmental variables, but averaging Severity
+# (aka Straw Test, Virulence). This function is necessary because the
+# data tends to randomly shuffle when being processed vai dplyr functions
+# for better or for worse. ¯\_(ツ)_/¯
+# 
+# @param DAT the full data set
+# @param CC The clone-corrected data set
+# 
+# @return A data frame containing Severity, Host, Source, Year, Region, and MLG
 makeENV <- function(DAT, CC){
@@
 
+# Get environmental variables for 11 loci
 ENV11 <- makeENV(dat11, dat11cc)
@@
 ENV11 <- select(ENV11, -MLG)
+
+# Get environmental variables for 16 loci
 ENV16 <- makeENV(dat, dat16cc)
 ENV16
-stopifnot(identical(ENV11$MLG, mll(dat11cc)))
+stopifnot(identical(ENV16$MLG, mll(dat16cc)))
 ENV16 <- select(ENV16, -MLG)
@@
 
-## Function to tie everything together
+## Functions to tie everything together
 
-```{r}
+The analysis has a couple of steps
+
+1. Model choice. Since we don't want to overparamaterize the model, we will use 
+*vegan*'s built in model choice function `ordistep()` to to forward-backward 
+selection of the appropriate model to fit our data. Yes, some (particularly 
+Bayesians) believe that model choice is evil and the proper analysis will find 
+the true pattern in the underlying data, but hey, I'm just trying to make sure 
+that I'm not making overconfident judgements.
+
+2. Plot the results of the model choice using ggplot2, overlaying the top 8 
+explanitory vectors
+
+```{r modelfun}
+# model choice for dbrda/capscale
+# 
+# @param bdist a distance matrix (here, we use Bruvo's distances, which is why
+#        it's called 'bdist').
+# @param ENV a data frame of environmental variables with the same number of
+#        observatios as bdist
+# @param CHOOSER the name of the function to perform the forwards-backwards selection
+# @param ... arguments to be passed on to CHOOSER
+# 
+# @return a capscale object
 choose_dbrda <- function(bdist, ENV, CHOOSER = "ordistep", ...){
+  # Step 1: create null model
   mod0  <- capscale(bdist ~ 1, data = ENV, add = TRUE)
+  # Step 2: create full model (all variables in ENV)
   mod1  <- capscale(bdist ~ ., data = ENV, add = TRUE)
+  # Step 3: Run forward-backwards selection (this can take a while)
   CHOOSER   <- match.fun(CHOOSER)
   the_model <- CHOOSER(mod0, scope = formula(mod1), ...)
+  # Return the best model with the anova results
   return(the_model)
 }
+
+# This was ganked from https://github.com/gavinsimpson/ggvegan/blob/59d233977a5b2d15d4de150b782fb1794aa1de8b/R/utils.R
+# 
+# @title Scale Vectors to Data
+# @description Scale vector arrows to \code{fill} proportion of the data.
+# @param arrows a two-column matrix-like object containing coordinates for the arrows/vectors on x and y axes.
+# @param data a two-column matrix-like object containing coordinates of the data on the x and y axes.
+# @param at numeric vector of length 2; location of the origin of the arrows.
+# @param fill numeric; what proportion of the range of the data to fill
+# @return a numeric multiplier that will scale the arrows
+# @author Gavin L. Simpson
 arrowMul <- function(arrows, data, at = c(0, 0), fill = 0.75) {
@@
 
+Here come the calculations. Note, since I am verifying that we get the same
+results from the 16 loci as we do for the 11 loci, we have to do this twice.
+
+1. calculate Bruvo's genetic distance.
+2. model choice
+
 ```{r rda_calculation, cache = TRUE}
+# 11 loci
 dat11cc.bruvo <- dat11cc %>% bruvo.dist(replen = other(.)$REPLEN)
 cap11cc       <- choose_dbrda(dat11cc.bruvo, ENV = ENV11, CHOOSER = "ordistep")
+# 16 loci
 dat16cc.bruvo <- dat16cc %>% bruvo.dist(replen = other(.)$REPLEN)
@@
 
+# Plot the results
 ```{r}

### 2017-04-17 doc/RMD/RDA-analysis.Rmd
fix friggen holdovers from old code
This is why you shouldn't copy and paste, kids.
Your code will work one day, but not the next
because you foolishly changed a variable name.

@@
 }
+# Plotting the dbRDA results
+# 
+# @param db a capscale object
+# @return a ggplot2 object from the scores 
 plot_dbrda <- function(db){
@@
 
-  mul     <- arrowMul(dbsum$biplot[, 1:2], dbsum$sites) * 10
+  mul     <- arrowMul(dbsum$biplot[, 1:2], dbsum$sites)
   Arrows  <- data.frame(dbsum$biplot * mul)
@@
   Arrows  <- top_n(Arrows, 8)
-  ggplot(DecendCentro, aes(x = CAP1, y = CAP2))+
-   geom_point(data = site_centroids, alpha = 1/2, fill = "dark orange", color = "black", size = 2.5, pch = 21)+
+  ggplot(Centroids, aes(x = CAP1, y = CAP2))+
+   geom_point(data = SampleCentroids, alpha = 1/2, fill = "dark orange", color = "black", size = 2.5, pch = 21)+
    coord_cartesian() +

### 2017-04-17 doc/RMD/RDA-analysis.Rmd
update dbRDA figure

@@
 
+  terms   <- paste0("(", paste(attr(db$terms, "term.labels"), collapse = "|"), ")")
   mul     <- arrowMul(dbsum$biplot[, 1:2], dbsum$sites)
@@
   Arrows  <- rownames_to_column(Arrows, var = "class")
+  # Making the classes presentable
+  Arrows$class <- gsub(terms, "\\1: ", Arrows$class)
+  Arrows$class <- gsub(": $", "", Arrows$class)
+  Arrows$class <- gsub("unk", "unknown", Arrows$class)
   Arrows  <- mutate_(Arrows, .dots = list(Length = ~sqrt(CAP1^2 * CAP2^2)))
@@
                  ) + 
-   geom_label_repel(aes(x = CAP1, y = CAP2, label = class), data = Arrows)
+    geom_label_repel(aes(x = CAP1, y = CAP2, label = class), data = Arrows) +
+    ylab("Eigenvalue 1") +
+    xlab("Eigenvalue 2")
 }
@@
 # Plot the results
-```{r}
-plot_dbrda(cap11cc) + theme_bw() + theme(text = element_text(size = 14))
-plot_dbrda(cap16cc) + theme_bw() + theme(text = element_text(size = 14))
+```{r resultplot, fig.width = 5, fig.height = 5}
+cap11cc
+cap11cc$anova
+plot_dbrda(cap11cc) + 
+  theme_classic() + 
+  theme(text = element_text(size = 14, color = "black")) +
+  theme(aspect.ratio = 1)
+ggsave(filename = "results/figures/publication/Figure7Z.svg", width = 88, height = 88, units = "mm", scale = 2)
+cap16cc
+cap16cc$anova
+plot_dbrda(cap16cc) +
+  theme_classic() + 
+  theme(text = element_text(size = 14, color = "black")) +
+  theme(aspect.ratio = 1)
 ```

### 2017-04-17 doc/RMD/RDA-analysis.Rmd
Add variance partitioning
I'm still not entirely sure about this

@@
 ```
+
+## Variance partitioning
+
+First, we want to examine how the entire data fits with the model. We will use
+the alleles as the representative of the entire data.
+
+```{r}
+dat11raw <- genind2df(dat11cc, usepop = FALSE) %>% mutate_all(funs(as.integer(.)))
+dat16raw <- genind2df(dat16cc, usepop = FALSE) %>% mutate_all(funs(as.integer(.)))
+vp11     <- varpart(dat11cc.bruvo, ~Severity + Year + Region + Host, data = ENV11, comm = dat11raw, add = TRUE)
+plot(vp11, Xnames = c("Full Model", "No Model"))
+try(vp16     <- varpart(dat16cc.bruvo, ~Year + Region + Host, data = ENV16, comm = dat16raw, add = TRUE))
+```
+
+```{r}
+vp11 <- varpart(dat11cc.bruvo, ~Severity, ~Year, ~Region, ~Host, data = ENV11, add = TRUE)
+vp16 <- varpart(dat16cc.bruvo, ~Year, ~Region, ~Host, data = ENV16, add = TRUE)
+
+plot(vp11, digits = 2, Xnames = c("Severity", "Year", "Region", "Host"))
+
+plot(vp16, digits = 2, Xnames = c("Year", "Region", "Host"))
+```

### 2017-04-18 doc/RMD/RDA-analysis.Rmd
Add more arrows, make prettier picture

@@
 # @return a ggplot2 object from the scores 
-plot_dbrda <- function(db){
+plot_dbrda <- function(db, arrows = 10){
   dbsum     <- scores(db, display = c("cn", "bp", "sites"), scaling = "sites")
@@
   Arrows  <- arrange(Arrows, Length)
-  Arrows  <- top_n(Arrows, 8)
+  Arrows  <- top_n(Arrows, arrows)
   ggplot(Centroids, aes(x = CAP1, y = CAP2))+
-    geom_point(data = SampleCentroids, alpha = 1/2, fill = "dark orange", color = "black", size = 2.5, pch = 21)+
+    geom_point(data = SampleCentroids, 
+               # alpha = 1/2, 
+               # fill = "white",
+               # fill = "dark orange", 
+               color = "grey45", 
+               size = 2.5, 
+               pch = 21)+
     coord_cartesian() +
@@
                  ) + 
-    geom_label_repel(aes(x = CAP1, y = CAP2, label = class), data = Arrows) +
+    geom_label_repel(aes(x = CAP1, y = CAP2, label = class), 
+                     point.padding = unit(0.5, "lines"),
+                     segment.color = "grey25",
+                     data = Arrows) +
     ylab("Eigenvalue 1") +
@@
 cap11cc$anova
+set.seed(99)
 plot_dbrda(cap11cc) + 
-  theme_classic() + 
-  theme(text = element_text(size = 14, color = "black")) +
+  theme_classic(base_size = 16, base_family = "Helvetica") + 
+  theme(axis.text = element_text(color = "black")) +
   theme(aspect.ratio = 1)
-ggsave(filename = "results/figures/publication/Figure7Z.svg", width = 88, height = 88, units = "mm", scale = 2)
+ggsave(filename = "results/figures/publication/Figure7Z.svg", width = 88, height = 88, units = "mm", scale = 1.5)
 cap16cc
@@
 plot_dbrda(cap16cc) +
-  theme_classic() + 
-  theme(text = element_text(size = 14, color = "black")) +
+  theme_classic(base_size = 16, base_family = "Helvetica") + 
+  theme(axis.text = element_text(color = "black")) +
   theme(aspect.ratio = 1)

### 2017-04-18 doc/RMD/RDA-analysis.Rmd
Add option to make wmn binary
This did not change the results of the analysis.

@@
 # @param CC The clone-corrected data set
+# @param wmn should the "Source" column be converted to a binary wmn
+#   factor? When this is true, everything that is not wmn will be converted
+#   to "other"
 # 
 # @return A data frame containing Severity, Host, Source, Year, Region, and MLG
-makeENV <- function(DAT, CC){
+makeENV <- function(DAT, CC, wmn = FALSE){
   # Creating the data frame with severity
@@
   STRATA <- select(strata(DAT), -MCG)
+  STRATA <- if (wmn) mutate(STRATA, Source = ifelse(Source == "wmn", "wmn", "other")) else STRATA
   MLL    <- data.frame(MLG = mll(DAT))
@@
 ```
+
+```{r}
+wmn11 <- seppop(dat11, ~Source)$wmn %>% setPop(~Region)
+wmn11_sub <- popsub(wmn11, blacklist = character(0))
+min_span_net <- bruvo.msn(wmn11_sub, replen = other(wmn11)$REPLEN, add = TRUE, loss = TRUE, showplot = FALSE, include.ties = TRUE)
+set.seed(67)
+plot_poppr_msn(wmn11,
+               min_span_net,
+               inds = "none",
+               mlg = TRUE,
+               gadj = 6,
+               nodelab = 100,
+               nodebase = 1.15,
+               palette = funky,
+               cutoff = (0.5/11)*3,
+               quantiles = FALSE,
+               beforecut = TRUE,
+               layfun = igraph::layout_with_lgl)
+```

### 2017-04-27 doc/RMD/RDA-analysis.Rmd
update RDA analysis

@@
 ```{r, load_data}
-load("data/sclerotinia_16_loci.rda")
+load(file.path(PROJHOME, "data", "sclerotinia_16_loci.rda"))
 setPop(dat)   <- ~Host/Source/Region/Year
@@
   SampleCentroids <- rownames_to_column(data.frame(dbsum$sites), var = "isolate_names")
-
+  labs    <- summary(db, axes = 0)[["cont"]][["importance"]]["Proportion Explained", 1:2]
   terms   <- paste0("(", paste(attr(db$terms, "term.labels"), collapse = "|"), ")")
@@
                      data = Arrows) +
-    ylab("Eigenvalue 1") +
-    xlab("Eigenvalue 2")
+    xlab(paste0("Eig 1 (", round(labs[[1]]*100, 2), "% variance explained)")) +
+    ylab(paste0("Eig 2 (", round(labs[[2]]*100, 2), "% variance explained)"))
 }
@@
 # Plot the results
+
 ```{r resultplot, fig.width = 5, fig.height = 5}
@@
   theme(aspect.ratio = 1)
-ggsave(filename = "results/figures/publication/Figure7Z.svg", width = 88, height = 88, units = "mm", scale = 1.5)
+FILE <- file.path(PROJHOME, "results", "figures", "publication", "Figure7Z.svg")
+ggsave(filename = FILE, width = 88, height = 88, units = "mm", scale = 1.5)
 cap16cc
@@
   theme(aspect.ratio = 1)
+```
+
+This is all well and good, but what exactly does it actually mean to have only 
+1.93 variance explained (for 11 loci, that is)? I was a bit stumped as well. 
+Luckily, Gavin Simpson has some answers: http://stackoverflow.com/a/22537820/2752888
+
+Basically, we have a situation where we have all of the variables only
+explaining a whopping 16 - 17% of the variance and the rest goes unexplained.
+
+```{r, echo = FALSE, results = "asis"}
+cap11sum <- summary(cap11cc, axes = 0, scaling = "sites")
+cap16sum <- summary(cap16cc, axes = 0, scaling = "sites")
+vartab <- function(x){
+  tot <- x$tot.chi
+  con <- x$constr.chi
+  uco <- x$unconst.chi
+  data.frame(list(Inertia = c(tot, con, uco),
+                  Proportion = c(tot, con, uco)/tot))
+}
+cat("\n\n11 loci:\n\n")
+knitr::kable(vartab(cap11sum), digits = 3)
+cat("\n\n16 loci:\n\n")
+knitr::kable(vartab(cap16sum), digits = 3)
 ```

### 2017-04-27 doc/RMD/RDA-analysis.Rmd
add some text
add session info

@@
 
+Out of both of the models, only Region explains the most variance, almost double
+that of Host or Year and Severity contributes almost nothing.
+
 ```{r}
@@
 ```
+
+```{r, echo = FALSE, results = "asis"}
+options(width = 100)
+cat("<details><summary>Session Information</summary>\n```\n")
+devtools::session_info()
+cat("```\n</details>")
+```

### 2017-04-27 doc/RMD/RDA-analysis.Rmd
fix boo boo

--- 2017-04-27 doc/RMD/RDA-analysis.Rmd
@@
 ```
+
+```{r, echo = FALSE, results = "asis"}
+options(width = 100)
+cat("<details><summary>Session Information</summary>\n\n```\n\n")
+devtools::session_info()
+cat("\n\n```\n\n</details>")
+```

### 2017-04-27 doc/RMD/RDA-analysis.Rmd
add ANOVA

@@
 
+# ANOVA
+
+The results of our modeling and the marginal effects for each parameter on the
+model itself.
+
+```{r}
+cap11cc
+cap11cc$anova
+cap16cc
+cap16cc$anova
+```
+
+We can test the effects of how much of the model explains the variance
+
+```{r total_anova, cache = TRUE}
+anova(cap11cc)
+anova(cap16cc)
+```
+
+We can also see the marginal effects with `by = "margin"`
+
+```{r marginal_anova, cache = TRUE}
+anova(cap11cc, by = "margin")
+anova(cap16cc, by = "margin")
+```
+
 # Plot the results
@@
 ```{r resultplot, fig.width = 5, fig.height = 5}
-cap11cc
-cap11cc$anova
 set.seed(99)
@@
 ggsave(filename = FILE, width = 88, height = 88, units = "mm", scale = 1.5)
-cap16cc
-cap16cc$anova
 plot_dbrda(cap16cc) +

### 2017-05-02 doc/RMD/pop-diff.Rmd
add popdiff

---
title: "Population Differentiation Analysis"
output: 
  html_notebook:
    toc: true
---

```{r setup, include = FALSE}
knitr::opts_knit$set(root.dir = PROJHOME)
```

## Packages and Data

```{r load_packages}
library('tidyverse')
library('poppr')
library('ggcompoplot')
```

```{r load_data}
load(file.path(PROJHOME, "data", "sclerotinia_16_loci.rda"))
setPop(dat11) <- ~Host/Source/Region/Year
dat11cc <- clonecorrect(dat11, ~Host/Source/Region/Year, keep = 1:4)
dat11cc

# Asserting that nothing messed up with the metadata.
stopifnot(identical(indNames(dat11cc), other(dat11cc)$meta$Isolate))
dat11.bruvo   <- bruvo.dist(dat11, replen = other(dat11)$REPLEN)
dat11cc.bruvo <- bruvo.dist(dat11cc, replen = other(dat11)$REPLEN)
```

# AMOVA

First, we can conduct an AMOVA analysis across all populations, with respect to
year.

```{r amova-no-cc}
(res   <- poppr.amova(dat11cc, ~Region/Host/Year, dist = dat11cc.bruvo))
set.seed(999)
(res.p <- randtest(res, nrepet = 999))
```

## DAPC of Region

This can tell us how well our regions separate

```{r dapc-region, fig.width = 7, fig.height = 7}
region.dapc <- dapc(dat11, strata(dat11)$Region, n.pca = 25, n.da = 14)
scatter(region.dapc)
ggcompoplot(region.dapc, setPop(dat11, ~Region), pal = funky, cols = 2)
```

## DAPC predictions

Since we saw MLGs crossing populations, we want to find out what happens when
we remove these MLGs, run DAPC on the regions, and then predict where they came
from.

```{r dapc-mlg, fig.width = 7, fig.height = 7}
minds <- mlg.crosspop(dat11, ~Region, indexreturn = TRUE)
dat11.nocross <- dat11[!mll(dat11) %in% minds]
dat11.nocross
dat11.nc.dapc <- dapc(dat11.nocross, strata(dat11.nocross)$Region, n.pca = 11, n.da = 14)
scatter(dat11.nc.dapc)
dat11.cross <- clonecorrect(dat11[mll(dat11) %in% minds], NA)
pred <- predict.dapc(dat11.nc.dapc, dat11.cross)
ggcompoplot(pred, setPop(dat11.cross, ~Region), cols = 2, pal = funky)
posterior <- as.data.frame(pred$posterior) %>% 
  rownames_to_column("Sample") %>%
  bind_cols(data_frame(MLG = mll(dat11.cross))) %>%
  gather(key = "population", value = "posterior", -Sample, -MLG) %>%
  group_by(Sample) %>%
  mutate(entropy = vegan::diversity(posterior)) %>%
  mutate(label = ifelse(posterior == max(posterior), MLG, NA)) %>%
  ungroup() %>%
  mutate(bins = cut(entropy, quantile(entropy)))
ggplot(posterior, aes(x = population, y = posterior, color = entropy, group = Sample)) +
  geom_line() +
  ggrepel::geom_label_repel(aes(label = label)) +
  viridis::scale_color_viridis() +
  facet_wrap(~bins, ncol = 1)
```

```{r}
top5 <- table(mll(dat11)) %>% 
  sort(decreasing = TRUE) %>% 
  head(5) %>%
  names() %>%
  as.integer()
top5
center_point <- function(n){
  fromto <- c(-1, 1)
  if (n < 4){
    fromto <- fromto/8
  } else if (n < 10){
    fromto <- fromto/4
  } else {
    fromto <- fromto/2
  }
  seq(from = fromto[1], to = fromto[2], length.out = n)
}
origins <- mlg.crosspop(dat11, ~Region, mlgsub = top5, quiet = TRUE, df = TRUE) %>%
  mutate(MLG = gsub("MLG.", "", MLG)) %>%
  mutate(population = match(as.character(Population), posterior$population %>% unique %>% sort)) %>%
  # group_by(Population) %>%
  # mutate(population = population + center_point(n())) %>%
  I()
posterior %>%
  filter(MLG %in% top5) %>%
  ggplot(aes(x = population, y = posterior, color = factor(MLG), group = MLG)) +
  geom_line() +
  geom_vline(aes(xintercept = population), data = origins, lty = 2) +
  scale_color_brewer(palette = "Dark2") +
  facet_wrap(~MLG, ncol = 1) +
  labs(list(
    lty = "Populations discovered in",
    color = "MLG"
  ))
scatter(dat11.nc.dapc)
```

<details>
<summary>Session Information</summary>

```{r, echo = FALSE}
options(width = 100)
devtools::session_info()
```

</details>

### 2017-05-02 doc/RMD/mlg-mcg.Rmd
Add a fledgling start to the MLG figure

---
title: "Multilocus Genotype And Mycelial Compatability Group Assessment"
output: 
  html_notebook:
    toc: true
---

```{r setup, include = FALSE}
knitr::opts_knit$set(root.dir = PROJHOME)
```

# Introduction

The purpose of this document is to assess the distribution of the MLGs and MCGs
within the data. Specifically, we want to know if there are any MLGs that 
consistently coordinate with a single MCG, or if there are anything close.

## Packages and Data

```{r load_packages}
library('tidyverse')
library('poppr')
library('viridis')
```

```{r load_data}
load(file.path(PROJHOME, "data", "sclerotinia_16_loci.rda"))
load(file.path(PROJHOME, "data", "mlg-crosspop-graph.rda"))
dat11
g11o <- igraph::cluster_optimal(graph11loc$total)
comm <- igraph::communities(g11o)
names(comm) <- c("International", "Costal", "Midwest")
comm
strata(dat11) <- strata(dat11) %>%
  mutate(MLGRegion = case_when(
    .$Region %in% comm$International ~ "International",
    .$Region %in% comm$Costal ~ "Costal",
    .$Region %in% comm$Midwest ~ "Midwest",
    TRUE ~ as.character(.$Region)
  ))

setPop(dat11) <- ~MLGRegion
pal <- unlist(map2(LETTERS[2:4], comm, function(a, b) setNames(viridis::viridis(length(b), option = a), b)))
pal <- c(pal, setNames(rep("#F6F6F6FF", 3), c("Mexico", "ID", "WI")))
```

## MLG table

First, it would be nice to visualize the MLGs across populations

```{r mlg_table, fig.width = 20, fig.height = 5}
mtab <- mlg.table(dat11, color = TRUE)
mplot <- last_plot()
mplot + scale_fill_viridis(discrete = TRUE, direction = -1, option = "C") +
  aes(color = I("black")) 
```

### 2017-05-03 doc/RMD/mlg-mcg.Rmd
Add MCG/MLG graph
And it is HAIRY

@@
 ```{r load_packages}
+library('igraph')
+library('ggraph')
 library('tidyverse')
@@
 
-```{r mlg_table, fig.width = 20, fig.height = 5}
+```{r mlg_table, fig.width = 20, fig.height = 5, keep = "last"}
 mtab <- mlg.table(dat11, color = TRUE)
@@
 ```
+
+Now we can take a look at the concordance of MLGs to MCGs
+
+```{r}
+mll.custom(dat11) <- strata(dat11)$MCG
+mcgmlg <- as.data.frame(table(mll(dat11, "original"), mll(dat11, "custom"))) %>%
+  setNames(c("MLG", "MCG", "Freq")) %>%
+  as_data_frame() %>%
+  filter(Freq > 0)
+mcgs <- mcgmlg %>%
+  group_by(MCG) %>%
+  summarize(MLGs = sum(Freq > 0), Samples = sum(Freq), Entropy = vegan::diversity(Freq), data = list(data_frame(MLG = MLG, Freq = Freq))) %>%
+  arrange(desc(MLGs))
+mlgs <- mcgmlg %>%
+  group_by(MLG) %>%
+  summarize(MCGs = sum(Freq > 0), Samples = sum(Freq), Entropy = vegan::diversity(Freq), data = list(data_frame(MCG = MCG, Freq = Freq))) %>%
+  arrange(desc(Samples), desc(MCGs))
+mcgs
+mlgs
+any(mcgs$MLGs > 1)
+```
+
+We can see that there are plenty of MLGs that contain multiple MCGs. This
+
+```{r}
+mlg.filter(dat11, dist = bruvo.dist, replen = other(dat11)$REPLEN) <- .Machine$double.eps + (0.5/11)
+dat11
+
+fmcgmlg <- as.data.frame(table(mll(dat11, "contracted"), mll(dat11, "custom"))) %>%
+  setNames(c("MLG", "MCG", "Freq")) %>%
+  as_data_frame() %>%
+  filter(Freq > 0)
+fmcgs <- fmcgmlg %>%
+  group_by(MCG) %>%
+  summarize(MLGs = sum(Freq > 0), Samples = sum(Freq), Entropy = vegan::diversity(Freq), data = list(data_frame(MLG = MLG, Freq = Freq))) %>%
+  arrange(desc(MLGs))
+fmlgs <- fmcgmlg %>%
+  group_by(MLG) %>%
+  summarize(MCGs = sum(Freq > 0), Samples = sum(Freq), Entropy = vegan::diversity(Freq), data = list(data_frame(MCG = MCG, Freq = Freq))) %>%
+  arrange(desc(Samples), desc(MCGs))
+fmcgs
+fmlgs
+any(fmcgs$MLGs > 1)
+```
+
+## making a graph
+
+I believe that making a graph to visualize this might help me understand what the h\*ck is going on.
+
+```{r}
+gdf <- mcgmlg %>% 
+  mutate(MLG = paste0('MLG.', MLG))
+MLGS <- gdf %>% 
+  group_by(MLG) %>%
+  summarize(size = sum(Freq)) %>%
+  rename(vertex = MLG)
+MCGS <- gdf %>% 
+  group_by(MCG) %>%
+  summarize(size = sum(Freq)) %>%
+  rename(vertex = MCG)
+VAT <- bind_rows(MLGS, MCGS)
+g <- gdf %>% 
+  select(MCG, MLG) %>%
+  graph_from_data_frame(vertices = VAT)
+osize <- V(g)$size
+```
+
+Because I have more control over the size and feel of the graph, I'm going to use
+ggraph. Of course, since this IS a complicated data set, It's not going to be
+very pretty to look at, but I'm going to save it as supplementary materials
+because it's valuable to at least look this ugliness in the face and say, "Yeah,
+I guess it's not so simple after all."
+
+```{r, fig.width = 10, fig.height = 10}
+
+V(g)$size <- sqrt(osize)/10
+V(g)$type <- ifelse(grepl("MLG", V(g)$name), "Multilocus Genotype", "Mycelial Compatibility Group")
+set.seed(2017-05-03)
+lay2 <- create_layout(g, layout = "igraph", algorithm = "nicely")
+
+mcg_mlg_graph <- ggraph(lay2) +
+  geom_node_circle(aes(r = size, fill = type)) +
+  geom_node_text(aes(label = gsub("MLG.", "", name), color = type, size = size/4), show.legend = FALSE) +
+  geom_edge_link(aes(start_cap = circle(node1.size, unit = "native"), 
+                     end_cap = circle(node2.size, unit = "native")), 
+                 arrow = arrow(length = unit(0.0125, "native"))) +
+  coord_fixed() +
+  scale_color_manual(values = c("white", "black")) +
+  scale_fill_manual(values = c("black", "white")) +
+  theme_graph(base_size = 16, base_family = "Helvetica") +
+  theme(legend.position = "bottom") +
+  ggtitle("Relation of Multilocus Genotypes and MCGs")
+
+mcg_mlg_graph
+ggsave(file.path(PROJHOME, "results/figures/publication/FigureS2.svg"), 
+       width = 88*3, height = 88*3, units = "mm")
+```

### 2017-05-03 doc/RMD/mlg-mcg.Rmd
update to inclue top 5 graph

@@
 
-Now we can take a look at the concordance of MLGs to MCGs
+Now we can take a look at the concordance of MLGs to MCGs. We can do this by 
+creating a contigency table. Of course, since we have 87 and well over 100 MLGs,
+this means that the contingency table is going to be big, so to summarize it
+further, I'm creating two tables, one based on MLGs that will count the number
+of MCGs within each MLG and vice-versa. Of course we lose information like, if
+an MCG contains several MLGs, how can we tell what the abundance is? A handy 
+measure is Evenness, which scales from 0 to 1, indicating how skewed the 
+observations are.
 
@@
   group_by(MCG) %>%
-  summarize(MLGs = sum(Freq > 0), Samples = sum(Freq), Entropy = vegan::diversity(Freq), data = list(data_frame(MLG = MLG, Freq = Freq))) %>%
+  summarize(MLGs = sum(Freq > 0), 
+            Samples = sum(Freq), 
+            Evenness = diversity_stats(Freq)["E.5"], 
+            data = list(data_frame(MLG = MLG, Freq = Freq))) %>%
   arrange(desc(MLGs))
@@
   group_by(MLG) %>%
-  summarize(MCGs = sum(Freq > 0), Samples = sum(Freq), Entropy = vegan::diversity(Freq), data = list(data_frame(MCG = MCG, Freq = Freq))) %>%
+  summarize(MCGs = sum(Freq > 0), 
+            Samples = sum(Freq), 
+            Evenness = diversity_stats(Freq)["E.5"], 
+            data = list(data_frame(MCG = MCG, Freq = Freq))) %>%
   arrange(desc(Samples), desc(MCGs))
@@
 mlgs
-any(mcgs$MLGs > 1)
+
+```
+
+It might be better to visualize these data as barplots. Here we are mapping the
+type (MCG/Count) to color and the opacity (alpha) to Evenness.
+
+```{r barplots, fig.width = 10, fig.height = 4}
+mcgs %>% 
+  gather(type, count, MLGs, Samples, -Evenness) %>%
+  arrange(desc(type), desc(count)) %>%
+  mutate(MCG = forcats::fct_inorder(MCG, ordered = TRUE)) %>%
+  ggplot(aes(x = MCG, y = count, group = type, fill = type, alpha = Evenness)) +
+  geom_col(aes(width = ifelse(type == "MLGs", 0.5, 0.85)), color = "black", position = "identity") +
+  annotate(geom = "text", x = 13, y = 51, label = sprintf("Mean Evenness: %.3f", mean(mcgs$Evenness, na.rm = TRUE))) +
+  scale_fill_viridis(end = 0.75, discrete = TRUE, direction = -1) +
+  scale_y_continuous(expand = c(0, 2)) +
+  theme_minimal() +
+  theme(panel.grid.major.x = element_blank()) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0, hjust = 1)) +
+  ggtitle("How Evenly are Multilocus Genotypes (MLGs) spread across MCGs?")
+mlgs %>% 
+  gather(type, count, MCGs, Samples, -Evenness) %>%
+  arrange(desc(type), desc(count)) %>%
+  mutate(MLG = forcats::fct_inorder(MLG, ordered = TRUE)) %>%
+  ggplot(aes(x = MLG, y = count, group = type, fill = type)) +
+  geom_col(aes(width = ifelse(type == "MCGs", 0.5, 0.85), alpha = Evenness), color = "black", position = "identity") +
+  annotate(geom = "text", x = 20, y = 21, label = sprintf("Mean Evenness: %.3f", mean(mlgs$Evenness, na.rm = TRUE))) +
+  scale_fill_viridis(end = 0.75, discrete = TRUE, direction = -1) +
+  scale_y_continuous(expand = c(0, 2)) +
+  theme_minimal() +
+  theme(panel.grid.major.x = element_blank()) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0, hjust = 1)) +
+  ggtitle("How Evenly are MCGs spread across Multilocus Genotypes (MLGs)?")
 ```
@@
 ```{r}
-mlg.filter(dat11, dist = bruvo.dist, replen = other(dat11)$REPLEN) <- .Machine$double.eps + (0.5/11)
+mlg.filter(dat11, dist = bruvo.dist, replen = other(dat11)$REPLEN) <- .Machine$double.eps + (3*(0.5/11))
 dat11
@@
 ```{r}
-gdf <- mcgmlg %>% 
-  mutate(MLG = paste0('MLG.', MLG))
+make_mcgmlg_graph <- function(x){
+  gdf <- mutate(x, MLG = paste0('MLG.', MLG))
   MLGS <- gdf %>% 
@@
     graph_from_data_frame(vertices = VAT)
+  V(g)$type <- ifelse(grepl("MLG", V(g)$name), "Multilocus Genotype", "Mycelial Compatibility Group")
+  g
+}
+g <- make_mcgmlg_graph(mcgmlg)
+gf <- make_mcgmlg_graph(fmcgmlg)
 osize <- V(g)$size
+fosize <- V(gf)$size
 ```
@@
 V(g)$size <- sqrt(osize)/10
-V(g)$type <- ifelse(grepl("MLG", V(g)$name), "Multilocus Genotype", "Mycelial Compatibility Group")
+V(gf)$size <- sqrt(fosize)/10
 set.seed(2017-05-03)
 lay2 <- create_layout(g, layout = "igraph", algorithm = "nicely")
+flay2 <- create_layout(gf, layout = "igraph", algorithm = "nicely")
 
@@
 mcg_mlg_graph
+mcg_mlg_graph %+% flay2
 ggsave(file.path(PROJHOME, "results/figures/publication/FigureS2.svg"), 
@@
 ```
+
+## Subgraph
+
+So that's a big hairy graph. What happens when we look at a subgraph of the
+top 5 MCGs?
+
+```{r top5graph, fig.height = 7, fig.width = 7}
+top5 <- filter(mcgmlg, as.character(MCG) %in% mcgs$MCG[1:5])
+top5g <- make_mcgmlg_graph(top5)
+tosize <- V(top5g)$size
+V(top5g)$size <- sqrt(tosize)/10
+set.seed(2017-05-03)
+top5lay <- create_layout(top5g, layout = "igraph", algorithm = "nicely")
+mcg_mlg_graph %+% top5lay + ggtitle("Top 5 Mycelial Compatibility Groups and associated MLGs")
+```
+
+Vey nice!

### 2017-05-03 doc/RMD/mlg-mcg.Rmd
add severity graph

@@
 Vey nice!
+
+There are a whole buttload of MLGs for those 5 MCGs. What are the severity 
+ratings for those?
+
+```{r severity, fig.width = 7, fig.height = 7}
+
+strat <- bind_cols(strata(dat11), 
+                   other(dat11)$meta, 
+                   data_frame(MLG = mll(dat11, "original")))
+Severity <- filter(strat, MCG %in% mcgs$MCG[1:5])
+ggplot(Severity, aes(x = MCG, y = Severity)) +
+  geom_point(aes(fill = ifelse(MLG %in% mlgs$MLG[1:5], MLG, "Other")), 
+             position = position_jitter(width = 0.25),
+             alpha = 0.75,
+             pch = 21) +
+  scale_fill_viridis(discrete = TRUE, direction = -1) +
+  facet_wrap(~Region, ncol = 2) +
+  labs(list(
+    title = "Severity by MCG and Region",
+    fill = "Multilocus Genotype",
+    caption = "Five most abundant multilocus genotypes shown"
+  ))
+```

### 2017-05-04 doc/RMD/mlg-mcg.Rmd
update mlgmcg

--- 2017-05-03 doc/RMD/mlg-mcg.Rmd
@@
 
-We can see that there are plenty of MLGs that contain multiple MCGs. This
-
-```{r}
-mlg.filter(dat11, dist = bruvo.dist, replen = other(dat11)$REPLEN) <- .Machine$double.eps + (3*(0.5/11))
-dat11
-
-fmcgmlg <- as.data.frame(table(mll(dat11, "contracted"), mll(dat11, "custom"))) %>%
-  setNames(c("MLG", "MCG", "Freq")) %>%
-  as_data_frame() %>%
-  filter(Freq > 0)
-fmcgs <- fmcgmlg %>%
-  group_by(MCG) %>%
-  summarize(MLGs = sum(Freq > 0), Samples = sum(Freq), Entropy = vegan::diversity(Freq), data = list(data_frame(MLG = MLG, Freq = Freq))) %>%
-  arrange(desc(MLGs))
-fmlgs <- fmcgmlg %>%
-  group_by(MLG) %>%
-  summarize(MCGs = sum(Freq > 0), Samples = sum(Freq), Entropy = vegan::diversity(Freq), data = list(data_frame(MCG = MCG, Freq = Freq))) %>%
-  arrange(desc(Samples), desc(MCGs))
-fmcgs
-fmlgs
-any(fmcgs$MLGs > 1)
-```
-
 ## making a graph
@@
   g <- gdf %>% 
-    select(MCG, MLG) %>%
+    select(MCG, MLG, Freq) %>%
+    rename(weight = Freq) %>%
     graph_from_data_frame(vertices = VAT)
@@
 g <- make_mcgmlg_graph(mcgmlg)
-gf <- make_mcgmlg_graph(fmcgmlg)
 osize <- V(g)$size
-fosize <- V(gf)$size
 ```
@@
 V(g)$size <- sqrt(osize)/10
-V(gf)$size <- sqrt(fosize)/10
-set.seed(2017-05-03)
+set.seed(2017-05-05)
 lay2 <- create_layout(g, layout = "igraph", algorithm = "nicely")
-flay2 <- create_layout(gf, layout = "igraph", algorithm = "nicely")
 
@@
   geom_node_circle(aes(r = size, fill = type)) +
-  geom_node_text(aes(label = gsub("MLG.", "", name), color = type, size = size/4), show.legend = FALSE) +
   geom_edge_link(aes(start_cap = circle(node1.size, unit = "native"), 
-                     end_cap = circle(node2.size, unit = "native")), 
-                 arrow = arrow(length = unit(0.0125, "native"))) +
+                     end_cap = circle(node2.size, unit = "native"), 
+                     width = weight),
+                 arrow = arrow(length = unit(0.01, "native"))) +
+  geom_node_text(aes(label = gsub("MLG.", "", name), color = type, size = size/10), show.legend = FALSE) +
   coord_fixed() +
+  scale_edge_width(range = c(0.25, 1.5)) +
+  # scale_edge_color_viridis(option = "C", end = 0.9) +
   scale_color_manual(values = c("white", "black")) +
@@
   theme_graph(base_size = 16, base_family = "Helvetica") +
-  theme(legend.position = "bottom") +
+  theme(legend.position = "bottom", legend.direction = "vertical") +
   ggtitle("Relation of Multilocus Genotypes and MCGs")
@@
 mcg_mlg_graph
-mcg_mlg_graph %+% flay2
 ggsave(file.path(PROJHOME, "results/figures/publication/FigureS2.svg"), 
@@
 Vey nice!
+
+There are a whole buttload of MLGs for those 5 MCGs. What are the severity 
+ratings for those? Again, we take take a look at these AND simultaneously query
+the top 5 MLGs for this.
+
+```{r severity, fig.width = 7, fig.height = 4}
+
+strat <- bind_cols(strata(dat11), 
+                   other(dat11)$meta, 
+                   data_frame(MLG = mll(dat11, "original")))
+count_group <- . %>% 
+  mutate(nobs = n()) %>%  # count the number of samples/MCG
+  ungroup() %>%
+  arrange(desc(nobs)) # arrange by number of samples and reorder factors 
+Severity <- filter(strat, MCG %in% mcgs$MCG[1:5]) %>%
+  group_by(MCG) %>% 
+  count_group %>%
+  mutate(MCG = forcats::fct_inorder(factor(MCG), ordered = TRUE)) %>%
+  mutate(MLG = ifelse(MLG %in% mlgs$MLG[1:5], paste("MLG", MLG), "Other"))
+
+severity_plot <- ggplot(Severity, aes(x = MCG, y = Severity)) +
+  geom_point(aes(fill = MLG), 
+             position = position_jitter(width = 0.2),
+             alpha = 0.75,
+             pch = 21) +
+  scale_fill_viridis(discrete = TRUE, direction = -1) +
+  theme_bw() +
+  theme(legend.position = "bottom") +
+  theme(aspect.ratio = 0.6) +
+  facet_wrap(~Region, nrow = 2) +
+  ylim(c(3.5, 8)) +
+  labs(list(
+    title = "Severity by MCG and Region",
+    fill = "Multilocus Genotype",
+    subtitle = "Five most abundant multilocus genotypes shown"
+  ))
+severity_plot
+ggsave(severity_plot, filename = "results/figures/publication/FigureS3.svg", width = 183, unit = "mm")
+```

### 2017-05-04 doc/RMD/mlg-mcg.Rmd
add subgraph figure

@@
 top5lay <- create_layout(top5g, layout = "igraph", algorithm = "nicely")
-mcg_mlg_graph %+% top5lay + ggtitle("Top 5 Mycelial Compatibility Groups and associated MLGs")
+t5g <- mcg_mlg_graph %+% top5lay + ggtitle("Top 5 Mycelial Compatibility Groups and associated MLGs")
+ggsave(t5g, file = file.path("results", "figures", "publication", "FigureS2B.svg"), width = 88, height = 88, units = "mm")
 ```

### 2017-05-04 doc/RMD/mlg-mcg.Rmd
change scale

--- 2017-05-04 doc/RMD/mlg-mcg.Rmd
@@
 top5lay <- create_layout(top5g, layout = "igraph", algorithm = "nicely")
-mcg_mlg_graph %+% top5lay + ggtitle("Top 5 Mycelial Compatibility Groups and associated MLGs")
+t5g <- mcg_mlg_graph %+% top5lay + ggtitle("Top 5 Mycelial Compatibility Groups and associated MLGs")
+ggsave(t5g, file = file.path("results", "figures", "publication", "FigureS2B.svg"), width = 88, height = 88, scale = 3, units = "mm")
 ```

### 2017-05-04 doc/RMD/mlg-mcg.Rmd
reset seed

--- 2017-05-04 doc/RMD/mlg-mcg.Rmd
@@
 V(top5g)$size <- sqrt(tosize)/10
-set.seed(2017-05-03)
+set.seed(2017-05-04)
 top5lay <- create_layout(top5g, layout = "igraph", algorithm = "nicely")
-mcg_mlg_graph %+% top5lay + ggtitle("Top 5 Mycelial Compatibility Groups and associated MLGs")
+t5g <- mcg_mlg_graph %+% top5lay + ggtitle("Top 5 Mycelial Compatibility Groups and associated MLGs")
+t5g
+ggsave(t5g, file = file.path("results", "figures", "publication", "FigureS2B.svg"), width = 88, height = 88, scale = 3, units = "mm")
 ```

### 2017-05-04 doc/RMD/mlg-mcg.Rmd
add community detection and indication on graphs

--- 2017-05-04 doc/RMD/mlg-mcg.Rmd
@@
 }
+
+add_communities <- function(g, clusters){
+  comm <- communities(clusters)
+  commlist <- setNames(rep(names(comm), lengths(comm)), unlist(comm, use.names = FALSE))
+  V(g)$community <- commlist[V(g)$name]
+  g
+}
 g <- make_mcgmlg_graph(mcgmlg)
+g <- add_communities(g, cluster_walktrap(g))
 osize <- V(g)$size
@@
 
+the_communities <- data_frame(vertex = lay2$name, community = lay2$community) %>%
+  group_by(community) %>%
+  mutate(comm = ifelse(n() > 10, paste("Community", community), "Other Communities (n < 10)")) %>%
+  mutate(comm = forcats::fct_infreq(comm) %>% forcats::fct_shift(-1L))
+lay2$community <- the_communities$comm
+
 mcg_mlg_graph <- ggraph(lay2) +
-  geom_node_circle(aes(r = size, fill = type)) +
+  geom_node_circle(aes(r = size, lty = type, fill = community)) +
   geom_edge_link(aes(start_cap = circle(node1.size, unit = "native"), 
@@
                  arrow = arrow(length = unit(0.01, "native"))) +
-  geom_node_text(aes(label = gsub("MLG.", "", name), color = type, size = size/10), show.legend = FALSE) +
+  geom_node_text(aes(label = gsub("MLG.", "", name), 
+                     color = ifelse(grepl("ther", community), "A", ifelse(grepl("3", community), "B", "C")), 
+                     size = size/10), 
+                 show.legend = FALSE) +
   coord_fixed() +
   scale_edge_width(range = c(0.25, 1.5)) +
-  # scale_edge_color_viridis(option = "C", end = 0.9) +
-  scale_color_manual(values = c("white", "black")) +
-  scale_fill_manual(values = c("black", "white")) +
+  scale_color_manual(values = c("white", "grey75", "black")) +
+  scale_fill_viridis(direction = -1, discrete = TRUE) +
   theme_graph(base_size = 16, base_family = "Helvetica") +
@@
 top5 <- filter(mcgmlg, as.character(MCG) %in% mcgs$MCG[1:5])
-top5g <- make_mcgmlg_graph(top5)
+top5g <- make_mcgmlg_graph(top5) %>% add_communities(cluster_walktrap(g))
 tosize <- V(top5g)$size
 V(top5g)$size <- sqrt(tosize)/10
-set.seed(2017-05-03)
+set.seed(2017-05-04)
 top5lay <- create_layout(top5g, layout = "igraph", algorithm = "nicely")
+top5lay$community <- inner_join(data_frame(vertex = top5lay$name), the_communities)$comm
 t5g <- mcg_mlg_graph %+% top5lay + ggtitle("Top 5 Mycelial Compatibility Groups and associated MLGs")
-ggsave(t5g, file = file.path("results", "figures", "publication", "FigureS2B.svg"), width = 88, height = 88, units = "mm")
+t5g
+ggsave(t5g, file = file.path("results", "figures", "publication", "FigureS2B.svg"), width = 88, height = 88, scale = 3, units = "mm")
 ```

### 2017-05-04 doc/RMD/mlg-mcg.Rmd
fix group colors

@@
 
+cnames <- LETTERS[1:10]
+cnames[c(7, 5, 1)] <- LETTERS[1:3]
 the_communities <- data_frame(vertex = lay2$name, community = lay2$community) %>%
+  mutate(community = case_when(
+    .$community == 7 ~ "A",
+    .$community == 5 ~ "B",
+    .$community == 1 ~ "C",
+    TRUE ~ as.character(.$community)
+  )) %>%
   group_by(community) %>%
   mutate(comm = ifelse(n() > 10, paste("Community", community), "Other Communities (n < 10)")) %>%
-  mutate(comm = forcats::fct_infreq(comm) %>% forcats::fct_shift(-1L))
+  mutate(comm = comm) # It just so happens that the communities are in order by size.
 lay2$community <- the_communities$comm
-
 mcg_mlg_graph <- ggraph(lay2) +
@@
   geom_node_text(aes(label = gsub("MLG.", "", name), 
-                     color = ifelse(grepl("ther", community), "A", ifelse(grepl("3", community), "B", "C")), 
+                     color = community, 
                      size = size/10), 
@@
   coord_fixed() +
+  scale_linetype_manual(values = c(1, 3)) +
   scale_edge_width(range = c(0.25, 1.5)) +
-  scale_color_manual(values = c("white", "grey75", "black")) +
+  scale_color_manual(values = c("black", "black", "white", "white")) +
   scale_fill_viridis(direction = -1, discrete = TRUE) +
@@
 top5 <- filter(mcgmlg, as.character(MCG) %in% mcgs$MCG[1:5])
-top5g <- make_mcgmlg_graph(top5) %>% add_communities(cluster_walktrap(g))
+top5g <- make_mcgmlg_graph(top5)
 tosize <- V(top5g)$size

### 2017-05-04 doc/RMD/mlg-mcg.Rmd
Add trees to mlg-mcg debacle

@@
 
-cnames <- LETTERS[1:10]
-cnames[c(7, 5, 1)] <- LETTERS[1:3]
 the_communities <- data_frame(vertex = lay2$name, community = lay2$community) %>%
-  mutate(community = case_when(
+  mutate(comm = case_when(
     .$community == 7 ~ "A",
@@
   group_by(community) %>%
-  mutate(comm = ifelse(n() > 10, paste("Community", community), "Other Communities (n < 10)")) %>%
+  mutate(comm = ifelse(n() > 10, paste("Community", comm), "Other Communities (n < 10)")) %>%
   mutate(comm = comm) # It just so happens that the communities are in order by size.
@@
 ```
+
+## Assessing MCG/MLG community structure/cohesiveness
+
+The communities defined above are defined by the connections between different
+MCGs. There is still the question of how cohesive these communities are. I 
+believe one way to investigate this is to add the communities assignments back
+to the data and assess how well they split up on a dendrogram created with 
+genetic distance.
+
+Note in this case we are assigning MLGs to the communities indicated. This means
+that even though they share a connection with an MCG in a different community,
+they belong to the MCG community they were initially assigned to. We will also
+assign communities based on what MCG they belong to.
+
+```{r treebuilding, cache = TRUE}
+the_communities
+strat_mlg <- strat %>% 
+  mutate(MLG = paste0("MLG.", MLG)) %>%
+  left_join(the_communities, by = c("MLG" = "vertex"))
+strat_mcg <- strat %>% 
+  left_join(the_communities, by = c("MCG" = "vertex"))
+
+datdist <- bruvo.dist(dat11, replen = other(dat)$REPLEN)
+datree <- phangorn::upgma(datdist)
+```
+
+Now we can visualize that tree.
+
+```{r, fig.width = 7, fig.height = 7}
+ape::plot.phylo(ape::ladderize(datree), type = "fan", no.margin = FALSE, tip.col = viridis(4, direction = -1)[as.integer(factor(strat_mlg$comm))])
+legend(-0.015, 0.11, legend = levels(factor(strat_mlg$comm)), fill = viridis(4, direction = -1), bty = "n")
+title("MLG-based assignment")
+ape::plot.phylo(ape::ladderize(datree), type = "fan", no.margin = FALSE, tip.col = viridis(4, direction = -1)[as.integer(factor(strat_mcg$comm))])
+legend(-0.015, 0.11, legend = levels(factor(strat_mcg$comm)), fill = viridis(4, direction = -1), bty = "n")
+title("MCG-based assignment")
+```
+
+From what I can see, it appears that there's not much of a tight correlation of 
+genetic structure and MCG.
+
+<details>
+<summary>Session Information</summary>
+
+```{r, echo = FALSE}
+options(width = 100)
+devtools::session_info()
+```
+
+</details>

### 2017-05-05 doc/RMD/mlg-mcg.Rmd
fix issue with mlgs

@@
   setNames(c("MLG", "MCG", "Freq")) %>%
+  mutate(MLG = as.character(MLG)) %>%
+  mutate(MCG = as.character(MCG)) %>%
   as_data_frame() %>%
@@
 mcgs <- mcgmlg %>%
+  arrange(MCG) %>%
   group_by(MCG) %>%
-  summarize(MLGs = sum(Freq > 0), 
+  mutate(MLGs = sum(Freq > 0), 
          Samples = sum(Freq), 
          Evenness = diversity_stats(Freq)["E.5"], 
-            data = list(data_frame(MLG = MLG, Freq = Freq))) %>%
-  arrange(desc(MLGs))
+         data = list(data_frame(MLG = MLG, Freq = Freq) %>% arrange(desc(Freq)))) %>%
+  arrange(desc(MLGs)) %>%
+  ungroup() %>%
+  distinct(MCG, MLGs, Samples, Evenness, data)
 mlgs <- mcgmlg %>%
+  arrange(MLG) %>%
   group_by(MLG) %>%
-  summarize(MCGs = sum(Freq > 0), 
+  mutate(MCGs     = sum(Freq > 0), 
          Samples  = sum(Freq), 
          Evenness = diversity_stats(Freq)["E.5"], 
-            data = list(data_frame(MCG = MCG, Freq = Freq))) %>%
-  arrange(desc(Samples), desc(MCGs))
+         data     = list(data_frame(MCG = MCG, Freq = Freq) %>% arrange(desc(Freq)))) %>%
+  arrange(desc(Samples), desc(MCGs)) %>%
+  ungroup() %>%
+  distinct(MLG, MCGs, Samples, Evenness, data)
 mcgs
@@
 
+```
+
+These tables are a good start, but we want to have a publication-ready table:
+
+```{r production-tables, results = 'asis'}
+fct2int <- function(x) as.integer(as.character(x))
+
+psex_from_graph <- function(g, MLG){
+  weights <- setNames(E(g)$weight, E(g)$label)
+  weights <- split(weights, names(weights))[MLG]
+  1 -vapply(weights, unique, numeric(1))
+}
+
+mlg_table <- mlgs %>%
+  slice(1:5L) %>%
+  mutate(MLG = as.character(MLG)) %>%
+  mutate(Psex = psex_from_graph(graph11loc$total, MLG)) %>%
+  unnest() %>%
+  group_by(MLG, Samples, Psex) %>%
+  summarize(MCG = paste(MCG, collapse = ", ")) %>%
+  rename(N = Samples) %>%
+  arrange(desc(N))
+
+kable(mlg_table, caption = "Top 5 MLGs", format.args = list(digits = 3))
+
+mcg_table <- mcgs %>%
+  slice(1:5L) %>%
+  mutate(MCG = as.character(MCG)) %>%
+  unnest() %>%
+  group_by(MCG, Samples, Evenness) %>%
+  summarize(MLG = list(strwrap(paste(MLG, collapse = ", "), 28))) %>%
+  rename(N = Samples) %>%
+  arrange(desc(N)) %>% 
+  unnest() %>%
+  ungroup() %>%
+  mutate(Evenness = ifelse(!duplicated(MCG), format(Evenness, digits = 3), "")) %>%
+  mutate(N = ifelse(!duplicated(MCG), N, "")) %>%
+  mutate(MCG = ifelse(!duplicated(MCG), MCG, ""))
+kable(mcg_table, caption = "Top 5 MCGs")
 ```
@@
   theme(panel.grid.major.x = element_blank()) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0, hjust = 1)) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
   ggtitle("How Evenly are Multilocus Genotypes (MLGs) spread across MCGs?")
@@
   theme(panel.grid.major.x = element_blank()) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0, hjust = 1)) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
   ggtitle("How Evenly are MCGs spread across Multilocus Genotypes (MLGs)?")
@@
 V(g)$size <- sqrt(osize)/10
-set.seed(2017-05-05)
+set.seed(2017-05-03)
 lay2 <- create_layout(g, layout = "igraph", algorithm = "nicely")
@@
 V(top5g)$size <- sqrt(tosize)/10
-set.seed(2017-05-04)
+set.seed(2017-05-08)
 top5lay <- create_layout(top5g, layout = "igraph", algorithm = "nicely")

### 2017-05-06 doc/RMD/mlg-mcg.Rmd
add tables and graphs

@@
 pal <- c(pal, setNames(rep("#F6F6F6FF", 3), c("Mexico", "ID", "WI")))
+
+strat <- bind_cols(strata(dat11), 
+                   other(dat11)$meta, 
+                   data_frame(MLG = mll(dat11, "original")))
 ```
@@
 
-These tables are a good start, but we want to have a publication-ready table:
+These tables are a good start, but we want to have a publication-ready table.
 
@@
   weights <- setNames(E(g)$weight, E(g)$label)
-  weights <- split(weights, names(weights))[MLG]
-  1 -vapply(weights, unique, numeric(1))
+  weights <- split(weights, names(weights))
+  if (length(MLG) > 1){
+    weights <- weights[MLG]
+    res <- 1 - vapply(weights, unique, numeric(1))
+  } else {
+    res <- 1 - unique(weights[[MLG]])
+  }
+  res
 }
 
-mlg_table <- mlgs %>%
-  slice(1:5L) %>%
-  mutate(MLG = as.character(MLG)) %>%
-  mutate(Psex = psex_from_graph(graph11loc$total, MLG)) %>%
-  unnest() %>%
-  group_by(MLG, Samples, Psex) %>%
-  summarize(MCG = paste(MCG, collapse = ", ")) %>%
-  rename(N = Samples) %>%
-  arrange(desc(N))
+# Paste region and count vectors
+pasteRC <- function(R, C){
+  ord <- rev(order(C))
+  RC <- paste(R[ord], C[ord], sep = " (")
+  RC <- paste(RC, collapse = "), ")
+  paste0(RC, ")")
+}
 
-kable(mlg_table, caption = "Top 5 MLGs", format.args = list(digits = 3))
+mlg_table <- mcgmlg %>%
+  filter(MLG %in% mlgs$MLG[1:5]) %>%
+  mutate(MLG = as.integer(MLG)) %>%
+  select(-Freq) %>%
+  inner_join(strat) %>%
+  group_by(MLG, MCG, Region) %>%
+  summarize(N = n()) %>%
+  ungroup() %>%
+  group_by(MLG, MCG) %>%
+  summarize(Region = pasteRC(Region, N), N = sum(N)) %>%
+  ungroup() %>%
+  mutate(mlmatch = match(MLG, mlgs$MLG[1:5])) %>%
+  arrange(mlmatch, desc(N), MCG) %>%
+  group_by(MLG) %>%
+  mutate(Psex = psex_from_graph(graph11loc$total, as.character(unique(MLG)))) %>%
+  ungroup() %>%
+  select(MLG, Psex, MCG, Region) %>%
+  mutate(Psex = ifelse(duplicated(MLG), "", format(Psex, scientific = FALSE, digits = 2))) %>%
+  mutate(MLG = ifelse(duplicated(MLG), "", as.character(MLG)))
 
-mcg_table <- mcgs %>%
-  slice(1:5L) %>%
-  mutate(MCG = as.character(MCG)) %>%
-  unnest() %>%
-  group_by(MCG, Samples, Evenness) %>%
-  summarize(MLG = list(strwrap(paste(MLG, collapse = ", "), 28))) %>%
-  rename(N = Samples) %>%
-  arrange(desc(N)) %>% 
-  unnest() %>%
+mcg_table <- mcgmlg %>%
+  filter(MCG %in% mcgs$MCG[1:5]) %>%
+  mutate(MLG = as.integer(MLG)) %>%
+  select(-Freq) %>%
+  inner_join(strat) %>%
+  group_by(MLG, MCG, Region) %>%
+  summarize(N = n()) %>%
   ungroup() %>%
-  mutate(Evenness = ifelse(!duplicated(MCG), format(Evenness, digits = 3), "")) %>%
-  mutate(N = ifelse(!duplicated(MCG), N, "")) %>%
-  mutate(MCG = ifelse(!duplicated(MCG), MCG, ""))
-kable(mcg_table, caption = "Top 5 MCGs")
+  group_by(MLG, MCG) %>%
+  summarize(Region = pasteRC(Region, N), N = sum(N)) %>%
+  ungroup() %>%
+  mutate(mlmatch = match(MCG, mcgs$MCG[1:5])) %>%
+  arrange(mlmatch, desc(N), MLG) %>%
+  select(MCG, MLG, Region) %>%
+  mutate(MCG = ifelse(duplicated(MCG), "", as.character(MCG))) %>%
+  mutate(MLG = as.character(MLG))
+
+kable(mlg_table, caption = "Top 5 MLGs with associated MCGs and Regions (number of samples in parenthesis)")
+if (!dir.exists(file.path(PROJHOME, "results/tables"))) dir.create(file.path(PROJHOME, "results/tables"))
+readr::write_csv(mlg_table, path = file.path(PROJHOME, "results/tables/mlg_table.csv"))
+
+kable(mcg_table, caption = "Top 5 MCGs with associated MLGs and Regions (number of samples in parenthesis)")
+if (!dir.exists(file.path(PROJHOME, "results/tables"))) dir.create(file.path(PROJHOME, "results/tables"))
+readr::write_csv(mcg_table, path = file.path(PROJHOME, "results/tables/mcg_table.csv"))
 ```
@@
   annotate(geom = "text", x = 13, y = 51, label = sprintf("Mean Evenness: %.3f", mean(mcgs$Evenness, na.rm = TRUE))) +
-  scale_fill_viridis(end = 0.75, discrete = TRUE, direction = -1) +
+  # scale_fill_viridis(end = 0.75, discrete = TRUE, direction = -1) +
+  scale_fill_manual(values = c("black", "white")) +
   scale_y_continuous(expand = c(0, 2)) +
@@
   annotate(geom = "text", x = 20, y = 21, label = sprintf("Mean Evenness: %.3f", mean(mlgs$Evenness, na.rm = TRUE))) +
-  scale_fill_viridis(end = 0.75, discrete = TRUE, direction = -1) +
+  # scale_fill_viridis(end = 0.75, discrete = TRUE, direction = -1) +
+  scale_fill_manual(values = c("black", "white")) +
   scale_y_continuous(expand = c(0, 2)) +
@@
 ```{r severity, fig.width = 7, fig.height = 4}
-
-strat <- bind_cols(strata(dat11), 
-                   other(dat11)$meta, 
-                   data_frame(MLG = mll(dat11, "original")))
 count_group <- . %>% 
@@
 
-Note in this case we are assigning MLGs to the communities indicated. This means
-that even though they share a connection with an MCG in a different community,
-they belong to the MCG community they were initially assigned to. We will also
-assign communities based on what MCG they belong to.
-
 ```{r treebuilding, cache = TRUE}
-the_communities
-strat_mlg <- strat %>% 
-  mutate(MLG = paste0("MLG.", MLG)) %>%
-  left_join(the_communities, by = c("MLG" = "vertex"))
-strat_mcg <- strat %>% 
-  left_join(the_communities, by = c("MCG" = "vertex"))
-
+strat_mcg <- left_join(strat, the_communities, by = c("MCG" = "vertex"))
 datdist   <- bruvo.dist(dat11, replen = other(dat)$REPLEN)
@@
 
-Now we can visualize that tree.
+Now we can visualize that trees.
 
-```{r, fig.width = 7, fig.height = 7}
-ape::plot.phylo(ape::ladderize(datree), type = "fan", no.margin = FALSE, tip.col = viridis(4, direction = -1)[as.integer(factor(strat_mlg$comm))])
-legend(-0.015, 0.11, legend = levels(factor(strat_mlg$comm)), fill = viridis(4, direction = -1), bty = "n")
-title("MLG-based assignment")
+```{r trees, fig.width = 7, fig.height = 7}
 ape::plot.phylo(ape::ladderize(datree), type = "fan", no.margin = FALSE, tip.col = viridis(4, direction = -1)[as.integer(factor(strat_mcg$comm))])
@@
 title("MCG-based assignment")
+
+ape::plot.phylo(ape::ladderize(datree), type = "fan", no.margin = FALSE, tip.col = plasma(366)[rank(strat$Severity, ties.method = "first")])
+legend(-0.015, 0.125, legend = quantile(strat$Severity), fill = plasma(5), bty = "n")
+title("Severity")
 ```
@@
 From what I can see, it appears that there's not much of a tight correlation of 
-genetic structure and MCG.
+genetic structure and MCG. There is the clear cluster of MLG 66 in Community C,
+but there are other isolates in distant parts of the tree. I'm wondering what
+happens if we average the distances between groups.
+
+```{r group_dist}
+datmat <- as.matrix(datdist)
+distmcg <- strat_mcg %>% 
+  group_by(comm) %>%
+  summarize(dist = list(as.dist(datmat[Isolate, Isolate])), N = n())
+distmcg %>%
+  group_by(comm) %>%
+  mutate(meandist = mean(dist[[1]]),
+         mediandist = median(dist[[1]])) %>%
+  ungroup() %>%
+  select(comm, N, meandist, mediandist)
+distmcg %>% 
+  unnest() %>%
+  ggplot(aes(x = dist, fill = comm)) +
+  geom_density(alpha = 0.5) +
+  geom_rug(aes(color = comm), alpha = 0.12) +
+  scale_fill_viridis(direction = -1, discrete = TRUE) +
+  scale_color_viridis(direction = -1, discrete = TRUE) +
+  ggtitle("Bruvo's distance by MCG community") +
+  theme_bw(base_size = 14)
+```
+
+This is quite distant. For reference, a distance of `r round((0.5/11)*7, 2)` is
+on average 7 steps.
 

### 2017-05-10 doc/RMD/mlg-mcg.Rmd
add visNetwork
need to work on the presentation, but that can be
done later

@@
 library('viridis')
+library('visNetwork')
 ```
@@
 
+### Interactive Visualizations
+
+The *visNetwork* package wraps a java library that has very good network
+visualization routines. I can conver this graph to that.
+
+```{r}
+
+make_rgb <- function(x, alpha = 1){
+  out <- col2rgb(x, alpha = TRUE)
+  out[4, ] <- floor(out[4, ]*alpha)
+  out      <- apply(out, 2, paste, collapse = ", ")
+  paste0("rgba(", out, ")")
+}
+vg <- g %>%
+  set_vertex_attr("size", value = osize) %>%
+  set_vertex_attr("value", value = osize) %>%
+  set_vertex_attr("color", value = viridis(4)[as.integer(factor(the_communities$comm))]) %>%
+  set_vertex_attr("shape", value = ifelse(V(.)$type == "Multilocus Genotype", "triangle", "circle")) %>%
+  set_edge_attr("width", value = E(.)$weight) %>%
+  toVisNetworkData()
+vg$nodes <- vg$nodes %>% 
+  group_by(id) %>%
+  mutate(color = list(list(background = make_rgb(color, 0.8), 
+                      border = make_rgb(rep("black", n()), 0.8),
+                      highlight = list(background = make_rgb(color),
+                                       border = make_rgb(rep("black", n()))
+                                       )
+                      )))
+vgn <- visNetwork(nodes = vg$nodes, edges = vg$edges,# height = "500px", 
+                  main = "Relation of Multilocus Genotypes and MCGs")
+set.seed(2017-05-03)
+vgn
+# vgn %>%
+#   visIgraphLayout("layout_nicely") %>%
+#   visOptions(highlightNearest = list(enabled = TRUE, hover = TRUE), 
+#              nodesIdSelection = TRUE)
+#   
+```
+
 ## Subgraph

### 2017-06-11 doc/RMD/data-comparison.Rmd
add poppr and remove slice to processing
The data processing document was failing in places because I hadn't
loaded poppr. Additionally, readxl updated with the ability to
finally ignore empty rows! Now, slice is no longer needed and
commented out for posterity.

@@
 library('assertr')
+library('poppr')
 ```
@@
 
+> 2017-06-11: This has been fixed in readxl version 1.0
+
 Quirks specific to thse data:
@@
   mutate(Source = ifelse(Source == "", "unk", Source)) %>%
-  slice(-n()) %>% # remove last row
+  # slice(-n()) %>% # remove last row
   arrange(Isolate) %>%
@@
   spread(locus, allele) %>%
-  slice(-n())
+  # slice(-n()) %>%
+  I()
 ex <- ex[!names(ex) %in% locNames(dat)]

### 2017-06-23 doc/RMD/data-comparison.Rmd
update repeat length parsing in data comparison

@@
 
-First we need to transform the data to something that
+First we need to transform the data to something that is useable, aka a genclone
+object. To avoid confustion with states, we are also transforming the two-letter
+abbreviations for France, Mexico, and Australia to the full names.
 
@@
 locNames(dat)
+```
 
-repeat_lengths <-
-  c(
-  `5-2(F)` = 2.000000,
-  `5-3(F)` = 0.500000,
-  `6-2(F)` = 6.000000,
-  `7-2(F)` = 2.000000,
-  `8-3(H)` = 2.000000,
-  `9-2(F)` = 2.000000,
-  `12-2(H)` = 2.000000,
-  `17-3(H)` = 3.000000,
-  `20-3(F)` = 2.000000,
-  `36-4(F)` = 0.500000,
-  `50-4(F)` = 0.500000,
-  `55-4(F)` = 4.000000,
-  `92-4(F)` = 2.000000,
-  `106-4(H)` = 4.000000,
-  `110-4(H)` = 4.000000,
-  `114-4(H)` = 4.000000
-  )
+### Gathering Repeat Lengths
+
+The repeat lengths were contained in the `A1_Copy of binned-genotypes_SE.xlsx`
+spreadsheet in the "replen est" sheet. Because this sheet contained several
+merged cells and other oddities, I copy+pasted the relevant data here and 
+parsed it from the text.
+
+```{r}
+rl <- read_tsv("dinuc			compound			hexinuc			dinuc			dinuc			dinuc			dinuc			trinuc			dinuc			compound			compound			tetranuc			dinuc			tetranuc			tetranuc			tetranuc
+(GT)8			[(GT)2GAT]3(GT)14GAT(GT)5[GAT(GT)4]3(GAT)3			(TTTTTC)2(TTTTTG)2(TTTTTC)			(GA)14			(CA)12			(CA)9(CT)9			(CA)9			(TTA)9			(GT)7GG(GT)5			(CA)6(CGCA)2(CAT)2			(CA)7(TACA)2			(TACA)10			(CT)12			(CATA)25			(TATG)9			(AGAT)14(AAGC)4
+ 5-2(F)	bins	count	 5-3(F)	bins	count	 6-2(F)	bins	count	 7-2(F)	bins	count	 8-3(H)	bins	count	 9-2(F)	bins	count	 12-2(H)	bins	count	 17-3(H)	bins	count	 20-3(F)	bins	count	 36-4(F)	bins	count	 50-4(F)	bins	count	 55-4(F)	bins	count	 92-4(F)	bins	count	 106-4(H)	bins	count	110-4(H)	bins	count	114-4(H)")
+
+x <- rl %>%
+  select(-starts_with("X")) %>% # removing non-informative columns
+  slice(-1) %>%                 # removing first row (repeat patterns)
+  gather(replen, locus) %>%
+  mutate(replen = strsplit(replen, "_") %>% map_chr(1)) %>% # fix duplicated column names
+  mutate(replen = case_when(
+    grepl("^di", .$replen)    ~ 2,
+    grepl("^tri", .$replen)   ~ 3,
+    grepl("^tetra", .$replen) ~ 4,
+    grepl("^hex", .$replen)   ~ 6,
+    TRUE                      ~ 0.5
+  ))
+
+x <- paste("c(", paste(paste0("`", x$locus, "`", " = ", x$replen), collapse = ",\n"), ")", sep = "\n")
+cat(x)
+
+repeat_lengths <- eval(parse(text = x))
 repeat_lengths <- ifelse(repeat_lengths < 1, 4, repeat_lengths)
+repeat_lengths
 ```

### 2017-06-23 doc/RMD/data-comparison.Rmd
write raw and cleaned data to csv

@@
   mutate(allele = as.integer(allele)) %>% # force alleles to integers
-  spread(locus, allele) %>%
-  # slice(-n()) %>%
-  I()
+  spread(locus, allele)
+
+readr::write_csv(ex, "data/raw_data.csv", col_names = TRUE)
+
 ex <- ex[!names(ex) %in% locNames(dat)]
@@
   as.genclone()
+
+strata(dat) %>%
+  bind_cols(datdf) %>%
+  as_data_frame() %>%
+  readr::write_csv("data/clean_data.csv", col_names = TRUE)
 ```

### 2017-06-23 doc/manuscript/manuscript.Rmd
why were these even nested :see_no_evil:

---
title: Population structure and phenotypic variation of *Sclerotinia sclerotiorum* from dry bean throughout the United States, and from Mexico, France, and Australia
author:
  - name: Zhian N. Kamvar
    email: zkamvar@unl.edu
    affiliation: University of Nebraska-Lincoln
  -name: Bimal Sajeewa Amadarasa
    email: bamadarasa2@unl.edu
    affiliation: University of Nebraska-Lincoln
  -name: R. Jhala
    email: rjhala2@unl.edu
    affiliation: University of Nebraska-Lincoln
  -name: James Steadman
    email: jsteadman1@unl.edu
    affiliation: University of Nebraska-Lincoln
  - name: Sydney E. Everhart
    email: everhart@unl.edu
    affiliation: University of Nebraska-Lincoln
    corresponding: everhart@unl.edu
address:
  - code: University of Nebraska-Lincoln
    address: Department of Plant Pathology, University of Nebraska-Lincoln, Lincoln 68583
abstract: |
  Characterization of pathogen population structure can yield important insight into pathogen biology and epidemiology. A worldwide collection of 366 *S. sclerotiorum* isolates from 26 dry and snap bean cultivars and experimental lines in Australia, France, Mexico, and 10 bean production states in the U.S. collected from 2003 to 2012 were genotyped at 16 microsatellite loci. A total of 165 multilocus genotypes were identified using a subset of 11 loci that followed an expected repeat motif.  There were a total of 87 mycelial compatibility groups (MCGs) in the collection of isolates. We did not detect an association of MLGs and MCGs. Discriminant analysis of principal components using clone corrected data showed region of origin was the best separator (AMOVA P = 0.001). The Mexican population was highly divergent from the rest, showing closest association with the California population. Despite overall genetic linkage, none was observed for populations from Washington in 2008, Oregon in 2003, California in 2004, or France in 2004 and 2005. AMOVA for populations based on year of sampling had a low PhiPT value of 0.07 (P = 0.001) while host cultivar, isolate virulence and sampling field did not have a significant impact on the microsatellite genotype.   The analysis of populations from same field over different years for linkage disequilibrium did not provide evidence for sexual outcrossing. However, analysis of molecular variance did show populations were significantly different in some locations between some years (eg. MI 2003 and 2005 populations; WA 2008 compared to both 2003 and 2005).  These results suggest that despite large physical separation, divergence of populations is low, which is not surprising given the low rate of heterothallic recombination thought to occur. However, some level of heterothallic recombination is also taking place in these fields in addition to natural mutations.  The lack of association of MLGs and MCGs in dry bean isolates is indicative of sexual outcrossing, and distribution of microsatellite MLGs across distant regions due to agricultural practices and human-mediated migration.

author_summary: |
  Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Nullam sapien tellus, commodo id velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.

bibliography: mybibfile.bib
output: rticles::plos_article
csl: plos.csl
---

_Text based on plos sample manuscript, see [http://journals.plos.org/ploscompbiol/s/latex](http://journals.plos.org/ploscompbiol/s/latex)_

# Introduction

Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut dui eu ante cursus gravida non sed sem. Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.

A list

- Item 1
- Item 2

Here are two sample references: @Feynman1963118 [@Dirac1953888].

# References {#references .unnumbered}

### 2017-06-23 doc/manuscript/manuscript.Rmd
fix silly manuscript mistake :chart_with_upwards_trend:

@@
     affiliation: University of Nebraska-Lincoln
-  -name: Bimal Sajeewa Amadarasa
+  - name: Bimal Sajeewa Amadarasa
     email: bamadarasa2@unl.edu
     affiliation: University of Nebraska-Lincoln
-  -name: R. Jhala
+  - name: R. Jhala
     email: rjhala2@unl.edu
     affiliation: University of Nebraska-Lincoln
-  -name: James Steadman
+  - name: James Steadman
     email: jsteadman1@unl.edu

### 2017-06-27 doc/manuscript/manuscript.Rmd
Add manuscript
 - moved abstract to separate file and read in via
   readLines
 - Add manuscript text via copy+paste
 - Make each sentence one line
 - Italicize all S. sclerotiorum
 - replace UTF-8 encoded apostrophies and hypens

@@
     address: Department of Plant Pathology, University of Nebraska-Lincoln, Lincoln 68583
-abstract: |
-  Characterization of pathogen population structure can yield important insight into pathogen biology and epidemiology.
-A worldwide collection of 366 *S.
-sclerotiorum* isolates from 26 dry and snap bean cultivars and experimental lines in Australia, France, Mexico, and 10 bean production states in the U.S.
-collected from 2003 to 2012 were genotyped at 16 microsatellite loci.
-A total of 165 multilocus genotypes were identified using a subset of 11 loci that followed an expected repeat motif.
-There were a total of 87 mycelial compatibility groups (MCGs) in the collection of isolates.
-We did not detect an association of MLGs and MCGs.
-Discriminant analysis of principal components using clone corrected data showed region of origin was the best separator (AMOVA P = 0.001).
-The Mexican population was highly divergent from the rest, showing closest association with the California population.
-Despite overall genetic linkage, none was observed for populations from Washington in 2008, Oregon in 2003, California in 2004, or France in 2004 and 2005.
-AMOVA for populations based on year of sampling had a low PhiPT value of 0.07 (P = 0.001) while host cultivar, isolate virulence and sampling field did not have a significant impact on the microsatellite genotype.
-The analysis of populations from same field over different years for linkage disequilibrium did not provide evidence for sexual outcrossing.
-However, analysis of molecular variance did show populations were significantly different in some locations between some years (eg.
-MI 2003 and 2005 populations; WA 2008 compared to both 2003 and 2005).
-These results suggest that despite large physical separation, divergence of populations is low, which is not surprising given the low rate of heterothallic recombination thought to occur.
-However, some level of heterothallic recombination is also taking place in these fields in addition to natural mutations.
-The lack of association of MLGs and MCGs in dry bean isolates is indicative of sexual outcrossing, and distribution of microsatellite MLGs across distant regions due to agricultural practices and human-mediated migration.
-
+abstract: >
+  `r paste(readLines("abstract.md"), collapse = "\n  ")`
 author_summary: |
   Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Curabitur eget porta erat.
-Morbi consectetur est vel gravida pretium.
-Suspendisse ut dui eu ante cursus gravida non sed sem.
-Nullam sapien tellus, commodo id velit id, eleifend volutpat quam.
-Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus.
-Nunc pellentesque pretium diam, quis maximus dolor faucibus id.
-Nunc convallis sodales ante, ut ullamcorper est egestas vitae.
-Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.
 
@@
 
-_Text based on plos sample manuscript, see [http://journals.plos.org/ploscompbiol/s/latex](http://journals.plos.org/ploscompbiol/s/latex)_
+# Introduction {-}
 
-# Introduction
+*Sclerotinia sclerotiorum* (Lib.) de Bary is an ascomycete plant pathogen found worldwide.
+This is a necrotrophic pathogen that is primarily homothallic (self-fertilization) (Bolton et al., 2006) (Sexton et al., 2006).
+It has a host range of more than 400 species of plants belonging to 75 families (Boland and Hall, 1994).
+Economically important crops affected by this pathogen include sunflower, soybean, oilseed rape, dry bean, chickpea, peanut, dry pea, lentils and vegetables (Bolton et al., 2006).
+*Sclerotinia sclerotiorum* causes white mold disease on dry bean (Phaseolus vulgaris L.), which can be a yield-limiting in temperate climates (Steadman, 1983).
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Curabitur eget porta erat.
-Morbi consectetur est vel gravida pretium.
-Suspendisse ut dui eu ante cursus gravida non sed sem.
-Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus.
-Nunc pellentesque pretium diam, quis maximus dolor faucibus id.
-Nunc convallis sodales ante, ut ullamcorper est egestas vitae.
-Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.
+Sclerotia are capable of surviving more than five years in soil (Steadman, 1983) and germinate carpogenically or myceliogenically to infect the plant when environmental conditions are favorable.
+A carpogenically geminated sclerotia gives rise to apothecia in which ascospores are formed.
+Ascospores infect senescent plant tissues that, when supplemented with cool and humid microclimate within the plant canopy, leads to mycelial infection of the host plant.
+Consequently, flowering, which coincide with canopy closure is a critical factor of disease initiation since sensing flower parts fallen on to leaves, petioles or stems provide the required energy source for plant infection.
+Myceliogenically geminated sclerotia result in formation of hyphae, which also have potential for directly plant infection, though this is considered rare and less important within the disease cycle of the pathogen (Abawi and Grogan, 1975).
+Diseased bean fields initially show scattered wilted leaves.
+Symptoms on individual plants are water-soaked lesions on leaves, pods and stems, which expand over time and show signs of white fungal mycelia.
+Lesions on stems and branches become necrotic as they age and dead tissues appear bleached.
+Sclerotia are often produced on dead or dying plants once conditions become unfavorable for fungal growth.
 
-A list
+Previous studies of *S.
+sclerotiorum* population biology in North America and Australia have reported clonal propagation of the pathogen (Cubeta et al., 1997; Ekins et al., 2011; Kohli and Kohn, 1998; Kohli et al., 1995) with evidence of outcrossing observed in some areas (Atallah et al., 2004; Kohli and Kohn, 1998; Malvarez et al., 2007; Sexton and Howlett, 2004).
+One such study examined *S. sclerotiorum* population structure using microsatellite markers from potato fields in the Columbia Basin of Washington state.
+Due to high crop diversity and high presence of the pathogen, it was important to understand the potential for outcrossing in order to develop disease management strategies.
+Molecular analysis revealed 92% of the isolate variability was found within subpopulations (Atallah et al., 2004).
+However, there was evidence of linkage equilibrium of microsatellite loci in 2 locations and illustrated the possibility of outcrossing in this region.
 
-- Item 1
-- Item 2
+A population study from Australian canola fields showed evidence of sexual outcrossing of *S. sclerotiorum* isolates (Sexton and Howlett, 2004).
+Another study characterized three populations of *S. sclerotiorum* from lettuce in California (CA) and Ontario (ON), and pea/lentil in Washington (WA) (Malvarez et al., 2007).
+While all three populations were  clonal, the index of association (IA) based on DNA fingerprint was closer to zero for CA than it was for the other populations.
+Populations from Australian sunflower with evidence of clonal reproduction were reported, and despite populations sampled from fields separated 50 to 400 km and from both head and basal rots, all showed similar gene diversity, with authors concluding isolates representing different geographies belong to one large genetic population of *S.
+sclerotiorum* (Ekins et al., 2011).
 
-Here are two sample references: @Feynman1963118 [@Dirac1953888].
+Attanayake et al. (2013) studied population differentiation of *S. sclerotiorum* isolates infecting canola from China and the USA.
+There were no shared microsatellite multilocus genotypes (MLG) between the two populations.
+Sexual outcrossing was detected in both populations with higher recombination rate in Chinese population than the US population.
+Additionally, phenotypic traits such as production of sclerotia, mycelial compatibility group (MCG), growth rate were also significantly different between the two populations.
+Genetic diversity of *S. sclerotiorum* from soy bean, canola, sunflower, and dry bean in the North Central region of the USA was studied by (Aldrich-Wolfe et al., 2015).
+Microsatellite analysis revealed little host- specificity or genetic-isolation of *S.
+sclerotiorum* isolates among host crops as 93% of the observed genetic variation was explained by within-host differences.
+Isolates from all host crops were primarily reproducing clonally while there were also evidence of genetic recombination of isolates from canola and sunflower.
 
-# References {#references .unnumbered}
+Most of the previous studies on *S. sclerotiorum* had been conducted in the USA, Canada and Australia but not in Europe.
+In those studies, different dry bean cultivars had been rarely used as  hosts to investigate any host-pathogen relationships or specificity.
+No white mold resistant dry bean cultivars exist.
+Canopy architecture and growth pattern of some cultivars enable improved disease avoidance (Kolkman and Kelly, 2002).
+Field trials in some regions of the US have shown G122 is white mold resistant and Beryl is susceptible (pers. comm. with J.
+Steadman).
+Understanding genetic relationships and reproduction behavior of *S.
+sclerotiorum* populations from locations worldwide is beneficial for plant breeders seeking to develop white mold resistance cultivars for deployment worldwide.
+This information is also helpful in making disease management decisions and developing disease resistance screening methods.
+
+Microsatellite markers (simple sequence repeats or SSR) are codominant, polymorphic and also able to detect multiple alleles per locus (Atallah et al., 2004; Sirjusingh and Kohn, 2001a).
+Reproducibility of microsatellites is higher compared other fingerprinting techniques such as random amplified polymorphic DNA (RAPD) (Atallah et al., 2004).
+High mutagenic rate and multiallelic nature of microsatellites render suitability for inferring population structure of organisms and their potential for outcrossing.
+In the case of plant pathogens like *S.
+sclerotiorum* that is predominantly clonal and occasionally outcrossing, microsatellites have been used successfully to ascertain their ability to reproduce by sexual outcrossing (Atallah et al., 2004; Sexton and Howlett, 2004) and also to detect differences among clonally propagated individuals.
+
+We analyzed *S.
+sclerotiorum* isolates collected from dry bean cultivars from 2003 to 2012 from different geographic locations in America, Europe and Australia using microsatellite markers to delineate genetic differences and reproduction behavior.
+The specific aims of our study were:  a) To assess association of phenotypic data (virulence) and environmental data (bean cultivar, region, field, year) on genotypic variation of *S.
+sclerotiorum* isolates, b) To test the hypothesis that *S.
+sclerotiorum* populations of dry beans originated from different continents were genetically different, and c) although previous studies show that *S.
+sclerotiorum* populations are largely clonal, few compare populations from different continents (Lehner and Mizubuti, 2017), c) To identify localized genotypes representative of each location and also investigate if there are widespread genotypes common among locations.
+d) To test potential for sexual reproduction in sub populations of the pathogen isolates under different conditions such as year or field in each region, year/region, field/region/year, and e) Test if pathogen within-field populations differentiate over time.
+
+The overall goal of this work was to identify representative isolates for use in breeding programs.
+
+# Materials and Methods {-}
+
+## Isolate collection and DNA extraction {-}
+
+Sclerotia of *S. sclerotiorum* were collected over several years from grower fields and/or dry bean nurseries in Mexico, France, Australia, and 11 states of the United States of America (Table 1).
+The dry bean nurseries in different regions are indicated by wmn (white mold nursery) or mtc (Mitchell) and all other isolates came from commercial fields (Table 1).
+The bean nurseries were 5 m x 10 m in size and maintained without applying fungicides to observe natural incidence of white mold disease.
+The nursery plots were incorporated with a basal dressing of N: P: K = 1: 3: 2 and side dressing of 0:3:2 during the season.
+Although the nursery locations were the same over sampling years, sampling fields within a location varied for sampling years.
+Sampling was carried out on a 'W' or 'V' shape transect, or randomly.
+Sampling year and number of isolates collected were necessarily influenced by incidence and severity of white mold disease.
+
+Mycelial growth of each isolate was generated by activating sclerotia according to the method described previously (Matheron and Porchas, 2005).
+These isolates (N = 366) were phenotypically characterized by determining mycelial compatibility group MCG and virulence.
+MCG was determined through co-culturing isolates in pairs on PDA plates, where incompatibility of different MCG resulted in formation of a mycelial barrage (Kohn et al., 1990) and isolates belonging to the same MCG did not form a mycelial barrage.
+Virulence of each isolate was assessed using a straw test (Petzoldt and Dickson, 1996; Teran et al., 2006) that rates necrotic lesion on a scale of 1-9, with ascending numbers assigned for increasing aggressiveness.
+DNA was purified from mycelium of each isolate using a Qiagen DNeasy Plant Mini Kit (Qiagen Inc., Valencia, CA) according to the manufacturer's instructions.
+
+## Microsatellite genotyping Analysis {-}
+
+We characterized *S. sclerotiorum* isolates for microsatellite polymorphism using 16 primer pairs developed by (Sirjusingh and Kohn, 2001b).
+Polymerase chain reaction (PCR) was carried out as described using primers labeled with FAM fluorophore, and resulting PCR products were first resolved in a 1.5% agarose gel stained with ethidium bromide to ensure product was within the expected size range.
+Capillary electrophoresis (fragment analysis) of PCR products added with size standard GeneScan™ 500 LIZ® was performed using an ABI 3730 genetic analyzer (Life Technologies Corporation, Carlsbad, CA) at the Michigan State University Genomic Sequencing Center.
+
+## Data analysis {-}
+
+Isolates that fell into the 10 most abundant MCGs were subjected to an analysis of variance (ANOVA) in JMP version 12 to determine if their virulence levels were significantly different.
+The mean separation was carried out using Tukey's HSD test.
+Five microsatellite loci included compound repeats, where genotypes could not be unequivocally assigned based on the known repeat motifs and were consequently removed, resulting in 11 polymorphic loci selected for analysis (Table 2).
+Since it could not be ruled out that isolates possessing the same genotype within the same field, year, and host were duplicate collections from the same plant, the data were first clone-corrected to remove these genotypes.
+The data set was hierarchically reformatted to group isolates within each location/host/region/year and clones were removed at that level.
+The clone-corrected data set included 318 isolates.
+All analyses were performed on this data set unless otherwise mentioned.
+
+## Distance-based redundancy analysis  {-}
+
+A distance-based redundancy analysis (dbRDA) (Legendre and Anderson, 1999) was performed with the function capscale in the vegan 2.4-3 package in R (https://CRAN.R-project.org/package=vegan) to delineate relative contribution of phenotypic, spatial, and temporal components in influencing genetic composition of the populations.
+dbRDA is a method of conducting constrained ordinations on data matrices consisting of non-Euclidean distance measures.
+RDA is a multivariate linear regression analysis of independent (explanatory: eg. environmental and phenotypic) variables on dependent (non-explanatory or response) variables.
+In our case, a pairwise Bruvo's (Bruvo et al., 2004) distance matrix was generated from microsatellite data using poppr 2.4.1 (Kamvar et al., 2014; Kamvar et al., 2015) and employed it as the dependent matrix for RDA analysis.
+To identify environmental variable(s) correlated with genetic variation, a forward-backward selection process was applied with the vegan function ordistep to identify the set of variables that explained more variation than predicted with the global model containing all variables.
+An ANOVA was then performed to test for significance of the reduced model and marginal effects using 999 permutations.
+The varpart function of vegan was used to determine variation partitioning of explanatory variables.
+
+## Shared Multilocus Genotype Community Analysis {-}
+
+To assess the connectivity of shared genotypes across populations, we first tabulated the multilocus genotypes shared between at least two populations (defined as states or countries) with the poppr function mlg.crosspop() (Kamvar et al.
+2014).
+From these data, we were able to construct a graph with populations as nodes and shared genotypes as edges (connections) between the nodes using the R packages igraph version 1.0.1 (Csardi and Nepusz 2006), dplyr version 0.5.0 (Wickham and Francois 2016), and purrr version 0.2.2 (Wickham 2016).
+Each node was weighted by the fraction of shared MLGs.
+Each edge represents a single MLG, but because a single MLG could cross populations, that MLG may be represented by (n*(n-1))/2 edges where n represents the number of populations it crosses.
+Edges were weighted by 1- Psex, where Psex is the probability of encountering the same genotype via two independent meiotic events (Arnaud-Haond 2007, Parks and Werth 1993).
+We then used the cluster_optimal() function from igraph to define communities (Csardi and Nepusz 2006).
+The graph was plotted using the R packages ggplot2 version 2.2.1 (Wickham 2009) and ggraph version 1.0.0 (Pedersen 2017).
+To ensure that we were able to capture the same community signal, we additionally ran this analysis including the 5 polymorphic markers described above.
+
+## Discriminant analysis of principal components {-}
+
+DAPC is a multivariate method designed to identify and describe clusters of genetically related individuals.
+Discriminant analysis maximizes separation between groups while minimizing variation within groups (Jombart et al., 2010).
+The find.clusters function in Adegenet 2.0.1 (Jombart, 2008) uses a k-means clustering algorithm to identify the optimum number of clusters.
+It was not able to identify a single optimum and identified a range of optimal clusters between 5-15.
+Consequently, DAPC was performed several times with different optimal number of clusters to determine whether isolates (N = 318) fell into each cluster have any biological meaning such as belonging to a particular host or came from the same region etc.
+the best summaries for the total isolates (N = 318).
+Since dbRDA identified region, host, and sampling year as the top three variables influencing microsatellite data, DAPC was also performed on data defined according to each of these variables, separately.
+Isolates (N = 318) were assigned to different levels of the three factors (region, host, year) and assessed predefined group size (original number of isolates at each level of the factor), post group size (number of isolates assigned by the DAPC algorithm) and the proportion of the original isolates assigned to the original group after analysis.
+This was performed to determine the suitability of explanatory variables (i.e. region, host, and year) in describing the clusters generated.
+Clone-corrected isolates sampled in different years from the white mold nurseries (wmn) were tested for population differentiation over time by DAPC analysis.
+Only WA, MI, and NE isolates were qualified since we filtered out the states that had at least ≥7 samples/year with at least two years of data for the analysis.
+Lastly, we pooled all isolates, clone corrected them (N = 165) and performed DAPC using the function find.clusters as described above to delineate how many clusters were formed by these unique genotypes.
+AMOVA was used to test the significance of clusters resulted by different DAPC analyses using GenAlEx version 6.41 (Peakall and Smouse, 2012).
+PhiPT which is an analogue of Wright's FST was used to ascertain the significance of populations.
+An FST value of 0.05 or less is generally considered as reasonably low and may be interpreted to mean that structuring between populations is weak (Balloux and Ugon-Moulin, 2002; Wright, 1978).
+
+## Molecular population statistics and linkage disequilibrium {-}
+
+Genotypic evenness (Grunwald et al., 2003), which is a measure of distribution of genotype abundances and unbiased gene diversity (Hexp) (Nei, 1978) were calculated using poppr package in R.
+Microsatellite data were also tested for evidence of linkage disequilibrium, where significant disequilibrium between pairs of loci in a population is attributed to linkage between loci and may be an indication of clonal reproduction if loci are physically unlinked, whereas linkage equilibrium can be an indication of outcrossing populations.
+The linkage disequilibrium of clone-corrected data was determined by calculating the index of association (IA) and its standardized version () and a permutation approach with 999 permutations was used to test whether observed values were significantly different from data randomized at each locus (Agapow and Burt, 2001; Brown et al., 1980).
+Value of IA tends to increase with the number of loci and therefore we made our inferences based on the , which standardizes the number of loci tested and considered a more robust method (Agapow and Burt, 2001).
+Higher observed  or IA values would indicate clonally reproducing populations.
+
+## Results {-}
+
+A total of 366 isolates were collected from 2003 to 2012 from diseased dry bean plants in 11 states in the US, Australia (2 mainland; 4 Tasmania), Mexico, and France (Fig. 1).
+Most isolates came from MI (62), ND (60), WA (59), NE (47), and CO (42).
+Other regions in lesser representation were France (24), MX (18), OR (17), CA (16), MN (11), Australia (6), WI (2), ID (1), and NY (1).
+Four of the Australian isolates were collected from Tasmania while the other two came from Victoria state.
+There were no samples collected in 2006 and 2011.
+The highest number of samples were collected in 2005 (N = 63) and lowest in 2012 (N = 4).
+Pathogen isolates represented 26 commercial cultivars and breeding lines with some representing unknown dry bean hosts (Fig. 2).
+We investigated the population association with host for cultivars where more than 10 isolates were obtained, which included G122, Beryl, Bunsi, Yellow, Pinto, and Merlot.
+These six cultivars had a total of 238 *S. sclerotiorum* isolates (of 366 total isolates collected) belonging to 130 MLGs.
+Of these MLGs, 53 were singletons, while the rest were obtained from more than one host cultivar; only one private allele on locus 55-4 was identified, belonging to the *S.
+sclerotiorum* population from the cultivar 'Bunsi'.
+The number of unique MLGs recorded from each host is shown in Table 3.
+The majority of MLGs reported were found on multiple host cultivars and only a few unique to a cultivar.
+For instance, there were only 15 unique MLGs within the population of 68 *S. sclerotiorum* isolates obtained from the 'Beryl' cultivar.
+
+There were a total of 165 multilocus genotypes in the isolate collection.
+The most abundant MLG represented 27 isolates (Fig. 3 and Table 4) across five regions (NE, MI, WA, CO, and ND).
+Twenty isolates of this MLG were identified from ND.
+The genotypes that had either highest geographical distribution or found in high frequency are given in Table 4.
+MLG 163 was identified in NE, MI, CO and ND, while MLG 165 was identified in MN and all states aforementioned, except for ND.
+Isolates with MLG 15 were identified from populations in MI, WA, CO, and ND.
+The MLGs 163, 165 and 15 had 15, 7, and 4 individuals across these populations, respectively.
+The highest number of clones was recorded from ND, NE, and MI, which included 20, 10 and 7 isolates, respectively (Table 4).
+
+Community analysis of the MLGs shared between populations revealed 4 communities with a modularity of 0.17 (Fig.
+3Z): A costal community (CA, OR, WA, and NY), a mid-west community (CO, NE, ND, MI), and an international community (MN, France, Australia).
+Analysis with 16 loci resulted in the removal of the NY node because it no longer shared a genotype with OR, but the same overall community structure was present with a modularity of 0.2 (Supplementary Figure S1Z).
+The international community appears to be driven by MLG 4, which is shared between all three populations and has a Psex value of 2.87 x 10-5, in contrast to the abundant MLG 25, which has a Psex value of 0.0168.
+
+The MCGs with the highest frequency were MCG 5 and MCG 44 representing 73 and 44 isolates, respectively (Fig. 4).
+MCGs that were unique to a region are shown in Fig.
+5 while MCGs that represented at least 2 or more regions are in Fig 6.
+20 MCGs were common in two or more regions and the rest were distributed uniquely in different regions.
+ID, MN, NE, NY, and WI did not have any unique/isolated MCGs.
+The virulence of isolates ranged from 1.4 to 7.9 with a mean of 5.01.
+There were six isolates that fell into upper outlier category (virulence > 7.6) and one isolate in the lower outlier range (virulence < 2.3) (box plot results not shown).
+The mean virulence of the 10 most abundant MCGs ranged from 4.6 (MCG 49) to 6.02 (MCG 44) (see Fig 4 for 10 most abundant MCGs).
+ANOVA of these 10 MCGs indicated significant difference for virulence levels at P <0.0001.
+Mean separation of 10 most common MCGs showed MCG 44 is different from MCGs 9, 1, 45, 4, 53, and 49 while the virulence levels of MCGs 2, 5, 3 failed to show any difference to MCG 44 or the other MCGs.
+
+The number of observed alleles per locus ranged anywhere from 2 to 10 (Table 2).
+All loci were informative at cut off level of 2/N where N was the total number of individuals.
+The genetic diversity indices of *S. sclerotiorum* populations from different states of the USA, France, Mexico and Australia based on SSR analysis are given in Table 5.
+The 165 unique MLGs identified among all 366 isolates are given in Supplementary Table 1.
+Gene diversity (Hexp) of populations from different regions ranged from 0.259 to 0.596 (Table 5).
+Evenness of genotypes of all regions was high, with the exception of ND, where a single genotype was represented by 20 isolates.
+There was one private allele in the WA population (204 bp allele of 55-4 locus) and two private alleles in the ND population (two 366 bp alleles in 9-2 locus and two 220 bp alleles in 12-2 locus).
+
+## dbRDA {-}
+
+The dbRDA was performed on clone-corrected isolates (N = 318) after grouping them by location, host, region, and year.
+The forward-backward selection process revealed severity, year, region, and host to be the optimal variables for the reduced model, accounting for 17.6% of the variation.
+ANOVA revealed that the reduced model was significant with an adjusted R^2^ of 0.032 (P = 0.001).
+The marginal effects of state, host, and year were also significant (P = 0.001), but virulence was not significant (P = 0.045).
+Much of the variation appears to be driven by isolates from Mexico (Fig. 7) and 2005, but it's worth noting that all Mexican isolates were collected in 2005.
+Variance partitioning of the independent variables indicated virulence to be the least influential factor with 0.1% contributing to explaining the variation of molecular data whereas the combination of variables accounted for 3.3%.
+
+## DAPC analysis {-}
+
+DAPC analysis for different regions and hosts are depicted in Fig. 8.
+These isolates were clone-corrected at location/host/region/year level prior to the analysis (N =318).
+Accordingly, MX (N = 15) and CA (N =18) populations grouped separately from all other populations with MX isolates showing the highest deviation (Fig. 8A).
+Comparison of genotypes by region showed all of the nine MLGs in the Mexican population were unique, with no genotypes crossing over to other populations.
+MLGs of Californian isolates were found only in OR and WA populations.
+MLG 6, 84 and 100 were (see Supplementary Table 1 for SSR genotypes related to 165 MLGs of total isolates) unique to CA isolates.
+
+Populations from different regions were compared with AMOVA, where differentiation was low among regions (PhiPT  = 0.11; P = 0.001).
+Pairwise comparisons of all populations according to region showed MX and CA populations were significantly different across all populations at P ≤ 0.001.
+The PhiPT values of Pairwise comparisons between CA and other populations ranged from 0.116 to 0.389 and from 0.355 to 0.493 between MX and other population.
+The CA population showed evidence of sexual recombination ( = 0.008; P = 0.27, Ia = 0.08; P = 0.3) while the  value of AU was 0.12 with a high P of 0.05 (Ia = 0.70; P = 0.05), whereas all other populations showed linkage disequilibrium for loci tested indicating high possibility of clonal reproduction.
+
+The dapc for hosts was limited to six most common cultivars: G122 (N = 55), Beryl (N = 62), Bunsi (N = 47), Yellow (N = 26), Pinto (N = 15), and Merlot (N = 11).
+According to Fig.
+8B, cultivar Merlot isolates group distant from others, but were not significantly different (P =0.325).
+Interestingly, *S.
+sclerotiorum* populations from cultivars 'Beryl' and 'G122', which are known to have different level of white mold susceptibility, showed no difference in population genetic data.
+
+AMOVA revealed genetic differences among isolates sampled in different years (DAPC biplot not shown).
+The PhiPT was a low value of 0.07 with P = 0.001.
+Isolates from 2012 were not included in the analysis since there were only 4 samples.
+Samples for other years ranged from 19 in 2009 to 63 in 2005.
+Sampling year 2003 to 2005 and 2009 were different across every Pairwise comparison.
+None of the above sampling variables (i.e.
+region, year, host) seemed to influence distinct molecular composition of *S.
+sclerotiorum* populations since pre-defined (i.e.
+region, host, and year) isolates and isolates that were assigned into post groups by DAPC varied largely (Figs.
+8A, 8B and 9).
+However, assignment of isolates by DAPC analysis of regions did show some promising results for MX and CA isolates (Fig. 8A).
+The original group size of each region and DAPC defined group sizes for different regions are given in Table 6.
+
+Since find.clusters function of Adegenet library did not result in a clear-cut number of clusters for 318 isolates, we arrived at six (6) as the best number of populations (Fig.
+9) for our data after examining results of DAPC for different number of clusters.
+Three clusters, namely clusters one, three, and six were formed separately from the rest.
+The cluster 3 of the DAPC plot consisted entirely of MX isolates (N =13/15).
+Rest of the two MX isolates were grouped in cluster 1.
+Cluster 6 had 50 isolates in total representing mainly of NE (N = 15/37), MI (N = 12/58), and CO (N = 12/34).
+All CA (N = 18) isolates and 13/58 WA isolates dominated the cluster 1.
+AMOVA on these six clusters resulted 44% molecular variance among populations with P = 0.001.
+Pairwise comparison showed significant difference (P = 0.001) for all pairs.
+The DAPC plot after pooling all isolates together and removing clones (N =165) was compared with the plot of 318 isolates looked similar to Fig. 8C.
+The number of clusters was set to 6 for the analysis which resulted in a DAPC having similar topology to plot in Fig. 9 with 318 isolates(results not shown).
+Isolates of MX grouped separately while another distinct group had 24 isolates representing CO, NE, and ND.
+
+We tested the clone corrected data for evidence of sexual outcrossing within *S. sclerotiorum* populations in different regions and different years/region.
+CA showed evidence of sexual outcrossing (Ia = 0.08, P.Ia = 0.317,  = 0.008, P. = 0.317) while all other regions implied clonal propagation (data not shown).
+Five populations showed evidence of sexual outcrossing: CA in 2004, FR in 2004 and 2005, OR in 2003, and WA in 2008.
+However, after clone-correction these populations had 6-7 isolates/population (Table 7) and therefore, may not be sufficient to make inferences about reproductive mode.
+Only exception was the WA_2008 population, which consisted of 12 isolates.
+Within sampling locations white mold nurseries (wmn) were tested for linkage disequilibrium since same wmn fields were used for sampling over different years.
+Only CA wmn showed evidence of sexual reproduction ( = 0.008; P = 0.29) while all other wmn locations suggested clonal reproduction.
+The DAPC analyses of WA, and MI isolates from same wmn fields over different years are in Fig. 10.
+Since *S. sclerotiorum* isolates are predominantly clonal, we wanted to explore whether there was any introduction of new isolates from other fields or if there were evidence of sexual reproduction.
+The AMOVA results for different clusters of Fig. 10A revealed significance difference between WA 2008 isolates from the rest (PhiPT ≥ 0.206, P = 0.001).
+The 2003 and 2005 isolates were genetically similar (PhiPT = 0.056, P = 0.129).
+MI isolates from 2008 and 2003 (Fig.
+10B) showed significant difference (PhiPT = 0.127, P = 0.04) while 2005 isolates did not differ molecularly from both 2008 and 2003 isolates (PhiPT ≤ 0.078, P ≥ 0.076).
+NE wmn isolates sampled in 2004 and 2008 were genetically similar (PhiPT = 0.042, P = 0.179) (DAPC results not shown).
+The standardized index of association of wmn isolates from each region (i.e.
+MI, WA, and NE) for above mentioned sampling years suggested clonal reproduction ( ≥ 0.05, P ≤ 0.002).
+
+# Discussion {-}
+
+Our study was conducted to characterize *S.
+sclerotiorum* isolates from different geographic areas to determine important phenotypic variables influencing genotypes; to ascertain reproductive behavior of the pathogen; to test whether different sampling years or locations have any effect on genotype/sexuality; and to identify abundant and unique genotypes across isolates and geographies.
+The dbRDA revealed region, year, and host as the most influential phenotypic variables on genotypes.
+However, further molecular analysis of pathogen isolates from selected hosts did not show any significant genetic difference among populations.
+According to dbRDA, virulence and location of isolates had little or no effect.
+Previous studies on *S.
+sclerotiorum* from alfalfa soil (Attanayake et al., 2012) and potato (Atallah et al., 2004), and canola (Attanayake et al., 2013) plants reported little or no correlation of genotypes on pathogen virulence.
+This is important information for resistant breeding since genotypes of *S.
+sclerotiorum* isolates seem to depend to a certain degree on different levels of region and also on environmental conditions of the sampling year.
+For instance our analysis indicated that MX, 2009, and 2005 were highly influential predictors of *S. sclerotiorum* genotypes (Fig. 7).
+Conversely, France and Australia, which are geographically distant to North America, were not within the top 10 predictors of genetic composition of isolates (Fig. 7).
+According to dbRDA, Mexican region imposed the highest impact on microsatellite loci of *S. sclerotiorum* isolates.
+We used Bruvo's distance for calculating genetic distances of different isolates since it incorporates stepwise mutation model and applicable to microsatellite analysis.
+
+The 366 isolates of our study belonged to 87 MCGs and 165 MLGs.
+MCGs with multiple isolates were represented by several to many microsatellite genotypes and demonstrated lack of similarity between MCGs and microsatellite genotypes.
+For instance, MCG 5 with the highest number of isolates (N = 73) represented 37 MLGs when characterized by 11 microsatellites.
+The most common genotype among these 73 isolates was MLG 25 with 19 isolates.
+Similarly, second most common MCG with 44 isolates was represented by 19 genotypes.
+Some studies on *S.
+sclerotiorum* found a strong association between MCG and multilocus genotypes (Aldrich-Wolfe et al., 2015) while others reported no evidence for such consistent association (Atallah et al., 2004; Attanayake et al., 2012) or either reported an association for a part of the sample (Kohn, 1995).
+Attanayake et al. (2013) carried out a genetic diversity study of *S. sclerotiorum* populations from canola in China and USA.
+They reported the MCGs were in accordance with the microsatellite MLGs in the U.S. population but not in the Chinese population.
+A study conducted on Canadian field populations of cultivated canola revealed MCGs and RFLP fingerprinting of *S.
+sclerotiorum* isolates were associated but MCGs of Norwegian isolates from wild perennial plant Ranunculua ficaria did not show such an association (Kohn, 1995).
+Aldrich-Wolfe et al.(2015) characterized 145 isolates from four different host crops in north central region of the USA.
+Isolates belonging to same MCG either shared identical genotypes or differed at no more than 2 of the 12 microsatellite loci.
+They reported only 14% of MCGs contained isolates with different genotypes.
+Different results were reported by Attanayake et al. (2012), who characterized 40 *S. sclerotiorum* isolates from 1 m^2^ area of alfalfa seed production soil using microsatellite markers.
+There were 15 MCGs with seven representing just a single isolate.
+Though single-isolate MCGs contained distinct multilocus genotype, they also found several instances where MCGs and microsatellite MLGs did not similarly correspond.
+Some isolates of the same multilocus genotype belonged to different MCGs.
+Our results agree with this type of relationship between MCGs and MLGs.
+Similarly, Attallah et al.
+(2004) did not find any strong correlation between microsatellite haplotypes and MCGs in 167 *S.
+sclerotiorum* isolates from four potato fields in the Columbia basin of WA state.
+These 167 isolates belonged to 82 MCGs with 64 single-isolate MCGs.
+We came across 43 single-isolate MCGs among 366 isolates from dry bean cultivars.
+
+Our study did not find any multilocus genotype/s common across all or majority of sampling geographies.
+The most widespread and common genotype (MLG 25, Supplemental Table 1) was found only in five states of the USA with majority of isolates confined to ND.
+However, we did find some MLGs of France and Australian isolates to be present in the USA samples.
+Some studies on *S. sclerotiorum* did not detect any shared MLGs between geographically distant populations.
+Attanayake et al. (2013) were unable to detect any shared microsatellite MLGs in *S. sclerotiorum* populations from infected canola fields in China and USA.
+Isolate collections were limited to a single county in China whereas USA isolates were from ND state only.
+This may have limited the chance of discovering genotypes common to both regions.
+Atallah et al.
+(2004) examined the population structure of *S.
+sclerotiorum* in different potato fields of Columbia basin and reported only 15 isolates shared the most common microsatellite haplotype in a sample of 167.
+Similarly, several other studies did not detect common fingerprints of *S.
+sclerotiorum* isolates in geographically distant populations: Australia and North America (Ekins, 1999);  California, Washington, and Ontario (Malvarez et al., 2007); Canada, North Carolina, and Louisiana (Cubeta et al., 1997).
+
+Contrary to above results Aldrich-Wolfe et al.
+(2015) who did a survey on *S.
+sclerotiorum* from different host crops in the North Central states of the United States found a single microsatellite haplotype to be prevalent throughout the region.
+A study in the UK that examined this pathogen from 12 populations in different crop plants and weed species *Ranunculus acris* (meadow buttercup) also reported one microsatellite haplotype was widely distributed across hosts and areas (Clarkson et al., 2013).
+Reason for our isolates from dry bean not to show a widespread MLG may be due to extensive sexual outcrossing and movement sclerotia by human activities.
+Though conditions of outcrossing for *S.
+sclerotiorum* is not well studied, it is believed weather and agricultural conditions play a role in production and release of ascospores of this pathogen (Huang, 1991; Sexton et al., 2006).
+One study found release of spores from apothecia declined over 36- 168 H at low RH of 65-75% but not at high RH of 90-95% (Clarkson et al., 2003).
+The same study reported increased ascospores discharge at 20-25° C compared to 15° C and the survival of *S. sclerotiorum* ascospores was reduced at high temperature and RH.
+Our samples were collected from 2003 to 2012 and except for the dry bean nurseries, came from different fields.
+Different agro-ecological conditions prevailed during sampling seasons may have resulted in different rates of sexual outcrossing of the pathogen and hence reduced the chance of extracting a common genotype.
+When conditions are favorable for ascospore production, release, and survival, coinfection of a single site on a plant by multiple ascospores is possible (Sexton et al., 2006).
+This may eventually result in facilitating outcrossing and recombination.
+
+Interestingly, all Mexican genotypes were unique to that region.
+On the other hand CA genotypes were present in the West Coast states of OR and WA which may be due to movement of sclerotia with planting material.
+Surprisingly, France and Australian genotypes were also isolated in the USA.
+MLG 4 (see supplemental Table 1) was recorded from France, Australia and MN while France and USA isolates had four common MLGs.
+Similarly, five MLGs were recorded common to both Australia and USA isolates.
+However, there were six unique genotypes in clone-corrected France population (N= 6/14) and three genotypes in CA population (N = 3/15).
+MLGs from Australian region, our samples were limited to 6, which is not a sufficient number to make inferences.
+Our results indicated that different continents were not a defining factor for microsatellite genotypes of *S. sclerotiorum* isolated from dry bean.
+This may be due to involuntary movement of sclerotia by human activities and probably because of recent speciation of *S. sclerotiorum*. 
+A phylogenetic study on Sclerotinaceae showed that most of the sclerotial lineages including different species of Sclerotinia had evolved relatively recently from a common ancestor (Carbone and Kohn, 1993).
+Few multilocus genotypes were unique to each sampling state of the USA but most of them were common among several states (Supplemental Table 1).
+For instance, Michigan state from which we had the largest collection of isolates (N = 62) consisted of 43 MLGs.
+Of these, 32 MLGs were present in at least one of the 5 states of NE, ND, WA, CO, MN; 4 MLGs were present in France isolates; and 11 were unique to MI sites (see Supplemental Table 1 for unique MLGs in different regions).
+When selecting *S. sclerotiorum* isolates for screening dry bean germplasm for disease resistance, plant breeders may find above information useful.
+A study on genetic diversity of *S.
+sclerotiorum* from multiple host crops in the North central region of the USA revealed there was little evidence for genetic isolation of this pathogen by host crop or by distance (Aldrich-Wolfe et al., 2015).
+Though our isolates didn't come from different crops, we had isolates from many resistant and susceptible dry bean cultivars, which also didn't show genetic differentiation according to AMOVA.
+It seems Mexican pathogen isolates are different from the rest and warrants further investigation using a larger population.
+Some France genotypes were common in USA isolates but at the same time 43% (6/14) of the France genotypes were unique to that region indicating white mold resistant cultivars produced in the USA might not work in that region.
+Geographic distance between different states of the USA seem not a barrier for gene transfer of *S. sclerotiorum* isolates though unique genotypes did prevail at a low frequency in each state.
+The genetic diversity of regions with >10 isolates had similar values (0.43 - 0.59) except for Mexican isolates.
+Hexp of MX was 0.26 and indicated less exposure of sampling sites to outside inocula.
+Test for linkage disequilibrium showed evidence of clonal reproduction for MX isolates (Ia = 2.61, P.Ia = 0.001,  = 0.37, P. = 0.001).
+This may have been another reason for low gene diversity.
+
+DAPC analysis by region followed by AMOVA corroborated the uniqueness of Mexican genotypes by grouping them away from the rest of the isolates (Fig. 8).
+Other populations based on geography were unable to group in separate clusters except for the Californian population.
+We tested whether genotypes of *S. sclerotiorum* populations were influenced by different dry bean host cultivars and year of sampling by DAPC and AMOVA.
+There was no host specificity but year of sampling did show a significant impact on genotypes.
+Other studies also reported host identity is no barrier for gene flow of this pathogen (Aldrich-Wolfe et al. 2015; Clarkson et al. 2013).
+Atallah et al. (2004) revealed statistically significant impact of the collection date of *S. sclerotiorum* isolates from four potato fields in the Columbia Basin on molecular variance.
+But there was no impact of the year or location of collection.
+We also did not detect location as a significant predictor of molecular variance but year had a significant impact.
+Our samples were collected on a single day in each year from each field and therefore did not warrant for an analysis of date of sample collection.
+
+We performed DAPC without assigning isolates to different populations based on sampling information to see the how isolates fall into clusters according to their genetic relatedness.
+We decided the most appropriate number as six since the DAPC results had 87% (N = 13/15) of Mexican isolates in a single cluster (Fig. 9 cluster 3).
+When number of clusters was assigned less than that MX isolates did not group separately while higher number of cluster assignment resulted in breaking down of other clusters in Fig.
+9 into smaller groups.
+DAPC analysis without population assignments revealed that in general geography is not a barrier for gene flow since isolates of different geographies from USA, Australia and France clustered together indicating genetic relatedness and possible sexual recombination within each region.
+For instance France isolates were present in cluster 1, 2, 4 and 5.
+As mentioned above only exception was the Mexican isolates.
+Njambere et al. (2014) studied S. trifoliorum, a close relative *S. sclerotiorum* and reported four cryptic genetic populations by Bayesian clustering analysis of microsatellite loci.
+Those clusters had no correlation to geographic locations of samples.
+Similarly, a cluster analysis of microsatellite data of *S.
+sclerotiorum* collected from 12 geographically distant host populations in  England and Wales resulted only two main clusters pooling isolates from different populations together (Clarkson et al., 2013).
+However, some other studies found *S.
+sclerotiorum* isolates from distant locations did differ significantly when molecularly characterized (Attanayake et al., 2013; Cubeta et al., 1997; Ekins, 1999; Malvarez et al., 2007).
+
+There were isolates that shared the same MLG belonging to different MCGs.
+For instance MLG 25 consisted of 27 isolates in 5 MCGs.
+This phenomenon is suggestive of a recombination by sexual outcrossing (Njambere et al., 2014; Sexton et al., 2006).
+In general high level of MLG diversity; and lack of association of MCG and microsatellite genotype are evidence of *S.
+sclerotiorum* undergoing sexual outcrossing and movement of isolates due to agricultural practices(Njambere et al., 2014).
+Our isolates had a total of 165 MLGs and they did not correspond well with 87 MCGs.
+Linkage disequilibrium analysis of populations in white mold nurseries indicated the possibility of sexual outcrossing in some regions (ex.
+CA, FR, OR and WA) during certain sampling years while most locations indicated clonal reproduction.
+For example wmn of WA was sampled in 2003, 2004, 2005, and 2008 but only 2008 isolates had evidence of outcrossing.
+Similar to our results many studies have reported primarily clonal reproduction of *S.
+sclerotiorum* populations with evidence of occasional outcrossing:  canola and sunflower pathogen populations in north central region of the USA (Aldrich-Wolfe et al., 2015); cabbage population in North Carolina (Cubeta et al., 1997); Canadian canola and North Carolina cabbage populations (Kohli and Kohn, 1998); and 12 populations from five crops in England and Wales (Clarkson et al., 2013).
+Contrary to above, one study from an experimental cabbage field plot in Louisiana, found a population of 16 isolates of *S. sclerotiorum* were strictly clonal (Cubeta et al., 1997).
+Attalah et al. (2004) studied the genetic structure of *S. sclerotiorum* in four potato fields in the Columbia basin and reported evidence of potential outcrossing of populations in two fields.
+When all isolates were combined, there was significant linkage disequilibrium and hence the population was highly clonal.
+Our study supports this phenomenon since linkage disequilibrium analysis revealed evidence of sexual outcrossing only in CA when populations were based on region.
+However, when populations were sub divided according each year/region, we could detect evidence of sexual out crossing in some populations.
+One reason for this is the overshadowing of low frequency genotypes of heterothallic recombination by a few successful genotypes of clonal nature (Atallah et al, 2004).
+
+*Sclerotinia sclerotiorum* isolates from dry bean represented high microsatellite diversity and lack of association MCGs and microsatellite MLGs.
+Analysis of molecular variance of populations based on region and sampling year showed low level of differentiation with majority of variance detected within populations.
+These features are indicative of sexual outcrossing and perhaps human mediated migration of *S. sclerotiorum* in dry bean.
+However, we also detected clones or identical microsatellite MLGs within and among populations indicating asexual reproduction.
+Further, linkage disequilibrium and index of association calculations suggested *S. sclerotiorum* populations are largely clonal.
+Although, there was low impact of region and year on population differentiation, we did not detect host cultivar as a barrier to gene flow of the pathogen.
+Our study did not find a widespread genotype among sampled regions from dry beans but some Australian and France MLGs were found in the USA isolates.
+The analysis of populations from same field over different years for linkage disequilibrium did not provide evidence for sexual outcrossing.
+However, analysis of molecular variance did show populations were significantly different between some years (eg.
+MI 2003 and 2005 populations; WA 2008 compared to both 2003 and 2005).
+These results indicate some level of heterothalic recombination is taking place in these fields in addition to natural mutations.
+An understanding of sexual behavior and ability of form new genotypes by *S. sclerotiorum* in dry bean cropping systems is useful for formulating disease management strategies.

### 2017-06-27 doc/manuscript/manuscript.Rmd
minor edits

@@
 The bean nurseries were 5 m x 10 m in size and maintained without applying fungicides to observe natural incidence of white mold disease.
-The nursery plots were incorporated with a basal dressing of N: P: K = 1: 3: 2 and side dressing of 0:3:2 during the season.
+The nursery plots were incorporated with a basal dressing of N:P:K = 1:3:2 and side dressing of 0:3:2 during the season.
 Although the nursery locations were the same over sampling years, sampling fields within a location varied for sampling years.
@@
 Community analysis of the MLGs shared between populations revealed 4 communities with a modularity of 0.17 (Fig.
-3Z): A costal community (CA, OR, WA, and NY), a mid-west community (CO, NE, ND, MI), and an international community (MN, France, Australia).
+3Z): A coastal community (CA, OR, WA, and NY), a mid-west community (CO, NE, ND, MI), and an international community (MN, France, Australia).
 Analysis with 16 loci resulted in the removal of the NY node because it no longer shared a genotype with OR, but the same overall community structure was present with a modularity of 0.2 (Supplementary Figure S1Z).
@@
 
-The dapc for hosts was limited to six most common cultivars: G122 (N = 55), Beryl (N = 62), Bunsi (N = 47), Yellow (N = 26), Pinto (N = 15), and Merlot (N = 11).
+The DAPC for hosts was limited to six most common cultivars: G122 (N = 55), Beryl (N = 62), Bunsi (N = 47), Yellow (N = 26), Pinto (N = 15), and Merlot (N = 11).
 According to Fig.
@@
 Different agro-ecological conditions prevailed during sampling seasons may have resulted in different rates of sexual outcrossing of the pathogen and hence reduced the chance of extracting a common genotype.
-When conditions are favorable for ascospore production, release, and survival, coinfection of a single site on a plant by multiple ascospores is possible (Sexton et al., 2006).
+When conditions are favorable for ascospore production, release, and survival, co-infection of a single site on a plant by multiple ascospores is possible (Sexton et al., 2006).
 This may eventually result in facilitating outcrossing and recombination.
@@
 As mentioned above only exception was the Mexican isolates.
-Njambere et al. (2014) studied S. trifoliorum, a close relative *S. sclerotiorum* and reported four cryptic genetic populations by Bayesian clustering analysis of microsatellite loci.
+Njambere et al. (2014) studied *S. trifoliorum*, a close relative *S. sclerotiorum* and reported four cryptic genetic populations by Bayesian clustering analysis of microsatellite loci.
 Those clusters had no correlation to geographic locations of samples.
@@
 MI 2003 and 2005 populations; WA 2008 compared to both 2003 and 2005).
-These results indicate some level of heterothalic recombination is taking place in these fields in addition to natural mutations.
+These results indicate some level of heterothallic recombination is taking place in these fields in addition to natural mutations.
 An understanding of sexual behavior and ability of form new genotypes by *S. sclerotiorum* in dry bean cropping systems is useful for formulating disease management strategies.

### 2017-06-27 doc/manuscript/manuscript.Rmd
update manuscript to be peerj style

@@
 title: Population structure and phenotypic variation of *Sclerotinia sclerotiorum* from dry bean throughout the United States, and from Mexico, France, and Australia
+preprint: false
 author:
@@
     email: zkamvar@unl.edu
-    affiliation: University of Nebraska-Lincoln
+    affiliation: 1
   - name: Bimal Sajeewa Amadarasa
     email: bamadarasa2@unl.edu
-    affiliation: University of Nebraska-Lincoln
+    affiliation: 1
   - name: R. Jhala
     email: rjhala2@unl.edu
-    affiliation: University of Nebraska-Lincoln
+    affiliation: 1
   - name: James Steadman
     email: jsteadman1@unl.edu
-    affiliation: University of Nebraska-Lincoln
+    affiliation: 1
   - name: Sydney E. Everhart
     email: everhart@unl.edu
-    affiliation: University of Nebraska-Lincoln
-    corresponding: everhart@unl.edu
-address:
-  - code: University of Nebraska-Lincoln
+    affiliation: 1
+    corresponding: true
+affiliation:
+  - code: 1
     address: Department of Plant Pathology, University of Nebraska-Lincoln, Lincoln 68583
@@
   `r paste(readLines("abstract.md"), collapse = "\n  ")`
-author_summary: |
-  Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-
+# author_summary: |
+#   Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+header-includes: >
+  \usepackage{indentfirst}
 bibliography: mybibfile.bib
-output: rticles::plos_article
-csl: plos.csl
+biblio-style: apalike
+output: 
+  rticles::peerj_article:
+    citation_package: natbib
+    base_format: bookdown::pdf_document2 # for using \@ref()
 ---
@@
 Economically important crops affected by this pathogen include sunflower, soybean, oilseed rape, dry bean, chickpea, peanut, dry pea, lentils and vegetables (Bolton et al., 2006).
-*Sclerotinia sclerotiorum* causes white mold disease on dry bean (Phaseolus vulgaris L.), which can be a yield-limiting in temperate climates (Steadman, 1983).
+*Sclerotinia sclerotiorum* causes white mold disease on dry bean (*Phaseolus vulgaris* L.), which can be a yield-limiting in temperate climates (Steadman, 1983).
 
@@
 However, there was evidence of linkage equilibrium of microsatellite loci in 2 locations and illustrated the possibility of outcrossing in this region.
+
+deleteme: @Dirac1953888
 
@@
 RDA is a multivariate linear regression analysis of independent (explanatory: eg. environmental and phenotypic) variables on dependent (non-explanatory or response) variables.
-In our case, a pairwise Bruvo's (Bruvo et al., 2004) distance matrix was generated from microsatellite data using poppr 2.4.1 (Kamvar et al., 2014; Kamvar et al., 2015) and employed it as the dependent matrix for RDA analysis.
+In our case, a pairwise Bruvo's [@bruvo2004simple] distance matrix was generated from microsatellite data using poppr 2.4.1 [@kamvar2014poppr; @kamvar2015novel] and employed it as the dependent matrix for RDA analysis.
 To identify environmental variable(s) correlated with genetic variation, a forward-backward selection process was applied with the vegan function ordistep to identify the set of variables that explained more variation than predicted with the global model containing all variables.
@@
 
-To assess the connectivity of shared genotypes across populations, we first tabulated the multilocus genotypes shared between at least two populations (defined as states or countries) with the poppr function mlg.crosspop() (Kamvar et al.
-2014).
+To assess the connectivity of shared genotypes across populations, we first tabulated the multilocus genotypes shared between at least two populations (defined as states or countries) with the poppr function mlg.crosspop() [@kamvar2014poppr].
 From these data, we were able to construct a graph with populations as nodes and shared genotypes as edges (connections) between the nodes using the R packages igraph version 1.0.1 (Csardi and Nepusz 2006), dplyr version 0.5.0 (Wickham and Francois 2016), and purrr version 0.2.2 (Wickham 2016).
@@
 Clone-corrected isolates sampled in different years from the white mold nurseries (wmn) were tested for population differentiation over time by DAPC analysis.
-Only WA, MI, and NE isolates were qualified since we filtered out the states that had at least ≥7 samples/year with at least two years of data for the analysis.
+Only WA, MI, and NE isolates were qualified since we filtered out the states that had at least $\geq$ 7 samples/year with at least two years of data for the analysis.
 Lastly, we pooled all isolates, clone corrected them (N = 165) and performed DAPC using the function find.clusters as described above to delineate how many clusters were formed by these unique genotypes.
@@
 Populations from different regions were compared with AMOVA, where differentiation was low among regions (PhiPT  = 0.11; P = 0.001).
-Pairwise comparisons of all populations according to region showed MX and CA populations were significantly different across all populations at P ≤ 0.001.
+Pairwise comparisons of all populations according to region showed MX and CA populations were significantly different across all populations at P $\leq$ 0.001.
 The PhiPT values of Pairwise comparisons between CA and other populations ranged from 0.116 to 0.389 and from 0.355 to 0.493 between MX and other population.
@@
 Since *S. sclerotiorum* isolates are predominantly clonal, we wanted to explore whether there was any introduction of new isolates from other fields or if there were evidence of sexual reproduction.
-The AMOVA results for different clusters of Fig. 10A revealed significance difference between WA 2008 isolates from the rest (PhiPT ≥ 0.206, P = 0.001).
+The AMOVA results for different clusters of Fig. 10A revealed significance difference between WA 2008 isolates from the rest (PhiPT $\geq$ 0.206, P = 0.001).
 The 2003 and 2005 isolates were genetically similar (PhiPT = 0.056, P = 0.129).
 MI isolates from 2008 and 2003 (Fig.
-10B) showed significant difference (PhiPT = 0.127, P = 0.04) while 2005 isolates did not differ molecularly from both 2008 and 2003 isolates (PhiPT ≤ 0.078, P ≥ 0.076).
+10B) showed significant difference (PhiPT = 0.127, P = 0.04) while 2005 isolates did not differ molecularly from both 2008 and 2003 isolates (PhiPT $\leq$ 0.078, P $\geq$ 0.076).
 NE wmn isolates sampled in 2004 and 2008 were genetically similar (PhiPT = 0.042, P = 0.179) (DAPC results not shown).
 The standardized index of association of wmn isolates from each region (i.e.
-MI, WA, and NE) for above mentioned sampling years suggested clonal reproduction ( ≥ 0.05, P ≤ 0.002).
+MI, WA, and NE) for above mentioned sampling years suggested clonal reproduction ( $\geq$ 0.05, P $\leq$ 0.002).
 
@@
 Previous studies on *S.
-sclerotiorum* from alfalfa soil (Attanayake et al., 2012) and potato (Atallah et al., 2004), and canola (Attanayake et al., 2013) plants reported little or no correlation of genotypes on pathogen virulence.
+sclerotiorum* from alfalfa soil (Attanayake et al., 2012) and potato (Atallah et al., 2004), and canola (Attanayake et al., 2013) plants reported little or no correlation of genotypes on pa25thogen virulence.
 This is important information for resistant breeding since genotypes of *S.
@@
 One study found release of spores from apothecia declined over 36- 168 H at low RH of 65-75% but not at high RH of 90-95% (Clarkson et al., 2003).
-The same study reported increased ascospores discharge at 20-25° C compared to 15° C and the survival of *S. sclerotiorum* ascospores was reduced at high temperature and RH.
+The same study reported increased ascospores discharge at 20-25$^{\circ}$ C compared to 15$^{\circ}$ C and the survival of *S. sclerotiorum* ascospores was reduced at high temperature and RH.
 Our samples were collected from 2003 to 2012 and except for the dry bean nurseries, came from different fields.

### 2017-06-28 doc/RMD/MLG-distribution.Rmd
rerun with correct locus configuration
This ensures the MLGs are the same as the paper.
Save figures as pdf instead of svg

@@
 load("data/sclerotinia_16_loci.rda")
-setPop(dat) <- ~Region
-dat
-locNames(dat)
-other(dat)$REPLEN
-head(other(dat)$meta)
-keeploci <- !locNames(dat) %in% colnames(corrected_loci)
-dat11 <- dat[loc = keeploci, mlg.reset = TRUE] # reducing to 11 loci and recalculating mlgs
-dat11
+
 ```
@@
 plot_mlg_graph(graph16loc$total, alt_layout) + labs(list(subtitle = "(16 loci)"))
-ggsave(filename = "results/figures/publication/FigureS1Z.svg", width = 88, height = 88, units = "mm", scale = 2)
+ggsave(filename = "results/figures/publication/FigureS1Z.pdf", width = 88, height = 88, units = "mm", scale = 2)
 ```
@@
 plot_mlg_graph(graph11loc$total, alt_layout) + labs(list(subtitle = "(11 loci)"))
-ggsave(filename = "results/figures/publication/Figure3Z.svg", width = 88, height = 88, units = "mm", scale = 2)
+ggsave(filename = "results/figures/publication/Figure3Z.pdf", width = 88, height = 88, units = "mm", scale = 2)
 ```

### 2017-06-28 doc/manuscript/manuscript.Rmd
Add community-graph figure

@@
 bibliography: mybibfile.bib
-biblio-style: apalike
 output: 
@@
 ---
+
+```{r setup, include = FALSE, echo = FALSE}
+knitr::opts_knit$set(eval.after = c('fig.cap', 'fig.scap'))
+knitr::opts_chunk$set(message = FALSE, warning = FALSE)
+```
 
@@
 Sclerotia are capable of surviving more than five years in soil (Steadman, 1983) and germinate carpogenically or myceliogenically to infect the plant when environmental conditions are favorable.
-A carpogenically geminated sclerotia gives rise to apothecia in which ascospores are formed.
-Ascospores infect senescent plant tissues that, when supplemented with cool and humid microclimate within the plant canopy, leads to mycelial infection of the host plant.
+A carpogenically germinated sclerotia gives rise to apothecia in which ascospores are formed.
+Ascospores infect senescent plant tissues that, when supplemented with cool and humid micro-climate within the plant canopy, leads to mycelial infection of the host plant.
 Consequently, flowering, which coincide with canopy closure is a critical factor of disease initiation since sensing flower parts fallen on to leaves, petioles or stems provide the required energy source for plant infection.
-Myceliogenically geminated sclerotia result in formation of hyphae, which also have potential for directly plant infection, though this is considered rare and less important within the disease cycle of the pathogen (Abawi and Grogan, 1975).
+Myceliogenically germinated sclerotia result in formation of hyphae, which also have potential for directly plant infection, though this is considered rare and less important within the disease cycle of the pathogen (Abawi and Grogan, 1975).
 Diseased bean fields initially show scattered wilted leaves.
@@
 However, there was evidence of linkage equilibrium of microsatellite loci in 2 locations and illustrated the possibility of outcrossing in this region.
-
-deleteme: @Dirac1953888
 
@@
 Canopy architecture and growth pattern of some cultivars enable improved disease avoidance (Kolkman and Kelly, 2002).
-Field trials in some regions of the US have shown G122 is white mold resistant and Beryl is susceptible (pers. comm. with J.
-Steadman).
+Field trials in some regions of the US have shown G122 is white mold resistant and Beryl is susceptible (pers. comm. with J. Steadman).
 Understanding genetic relationships and reproduction behavior of *S.
@@
 Community analysis of the MLGs shared between populations revealed 4 communities with a modularity of 0.17 (Fig.
-3Z): A coastal community (CA, OR, WA, and NY), a mid-west community (CO, NE, ND, MI), and an international community (MN, France, Australia).
+\@ref(community-graph)): A coastal community (CA, OR, WA, and NY), a mid-west community (CO, NE, ND, MI), and an international community (MN, France, Australia).
 Analysis with 16 loci resulted in the removal of the NY node because it no longer shared a genotype with OR, but the same overall community structure was present with a modularity of 0.2 (Supplementary Figure S1Z).
 The international community appears to be driven by MLG 4, which is shared between all three populations and has a Psex value of 2.87 x 10-5, in contrast to the abundant MLG 25, which has a Psex value of 0.0168.
+
+![Network of populations (nodes/circles) and their shared multilocus haplotypes (MLG) (edges/lines) genotyped over 11 loci. 
+Each node is labeled with **name (number of MLGs shared/number of MLGs total).** 
+The shade and area of the nodes are proportional to the number of unique MLGs within the node and the inner nodes are proportional to the number of private MLGs to the region. 
+Each edge represents a single MLG where its thickness represents the number of populations that share the MLG and the shade represents the value of $P_{sex}$, or the probability of encountering that MLG from two independent meiotic events.
+](../../results/figures/publication/Figure3Z.pdf){#community-graph width=50%}
 

### 2017-06-28 doc/manuscript/manuscript.Rmd
update manuscript with correct bibliography

@@
   \usepackage{indentfirst}
-bibliography: mybibfile.bib
+bibliography: ssc_bibliography.bib
 output: 
@@
 
-![Network of populations (nodes/circles) and their shared multilocus haplotypes (MLG) (edges/lines) genotyped over 11 loci. 
+<!-- community graph figure -->
+
+![
+Network of populations (nodes/circles) and their shared multilocus haplotypes (MLG) (edges/lines) genotyped over 11 loci. 
 Each node is labeled with **name (number of MLGs shared/number of MLGs total).** 
@@
 ](../../results/figures/publication/Figure3Z.pdf){#community-graph width=50%}
+
+<!-- end fig -->
 

### 2017-06-28 doc/RMD/MLG-distribution.Rmd
Add seeds to MLG-distribution

@@
 locus_table(dat)
+set.seed(2017-06-28)
 genotype_curve(dat, sample = 1000, quiet = TRUE)
@@
 
-plot_mlg_graph <- function(g, glayout = NULL, label = TRUE){
+plot_mlg_graph <- function(g, glayout = NULL, label = TRUE, seed = 2017-06-28){
+  set.seed(seed)
   if (is.null(glayout)){

### 2017-06-28 doc/RMD/RDA-analysis.Rmd
add popdiff
changed all svg figures to pdf

@@
     toc: true
+editor_options: 
+  chunk_output_type: inline
 ---
@@
 # @return a ggplot2 object from the scores 
-plot_dbrda <- function(db, arrows = 10){
+plot_dbrda <- function(db, arrows = 10, seed = 2017-06-28){
+  set.seed(seed)
   dbsum     <- scores(db, display = c("cn", "bp", "sites"), scaling = "sites")
@@
   theme(aspect.ratio = 1)
-FILE <- file.path(PROJHOME, "results", "figures", "publication", "Figure7Z.svg")
+FILE <- file.path(PROJHOME, "results", "figures", "publication", "Figure7Z.pdf")
 ggsave(filename = FILE, width = 88, height = 88, units = "mm", scale = 1.5)
@@
 
-```{r, echo = FALSE, results = "asis"}
+<details>
+<summary>Session Information</summary>
+
+```{r, echo = FALSE}
 options(width = 100)
-cat("<details><summary>Session Information</summary>\n\n```\n\n")
 devtools::session_info()
-cat("\n\n```\n\n</details>")
 ```
+
+</details>

### 2017-06-28 doc/RMD/mlg-mcg.Rmd
changed all svg figures to pdf

@@
 mcg_mlg_graph
-ggsave(file.path(PROJHOME, "results/figures/publication/FigureS2.svg"), 
+ggsave(file.path(PROJHOME, "results/figures/publication/FigureS2.pdf"), 
        width = 88*3, height = 88*3, units = "mm")
@@
 t5g
-ggsave(t5g, file = file.path("results", "figures", "publication", "FigureS2B.svg"), width = 88, height = 88, scale = 3, units = "mm")
+ggsave(t5g, file = file.path("results", "figures", "publication", "FigureS2B.pdf"), width = 88, height = 88, scale = 3, units = "mm")
 ```
@@
 severity_plot
-ggsave(severity_plot, filename = "results/figures/publication/FigureS3.svg", width = 183, unit = "mm")
+ggsave(severity_plot, filename = "results/figures/publication/FigureS3.pdf", width = 183, unit = "mm")
 ```

### 2017-06-28 doc/manuscript/manuscript.Rmd
Add seeds to MLG-distribution
add fig 7 to manuscript

@@
 The marginal effects of state, host, and year were also significant (P = 0.001), but virulence was not significant (P = 0.045).
-Much of the variation appears to be driven by isolates from Mexico (Fig. 7) and 2005, but it's worth noting that all Mexican isolates were collected in 2005.
+Much of the variation appears to be driven by isolates from Mexico (Fig. \@ref(rda-biplot)) and 2005, but it's worth noting that all Mexican isolates were collected in 2005.
 Variance partitioning of the independent variables indicated virulence to be the least influential factor with 0.1% contributing to explaining the variation of molecular data whereas the combination of variables accounted for 3.3%.
+
+![
+Biplot showing 10 most influential explanatory variables (arrows) overlayed on the first two eigenvectors of distance based redundancy analysis of *S. sclerotiorum* isolates. 
+The length of the arrows are directly proportional to the strength of the correlation between explanatory and molecular variables. 
+Points represent the 318 clone-corrected samples in ordination space. 
+](../../results/figures/publication/Figure7Z.pdf){#rda-biplot width=50%}
 

### 2017-06-28 doc/manuscript/manuscript.Rmd
Update bibliography and citations
Some issues I noticed:

 - There were differences in spelling of some author
   names (e.g. Atallah and Attalah)
 - There were two versions of sirjusingh2001characterisation
   because of the difference between the British
   and American spellings of "characterization"
 - Sometimes there weren't commas before the year
 - Sometimes there was no period after et al.

@@
 *Sclerotinia sclerotiorum* (Lib.) de Bary is an ascomycete plant pathogen found worldwide.
-This is a necrotrophic pathogen that is primarily homothallic (self-fertilization) (Bolton et al., 2006) (Sexton et al., 2006).
-It has a host range of more than 400 species of plants belonging to 75 families (Boland and Hall, 1994).
-Economically important crops affected by this pathogen include sunflower, soybean, oilseed rape, dry bean, chickpea, peanut, dry pea, lentils and vegetables (Bolton et al., 2006).
-*Sclerotinia sclerotiorum* causes white mold disease on dry bean (*Phaseolus vulgaris* L.), which can be a yield-limiting in temperate climates (Steadman, 1983).
+This is a necrotrophic pathogen that is primarily homothallic (self-fertilization) [@bolton2006sclerotinia; @sexton2006population].
+It has a host range of more than 400 species of plants belonging to 75 families [@boland1994index].
+Economically important crops affected by this pathogen include sunflower, soybean, oilseed rape, dry bean, chickpea, peanut, dry pea, lentils and vegetables [@bolton2006sclerotinia].
+*Sclerotinia sclerotiorum* causes white mold disease on dry bean (*Phaseolus vulgaris* L.), which can be a yield-limiting in temperate climates [@steadman1983white].
 
-Sclerotia are capable of surviving more than five years in soil (Steadman, 1983) and germinate carpogenically or myceliogenically to infect the plant when environmental conditions are favorable.
+Sclerotia are capable of surviving more than five years in soil [@steadman1983white] and germinate carpogenically or myceliogenically to infect the plant when environmental conditions are favorable.
 A carpogenically germinated sclerotia gives rise to apothecia in which ascospores are formed.
@@
 Consequently, flowering, which coincide with canopy closure is a critical factor of disease initiation since sensing flower parts fallen on to leaves, petioles or stems provide the required energy source for plant infection.
-Myceliogenically germinated sclerotia result in formation of hyphae, which also have potential for directly plant infection, though this is considered rare and less important within the disease cycle of the pathogen (Abawi and Grogan, 1975).
+Myceliogenically germinated sclerotia result in formation of hyphae, which also have potential for directly plant infection, though this is considered rare and less important within the disease cycle of the pathogen [@abawi1975source].
 Diseased bean fields initially show scattered wilted leaves.
@@
 Previous studies of *S.
-sclerotiorum* population biology in North America and Australia have reported clonal propagation of the pathogen (Cubeta et al., 1997; Ekins et al., 2011; Kohli and Kohn, 1998; Kohli et al., 1995) with evidence of outcrossing observed in some areas (Atallah et al., 2004; Kohli and Kohn, 1998; Malvarez et al., 2007; Sexton and Howlett, 2004).
+sclerotiorum* population biology in North America and Australia have reported clonal propagation of the pathogen [@cubeta1997clonality; @ekins2011population; @kohli1998random; @kohli1995clonal] with evidence of outcrossing observed in some areas [@atallah2004high; @kohli1998random; @malvarez2007new; @sexton2004microsatellite].
 One such study examined *S. sclerotiorum* population structure using microsatellite markers from potato fields in the Columbia Basin of Washington state.
 Due to high crop diversity and high presence of the pathogen, it was important to understand the potential for outcrossing in order to develop disease management strategies.
-Molecular analysis revealed 92% of the isolate variability was found within subpopulations (Atallah et al., 2004).
+Molecular analysis revealed 92% of the isolate variability was found within subpopulations [@atallah2004high].
 However, there was evidence of linkage equilibrium of microsatellite loci in 2 locations and illustrated the possibility of outcrossing in this region.
 
-A population study from Australian canola fields showed evidence of sexual outcrossing of *S. sclerotiorum* isolates (Sexton and Howlett, 2004).
-Another study characterized three populations of *S. sclerotiorum* from lettuce in California (CA) and Ontario (ON), and pea/lentil in Washington (WA) (Malvarez et al., 2007).
+A population study from Australian canola fields showed evidence of sexual outcrossing of *S. sclerotiorum* isolates [@sexton2004microsatellite].
+Another study characterized three populations of *S. sclerotiorum* from lettuce in California (CA) and Ontario (ON), and pea/lentil in Washington (WA) [@malvarez2007new].
 While all three populations were  clonal, the index of association (IA) based on DNA fingerprint was closer to zero for CA than it was for the other populations.
 Populations from Australian sunflower with evidence of clonal reproduction were reported, and despite populations sampled from fields separated 50 to 400 km and from both head and basal rots, all showed similar gene diversity, with authors concluding isolates representing different geographies belong to one large genetic population of *S.
-sclerotiorum* (Ekins et al., 2011).
+sclerotiorum* [@ekins2011population].
 
-Attanayake et al. (2013) studied population differentiation of *S. sclerotiorum* isolates infecting canola from China and the USA.
+@attanayake2013sclerotinia studied population differentiation of *S. sclerotiorum* isolates infecting canola from China and the USA.
 There were no shared microsatellite multilocus genotypes (MLG) between the two populations.
@@
 Additionally, phenotypic traits such as production of sclerotia, mycelial compatibility group (MCG), growth rate were also significantly different between the two populations.
-Genetic diversity of *S. sclerotiorum* from soy bean, canola, sunflower, and dry bean in the North Central region of the USA was studied by (Aldrich-Wolfe et al., 2015).
+Genetic diversity of *S. sclerotiorum* from soy bean, canola, sunflower, and dry bean in the North Central region of the USA was studied by [@aldrich-wolfe2015genetic].
 Microsatellite analysis revealed little host- specificity or genetic-isolation of *S.
@@
 No white mold resistant dry bean cultivars exist.
-Canopy architecture and growth pattern of some cultivars enable improved disease avoidance (Kolkman and Kelly, 2002).
+Canopy architecture and growth pattern of some cultivars enable improved disease avoidance [@kolkman2002agronomic].
 Field trials in some regions of the US have shown G122 is white mold resistant and Beryl is susceptible (pers. comm. with J. Steadman).
@@
 
-Microsatellite markers (simple sequence repeats or SSR) are codominant, polymorphic and also able to detect multiple alleles per locus (Atallah et al., 2004; Sirjusingh and Kohn, 2001a).
-Reproducibility of microsatellites is higher compared other fingerprinting techniques such as random amplified polymorphic DNA (RAPD) (Atallah et al., 2004).
+Microsatellite markers (simple sequence repeats or SSR) are codominant, polymorphic and also able to detect multiple alleles per locus [@atallah2004high; @sirjusingh2001characterisation].
+Reproducibility of microsatellites is higher compared other fingerprinting techniques such as random amplified polymorphic DNA (RAPD) [@atallah2004high].
 High mutagenic rate and multiallelic nature of microsatellites render suitability for inferring population structure of organisms and their potential for outcrossing.
 In the case of plant pathogens like *S.
-sclerotiorum* that is predominantly clonal and occasionally outcrossing, microsatellites have been used successfully to ascertain their ability to reproduce by sexual outcrossing (Atallah et al., 2004; Sexton and Howlett, 2004) and also to detect differences among clonally propagated individuals.
+sclerotiorum* that is predominantly clonal and occasionally outcrossing, microsatellites have been used successfully to ascertain their ability to reproduce by sexual outcrossing [@atallah2004high; @sexton2004microsatellite] and also to detect differences among clonally propagated individuals.
 
@@
 sclerotiorum* populations of dry beans originated from different continents were genetically different, and c) although previous studies show that *S.
-sclerotiorum* populations are largely clonal, few compare populations from different continents (Lehner and Mizubuti, 2017), c) To identify localized genotypes representative of each location and also investigate if there are widespread genotypes common among locations.
+sclerotiorum* populations are largely clonal, few compare populations from different continents **(Lehner and Mizubuti, 2017)**, c) To identify localized genotypes representative of each location and also investigate if there are widespread genotypes common among locations.
 d) To test potential for sexual reproduction in sub populations of the pathogen isolates under different conditions such as year or field in each region, year/region, field/region/year, and e) Test if pathogen within-field populations differentiate over time.
@@
 
-Mycelial growth of each isolate was generated by activating sclerotia according to the method described previously (Matheron and Porchas, 2005).
+Mycelial growth of each isolate was generated by activating sclerotia according to the method described previously [@matheron1989in-vitro].
 These isolates (N = 366) were phenotypically characterized by determining mycelial compatibility group MCG and virulence.
-MCG was determined through co-culturing isolates in pairs on PDA plates, where incompatibility of different MCG resulted in formation of a mycelial barrage (Kohn et al., 1990) and isolates belonging to the same MCG did not form a mycelial barrage.
-Virulence of each isolate was assessed using a straw test (Petzoldt and Dickson, 1996; Teran et al., 2006) that rates necrotic lesion on a scale of 1-9, with ascending numbers assigned for increasing aggressiveness.
+MCG was determined through co-culturing isolates in pairs on PDA plates, where incompatibility of different MCG resulted in formation of a mycelial barrage [@kohn1990mycelial] and isolates belonging to the same MCG did not form a mycelial barrage.
+Virulence of each isolate was assessed using a straw test [@petzoldt1996straw; @teran2006modified] that rates necrotic lesion on a scale of 1-9, with ascending numbers assigned for increasing aggressiveness.
 DNA was purified from mycelium of each isolate using a Qiagen DNeasy Plant Mini Kit (Qiagen Inc., Valencia, CA) according to the manufacturer's instructions.
@@
 
-We characterized *S. sclerotiorum* isolates for microsatellite polymorphism using 16 primer pairs developed by (Sirjusingh and Kohn, 2001b).
+We characterized *S. sclerotiorum* isolates for microsatellite polymorphism using 16 primer pairs developed by @sirjusingh2001characterisation.
 Polymerase chain reaction (PCR) was carried out as described using primers labeled with FAM fluorophore, and resulting PCR products were first resolved in a 1.5% agarose gel stained with ethidium bromide to ensure product was within the expected size range.
@@
 
-A distance-based redundancy analysis (dbRDA) (Legendre and Anderson, 1999) was performed with the function capscale in the vegan 2.4-3 package in R (https://CRAN.R-project.org/package=vegan) to delineate relative contribution of phenotypic, spatial, and temporal components in influencing genetic composition of the populations.
+A distance-based redundancy analysis (dbRDA) [@legendre1999distance] was performed with the function capscale in the vegan 2.4-3 package in R (https://CRAN.R-project.org/package=vegan) to delineate relative contribution of phenotypic, spatial, and temporal components in influencing genetic composition of the populations.
 dbRDA is a method of conducting constrained ordinations on data matrices consisting of non-Euclidean distance measures.
@@
 To assess the connectivity of shared genotypes across populations, we first tabulated the multilocus genotypes shared between at least two populations (defined as states or countries) with the poppr function mlg.crosspop() [@kamvar2014poppr].
-From these data, we were able to construct a graph with populations as nodes and shared genotypes as edges (connections) between the nodes using the R packages igraph version 1.0.1 (Csardi and Nepusz 2006), dplyr version 0.5.0 (Wickham and Francois 2016), and purrr version 0.2.2 (Wickham 2016).
+From these data, we were able to construct a graph with populations as nodes and shared genotypes as edges (connections) between the nodes using the R packages igraph version 1.0.1 **(Csardi and Nepusz 2006)**, dplyr version 0.5.0 **(Wickham and Francois 2016)**, and purrr version 0.2.2 **(Wickham 2016)**.
 Each node was weighted by the fraction of shared MLGs.
@@
 Edges were weighted by 1- Psex, where Psex is the probability of encountering the same genotype via two independent meiotic events (Arnaud-Haond 2007, Parks and Werth 1993).
-We then used the cluster_optimal() function from igraph to define communities (Csardi and Nepusz 2006).
-The graph was plotted using the R packages ggplot2 version 2.2.1 (Wickham 2009) and ggraph version 1.0.0 (Pedersen 2017).
+We then used the cluster_optimal() function from igraph to define communities **(Csardi and Nepusz 2006)**.
+The graph was plotted using the R packages ggplot2 version 2.2.1 **(Wickham 2009)** and ggraph version 1.0.0 **(Pedersen 2017)**.
 To ensure that we were able to capture the same community signal, we additionally ran this analysis including the 5 polymorphic markers described above.
@@
 DAPC is a multivariate method designed to identify and describe clusters of genetically related individuals.
-Discriminant analysis maximizes separation between groups while minimizing variation within groups (Jombart et al., 2010).
-The find.clusters function in Adegenet 2.0.1 (Jombart, 2008) uses a k-means clustering algorithm to identify the optimum number of clusters.
+Discriminant analysis maximizes separation between groups while minimizing variation within groups [@jombart2010discriminant].
+The find.clusters function in Adegenet 2.0.1 [@jombart2008adegenet] uses a k-means clustering algorithm to identify the optimum number of clusters.
 It was not able to identify a single optimum and identified a range of optimal clusters between 5-15.
@@
 Lastly, we pooled all isolates, clone corrected them (N = 165) and performed DAPC using the function find.clusters as described above to delineate how many clusters were formed by these unique genotypes.
-AMOVA was used to test the significance of clusters resulted by different DAPC analyses using GenAlEx version 6.41 (Peakall and Smouse, 2012).
+AMOVA was used to test the significance of clusters resulted by different DAPC analyses using GenAlEx version 6.41 [@peakall2012genalex].
 PhiPT which is an analogue of Wright's FST was used to ascertain the significance of populations.
-An FST value of 0.05 or less is generally considered as reasonably low and may be interpreted to mean that structuring between populations is weak (Balloux and Ugon-Moulin, 2002; Wright, 1978).
+An FST value of 0.05 or less is generally considered as reasonably low and may be interpreted to mean that structuring between populations is weak [@jost2008gst].
+<!-- **(Balloux and Ugon-Moulin, 2002; Wright, 1978)**.   -->
 
@@
 
-Genotypic evenness (Grunwald et al., 2003), which is a measure of distribution of genotype abundances and unbiased gene diversity (Hexp) (Nei, 1978) were calculated using poppr package in R.
+Genotypic evenness [@grunwald2003analysis], which is a measure of distribution of genotype abundances and unbiased gene diversity (Hexp) [@nei1978estimation] were calculated using poppr package in R.
 Microsatellite data were also tested for evidence of linkage disequilibrium, where significant disequilibrium between pairs of loci in a population is attributed to linkage between loci and may be an indication of clonal reproduction if loci are physically unlinked, whereas linkage equilibrium can be an indication of outcrossing populations.
-The linkage disequilibrium of clone-corrected data was determined by calculating the index of association (IA) and its standardized version () and a permutation approach with 999 permutations was used to test whether observed values were significantly different from data randomized at each locus (Agapow and Burt, 2001; Brown et al., 1980).
-Value of IA tends to increase with the number of loci and therefore we made our inferences based on the , which standardizes the number of loci tested and considered a more robust method (Agapow and Burt, 2001).
+The linkage disequilibrium of clone-corrected data was determined by calculating the index of association ($I_{A}$) and its standardized version ($\bar{r}_d$) and a permutation approach with 999 permutations was used to test whether observed values were significantly different from data randomized at each locus [@agapow2001indices; @brown1980multilocus].
+Value of IA tends to increase with the number of loci and therefore we made our inferences based on the , which standardizes the number of loci tested and considered a more robust method [@agapow2001indices].
 Higher observed  or IA values would indicate clonally reproducing populations.
@@
 Previous studies on *S.
-sclerotiorum* from alfalfa soil (Attanayake et al., 2012) and potato (Atallah et al., 2004), and canola (Attanayake et al., 2013) plants reported little or no correlation of genotypes on pa25thogen virulence.
+sclerotiorum* from alfalfa soil [@attanayake2012genetic] and potato [@atallah2004high], and canola [@attanayake2013sclerotinia] plants reported little or no correlation of genotypes on pa25thogen virulence.
 This is important information for resistant breeding since genotypes of *S.
@@
 Some studies on *S.
-sclerotiorum* found a strong association between MCG and multilocus genotypes (Aldrich-Wolfe et al., 2015) while others reported no evidence for such consistent association (Atallah et al., 2004; Attanayake et al., 2012) or either reported an association for a part of the sample (Kohn, 1995).
-Attanayake et al. (2013) carried out a genetic diversity study of *S. sclerotiorum* populations from canola in China and USA.
+sclerotiorum* found a strong association between MCG and multilocus genotypes [@aldrich-wolfe2015genetic] while others reported no evidence for such consistent association [@atallah2004high; @attanayake2012genetic] or either reported an association for a part of the sample [@kohn1995clonal].
+@attanayake2013sclerotinia carried out a genetic diversity study of *S. sclerotiorum* populations from canola in China and USA.
 They reported the MCGs were in accordance with the microsatellite MLGs in the U.S. population but not in the Chinese population.
 A study conducted on Canadian field populations of cultivated canola revealed MCGs and RFLP fingerprinting of *S.
-sclerotiorum* isolates were associated but MCGs of Norwegian isolates from wild perennial plant Ranunculua ficaria did not show such an association (Kohn, 1995).
-Aldrich-Wolfe et al.(2015) characterized 145 isolates from four different host crops in north central region of the USA.
+sclerotiorum* isolates were associated but MCGs of Norwegian isolates from wild perennial plant Ranunculua ficaria did not show such an association [@kohn1995clonal].
+@aldrich-wolfe2015genetic characterized 145 isolates from four different host crops in north central region of the USA.
 Isolates belonging to same MCG either shared identical genotypes or differed at no more than 2 of the 12 microsatellite loci.
 They reported only 14% of MCGs contained isolates with different genotypes.
-Different results were reported by Attanayake et al. (2012), who characterized 40 *S. sclerotiorum* isolates from 1 m^2^ area of alfalfa seed production soil using microsatellite markers.
+Different results were reported by @attanayake2012genetic, who characterized 40 *S. sclerotiorum* isolates from 1 m^2^ area of alfalfa seed production soil using microsatellite markers.
 There were 15 MCGs with seven representing just a single isolate.
@@
 Our results agree with this type of relationship between MCGs and MLGs.
-Similarly, Attallah et al.
-(2004) did not find any strong correlation between microsatellite haplotypes and MCGs in 167 *S.
-sclerotiorum* isolates from four potato fields in the Columbia basin of WA state.
+Similarly, @atallah2004high did not find any strong correlation between microsatellite haplotypes and MCGs in 167 *S. sclerotiorum* isolates from four potato fields in the Columbia basin of WA state.
 These 167 isolates belonged to 82 MCGs with 64 single-isolate MCGs.
@@
 Some studies on *S. sclerotiorum* did not detect any shared MLGs between geographically distant populations.
-Attanayake et al. (2013) were unable to detect any shared microsatellite MLGs in *S. sclerotiorum* populations from infected canola fields in China and USA.
+@attanayake2013sclerotinia were unable to detect any shared microsatellite MLGs in *S. sclerotiorum* populations from infected canola fields in China and USA.
 Isolate collections were limited to a single county in China whereas USA isolates were from ND state only.
 This may have limited the chance of discovering genotypes common to both regions.
-Atallah et al.
-(2004) examined the population structure of *S.
+@atallah2004high examined the population structure of *S.
 sclerotiorum* in different potato fields of Columbia basin and reported only 15 isolates shared the most common microsatellite haplotype in a sample of 167.
 Similarly, several other studies did not detect common fingerprints of *S.
-sclerotiorum* isolates in geographically distant populations: Australia and North America (Ekins, 1999);  California, Washington, and Ontario (Malvarez et al., 2007); Canada, North Carolina, and Louisiana (Cubeta et al., 1997).
+sclerotiorum* isolates in geographically distant populations: Australia and North America [@ekins1999genetic];  California, Washington, and Ontario [@malvarez2007new]; Canada, North Carolina, and Louisiana [@cubeta1997clonality].
 
-Contrary to above results Aldrich-Wolfe et al.
-(2015) who did a survey on *S.
+Contrary to above results, @aldrich-wolfe2015genetic who did a survey on *S.
 sclerotiorum* from different host crops in the North Central states of the United States found a single microsatellite haplotype to be prevalent throughout the region.
-A study in the UK that examined this pathogen from 12 populations in different crop plants and weed species *Ranunculus acris* (meadow buttercup) also reported one microsatellite haplotype was widely distributed across hosts and areas (Clarkson et al., 2013).
+A study in the UK that examined this pathogen from 12 populations in different crop plants and weed species *Ranunculus acris* (meadow buttercup) also reported one microsatellite haplotype was widely distributed across hosts and areas [@clarkson2013population].
 Reason for our isolates from dry bean not to show a widespread MLG may be due to extensive sexual outcrossing and movement sclerotia by human activities.
 Though conditions of outcrossing for *S.
-sclerotiorum* is not well studied, it is believed weather and agricultural conditions play a role in production and release of ascospores of this pathogen (Huang, 1991; Sexton et al., 2006).
-One study found release of spores from apothecia declined over 36- 168 H at low RH of 65-75% but not at high RH of 90-95% (Clarkson et al., 2003).
+sclerotiorum* is not well studied, it is believed weather and agricultural conditions play a role in production and release of ascospores of this pathogen [@huang1991induction; @sexton2006population].
+One study found release of spores from apothecia declined over 36- 168 H at low RH of 65-75% but not at high RH of 90-95% [@clarkson2003ascospore].
 The same study reported increased ascospores discharge at 20-25$^{\circ}$ C compared to 15$^{\circ}$ C and the survival of *S. sclerotiorum* ascospores was reduced at high temperature and RH.
@@
 Different agro-ecological conditions prevailed during sampling seasons may have resulted in different rates of sexual outcrossing of the pathogen and hence reduced the chance of extracting a common genotype.
-When conditions are favorable for ascospore production, release, and survival, co-infection of a single site on a plant by multiple ascospores is possible (Sexton et al., 2006).
+When conditions are favorable for ascospore production, release, and survival, co-infection of a single site on a plant by multiple ascospores is possible [@sexton2006population].
 This may eventually result in facilitating outcrossing and recombination.
@@
 This may be due to involuntary movement of sclerotia by human activities and probably because of recent speciation of *S. sclerotiorum*. 
-A phylogenetic study on Sclerotinaceae showed that most of the sclerotial lineages including different species of Sclerotinia had evolved relatively recently from a common ancestor (Carbone and Kohn, 1993).
+A phylogenetic study on Sclerotinaceae showed that most of the sclerotial lineages including different species of Sclerotinia had evolved relatively recently from a common ancestor [@carbone1993ribosomal].
 Few multilocus genotypes were unique to each sampling state of the USA but most of them were common among several states (Supplemental Table 1).
@@
 A study on genetic diversity of *S.
-sclerotiorum* from multiple host crops in the North central region of the USA revealed there was little evidence for genetic isolation of this pathogen by host crop or by distance (Aldrich-Wolfe et al., 2015).
+sclerotiorum* from multiple host crops in the North central region of the USA revealed there was little evidence for genetic isolation of this pathogen by host crop or by distance [@aldrich-wolfe2015genetic].
 Though our isolates didn't come from different crops, we had isolates from many resistant and susceptible dry bean cultivars, which also didn't show genetic differentiation according to AMOVA.
@@
 There was no host specificity but year of sampling did show a significant impact on genotypes.
-Other studies also reported host identity is no barrier for gene flow of this pathogen (Aldrich-Wolfe et al. 2015; Clarkson et al. 2013).
-Atallah et al. (2004) revealed statistically significant impact of the collection date of *S. sclerotiorum* isolates from four potato fields in the Columbia Basin on molecular variance.
+Other studies also reported host identity is no barrier for gene flow of this pathogen [@aldrich-wolfe2015genetic; @clarkson2013population].
+@atallah2004high revealed statistically significant impact of the collection date of *S. sclerotiorum* isolates from four potato fields in the Columbia Basin on molecular variance.
 But there was no impact of the year or location of collection.
@@
 As mentioned above only exception was the Mexican isolates.
-Njambere et al. (2014) studied *S. trifoliorum*, a close relative *S. sclerotiorum* and reported four cryptic genetic populations by Bayesian clustering analysis of microsatellite loci.
+@njambere2014genotypic studied *S. trifoliorum*, a close relative *S. sclerotiorum* and reported four cryptic genetic populations by Bayesian clustering analysis of microsatellite loci.
 Those clusters had no correlation to geographic locations of samples.
 Similarly, a cluster analysis of microsatellite data of *S.
-sclerotiorum* collected from 12 geographically distant host populations in  England and Wales resulted only two main clusters pooling isolates from different populations together (Clarkson et al., 2013).
+sclerotiorum* collected from 12 geographically distant host populations in  England and Wales resulted only two main clusters pooling isolates from different populations together [@clarkson2013population].
 However, some other studies found *S.
-sclerotiorum* isolates from distant locations did differ significantly when molecularly characterized (Attanayake et al., 2013; Cubeta et al., 1997; Ekins, 1999; Malvarez et al., 2007).
+sclerotiorum* isolates from distant locations did differ significantly when molecularly characterized [@attanayake2013sclerotinia; @cubeta1997clonality; @ekins1999genetic @malvarez2007new].
 
@@
 For instance MLG 25 consisted of 27 isolates in 5 MCGs.
-This phenomenon is suggestive of a recombination by sexual outcrossing (Njambere et al., 2014; Sexton et al., 2006).
+This phenomenon is suggestive of a recombination by sexual outcrossing [@njambere2014genotypic; @sexton2006population].
 In general high level of MLG diversity; and lack of association of MCG and microsatellite genotype are evidence of *S.
-sclerotiorum* undergoing sexual outcrossing and movement of isolates due to agricultural practices(Njambere et al., 2014).
+sclerotiorum* undergoing sexual outcrossing and movement of isolates due to agricultural practices [@njambere2014genotypic].
 Our isolates had a total of 165 MLGs and they did not correspond well with 87 MCGs.
@@
 Similar to our results many studies have reported primarily clonal reproduction of *S.
-sclerotiorum* populations with evidence of occasional outcrossing:  canola and sunflower pathogen populations in north central region of the USA (Aldrich-Wolfe et al., 2015); cabbage population in North Carolina (Cubeta et al., 1997); Canadian canola and North Carolina cabbage populations (Kohli and Kohn, 1998); and 12 populations from five crops in England and Wales (Clarkson et al., 2013).
-Contrary to above, one study from an experimental cabbage field plot in Louisiana, found a population of 16 isolates of *S. sclerotiorum* were strictly clonal (Cubeta et al., 1997).
-Attalah et al. (2004) studied the genetic structure of *S. sclerotiorum* in four potato fields in the Columbia basin and reported evidence of potential outcrossing of populations in two fields.
+sclerotiorum* populations with evidence of occasional outcrossing:  canola and sunflower pathogen populations in north central region of the USA [@aldrich-wolfe2015genetic]; cabbage population in North Carolina [@cubeta1997clonality]; Canadian canola and North Carolina cabbage populations [@kohli1998random]; and 12 populations from five crops in England and Wales [@clarkson2013population].
+Contrary to above, one study from an experimental cabbage field plot in Louisiana, found a population of 16 isolates of *S. sclerotiorum* were strictly clonal [@cubeta1997clonality].
+@atallah2004high studied the genetic structure of *S. sclerotiorum* in four potato fields in the Columbia basin and reported evidence of potential outcrossing of populations in two fields.
 When all isolates were combined, there was significant linkage disequilibrium and hence the population was highly clonal.
@@
 However, when populations were sub divided according each year/region, we could detect evidence of sexual out crossing in some populations.
-One reason for this is the overshadowing of low frequency genotypes of heterothallic recombination by a few successful genotypes of clonal nature (Atallah et al, 2004).
+One reason for this is the overshadowing of low frequency genotypes of heterothallic recombination by a few successful genotypes of clonal nature [@atallah2004high].
 

### 2017-06-28 doc/manuscript/manuscript.Rmd
update general formatting

@@
 Another study characterized three populations of *S. sclerotiorum* from lettuce in California (CA) and Ontario (ON), and pea/lentil in Washington (WA) [@malvarez2007new].
-While all three populations were  clonal, the index of association (IA) based on DNA fingerprint was closer to zero for CA than it was for the other populations.
+While all three populations were  clonal, the index of association ($I_A$) based on DNA fingerprint was closer to zero for CA than it was for the other populations.
 Populations from Australian sunflower with evidence of clonal reproduction were reported, and despite populations sampled from fields separated 50 to 400 km and from both head and basal rots, all showed similar gene diversity, with authors concluding isolates representing different geographies belong to one large genetic population of *S.
@@
 sclerotiorum* populations of dry beans originated from different continents were genetically different, and c) although previous studies show that *S.
-sclerotiorum* populations are largely clonal, few compare populations from different continents **(Lehner and Mizubuti, 2017)**, c) To identify localized genotypes representative of each location and also investigate if there are widespread genotypes common among locations.
+sclerotiorum* populations are largely clonal, few compare populations from different continents [@lehner2017sclerotinia], c) To identify localized genotypes representative of each location and also investigate if there are widespread genotypes common among locations.
 d) To test potential for sexual reproduction in sub populations of the pathogen isolates under different conditions such as year or field in each region, year/region, field/region/year, and e) Test if pathogen within-field populations differentiate over time.
@@
 
-A distance-based redundancy analysis (dbRDA) [@legendre1999distance] was performed with the function capscale in the vegan 2.4-3 package in R (https://CRAN.R-project.org/package=vegan) to delineate relative contribution of phenotypic, spatial, and temporal components in influencing genetic composition of the populations.
+A distance-based redundancy analysis (dbRDA) [@legendre1999distance] was performed with the function `capscale()` in the *vegan* 2.4-3 package in R (https://CRAN.R-project.org/package=vegan) to delineate relative contribution of phenotypic, spatial, and temporal components in influencing genetic composition of the populations.
 dbRDA is a method of conducting constrained ordinations on data matrices consisting of non-Euclidean distance measures.
 RDA is a multivariate linear regression analysis of independent (explanatory: eg. environmental and phenotypic) variables on dependent (non-explanatory or response) variables.
-In our case, a pairwise Bruvo's [@bruvo2004simple] distance matrix was generated from microsatellite data using poppr 2.4.1 [@kamvar2014poppr; @kamvar2015novel] and employed it as the dependent matrix for RDA analysis.
-To identify environmental variable(s) correlated with genetic variation, a forward-backward selection process was applied with the vegan function ordistep to identify the set of variables that explained more variation than predicted with the global model containing all variables.
+In our case, a pairwise Bruvo's [@bruvo2004simple] distance matrix was generated from microsatellite data using *poppr* 2.4.1 [@kamvar2014poppr; @kamvar2015novel] and employed it as the dependent matrix for RDA analysis.
+To identify environmental variable(s) correlated with genetic variation, a forward-backward selection process was applied with the vegan function `ordistep()` to identify the set of variables that explained more variation than predicted with the global model containing all variables.
 An ANOVA was then performed to test for significance of the reduced model and marginal effects using 999 permutations.
-The varpart function of vegan was used to determine variation partitioning of explanatory variables.
+The `varpart()` function of vegan was used to determine variation partitioning of explanatory variables.
 
@@
 
-To assess the connectivity of shared genotypes across populations, we first tabulated the multilocus genotypes shared between at least two populations (defined as states or countries) with the poppr function mlg.crosspop() [@kamvar2014poppr].
-From these data, we were able to construct a graph with populations as nodes and shared genotypes as edges (connections) between the nodes using the R packages igraph version 1.0.1 **(Csardi and Nepusz 2006)**, dplyr version 0.5.0 **(Wickham and Francois 2016)**, and purrr version 0.2.2 **(Wickham 2016)**.
+To assess the connectivity of shared genotypes across populations, we first tabulated the multilocus genotypes shared between at least two populations (defined as states or countries) with the *poppr* function `mlg.crosspop()` [@kamvar2014poppr].
+From these data, we were able to construct a graph with populations as nodes and shared genotypes as edges (connections) between the nodes using the R packages *igraph* version 1.0.1 **(Csardi and Nepusz 2006)**, *dplyr* version 0.5.0 **(Wickham and Francois 2016)**, and *purrr* version 0.2.2 **(Wickham 2016)**.
 Each node was weighted by the fraction of shared MLGs.
-Each edge represents a single MLG, but because a single MLG could cross populations, that MLG may be represented by (n*(n-1))/2 edges where n represents the number of populations it crosses.
-Edges were weighted by 1- Psex, where Psex is the probability of encountering the same genotype via two independent meiotic events (Arnaud-Haond 2007, Parks and Werth 1993).
-We then used the cluster_optimal() function from igraph to define communities **(Csardi and Nepusz 2006)**.
-The graph was plotted using the R packages ggplot2 version 2.2.1 **(Wickham 2009)** and ggraph version 1.0.0 **(Pedersen 2017)**.
+Each edge represents a single MLG, but because a single MLG could cross populations, that MLG may be represented by (n\*(n-1))/2 edges where n represents the number of populations it crosses.
+Edges were weighted by $1-P_{sex}$, where $P_{sex}$ is the probability of encountering the same genotype via two independent meiotic events [@arnaud2007standardizing, @parks1993study].
+We then used the `cluster_optimal()` function from *igraph* to define communities **(Csardi and Nepusz 2006)**.
+The graph was plotted using the R packages *ggplot2* version 2.2.1 **(Wickham 2009)** and *ggraph* version 1.0.0 **(Pedersen 2017)**.
 To ensure that we were able to capture the same community signal, we additionally ran this analysis including the 5 polymorphic markers described above.
@@
 Discriminant analysis maximizes separation between groups while minimizing variation within groups [@jombart2010discriminant].
-The find.clusters function in Adegenet 2.0.1 [@jombart2008adegenet] uses a k-means clustering algorithm to identify the optimum number of clusters.
+The `find.clusters()` function in *adegenet* 2.0.1 [@jombart2008adegenet] uses a k-means clustering algorithm to identify the optimum number of clusters.
 It was not able to identify a single optimum and identified a range of optimal clusters between 5-15.
@@
 PhiPT which is an analogue of Wright's FST was used to ascertain the significance of populations.
-An FST value of 0.05 or less is generally considered as reasonably low and may be interpreted to mean that structuring between populations is weak [@jost2008gst].
+An $F_{st}$ value of 0.05 or less is generally considered as reasonably low and may be interpreted to mean that structuring between populations is weak [@jost2008gst].
 <!-- **(Balloux and Ugon-Moulin, 2002; Wright, 1978)**.   -->
@@
 
-Genotypic evenness [@grunwald2003analysis], which is a measure of distribution of genotype abundances and unbiased gene diversity (Hexp) [@nei1978estimation] were calculated using poppr package in R.
+Genotypic evenness [@grunwald2003analysis], which is a measure of distribution of genotype abundances and unbiased gene diversity ($H_{exp}$) [@nei1978estimation] were calculated using *poppr* package in R [@kamvar2014poppr].
 Microsatellite data were also tested for evidence of linkage disequilibrium, where significant disequilibrium between pairs of loci in a population is attributed to linkage between loci and may be an indication of clonal reproduction if loci are physically unlinked, whereas linkage equilibrium can be an indication of outcrossing populations.
 The linkage disequilibrium of clone-corrected data was determined by calculating the index of association ($I_{A}$) and its standardized version ($\bar{r}_d$) and a permutation approach with 999 permutations was used to test whether observed values were significantly different from data randomized at each locus [@agapow2001indices; @brown1980multilocus].
-Value of IA tends to increase with the number of loci and therefore we made our inferences based on the , which standardizes the number of loci tested and considered a more robust method [@agapow2001indices].
-Higher observed  or IA values would indicate clonally reproducing populations.
+Value of $I_A$ tends to increase with the number of loci and therefore we made our inferences based on the , which standardizes the number of loci tested and considered a more robust method [@agapow2001indices].
+Higher observed  or $I_A$ values would indicate clonally reproducing populations.
 
@@
 Analysis with 16 loci resulted in the removal of the NY node because it no longer shared a genotype with OR, but the same overall community structure was present with a modularity of 0.2 (Supplementary Figure S1Z).
-The international community appears to be driven by MLG 4, which is shared between all three populations and has a Psex value of 2.87 x 10-5, in contrast to the abundant MLG 25, which has a Psex value of 0.0168.
+The international community appears to be driven by MLG 4, which is shared between all three populations and has a $P_{sex}$ value of 2.87 x 10^-5^, in contrast to the abundant MLG 25, which has a $P_{sex}$ value of 0.0168.
 
@@
 The 165 unique MLGs identified among all 366 isolates are given in Supplementary Table 1.
-Gene diversity (Hexp) of populations from different regions ranged from 0.259 to 0.596 (Table 5).
+Gene diversity ($H_{exp}$) of populations from different regions ranged from 0.259 to 0.596 (Table 5).
 Evenness of genotypes of all regions was high, with the exception of ND, where a single genotype was represented by 20 isolates.
@@
 
-Since find.clusters function of Adegenet library did not result in a clear-cut number of clusters for 318 isolates, we arrived at six (6) as the best number of populations (Fig.
+Since `find.clusters()` function of *adegenet* did not result in a clear-cut number of clusters for 318 isolates, we arrived at six (6) as the best number of populations (Fig.
 9) for our data after examining results of DAPC for different number of clusters.
@@
 The genetic diversity of regions with >10 isolates had similar values (0.43 - 0.59) except for Mexican isolates.
-Hexp of MX was 0.26 and indicated less exposure of sampling sites to outside inocula.
+$H_{exp}$ of MX was 0.26 and indicated less exposure of sampling sites to outside inocula.
 Test for linkage disequilibrium showed evidence of clonal reproduction for MX isolates (Ia = 2.61, P.Ia = 0.001,  = 0.37, P. = 0.001).

### 2017-06-28 doc/manuscript/manuscript.Rmd
Added notes

@@
 
+<!-- SEE: Avoid sentences that start with the authors/reference. -->
 @attanayake2013sclerotinia studied population differentiation of *S. sclerotiorum* isolates infecting canola from China and the USA.
@@
 The dry bean nurseries in different regions are indicated by wmn (white mold nursery) or mtc (Mitchell) and all other isolates came from commercial fields (Table 1).
-The bean nurseries were 5 m x 10 m in size and maintained without applying fungicides to observe natural incidence of white mold disease.
+The bean nurseries were 5m x 10 m in size and maintained without applying fungicides to observe natural incidence of white mold disease.
 The nursery plots were incorporated with a basal dressing of N:P:K = 1:3:2 and side dressing of 0:3:2 during the season.
@@
 Consequently, DAPC was performed several times with different optimal number of clusters to determine whether isolates (N = 318) fell into each cluster have any biological meaning such as belonging to a particular host or came from the same region etc.
-the best summaries for the total isolates (N = 318).
 Since dbRDA identified region, host, and sampling year as the top three variables influencing microsatellite data, DAPC was also performed on data defined according to each of these variables, separately.
@@
 AMOVA was used to test the significance of clusters resulted by different DAPC analyses using GenAlEx version 6.41 [@peakall2012genalex].
-PhiPT which is an analogue of Wright's FST was used to ascertain the significance of populations.
+$\Phi_{PT}$ which is an analogue of Wright's $F_{st}$ was used to ascertain the significance of populations.
 An $F_{st}$ value of 0.05 or less is generally considered as reasonably low and may be interpreted to mean that structuring between populations is weak [@jost2008gst].
@@
 The linkage disequilibrium of clone-corrected data was determined by calculating the index of association ($I_{A}$) and its standardized version ($\bar{r}_d$) and a permutation approach with 999 permutations was used to test whether observed values were significantly different from data randomized at each locus [@agapow2001indices; @brown1980multilocus].
-Value of $I_A$ tends to increase with the number of loci and therefore we made our inferences based on the , which standardizes the number of loci tested and considered a more robust method [@agapow2001indices].
-Higher observed  or $I_A$ values would indicate clonally reproducing populations.
+The value of $I_A$ tends to increase with the number of loci and therefore we made our inferences based on $\bar{r}_d$, which standardizes the number of loci tested and considered a more robust method [@agapow2001indices].
+Higher observed $\bar{r}_d$ or $I_A$ values would indicate clonally reproducing populations.
 
@@
 
-A total of 366 isolates were collected from 2003 to 2012 from diseased dry bean plants in 11 states in the US, Australia (2 mainland; 4 Tasmania), Mexico, and France (Fig. 1).
+<!--  ZNK: I think figures 1-4 are a bit redundant with the tables in the text. It might be better to simply remove them. -->
+A total of 366 isolates were collected from 2003 to 2012 from diseased dry bean plants in 11 states in the US, Australia (2 mainland; 4 Tasmania), Mexico, and France **(Fig. 1)**.
 Most isolates came from MI (62), ND (60), WA (59), NE (47), and CO (42).
@@
 The highest number of samples were collected in 2005 (N = 63) and lowest in 2012 (N = 4).
-Pathogen isolates represented 26 commercial cultivars and breeding lines with some representing unknown dry bean hosts (Fig. 2).
+Pathogen isolates represented 26 commercial cultivars and breeding lines with some representing unknown dry bean hosts **(Fig. 2)**.
 We investigated the population association with host for cultivars where more than 10 isolates were obtained, which included G122, Beryl, Bunsi, Yellow, Pinto, and Merlot.
@@
 
-The MCGs with the highest frequency were MCG 5 and MCG 44 representing 73 and 44 isolates, respectively (Fig. 4).
+The MCGs with the highest frequency were MCG 5 and MCG 44 representing 73 and 44 isolates, respectively **(Fig. 4)**. <!-- Maybe add aggressiveness information here. -->
 MCGs that were unique to a region are shown in Fig.
@@
 
-Populations from different regions were compared with AMOVA, where differentiation was low among regions (PhiPT  = 0.11; P = 0.001).
+Populations from different regions were compared with AMOVA, where differentiation was low among regions ($\Phi_{PT}$  = 0.11; P = 0.001).
 Pairwise comparisons of all populations according to region showed MX and CA populations were significantly different across all populations at P $\leq$ 0.001.
-The PhiPT values of Pairwise comparisons between CA and other populations ranged from 0.116 to 0.389 and from 0.355 to 0.493 between MX and other population.
-The CA population showed evidence of sexual recombination ( = 0.008; P = 0.27, Ia = 0.08; P = 0.3) while the  value of AU was 0.12 with a high P of 0.05 (Ia = 0.70; P = 0.05), whereas all other populations showed linkage disequilibrium for loci tested indicating high possibility of clonal reproduction.
+The $\Phi_{PT}$ values of Pairwise comparisons between CA and other populations ranged from 0.116 to 0.389 and from 0.355 to 0.493 between MX and other population.
+The CA population showed evidence of sexual <!-- ZNK: There is the sample size caveat here that should be discussed. -->
+recombination ($\bar{r}_d$ = 0.008; P = 0.27, $I_A$ = 0.08; P = 0.3) while the $\bar{r}_d$ value of AU was 0.12 with a high P of 0.05 ($I_A$ = 0.70; P = 0.05), whereas all other populations showed linkage disequilibrium for loci tested indicating high possibility of clonal reproduction.
 
@@
 AMOVA revealed genetic differences among isolates sampled in different years (DAPC biplot not shown).
-The PhiPT was a low value of 0.07 with P = 0.001.
+The $\Phi_{PT}$ was a low value of 0.07 with P = 0.001.
 Isolates from 2012 were not included in the analysis since there were only 4 samples.
@@
 We tested the clone corrected data for evidence of sexual outcrossing within *S. sclerotiorum* populations in different regions and different years/region.
-CA showed evidence of sexual outcrossing (Ia = 0.08, P.Ia = 0.317,  = 0.008, P. = 0.317) while all other regions implied clonal propagation (data not shown).
+CA showed evidence of sexual outcrossing (Ia = 0.08, P.Ia = 0.317, $\bar{r}_d$ = 0.008, P.$\bar{r}_d$ = 0.317) while all other regions implied clonal propagation (data not shown).
 Five populations showed evidence of sexual outcrossing: CA in 2004, FR in 2004 and 2005, OR in 2003, and WA in 2008.
@@
 Since *S. sclerotiorum* isolates are predominantly clonal, we wanted to explore whether there was any introduction of new isolates from other fields or if there were evidence of sexual reproduction.
-The AMOVA results for different clusters of Fig. 10A revealed significance difference between WA 2008 isolates from the rest (PhiPT $\geq$ 0.206, P = 0.001).
-The 2003 and 2005 isolates were genetically similar (PhiPT = 0.056, P = 0.129).
+The AMOVA results for different clusters of Fig. 10A revealed significance difference between WA 2008 isolates from the rest ($\Phi_{PT}$ $\geq$ 0.206, P = 0.001).
+The 2003 and 2005 isolates were genetically similar ($\Phi_{PT}$ = 0.056, P = 0.129).
 MI isolates from 2008 and 2003 (Fig.
-10B) showed significant difference (PhiPT = 0.127, P = 0.04) while 2005 isolates did not differ molecularly from both 2008 and 2003 isolates (PhiPT $\leq$ 0.078, P $\geq$ 0.076).
-NE wmn isolates sampled in 2004 and 2008 were genetically similar (PhiPT = 0.042, P = 0.179) (DAPC results not shown).
+10B) showed significant difference ($\Phi_{PT}$ = 0.127, P = 0.04) while 2005 isolates did not differ molecularly from both 2008 and 2003 isolates ($\Phi_{PT}$ $\leq$ 0.078, P $\geq$ 0.076).
+NE wmn isolates sampled in 2004 and 2008 were genetically similar ($\Phi_{PT}$ = 0.042, P = 0.179) (DAPC results not shown).
 The standardized index of association of wmn isolates from each region (i.e.
@@
 Our study did not find any multilocus genotype/s common across all or majority of sampling geographies.
+<!-- ZNK: I think this statement is a bit misleading.
+Yes, we didn't find any individual MLGs, but 46% of these are present across at least two populations.
+-->
 The most widespread and common genotype (MLG 25, Supplemental Table 1) was found only in five states of the USA with majority of isolates confined to ND.
@@
 On the other hand CA genotypes were present in the West Coast states of OR and WA which may be due to movement of sclerotia with planting material.
+<!-- ZNK: Do we have some sort of evidence for this, or is this pure speculation?
+-->
+<!-- 
+I'm not sure what the point of this paragraph is.
+I believe I can put the discussion of the graph results here. 
+-->
 Surprisingly, France and Australian genotypes were also isolated in the USA.
@@
 Some France genotypes were common in USA isolates but at the same time 43% (6/14) of the France genotypes were unique to that region indicating white mold resistant cultivars produced in the USA might not work in that region.
-Geographic distance between different states of the USA seem not a barrier for gene transfer of *S. sclerotiorum* isolates though unique genotypes did prevail at a low frequency in each state.
+<!-- ZNK: I'm not entirely sold on this conclusion.
+-->
+Geographic distance between different states of the USA seem not a barrier for gene transfer of *S.
+sclerotiorum* isolates though unique genotypes did prevail at a low frequency in each state.
+<!-- ZNK: I don't know that this adds anything.
+-->
 The genetic diversity of regions with >10 isolates had similar values (0.43 - 0.59) except for Mexican isolates.
-$H_{exp}$ of MX was 0.26 and indicated less exposure of sampling sites to outside inocula.
-Test for linkage disequilibrium showed evidence of clonal reproduction for MX isolates (Ia = 2.61, P.Ia = 0.001,  = 0.37, P. = 0.001).
+$H_{exp}$ of MX was 0.26 and indicated less exposure of sampling sites to outside inocula. <!-- ZNK: Does this mean that there's heavy inbreeding here? -->
+Test for linkage disequilibrium showed evidence of clonal reproduction for MX isolates (Ia = 2.61, P.Ia = 0.001, $\bar{r}_d$ = 0.37, P.$\bar{r}_d$  = 0.001).
 This may have been another reason for low gene diversity.
@@
 We also did not detect location as a significant predictor of molecular variance but year had a significant impact.
-Our samples were collected on a single day in each year from each field and therefore did not warrant for an analysis of date of sample collection.
+Our samples were collected on a single day in each year from each field and therefore did not warrant for an analysis of date of sample collection. <!-- BSA: Is this true, coauthors? -->
 

### 2017-06-28 doc/manuscript/manuscript.Rmd
update package citations

@@
 To assess the connectivity of shared genotypes across populations, we first tabulated the multilocus genotypes shared between at least two populations (defined as states or countries) with the *poppr* function `mlg.crosspop()` [@kamvar2014poppr].
-From these data, we were able to construct a graph with populations as nodes and shared genotypes as edges (connections) between the nodes using the R packages *igraph* version 1.0.1 **(Csardi and Nepusz 2006)**, *dplyr* version 0.5.0 **(Wickham and Francois 2016)**, and *purrr* version 0.2.2 **(Wickham 2016)**.
+From these data, we were able to construct a graph with populations as nodes and shared genotypes as edges (connections) between the nodes using the R packages *igraph* version 1.0.1 [@csardi2006igraph], *dplyr* version 0.7.1 [@dplyr], and *purrr* version 0.2.2.2 [@purrr].
 Each node was weighted by the fraction of shared MLGs.
@@
 Edges were weighted by $1-P_{sex}$, where $P_{sex}$ is the probability of encountering the same genotype via two independent meiotic events [@arnaud2007standardizing, @parks1993study].
-We then used the `cluster_optimal()` function from *igraph* to define communities **(Csardi and Nepusz 2006)**.
-The graph was plotted using the R packages *ggplot2* version 2.2.1 **(Wickham 2009)** and *ggraph* version 1.0.0 **(Pedersen 2017)**.
+We then used the `cluster_optimal()` function from *igraph* to define communities [@csardi2006igraph].
+The graph was plotted using the R packages *ggplot2* version 2.2.1 [@ggplot2] and *ggraph* version 1.0.0 [@ggraph].
 To ensure that we were able to capture the same community signal, we additionally ran this analysis including the 5 polymorphic markers described above.

### 2017-06-29 doc/RMD/table-1.Rmd
add huxtable to packages and create table 1

---
title: "Table 1"
output: 
  html_notebook:
    toc: true
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = PROJHOME)
```
# Creation of Table 1

Table one is one of those long, descriptive tables that requires a lot of
fiddling. The *huxtable* package makes working with these tables slightly less
of a pain in the neck. Because this table involves wrapping and the text
can get split up, I am manually copying and pasting the result into my document.

```{r isolate_table, results = "asis"}
library("readr")
library("dplyr")
library("huxtable")

clmn <- cols(
  .default = col_integer(),
  Region = col_character(),
  Source = col_character(),
  Host = col_character()
)

# Helper function to generate positions of blanks in a column for pretty printing
blank_it <- function(x){
  res <- rep(TRUE, length(x))
  res[1] <- FALSE
  res
}

dat <- readr::read_csv(file.path(PROJHOME, "data/clean_data.csv"), col_types = clmn) %>%
  mutate(Country = ifelse(nchar(Region) == 2, "USA", Region)) %>% # Defining country column
  mutate(Region = ifelse(nchar(Region) == 2, Region, "")) %>%     # Adding blanks for international "states"
  select(Country, Region, Source, Year, Host) %>%                 # Summarizing the number of isolates
  group_by_all() %>%                                              #
  summarize(N = n()) %>%                                          #
  arrange(desc(Country), Region, Source, Year) %>%                # Rearranging rows
  group_by(Country, Region, Source) %>%                           # Collapsing Years and Hosts
  summarize(Year = paste(sort(unique(Year)), collapse = ", "),     #
            Host = paste(sort(unique(Host)), collapse = ", "),    #
            N = sum(N)) %>%                                       #
  arrange(desc(Country), Region, Source, N) %>%                   # Rearranging rows
  group_by(Country, Region) %>%                                   # Adding blanks in repeated Country and Region Names
  mutate(set_blank = blank_it(Region)) %>%                        #
  ungroup() %>%                                                   #
  mutate(Country = ifelse(set_blank, "", Country)) %>%            #
  mutate(Region = ifelse(set_blank, "", Region)) %>%              #
  select(Country, Region, Source, Year, Host, N) %>%              #
  rename(State = Region) %>%                                      # Renaming columns
  rename(`Field Code` = Source)                                   #

dt <- huxtable::as_huxtable(dat, add_colnames = TRUE) %>% # Creating the huxtable
  set_align(everywhere, col = 4, value = "right") %>%     # Aligning numeric columns to the right
  set_align(everywhere, col = 6, value = "right") %>%     # 
  set_wrap(TRUE) %>%                                      # Wrapping columns
  set_number_format(value = 0)                            # Print numbers without decimals
col_width(dt) <- c(0.055, 0.035, 0.06, 0.1, 0.151, 0.025) # Specify width for each column
print_md(dt)                                              # print in pandoc markdown format
```

## Session Information

```{r session_info}
options(width = 100)
devtools::session_info()
```

### 2017-06-29 doc/manuscript/manuscript.Rmd
Add table 1 to manuscript

@@
 
-Sclerotia of *S. sclerotiorum* were collected over several years from grower fields and/or dry bean nurseries in Mexico, France, Australia, and 11 states of the United States of America (Table 1).
-The dry bean nurseries in different regions are indicated by wmn (white mold nursery) or mtc (Mitchell) and all other isolates came from commercial fields (Table 1).
+Sclerotia of *S.
+sclerotiorum* were collected over several years from grower fields and/or dry bean nurseries in Mexico, France, Australia, and 11 states of the United States of America (Table \@ref(tab:isolate-table)).
+The dry bean nurseries in different regions are indicated by wmn (white mold nursery) or mtc (Mitchell) and all other isolates came from commercial fields (Table \@ref(tab:isolate-table)).
 The bean nurseries were 5m x 10 m in size and maintained without applying fungicides to observe natural incidence of white mold disease.
@@
 Sampling year and number of isolates collected were necessarily influenced by incidence and severity of white mold disease.
+
+-----------------------------------------------------------------------------
+Country   State  Field Code              Year Host                          N 
+--------- ------ ---------- ----------------- -------------------------- ----
+USA       CA     wmn               2004, 2005 Beryl, Bunsi, G122           18
+
+USA       CO     eat               2007, 2010 Yellow                       13
+
+                 gree              2007, 2010 Pinto, Yellow                14
+
+                 luc               2007, 2010 Yellow                       14
+
+                 wmn                     2003 GH                            1
+
+USA       ID     unk                     2003 GH                            1
+
+USA       MI     eln                     2009 unk                           1
+
+                 hur               2008, 2009 BL, Black, SR06233, Vista     5
+
+                 mung                    2009 Zorro                         1
+
+                 sanc                    2008 Merlot                        3
+
+                 Sebewaing               2009 Black                         1
+
+                 tuc               2008, 2009 Black, Fuji, Vista            6
+
+                 unk               2003, 2009 GH, unk                       2
+
+                 wmn        2003, 2004, 2005, 11A, 37, 38, B07104,         43 
+                                   2008, 2009 Beryl, Bunsi, cornell,          
+                                              G122, Orion, PO7863, WM31
+
+USA       MN     wmn               2003, 2004 Beryl, Bunsi, G122           11
+
+USA       ND     gfc               2007, 2010 unk                           9
+
+                 nec                     2010 unk                           1
+
+                 pmc                     2010 unk                           7
+
+                 rrv                     2007 unk                          21
+
+                 stc                     2010 unk                           2
+
+                 tlc                     2010 unk                           9
+
+                 wlc                     2010 unk                           4
+
+                 wmn                     2005 Beryl, Bunsi, G122            7
+
+USA       NE     mtc               2007, 2010 Beryl, Emerson, Orion        12
+
+                 sbf               2007, 2009 Beryl, Pinto, Weihing         6
+
+                 unk                     2003 GH                            2
+
+                 wmn        2004, 2005, 2008, Beryl, Bunsi, G122,          27 
+                                         2010 PO7683, unk
+
+USA       NY     unk                     2003 GH                            1
+
+USA       OR     corv                    2003 G122                          1
+
+                 unk                     2003 GH                            1
+
+                 wmn               2003, 2004 Beryl, Bunsi, G122           15
+
+USA       WA     prsr                    2007 Merlot, Pinto, redkid        22
+
+                 unk                     2003 GH                            1
+
+                 wmn        2003, 2004, 2005, 11A, 37, 38, Beryl, Bunsi,   36 
+                                         2008 cornell, G122, Orion,           
+                                              PO7104, PO7863, WM31
+
+USA       WI     dfor                    2003 GH                            1
+
+                 unk                     2003 GH                            1
+
+Mexico           wmn                     2005 Beryl, Bunsi, G122           18
+
+France           flds                    2012 unk                           4
+
+                 wmn               2004, 2005 Beryl, Bunsi, G122           18
+
+Australia        vic                     2004 Beryl                         2
+
+                 wmn                     2004 Beryl, Bunsi, G122            4
+
+-----------------------------------------------------------------------------
+
+Table: (\#tab:isolate-table) Description of *S. sclerotiorum* isolates used in this study. 
+N = Number of Isolates. 
+Key abbreviations: wmn = white mold nursery, unk = unknown.
 

### 2017-06-29 doc/RMD/data-comparison.Rmd
Update data_comparison
I noticed that data/clean_data.csv did not have a
column for severity, so I gave it one. This allowed
me to do a comparison for the MCG and Severity.

I made changes sufficiently enough that the hacks I
used in previous documents weren't screwed up.

@@
   df2genind(ind.names = saj$Isolate, 
-            strata = select(saj, MCG:Host), # Filtering out severity and Isolate
+            strata = select(saj, Severity:Host), # Filtering out severity and Isolate
             ploidy = 1) %>%
   as.genclone()
-nameStrata(dat)[2] <- "Region"
+nameStrata(dat)[3] <- "Region"
 pops <- levels(strata(dat)$Region)
@@
   arrange(Isolate) %>%
-  select(-(Severity:Host))
+  select(-(MCG:Host))
 datdf <- genind2df(dat, usepop = FALSE) %>% 
@@
   bind_cols(datdf) %>%
-  as_data_frame() %>%
+  dplyr::as_data_frame() %>%
   readr::write_csv("data/clean_data.csv", col_names = TRUE)
@@
 
+> Note 2017-06-29: I realized that the severity data was not present in the
+> clean data, so I added that in the strata above. To avoid downstream effects,
+> I'm additionally removing it from the data set here:
+
 ```{r fix_strata}
@@
 other(dat)$meta <- select(saj, Severity, Isolate)
+strata(dat) <- select(strata(dat), -Severity)
 other(dat)$REPLEN <- fix_replen(dat, repeat_lengths)

### 2017-06-29 doc/RMD/MCG-virulence.Rmd
Add vignette to assess Virulence by MCG

---
title: "Assessing Virulence by MCG"
output: 
  html_notebook:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = PROJHOME)
```

```{r}
library("tidyverse")
```

# Purpose

This document is to assess virulence associated with the 10 most common MCGs. We
can get this from the cleaned data set that we saved in the
`data-comparison.Rmd` file.

```{r load_data}
cols(
  .default = col_integer(),
  Severity = col_double(),
  Region = col_character(),
  Source = col_character(),
  Host = col_character()
)
dat <- read_csv(file.path(PROJHOME, "data", "clean_data.csv")) %>%
  select(MCG, Severity)
dat
```

Now, I can filter out the top 10 MCGs:

```{r filtering}
top_mcg <- dat %>% 
  group_by(MCG) %>%
  summarize(N = n()) %>%
  arrange(desc(N)) %>%
  slice(1:10) %>% 
  inner_join(dat, by = "MCG") %>%
  select(-N) %>%
  mutate(MCG = forcats::fct_inorder(as.character(MCG)))
top_mcg
```

## Visualizing distributions

```{r vis, fig.width = 5, fig.height = 4}
ggplot(top_mcg, aes(x = MCG, y = Severity)) +
  geom_boxplot() +
  geom_point(position = position_jitter(width = 0.25), alpha = 0.5) +
  theme_bw(base_size = 16, base_family = "Helvetica") +
  theme(aspect.ratio = 1/2) +
  labs(list(
    title = "Severity for the top 10 MCGs",
    caption = "MCG = Mycelial Compatibility Group"
    ))
```

## AMOVA

The default AMOVA in R sets contrasts as `contrast.treatment`, which compares
everything to the first factor, considered the treatment. Since we are
interested in whether or not there IS a difference between samples, this will
be sufficient.

```{r amova}
AMOVA <- aov(Severity ~ MCG, data = top_mcg)
AMOVA
summary(AMOVA)
```

There appears to be a significant effect at p = `r formatC(broom::tidy(AMOVA)$p.value[1])`.

## Session Information

```{r session_info}
options(width = 100)
devtools::session_info()
```

### 2017-06-29 doc/RMD/table-1.Rmd
remove as-is output so it's easier to see on github
update table-1.Rmd to account for the new column

@@
 
-```{r isolate_table, results = "asis"}
+```{r isolate_table, comment = NA}
 library("readr")
@@
   .default = col_integer(),
+  Severity = col_double(),
   Region = col_character(),

### 2017-06-29 doc/RMD/MCG-virulence.Rmd
Add a summary table for severity

@@
 ```{r vis, fig.width = 5, fig.height = 4}
+set.seed(2017-06-29)
 ggplot(top_mcg, aes(x = MCG, y = Severity)) +
@@
     ))
+```
+
+```{r virulence_table}
+top_mcg %>% 
+  group_by(MCG) %>%
+  summarize(N = n(), 
+            `Min Severity` = min(Severity),
+            `Max Severity` = max(Severity),
+            `Average Severity` = mean(Severity)
+            ) %>%
+  knitr::kable(digits = 2)
 ```

### 2017-06-29 doc/RMD/MCG-virulence.Rmd
add Tukey Test

@@
 
-## AMOVA
+## ANOVA
 
-The default AMOVA in R sets contrasts as `contrast.treatment`, which compares
+The default ANOVA in R sets contrasts as `contrast.treatment`, which compares
 everything to the first factor, considered the treatment. Since we are
@@
 
-```{r amova}
-AMOVA <- aov(Severity ~ MCG, data = top_mcg)
-AMOVA
-summary(AMOVA)
+After the ANOVA, we performed a Tukey's Honest Significant Difference test to
+see exactly what groups these fell into.
+
+```{r anova, fig.width = 4, fig.height = 8}
+ANOVA <- aov(Severity ~ MCG, data = top_mcg)
+ANOVA
+summary(ANOVA)
+plot(TukeyHSD(ANOVA), las = 2)
+agricolae::HSD.test(ANOVA, "MCG")$groups
 ```
 
-There appears to be a significant effect at p = `r formatC(broom::tidy(AMOVA)$p.value[1])`.
+There appears to be a significant effect at p = `r formatC(broom::tidy(ANOVA)$p.value[1])`.
 

### 2017-06-29 doc/manuscript/manuscript.Rmd
begin rewrite of methods

@@
   `r paste(readLines("abstract.md"), collapse = "\n  ")`
-# author_summary: |
-#   Lorem ipsum dolor sit amet, consectetur adipiscing elit.
 header-includes: >
@@
 
-## Microsatellite genotyping Analysis {-}
+## Microsatellite genotyping {-}
 
@@
 
-## Data analysis {-}
+## Data processing and Analyses {-}
 
-Isolates that fell into the 10 most abundant MCGs were subjected to an analysis of variance (ANOVA) in JMP version 12 to determine if their virulence levels were significantly different.
-The mean separation was carried out using Tukey's HSD test.
-Five microsatellite loci included compound repeats, where genotypes could not be unequivocally assigned based on the known repeat motifs and were consequently removed, resulting in 11 polymorphic loci selected for analysis (Table 2).
-Since it could not be ruled out that isolates possessing the same genotype within the same field, year, and host were duplicate collections from the same plant, the data were first clone-corrected to remove these genotypes.
-The data set was hierarchically reformatted to group isolates within each location/host/region/year and clones were removed at that level.
+All data processing and analyses were performed in R version 3.4.0 [@R].
+Of the 16 microsatellite loci, five included compound repeats, making allele scoring difficult. 
+These loci were removed for the reported statistics, but were processed and included in data analyses to ensure the integrity of the results with 11 loci as compared to 16.
+To control for non-independent samples (i.e. those sampled from the same location, region, year, and plant), data were clone-corrected for subsequent analyses.
 The clone-corrected data set included 318 isolates.
 All analyses were performed on this data set unless otherwise mentioned.
+All data and analyses are openly available and reproducible at https://github.com/everhartlab/sclerotinia_366/
 
-## Distance-based redundancy analysis  {-}
+### Virulence Assessment {-}
 
-A distance-based redundancy analysis (dbRDA) [@legendre1999distance] was performed with the function `capscale()` in the *vegan* 2.4-3 package in R (https://CRAN.R-project.org/package=vegan) to delineate relative contribution of phenotypic, spatial, and temporal components in influencing genetic composition of the populations.
+We used analysis of variance (ANOVA) to assess if MCGs differed in severity.
+We chose the top 10 MCGs, representing 57% of the isolates collected.
+Between-group differences were assessed using Tukey's HSD test using the `HSD.test()` function in the package *agricolae* version 1.2.4 [@mendiburu2015agricolae].
+
+### Distance-based redundancy analysis  {-}
+
+A distance-based redundancy analysis (dbRDA) [@legendre1999distance] was performed with the function `capscale()` in the *vegan* package version 2.4-3 [@vegan] to delineate relative contribution of phenotypic, spatial, and temporal components in influencing genetic composition of the populations.
 dbRDA is a method of conducting constrained ordinations on data matrices consisting of non-Euclidean distance measures.
@@
 
-## Shared Multilocus Genotype Community Analysis {-}
+### Shared Multilocus Genotype Community Analysis {-}
 
@@
 
-## Discriminant analysis of principal components {-}
+### Discriminant analysis of principal components {-}
 
@@
 
-## Molecular population statistics and linkage disequilibrium {-}
+### Molecular population statistics and linkage disequilibrium {-}
 
@@
 
-## Results {-}
+# Results {-}
+
+## Genotypic Structure {-}
 
@@
 
+## Virulence {-}
+
 The MCGs with the highest frequency were MCG 5 and MCG 44 representing 73 and 44 isolates, respectively **(Fig. 4)**. <!-- Maybe add aggressiveness information here. -->
@@
 
-## dbRDA {-}
+## Variable Importance {-}
 

### 2017-06-29 doc/RMD/locus-stats.Rmd
add locus-stats

---
title: "Locus Statistics"
output: 
  html_notebook:
    toc: true
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = PROJHOME)
```

# Purpose

This will tabulate statistics per locus for presentation in a table.

```{r}
library("poppr")
load(file.path(PROJHOME, "data", "sclerotinia_16_loci.rda"))
```

```{r locus_table, comment = NA}
makerange <- . %>% as.integer() %>% range() %>% paste(collapse = "-")

ranges <- map_chr(alleles(dat11), makerange)
ranges <- c(ranges, alleles(dat11) %>% unlist(use.names = FALSE) %>% makerange)
locus_table(dat11, information = FALSE) %>% 
  as.data.frame() %>%
  rownames_to_column("Locus") %>%
  mutate(Locus = gsub("\\([FH]\\)", "", Locus)) %>%
  add_column(Range = ranges) %>%
  select(Locus, Range, allele, Hexp, Evenness) %>%
  mutate(allele = round(allele, 2)) %>%
  rename(`$H_{exp}$` = Hexp, `No. alleles` = allele) %>%
  huxtable::as_huxtable(add_colnames = TRUE) %>%
  huxtable::set_number_format(1:12, 3, 0) %>%
  huxtable::set_number_format(everywhere, 4:5, 2) %>%
  huxtable::set_align(everywhere, 3:5, "right") %>%
  huxtable::set_col_width(c(0.06, 0.07, 0.123, 0.09, 0.08)) %>%
  huxtable::print_md(max_width = 47)
```

## Session Information

```{r session_info}
options(width = 100)
devtools::session_info()
```

### 2017-06-29 doc/RMD/RDA-analysis.Rmd
reduce arrows in rda analysis

@@
 ```{r resultplot, fig.width = 5, fig.height = 5}
-set.seed(99)
-plot_dbrda(cap11cc) + 
+set.seed(999)
+plot_dbrda(cap11cc, arrows = 5) + 
   theme_classic(base_size = 16, base_family = "Helvetica") + 
@@
 ggsave(filename = FILE, width = 88, height = 88, units = "mm", scale = 1.5)
-plot_dbrda(cap16cc) +
+plot_dbrda(cap16cc, arrows = 10) +
   theme_classic(base_size = 16, base_family = "Helvetica") + 

### 2017-06-29 doc/RMD/mlg-mcg.Rmd
add mlg-table plot (with insets!

@@
   mutate(MCG = as.character(MCG)) %>%
-  as_data_frame() %>%
+  as_tibble() %>%
   filter(Freq > 0)
@@
 
-```{r barplots, fig.width = 10, fig.height = 4}
-mcgs %>% 
+```{r barplots, fig.width = 5, fig.height = 15}
+mcg_graph <- mcgs %>% 
   gather(type, count, MLGs, Samples, -Evenness) %>%
-  arrange(desc(type), desc(count)) %>%
+  arrange(desc(type), count) %>%
+  rename(Type = type) %>%
   mutate(MCG = forcats::fct_inorder(MCG, ordered = TRUE)) %>%
-  ggplot(aes(x = MCG, y = count, group = type, fill = type, alpha = Evenness)) +
-  geom_col(aes(width = ifelse(type == "MLGs", 0.5, 0.85)), color = "black", position = "identity") +
-  annotate(geom = "text", x = 13, y = 51, label = sprintf("Mean Evenness: %.3f", mean(mcgs$Evenness, na.rm = TRUE))) +
-  # scale_fill_viridis(end = 0.75, discrete = TRUE, direction = -1) +
+  ggplot(aes(x = MCG, y = count, group = Type, fill = Type, alpha = Evenness)) +
+  geom_col(aes(width = ifelse(Type == "MLGs", 0.5, 0.85)), color = "black", position = "identity") +
+  # annotate(geom = "text", x = 18, y =35, 
+  #          label = sprintf("Mean Evenness: %.3f", mean(mcgs$Evenness, na.rm = TRUE))) +
   scale_fill_manual(values = c("black", "white")) +
   scale_y_continuous(expand = c(0, 2)) +
-  theme_minimal() +
-  theme(panel.grid.major.x = element_blank()) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
-  ggtitle("How Evenly are Multilocus Genotypes (MLGs) spread across MCGs?")
+  theme_minimal(base_size = 16, base_family = "Helvetica") +
+  coord_flip() +
+  theme(panel.grid.major.y = element_blank()) +
+  theme(legend.position = c(.745, .85)) +
+  # theme(legend.box = "horizontal") +
+  theme(aspect.ratio = 3) +
+  labs(list(
+    alpha = "MLG Evenness"
+  ))
+
+uneven <- mcgs %>% 
+  filter(MCG == 5) %>% 
+  unnest() %>% 
+  ggplot(aes(x = forcats::fct_inorder(MLG), y = Freq)) + 
+  geom_col(color = "grey62", fill = "grey72", width = 0.5) + 
+  theme_void() +
+  ggtitle("Evenness = 0.48") +
+  theme(aspect.ratio = 1) +
+  theme(title = element_text(hjust = 0.5, 
+                             size = rel(1.25),
+                             family = "Helvetica",
+                             color = "grey30"))
+
+even <- mcgs %>% 
+  filter(MCG == 53) %>% 
+  unnest() %>% 
+  ggplot(aes(x = forcats::fct_inorder(MLG), y = Freq)) + 
+  geom_col(color = "grey20", fill = "grey20", width = 0.75) + 
+  theme_void() +
+  ggtitle("Evenness = 1") +
+  theme(aspect.ratio = 1) +
+  theme(title = element_text(hjust = 0.5, 
+                             size = rel(1.25),
+                             family = "Helvetica",
+                             color = "grey30"))
+
+vp1 <- grid::viewport(width = 0.2, height = 0.2, x = 0.78, y = 0.67, just = "right")
+vp2 <- grid::viewport(width = 0.2, height = 0.2, x = 0.78, y = 0.555, just = "right")
+
+pdf(file.path(PROJHOME, "results/figures/publication/Figure1Z.pdf"), width = 5, height = 12)
+print(mcg_graph)
+print(uneven, vp = vp1)
+print(even, vp = vp2)
+dev.off()
+
+print(mcg_graph)
+print(uneven, vp = vp1)
+print(even, vp = vp2)
+```
+
+```{r mlg_barplots, fig.width = 4, fig.height = 10}
 mlgs %>% 
   gather(type, count, MCGs, Samples, -Evenness) %>%
-  arrange(desc(type), desc(count)) %>%
+  arrange(desc(type), count) %>%
   mutate(MLG = forcats::fct_inorder(MLG, ordered = TRUE)) %>%
@@
   theme(panel.grid.major.x = element_blank()) +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
+  coord_flip()
+  # theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
   ggtitle("How Evenly are MCGs spread across Multilocus Genotypes (MLGs)?")
@@
 
-```{r}
+```{r, eval = FALSE}
 

### 2017-06-29 doc/manuscript/manuscript.Rmd
update publication

@@
 All analyses were performed on this data set unless otherwise mentioned.
-All data and analyses are openly available and reproducible at https://github.com/everhartlab/sclerotinia_366/
+All data and analyses are openly available and reproducible at https://github.com/everhartlab/sclerotinia-366/
 
@@
 
+![
+Barplot showing abundance of each Mycelial Compatibility Group (MCG) in descending order.
+White bars represent sample counts and grey bars represent counts of unique multilocus genotypes (MLG).
+The transparency associated with the grey bars represent the evenness of the distribution of the MLGs within a given MCG as shown in the inset histograms.
+](../../results/figures/publication/Figure1Z.pdf){#barplot width=50%}
+
 <!--  ZNK: I think figures 1-4 are a bit redundant with the tables in the text. It might be better to simply remove them. -->
@@
 The highest number of samples were collected in 2005 (N = 63) and lowest in 2012 (N = 4).
+
+```{r table3, echo = FALSE, warning = FALSE, message = FALSE, results = "asis"}
+suppressPackageStartupMessages(library("tidyverse"))
+read_csv("../../results/tables/mlg_table.csv", 
+         na = "NA", # spaces read in as spaces
+         col_types = cols(.default = col_character())) %>%
+  rename(`$P_{sex}$` = Psex) %>% 
+  knitr::kable(format = "pandoc")
+```
+
+Table: (\#tab:mlg-table) The 5 most abundant Multilocus Genotypes (MLG) with
+the probability of second encounter ($P_{sex}$), Mycelial Compatibility Groups (MCG), and Regions with sample sizes in parentheses.
+
 Pathogen isolates represented 26 commercial cultivars and breeding lines with some representing unknown dry bean hosts **(Fig. 2)**.
@@
 sclerotiorum* population from the cultivar 'Bunsi'.
-The number of unique MLGs recorded from each host is shown in Table 3.
+The five most abundant MLGs with associated MCGs are shown in Table \@ref(tab:mlg-table).
 The majority of MLGs reported were found on multiple host cultivars and only a few unique to a cultivar.
@@
 There were a total of 165 multilocus genotypes in the isolate collection.
-The most abundant MLG represented 27 isolates (Fig. 3 and Table 4) across five regions (NE, MI, WA, CO, and ND).
+The most abundant MLG represented 27 isolates (Fig. \@ref(barplot) and Table 4) across five regions (NE, MI, WA, CO, and ND).
 Twenty isolates of this MLG were identified from ND.
@@
 The MLGs 163, 165 and 15 had 15, 7, and 4 individuals across these populations, respectively.
-The highest number of clones was recorded from ND, NE, and MI, which included 20, 10 and 7 isolates, respectively (Table 4).
+The highest number of clones was recorded from ND, NE, and MI, which included 20, 10 and 7 isolates, respectively **(Table 4)**.
 
@@
 
-The number of observed alleles per locus ranged anywhere from 2 to 10 (Table 2).
+----------------------------------------------
+Locus  Range    No. alleles $H_{exp}$ Evenness 
+------ ------- ------------ --------- --------
+5-2    318-324            4      0.45     0.62
+
+6-2    483-495            3      0.64     0.95
+
+7-2    158-174            7      0.73     0.76
+
+8-3    244-270            7      0.74     0.79
+
+9-2    360-382            9      0.35     0.41
+
+12-2   214-222            5      0.58     0.78
+
+17-3   342-363            7      0.55     0.53
+
+20-3   280-282            2      0.05     0.42
+
+55-4   153-216           10      0.72     0.66
+
+110-4  370-386            5      0.76     0.91
+
+114-4  339-416           10      0.83     0.80
+
+*mean* 153-495         6.27      0.58     0.69
+
+----------------------------------------------
+
+Table: (\#tab:locus-stats) Allelic diversity on full data set at loci used in this study. $H_{exp}$ = Nei's Gene Diversity [@nei1978estimation].
+
+The number of observed alleles per locus ranged anywhere from 2 to 10 (Table \@ref(tab:locus-stats)).
 All loci were informative at cut off level of 2/N where N was the total number of individuals.
@@
 The 165 unique MLGs identified among all 366 isolates are given in Supplementary Table 1.
-Gene diversity ($H_{exp}$) of populations from different regions ranged from 0.259 to 0.596 (Table 5).
+Gene diversity ($H_{exp}$) of populations from different regions ranged from 0.259 to 0.596 **(Table 5)**.
 Evenness of genotypes of all regions was high, with the exception of ND, where a single genotype was represented by 20 isolates.
@@
 Five populations showed evidence of sexual outcrossing: CA in 2004, FR in 2004 and 2005, OR in 2003, and WA in 2008.
-However, after clone-correction these populations had 6-7 isolates/population (Table 7) and therefore, may not be sufficient to make inferences about reproductive mode.
+However, after clone-correction these populations had 6-7 isolates/population **(Table 7)** and therefore, may not be sufficient to make inferences about reproductive mode.
 Only exception was the WA_2008 population, which consisted of 12 isolates.

### 2017-06-30 doc/RMD/mlg-mcg.Rmd
update barplot figure to fix #2

@@
 
-```{r barplots, fig.width = 5, fig.height = 15}
+```{r barplots, fig.width = 5, fig.height = 10}
 mcg_graph <- mcgs %>% 
+  filter(MLGs > 1) %>%
   gather(type, count, MLGs, Samples, -Evenness) %>%
@@
   theme(panel.grid.major.y = element_blank()) +
-  theme(legend.position = c(.745, .85)) +
+  theme(legend.box.just = "left") +
+  theme(legend.box.margin = unit(c(0, 0.55, 11, 0), "lines")) +
+  theme(legend.box.background = element_rect(fill = NA, color = "black")) +
   # theme(legend.box = "horizontal") +
-  theme(aspect.ratio = 3) +
+  theme(aspect.ratio = 2) +
   labs(list(
-    alpha = "MLG Evenness"
+    alpha = "MLG\nEvenness"
   ))
 
-uneven <- mcgs %>% 
-  filter(MCG == 5) %>% 
+even <- mcgs %>% 
+  filter(MCG %in% c(53, 9)) %>% 
+  mutate(MCG = ifelse(MCG == 53, "Evenness = 1.00", "Evenness = 0.55")) %>%
   unnest() %>% 
-  ggplot(aes(x = forcats::fct_inorder(MLG), y = Freq)) + 
-  geom_col(color = "grey62", fill = "grey72", width = 0.5) + 
+  ggplot(aes(x = forcats::fct_inorder(MLG), y = Freq, alpha = Evenness)) + 
+  geom_col(color = "grey30", width = 0.75) + 
   theme_void() +
-  ggtitle("Evenness = 0.48") +
-  theme(aspect.ratio = 1) +
-  theme(title = element_text(hjust = 0.5, 
-                             size = rel(1.25),
-                             family = "Helvetica",
-                             color = "grey30"))
+  theme(strip.text = element_text(family = "Helvetica", color = "grey30", face = "bold")) +
+  theme(aspect.ratio = 0.66) +
+  theme(panel.spacing = unit(0, "line")) +
+  theme(legend.position = "none") +
+  # theme(plot.background = element_rect(fill = NA, colour = "black")) +
+  facet_wrap(~MCG, ncol = 1, scale = "free")
 
-even <- mcgs %>% 
-  filter(MCG == 53) %>% 
-  unnest() %>% 
-  ggplot(aes(x = forcats::fct_inorder(MLG), y = Freq)) + 
-  geom_col(color = "grey20", fill = "grey20", width = 0.75) + 
-  theme_void() +
-  ggtitle("Evenness = 1") +
-  theme(aspect.ratio = 1) +
-  theme(title = element_text(hjust = 0.5, 
-                             size = rel(1.25),
-                             family = "Helvetica",
-                             color = "grey30"))
+vp1 <- grid::viewport(width = 0.25, height = 0.4, x = 0.795, y = 0.64, just = c("center", "top"))
+mcg_graph <- mcg_graph + theme(legend.position = c(.789, .585))
 
-vp1 <- grid::viewport(width = 0.2, height = 0.2, x = 0.78, y = 0.67, just = "right")
-vp2 <- grid::viewport(width = 0.2, height = 0.2, x = 0.78, y = 0.555, just = "right")
-
-pdf(file.path(PROJHOME, "results/figures/publication/Figure1Z.pdf"), width = 5, height = 12)
+pdf(file.path(PROJHOME, "results/figures/publication/Figure1Z.pdf"), width = 5, height = 10)
 print(mcg_graph)
-print(uneven, vp = vp1)
-print(even, vp = vp2)
+print(even, vp = vp1)
 dev.off()
@@
 print(mcg_graph)
-print(uneven, vp = vp1)
-print(even, vp = vp2)
+print(even, vp = vp1)
 ```
@@
 ```{r}
+
 make_mcgmlg_graph <- function(x){

### 2017-06-30 doc/manuscript/manuscript.Rmd
rerun make
begin pruning introduction

@@
 
+<!-- 
+ZNK: The introduction needs a lot of work. Here's an outline of what I think
+     needs to go in here.
+Outline:
+
+ - *S. sclerotiorum* basic biology
+ - What have genetic studies and phenotypic studies found
+ - What are we missing? A detailed analysis of *S. sclerotiorum* from diverse
+   regions that may potentially identify genotypes or MCGs that may be good
+   tools for testing breeding efforts.
+ - Our hypotheses and specific questions.
+ -->
+
 *Sclerotinia sclerotiorum* (Lib.) de Bary is an ascomycete plant pathogen found worldwide.
@@
 
-Sclerotia are capable of surviving more than five years in soil [@steadman1983white] and germinate carpogenically or myceliogenically to infect the plant when environmental conditions are favorable.
+<!-- ZNK: TMI -->
+<!-- Sclerotia are capable of surviving more than five years in soil [@steadman1983white] and germinate carpogenically or myceliogenically to infect the plant when environmental conditions are favorable.
 A carpogenically germinated sclerotia gives rise to apothecia in which ascospores are formed.
@@
 Lesions on stems and branches become necrotic as they age and dead tissues appear bleached.
-Sclerotia are often produced on dead or dying plants once conditions become unfavorable for fungal growth.
+Sclerotia are often produced on dead or dying plants once conditions become unfavorable for fungal growth. -->
 
@@
 
-A population study from Australian canola fields showed evidence of sexual outcrossing of *S. sclerotiorum* isolates [@sexton2004microsatellite].
+<!-- ZNK: This is more discussion-material -->
+<!-- A population study from Australian canola fields showed evidence of sexual outcrossing of *S. sclerotiorum* isolates [@sexton2004microsatellite].
 Another study characterized three populations of *S. sclerotiorum* from lettuce in California (CA) and Ontario (ON), and pea/lentil in Washington (WA) [@malvarez2007new].
@@
 sclerotiorum* [@ekins2011population].
+-->
 
@@
 
-Most of the previous studies on *S. sclerotiorum* had been conducted in the USA, Canada and Australia but not in Europe.
+<!-- Most of the previous studies on *S. sclerotiorum* had been conducted in the USA, Canada and Australia but not in Europe. 
 In those studies, different dry bean cultivars had been rarely used as  hosts to investigate any host-pathogen relationships or specificity.
 No white mold resistant dry bean cultivars exist.
-Canopy architecture and growth pattern of some cultivars enable improved disease avoidance [@kolkman2002agronomic].
-Field trials in some regions of the US have shown G122 is white mold resistant and Beryl is susceptible (pers. comm. with J. Steadman).
+Canopy architecture and growth pattern of some cultivars enable improved disease avoidance [@kolkman2002agronomic]. -->
+To date, no commercial white mold-resistant dry bean cultivars exist. 
+Field trials in some regions of the US have shown the G122 cultivar is potentially resistant while the Beryl cultivar is susceptible (pers. comm. with J. Steadman).
 Understanding genetic relationships and reproduction behavior of *S.
@@
 
-Microsatellite markers (simple sequence repeats or SSR) are codominant, polymorphic and also able to detect multiple alleles per locus [@atallah2004high; @sirjusingh2001characterisation].
+<!-- Microsatellite markers (simple sequence repeats or SSR) are codominant, polymorphic and also able to detect multiple alleles per locus [@atallah2004high; @sirjusingh2001characterisation].
 Reproducibility of microsatellites is higher compared other fingerprinting techniques such as random amplified polymorphic DNA (RAPD) [@atallah2004high].
@@
 sclerotiorum* that is predominantly clonal and occasionally outcrossing, microsatellites have been used successfully to ascertain their ability to reproduce by sexual outcrossing [@atallah2004high; @sexton2004microsatellite] and also to detect differences among clonally propagated individuals.
+-->
 
@@
 Barplot showing abundance of each Mycelial Compatibility Group (MCG) in descending order.
+Singletons (46) have been removed for clarity.
 White bars represent sample counts and grey bars represent counts of unique multilocus genotypes (MLG).
 The transparency associated with the grey bars represent the evenness of the distribution of the MLGs within a given MCG as shown in the inset histograms.
-](../../results/figures/publication/Figure1Z.pdf){#barplot width=50%}
+](../../results/figures/publication/Figure1Z.pdf){#barplot width=70%}
 
@@
 
-Community analysis of the MLGs shared between populations revealed 4 communities with a modularity of 0.17 (Fig.
-\@ref(community-graph)): A coastal community (CA, OR, WA, and NY), a mid-west community (CO, NE, ND, MI), and an international community (MN, France, Australia).
-Analysis with 16 loci resulted in the removal of the NY node because it no longer shared a genotype with OR, but the same overall community structure was present with a modularity of 0.2 (Supplementary Figure S1Z).
-The international community appears to be driven by MLG 4, which is shared between all three populations and has a $P_{sex}$ value of 2.87 x 10^-5^, in contrast to the abundant MLG 25, which has a $P_{sex}$ value of 0.0168.
-
 <!-- community graph figure -->
@@
 Each edge represents a single MLG where its thickness represents the number of populations that share the MLG and the shade represents the value of $P_{sex}$, or the probability of encountering that MLG from two independent meiotic events.
-](../../results/figures/publication/Figure3Z.pdf){#community-graph width=50%}
+](../../results/figures/publication/Figure3Z.pdf){#community-graph width=100%}
 
 <!-- end fig -->
+
+Community analysis of the MLGs shared between populations revealed 4 communities with a modularity of 0.17 (Fig.
+\@ref(community-graph)): A coastal community (CA, OR, WA, and NY), a mid-west community (CO, NE, ND, MI), and an international community (MN, France, Australia).
+Analysis with 16 loci resulted in the removal of the NY node because it no longer shared a genotype with OR, but the same overall community structure was present with a modularity of 0.2 (Supplementary Figure S1Z).
+The international community appears to be driven by MLG 4, which is shared between all three populations and has a $P_{sex}$ value of 2.87 x 10^-5^, in contrast to the abundant MLG 25, which has a $P_{sex}$ value of 0.0168.
 
@@
 Mean separation of 10 most common MCGs showed MCG 44 is different from MCGs 9, 1, 45, 4, 53, and 49 while the virulence levels of MCGs 2, 5, 3 failed to show any difference to MCG 44 or the other MCGs.
+
+The number of observed alleles per locus ranged anywhere from 2 to 10 (Table \@ref(tab:locus-stats)).
+All loci were informative at cut off level of 2/N where N was the total number of individuals.
+The genetic diversity indices of *S. sclerotiorum* populations from different states of the USA, France, Mexico and Australia based on SSR analysis are given in Table 5.
+The 165 unique MLGs identified among all 366 isolates are given in Supplementary Table 1.
+Gene diversity ($H_{exp}$) of populations from different regions ranged from 0.259 to 0.596 **(Table 5)**.
+Evenness of genotypes of all regions was high, with the exception of ND, where a single genotype was represented by 20 isolates.
+There was one private allele in the WA population (204 bp allele of 55-4 locus) and two private alleles in the ND population (two 366 bp alleles in 9-2 locus and two 220 bp alleles in 12-2 locus).
 
@@
 Table: (\#tab:locus-stats) Allelic diversity on full data set at loci used in this study. $H_{exp}$ = Nei's Gene Diversity [@nei1978estimation].
-
-The number of observed alleles per locus ranged anywhere from 2 to 10 (Table \@ref(tab:locus-stats)).
-All loci were informative at cut off level of 2/N where N was the total number of individuals.
-The genetic diversity indices of *S. sclerotiorum* populations from different states of the USA, France, Mexico and Australia based on SSR analysis are given in Table 5.
-The 165 unique MLGs identified among all 366 isolates are given in Supplementary Table 1.
-Gene diversity ($H_{exp}$) of populations from different regions ranged from 0.259 to 0.596 **(Table 5)**.
-Evenness of genotypes of all regions was high, with the exception of ND, where a single genotype was represented by 20 isolates.
-There was one private allele in the WA population (204 bp allele of 55-4 locus) and two private alleles in the ND population (two 366 bp alleles in 9-2 locus and two 220 bp alleles in 12-2 locus).
 

### 2017-06-30 doc/RMD/locus-stats.Rmd
update locus-stats to use syd's preferred *h*

@@
 ```{r}
+library("tidyverse")
 library("poppr")
@@
   mutate(allele = round(allele, 2)) %>%
-  rename(`$H_{exp}$` = Hexp, `No. alleles` = allele) %>%
+  rename(`*h*` = Hexp, `No. alleles` = allele) %>%
   huxtable::as_huxtable(add_colnames = TRUE) %>%
   huxtable::set_number_format(1:12, 3, 0) %>%
-  huxtable::set_number_format(everywhere, 4:5, 2) %>%
-  huxtable::set_align(everywhere, 3:5, "right") %>%
-  huxtable::set_col_width(c(0.06, 0.07, 0.123, 0.09, 0.08)) %>%
+  huxtable::set_number_format(huxtable::everywhere, 4:5, 2) %>%
+  huxtable::set_align(huxtable::everywhere, 3:5, "right") %>%
+  huxtable::set_col_width(c(0.06, 0.07, 0.12, 0.06, 0.09)) %>%
   huxtable::print_md(max_width = 47)

### 2017-06-30 doc/RMD/mlg-mcg.Rmd
update documents to shush

@@
 
-```{r load_packages}
+```{r load_packages, message = FALSE, warning = FALSE}
 library('igraph')
@@
 mcg_graph <- mcgs %>% 
-  filter(MLGs > 1) %>%
+  filter(Samples > 1) %>%
   gather(type, count, MLGs, Samples, -Evenness) %>%

### 2017-06-30 doc/RMD/pop-diff.Rmd
update documents to shush

@@
 
-```{r load_packages}
+```{r load_packages, warning = FALSE, message = FALSE}
 library('tidyverse')

### 2017-06-30 doc/RMD/wmn-differentiation.Rmd
Add wmn-differentiation

---
title: "R Notebook"
output: 
  html_notebook:
    toc: true
editor_options: 
  chunk_output_type: inline
---

```{r setup, include = FALSE}
knitr::opts_knit$set(root.dir = PROJHOME)
```

```{r packages, message = FALSE, warning = FALSE}
library("tidyverse")
library("poppr")
(load("data/sclerotinia_16_loci.rda"))
```

# Purpose

The white mold nursery populations are unique because they are not fungicide
treated and have the same cultivars planted in them year after year.

The question becomes, are white mold nurseries differentiated from each other or
are they more or less homogeneous? We could use AMOVA to test for these with
location and binary source (wmn or non-wmn) as the hierarchy.

# Data Setup

First, we want to clone-correct our data down to the field level so that we
don't accidentally include non-independent samples.

```{r correction}
dat11
dat11cc <- clonecorrect(dat11, ~Region/Source/Host/Year)
dat11cc
```

Now that we've done that, we should make a new variable in the strata that 
separates the white mold nurseries from the others. We'll call this stratum 
"Source Type".

```{r new_stratum}
addStrata(dat11cc) <- strata(dat11cc) %>% 
  mutate(SourceType = forcats::fct_inorder(ifelse(Source == "wmn", "wmn", "other"))) %>%
  select(SourceType)
setPop(dat11cc) <- ~SourceType
dat11cc
```

I can perform AMOVA on the newly defined variable using Bruvo's distance.

```{r AMOVA}
other(dat11cc)$REPLEN
bd <- bruvo.dist(dat11cc, replen = other(dat11cc)$REPLEN)
(ssc_amova <- poppr.amova(dat11cc, ~SourceType, dist = bd, quiet = TRUE))
ssc_amova_test <- randtest(ssc_amova, nrepet = 999)
plot(ssc_amova_test)
ssc_amova_test
```

This result is telling us that there is some subdivision between white mold
nurseries and non-white mold nurseries. Of course, from previous analyses, we
know that Mexico is differentiated from other populations, so what happens if we
account for Region? Here, we are placing region lower in the heirarchy because
we specifically want to test the effect of region on the differentiation between
white mold nurseries within different regions.

```{r AMOVA-Region, cache = TRUE}
(ssc_amova_region <- poppr.amova(dat11cc, ~SourceType/Region, dist = bd, quiet = TRUE))
ssc_amova_region_test <- randtest(ssc_amova_region, nrepet = 999)
plot(ssc_amova_region_test)
ssc_amova_region_test
```

Okay! This shows that when we account for Region after accounting for Source
Type, we find that the differentiation is coming mainly from the Regions. What
happens when we remove Mexico?

```{r AMOVA-nomex, cache = TRUE}
datnomex <- setPop(dat11cc, ~Region) %>% popsub(blacklist = "Mexico")
bdnm     <- bruvo.dist(datnomex, replen = other(datnomex)$REPLEN)
(ssc_amova_nm <- poppr.amova(datnomex, ~SourceType/Region, dist = bdnm, quiet = TRUE))
ssc_amova_nm_test <- randtest(ssc_amova_nm, nrepet = 999)
plot(ssc_amova_nm_test)
ssc_amova_nm_test
```

When we remove the Mexican isolates (which only contained white mold nurseries
and shared no genotypes), we see that indeed, the degree of differentiation
went down.

<details>
<summary>Session Information</summary>

```{r, echo = FALSE}
options(width = 100)
devtools::session_info()
```

</details>

### 2017-06-30 doc/manuscript/manuscript.Rmd
update manuscript after discussion with Syd
Still needs tweaking

@@
 ---
-title: Population structure and phenotypic variation of *Sclerotinia sclerotiorum* from dry bean throughout the United States, and from Mexico, France, and Australia
+title: Population structure and phenotypic variation of *Sclerotinia sclerotiorum* from dry bean in the United States
 preprint: false
@@
     affiliation: 1
-  - name: R. Jhala
+  - name: Rachana Jhala
     email: rjhala2@unl.edu
@@
 Field trials in some regions of the US have shown the G122 cultivar is potentially resistant while the Beryl cultivar is susceptible (pers. comm. with J. Steadman).
+<!-- SEE: Evaluating cultivars, there should be some BIC reports that we can cite; personal communication is kind of eh. -->
 Understanding genetic relationships and reproduction behavior of *S.
@@
 sclerotiorum* isolates collected from dry bean cultivars from 2003 to 2012 from different geographic locations in America, Europe and Australia using microsatellite markers to delineate genetic differences and reproduction behavior.
-The specific aims of our study were:  a) To assess association of phenotypic data (virulence) and environmental data (bean cultivar, region, field, year) on genotypic variation of *S.
+The specific aims of our study were:  a) To assess association of phenotypic data (aggressiveness) and environmental data (bean cultivar, region, field, year) on genotypic variation of *S.
 sclerotiorum* isolates, b) To test the hypothesis that *S.
@@
 Mycelial growth of each isolate was generated by activating sclerotia according to the method described previously [@matheron1989in-vitro].
-These isolates (N = 366) were phenotypically characterized by determining mycelial compatibility group MCG and virulence.
+These isolates (N = 366) were phenotypically characterized by determining mycelial compatibility group MCG and aggressiveness.
 MCG was determined through co-culturing isolates in pairs on PDA plates, where incompatibility of different MCG resulted in formation of a mycelial barrage [@kohn1990mycelial] and isolates belonging to the same MCG did not form a mycelial barrage.
-Virulence of each isolate was assessed using a straw test [@petzoldt1996straw; @teran2006modified] that rates necrotic lesion on a scale of 1-9, with ascending numbers assigned for increasing aggressiveness.
+Aggressiveness of each isolate was assessed using a straw test [@petzoldt1996straw; @teran2006modified] that rates necrotic lesion on a scale of 1-9, with ascending numbers assigned for increasing aggressiveness.
 DNA was purified from mycelium of each isolate using a Qiagen DNeasy Plant Mini Kit (Qiagen Inc., Valencia, CA) according to the manufacturer's instructions.
@@
 
-### Virulence Assessment {-}
+### Aggressiveness Assessment {-}
 
@@
 
-Genotypic evenness [@grunwald2003analysis], which is a measure of distribution of genotype abundances and unbiased gene diversity ($H_{exp}$) [@nei1978estimation] were calculated using *poppr* package in R [@kamvar2014poppr].
+Genotypic evenness [@grunwald2003analysis], which is a measure of distribution of genotype abundances and unbiased gene diversity (*h*) [@nei1978estimation] were calculated using *poppr* package in R [@kamvar2014poppr].
 Microsatellite data were also tested for evidence of linkage disequilibrium, where significant disequilibrium between pairs of loci in a population is attributed to linkage between loci and may be an indication of clonal reproduction if loci are physically unlinked, whereas linkage equilibrium can be an indication of outcrossing populations.
@@
 
+The associations between MCGs and MLGs are complicated \@ref(biggraph)
+
 <!-- community graph figure -->
@@
 
-## Virulence {-}
+## Aggressiveness {-}
 
@@
 ID, MN, NE, NY, and WI did not have any unique/isolated MCGs.
-The virulence of isolates ranged from 1.4 to 7.9 with a mean of 5.01.
-There were six isolates that fell into upper outlier category (virulence > 7.6) and one isolate in the lower outlier range (virulence < 2.3) (box plot results not shown).
-The mean virulence of the 10 most abundant MCGs ranged from 4.6 (MCG 49) to 6.02 (MCG 44) (see Fig 4 for 10 most abundant MCGs).
-ANOVA of these 10 MCGs indicated significant difference for virulence levels at P <0.0001.
-Mean separation of 10 most common MCGs showed MCG 44 is different from MCGs 9, 1, 45, 4, 53, and 49 while the virulence levels of MCGs 2, 5, 3 failed to show any difference to MCG 44 or the other MCGs.
+The aggressiveness of isolates ranged from 1.4 to 7.9 with a mean of 5.01.
+There were six isolates that fell into upper outlier category (aggressiveness > 7.6) and one isolate in the lower outlier range (aggressiveness < 2.3) (box plot results not shown).
+The mean aggressiveness of the 10 most abundant MCGs ranged from 4.6 (MCG 49) to 6.02 (MCG 44) (see Fig 4 for 10 most abundant MCGs).
+ANOVA of these 10 MCGs indicated significant difference for aggressiveness levels at P <0.0001.
+Mean separation of 10 most common MCGs showed MCG 44 is different from MCGs 9, 1, 45, 4, 53, and 49 while the aggressiveness levels of MCGs 2, 5, 3 failed to show any difference to MCG 44 or the other MCGs.
 
@@
 The 165 unique MLGs identified among all 366 isolates are given in Supplementary Table 1.
-Gene diversity ($H_{exp}$) of populations from different regions ranged from 0.259 to 0.596 **(Table 5)**.
+Gene diversity (*h*) of populations from different regions ranged from 0.259 to 0.596 **(Table 5)**.
 Evenness of genotypes of all regions was high, with the exception of ND, where a single genotype was represented by 20 isolates.
@@
 
-----------------------------------------------
-Locus  Range    No. alleles $H_{exp}$ Evenness 
------- ------- ------------ --------- --------
+--------------------------------------------
+Locus  Range    No. alleles    *h*  Evenness 
+------ ------- ------------ ------ ---------
 5-2    318-324            4   0.45      0.62
@@
 
-*mean* 153-495         6.27      0.58     0.69
+mean   153-495         6.27   0.58      0.69
 
-----------------------------------------------
+--------------------------------------------
 
-Table: (\#tab:locus-stats) Allelic diversity on full data set at loci used in this study. $H_{exp}$ = Nei's Gene Diversity [@nei1978estimation].
+Table: (\#tab:locus-stats) Allelic diversity on full data set at loci used in this study. *h* = Nei's Gene Diversity [@nei1978estimation].
 
@@
 ANOVA revealed that the reduced model was significant with an adjusted R^2^ of 0.032 (P = 0.001).
-The marginal effects of state, host, and year were also significant (P = 0.001), but virulence was not significant (P = 0.045).
+The marginal effects of state, host, and year were also significant (P = 0.001), but aggressiveness was not significant (P = 0.045).
 Much of the variation appears to be driven by isolates from Mexico (Fig. \@ref(rda-biplot)) and 2005, but it's worth noting that all Mexican isolates were collected in 2005.
-Variance partitioning of the independent variables indicated virulence to be the least influential factor with 0.1% contributing to explaining the variation of molecular data whereas the combination of variables accounted for 3.3%.
+Variance partitioning of the independent variables indicated aggressiveness to be the least influential factor with 0.1% contributing to explaining the variation of molecular data whereas the combination of variables accounted for 3.3%.
 
@@
 However, further molecular analysis of pathogen isolates from selected hosts did not show any significant genetic difference among populations.
-According to dbRDA, virulence and location of isolates had little or no effect.
+According to dbRDA, aggressiveness and location of isolates had little or no effect.
 Previous studies on *S.
-sclerotiorum* from alfalfa soil [@attanayake2012genetic] and potato [@atallah2004high], and canola [@attanayake2013sclerotinia] plants reported little or no correlation of genotypes on pa25thogen virulence.
+sclerotiorum* from alfalfa soil [@attanayake2012genetic] and potato [@atallah2004high], and canola [@attanayake2013sclerotinia] plants reported little or no correlation of genotypes on pa25thogen aggressiveness.
 This is important information for resistant breeding since genotypes of *S.
@@
 The genetic diversity of regions with >10 isolates had similar values (0.43 - 0.59) except for Mexican isolates.
-$H_{exp}$ of MX was 0.26 and indicated less exposure of sampling sites to outside inocula. <!-- ZNK: Does this mean that there's heavy inbreeding here? -->
+*h* of MX was 0.26 and indicated less exposure of sampling sites to outside inocula. <!-- ZNK: Does this mean that there's heavy inbreeding here? -->
 Test for linkage disequilibrium showed evidence of clonal reproduction for MX isolates (Ia = 2.61, P.Ia = 0.001, $\bar{r}_d$ = 0.37, P.$\bar{r}_d$  = 0.001).
@@
 An understanding of sexual behavior and ability of form new genotypes by *S. sclerotiorum* in dry bean cropping systems is useful for formulating disease management strategies.
+
+# Supplementary Information {-}
+\newpage
+
+<!-- This part is necessary for adding the "S" before supplementary figures.
+Will not work with HTML outputs, though. -->
+
+\setcounter{table}{0}
+\renewcommand{\thetable}{S\arabic{table}}
+\setcounter{figure}{0}
+\renewcommand{\thefigure}{S\arabic{figure}}
+
+![
+Graph showing associations between Mycelial Compatibility Groups dotted outline and Multilocus Genotypes full outline
+](../../results/figures/publication/FigureS2.pdf){#biggraph width=100%}
+
+\newpage

### 2017-06-30 doc/RMD/locus-stats.Rmd
wrap session info in details tag

@@
 
-```{r}
+```{r load_packages, message = FALSE, warning = FALSE}
 library("tidyverse")
@@
 
-## Session Information
+<details>
+<summary>Session Information</summary>
 
-```{r session_info}
+```{r, echo = FALSE}
 options(width = 100)
@@
 ```
+
+</details>

### 2017-06-30 doc/RMD/mlg-mcg.Rmd
remove title from S2 graph

@@
   theme_graph(base_size = 16, base_family = "Helvetica") +
-  theme(legend.position = "bottom", legend.direction = "vertical") +
-  ggtitle("Relation of Multilocus Genotypes and MCGs")
+  theme(legend.position = "bottom", legend.direction = "vertical")
+  # ggtitle("Relation of Multilocus Genotypes and MCGs")
 

### 2017-06-30 doc/RMD/table-1.Rmd
update sorting for table 1

--- 2017-06-29 doc/RMD/table-1.Rmd
@@
 
-```{r isolate_table, results = "asis"}
+```{r load_packages, warnings = FALSE, messages = FALSE}
 library("readr")
@@
 library("huxtable")
+```
 
+```{r isolate_table, comment = NA}
 clmn <- cols(
   .default = col_integer(),
+  Severity = col_double(),
   Region = col_character(),
@@
   mutate(Country = ifelse(nchar(Region) == 2, "USA", Region)) %>% # Defining country column
-  mutate(Region = ifelse(nchar(Region) == 2, Region, "")) %>%     # Adding blanks for international "states"
+  mutate(Region = ifelse(nchar(Region) == 2, Region, "-")) %>%     # Adding blanks for international "states"
   select(Country, Region, Source, Year, Host) %>%                 # Summarizing the number of isolates
@@
             N = sum(N)) %>%                                       #
-  arrange(desc(Country), Region, Source, N) %>%                   # Rearranging rows
+  arrange(desc(Country), Region, desc(N)) %>%                     # Rearranging rows
   group_by(Country, Region) %>%                                   # Adding blanks in repeated Country and Region Names

### 2017-06-30 doc/manuscript/manuscript.Rmd
update table 1 in manuscript

@@
 
-USA       CO     eat               2007, 2010 Yellow                       13
-
-                 gree              2007, 2010 Pinto, Yellow                14
+USA       CO     gree              2007, 2010 Pinto, Yellow                14
 
                  luc               2007, 2010 Yellow                       14
+
+                 eat               2007, 2010 Yellow                       13
 
@@
 
-USA       MI     eln                     2009 unk                           1
+USA       MI     wmn        2003, 2004, 2005, 11A, 37, 38, B07104,         43 
+                                   2008, 2009 Beryl, Bunsi, cornell,          
+                                              G122, ion, PO7863, WM31
+
+                 tuc               2008, 2009 Black, Fuji, Vista            6
 
@@
 
+                 sanc                    2008 Merlot                        3
+
+                 unk               2003, 2009 GH, unk                       2
+
+                 eln                     2009 unk                           1
+
                  mung                    2009 Zorro                         1
-
-                 sanc                    2008 Merlot                        3
 
@@
 
-                 tuc               2008, 2009 Black, Fuji, Vista            6
-
-                 unk               2003, 2009 GH, unk                       2
-
-                 wmn        2003, 2004, 2005, 11A, 37, 38, B07104,         43 
-                                   2008, 2009 Beryl, Bunsi, cornell,          
-                                              G122, Orion, PO7863, WM31
-
 USA       MN     wmn               2003, 2004 Beryl, Bunsi, G122           11
 
-USA       ND     gfc               2007, 2010 unk                           9
+USA       ND     rrv                     2007 unk                          21
+
+                 gfc               2007, 2010 unk                           9
+
+                 tlc                     2010 unk                           9
+
+                 pmc                     2010 unk                           7
+
+                 wmn                     2005 Beryl, Bunsi, G122            7
+
+                 wlc                     2010 unk                           4
+
+                 stc                     2010 unk                           2
 
@@
 
-                 pmc                     2010 unk                           7
+USA       NE     wmn        2004, 2005, 2008, Beryl, Bunsi, G122,          27 
+                                         2010 PO7683, unk
 
-                 rrv                     2007 unk                          21
-
-                 stc                     2010 unk                           2
-
-                 tlc                     2010 unk                           9
-
-                 wlc                     2010 unk                           4
-
-                 wmn                     2005 Beryl, Bunsi, G122            7
-
-USA       NE     mtc               2007, 2010 Beryl, Emerson, Orion        12
+                 mtc               2007, 2010 Beryl, Emerson, Orion        12
 
@@
 
-                 wmn        2004, 2005, 2008, Beryl, Bunsi, G122,          27 
-                                         2010 PO7683, unk
-
 USA       NY     unk                     2003 GH                            1
 
-USA       OR     corv                    2003 G122                          1
+USA       OR     wmn               2003, 2004 Beryl, Bunsi, G122           15
+
+                 corv                    2003 G122                          1
 
@@
 
-                 wmn               2003, 2004 Beryl, Bunsi, G122           15
+USA       WA     wmn        2003, 2004, 2005, 11A, 37, 38, Beryl, Bunsi,   36 
+                                         2008  cornell, G122, Orion,          
+                                              PO7104, PO7863, WM31
 
-USA       WA     prsr                    2007 Merlot, Pinto, redkid        22
+                 prsr                    2007 Merlot, Pinto, redkid        22
 
                  unk                     2003 GH                            1
-
-                 wmn        2003, 2004, 2005, 11A, 37, 38, Beryl, Bunsi,   36 
-                                         2008 cornell, G122, Orion,           
-                                              PO7104, PO7863, WM31
 
@@
 
-Mexico           wmn                     2005 Beryl, Bunsi, G122           18
+Mexico    -      wmn                     2005 Beryl, Bunsi, G122           18
 
-France           flds                    2012 unk                           4
+France    -      wmn               2004, 2005 Beryl, Bunsi, G122           18
 
-                 wmn               2004, 2005 Beryl, Bunsi, G122           18
+                 flds                    2012 unk                           4
 
-Australia        vic                     2004 Beryl                         2
+Australia -      wmn                     2004 Beryl, Bunsi, G122            4
 
-                 wmn                     2004 Beryl, Bunsi, G122            4
+                 vic                     2004 Beryl                         2
 

### 2017-06-30 doc/RMD/MCG-virulence.Rmd
wrap session info in details tag
Severity -> Aggressiveness

@@
 can get this from the cleaned data set that we saved in the
-`data-comparison.Rmd` file.
+`data-comparison.Rmd` file. Note, we will be labelling the plot with 
+"Aggressiveness" as that is the preferred term.
 
@@
 
-```{r vis, fig.width = 5, fig.height = 4}
+```{r vis, fig.width = 5, fig.height = 3}
 set.seed(2017-06-29)
@@
   labs(list(
-    title = "Severity for the top 10 MCGs",
-    caption = "MCG = Mycelial Compatibility Group"
+    # title = "Aggressiveness for the top 10 MCGs",
+    y = "Aggressiveness",
+    x = "Mycelial Compatibility Group"
     ))
@@
   summarize(N = n(), 
-            `Min Severity` = min(Severity),
-            `Max Severity` = max(Severity),
-            `Average Severity` = mean(Severity)
+            `Min Aggressiveness` = min(Severity),
+            `Max Aggressiveness` = max(Severity),
+            `Average Aggressiveness` = mean(Severity)
             ) %>%

### 2017-06-30 doc/RMD/mlg-mcg.Rmd
fix dimensions of graph

@@
   theme_graph(base_size = 16, base_family = "Helvetica") +
-  theme(legend.position = "bottom", legend.direction = "vertical")
+  theme(legend.position = "bottom", legend.direction = "vertical") +
+  theme(legend.box.spacing = unit(-1, "lines")) +
+  theme(plot.margin = unit(c(0, 0, 0, 0), "lines"))
   # ggtitle("Relation of Multilocus Genotypes and MCGs")
@@
 mcg_mlg_graph
-ggsave(file.path(PROJHOME, "results/figures/publication/FigureS2.pdf"), 
-       width = 88*3, height = 88*3, units = "mm")
+ggsave(plot = mcg_mlg_graph, filename = file.path(PROJHOME, "results/figures/publication/FigureS2.pdf"), 
+       width = 88*3, height = 88*3.5, units = "mm")
 ```

### 2017-06-30 doc/manuscript/manuscript.Rmd
Move graph to main section; demote bargraph

@@
 The transparency associated with the grey bars represent the evenness of the distribution of the MLGs within a given MCG as shown in the inset histograms.
-](../../results/figures/publication/Figure1Z.pdf){#barplot width=70%}
+](../../results/figures/publication/Figure1Z.pdf){#barplot width=50%}
 
@@
 
-The associations between MCGs and MLGs are complicated \@ref(biggraph)
+The associations between MCGs and MLGs are complicated (Fig. \@ref(biggraph))
+
+![
+Graph showing complex associations between Mycelial Compatibility Groups (dotted nodes) and Multilocus Genotypes (full nodes). 
+Node size reflect the number of samples represented by each node (circle).
+Edges (arrows) point from Mycelial Compatibility Group to Multilocus Genotype where the weight (thickness) of the edge represents the number of samples shared.
+Node color represents the community assignment based on the walktrap algorithm with a maximum of four steps [@pons2006computing].
+](../../results/figures/publication/FigureS2.pdf){#biggraph width=100%}
 
@@
 \renewcommand{\thefigure}{S\arabic{figure}}
-
-![
-Graph showing associations between Mycelial Compatibility Groups dotted outline and Multilocus Genotypes full outline
-](../../results/figures/publication/FigureS2.pdf){#biggraph width=100%}
-
-\newpage

### 2017-07-03 doc/RMD/MCG-virulence.Rmd
update mcg-virulence figure to have common theme

@@
 
-```{r vis, fig.width = 5, fig.height = 3}
+```{r vis, fig.width = 7, fig.height = 4}
 set.seed(2017-06-29)
@@
   geom_point(position = position_jitter(width = 0.25), alpha = 0.5) +
+  scale_y_continuous(limits = c(0, NA)) +
+  # scale_x_discrete(position = "top") +
   theme_bw(base_size = 16, base_family = "Helvetica") +
   theme(aspect.ratio = 1/2) +
+  theme(axis.text = element_text(color = "black")) +
+  # theme(axis.ticks.x = element_blank()) +
+  theme(panel.grid.major = element_line(colour = "grey20")) +
+  theme(panel.grid.minor = element_line(colour = "grey50")) +
+  theme(panel.grid.major.x = element_blank()) +
+  theme(panel.border = element_blank()) +
   labs(list(

### 2017-07-03 doc/RMD/wmn-differentiation.Rmd
add amova table and density plot

@@
 dat11cc
+
+make_amova_table <- function(am, amt, samples = "Region"){
+  tot <- nrow(am$results)
+  res <- data.frame(list(am$results[-tot, c("Df", "Sum Sq")], 
+                         Percent = am$componentsofcovariance[-tot, 2],
+                         Pval    = rev(amt$pvalue), 
+                         Phi     = am$statphi$Phi[-tot]))
+  res <- as.matrix(res)
+  colnames(res) <- c("d.f.", "Sum of Squares", "Percent variation", "P", 
+                     "Phi statistic")
+  names(dimnames(res)) <- c("levels", "statistic")
+  rownames(res) <- gsub("samples", samples, rownames(res))
+  return(res)
+}
+
+make_amova_printable <- function(amtab, amtabcc){
+  am_array <- array(dim      = c(dim(amtab), 2),
+                    dimnames = c(dimnames(amtab), 
+                                 list(c("full", "clone-corrected"))))
+  am_array[, , 1] <- amtab
+  am_array[, , 2] <- amtabcc
+  tabfun <- function(x){
+    x <- paste0(paste0(signif(x, 3), collapse = " ("), ")")
+    return(x)
+  }
+  res <- apply(am_array, c(1, 2), tabfun)
+  return(res)
+}
 ```
@@
 ```{r AMOVA-Region, cache = TRUE}
-(ssc_amova_region <- poppr.amova(dat11cc, ~SourceType/Region, dist = bd, quiet = TRUE))
+ssc_amova_region <- poppr.amova(dat11cc, ~SourceType/Region, dist = bd, quiet = TRUE)
 ssc_amova_region_test <- randtest(ssc_amova_region, nrepet = 999)
 plot(ssc_amova_region_test)
-ssc_amova_region_test
+(ssc_amova_table <- make_amova_table(ssc_amova_region, ssc_amova_region_test))
 ```
@@
 bdnm     <- bruvo.dist(datnomex, replen = other(datnomex)$REPLEN)
-(ssc_amova_nm <- poppr.amova(datnomex, ~SourceType/Region, dist = bdnm, quiet = TRUE))
+ssc_amova_nm <- poppr.amova(datnomex, ~SourceType/Region, dist = bdnm, quiet = TRUE)
 ssc_amova_nm_test <- randtest(ssc_amova_nm, nrepet = 999)
 plot(ssc_amova_nm_test)
-ssc_amova_nm_test
+(ssc_amova_nm_table <- make_amova_table(ssc_amova_nm, ssc_amova_nm_test))
+```
+
+```{r AMOVA_table, comment = NA}
+make_amova_printable(ssc_amova_table, ssc_amova_nm_table) %>%  
+  as_tibble() %>%
+  add_column(Hierarchy = c("Between Source", "Between Region within Source", "Within Region"), .before = 1) %>%
+  rename(ps = `Phi statistic`) %>%
+  mutate(ps = gsub("0\\.00(\\d{1})(\\d{2})", "\\1.\\2e^-3^", ps)) %>%
+  rename(`$\\Phi statistic$` = ps) %>%
+  rename(`% variation` = `Percent variation`) %>%
+  rename(S.S. = `Sum of Squares`) %>%
+  select(-P) %>%
+  huxtable::as_huxtable(add_colnames = TRUE) %>% 
+  huxtable::set_col_width(c(1.3, 0.7, 0.9, 1, 1.3)) %>% 
+  huxtable::set_align(huxtable::everywhere, 2:5, "center") %>% 
+  huxtable::print_md(max_width = 95)
 ```
@@
 went down.
+
+## Visualizing the partitions
+
+We can visualize the partitions if we create distributions showing the genetic
+distance.
+
+```{r distrib, fig.width = 6, fig.height = 3}
+wmn_inds <- (setPop(dat11cc, ~SourceType) %>% pop()) == "wmn"
+
+# Function to set the upper triangle to NA to avoid over-representation.
+set_upper_tri_NA <- function(d){d[upper.tri(d)] <- NA; diag(d) <- NA; d}
+
+# dist data converted to matrix fed into here. Removes NAs from previous function
+tidy_dist <- . %>%
+  as.data.frame() %>% 
+  rownames_to_column("from") %>% 
+  gather(to, distance, -from) %>% 
+  filter(!is.na(distance))
+wmn_distance   <- as.matrix(bd)[wmn_inds, wmn_inds] %>% set_upper_tri_NA() %>% tidy_dist
+nwmn_distance  <- as.matrix(bd)[!wmn_inds, !wmn_inds] %>% set_upper_tri_NA() %>% tidy_dist
+inter_distance <- as.matrix(bd)[wmn_inds, !wmn_inds] %>% tidy_dist
+
+dists <- bind_rows(`White Mold Nurseries` = wmn_distance, 
+                   `Between` = inter_distance,
+                   `Other Sources` = nwmn_distance, 
+                   .id = "Comparison") %>%
+  mutate(Comparison = forcats::fct_inorder(Comparison))
+dists
+ggplot(dists, aes(y = distance, x = Comparison, fill = Comparison)) +
+  geom_violin() +
+  geom_boxplot(width = 0.25) +
+  scale_fill_manual(values =  c(`Other Sources` = "grey35", `White Mold Nurseries` = "grey95", Between = "grey65")) +
+  scale_color_manual(values = c(`Other Sources` = "grey35", `White Mold Nurseries` = "grey95", Between = "grey65")) +
+  theme_bw(base_size = 16, base_family = "Helvetica") +
+  theme(aspect.ratio = 1/2) +
+  theme(legend.position = "none") +
+  theme(axis.title.x = element_blank()) +
+  theme(axis.text = element_text(color = "black")) +
+  theme(panel.grid.major = element_line(colour = "grey20")) +
+  theme(panel.grid.minor = element_line(colour = "grey50")) +
+  theme(panel.grid.major.x = element_blank()) +
+  theme(panel.border = element_blank()) +
+  ylab("Bruvo's Distance")
+```
+
+```{r dist_comparison, results = "asis"}
+summary_dists <- dists %>% 
+  group_by(Comparison) %>% 
+  summarize(`Average Distance` = mean(distance),
+            `Standard Deviation` = sd(distance)) 
+knitr::kable(summary_dists, digits = 3)
+```
+
+On first look, it appears that there's not much difference. Looking at the
+table, we can see that the average distance within White Mold Nurseries is 
+`r round(summary_dists[["Average Distance"]][1], 3)`, whereas the average distance within Other
+Sources is `r round(summary_dists[["Average Distance"]][3], 3)`. This is well within the standard
+deviation for both distributions, but we also recognize that a single step of
+Bruvo's distance for 11 loci in a haploid organism is `r round(0.5/11, 3)`.
+
+# Summary
 

### 2017-07-03 doc/manuscript/manuscript.Rmd
add AMOVA table to manuscript

@@
 
+-------------------------------------------------------------------------------
+Hierarchy               d.f.        S.S.       % variation    $\Phi statistic$  
+-------------------- ---------- ------------- -------------- ------------------
+Between Source         1 (1)    0.897 (0.753) -0.158 (0.116)   0.136 (0.0965)
+
+Between Region        20 (19)     12 (8.91)    13.8 (9.53)     **0.138 (0.0955)**   
+within Source
+
+Within Region        296 (282)   56.7 (55.1)   86.4 (90.3)   **-1.58e^-3^ (1.16e^-3^)**
+
+-------------------------------------------------------------------------------
+
+Table: (\#tab:amova) Analysis of Molecular Variance showing little differentiation between White Mold Nurseries and other sources.
+The hierarchy was constructed as Source/Region where source is defined as belonging to a white mold nursery or not.
+Bold $\Phi$ values indicate significant values.
+Numbers in parentheses represent AMOVA run without Mexican isolates included.
+S.S.
+= Sum of Squares, d.f.
+= degrees of freedom.
+
 Since `find.clusters()` function of *adegenet* did not result in a clear-cut number of clusters for 318 isolates, we arrived at six (6) as the best number of populations (Fig.

### 2017-07-03 doc/RMD/wmn-differentiation.Rmd
Add questions related to host

@@
 
-# Summary
+# Host differentiation
+
+There are three main hosts shared within the white mold nurseries:
+
+| Cultivar | Susceptibility |
+| ----- | ---- |
+| Beryl | Susceptible |
+| Bunsi | ? |
+| G122  | Some Resistance |
+
+These hosts are grown in the nurseries at different times and it is of interest
+if cultivar can have an effect on population structure. Given the results of
+Aldrich-Wolfe 2015, I would suspect that cultivar has no effect, but it's
+important to verify this. We are going to use AMOVA and DAPC for this.
+
+```{r host_AMOVA}
+bebug <- dat11cc %>% 
+  setPop(~Source) %>% 
+  popsub("wmn") %>% 
+  setPop(~Host) %>% 
+  popsub(c("Beryl", "Bunsi", "G122"))
+bebugd <- bruvo.dist(bebug, replen = other(bebug)$REPLEN)
+(ssc_host_amova <- poppr.amova(bebug, ~Host, dist = bebugd, quiet = TRUE))
+(ssc_host_amova <- poppr.amova(bebug, ~Host/Region, dist = bebugd, quiet = TRUE))
+```
+
+The AMOVA makes it clear that there's no differentiation between hosts. Let's
+see what DAPC shows us.
+
+```{r dapc, fig.width = 5, fig.height = 5.5}
+bebug_dapc <- dapc(bebug, n.pca = 20, n.da = 2)
+scatter(bebug_dapc, clabel = 0,
+        legend = TRUE,
+        scree.pca = TRUE, 
+        scree.da = FALSE,
+        col = viridis::plasma(3), 
+        bg.inset = "grey90",
+        bg = "grey90")
+
+ggcompoplot::ggcompoplot(bebug_dapc, bebug, pal = viridis::plasma(3)) +
+  theme(legend.position = "top")
+```
 

### 2017-07-03 doc/RMD/pop-diff.Rmd
add poppr table to manuscript

@@
 
+# Genotypic Diversity
+
+```{r geno-div, cache = TRUE, fig.width = 10, fig.height = 10, comment = NA}
+ptab <- dat11cc %>% setPop(~Region) %>% poppr(sample = 999, total = FALSE)
+```
+
+```{r geno-div2, fig.width = 10, fig.height = 10, comment = NA}
+ptab %>%
+  arrange(desc(N), desc(MLG)) %>%
+  mutate(Ia = case_when(
+    !is.finite(.$p.Ia) ~ "-",
+    .$p.Ia == 0.001 ~ paste0(round(.$Ia, 2), "*"),
+    .$p.Ia <= 0.01 ~  paste0(round(.$Ia, 2), "-"),
+    .$p.Ia <= 0.05 ~  paste0(round(.$Ia, 2), "~"),
+    TRUE ~ as.character(round(.$Ia, 2))
+  )) %>%
+  mutate(rbarD = case_when(
+    !is.finite(.$p.rD) ~ "-",
+    .$p.rD == 0.001 ~ paste0(round(.$rbarD, 2), "*"),
+    .$p.rD <= 0.01 ~  paste0(round(.$rbarD, 2), "-"),
+    .$p.rD <= 0.05 ~  paste0(round(.$rbarD, 2), "~"),
+    TRUE ~ as.character(round(.$rbarD, 2))
+  )) %>%
+  mutate_if(is.numeric, round, digits = 2) %>%
+  mutate(N = paste(N, MLG, sep = " (") %>% paste0(")") %>% formatC()) %>%
+  mutate(eMLG = paste(eMLG, SE, sep = " (") %>%  paste0(")") %>% formatC()) %>%
+  rename(`*h*` = Hexp) %>%
+  rename(`$E_5$` = E.5) %>%
+  rename(`$I_A$` = Ia) %>%
+  rename(`$\\lambda$` = lambda) %>%
+  rename(`$\\bar{r}_d$` = rbarD) %>%
+  select(-File, -MLG, -SE, -p.rD, -p.Ia) %>% 
+  huxtable::as_huxtable(add_colnames = TRUE) %>%
+  huxtable::set_number_format(huxtable::everywhere, 4:5, 1) %>%
+  huxtable::set_align(huxtable::everywhere, -1, "right") %>%
+  huxtable::set_col_width(c(1.1, 0.9, 1.3, 0.5, 0.5, 1.2, 0.7, 0.6, 0.6, 0.8)) %>%
+  huxtable::print_md(max_width = 80)
+
+ptab %>% 
+  arrange(desc(N), desc(MLG)) %>% 
+  write_csv(file.path(PROJHOME, "results/tables/genotype_diversity_table_region.csv"), col_names = TRUE)
+
+mtab <- mlg.table(dat11, ~Region, background = TRUE)
+```
+
 # AMOVA
@@
 region.dapc <- dapc(dat11, strata(dat11)$Region, n.pca = 25, n.da = 14)
-scatter(region.dapc)
+scatter(region.dapc, scree.pca = TRUE,
+        bg = "grey80",
+        col = funky(nlevels(strata(dat11)$Region)), 
+        posi.pca = "topright"
+        )
 ggcompoplot(region.dapc, setPop(dat11, ~Region), pal = funky, cols = 2)

### 2017-07-03 doc/manuscript/manuscript.Rmd
add poppr table to manuscript

@@
 \renewcommand{\thefigure}{S\arabic{figure}}
+
+----------------------------------------------------------------------------
+Pop             N        eMLG    H    G  $\lambda$  $E_5$   *h* $I_A$ $\bar{ 
+                                                                       r}_d$ 
+--------- ------- ----------- ---- ---- ---------- ------ ----- ----- ------
+WA        58 (56) 9.95 (0.23)  4.0 54.3       0.98   0.98  0.60 0.65*  0.07*
+
+MI        58 (43)  9.3 (0.79)  3.6 29.0       0.97   0.78  0.54 1.25*  0.14*
+
+ND        41 (35) 9.44 (0.73)  3.5 25.9       0.96   0.82  0.54    1*   0.1*
+
+NE        37 (28) 8.93 (0.94)  3.2 17.8       0.94   0.75  0.55 2.22*  0.25*
+
+CO        34 (28) 9.46 (0.67)  3.3 24.1       0.96   0.92  0.56 2.33*  0.27*
+
+France    21 (14)  8.5 (0.85)  2.6 12.6       0.92   0.95  0.48 0.94*  0.11*
+
+CA        18 (15) 9.12 (0.72)  2.7 13.5       0.93   0.94  0.51 0.28~  0.03~
+
+OR        17 (13) 8.52 (0.85)  2.5 10.7       0.91   0.89  0.47 0.92*   0.1*
+
+Mexico     15 (9)  7.1 (0.85)  2.1  7.3       0.86   0.89  0.28 2.57*  0.37*
+
+MN          9 (7)       7 (0)  1.9  6.2       0.84   0.93  0.47 1.34*  0.19*
+
+Australia   6 (6)       6 (0)  1.8  6.0       0.83   1.00  0.48  0.7~  0.12~
+
+WI          2 (2)       2 (0)  0.7  2.0       0.50   1.00  0.27     -      -
+
+NY          1 (1)       1 (0)  0.0  1.0       0.00    NaN   NaN     -      -
+
+ID          1 (1)       1 (0)  0.0  1.0       0.00    NaN   NaN     -      -
+
+----------------------------------------------------------------------------
+
+Table: (\#tab:iatab) Genotypic diversity and Linkage Disequilibrium summary for geographic populations arranged by abundance after clone-correction by a hierarchy of Region/Source/Host/Year. 
+Pop = Population, N = number of individuals (number of MLG in parentheses), 
+eMLG = expected number of MLGs based on rarefaction at 10 individuals (standard error in parentheses), 
+H = Shannon-Weiner Index, G = Stoddardt and Taylor's Index, 
+$\lambda$ = Simpson's Index, 
+*h* = Nei's 1978 gene diversity, 
+$I_A$ = the index of association, 
+$\bar{r}_d$ = standardized index of association. 
+P-value symbols: \* = 0.001, ~ = $\leq 0.05$.

### 2017-07-05 doc/RMD/pop-diff.Rmd
Add figure of DAPC Posterior Probabilities

@@
 knitr::opts_knit$set(root.dir = PROJHOME)
+knitr::opts_chunk$set(fig.align = "center")
 ```
@@
 ```{r dapc-region, fig.width = 7, fig.height = 7}
-region.dapc <- dapc(dat11, strata(dat11)$Region, n.pca = 25, n.da = 14)
+region.dapc <- dapc(dat11cc, strata(dat11cc)$Region, n.pca = 25, n.da = 14)
 scatter(region.dapc, scree.pca = TRUE,
         bg = "grey80",
-        col = funky(nlevels(strata(dat11)$Region)), 
+        col = funky(nlevels(strata(dat11cc)$Region)), 
         posi.pca = "topright"
         )
-ggcompoplot(region.dapc, setPop(dat11, ~Region), pal = funky, cols = 2)
+ggcompoplot(region.dapc, setPop(dat11cc, ~Region), pal = funky, cols = 2)
+```
+
+Another way to visualize this is to look at the probability of assignment by
+source pouplation.
+
+```{r dapc-region-prob, fig.width = 5.5, fig.height = 9}
+region.dapc$posterior %>% 
+  as.data.frame() %>% 
+  as_tibble() %>% 
+  add_column(Individual = indNames(dat11cc)) %>% 
+  add_column(`Original Population` = as.character(strata(dat11cc)$Region)) %>% 
+  group_by(`Original Population`) %>%
+  mutate(n = n()) %>%
+  gather(Population, Posterior, -Individual, -`Original Population`, -n) %>% 
+  group_by(Population, `Original Population`, n) %>% 
+  summarize(`Mean Assignment` = mean(Posterior)) %>% 
+  arrange(`Mean Assignment`) %>% 
+  mutate(Source = case_when(
+    Population == `Original Population` ~ "Original Population",
+    TRUE ~ "Other"
+    )) %>%
+  ungroup() %>%
+  arrange(desc(n)) %>%
+  mutate(`Original Population` = forcats::fct_inorder(`Original Population`)) %>%
+  mutate(Population = factor(Population, levels(`Original Population`))) %>%
+  mutate(`Original Population` = paste0(`Original Population`, " (", n, ")")) %>%
+  mutate(`Original Population` = forcats::fct_inorder(`Original Population`)) %>%
+  filter(n >= 10) %>%
+  ggplot(aes(x = Population, y = `Mean Assignment`)) +
+  geom_segment(aes(xend = Population, yend = 0)) +
+  geom_point(aes(fill = Source), pch = 21) + 
+  facet_wrap(~`Original Population`, ncol = 2) +
+  scale_fill_grey() +
+  scale_y_continuous(limits = c(0, 1)) +
+  theme_bw(base_size = 16, base_family = "Helvetica") +
+  theme(legend.position = "top") +
+  theme(aspect.ratio = 1/2) +
+  theme(axis.text = element_text(color = "black")) +
+  theme(axis.ticks.y = element_blank()) +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
+  theme(panel.grid.major.x = element_line(linetype = 0, color = "grey50")) +
+  theme(panel.grid.major = element_line(colour = "grey20")) +
+  theme(panel.grid.minor = element_line(linetype = 3, colour = "grey50")) +
+  theme(panel.spacing.y = unit(0, "line")) +
+  theme(strip.background = element_rect(color = NA, fill = "grey90")) +
+  theme(strip.text = element_text(face = "bold", hjust = 0.05)) +
+  theme(panel.border = element_blank()) +
+  xlab("Population Assignment") +
+  ylab("Mean Probability")
 ```

### 2017-07-05 doc/RMD/pop-diff.Rmd
add cross-validation; make dapc black and white

@@
 
-```{r dapc-region, fig.width = 7, fig.height = 7}
-region.dapc <- dapc(dat11cc, strata(dat11cc)$Region, n.pca = 25, n.da = 14)
-scatter(region.dapc, scree.pca = TRUE,
-        bg = "grey80",
-        col = funky(nlevels(strata(dat11cc)$Region)), 
+```{r dapc-region-calc, fig.width = 8, fig.height = 7, cache = TRUE}
+set.seed(2017-07-05)
+region.xval <- xvalDapc(tab(dat11cc), strata(dat11cc)$Region, n.pca = 15:25, 
+                        n.rep = 100, parallel = "multicore", ncpus = 4)
+region.dapc <- region.xval$DAPC
+```
+
+```{r dapc-region, fig.width = 9, fig.height = 8, cache = TRUE}
+scatter(region.dapc, 
+        scree.pca = TRUE,
+        bg = "grey95",
+        col = "black",
+        # col = rev(funky(nlevels(strata(dat11cc)$Region))), 
+        pch = seq(nlevels(strata(dat11cc)$Region)),
+        legend = "true",
+        posi.leg = "topleft",
         posi.pca = "topright"
         )
-ggcompoplot(region.dapc, setPop(dat11cc, ~Region), pal = funky, cols = 2)
+ggcompoplot(region.dapc, setPop(dat11cc, ~Region), pal = rev(funky(nlevels(strata(dat11cc)$Region))), cols = 2)
 ```

### 2017-07-06 doc/RMD/wmn-differentiation.Rmd
change DAPC to host/region

@@
 ```{r dapc, fig.width = 5, fig.height = 5.5}
+setPop(bebug) <- ~Host/Region
 bebug_dapc <- dapc(bebug, n.pca = 20, n.da = 2)
-scatter(bebug_dapc, clabel = 0,
+scatter(bebug_dapc, #clabel = 0,
         legend = TRUE,
@@
         scree.da = FALSE,
-        col = viridis::plasma(3), 
+        # col = viridis::plasma(3), 
         bg.inset = "grey90",
@@
 
-ggcompoplot::ggcompoplot(bebug_dapc, bebug, pal = viridis::plasma(3)) +
+ggcompoplot::ggcompoplot(bebug_dapc, bebug) + #, pal = viridis::plasma(3)) +
   theme(legend.position = "top")

### 2017-07-06 doc/RMD/pop-diff.Rmd
made compound dapc figure

--- 2017-07-03 doc/RMD/pop-diff.Rmd
@@
 
-```{r amova-no-cc}
+```{r amova-no-cc, cache = TRUE}
 (res   <- poppr.amova(dat11cc, ~Region/Host/Year, dist = dat11cc.bruvo))
@@
 
-```{r dapc-region, fig.width = 7, fig.height = 7}
-region.dapc <- dapc(dat11, strata(dat11)$Region, n.pca = 25, n.da = 14)
-scatter(region.dapc, scree.pca = TRUE,
-        bg = "grey80",
-        col = funky(nlevels(strata(dat11)$Region)), 
+```{r dapc-region-calc, fig.width = 8, fig.height = 7, cache = TRUE}
+set.seed(2017-07-05)
+region.xval <- xvalDapc(tab(dat11cc), strata(dat11cc)$Region, n.pca = 15:25, 
+                        n.rep = 100, parallel = "multicore", ncpus = 4)
+region.dapc <- region.xval$DAPC
+```
+
+```{r dapc-region, fig.width = 9, fig.height = 8, cache = TRUE}
+scatter(region.dapc, 
+        scree.pca = TRUE,
+        bg = "grey95",
+        col = "black",
+        # col = rev(funky(nlevels(strata(dat11cc)$Region))), 
+        pch = seq(nlevels(strata(dat11cc)$Region)),
+        legend = "true",
+        posi.leg = "topleft",
         posi.pca = "topright"
         )
-ggcompoplot(region.dapc, setPop(dat11, ~Region), pal = funky, cols = 2)
+ggcompoplot(region.dapc, setPop(dat11cc, ~Region), pal = rev(funky(nlevels(strata(dat11cc)$Region))), cols = 2)
+```
+
+```{r fig.width = 10, fig.height = 6}
+LDS <- bind_cols(Population = region.dapc$grp, as.data.frame(region.dapc$ind.coord)) %>%
+  as_tibble()
+LDS_pop <- LDS %>% 
+  group_by(Population) %>% 
+  summarize_all(mean) %>%
+  rename_all(function(x) gsub("LD", "mean", x))
+LDS <- full_join(LDS, LDS_pop)
+LDS_PLOT <- ggplot(LDS, aes(x = LD1, y = LD2, color = Population)) + 
+  geom_point(aes(fill = Population), alpha = 0.5, pch = 21, color = "black") +
+  geom_segment(aes(x = mean1, y = mean2, xend = LD1, yend = LD2), alpha = 0.5) +
+  stat_ellipse(type = "norm", level = 0.66, alpha = 0.75) + 
+  ggrepel::geom_label_repel(aes(x = mean1, y = mean2, label = Population), 
+                            data = LDS_pop, show.legend = FALSE, color = "black") +
+  theme_bw() +  
+  theme(aspect.ratio = 1/1.618) +
+  theme(legend.position = "bottom") +
+  theme(axis.text = element_blank()) + 
+  theme(axis.title = element_blank()) + 
+  theme(axis.ticks = element_blank()) + 
+  viridis::scale_color_viridis(discrete = TRUE) +
+  viridis::scale_fill_viridis(discrete = TRUE) +
+  scale_y_continuous(breaks = 0) + 
+  scale_x_continuous(breaks = 0) + 
+  theme(panel.background = element_rect(fill = "grey95")) +
+  theme(panel.grid.major = element_line(color = "black"))
+LDS_PLOT
+LDS_PLOT + facet_wrap(~Population)
+```
+
+Another way to visualize this is to look at the probability of assignment by
+source pouplation.
+
+```{r dapc-region-prob, fig.width = 7, fig.height = 6}
+region_summary <- region.dapc$posterior %>% 
+  as.data.frame() %>% 
+  as_tibble() %>% 
+  add_column(Individual = indNames(dat11cc)) %>% 
+  add_column(`Original Population` = as.character(strata(dat11cc)$Region)) %>% 
+  group_by(`Original Population`) %>%
+  mutate(n = n()) %>%
+  gather(Population, Posterior, -Individual, -`Original Population`, -n) %>% 
+  group_by(Population, `Original Population`, n) %>% 
+  summarize(`Mean Assignment` = mean(Posterior)) %>% 
+  arrange(`Mean Assignment`) %>% 
+  mutate(Source = case_when(
+    Population == `Original Population` ~ "Original Population",
+    TRUE ~ "Other"
+    )) %>%
+  ungroup() %>%
+  arrange(desc(n)) %>%
+  mutate(`Original Population` = forcats::fct_inorder(`Original Population`)) %>%
+  mutate(Population = factor(Population, levels(`Original Population`))) %>%
+  mutate(`Original Population` = paste0(`Original Population`, " (", n, ")")) %>%
+  mutate(`Original Population` = forcats::fct_inorder(`Original Population`))
+REGION_PLOT <- region_summary %>%
+  filter(n >= 10) %>%
+  ggplot(aes(x = Population, y = `Mean Assignment`)) +
+  geom_segment(aes(xend = Population, yend = 0)) +
+  geom_point(aes(fill = Source), pch = 21) + 
+  facet_wrap(~`Original Population`, ncol = 3, strip.position = "top") +
+  scale_fill_grey() +
+  # scale_x_discrete(position = "top") +
+  scale_y_continuous(limits = c(0, 1), minor_breaks = c(seq(0, 1, by = 0.125)), breaks = c(0, 0.5, 1)) +
+  theme_bw(base_size = 16, base_family = "Helvetica") +
+  theme(legend.position = "top") +
+  theme(aspect.ratio = 1/2) +
+  theme(axis.text = element_text(color = "black")) +
+  theme(axis.ticks.y = element_blank()) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
+  # theme(axis.text.x.top = element_text(angle = 90, vjust = 0.5, hjust = 0)) +
+  theme(panel.grid.major.x = element_line(linetype = 0, color = "grey50")) +
+  theme(panel.grid.major = element_line(colour = "grey20")) +
+  theme(panel.grid.minor = element_line(linetype = 3, colour = "grey50")) +
+  # theme(panel.spacing.y = unit(0, "line")) +
+  theme(strip.background = element_rect(color = NA, fill = "grey90")) +
+  theme(strip.text = element_text(face = "bold", hjust = 0.05)) +
+  theme(panel.border = element_blank()) +
+  xlab("Population Assignment") +
+  ylab("Mean Probability")
+REGION_PLOT
+
+```
+
+```{r dapc-region-dualplot, fig.width = 13, fig.height = 6}
+cowplot::plot_grid(REGION_PLOT, LDS_PLOT, 
+                   nrow = 1, 
+                   rel_widths = c(1, 0.85),
+                   label_size = 16,
+                   labels = "auto")
+ggsave(filename = "results/figures/publication/DAPC.pdf")
 ```

### 2017-07-06 doc/manuscript/manuscript.Rmd
Add compound DAPC figure to paper

@@
 
-## DAPC analysis {-}
+## DAPC {-}
+
+![
+Discriminant Analysis on regions showing that Mexico is differentiated from other populations. 
+**(a)** Mean population assignment (horizontal axis) probability (vertical axis) for all populations with N > 10 (facets).
+All values sum to one.
+**(b)** Scatter plot of first two components from DAPC.
+Points represent observed individuals connected to the population centroids with ellipses representing a 66% confidence interval for a normal distribution. 
+](../../results/figures/publication/DAPC.pdf){#DAPC width=100%}
 

### 2017-07-06 doc/RMD/pop-diff.Rmd
change direction of viridis palette

@@
   theme(axis.ticks = element_blank()) + 
-  viridis::scale_color_viridis(discrete = TRUE) +
-  viridis::scale_fill_viridis(discrete = TRUE) +
+  viridis::scale_color_viridis(discrete = TRUE, direction = -1) +
+  viridis::scale_fill_viridis(discrete = TRUE, direction = -1) +
   scale_y_continuous(breaks = 0) + 

### 2017-07-06 doc/manuscript/manuscript.Rmd
change direction of viridis palette

@@
 **(a)** Mean population assignment (horizontal axis) probability (vertical axis) for all populations with N > 10 (facets).
+Numbers next to source populations indicate population size.
 All values sum to one.
@@
 Pop = Population, N = number of individuals (number of MLG in parentheses), 
-eMLG = expected number of MLGs based on rarefaction at 10 individuals (standard error in parentheses), 
+eMLG = expected number of MLGs based on rarefaction at min(10, N) individuals (standard error in parentheses), 
 H = Shannon-Weiner Index, G = Stoddardt and Taylor's Index, 

### 2017-07-06 doc/manuscript/manuscript.Rmd
add comments
update first paragraph

@@
 *Sclerotinia sclerotiorum* (Lib.) de Bary is an ascomycete plant pathogen found worldwide.
-This is a necrotrophic pathogen that is primarily homothallic (self-fertilization) [@bolton2006sclerotinia; @sexton2006population].
-It has a host range of more than 400 species of plants belonging to 75 families [@boland1994index].
-Economically important crops affected by this pathogen include sunflower, soybean, oilseed rape, dry bean, chickpea, peanut, dry pea, lentils and vegetables [@bolton2006sclerotinia].
-*Sclerotinia sclerotiorum* causes white mold disease on dry bean (*Phaseolus vulgaris* L.), which can be a yield-limiting in temperate climates [@steadman1983white].
+This is a necrotrophic pathogen that is primarily homothallic (self-fertilization) [@bolton2006sclerotinia; @sexton2006population] that causes disease on more than 400 plant species belonging to 75 families [@boland1994index].
+These include crops of major economic importance such as sunflower (*Helianthus spp.*), soybean (*Glycine max* L.), canola (*Brassica napa* L., *Brassica campestris* L.), and dry bean (*Phaseolus vulgaris* L.) [@bolton2006sclerotinia].
+On dry bean, *S. sclerotiorum* is the causal agent of White Mold, a devastating disease that can be a yield-limiting in temperate climates [@steadman1983white].
 
@@
 
+<!-- MCG abundance figure -->
+
 ![
@@
 ](../../results/figures/publication/Figure1Z.pdf){#barplot width=50%}
+
+<!-- end figure -->
 
@@
 
+<!-- MCG graph figure -->
+
 ![
@@
 ](../../results/figures/publication/FigureS2.pdf){#biggraph width=100%}
+
+<!-- end fig -->
 
@@
 
+<!-- RDA figure -->
+
 ![
@@
 
+<!-- end figure -->
+
 ## DAPC {-}
+
+<!-- DAPC figure -->
 
@@
 Points represent observed individuals connected to the population centroids with ellipses representing a 66% confidence interval for a normal distribution. 
+The center of each component is represented as grid lines.
 ](../../results/figures/publication/DAPC.pdf){#DAPC width=100%}
+
+<!-- end figure -->
 

### 2017-07-06 doc/manuscript/manuscript.Rmd
update sampling paragraph

@@
 Sclerotia of *S.
-sclerotiorum* were collected over several years from grower fields and/or dry bean nurseries in Mexico, France, Australia, and 11 states of the United States of America (Table \@ref(tab:isolate-table)).
+sclerotiorum* were collected over several years from grower fields and/or dry bean nurseries in 11 states of the United States of America, Mexico, France, and Australia (Table \@ref(tab:isolate-table)).
 The dry bean nurseries in different regions are indicated by wmn (white mold nursery) or mtc (Mitchell) and all other isolates came from commercial fields (Table \@ref(tab:isolate-table)).
-The bean nurseries were 5m x 10 m in size and maintained without applying fungicides to observe natural incidence of white mold disease.
-The nursery plots were incorporated with a basal dressing of N:P:K = 1:3:2 and side dressing of 0:3:2 during the season.
-Although the nursery locations were the same over sampling years, sampling fields within a location varied for sampling years.
-Sampling was carried out on a 'W' or 'V' shape transect, or randomly.
-Sampling year and number of isolates collected were necessarily influenced by incidence and severity of white mold disease.
+The dry bean nurseries were 5m x 10m in size and maintained without applying fungicides to observe natural incidence of white mold.
+The nursery plots were incorporated with a basal dressing of N:P:K = 1:3:2 and side dressing of 0:3:2 during the growing season.
+Although the nursery locations were the same over sampling years, sampling plots within a location varied for sampling years.
+Sampling of diseased tissue was carried out on a 'W' or 'V' shape transect, or randomly.
+Because sampling depended on disease incidence, the number of samples isolated varied from year to year.
 

### 2017-07-06 doc/manuscript/manuscript.Rmd
fix growth and severity

@@
 These isolates (N = 366) were phenotypically characterized by determining mycelial compatibility group MCG and aggressiveness.
-MCG was determined through co-culturing isolates in pairs on PDA plates, where incompatibility of different MCG resulted in formation of a mycelial barrage [@kohn1990mycelial] and isolates belonging to the same MCG did not form a mycelial barrage.
-Aggressiveness of each isolate was assessed using a straw test [@petzoldt1996straw; @teran2006modified] that rates necrotic lesion on a scale of 1-9, with ascending numbers assigned for increasing aggressiveness.
+MCG was determined through co-culturing isolates in pairs on potato-dextrose agar (PDA) plates, where incompatibility of different MCG resulted in formation of a mycelial barrier [@kohn1990mycelial].
+Aggressiveness of each isolate was assessed using a straw test that rates necrotic lesion on a scale from 1--9 with increasing aggressiveness [@petzoldt1996straw; @teran2006modified].
 DNA was purified from mycelium of each isolate using a Qiagen DNeasy Plant Mini Kit (Qiagen Inc., Valencia, CA) according to the manufacturer's instructions.

### 2017-07-06 doc/RMD/wmn-differentiation.Rmd
don't print whole friggen table :zap:

@@
                    .id = "Comparison") %>%
-  mutate(Comparison = forcats::fct_inorder(Comparison))
+  mutate(Comparison = forcats::fct_inorder(Comparison)) %>%
+  as_tibble()
 dists

### 2017-07-06 doc/manuscript/manuscript.Rmd
update aims
I updated the specific aims and goals of the manuscript. These are
not different than the previous aims, they are just reworded to make
a bit more sense.

@@
 
-We analyzed *S.
-sclerotiorum* isolates collected from dry bean cultivars from 2003 to 2012 from different geographic locations in America, Europe and Australia using microsatellite markers to delineate genetic differences and reproduction behavior.
-The specific aims of our study were:  a) To assess association of phenotypic data (aggressiveness) and environmental data (bean cultivar, region, field, year) on genotypic variation of *S.
-sclerotiorum* isolates, b) To test the hypothesis that *S.
-sclerotiorum* populations of dry beans originated from different continents were genetically different, and c) although previous studies show that *S.
-sclerotiorum* populations are largely clonal, few compare populations from different continents [@lehner2017sclerotinia], c) To identify localized genotypes representative of each location and also investigate if there are widespread genotypes common among locations.
-d) To test potential for sexual reproduction in sub populations of the pathogen isolates under different conditions such as year or field in each region, year/region, field/region/year, and e) Test if pathogen within-field populations differentiate over time.
-
-The overall goal of this work was to identify representative isolates for use in breeding programs.
+In this study, we analyze and characterize the genetic diversity of 366 *S.
+sclerotiorum* isolates collected between 2003 and 2012 from dry bean cultivars among different geographic locations in the United States, France, Mexico, and Australia.
+Because the white mold nurseries serve as testing grounds for new varieties of dry bean, we wanted to know if the *S.
+sclerotiorum* populations from white mold nurseries were representative of the fields within the same region.
+As these nurseries are not treated with any chemical or cultural control of white mold, we hypothesized that these nurseries would represent the natural population of *S. sclerotiorum*.
+Furthermore, as three cultivars (Beryl, Bunsi, and G122) were grown among all white mold nurseries, we wanted to investigate the potential effect of cultivar on genetic diversity.
+We additionally wanted to assess what categorical or phenotypic variables best predict genetic structure and if there was correlation between multilocus genotype and mycelial compatibility group.
 
@@
 Because sampling depended on disease incidence, the number of samples isolated varied from year to year.
+In total, we collected 366 isolates.
 
@@
 Mycelial growth of each isolate was generated by activating sclerotia according to the method described previously [@matheron1989in-vitro].
-These isolates (N = 366) were phenotypically characterized by determining mycelial compatibility group MCG and aggressiveness.
+These isolates were phenotypically characterized by determining mycelial compatibility group MCG and aggressiveness.
 MCG was determined through co-culturing isolates in pairs on potato-dextrose agar (PDA) plates, where incompatibility of different MCG resulted in formation of a mycelial barrier [@kohn1990mycelial].
@@
 
+> ZNK: how were the data received and formatted?
+
 ## Data processing and Analyses {-}
 
-All data processing and analyses were performed in R version 3.4.0 [@R].
-Of the 16 microsatellite loci, five included compound repeats, making allele scoring difficult. 
+Of the 16 microsatellite loci, five included compound repeats, which made allele scoring difficult. 
 These loci were removed for the reported statistics, but were processed and included in data analyses to ensure the integrity of the results with 11 loci as compared to 16.
@@
 All analyses were performed on this data set unless otherwise mentioned.
-All data and analyses are openly available and reproducible at https://github.com/everhartlab/sclerotinia-366/
+All data processing and analyses were performed in R version 3.4.0 [@R] and are openly available and reproducible at https://github.com/everhartlab/sclerotinia-366/
 

### 2017-07-07 doc/manuscript/manuscript.Rmd
clarify intro

@@
 
-In this study, we analyze and characterize the genetic diversity of 366 *S.
+In this study, we analyze and characterize the genetic and phenotypic diversity of 366 *S.
 sclerotiorum* isolates collected between 2003 and 2012 from dry bean cultivars among different geographic locations in the United States, France, Mexico, and Australia.
@@
 We used analysis of variance (ANOVA) to assess if MCGs differed in severity.
-We chose the top 10 MCGs, representing 57% of the isolates collected.
+To minimize complications due to small sample sizes, we chose the top 10 MCGs, representing 57% of the isolates collected.
 Between-group differences were assessed using Tukey's HSD test using the `HSD.test()` function in the package *agricolae* version 1.2.4 [@mendiburu2015agricolae].

### 2017-07-07 doc/manuscript/manuscript.Rmd
add more context to intro
add statement about GAC in processing

@@
 We additionally wanted to assess what categorical or phenotypic variables best predict genetic structure and if there was correlation between multilocus genotype and mycelial compatibility group.
+By investigating these aims, we hope to effectively describe the population dynamics of *S. sclerotiorum* and make available our database of isolates for use in future dry bean breeding efforts.
 
@@
 These loci were removed for the reported statistics, but were processed and included in data analyses to ensure the integrity of the results with 11 loci as compared to 16.
+We assessed the power of our 11 markers by generating a genotype accumulation curve in the R package *poppr* version 2.4.1, looking for evidence of saturation, which would indicate that enough loci have been sampled to sufficiently characterize the full set of genotypes [@arnaud2007standardizing, @kamvar2015novel].
 To control for non-independent samples (i.e. those sampled from the same location, region, year, and plant), data were clone-corrected for subsequent analyses.

### 2017-07-07 doc/manuscript/manuscript.Rmd
Update description of dbRDA section

@@
 
-A distance-based redundancy analysis (dbRDA) [@legendre1999distance] was performed with the function `capscale()` in the *vegan* package version 2.4-3 [@vegan] to delineate relative contribution of phenotypic, spatial, and temporal components in influencing genetic composition of the populations.
-dbRDA is a method of conducting constrained ordinations on data matrices consisting of non-Euclidean distance measures.
-RDA is a multivariate linear regression analysis of independent (explanatory: eg. environmental and phenotypic) variables on dependent (non-explanatory or response) variables.
-In our case, a pairwise Bruvo's [@bruvo2004simple] distance matrix was generated from microsatellite data using *poppr* 2.4.1 [@kamvar2014poppr; @kamvar2015novel] and employed it as the dependent matrix for RDA analysis.
-To identify environmental variable(s) correlated with genetic variation, a forward-backward selection process was applied with the vegan function `ordistep()` to identify the set of variables that explained more variation than predicted with the global model containing all variables.
+A distance-based redundancy analysis (dbRDA) [@legendre1999distance] was performed with the function `capscale()` in the *vegan* package version 2.4-3 [@vegan].
+This method uses constrained ordinations on a distance matrix representing the response variable to delineate relative contribution of any number of independent explanatory variables. 
+We used this method to delineate the phenotypic, geographic, and temporal components in influencing genetic composition of the populations.
+<!-- RDA is a multivariate linear regression analysis of independent (explanatory: eg. environmental and phenotypic) variables on dependent (non-explanatory or response) variables. -->
+The distance matrix we used as the response variable was generated using Bruvo's genetic distance as implemented in *poppr* version 2.4.1, which employs a stepwise mutation model for microsatellite data [@bruvo2004simple, @kamvar2014poppr].
+<!-- In our case, a pairwise Bruvo's [@bruvo2004simple] distance matrix was generated from microsatellite data using *poppr* 2.4.1 [@kamvar2014poppr; @kamvar2015novel] and employed it as the dependent matrix for RDA analysis.
+-->
+To identify explanatory variable(s) correlated with genetic variation, a forward-backward selection process was applied with the vegan function `ordistep()` to identify the set of variables that explained more variation than predicted with the global model containing all variables.
 An ANOVA was then performed to test for significance of the reduced model and marginal effects using 999 permutations.

### 2017-07-07 doc/manuscript/manuscript.Rmd
Create broad category; move agressiveness
add initial sentence for Genotype community analysis

@@
 
-### Aggressiveness Assessment {-}
-
-We used analysis of variance (ANOVA) to assess if MCGs differed in severity.
-To minimize complications due to small sample sizes, we chose the top 10 MCGs, representing 57% of the isolates collected.
-Between-group differences were assessed using Tukey's HSD test using the `HSD.test()` function in the package *agricolae* version 1.2.4 [@mendiburu2015agricolae].
+## Assessing importance of categorical and phenotypic variables  {-}
 
@@
 
+### Aggressiveness assessment {-}
+
+We used analysis of variance (ANOVA) to assess if MCGs differed in severity.
+To minimize complications due to small sample sizes, we chose the top 10 MCGs, representing 57% of the isolates collected.
+Between-group differences were assessed using Tukey's HSD test using the `HSD.test()` function in the package *agricolae* version 1.2.4 [@mendiburu2015agricolae].
+
 ### Shared Multilocus Genotype Community Analysis {-}
 
-To assess the connectivity of shared genotypes across populations, we first tabulated the multilocus genotypes shared between at least two populations (defined as states or countries) with the *poppr* function `mlg.crosspop()` [@kamvar2014poppr].
+As many disparate geographic regions shared multilocus genotypes, we wanted to know if we could find any patterns in connectivity. 
+To assess the connectivity of shared genotypes across regions, we first tabulated the multilocus genotypes shared between at least two populations (defined as states or countries) with the *poppr* function `mlg.crosspop()` [@kamvar2014poppr].
 From these data, we were able to construct a graph with populations as nodes and shared genotypes as edges (connections) between the nodes using the R packages *igraph* version 1.0.1 [@csardi2006igraph], *dplyr* version 0.7.1 [@dplyr], and *purrr* version 0.2.2.2 [@purrr].

### 2017-07-07 doc/manuscript/manuscript.Rmd
Add broad section for Genetic Diversity
Add description of MLG-MCG correlation

@@
 Between-group differences were assessed using Tukey's HSD test using the `HSD.test()` function in the package *agricolae* version 1.2.4 [@mendiburu2015agricolae].
+
+### Correlating Multilocus Genotypes with Mycelial Compatibility Groups {-}
+
+We wanted to assess if there was any correlation between multilocus genotypes (MLG) and mycelial compatibility groups (MCG). 
+This was done using a graph-based approach with community analysis.
+We first created a contingency table with MLGs and MCGs and converted it to a directed and weighted edgelist where each edge was a connection from an MCG to an MLG weighted by the number of samples shared in the connection.
+The cluster walktrap community detection algorithm as implemented in the `cluster_walktrap()` function in *igraph* version 1.0.1 was used to define clusters of MLGs and MCGs [@csardi2006igraph, @pons2006computing].
+All communities that had fewer than 10 members were then consolidated into one.
+These community definitions was then used to assess the average genetic distance (as defined by Bruvo's distance) within members of the community [@bruvo2004simple].
+
+## Genetic Diversity
 

### 2017-07-07 doc/manuscript/manuscript.Rmd
fix citations (use ; not ,)

@@
 These loci were removed for the reported statistics, but were processed and included in data analyses to ensure the integrity of the results with 11 loci as compared to 16.
-We assessed the power of our 11 markers by generating a genotype accumulation curve in the R package *poppr* version 2.4.1, looking for evidence of saturation, which would indicate that enough loci have been sampled to sufficiently characterize the full set of genotypes [@arnaud2007standardizing, @kamvar2015novel].
+We assessed the power of our 11 markers by generating a genotype accumulation curve in the R package *poppr* version 2.4.1, looking for evidence of saturation, which would indicate that enough loci have been sampled to sufficiently characterize the full set of genotypes [@arnaud2007standardizing; @kamvar2015novel].
 To control for non-independent samples (i.e. those sampled from the same location, region, year, and plant), data were clone-corrected for subsequent analyses.
@@
 <!-- RDA is a multivariate linear regression analysis of independent (explanatory: eg. environmental and phenotypic) variables on dependent (non-explanatory or response) variables. -->
-The distance matrix we used as the response variable was generated using Bruvo's genetic distance as implemented in *poppr* version 2.4.1, which employs a stepwise mutation model for microsatellite data [@bruvo2004simple, @kamvar2014poppr].
+The distance matrix we used as the response variable was generated using Bruvo's genetic distance as implemented in *poppr* version 2.4.1, which employs a stepwise mutation model for microsatellite data [@bruvo2004simple; @kamvar2014poppr].
 <!-- In our case, a pairwise Bruvo's [@bruvo2004simple] distance matrix was generated from microsatellite data using *poppr* 2.4.1 [@kamvar2014poppr; @kamvar2015novel] and employed it as the dependent matrix for RDA analysis.
@@
 We first created a contingency table with MLGs and MCGs and converted it to a directed and weighted edgelist where each edge was a connection from an MCG to an MLG weighted by the number of samples shared in the connection.
-The cluster walktrap community detection algorithm as implemented in the `cluster_walktrap()` function in *igraph* version 1.0.1 was used to define clusters of MLGs and MCGs [@csardi2006igraph, @pons2006computing].
+The cluster walktrap community detection algorithm as implemented in the `cluster_walktrap()` function in *igraph* version 1.0.1 was used to define clusters of MLGs and MCGs [@csardi2006igraph; @pons2006computing].
 All communities that had fewer than 10 members were then consolidated into one.
@@
 
-## Genetic Diversity
+## Genetic Diversity {-}
 
@@
 Each edge represents a single MLG, but because a single MLG could cross populations, that MLG may be represented by (n\*(n-1))/2 edges where n represents the number of populations it crosses.
-Edges were weighted by $1-P_{sex}$, where $P_{sex}$ is the probability of encountering the same genotype via two independent meiotic events [@arnaud2007standardizing, @parks1993study].
+Edges were weighted by $1-P_{sex}$, where $P_{sex}$ is the probability of encountering the same genotype via two independent meiotic events [@arnaud2007standardizing; @parks1993study].
 We then used the `cluster_optimal()` function from *igraph* to define communities [@csardi2006igraph].

### 2017-07-10 doc/manuscript/manuscript.Rmd
add section on population differentiation

@@
 
+### Population Differentiation {-}
+
+We used analysis of molecular variance (AMOVA) with Bruvo's genetic distance to quantify the differentiation conferred by source and region [@excoffier1992analysis; @bruvo2004simple].
+Source in this case was defined as a population belonging to a wmn or not and region defined as state. 
+To identify specific regions that were more differentiated than others, we used discriminant analysis of principal components (DAPC) as implemented in *adegenet* version 2.1.0, assessing the per-sample posterior group assignment probability [@jombart2008adegenet].
+This method decomposes the genetic data into principal components, and then uses a subset of these as the inputs for discriminant analysis, which attempts to minimize within-group variation and maximize among-group variation [@jombart2010discriminant].
+To avoid over-fitting the data, the optimal number of principal components was selected by using the *adegenet* function `xvalDapc()`. 
+This function implements a cross-validation procedure to iterate over an increasing number of principal components on a subset (90%) of the data, trying to find the minimum number of principal components that maximizes the rate of successful group reassignment.
+We additionally assessed genotypic diversity by assessing Stoddart and Taylor's index (*G*) [@stoddart1988genotypic], Shannon's index (*H*) [@shannon2001mathematical], Simpson's index ($\lambda$) [@simpson1949measurement], and evenness ($E_5$, *eMLG*) [@heck1975explicit; @hurlbert1971nonconcept; @pielou1975ecological; @grunwald2003analysis].
+To assess the potential for random mating, we tested for linkage disequilibrium with the index of association, $I_A$ and its standardized version, $\bar{r}_d$ using 999 permutations [@agapow2001indices; @smith1993how; @brown1980multilocus].
+Both genotypic diversity and linkage disequilibrium were calculated in *poppr* version 2.4.1 [@kamvar2014poppr].
+
 ### Shared Multilocus Genotype Community Analysis {-}
@@
 
+<!-- 
 ### Discriminant analysis of principal components {-}
@@
 An $F_{st}$ value of 0.05 or less is generally considered as reasonably low and may be interpreted to mean that structuring between populations is weak [@jost2008gst].
-<!-- **(Balloux and Ugon-Moulin, 2002; Wright, 1978)**.   -->
+**(Balloux and Ugon-Moulin, 2002; Wright, 1978)**.   -->
 
@@
 
-Table: (\#tab:amova) Analysis of Molecular Variance showing little differentiation between White Mold Nurseries and other sources.
+Table: (\#tab:amova) Analysis of Molecular Variance on Bruvo's genetic distance showing little differentiation between White Mold Nurseries and other sources.
 The hierarchy was constructed as Source/Region where source is defined as belonging to a white mold nursery or not.
@@
 *h* = Nei's 1978 gene diversity, 
+$E_5$ = Evenness,
 $I_A$ = the index of association, 

### 2017-07-10 doc/RMD/pop-diff.Rmd
update analyses

@@
   mutate(Ia = case_when(
-    !is.finite(.$p.Ia) ~ "-",
-    .$p.Ia == 0.001 ~ paste0(round(.$Ia, 2), "*"),
-    .$p.Ia <= 0.01 ~  paste0(round(.$Ia, 2), "-"),
-    .$p.Ia <= 0.05 ~  paste0(round(.$Ia, 2), "~"),
-    TRUE ~ as.character(round(.$Ia, 2))
+    !is.finite(p.Ia) ~ "-",
+    p.Ia == 0.001 ~ paste0(round(Ia, 2), "*"),
+    p.Ia <= 0.01  ~ paste0(round(Ia, 2), "-"),
+    p.Ia <= 0.05  ~ paste0(round(Ia, 2), "~"),
+    TRUE          ~ as.character(round(Ia, 2))
   )) %>%
   mutate(rbarD = case_when(
-    !is.finite(.$p.rD) ~ "-",
-    .$p.rD == 0.001 ~ paste0(round(.$rbarD, 2), "*"),
-    .$p.rD <= 0.01 ~  paste0(round(.$rbarD, 2), "-"),
-    .$p.rD <= 0.05 ~  paste0(round(.$rbarD, 2), "~"),
-    TRUE ~ as.character(round(.$rbarD, 2))
+    !is.finite(p.rD) ~ "-",
+    p.rD == 0.001 ~ paste0(round(rbarD, 2), "*"),
+    p.rD <= 0.01  ~ paste0(round(rbarD, 2), "-"),
+    p.rD <= 0.05  ~ paste0(round(rbarD, 2), "~"),
+    TRUE          ~ as.character(round(rbarD, 2))
   )) %>%

### 2017-07-10 doc/RMD/wmn-differentiation.Rmd
update analyses

@@
 
-```{r dapc, fig.width = 5, fig.height = 5.5}
-setPop(bebug) <- ~Host/Region
+```{r dapc, fig.width = 5, fig.height = 4.5}
+setPop(bebug) <- ~Host
 bebug_dapc <- dapc(bebug, n.pca = 20, n.da = 2)
@@
         scree.da = FALSE,
-        # col = viridis::plasma(3), 
+        col = viridis::plasma(3, end = 0.85),
         bg.inset = "grey90",
@@
 
-ggcompoplot::ggcompoplot(bebug_dapc, bebug) + #, pal = viridis::plasma(3)) +
+ggcompoplot::ggcompoplot(bebug_dapc, setPop(bebug, ~Region), cols = 5, pal = viridis::plasma(3, end = 0.85)) +
   theme(legend.position = "top")
+```
+
+```{r dapc_hr, fig.width = 5, fig.height = 5.5}
+setPop(bebug) <- ~Host/Region
+bebug_dapc <- dapc(bebug, n.pca = 20, n.da = 20)
+bebug_dapc
+```
+
+```{r ggdapc, fig.width = 8, fig.height = 4}
+LDS <- bind_cols(Population = bebug_dapc$grp, as.data.frame(bebug_dapc$ind.coord)) %>%
+  as_tibble()
+LDS_pop <- LDS %>% 
+  group_by(Population) %>% 
+  summarize_all(mean) %>%
+  rename_all(function(x) gsub("LD", "mean", x))
+LDS <- full_join(LDS, LDS_pop, by = "Population") %>% separate(Population, c("Cultivar", "Population"))
+LDS_PLOT <- ggplot(LDS, aes(x = LD1, y = LD2, color = Cultivar)) + 
+  geom_point(aes(fill = Cultivar), alpha = 0.5, pch = 21, color = "black") +
+  geom_segment(aes(x = mean1, y = mean2, xend = LD1, yend = LD2), alpha = 0.5) +
+  stat_ellipse(type = "norm", level = 0.66, alpha = 0.75) + 
+  theme_bw(base_size = 16, base_family = "Helvetica") +  
+  theme(aspect.ratio = 1/1.618) +
+  theme(legend.position = "bottom") +
+  theme(axis.text = element_blank()) + 
+  theme(axis.title = element_blank()) + 
+  theme(axis.ticks = element_blank()) + 
+  viridis::scale_color_viridis(option = "B", discrete = TRUE, direction = -1, end = 0.85) +
+  viridis::scale_fill_viridis(option = "B", discrete = TRUE, direction = -1, end = 0.85) +
+  scale_y_continuous(breaks = 0) + 
+  scale_x_continuous(breaks = 0) + 
+  theme(panel.background = element_rect(fill = "grey95")) +
+  theme(panel.grid.major = element_line(color = "black")) +
+  facet_wrap(~Population, nrow = 2)
+LDS_PLOT
 ```

### 2017-07-10 doc/manuscript/manuscript.Rmd
moved comment
tentatively finished with methods.

@@
 
-We used analysis of molecular variance (AMOVA) with Bruvo's genetic distance to quantify the differentiation conferred by source and region [@excoffier1992analysis; @bruvo2004simple].
+We used analysis of molecular variance (AMOVA) with Bruvo's genetic distance in *poppr* version 2.4.1 to assess differentiation associated with wmn by quantifying the differentiation conferred by source and region [@excoffier1992analysis; @bruvo2004simple; @kamvar2014poppr].
 Source in this case was defined as a population belonging to a wmn or not and region defined as state. 
@@
 Both genotypic diversity and linkage disequilibrium were calculated in *poppr* version 2.4.1 [@kamvar2014poppr].
+
+### Cultivar Differentiation {-}
+
+To assess if cultivar had an influence on genetic diversity across wmn, we first subset the clone-corrected data to contain only samples from wmn and from the cultivars Beryl, Bunsi, and G122 and tested differentiation using AMOVA and DAPC as described above.
 
@@
 To ensure that we were able to capture the same community signal, we additionally ran this analysis including the 5 polymorphic markers described above.
-
-<!-- 
-### Discriminant analysis of principal components {-}
-
-DAPC is a multivariate method designed to identify and describe clusters of genetically related individuals.
-Discriminant analysis maximizes separation between groups while minimizing variation within groups [@jombart2010discriminant].
-The `find.clusters()` function in *adegenet* 2.0.1 [@jombart2008adegenet] uses a k-means clustering algorithm to identify the optimum number of clusters.
-It was not able to identify a single optimum and identified a range of optimal clusters between 5-15.
-Consequently, DAPC was performed several times with different optimal number of clusters to determine whether isolates (N = 318) fell into each cluster have any biological meaning such as belonging to a particular host or came from the same region etc.
-Since dbRDA identified region, host, and sampling year as the top three variables influencing microsatellite data, DAPC was also performed on data defined according to each of these variables, separately.
-Isolates (N = 318) were assigned to different levels of the three factors (region, host, year) and assessed predefined group size (original number of isolates at each level of the factor), post group size (number of isolates assigned by the DAPC algorithm) and the proportion of the original isolates assigned to the original group after analysis.
-This was performed to determine the suitability of explanatory variables (i.e. region, host, and year) in describing the clusters generated.
-Clone-corrected isolates sampled in different years from the white mold nurseries (wmn) were tested for population differentiation over time by DAPC analysis.
-Only WA, MI, and NE isolates were qualified since we filtered out the states that had at least $\geq$ 7 samples/year with at least two years of data for the analysis.
-Lastly, we pooled all isolates, clone corrected them (N = 165) and performed DAPC using the function find.clusters as described above to delineate how many clusters were formed by these unique genotypes.
-AMOVA was used to test the significance of clusters resulted by different DAPC analyses using GenAlEx version 6.41 [@peakall2012genalex].
-$\Phi_{PT}$ which is an analogue of Wright's $F_{st}$ was used to ascertain the significance of populations.
-An $F_{st}$ value of 0.05 or less is generally considered as reasonably low and may be interpreted to mean that structuring between populations is weak [@jost2008gst].
-**(Balloux and Ugon-Moulin, 2002; Wright, 1978)**.   -->
-
-### Molecular population statistics and linkage disequilibrium {-}
-
-Genotypic evenness [@grunwald2003analysis], which is a measure of distribution of genotype abundances and unbiased gene diversity (*h*) [@nei1978estimation] were calculated using *poppr* package in R [@kamvar2014poppr].
-Microsatellite data were also tested for evidence of linkage disequilibrium, where significant disequilibrium between pairs of loci in a population is attributed to linkage between loci and may be an indication of clonal reproduction if loci are physically unlinked, whereas linkage equilibrium can be an indication of outcrossing populations.
-The linkage disequilibrium of clone-corrected data was determined by calculating the index of association ($I_{A}$) and its standardized version ($\bar{r}_d$) and a permutation approach with 999 permutations was used to test whether observed values were significantly different from data randomized at each locus [@agapow2001indices; @brown1980multilocus].
-The value of $I_A$ tends to increase with the number of loci and therefore we made our inferences based on $\bar{r}_d$, which standardizes the number of loci tested and considered a more robust method [@agapow2001indices].
-Higher observed $\bar{r}_d$ or $I_A$ values would indicate clonally reproducing populations.
 

### 2017-07-10 doc/manuscript/manuscript.Rmd
Suggested revisions and comments for methods.

@@
 The dry bean nurseries in different regions are indicated by wmn (white mold nursery) or mtc (Mitchell) and all other isolates came from commercial fields (Table \@ref(tab:isolate-table)).
+> SEE: I was not sure what "dry bean nurseries" were because I'm use to seeing them referred to as the white mold nurseries.
+I think if the introduction uses one term for it, this would be acceptable.
+That or consider providing a description of the white mold nurseries in the methods section.
+This now reminds me..
+there is a publication by Lindsey Otto-Hanson and Steadman that gave a good description of these.
+In fact, she called them multisite screening locations.
+Here's the link: http://apsjournals.apsnet.org/doi/abs/10.1094/PDIS-11-10-0865
+
 The dry bean nurseries were 5m x 10m in size and maintained without applying fungicides to observe natural incidence of white mold.
@@
 Because sampling depended on disease incidence, the number of samples isolated varied from year to year.
-In total, we collected 366 isolates.
+In total, we collected 366 isolates that were used in the present study.
 
@@
 
-Table: (\#tab:isolate-table) Description of *S. sclerotiorum* isolates used in this study. 
+Table: (\#tab:isolate-table) Description of *Sclerotinia sclerotiorum* isolates used in this study.
 N = Number of Isolates. 
@@
 
-Mycelial growth of each isolate was generated by activating sclerotia according to the method described previously [@matheron1989in-vitro].
-These isolates were phenotypically characterized by determining mycelial compatibility group MCG and aggressiveness.
-MCG was determined through co-culturing isolates in pairs on potato-dextrose agar (PDA) plates, where incompatibility of different MCG resulted in formation of a mycelial barrier [@kohn1990mycelial].
-Aggressiveness of each isolate was assessed using a straw test that rates necrotic lesion on a scale from 1--9 with increasing aggressiveness [@petzoldt1996straw; @teran2006modified].
-DNA was purified from mycelium of each isolate using a Qiagen DNeasy Plant Mini Kit (Qiagen Inc., Valencia, CA) according to the manufacturer's instructions.
+Each isolate was regenerated by activating one or several sclerotia according to the method described previously [@matheron1989in-vitro].
+Once actively growing, mycelial compatibility group (MCG) and aggressiveness was determined for each isolate.
+MCG was determined through co-culturing isolates in pairs on potato-dextrose agar (PDA) plates, where incompatibility of different MCG resulted in formation of a mycelial barrier or barrage line [@kohn1990mycelial].
+Since MCG are considered mutually exclusive, pairing was not required between all isolates in the study.
+> SEE: Is that correct?  Perhaps consult with Becky or Rachana and then elaborate upon the sentence above.
+
+Aggressiveness of each isolate was assessed using a straw test that rated necrotic lesion size on a scale from 1--9 with increasing aggressiveness [@petzoldt1996straw; @teran2006modified].
+> SEE: Rating is not clear from the single sentence above.  On what plant host was it evaluated?  In the greenhouse?  What is a rating of 1 compared to a rating of 9 on the straw test?
+
+DNA was purified from 100 mg mycelium collected from an actively growing culture of each isolate using a Qiagen DNeasy Plant Mini Kit (Qiagen Inc., Valencia, CA) according to the manufacturer's instructions.
+> SEE: More details on how tissue was disrupted may be needed here.
 
@@
 
-We characterized *S. sclerotiorum* isolates for microsatellite polymorphism using 16 primer pairs developed by @sirjusingh2001characterisation.
+We genotyped *S. sclerotiorum* isolates using 16 microsatellite primer pairs developed previously [@sirjusingh2001characterisation].
 Polymerase chain reaction (PCR) was carried out as described using primers labeled with FAM fluorophore, and resulting PCR products were first resolved in a 1.5% agarose gel stained with ethidium bromide to ensure product was within the expected size range.
-Capillary electrophoresis (fragment analysis) of PCR products added with size standard GeneScan™ 500 LIZ® was performed using an ABI 3730 genetic analyzer (Life Technologies Corporation, Carlsbad, CA) at the Michigan State University Genomic Sequencing Center.
+> SEE: Pretty rare to see PCR spelled out in full these days. I suggest just using the abbreviation and if PeerJ asks us to spell it out, we will.
+
+Capillary electrophoresis (fragment analysis) of amplicons, with size standard GeneScan™ 500 LIZ®, was performed using an ABI 3730 genetic analyzer (Life Technologies Corporation, Carlsbad, CA) at the Michigan State University Genomic Sequencing Center.
 
@@
 
+> SEE: data were received as .fsa files, which are fasta-type files.
+I have the raw data if you want them.
+There is, apparently, an R-package called Fragman used for reading the files and can be used for scoring.
+Anyway, in this case, I believe Rachana used PeakScanner (free software) to read the files and she recorded the fragment sizes manually (ie.
+read the chromatogram and wrote down in Excel what size the peak was).
+
 ## Data processing and Analyses {-}
 
-Of the 16 microsatellite loci, five included compound repeats, which made allele scoring difficult. 
-These loci were removed for the reported statistics, but were processed and included in data analyses to ensure the integrity of the results with 11 loci as compared to 16.
-We assessed the power of our 11 markers by generating a genotype accumulation curve in the R package *poppr* version 2.4.1, looking for evidence of saturation, which would indicate that enough loci have been sampled to sufficiently characterize the full set of genotypes [@arnaud2007standardizing; @kamvar2015novel].
-To control for non-independent samples (i.e. those sampled from the same location, region, year, and plant), data were clone-corrected for subsequent analyses.
+Of the 16 microsatellite loci, five included compound repeats, which made it difficult to accurately/confidently bin alleles into fragment sizes expected for each locus based on the described repeat motif.
+> SEE: I would not say allele scoring was difficult, because scoring an allele that is different than the size expected is really the same difficulty level as one that is at the expected size.
+It's the accurate binning of alleles into the expected fragment size that is difficult.
+
+Loci with compound repeats were removed for the reported statistics, but were processed and included in data analyses to ensure the integrity of the results with 11 loci as compared to 16.
+We assessed the power of our 11 markers by generating a genotype accumulation curve in the R package *poppr* version 2.4.1, looking for evidence of saturation, which would indicate that loci were sufficiently sampled to adequately represent the full set of genotypes [@arnaud2007standardizing; @kamvar2015novel].
+To control for non-independent samples (i.e. samples from the same location, region, year, and plant), data were clone-corrected for subsequent analyses.
+> SEE: it is not clear what non-independent samples are.  Although your description is accurate and concise, a more general description might enable the reader to better understand what is meant here.
+
 The clone-corrected data set included 318 isolates.
@@
 <!-- RDA is a multivariate linear regression analysis of independent (explanatory: eg. environmental and phenotypic) variables on dependent (non-explanatory or response) variables. -->
-The distance matrix we used as the response variable was generated using Bruvo's genetic distance as implemented in *poppr* version 2.4.1, which employs a stepwise mutation model for microsatellite data [@bruvo2004simple; @kamvar2014poppr].
+The distance matrix we used as the response variable was generated using Bruvo's genetic distance as implemented in *poppr* version 2.4.1, which employed a stepwise mutation model for microsatellite data [@bruvo2004simple; @kamvar2014poppr].
 <!-- In our case, a pairwise Bruvo's [@bruvo2004simple] distance matrix was generated from microsatellite data using *poppr* 2.4.1 [@kamvar2014poppr; @kamvar2015novel] and employed it as the dependent matrix for RDA analysis.
@@
 
-We used analysis of variance (ANOVA) to assess if MCGs differed in severity.
+We used analysis of variance (ANOVA) to assess if MCGs differed in aggressiveness.
 To minimize complications due to small sample sizes, we chose the top 10 MCGs, representing 57% of the isolates collected.
 Between-group differences were assessed using Tukey's HSD test using the `HSD.test()` function in the package *agricolae* version 1.2.4 [@mendiburu2015agricolae].
+> SEE: It's not clear what groups were used to compare aggressiveness in the ANOVA.  What alpha-level was used for the cut-off?
 
@@
 
-We wanted to assess if there was any correlation between multilocus genotypes (MLG) and mycelial compatibility groups (MCG). 
+We wanted to assess if there was correlation between multilocus genotypes (MLG) and mycelial compatibility groups (MCG). 
 This was done using a graph-based approach with community analysis.
@@
 All communities that had fewer than 10 members were then consolidated into one.
-These community definitions was then used to assess the average genetic distance (as defined by Bruvo's distance) within members of the community [@bruvo2004simple].
+These community definitions were then used to assess the average genetic distance (as defined by Bruvo's distance) within members of the community [@bruvo2004simple].
 
 ## Genetic Diversity {-}
+> SEE: deleting this header?
 
@@
 Source in this case was defined as a population belonging to a wmn or not and region defined as state. 
+> SEE: is mtc included with "wmn"? At the beginning of the methods, I thought both were considered dry bean nurseries.
+
 To identify specific regions that were more differentiated than others, we used discriminant analysis of principal components (DAPC) as implemented in *adegenet* version 2.1.0, assessing the per-sample posterior group assignment probability [@jombart2008adegenet].
-This method decomposes the genetic data into principal components, and then uses a subset of these as the inputs for discriminant analysis, which attempts to minimize within-group variation and maximize among-group variation [@jombart2010discriminant].
+This method decomposed the genetic data into principal components, and then used a subset of these as the inputs for discriminant analysis, which attempts to minimize within-group variation and maximize among-group variation [@jombart2010discriminant].
 To avoid over-fitting the data, the optimal number of principal components was selected by using the *adegenet* function `xvalDapc()`. 
-This function implements a cross-validation procedure to iterate over an increasing number of principal components on a subset (90%) of the data, trying to find the minimum number of principal components that maximizes the rate of successful group reassignment.
-We additionally assessed genotypic diversity by assessing Stoddart and Taylor's index (*G*) [@stoddart1988genotypic], Shannon's index (*H*) [@shannon2001mathematical], Simpson's index ($\lambda$) [@simpson1949measurement], and evenness ($E_5$, *eMLG*) [@heck1975explicit; @hurlbert1971nonconcept; @pielou1975ecological; @grunwald2003analysis].
+This function implemented a cross-validation procedure to iterate over an increasing number of principal components on a subset (90%) of the data, trying to find the minimum number of principal components that maximized the rate of successful group reassignment.
+We additionally assessed genotypic diversity by calculating Stoddart and Taylor's index (*G*) [@stoddart1988genotypic], Shannon's index (*H*) [@shannon2001mathematical], Simpson's index ($\lambda$) [@simpson1949measurement], and evenness ($E_5$, *eMLG*) [@heck1975explicit; @hurlbert1971nonconcept; @pielou1975ecological; @grunwald2003analysis].
 To assess the potential for random mating, we tested for linkage disequilibrium with the index of association, $I_A$ and its standardized version, $\bar{r}_d$ using 999 permutations [@agapow2001indices; @smith1993how; @brown1980multilocus].
@@
 As many disparate geographic regions shared multilocus genotypes, we wanted to know if we could find any patterns in connectivity. 
+> SEE: avoid beginning a paragraph with a preposition because it forces the reader to read and remember the first half of the sentence before understanding it by reading the second half of the sentence.
+It's better to reorganize and start with the subject, such as "Connectivity between shared multilocus genotypes identified in disparate geographic regions was evaluated."
+
 To assess the connectivity of shared genotypes across regions, we first tabulated the multilocus genotypes shared between at least two populations (defined as states or countries) with the *poppr* function `mlg.crosspop()` [@kamvar2014poppr].
-From these data, we were able to construct a graph with populations as nodes and shared genotypes as edges (connections) between the nodes using the R packages *igraph* version 1.0.1 [@csardi2006igraph], *dplyr* version 0.7.1 [@dplyr], and *purrr* version 0.2.2.2 [@purrr].
+From these data, we constructed a graph with populations as nodes and shared genotypes as edges (connections) between nodes using the R packages *igraph* version 1.0.1 [@csardi2006igraph], *dplyr* version 0.7.1 [@dplyr], and *purrr* version 0.2.2.2 [@purrr].
 Each node was weighted by the fraction of shared MLGs.
-Each edge represents a single MLG, but because a single MLG could cross populations, that MLG may be represented by (n\*(n-1))/2 edges where n represents the number of populations it crosses.
-Edges were weighted by $1-P_{sex}$, where $P_{sex}$ is the probability of encountering the same genotype via two independent meiotic events [@arnaud2007standardizing; @parks1993study].
+Each edge represented a single MLG, but because a single MLG could cross populations, that MLG could be represented by (n\*(n-1))/2 edges where n represented the number of populations it crossed.
+Edges were weighted by $1-P_{sex}$, where $P_{sex}$ was the probability of encountering the same genotype via two independent meiotic events [@arnaud2007standardizing; @parks1993study].
 We then used the `cluster_optimal()` function from *igraph* to define communities [@csardi2006igraph].
 The graph was plotted using the R packages *ggplot2* version 2.2.1 [@ggplot2] and *ggraph* version 1.0.0 [@ggraph].
-To ensure that we were able to capture the same community signal, we additionally ran this analysis including the 5 polymorphic markers described above.
+To ensure that we captured the same community signal, we additionally performed this analysis including the 5 polymorphic markers described above.
 
@@
 The dbRDA was performed on clone-corrected isolates (N = 318) after grouping them by location, host, region, and year.
-The forward-backward selection process revealed severity, year, region, and host to be the optimal variables for the reduced model, accounting for 17.6% of the variation.
+The forward-backward selection process revealed aggressiveness, year, region, and host to be the optimal variables for the reduced model, accounting for 17.6% of the variation.
 ANOVA revealed that the reduced model was significant with an adjusted R^2^ of 0.032 (P = 0.001).

### 2017-07-11 doc/RMD/RDA-analysis.Rmd
update documents to shush
update RDA analysis to include RsquareAdjust

@@
 
-```{r, load_packages}
+```{r, load_packages, message = FALSE, warning = FALSE}
 library('tidyverse')
@@
 cap11cc$anova
+vegan::RsquareAdj(cap11cc)
 cap16cc
 cap16cc$anova
+vegan::RsquareAdj(cap16cc)
 ```

### 2017-07-11 doc/manuscript/manuscript.Rmd
Begin results section.

--- 2017-07-10 doc/manuscript/manuscript.Rmd
@@
 
+A total of 366 isolates were collected from 2003 to 2012 (except 2006 and 2011) from diseased dry bean plants in 11 states in the US, Australia, Mexico, and France (Table \@ref(tab:isolate-table)).
+Over 50% of the isolates came from four states, MI (62), ND (60), WA (59), NE (47). 
+Four regions had fewer than 10 isolates, Australia (6), WI (2), NY (1), ID (1).
+After clone-correction on the hierarchy of Region/Source/Host/Year, we were left with 318 isolates representing 165 MLGs \@ref(tab:iatab).
+Analysis of the genotype accumulation curve showed no clear plateau, indicating that we would likely obtain more multilocus genotypes if we were to collect more loci.
+The number of observed alleles per locus ranged from 2 to 10 with an average of 6.27 (Table \@ref(tab:locus-stats)).
+
+--------------------------------------------
+Locus  Range    No. alleles    *h*  Evenness 
+------ ------- ------------ ------ ---------
+5-2    318-324            4   0.45      0.62
+
+6-2    483-495            3   0.64      0.95
+
+7-2    158-174            7   0.73      0.76
+
+8-3    244-270            7   0.74      0.79
+
+9-2    360-382            9   0.35      0.41
+
+12-2   214-222            5   0.58      0.78
+
+17-3   342-363            7   0.55      0.53
+
+20-3   280-282            2   0.05      0.42
+
+55-4   153-216           10   0.72      0.66
+
+110-4  370-386            5   0.76      0.91
+
+114-4  339-416           10   0.83      0.80
+
+mean   153-495         6.27   0.58      0.69
+
+--------------------------------------------
+
+Table: (\#tab:locus-stats) Allelic diversity on full data set at loci used in this study. *h* = Nei's Gene Diversity [@nei1978estimation].
+
+## Variable Importance {-}
+
+The dbRDA was performed on clone-corrected isolates (N = 318) after grouping them by location, host, region, and year.
+The forward-backward selection process revealed severity, year, region, and host to be the optimal variables for the reduced model, accounting for 17.6% of the total variation.
+ANOVA revealed that the reduced model was significant with an adjusted R^2^ of 0.033 (P = 0.001).
+Assessment of the marginal effects showed that state, host, and year significantly explained genetic variation (P = 0.001), but aggressiveness was only mildly significant (P = 0.045). 
+Much of the variation appears to be driven by isolates from Mexico (Fig. \@ref(rda-biplot)) and 2005; it's worth noting that all Mexican isolates were collected in 2005.
+Variance partitioning of the independent variables indicated aggressiveness to be the least influential factor with 0.1% contributing to explaining the variation of molecular data whereas the combination of variables accounted for 3.3%.
+
+<!-- RDA figure -->
+
+![
+Biplot showing 10 most influential explanatory variables (arrows) overlayed on the first two eigenvectors of distance based redundancy analysis of *S. sclerotiorum* isolates. 
+The length of the arrows are directly proportional to the strength of the correlation between explanatory and molecular variables. 
+Points represent the 318 clone-corrected samples in ordination space. 
+](../../results/figures/publication/Figure7Z.pdf){#rda-biplot width=50%}
+
+<!-- end figure -->
+
 ## Genotypic Structure {-}
@@
 
-<!--  ZNK: I think figures 1-4 are a bit redundant with the tables in the text. It might be better to simply remove them. -->
-A total of 366 isolates were collected from 2003 to 2012 from diseased dry bean plants in 11 states in the US, Australia (2 mainland; 4 Tasmania), Mexico, and France **(Fig. 1)**.
-Most isolates came from MI (62), ND (60), WA (59), NE (47), and CO (42).
-Other regions in lesser representation were France (24), MX (18), OR (17), CA (16), MN (11), Australia (6), WI (2), ID (1), and NY (1).
-Four of the Australian isolates were collected from Tasmania while the other two came from Victoria state.
-There were no samples collected in 2006 and 2011.
-The highest number of samples were collected in 2005 (N = 63) and lowest in 2012 (N = 4).
-
 ```{r table3, echo = FALSE, warning = FALSE, message = FALSE, results = "asis"}
@@
 
-Pathogen isolates represented 26 commercial cultivars and breeding lines with some representing unknown dry bean hosts **(Fig. 2)**.
-We investigated the population association with host for cultivars where more than 10 isolates were obtained, which included G122, Beryl, Bunsi, Yellow, Pinto, and Merlot.
+There were a total of 165 multilocus genotypes in the isolate collection.
+The most abundant MLG represented 27 isolates (Table \@ref(tab:mlg-table)) across five regions (NE, MI, WA, CO, and ND).
+
 These six cultivars had a total of 238 *S. sclerotiorum* isolates (of 366 total isolates collected) belonging to 130 MLGs.
@@
 For instance, there were only 15 unique MLGs within the population of 68 *S. sclerotiorum* isolates obtained from the 'Beryl' cultivar.
-
-There were a total of 165 multilocus genotypes in the isolate collection.
-The most abundant MLG represented 27 isolates (Fig. \@ref(barplot) and Table 4) across five regions (NE, MI, WA, CO, and ND).
-Twenty isolates of this MLG were identified from ND.
-The genotypes that had either highest geographical distribution or found in high frequency are given in Table 4.
-MLG 163 was identified in NE, MI, CO and ND, while MLG 165 was identified in MN and all states aforementioned, except for ND.
-Isolates with MLG 15 were identified from populations in MI, WA, CO, and ND.
-The MLGs 163, 165 and 15 had 15, 7, and 4 individuals across these populations, respectively.
-The highest number of clones was recorded from ND, NE, and MI, which included 20, 10 and 7 isolates, respectively **(Table 4)**.
 
@@
 Mean separation of 10 most common MCGs showed MCG 44 is different from MCGs 9, 1, 45, 4, 53, and 49 while the aggressiveness levels of MCGs 2, 5, 3 failed to show any difference to MCG 44 or the other MCGs.
-
-The number of observed alleles per locus ranged anywhere from 2 to 10 (Table \@ref(tab:locus-stats)).
-All loci were informative at cut off level of 2/N where N was the total number of individuals.
-The genetic diversity indices of *S. sclerotiorum* populations from different states of the USA, France, Mexico and Australia based on SSR analysis are given in Table 5.
-The 165 unique MLGs identified among all 366 isolates are given in Supplementary Table 1.
-Gene diversity (*h*) of populations from different regions ranged from 0.259 to 0.596 **(Table 5)**.
-Evenness of genotypes of all regions was high, with the exception of ND, where a single genotype was represented by 20 isolates.
-There was one private allele in the WA population (204 bp allele of 55-4 locus) and two private alleles in the ND population (two 366 bp alleles in 9-2 locus and two 220 bp alleles in 12-2 locus).
-
---------------------------------------------
-Locus  Range    No. alleles    *h*  Evenness 
------- ------- ------------ ------ ---------
-5-2    318-324            4   0.45      0.62
-
-6-2    483-495            3   0.64      0.95
-
-7-2    158-174            7   0.73      0.76
-
-8-3    244-270            7   0.74      0.79
-
-9-2    360-382            9   0.35      0.41
-
-12-2   214-222            5   0.58      0.78
-
-17-3   342-363            7   0.55      0.53
-
-20-3   280-282            2   0.05      0.42
-
-55-4   153-216           10   0.72      0.66
-
-110-4  370-386            5   0.76      0.91
-
-114-4  339-416           10   0.83      0.80
-
-mean   153-495         6.27   0.58      0.69
-
---------------------------------------------
-
-Table: (\#tab:locus-stats) Allelic diversity on full data set at loci used in this study. *h* = Nei's Gene Diversity [@nei1978estimation].
-
-## Variable Importance {-}
-
-The dbRDA was performed on clone-corrected isolates (N = 318) after grouping them by location, host, region, and year.
-The forward-backward selection process revealed severity, year, region, and host to be the optimal variables for the reduced model, accounting for 17.6% of the variation.
-ANOVA revealed that the reduced model was significant with an adjusted R^2^ of 0.032 (P = 0.001).
-The marginal effects of state, host, and year were also significant (P = 0.001), but aggressiveness was not significant (P = 0.045).
-Much of the variation appears to be driven by isolates from Mexico (Fig. \@ref(rda-biplot)) and 2005, but it's worth noting that all Mexican isolates were collected in 2005.
-Variance partitioning of the independent variables indicated aggressiveness to be the least influential factor with 0.1% contributing to explaining the variation of molecular data whereas the combination of variables accounted for 3.3%.
-
-<!-- RDA figure -->
-
-![
-Biplot showing 10 most influential explanatory variables (arrows) overlayed on the first two eigenvectors of distance based redundancy analysis of *S. sclerotiorum* isolates. 
-The length of the arrows are directly proportional to the strength of the correlation between explanatory and molecular variables. 
-Points represent the 318 clone-corrected samples in ordination space. 
-](../../results/figures/publication/Figure7Z.pdf){#rda-biplot width=50%}
-
-<!-- end figure -->
 

### 2017-07-11 doc/manuscript/manuscript.Rmd
Used otto-hanson citation for resistance

--- 2017-07-10 doc/manuscript/manuscript.Rmd
@@
 To date, no commercial white mold-resistant dry bean cultivars exist. 
-Field trials in some regions of the US have shown the G122 cultivar is potentially resistant while the Beryl cultivar is susceptible (pers. comm. with J. Steadman).
-<!-- SEE: Evaluating cultivars, there should be some BIC reports that we can cite; personal communication is kind of eh. -->
+Field trials in some regions of the US have shown the G122 cultivar is partially resistant while the Beryl cultivar is susceptible [@otto-hanson2011variation].
+
 Understanding genetic relationships and reproduction behavior of *S.

### 2017-07-11 doc/manuscript/manuscript.Rmd
Address Sydney's comments

@@
 To date, no commercial white mold-resistant dry bean cultivars exist. 
-Field trials in some regions of the US have shown the G122 cultivar is partially resistant while the Beryl cultivar is susceptible [@otto-hanson2011variation].
+Field trials in some regions of the US have shown the G122 cultivar is partially resistant while the Beryl cultivar is susceptible [@otto-hanson2011variation; @steadman2004identification].
 
@@
 Sclerotia of *S.
-sclerotiorum* were collected over several years from grower fields and/or dry bean nurseries in 11 states of the United States of America, Mexico, France, and Australia (Table \@ref(tab:isolate-table)).
-The dry bean nurseries in different regions are indicated by wmn (white mold nursery) or mtc (Mitchell) and all other isolates came from commercial fields (Table \@ref(tab:isolate-table)).
+sclerotiorum* were collected over several years from grower fields and/or white mold nurseries in 11 states of the United States of America, Mexico, France, and Australia (Table \@ref(tab:isolate-table)).
+The white mold nurseries in different regions are indicated by wmn (white mold nursery) or mtc (Mitchell) and all other isolates came from commercial fields (Table \@ref(tab:isolate-table)).
+
 > SEE: I was not sure what "dry bean nurseries" were because I'm use to seeing them referred to as the white mold nurseries.
@@
 
-The dry bean nurseries were 5m x 10m in size and maintained without applying fungicides to observe natural incidence of white mold.
+The white mold nurseries were 5m x 10m in size and maintained without applying fungicides to observe natural incidence of white mold.
 The nursery plots were incorporated with a basal dressing of N:P:K = 1:3:2 and side dressing of 0:3:2 during the growing season.
@@
 Since MCG are considered mutually exclusive, pairing was not required between all isolates in the study.
+
 > SEE: Is that correct?  Perhaps consult with Becky or Rachana and then elaborate upon the sentence above.
@@
 Aggressiveness of each isolate was assessed using a straw test that rated necrotic lesion size on a scale from 1--9 with increasing aggressiveness [@petzoldt1996straw; @teran2006modified].
+
 > SEE: Rating is not clear from the single sentence above.  On what plant host was it evaluated?  In the greenhouse?  What is a rating of 1 compared to a rating of 9 on the straw test?
@@
 DNA was purified from 100 mg mycelium collected from an actively growing culture of each isolate using a Qiagen DNeasy Plant Mini Kit (Qiagen Inc., Valencia, CA) according to the manufacturer's instructions.
+
 > SEE: More details on how tissue was disrupted may be needed here.
@@
 We genotyped *S. sclerotiorum* isolates using 16 microsatellite primer pairs developed previously [@sirjusingh2001characterisation].
-Polymerase chain reaction (PCR) was carried out as described using primers labeled with FAM fluorophore, and resulting PCR products were first resolved in a 1.5% agarose gel stained with ethidium bromide to ensure product was within the expected size range.
-> SEE: Pretty rare to see PCR spelled out in full these days. I suggest just using the abbreviation and if PeerJ asks us to spell it out, we will.
+PCR was carried out as described using primers labeled with FAM fluorophore, and resulting PCR products were first resolved in a 1.5% agarose gel stained with ethidium bromide to ensure product was within the expected size range.
 
 Capillary electrophoresis (fragment analysis) of amplicons, with size standard GeneScan™ 500 LIZ®, was performed using an ABI 3730 genetic analyzer (Life Technologies Corporation, Carlsbad, CA) at the Michigan State University Genomic Sequencing Center.
-
-> ZNK: how were the data received and formatted?
-
-> SEE: data were received as .fsa files, which are fasta-type files.
-I have the raw data if you want them.
-There is, apparently, an R-package called Fragman used for reading the files and can be used for scoring.
-Anyway, in this case, I believe Rachana used PeakScanner (free software) to read the files and she recorded the fragment sizes manually (ie.
-read the chromatogram and wrote down in Excel what size the peak was).
+Alleles were binned using PeakScanner version 1.0 (Life Technologies Corporation, Carlsbad, CA) and recorded manually in an excel sheet.
 
@@
 
-Of the 16 microsatellite loci, five included compound repeats, which made it difficult to accurately/confidently bin alleles into fragment sizes expected for each locus based on the described repeat motif.
-> SEE: I would not say allele scoring was difficult, because scoring an allele that is different than the size expected is really the same difficulty level as one that is at the expected size.
-It's the accurate binning of alleles into the expected fragment size that is difficult.
-
+Of the 16 microsatellite loci, five included compound repeats, which made it challenging to accurately/confidently bin alleles into fragment sizes expected for each locus based on the described repeat motif.
 Loci with compound repeats were removed for the reported statistics, but were processed and included in data analyses to ensure the integrity of the results with 11 loci as compared to 16.
 We assessed the power of our 11 markers by generating a genotype accumulation curve in the R package *poppr* version 2.4.1, looking for evidence of saturation, which would indicate that loci were sufficiently sampled to adequately represent the full set of genotypes [@arnaud2007standardizing; @kamvar2015novel].
-To control for non-independent samples (i.e. samples from the same location, region, year, and plant), data were clone-corrected for subsequent analyses.
-> SEE: it is not clear what non-independent samples are.  Although your description is accurate and concise, a more general description might enable the reader to better understand what is meant here.
+To avoid including samples that were potentially collected from the same plant (which increases the probability of collecting the same sample more than once), data were clone-corrected on the region/source/host/year hierarchy for subsequent analyses.
 
@@
 To minimize complications due to small sample sizes, we chose the top 10 MCGs, representing 57% of the isolates collected.
-Between-group differences were assessed using Tukey's HSD test using the `HSD.test()` function in the package *agricolae* version 1.2.4 [@mendiburu2015agricolae].
-> SEE: It's not clear what groups were used to compare aggressiveness in the ANOVA.  What alpha-level was used for the cut-off?
+Between-MCG differences were assessed using Tukey's HSD test with an alpha of 0.05 using the `HSD.test()` function in the package *agricolae* version 1.2.4 [@mendiburu2015agricolae].
 
@@
 ## Genetic Diversity {-}
-> SEE: deleting this header?
 
@@
 Source in this case was defined as a population belonging to a wmn or not and region defined as state.
-> SEE: is mtc included with "wmn"? At the beginning of the methods, I thought both were considered dry bean nurseries.
 
 To identify specific regions that were more differentiated than others, we used discriminant analysis of principal components (DAPC) as implemented in *adegenet* version 2.1.0, assessing the per-sample posterior group assignment probability [@jombart2008adegenet].
-This method decomposed the genetic data into principal components, and then used a subset of these as the inputs for discriminant analysis, which attempts to minimize within-group variation and maximize among-group variation [@jombart2010discriminant].
+This method decomposes the genetic data into principal components, and then uses a subset of these as the inputs for discriminant analysis, which attempts to minimize within-group variation and maximize among-group variation [@jombart2010discriminant].
 To avoid over-fitting the data, the optimal number of principal components was selected by using the *adegenet* function `xvalDapc()`. 
@@
 
-As many disparate geographic regions shared multilocus genotypes, we wanted to know if we could find any patterns in connectivity. 
-> SEE: avoid beginning a paragraph with a preposition because it forces the reader to read and remember the first half of the sentence before understanding it by reading the second half of the sentence.
-It's better to reorganize and start with the subject, such as "Connectivity between shared multilocus genotypes identified in disparate geographic regions was evaluated."
-
-To assess the connectivity of shared genotypes across regions, we first tabulated the multilocus genotypes shared between at least two populations (defined as states or countries) with the *poppr* function `mlg.crosspop()` [@kamvar2014poppr].
+We wanted to evaluate patterns of connectivity between shared multilocus genotypes across geographic regions.
+We first tabulated the multilocus genotypes shared between at least two populations (defined as states or countries) with the *poppr* function `mlg.crosspop()` [@kamvar2014poppr].
 From these data, we constructed a graph with populations as nodes and shared genotypes as edges (connections) between nodes using the R packages *igraph* version 1.0.1 [@csardi2006igraph], *dplyr* version 0.7.1 [@dplyr], and *purrr* version 0.2.2.2 [@purrr].
@@
 Each edge represented a single MLG, but because a single MLG could cross populations, that MLG could be represented by (n\*(n-1))/2 edges where n represented the number of populations it crossed.
-Edges were weighted by $1-P_{sex}$, where $P_{sex}$ was the probability of encountering the same genotype via two independent meiotic events [@arnaud2007standardizing; @parks1993study].
+Edges were weighted by $1-P_{sex}$, where $P_{sex}$ is the probability of encountering the same genotype via two independent meiotic events [@arnaud2007standardizing; @parks1993study].
 We then used the `cluster_optimal()` function from *igraph* to define communities [@csardi2006igraph].
@@
 The same study reported increased ascospores discharge at 20-25$^{\circ}$ C compared to 15$^{\circ}$ C and the survival of *S. sclerotiorum* ascospores was reduced at high temperature and RH.
-Our samples were collected from 2003 to 2012 and except for the dry bean nurseries, came from different fields.
+Our samples were collected from 2003 to 2012 and except for the white mold nurseries, came from different fields.
 Different agro-ecological conditions prevailed during sampling seasons may have resulted in different rates of sexual outcrossing of the pathogen and hence reduced the chance of extracting a common genotype.

### 2017-07-11 doc/manuscript/manuscript.Rmd
remove past tense from method description.

@@
 To avoid over-fitting the data, the optimal number of principal components was selected by using the *adegenet* function `xvalDapc()`. 
-This function implemented a cross-validation procedure to iterate over an increasing number of principal components on a subset (90%) of the data, trying to find the minimum number of principal components that maximized the rate of successful group reassignment.
+This function implements a cross-validation procedure to iterate over an increasing number of principal components on a subset (90%) of the data, trying to find the minimum number of principal components that maximizes the rate of successful group reassignment.
 We additionally assessed genotypic diversity by calculating Stoddart and Taylor's index (*G*) [@stoddart1988genotypic], Shannon's index (*H*) [@shannon2001mathematical], Simpson's index ($\lambda$) [@simpson1949measurement], and evenness ($E_5$, *eMLG*) [@heck1975explicit; @hurlbert1971nonconcept; @pielou1975ecological; @grunwald2003analysis].

### 2017-07-11 doc/manuscript/manuscript.Rmd

@@
 Mean separation of 10 most common MCGs showed MCG 44 is different from MCGs 9, 1, 45, 4, 53, and 49 while the aggressiveness levels of MCGs 2, 5, 3 failed to show any difference to MCG 44 or the other MCGs.
-
-The number of observed alleles per locus ranged anywhere from 2 to 10 (Table \@ref(tab:locus-stats)).
-All loci were informative at cut off level of 2/N where N was the total number of individuals.
-The genetic diversity indices of *S. sclerotiorum* populations from different states of the USA, France, Mexico and Australia based on SSR analysis are given in Table 5.
-The 165 unique MLGs identified among all 366 isolates are given in Supplementary Table 1.
-Gene diversity (*h*) of populations from different regions ranged from 0.259 to 0.596 **(Table 5)**.
-Evenness of genotypes of all regions was high, with the exception of ND, where a single genotype was represented by 20 isolates.
-There was one private allele in the WA population (204 bp allele of 55-4 locus) and two private alleles in the ND population (two 366 bp alleles in 9-2 locus and two 220 bp alleles in 12-2 locus).
-
---------------------------------------------
-Locus  Range    No. alleles    *h*  Evenness 
------- ------- ------------ ------ ---------
-5-2    318-324            4   0.45      0.62
-
-6-2    483-495            3   0.64      0.95
-
-7-2    158-174            7   0.73      0.76
-
-8-3    244-270            7   0.74      0.79
-
-9-2    360-382            9   0.35      0.41
-
-12-2   214-222            5   0.58      0.78
-
-17-3   342-363            7   0.55      0.53
-
-20-3   280-282            2   0.05      0.42
-
-55-4   153-216           10   0.72      0.66
-
-110-4  370-386            5   0.76      0.91
-
-114-4  339-416           10   0.83      0.80
-
-mean   153-495         6.27   0.58      0.69
-
---------------------------------------------
-
-Table: (\#tab:locus-stats) Allelic diversity on full data set at loci used in this study. *h* = Nei's Gene Diversity [@nei1978estimation].
-
-## Variable Importance {-}
-
-The dbRDA was performed on clone-corrected isolates (N = 318) after grouping them by location, host, region, and year.
-The forward-backward selection process revealed aggressiveness, year, region, and host to be the optimal variables for the reduced model, accounting for 17.6% of the variation.
-ANOVA revealed that the reduced model was significant with an adjusted R^2^ of 0.032 (P = 0.001).
-The marginal effects of state, host, and year were also significant (P = 0.001), but aggressiveness was not significant (P = 0.045).
-Much of the variation appears to be driven by isolates from Mexico (Fig. \@ref(rda-biplot)) and 2005, but it's worth noting that all Mexican isolates were collected in 2005.
-Variance partitioning of the independent variables indicated aggressiveness to be the least influential factor with 0.1% contributing to explaining the variation of molecular data whereas the combination of variables accounted for 3.3%.
-
-<!-- RDA figure -->
-
-![
-Biplot showing 10 most influential explanatory variables (arrows) overlayed on the first two eigenvectors of distance based redundancy analysis of *S. sclerotiorum* isolates. 
-The length of the arrows are directly proportional to the strength of the correlation between explanatory and molecular variables. 
-Points represent the 318 clone-corrected samples in ordination space. 
-](../../results/figures/publication/Figure7Z.pdf){#rda-biplot width=50%}
-
-<!-- end figure -->
 

### 2017-07-11 doc/RMD/mlg-mcg.Rmd
update MCG barplot

@@
 
-```{r barplots, fig.width = 5, fig.height = 10}
+```{r barplots, fig.width = 10, fig.height = 5}
 mcg_graph <- mcgs %>% 
@@
   gather(type, count, MLGs, Samples, -Evenness) %>%
-  arrange(desc(type), count) %>%
+  arrange(desc(type), desc(count)) %>%
   rename(Type = type) %>%
@@
   theme_minimal(base_size = 16, base_family = "Helvetica") +
-  coord_flip() +
-  theme(panel.grid.major.y = element_blank()) +
-  theme(legend.box.just = "left") +
-  theme(legend.box.margin = unit(c(0, 0.55, 11, 0), "lines")) +
-  theme(legend.box.background = element_rect(fill = NA, color = "black")) +
-  # theme(legend.box = "horizontal") +
-  theme(aspect.ratio = 2) +
+  # coord_flip() +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
+  theme(panel.grid.major.x = element_blank()) +
+  theme(panel.grid.major.y = element_line(color = "grey50")) +
+  theme(panel.grid.minor.y = element_line(color = "grey70")) +
+  theme(legend.box.just = "top") +
+  theme(legend.box.margin = unit(c(0, 6.55, 0, 0), "lines")) +
+  theme(legend.box.background = element_rect(fill = "white", color = "black")) +
+  theme(legend.box = "horizontal") +
+  theme(aspect.ratio = 1/2) +
   labs(list(
@@
 
-vp1 <- grid::viewport(width = 0.25, height = 0.4, x = 0.795, y = 0.64, just = c("center", "top"))
-mcg_graph <- mcg_graph + theme(legend.position = c(.789, .585))
+vp1 <- grid::viewport(width = 0.25, height = 0.4, x = 1, y = 0.975, just = c("right", "top"))
+mcg_graph <- mcg_graph + theme(legend.position = c(.763, .75))
 
-pdf(file.path(PROJHOME, "results/figures/publication/Figure1Z.pdf"), width = 5, height = 10)
+print(mcg_graph)
+print(even, vp = vp1)
+
+if (!interactive()){
+  pdf(file.path(PROJHOME, "results/figures/publication/Figure1Z.pdf"), width = 10, height = 5)
   print(mcg_graph)
@@
   dev.off()
-
-print(mcg_graph)
-print(even, vp = vp1)
+}
 ```

### 2017-07-11 doc/RMD/mlg-mcg.Rmd
add counts of MLG/MCG pairs

@@
 
+Of interest is accounting for the communities that have a single MCG/MLG pair.
+We can detect these by filtering for communities that only have two members:
+
+```{r}
+singles <- the_communities %>% 
+  count(community) %>% 
+  arrange(desc(n)) %>% 
+  filter(n == 2) %>% 
+  inner_join(the_communities, by = "community") %>% 
+  mutate(type = case_when(grepl("MLG", vertex) ~ "MLG", TRUE ~ "MCG")) %>% 
+  spread(type, vertex) %>%
+  select(-n, -comm)
+# assessing that this truely contains the singletons
+nrow(inner_join(singles, the_communities, by = c("MCG" = "vertex")))
+nrow(inner_join(singles, the_communities, by = c("MLG" = "vertex")))
+
+knitr::kable(singles)
+```
+
 ### Interactive Visualizations

### 2017-07-11 doc/RMD/MCG-virulence.Rmd
export table from MCG-virulence analysis

@@
 summary(ANOVA)
-plot(TukeyHSD(ANOVA), las = 2)
-agricolae::HSD.test(ANOVA, "MCG")$groups
+plot(TukeyHSD(ANOVA, conf.level = 0.95), las = 2)
+grps <- agricolae::HSD.test(ANOVA, "MCG", alpha = 0.05)$groups
+readr::write_csv(grps, path = "results/tables/MCG-aggressiveness.csv", col_names = TRUE)
+
+grps %>% 
+  dplyr::rename(MCG = trt) %>%
+  dplyr::rename(Group = M) %>%
+  dplyr::rename(`Mean Aggressiveness` = means) %>%
+  huxtable::as_huxtable(add_colnames = TRUE) %>%
+  huxtable::set_number_format(row = huxtable::everywhere, col = 1, 0) %>% 
+  huxtable::set_col_width(c(0.05, 0.3, 0.08)) %>%
+  huxtable::set_align(row = huxtable::everywhere, col = 2, "right") %>%
+  huxtable::print_md(max_width = 31)
 ```

### 2017-07-11 doc/manuscript/manuscript.Rmd
Add more information to results

@@
 The white mold nurseries in different regions are indicated by wmn (white mold nursery) or mtc (Mitchell) and all other isolates came from commercial fields (Table \@ref(tab:isolate-table)).
-
-> SEE: I was not sure what "dry bean nurseries" were because I'm use to seeing them referred to as the white mold nurseries.
-I think if the introduction uses one term for it, this would be acceptable.
-That or consider providing a description of the white mold nurseries in the methods section.
-This now reminds me..
-there is a publication by Lindsey Otto-Hanson and Steadman that gave a good description of these.
-In fact, she called them multisite screening locations.
-Here's the link: http://apsjournals.apsnet.org/doi/abs/10.1094/PDIS-11-10-0865
 
@@
 
-## Variable Importance {-}
+## Phenotypic variable assessment {-}
+
+### Variable Importance {-}
 
@@
 Variance partitioning of the independent variables indicated aggressiveness to be the least influential factor with 0.1% contributing to explaining the variation of molecular data whereas the combination of variables accounted for 3.3%.
+
+### Mycelial Compatibility Groups {-}
+
+We observed 87 mycelial compatibility groups (MCG), the most abundant of which represented 73 isolates over 37 multilocus genotypes (MLG) (Fig. \@ref(barplot). 
+Our assessment of aggressiveness association with MCGs showed a significant effect (P < 0.001) indicating that MCG 44 is, on average, more aggressive than other MCGs (Table \@ref(tab:mcg-aggressiveness)).
 
@@
 
-<!-- MCG abundance figure -->
-
-![
-Barplot showing abundance of each Mycelial Compatibility Group (MCG) in descending order.
-Singletons (46) have been removed for clarity.
-White bars represent sample counts and grey bars represent counts of unique multilocus genotypes (MLG).
-The transparency associated with the grey bars represent the evenness of the distribution of the MLGs within a given MCG as shown in the inset histograms.
-](../../results/figures/publication/Figure1Z.pdf){#barplot width=50%}
-
-<!-- end figure -->
+We observed a total of 165 multilocus genotypes (MLG) (215 with 16 loci).
+The most abundant MLG represented 27 isolates (Table \@ref(tab:mlg-table)) across five regions (NE, MI, WA, CO, and ND). 
+The genotypes were relatively evenly distributed within regions with moderate to high diversity (Table \@ref(tab:iatab)). 
+Of the 165 MLGs, 76 (46%) were found across all geographic regions except WI (2), ID (1), and Mexico (18) (Fig. \@ref(community-graph)).
+Our graph-based approach to correlating MLGs with MCGs revealed a large and complex network (Fig. \@ref(biggraph)). 
+Community analysis showed 51 communities, 23 of which consisted of a single MLG/MCG pair. 
+The three communities with the most members contained 8 of the ten most abundant MCGs. 
+Overall, however, there was no clear pattern to the association between MLG and MCGs.
 
@@
 
-There were a total of 165 multilocus genotypes in the isolate collection.
-The most abundant MLG represented 27 isolates (Table \@ref(tab:mlg-table)) across five regions (NE, MI, WA, CO, and ND).
+<!-- MCG abundance figure -->
+
+![
+Barplot showing abundance of each Mycelial Compatibility Group (MCG) in descending order.
+Singletons (46) have been removed for clarity, leaving 41 MCGs.
+White bars represent sample counts and grey bars represent counts of unique multilocus genotypes (MLG).
+The transparency associated with the grey bars represent the evenness of the distribution of the MLGs within a given MCG as shown in the inset histograms.
+](../../results/figures/publication/Figure1Z.pdf){#barplot width=100%}
+
+<!-- end figure -->
 
@@
 For instance, there were only 15 unique MLGs within the population of 68 *S. sclerotiorum* isolates obtained from the 'Beryl' cultivar.
-
-The associations between MCGs and MLGs are complicated (Fig. \@ref(biggraph))
 
@@
 \@ref(community-graph)): A coastal community (CA, OR, WA, and NY), a mid-west community (CO, NE, ND, MI), and an international community (MN, France, Australia).
-Analysis with 16 loci resulted in the removal of the NY node because it no longer shared a genotype with OR, but the same overall community structure was present with a modularity of 0.2 (Supplementary Figure S1Z).
+Analysis with 16 loci resulted in the removal of the NY node because it no longer shared a genotype with OR, but the same overall community structure was present with a modularity of 0.2 (Fig.
+\@ref(community-graph16)).
 The international community appears to be driven by MLG 4, which is shared between all three populations and has a $P_{sex}$ value of 2.87 x 10^-5^, in contrast to the abundant MLG 25, which has a $P_{sex}$ value of 0.0168.
-
-## Aggressiveness {-}
-
-The MCGs with the highest frequency were MCG 5 and MCG 44 representing 73 and 44 isolates, respectively **(Fig. 4)**. <!-- Maybe add aggressiveness information here. -->
-MCGs that were unique to a region are shown in Fig.
-5 while MCGs that represented at least 2 or more regions are in Fig 6.
-20 MCGs were common in two or more regions and the rest were distributed uniquely in different regions.
-ID, MN, NE, NY, and WI did not have any unique/isolated MCGs.
-The aggressiveness of isolates ranged from 1.4 to 7.9 with a mean of 5.01.
-There were six isolates that fell into upper outlier category (aggressiveness > 7.6) and one isolate in the lower outlier range (aggressiveness < 2.3) (box plot results not shown).
-The mean aggressiveness of the 10 most abundant MCGs ranged from 4.6 (MCG 49) to 6.02 (MCG 44) (see Fig 4 for 10 most abundant MCGs).
-ANOVA of these 10 MCGs indicated significant difference for aggressiveness levels at P <0.0001.
-Mean separation of 10 most common MCGs showed MCG 44 is different from MCGs 9, 1, 45, 4, 53, and 49 while the aggressiveness levels of MCGs 2, 5, 3 failed to show any difference to MCG 44 or the other MCGs.
 
@@
 
+```{r table2, echo = FALSE, warning = FALSE, message = FALSE, results = "asis"}
+suppressPackageStartupMessages(library("tidyverse"))
+read_csv("../../results/tables/MCG-aggressiveness.csv", 
+         na = "NA", # spaces read in as spaces
+         col_types = cols(.default = col_character())) %>%
+  dplyr::rename(MCG = trt) %>%
+  dplyr::rename(Group = M) %>%
+  dplyr::rename(`Mean Aggressiveness` = means) %>%
+  huxtable::as_huxtable(add_colnames = TRUE) %>%
+  huxtable::set_number_format(row = huxtable::everywhere, col = 1, 0) %>% 
+  huxtable::set_col_width(c(0.05, 0.3, 0.08)) %>%
+  huxtable::set_align(row = huxtable::everywhere, col = 2, "right") %>%
+  huxtable::print_md(max_width = 31)
+```
+
+Table: (\#tab:mcg-aggressiveness) Mean aggressiveness ratings for the ten most abundant mycelial compatibility groups (MCG) with groupings according to 95% family-wise confidence interval.
+
+\newpage
+
 ----------------------------------------------------------------------------
@@
 P-value symbols: \* = 0.001, ~ = $\leq 0.05$.
+
+<!-- community graph figure -->
+
+![
+Network of populations (nodes/circles) and their shared multilocus haplotypes (MLG) (edges/lines) genotyped over 16 loci. 
+Each node is labeled with **name (number of MLGs shared/number of MLGs total).** 
+The shade and area of the nodes are proportional to the number of unique MLGs within the node and the inner nodes are proportional to the number of private MLGs to the region. 
+Each edge represents a single MLG where its thickness represents the number of populations that share the MLG and the shade represents the value of $P_{sex}$, or the probability of encountering that MLG from two independent meiotic events.
+](../../results/figures/publication/FigureS1Z.pdf){#community-graph16 width=100%}
+
+<!-- end fig -->
+
+\newpage

### 2017-07-11 doc/manuscript/manuscript.Rmd
overhaul results section

@@
 Assessment of the marginal effects showed that state, host, and year significantly explained genetic variation (P = 0.001), but aggressiveness was only mildly significant (P = 0.045). 
-Much of the variation appears to be driven by isolates from Mexico (Fig. \@ref(rda-biplot)) and 2005; it's worth noting that all Mexican isolates were collected in 2005.
+Much of the variation appears to be driven by isolates from Mexico and 2005 (Fig. \@ref(rda-biplot)); it's worth noting that all Mexican isolates were collected in 2005.
 Variance partitioning of the independent variables indicated aggressiveness to be the least influential factor with 0.1% contributing to explaining the variation of molecular data whereas the combination of variables accounted for 3.3%.
@@
 
-We observed 87 mycelial compatibility groups (MCG), the most abundant of which represented 73 isolates over 37 multilocus genotypes (MLG) (Fig. \@ref(barplot). 
+We observed 87 mycelial compatibility groups (MCG), the most abundant of which represented 73 isolates over 37 multilocus genotypes (MLG) (Fig. \@ref(barplot)). 
 Our assessment of aggressiveness association with MCGs showed a significant effect (P < 0.001) indicating that MCG 44 is, on average, more aggressive than other MCGs (Table \@ref(tab:mcg-aggressiveness)).
@@
 
+<!-- MCG abundance figure -->
+
+![
+Barplot showing abundance of each Mycelial Compatibility Group (MCG) in descending order.
+Singletons (46) have been removed for clarity, leaving 41 MCGs.
+White bars represent sample counts and grey bars represent counts of unique multilocus genotypes (MLG).
+The transparency associated with the grey bars represent the evenness of the distribution of the MLGs within a given MCG as shown in the inset histograms.
+](../../results/figures/publication/Figure1Z.pdf){#barplot width=100%}
+
+<!-- end figure -->
+
 ## Genotypic Structure {-}
@@
 Of the 165 MLGs, 76 (46%) were found across all geographic regions except WI (2), ID (1), and Mexico (18) (Fig. \@ref(community-graph)).
-Our graph-based approach to correlating MLGs with MCGs revealed a large and complex network (Fig. \@ref(biggraph)). 
-Community analysis showed 51 communities, 23 of which consisted of a single MLG/MCG pair. 
-The three communities with the most members contained 8 of the ten most abundant MCGs. 
-Overall, however, there was no clear pattern to the association between MLG and MCGs.
+Community analysis of the MLGs shared between populations revealed 4 communities with a modularity of 0.17 (Fig.
+\@ref(community-graph)): A coastal community (CA, OR, WA, and NY), a mid-west community (CO, NE, ND, MI), and an international community (MN, France, Australia).
+Analysis with 16 loci resulted in the removal of the NY node because it no longer shared a genotype with OR, but the same overall community structure was present with a modularity of 0.2 (Fig.
+\@ref(community-graph16)).
+The international community appears to be driven by MLG 4, which is shared between all three populations and has a $P_{sex}$ value of 2.87 x 10^-5^, in contrast to the abundant MLG 25, which has a $P_{sex}$ value of 0.0168.
 
@@
 
-<!-- MCG abundance figure -->
+<!-- community graph figure -->
 
 ![
-Barplot showing abundance of each Mycelial Compatibility Group (MCG) in descending order.
-Singletons (46) have been removed for clarity, leaving 41 MCGs.
-White bars represent sample counts and grey bars represent counts of unique multilocus genotypes (MLG).
-The transparency associated with the grey bars represent the evenness of the distribution of the MLGs within a given MCG as shown in the inset histograms.
-](../../results/figures/publication/Figure1Z.pdf){#barplot width=100%}
+Network of populations (nodes/circles) and their shared multilocus haplotypes (MLG) (edges/lines) genotyped over 11 loci. 
+Each node is labeled with **name (number of MLGs shared/number of MLGs total).** 
+The shade and area of the nodes are proportional to the number of unique MLGs within the node and the inner nodes are proportional to the number of private MLGs to the region. 
+Each edge represents a single MLG where its thickness represents the number of populations that share the MLG and the shade represents the value of $P_{sex}$, or the probability of encountering that MLG from two independent meiotic events.
+](../../results/figures/publication/Figure3Z.pdf){#community-graph width=100%}
 
-<!-- end figure -->
+<!-- end fig -->
 
-These six cultivars had a total of 238 *S. sclerotiorum* isolates (of 366 total isolates collected) belonging to 130 MLGs.
-Of these MLGs, 53 were singletons, while the rest were obtained from more than one host cultivar; only one private allele on locus 55-4 was identified, belonging to the *S.
-sclerotiorum* population from the cultivar 'Bunsi'.
-The five most abundant MLGs with associated MCGs are shown in Table \@ref(tab:mlg-table).
-The majority of MLGs reported were found on multiple host cultivars and only a few unique to a cultivar.
-For instance, there were only 15 unique MLGs within the population of 68 *S. sclerotiorum* isolates obtained from the 'Beryl' cultivar.
+Our network-based approach to correlating MLGs with MCGs revealed a large and complex network (Fig. \@ref(biggraph)). 
+Community analysis showed 51 communities, 23 of which consisted of a single MLG/MCG pair. 
+The three communities with the most members contained 8 of the ten most abundant MCGs. 
+Overall, however, there was no clear pattern to the association between MLG and MCGs.
 
@@
 
-<!-- community graph figure -->
+## Population Differentiation {-}
 
-![
-Network of populations (nodes/circles) and their shared multilocus haplotypes (MLG) (edges/lines) genotyped over 11 loci. 
-Each node is labeled with **name (number of MLGs shared/number of MLGs total).** 
-The shade and area of the nodes are proportional to the number of unique MLGs within the node and the inner nodes are proportional to the number of private MLGs to the region. 
-Each edge represents a single MLG where its thickness represents the number of populations that share the MLG and the shade represents the value of $P_{sex}$, or the probability of encountering that MLG from two independent meiotic events.
-](../../results/figures/publication/Figure3Z.pdf){#community-graph width=100%}
+### Analysis of Molecular Variance
 
-<!-- end fig -->
-
-Community analysis of the MLGs shared between populations revealed 4 communities with a modularity of 0.17 (Fig.
-\@ref(community-graph)): A coastal community (CA, OR, WA, and NY), a mid-west community (CO, NE, ND, MI), and an international community (MN, France, Australia).
-Analysis with 16 loci resulted in the removal of the NY node because it no longer shared a genotype with OR, but the same overall community structure was present with a modularity of 0.2 (Fig.
-\@ref(community-graph16)).
-The international community appears to be driven by MLG 4, which is shared between all three populations and has a $P_{sex}$ value of 2.87 x 10^-5^, in contrast to the abundant MLG 25, which has a $P_{sex}$ value of 0.0168.
-
-## DAPC {-}
-
-<!-- DAPC figure -->
-
-![
-Discriminant Analysis on regions showing that Mexico is differentiated from other populations. 
-**(a)** Mean population assignment (horizontal axis) probability (vertical axis) for all populations with N > 10 (facets).
-Numbers next to source populations indicate population size.
-All values sum to one.
-**(b)** Scatter plot of first two components from DAPC.
-Points represent observed individuals connected to the population centroids with ellipses representing a 66% confidence interval for a normal distribution. 
-The center of each component is represented as grid lines.
-](../../results/figures/publication/DAPC.pdf){#DAPC width=100%}
-
-<!-- end figure -->
-
-DAPC analysis for different regions and hosts are depicted in Fig. 8.
-These isolates were clone-corrected at location/host/region/year level prior to the analysis (N =318).
-Accordingly, MX (N = 15) and CA (N =18) populations grouped separately from all other populations with MX isolates showing the highest deviation (Fig. 8A).
-Comparison of genotypes by region showed all of the nine MLGs in the Mexican population were unique, with no genotypes crossing over to other populations.
-MLGs of Californian isolates were found only in OR and WA populations.
-MLG 6, 84 and 100 were (see Supplementary Table 1 for SSR genotypes related to 165 MLGs of total isolates) unique to CA isolates.
-
-Populations from different regions were compared with AMOVA, where differentiation was low among regions ($\Phi_{PT}$  = 0.11; P = 0.001).
-Pairwise comparisons of all populations according to region showed MX and CA populations were significantly different across all populations at P $\leq$ 0.001.
-The $\Phi_{PT}$ values of Pairwise comparisons between CA and other populations ranged from 0.116 to 0.389 and from 0.355 to 0.493 between MX and other population.
-The CA population showed evidence of sexual <!-- ZNK: There is the sample size caveat here that should be discussed. -->
-recombination ($\bar{r}_d$ = 0.008; P = 0.27, $I_A$ = 0.08; P = 0.3) while the $\bar{r}_d$ value of AU was 0.12 with a high P of 0.05 ($I_A$ = 0.70; P = 0.05), whereas all other populations showed linkage disequilibrium for loci tested indicating high possibility of clonal reproduction.
-
-The DAPC for hosts was limited to six most common cultivars: G122 (N = 55), Beryl (N = 62), Bunsi (N = 47), Yellow (N = 26), Pinto (N = 15), and Merlot (N = 11).
-According to Fig.
-8B, cultivar Merlot isolates group distant from others, but were not significantly different (P =0.325).
-Interestingly, *S.
-sclerotiorum* populations from cultivars 'Beryl' and 'G122', which are known to have different level of white mold susceptibility, showed no difference in population genetic data.
-
-AMOVA revealed genetic differences among isolates sampled in different years (DAPC biplot not shown).
-The $\Phi_{PT}$ was a low value of 0.07 with P = 0.001.
-Isolates from 2012 were not included in the analysis since there were only 4 samples.
-Samples for other years ranged from 19 in 2009 to 63 in 2005.
-Sampling year 2003 to 2005 and 2009 were different across every Pairwise comparison.
-None of the above sampling variables (i.e.
-region, year, host) seemed to influence distinct molecular composition of *S.
-sclerotiorum* populations since pre-defined (i.e.
-region, host, and year) isolates and isolates that were assigned into post groups by DAPC varied largely (Figs.
-8A, 8B and 9).
-However, assignment of isolates by DAPC analysis of regions did show some promising results for MX and CA isolates (Fig. 8A).
-The original group size of each region and DAPC defined group sizes for different regions are given in Table 6.
+The analysis of molecular variance (AMOVA) over the hierarchy of Source and Region showed no significant variation between white mold nurseries and non-white mold nurseries after accounting for variation between regions (Table \@ref(tab:amova)).
+Because Mexico shared no genotypes with any of the other populations and we suspected that it was contributing to some of the observed variance, we re-ran AMOVA without the Mexican isolates and found similar decreased, but still significant levels of differentiation between regions.
+In contrast, when we compared the three cultivars, Beryl, Bunsi, and G122, we found no evidence of population differentiation (Supplementary Data).
 
 -------------------------------------------------------------------------------
-Hierarchy               d.f.        S.S.       % variation    $\Phi statistic$  
+Hierarchy               d.f.        S.S.       % variation    $\Phi$ statistic  
 -------------------- ---------- ------------- -------------- ------------------
@@
 
-Since `find.clusters()` function of *adegenet* did not result in a clear-cut number of clusters for 318 isolates, we arrived at six (6) as the best number of populations (Fig.
-9) for our data after examining results of DAPC for different number of clusters.
-Three clusters, namely clusters one, three, and six were formed separately from the rest.
-The cluster 3 of the DAPC plot consisted entirely of MX isolates (N =13/15).
-Rest of the two MX isolates were grouped in cluster 1.
-Cluster 6 had 50 isolates in total representing mainly of NE (N = 15/37), MI (N = 12/58), and CO (N = 12/34).
-All CA (N = 18) isolates and 13/58 WA isolates dominated the cluster 1.
-AMOVA on these six clusters resulted 44% molecular variance among populations with P = 0.001.
-Pairwise comparison showed significant difference (P = 0.001) for all pairs.
-The DAPC plot after pooling all isolates together and removing clones (N =165) was compared with the plot of 318 isolates looked similar to Fig. 8C.
-The number of clusters was set to 6 for the analysis which resulted in a DAPC having similar topology to plot in Fig. 9 with 318 isolates(results not shown).
-Isolates of MX grouped separately while another distinct group had 24 isolates representing CO, NE, and ND.
+### Discriminant Analysis of Principal Components
 
-We tested the clone corrected data for evidence of sexual outcrossing within *S. sclerotiorum* populations in different regions and different years/region.
-CA showed evidence of sexual outcrossing (Ia = 0.08, P.Ia = 0.317, $\bar{r}_d$ = 0.008, P.$\bar{r}_d$ = 0.317) while all other regions implied clonal propagation (data not shown).
-Five populations showed evidence of sexual outcrossing: CA in 2004, FR in 2004 and 2005, OR in 2003, and WA in 2008.
-However, after clone-correction these populations had 6-7 isolates/population **(Table 7)** and therefore, may not be sufficient to make inferences about reproductive mode.
-Only exception was the WA_2008 population, which consisted of 12 isolates.
-Within sampling locations white mold nurseries (wmn) were tested for linkage disequilibrium since same wmn fields were used for sampling over different years.
-Only CA wmn showed evidence of sexual reproduction ( = 0.008; P = 0.29) while all other wmn locations suggested clonal reproduction.
-The DAPC analyses of WA, and MI isolates from same wmn fields over different years are in Fig. 10.
-Since *S. sclerotiorum* isolates are predominantly clonal, we wanted to explore whether there was any introduction of new isolates from other fields or if there were evidence of sexual reproduction.
-The AMOVA results for different clusters of Fig. 10A revealed significance difference between WA 2008 isolates from the rest ($\Phi_{PT}$ $\geq$ 0.206, P = 0.001).
-The 2003 and 2005 isolates were genetically similar ($\Phi_{PT}$ = 0.056, P = 0.129).
-MI isolates from 2008 and 2003 (Fig.
-10B) showed significant difference ($\Phi_{PT}$ = 0.127, P = 0.04) while 2005 isolates did not differ molecularly from both 2008 and 2003 isolates ($\Phi_{PT}$ $\leq$ 0.078, P $\geq$ 0.076).
-NE wmn isolates sampled in 2004 and 2008 were genetically similar ($\Phi_{PT}$ = 0.042, P = 0.179) (DAPC results not shown).
-The standardized index of association of wmn isolates from each region (i.e.
-MI, WA, and NE) for above mentioned sampling years suggested clonal reproduction ( $\geq$ 0.05, P $\leq$ 0.002).
+Discriminant Analysis of Principal Components (DAPC) was performed with the first 21 principal components, representing 88% of the total variance. 
+The first discriminant axis (representing 64% of the discriminatory power) separated the centroid for the Mexico isolates from the rest of the data, indicating strong differentiation (Fig.
+\@ref(DAPC)b).
+The second discriminant axis separated the centroid for the Californian isolates.
+The mean population assignment probabilities for all populations with n > 10 showed that only isolates from Mexico, CA, and France had > 50% probabilities of being reassigned to their source populations (Fig.
+\@ref(DAPC)a).
+
+<!-- DAPC figure -->
+
+![
+Discriminant Analysis of Principal Components on regions showing that Mexico is differentiated from other populations. 
+**(a)** Mean population assignment (horizontal axis) probability (vertical axis) for all populations with N > 10 (facets).
+Numbers next to source populations indicate population size.
+All values sum to one.
+**(b)** Scatter plot of first two components from DAPC.
+Points represent observed individuals connected to the population centroids with ellipses representing a 66% confidence interval for a normal distribution. 
+The center of each component is represented as grid lines.
+](../../results/figures/publication/DAPC.pdf){#DAPC width=100%}
+
+<!-- end figure -->
+
+### Linkage Analysis {-}
+
+Nearly all populations showed evidence for linkage (\@ref(fig:iatab)), which serves as evidence for clonal reproduction or other forms of non-random mating. 
+The only exceptions are CA and Australia. Both of these populations showed only moderate significance (P < 0.05) with $\bar{r}_d$ values of 0.03 and 0.12, respectively.
 

### 2017-07-11 doc/manuscript/manuscript.Rmd
nitpciky changes

@@
 
-### Analysis of Molecular Variance
+### Analysis of Molecular Variance {-}
 
@@
 
-### Discriminant Analysis of Principal Components
+### Discriminant Analysis of Principal Components {-}
 
@@
 
-Nearly all populations showed evidence for linkage (\@ref(fig:iatab)), which serves as evidence for clonal reproduction or other forms of non-random mating. 
+Nearly all populations showed evidence for linkage (Table \@ref(tab:iatab)), which serves as evidence for clonal reproduction or other forms of non-random mating. 
 The only exceptions are CA and Australia. Both of these populations showed only moderate significance (P < 0.05) with $\bar{r}_d$ values of 0.03 and 0.12, respectively.

### 2017-07-12 doc/manuscript/manuscript.Rmd
fixed some issues; add Bruvo

@@
 sclerotiorum* were collected over several years from grower fields and/or white mold nurseries in 11 states of the United States of America, Mexico, France, and Australia (Table \@ref(tab:isolate-table)).
-The white mold nurseries in different regions are indicated by wmn (white mold nursery) or mtc (Mitchell) and all other isolates came from commercial fields (Table \@ref(tab:isolate-table)).
+The white mold nurseries in different regions are indicated by wmn (white mold nursery) and all other isolates came from commercial fields (Table \@ref(tab:isolate-table)).
 
@@
 Once actively growing, mycelial compatibility group (MCG) and aggressiveness was determined for each isolate.
-MCG was determined through co-culturing isolates in pairs on potato-dextrose agar (PDA) plates, where incompatibility of different MCG resulted in formation of a mycelial barrier or barrage line [@kohn1990mycelial].
-Since MCG are considered mutually exclusive, pairing was not required between all isolates in the study.
+MCG was determined through co-culturing isolates in pairs on potato-dextrose agar (PDA) plates, where incompatibility of different MCGs resulted in formation of a mycelial barrier or barrage line [@kohn1990mycelial].
+Since MCGs are considered mutually exclusive, pairing was not required between all isolates in the study.
 
@@
 Capillary electrophoresis (fragment analysis) of amplicons, with size standard GeneScan™ 500 LIZ®, was performed using an ABI 3730 genetic analyzer (Life Technologies Corporation, Carlsbad, CA) at the Michigan State University Genomic Sequencing Center.
-Alleles were binned using PeakScanner version 1.0 (Life Technologies Corporation, Carlsbad, CA) and recorded manually in an excel sheet.
+Alleles were scored using PeakScanner version 1.0 (Life Technologies Corporation, Carlsbad, CA) and recorded manually in an excel sheet.
 
@@
 
-Of the 16 microsatellite loci, five included compound repeats, which made it challenging to accurately/confidently bin alleles into fragment sizes expected for each locus based on the described repeat motif.
+All analyses were performed on this data set unless otherwise mentioned.
+All data processing and analyses were performed in R version 3.4.0 [@R] and are openly available and reproducible at https://github.com/everhartlab/sclerotinia-366/.
+Of the 16 microsatellite loci genotyped, five included compound repeats, which made it challenging to accurately/confidently bin alleles into fragment sizes expected for each locus based on the described repeat motif.
+<!-- add something about repeat length consistency  -->
 Loci with compound repeats were removed for the reported statistics, but were processed and included in data analyses to ensure the integrity of the results with 11 loci as compared to 16.
@@
 To avoid including samples that were potentially collected from the same plant (which increases the probability of collecting the same sample more than once), data were clone-corrected on the region/source/host/year hierarchy for subsequent analyses.
-
 The clone-corrected data set included 318 isolates.
-All analyses were performed on this data set unless otherwise mentioned.
-All data processing and analyses were performed in R version 3.4.0 [@R] and are openly available and reproducible at https://github.com/everhartlab/sclerotinia-366/
 
@@
 ![
-Biplot showing 10 most influential explanatory variables (arrows) overlayed on the first two eigenvectors of distance based redundancy analysis of *S. sclerotiorum* isolates. 
+Biplot showing five most influential explanatory variables (arrows) overlayed on the first two eigenvectors of distance based redundancy analysis of *S. sclerotiorum* isolates. 
 The length of the arrows are directly proportional to the strength of the correlation between explanatory and molecular variables. 
@@
 The three communities with the most members contained 8 of the ten most abundant MCGs. 
-Overall, however, there was no clear pattern to the association between MLG and MCGs.
+Comparing these communities with Bruvo's genetic distance revealed and average distance of 0.451 among communities and an average distance of 0.437 within communities. 
+Overall, there was no clear pattern to the association between MLG and MCGs.
 

### 2017-07-12 doc/RMD/mlg-mcg.Rmd
add joyplot and info regarding bruvo dist b/n comm

@@
 
-```{r group_dist}
+```{r group_dist, fig.width = 10, fig.height = 10}
 datmat <- as.matrix(datdist)
@@
   group_by(comm) %>%
-  summarize(dist = list(as.dist(datmat[Isolate, Isolate])), N = n())
+  summarize(dist = list(datmat[Isolate, Isolate]), N = n())
+
+isolist <- strat_mcg %>% 
+  group_by(comm) %>% 
+  filter(!duplicated(MLG)) %>%
+  summarize(Isolates = list(Isolate)) %>% 
+  inner_join(expand(., comm, comm)) %>%
+  rowwise() %>%
+  mutate(COMMS = paste(sort(c(comm, comm1)), collapse = " ")) %>% 
+  ungroup() %>% 
+  filter(!duplicated(COMMS))
+isolist <- isolist %>%
+  filter(comm == comm1) %>% 
+  select(-comm, -COMMS) %>% 
+  rename(Isolates2 = Isolates) %>% 
+  inner_join(isolist)
+
+distmcg <- isolist  %>% 
+  group_by(comm, comm1, COMMS) %>% 
+  summarize(dist = list(datmat[Isolates[[1]], Isolates2[[1]]]), 
+            N = length(Isolates[[1]])) %>% 
+  rowwise() %>% 
+  mutate(dist = case_when(isSymmetric(dist) ~ list(dist[lower.tri(dist)]), 
+                          TRUE ~ list(as.numeric(dist))))
+
 distmcg %>%
-  group_by(comm) %>%
-  mutate(meandist = mean(dist[[1]]),
-         mediandist = median(dist[[1]])) %>%
+  rowwise() %>%
+  mutate(mean   = mean(dist),#mean(ifelse(dist == 0, NA, dist), na.rm = TRUE),
+         median = median(dist)) %>%#median(ifelse(dist == 0, NA, dist), na.rm = TRUE)) %>%
   ungroup() %>%
-  select(comm, N, meandist, mediandist)
-distmcg %>% 
-  unnest() %>%
-  ggplot(aes(x = dist, fill = comm)) +
-  geom_density(alpha = 0.5) +
-  geom_rug(aes(color = comm), alpha = 0.12) +
-  scale_fill_viridis(direction = -1, discrete = TRUE) +
-  scale_color_viridis(direction = -1, discrete = TRUE) +
+  select(COMMS, mean, median, N) %>%
+  rename(Comparison = COMMS) %>%
+  knitr::kable()
+
+distplot <- distmcg %>% 
+  mutate(mean = mean(dist)) %>%
+  mutate(COMMS = ifelse(comm == comm1, paste0(comm, "\n"), paste0(comm, "/\n", comm1))) %>%
+  mutate(comm = gsub(" \\(", "\n(", comm)) %>%
+  mutate(comm1 = gsub(" \\(", "\n(", comm1)) %>%
+  mutate(comparison = ifelse(comm == comm1, "Within", "Among")) %>%
+  ungroup() %>%
+  unnest()
+
+distplot %>%
+  group_by(comparison) %>%
+  summarize(mean = mean(dist), median = median(dist))
+
+if (require("ggjoy")){
+  ggplot(distplot, aes(x = dist, y = COMMS, group = COMMS, height = ..density..)) + 
+    geom_joy(aes(lty = comparison, fill = mean)) + 
+    theme_joy() + 
+    theme(axis.text.y = element_text(hjust = 0)) + 
+    viridis::scale_fill_viridis(option = "A", limits = quantile(distplot$dist, c(0.25, 0.75)))
+}
+
+ggplot(distplot, aes(x = dist, group = COMMS)) +
+  geom_density(aes(fill = mean, lty = comparison)) +
   ggtitle("Bruvo's distance by MCG community") +
-  theme_bw(base_size = 14)
+  xlab("Bruvo's distance") +
+  facet_grid(comm1~comm) +
+  theme_minimal(base_size = 16, base_family = "Helvetica") +
+  theme(aspect.ratio = 1) +
+  theme(panel.grid.major.y = element_blank()) +
+  theme(panel.background = element_rect(fill = "grey95")) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
+  viridis::scale_fill_viridis(option = "A", limits = quantile(distplot$dist, c(0.25, 0.75)))
+
+commamova <- dat11 %>% 
+  addStrata(data.frame(Community = strat_mcg$comm)) %>%
+  clonecorrect(~Community) %>%
+  poppr.amova(~Community, dist = as.dist(datmat[indNames(.), indNames(.)]), 
+              method = "pegas", nperm = 1000)
+commamova
 ```

### 2017-07-12 doc/manuscript/manuscript.Rmd
add joyplot and info regarding bruvo dist b/n comm

@@
 The three communities with the most members contained 8 of the ten most abundant MCGs. 
-Comparing these communities with Bruvo's genetic distance revealed and average distance of 0.451 among communities and an average distance of 0.437 within communities. 
+Comparing these communities with Bruvo's genetic distance revealed and average distance of 0.451 among communities and an average distance of 0.437 within communities, showing no significant difference.
 Overall, there was no clear pattern to the association between MLG and MCGs.

### 2017-07-12 doc/manuscript/manuscript.Rmd
Comment out current discussion

@@
 
+<!-- 
 Our study was conducted to characterize *S.
@@
 Our study did not find any multilocus genotype/s common across all or majority of sampling geographies.
-<!-- ZNK: I think this statement is a bit misleading.
+<!-~ ZNK: I think this statement is a bit misleading.
 Yes, we didn't find any individual MLGs, but 46% of these are present across at least two populations.
--->
+~->
 The most widespread and common genotype (MLG 25, Supplemental Table 1) was found only in five states of the USA with majority of isolates confined to ND.
@@
 On the other hand CA genotypes were present in the West Coast states of OR and WA which may be due to movement of sclerotia with planting material.
-<!-- ZNK: Do we have some sort of evidence for this, or is this pure speculation?
--->
-<!-- 
+<!-~ ZNK: Do we have some sort of evidence for this, or is this pure speculation?
+~->
+<!-~ 
 I'm not sure what the point of this paragraph is.
 I believe I can put the discussion of the graph results here. 
--->
+~->
 Surprisingly, France and Australian genotypes were also isolated in the USA.
@@
 Some France genotypes were common in USA isolates but at the same time 43% (6/14) of the France genotypes were unique to that region indicating white mold resistant cultivars produced in the USA might not work in that region.
-<!-- ZNK: I'm not entirely sold on this conclusion.
--->
+<!-~ ZNK: I'm not entirely sold on this conclusion.
+~->
 Geographic distance between different states of the USA seem not a barrier for gene transfer of *S.
 sclerotiorum* isolates though unique genotypes did prevail at a low frequency in each state.
-<!-- ZNK: I don't know that this adds anything.
--->
+<!-~ ZNK: I don't know that this adds anything.
+~->
 The genetic diversity of regions with >10 isolates had similar values (0.43 - 0.59) except for Mexican isolates.
-*h* of MX was 0.26 and indicated less exposure of sampling sites to outside inocula. <!-- ZNK: Does this mean that there's heavy inbreeding here? -->
+*h* of MX was 0.26 and indicated less exposure of sampling sites to outside inocula. <!-~ ZNK: Does this mean that there's heavy inbreeding here? ~->
 Test for linkage disequilibrium showed evidence of clonal reproduction for MX isolates (Ia = 2.61, P.Ia = 0.001, $\bar{r}_d$ = 0.37, P.$\bar{r}_d$  = 0.001).
@@
 We also did not detect location as a significant predictor of molecular variance but year had a significant impact.
-Our samples were collected on a single day in each year from each field and therefore did not warrant for an analysis of date of sample collection. <!-- BSA: Is this true, coauthors? -->
+Our samples were collected on a single day in each year from each field and therefore did not warrant for an analysis of date of sample collection. <!-~ BSA: Is this true, coauthors? ~->
 
@@
 These results indicate some level of heterothallic recombination is taking place in these fields in addition to natural mutations.
-An understanding of sexual behavior and ability of form new genotypes by *S. sclerotiorum* in dry bean cropping systems is useful for formulating disease management strategies.
+An understanding of sexual behavior and ability of form new genotypes by *S. sclerotiorum* in dry bean cropping systems is useful for formulating disease management strategies. -->
 

### 2017-07-12 doc/RMD/MLG-distribution.Rmd
wrap session info in details tag
add DAPC for region

@@
 
-```{r}
+```{r load_packages, message = FALSE, warning = FALSE}
 library('tidyverse')
@@
 
-## Session Information
+### Region Cohesion
 
-```{r}
+Here I am using DAPC to assess how cohesive these artificial regions are.
+
+```{r, region_cohesion}
+g11o <- igraph::cluster_optimal(graph11loc$total)
+comm <- igraph::communities(g11o)
+names(comm) <- c("International", "Costal", "Midwest")
+comm
+strata(dat11) <- strata(dat11) %>%
+  mutate(MLGRegion = case_when(
+    .$Region %in% comm$International ~ "International",
+    .$Region %in% comm$Costal ~ "Costal",
+    .$Region %in% comm$Midwest ~ "Midwest",
+    TRUE ~ as.character(.$Region)
+  ))
+
+setPop(dat11) <- ~MLGRegion
+```
+
+```{r region_dapc}
+set.seed(2017-07-12)
+reg <- xvalDapc(tab(dat11), pop(dat11), n.rep = 100)
+reg$DAPC
+scatter(reg$DAPC, col = viridis::viridis(6), bg = "grey90")
+ggcompoplot::ggcompoplot(reg$DAPC, dat11, pal = viridis::viridis(6), cols = 2)
+reg$DAPC$posterior %>%
+  as_tibble() %>%
+  bind_cols(data.frame(Population = reg$DAPC$grp)) %>%
+  group_by(Population) %>%
+  summarize_all(mean) %>%
+  gather(assignment, probability, -Population) %>%
+  ggplot(aes(x = assignment, y = probability)) +
+  geom_segment(aes(xend = assignment, yend = 0)) +
+  geom_point(pch = 21, fill = "grey") +
+  facet_wrap(~Population) +
+  scale_y_continuous(limits = c(0, 1), 
+                     minor_breaks = c(seq(0, 1, by = 0.125)), 
+                     breaks = c(0, 0.5, 1)) +
+  theme_bw(base_size = 16, base_family = "Helvetica") +
+  theme(legend.position = "top") +
+  theme(aspect.ratio = 1/2) +
+  theme(axis.text = element_text(color = "black")) +
+  theme(axis.ticks.y = element_blank()) +
+  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
+  theme(panel.grid.major.x = element_line(linetype = 0, color = "grey50")) +
+  theme(panel.grid.major = element_line(colour = "grey20")) +
+  theme(panel.grid.minor = element_line(linetype = 3, colour = "grey50")) +
+  theme(strip.background = element_rect(color = NA, fill = "grey90")) +
+  theme(strip.text = element_text(face = "bold", hjust = 0.05)) +
+  theme(panel.border = element_blank()) +
+  xlab("Population Assignment") +
+  ylab("Mean Probability")
+```
+
+<details>
+<summary>Session Information</summary>
+
+```{r, echo = FALSE}
 options(width = 100)

### 2017-07-12 doc/RMD/wmn-differentiation.Rmd
add DAPC for white mold nurseries

@@
 
-```{r AMOVA}
+```{r AMOVA, cache = TRUE}
 other(dat11cc)$REPLEN
@@
 
+## WMN DAPC
+
+DAPC is a nice way of visualizing these. First, I'm simply going to compare 
+between white mold nurseries and others:
+
+```{r wmndapc, cache = TRUE}
+filter <- strata(dat11cc) %>%
+  group_by(Region) %>%
+  mutate(filter = length(unique(SourceType)) > 1) %>%
+  pull(filter)
+wmn_compare <- dat11cc[filter]
+set.seed(2017-07-12)
+wmndapc <- xvalDapc(tab(wmn_compare), pop(wmn_compare), n.rep = 200, n.pca = 1:10)
+wmndapc
+scatter(wmndapc$DAPC)
+(wmntab <- table(wmndapc$DAPC$assign, wmndapc$DAPC$grp))
+chisq.test(wmntab, simulate.p.value = TRUE)
+```
+
+Okay, this shows that there's quite a bit of overlap between these two. What
+happens if I look at the DAPC when it includes the pouplation groupings. Mainly,
+I want to know if the white mold nurseries can be compared while considering the
+pouplations
+
+```{r wmndapcpop, fig.width = 8, fig.height = 10, cache = TRUE}
+set.seed(2017-07-12)
+wmndapc <- xvalDapc(tab(wmn_compare), strata(wmn_compare, ~SourceType/Region)[[2]], 
+                    n.pca = 5:20, n.rep = 100)$DAPC
+scatter(wmndapc)
+LDS <- bind_cols(Population = wmndapc$grp, as.data.frame(wmndapc$ind.coord)) %>%
+  as_tibble()
+LDS_pop <- LDS %>% 
+  group_by(Population) %>% 
+  summarize_all(mean) %>%
+  rename_all(function(x) gsub("LD", "mean", x))
+LDS <- full_join(LDS, LDS_pop) %>%
+  separate(Population, c("SourceType", "Pop"), remove = FALSE) 
+LDS_PLOT <- ggplot(LDS, aes(x = LD1, y = LD2, color = SourceType)) + 
+  geom_point(aes(fill = SourceType), alpha = 0.5, pch = 21, color = "black") +
+  geom_segment(aes(x = mean1, y = mean2, xend = LD1, yend = LD2), alpha = 0.5) +
+  stat_ellipse(type = "norm", level = 0.66, alpha = 0.75) + 
+  # ggrepel::geom_label_repel(aes(x = mean1, y = mean2, label = Population), 
+  #                           data = LDS_pop, show.legend = FALSE, color = "black") +
+  theme_bw() +  
+  theme(aspect.ratio = 1/1.618) +
+  theme(legend.position = "bottom") +
+  theme(axis.text = element_blank()) + 
+  theme(axis.title = element_blank()) + 
+  theme(axis.ticks = element_blank()) + 
+  viridis::scale_color_viridis(discrete = TRUE, direction = -1) +
+  viridis::scale_fill_viridis(discrete = TRUE, direction = -1) +
+  scale_y_continuous(breaks = 0) + 
+  scale_x_continuous(breaks = 0) + 
+  theme(panel.background = element_rect(fill = "grey95")) +
+  theme(panel.grid.major = element_line(color = "black"))
+
+LDS_PLOT + facet_wrap(~Pop, ncol = 3)
+
+```
+
 # Host differentiation

### 2017-07-12 doc/manuscript/manuscript.Rmd
rerun manuscript

@@
 
+In this study, we characterized the diversity of *Sclerotinia sclerotiorum* from dry bean fields across the United States.
+Our results suggest that populations from white mold nurseries may reflect the populations of the surrounding regions. 
+We found significant population differentiation by geographic region, mainly differentiated into four broad groups, a Coastal Region, Midwestern Region, International Region, and Mexico. 
+There were not significant associations between multilocus genotypes and mycelial compatibility groups.
+
 <!-- 

### 2017-07-13 doc/manuscript/manuscript.Rmd
add notes for future zhian

@@
 Once actively growing, mycelial compatibility group (MCG) and aggressiveness was determined for each isolate.
-MCG was determined through co-culturing isolates in pairs on potato-dextrose agar (PDA) plates, where incompatibility of different MCGs resulted in formation of a mycelial barrier or barrage line [@kohn1990mycelial].
+MCG was determined as described in @otto-hanson2011variation through co-culturing isolates in pairs on potato-dextrose agar (PDA) plates, where incompatibility of different MCGs resulted in formation of a mycelial barrier of dead tissue (barrage line) [@kohn1990mycelial].
 Since MCGs are considered mutually exclusive, pairing was not required between all isolates in the study.
@@
 
+<!-- 
+ - Big Picture reminder of importance
+   + This is an extension of a previous work [@otto-hanson2011variation]
+   + White Mold Nurseries serve as good way to screen potential resistant lines, avoiding environmental effects
+   + It's unclear how genetic structure can influence
+ - Critical Analysis of Major Findings
+   + This is the largest genetic analysis of *S. sclerotiorum* to date
+   + Similar to the findings of @otto-hanson2011variation, we found that
+     Region has a strong influence on both population structure and
+     severity.
+   + North Dakota appeared
+ - Additional Findings and how they fit into the literature
+ - Limitations
+ - Future Directions
+ - Overall Conclusion and Impact
+-->
+
 In this study, we characterized the diversity of *Sclerotinia sclerotiorum* from dry bean fields across the United States.
-Our results suggest that populations from white mold nurseries may reflect the populations of the surrounding regions. 
+Our results suggest that, in general, populations from white mold nurseries reflect the populations of the surrounding regions. 
 We found significant population differentiation by geographic region, mainly differentiated into four broad groups, a Coastal Region, Midwestern Region, International Region, and Mexico. 
-There were not significant associations between multilocus genotypes and mycelial compatibility groups.
+We found no significant associations between genotype and phenotype.
 

### 2017-07-22 doc/manuscript/manuscript.Rmd
add sclerotinia report references

@@
 Field trials in some regions of the US have shown the G122 cultivar is partially resistant while the Beryl cultivar is susceptible [@otto-hanson2011variation; @steadman2004identification].
+<!-- This puny paragraph needs a lot of work to more fully describe the operation of the wmn -->
 
@@
 As these nurseries are not treated with any chemical or cultural control of white mold, we hypothesized that these nurseries would represent the natural population of *S. sclerotiorum*.
-Furthermore, as three cultivars (Beryl, Bunsi, and G122) were grown among all white mold nurseries, we wanted to investigate the potential effect of cultivar on genetic diversity.
+Furthermore, as three cultivars (Beryl, Bunsi, and G122) were grown and tested previously by @otto-hanson2011variation among all white mold nurseries, we wanted to investigate the potential effect of cultivar on genetic diversity.
 We additionally wanted to assess what categorical or phenotypic variables best predict genetic structure and if there was correlation between multilocus genotype and mycelial compatibility group.

### 2017-07-25 doc/RMD/mlg-mcg.Rmd
update manuscript and calculation of singletons

@@
 mcg_mlg_graph
+if (!interactive()){
   ggsave(plot = mcg_mlg_graph, filename = file.path(PROJHOME, "results/figures/publication/FigureS2.pdf"), 
        width = 88*3, height = 88*3.5, units = "mm")
+}
 ```
 
-Of interest is accounting for the communities that have a single MCG/MLG pair.
-We can detect these by filtering for communities that only have two members:
+Of interest is accounting for the communities that have a single MLG within the
+community. This gives us a lower bound at how many MLGs are reproductively
+isolated. We detect these by filtering for communities that only have one MLG:
 
@@
 singles <- the_communities %>% 
-  count(community) %>% 
-  arrange(desc(n)) %>% 
-  filter(n == 2) %>% 
+  mutate(MLG = grepl("MLG", vertex)) %>% 
+  group_by(community) %>% 
+  summarize(n = sum(MLG)) %>% 
+  filter(n == 1) %>%
   inner_join(the_communities, by = "community") %>% 
   mutate(type = case_when(grepl("MLG", vertex) ~ "MLG", TRUE ~ "MCG")) %>% 
-  spread(type, vertex) %>%
-  select(-n, -comm)
+  group_by(community, type) %>% 
+  summarize(vertex = paste(as.character(vertex), collapse = ", ")) %>%
+  spread(type, vertex) 
 # assessing that this truely contains the singletons
-nrow(inner_join(singles, the_communities, by = c("MCG" = "vertex")))
 nrow(inner_join(singles, the_communities, by = c("MLG" = "vertex")))
-
 knitr::kable(singles)

### 2017-07-25 doc/manuscript/manuscript.Rmd
update manuscript and calculation of singletons

@@
 We first created a contingency table with MLGs and MCGs and converted it to a directed and weighted edgelist where each edge was a connection from an MCG to an MLG weighted by the number of samples shared in the connection.
+This was then converted to a bipartite graph where top nodes represented MCGs and bottom nodes represented MLGs.
 The cluster walktrap community detection algorithm as implemented in the `cluster_walktrap()` function in *igraph* version 1.0.1 was used to define clusters of MLGs and MCGs [@csardi2006igraph; @pons2006computing].
@@
 Our network-based approach to correlating MLGs with MCGs revealed a large and complex network (Fig. \@ref(biggraph)). 
-Community analysis showed 51 communities, 23 of which consisted of a single MLG/MCG pair. 
+Community analysis showed 51 communities, 26 of which consisted of a single MLG indicating that 15.76% of the 165 MLGs are unable to cross with any other MLG in this data set.
 The three communities with the most members contained 8 of the ten most abundant MCGs. 

### 2017-07-26 doc/manuscript/manuscript.Rmd
add stuff to discussion to be deleted later

@@
 Four regions had fewer than 10 isolates, Australia (6), WI (2), NY (1), ID (1).
-After clone-correction on the hierarchy of Region/Source/Host/Year, we were left with 318 isolates representing 165 MLGs \@ref(tab:iatab).
+After clone-correction on the hierarchy of Region/Source/Host/Year, we were left with 318 isolates representing 165 MLGs (Table \@ref(tab:iatab)).
 Analysis of the genotype accumulation curve showed no clear plateau, indicating that we would likely obtain more multilocus genotypes if we were to collect more loci.
@@
 
-We observed 87 mycelial compatibility groups (MCG), the most abundant of which represented 73 isolates over 37 multilocus genotypes (MLG) (Fig. \@ref(barplot)). 
+We observed 87 mycelial compatibility groups (MCG), the most abundant of which (MCG 5) represented 73 isolates over 37 multilocus genotypes (MLG) (Fig. \@ref(barplot)). 
 Our assessment of aggressiveness association with MCGs showed a significant effect (P < 0.001) indicating that MCG 44 is, on average, more aggressive than other MCGs (Table \@ref(tab:mcg-aggressiveness)).
@@
 In this study, we characterized the diversity of *Sclerotinia sclerotiorum* from dry bean fields across the United States.
-Our results suggest that, in general, populations from white mold nurseries reflect the populations of the surrounding regions. 
-We found significant population differentiation by geographic region, mainly differentiated into four broad groups, a Coastal Region, Midwestern Region, International Region, and Mexico. 
-We found no significant associations between genotype and phenotype.
+Our results suggest that, broadly, populations from white mold nurseries reflect the populations of the surrounding regions, indicating that resistance screening may be successful at a regional level.
+We found significant population differentiation by geographic region, mainly differentiated into three broad North American groups based on shared genotypes and posterior groupings, a Coastal Region, Midwestern Region, and Mexico.
+To date, with 366 isolates, this is the largest population genetic study of *S. sclerotiorum* assessing population structure within managed and unmanaged agricultural environments. 
+These findings indicate that the white mold nurseries can be effective at screening for potential resistant lines within growing regions.
+
+Of particular interest in this study was testing the association between mycelial compatibility groups (MCGs) and multilocus genotypes (MLGs).
+In the past MCGs have been used as an indicator for clonal lineages [@otto-hanson2011variation; @kohn1995clonal; @lehner2015analysis]. 
+As genetic markers became more widely available, studies began to find conflicting results regarding the association between MCG and MLG [@sirjusingh2001characterisation; @lehner2015analysis].
 

### 2017-07-26 doc/RMD/mlg-mcg.Rmd
update MCG analysis to only include truely unconnected MLGs

@@
 community. This gives us a lower bound at how many MLGs are reproductively
-isolated. We detect these by filtering for communities that only have one MLG:
+isolated. We detect these by filtering for communities that only have one MLG.
+But, because we do have communities with one MLG that are peripheral to larger
+communities, we can also check the degree of the MCG and MLG nodes to ensure 
+that they correspond one-to-one.
 
@@
   summarize(vertex = paste(as.character(vertex), collapse = ", ")) %>%
-  spread(type, vertex) 
+  spread(type, vertex) %>%
+  mutate(MCG_degree = map_dbl(strsplit(MCG, ", "), . %>% degree(g, .) %>% sum())) %>%
+  mutate(MLG_degree = degree(g, MLG)) %>%
+  filter(MLG_degree - MCG_degree == 0)
 # assessing that this truely contains the singletons
 nrow(inner_join(singles, the_communities, by = c("MLG" = "vertex")))
+singles
+
 knitr::kable(singles)
+```
+
+This result indicates that there are `r nrow(singles)` MLGs that are unable to
+pair with any other MLGs due to mycelial incompatibility. This represents 
+`r paste0(sprintf("%0.2f", nrow(singles)/nmll(dat11, "original")*100), "%")` 
+of the total MLGs.
+
+```{r}
+mll(dat11) <- "original"
+mlg.table(dat11, ~Region, color = TRUE, mlgsub = as.integer(gsub("MLG.", "", singles$MLG)))
+p <- last_plot()
+p + 
+  scale_fill_brewer(palette = "Set1") +
+  ggtitle("Incompatible Multilocus Genotypes") +
+  labs(list(caption = "These MLGs cannot pair with any other MLG due to\nmycelial incompatibility"))
 ```
@@
 set.seed(2017-05-03)
-vgn
-# vgn %>%
-#   visIgraphLayout("layout_nicely") %>%
-#   visOptions(highlightNearest = list(enabled = TRUE, hover = TRUE), 
-#              nodesIdSelection = TRUE)
-#   
+
+vgn %>%
+  visIgraphLayout("layout_nicely") %>%
+  visOptions(highlightNearest = list(enabled = TRUE, hover = TRUE), 
+             nodesIdSelection = TRUE)
+
 ```

### 2017-07-26 doc/manuscript/manuscript.Rmd
update MCG analysis to only include truely unconnected MLGs

@@
 Our network-based approach to correlating MLGs with MCGs revealed a large and complex network (Fig. \@ref(biggraph)). 
-Community analysis showed 51 communities, 26 of which consisted of a single MLG indicating that 15.76% of the 165 MLGs are unable to cross with any other MLG in this data set.
+Community analysis showed 51 communities, 15 of which consisted of a single MLG unconnected with any other community indicating that just 9.09% of the 165 MLGs are unable to cross with any other MLG in this data set.
 The three communities with the most members contained 8 of the ten most abundant MCGs. 

### 2017-07-26 doc/RMD/MLG-distribution.Rmd
add colnames parameter

@@
     glayout <- layout_nicely(g)
+    colnames(glayout) <- c("x", "y")
   } else {

### 2017-07-26 doc/RMD/mlg-mcg.Rmd
Add analysis of connection

@@
 
+The opposite question is, how connected is this huge hairy mess? The community
+analysis gave us an idea of where different centers lie, but it doesn't really
+do much for us in actually giving us an idea of what the possabilities for 
+connection are. One way we can look at this is by using the `ego()` function to
+assess the neighborhood of vertices for a given node order, which is to say the
+number of edges away from the node you must travel to visit all your neighbors. 
+Order 0 would give us the node itself, and order 1 would give us the degree plus 
+one. If we set this degree to a large number, we will find all the possible 
+connections.
+
+Here I'm getting a simple measure by counting all the vertices, I'm subtracting
+two here because I want to remove the genotype and the MCG associated.
+
+```{r node_order}
+# getting the sizes for all genotypes.
+sizes <- ego_size(g, 200, nodes = which(grepl("Genotype", V(g)$type)), mode = "all") - 2
+table(sizes)
+```
+
+Here we can see there are 12 genotypes that are only associated with themselves
+and their respective MCGs. This corresponds with the results we found above. Up
+next are the genotypes that are associated with another partner, which could
+represent either two MCGs to one genotype (there are three of these), or two 
+genotypes to one MCG. By logical deduction, we can see that there are 5 of these
+left. The rest of the entries are not so easy to parse, but it's clear that 
+there are 130 genotypes that all have the ability to reach each other through
+mutual mycelial compatibility groups.
+
+To focus on strictly the MLGs, we can see a more clear picture. Here, we are
+taking the results of `ego()` and counting how many MLGs are in the vertex list.
+From there, we can tabulate the results and divide the number of entries by the
+category, giving us the number of communities in each category.
+
+```{r community_counting2}
+ncommunities <- ego(g, 200, nodes = which(grepl("Genotype", V(g)$type)), mode = "all") %>%
+  map_int(. %>% names %>% grepl("MLG", .) %>% sum) %>%
+  table() %>%
+  {./as.integer(names(.))}
+ncommunities
+map_dbl(as.integer(names(ncommunities)), choose, 2)
+```
+
+From here, we can see that we have 15 MLGs that only interact with themselves,
+3 pairs of MLGs, 1 triplet, 1 hendeclet, and 1 community with 130 members, of
+which there are 8385 possible combinations.
+
+Of course, the 8385 possible combinations doesn't give us an idea of what the
+number of immediately possible combinations exist. We can examine that by 
+setting the mode to 2.
+
+```{r community_counting, fig.width = 5, fig.height = 3}
+in_reach <- ego(g, 2, nodes = which(grepl("Genotype", V(g)$type)), mode = "all") %>%
+  map(function(.x) .x[grepl("MLG", names(.x))])
+lengths(in_reach) %>%
+  enframe(name = "index", value = "N") %>%
+  ggplot(aes(x = N - 1)) +
+  geom_histogram(binwidth = 1, center = 0, color = "black", fill = "grey50") +
+  theme_bw(base_size = 16, base_family = "Helvetica") +
+  theme(aspect.ratio = 1/2) +
+  theme(axis.text = element_text(color = "black")) +
+  scale_y_continuous(expand = c(0, 0.1)) +
+  scale_x_continuous(expand = c(0, 0.1), breaks = c((0:6)*10), limits = c(-1, 60)) +
+  theme(panel.grid.major = element_line(colour = "grey20")) +
+  theme(panel.grid.minor = element_line(colour = "grey50", linetype = 3)) +
+  theme(panel.grid.major.x = element_blank()) +
+  theme(panel.grid.minor.x = element_blank()) +
+  theme(panel.border = element_blank()) +
+  xlab("Number of MLG Neigbors")
+```
+
+The maximum here is `r max(lengths(in_reach))`, which represents MLG 25.
+
 ### Interactive Visualizations

### 2017-07-26 doc/manuscript/manuscript.Rmd
Add more text to manuscript discussion

@@
 Community analysis showed 51 communities, 15 of which consisted of a single MLG unconnected with any other community indicating that just 9.09% of the 165 MLGs are unable to cross with any other MLG in this data set.
+Of all the MLGs, MLG 25 had 57 neighbors that shared the same mycelial compatibility group. 
 The three communities with the most members contained 8 of the ten most abundant MCGs. 
@@
 Of particular interest in this study was testing the association between mycelial compatibility groups (MCGs) and multilocus genotypes (MLGs).
-In the past MCGs have been used as an indicator for clonal lineages [@otto-hanson2011variation; @kohn1995clonal; @lehner2015analysis]. 
-As genetic markers became more widely available, studies began to find conflicting results regarding the association between MCG and MLG [@sirjusingh2001characterisation; @lehner2015analysis].
+In the past MCGs have been used as a proxy for clonal lineages [@otto-hanson2011variation; @kohn1995clonal; @carbone1999patterns; @lehner2015analysis]. 
+When using genetic markers with MCGs, there are three basic relationship patterns, 1) One MLG to one MCG, 2) Several MLGs belong to one MCG, and 3) Several MCGs belong to one MLG, noting that 2 and 3 are not mutually exclusive.
+Our results show all three (Fig. \@ref(biggraph)) where only 12 of the observed MLGs and MCGs fit into the first pattern.
+
+All three of these patterns have been observed together in separate studies [@aldrich-wolfe2015genetic; @carbone1999patterns; @attanayake2012genetic].
 

### 2017-07-26 doc/manuscript/manuscript.Rmd
Add some jazz to manuscript

@@
 When using genetic markers with MCGs, there are three basic relationship patterns, 1) One MLG to one MCG, 2) Several MLGs belong to one MCG, and 3) Several MCGs belong to one MLG, noting that 2 and 3 are not mutually exclusive.
-Our results show all three (Fig. \@ref(biggraph)) where only 12 of the observed MLGs and MCGs fit into the first pattern.
-
-All three of these patterns have been observed together in separate studies [@aldrich-wolfe2015genetic; @carbone1999patterns; @attanayake2012genetic].
+Similar to what has been observed in other studies, our results show all three patterns (Fig. \@ref(biggraph)) [@aldrich-wolfe2015genetic; @carbone1999patterns; @attanayake2012genetic]. 
+Because we took a bipartite network approach to assessing the relationship between MCG and MLG, we were able to assess just how concordant MCGs were with MLGs.
 

### 2017-07-27 doc/RMD/mlg-mcg.Rmd
add code to further examine mlg/mcg relationship

@@
 
-## MLG table
+# MLG table
 
@@
 
-## making a graph
+# Making a Graph
 
@@
 
-```{r community_counting, fig.width = 5, fig.height = 3}
+```{r community_counting, fig.width = 6, fig.height = 3.5}
 in_reach <- ego(g, 2, nodes = which(grepl("Genotype", V(g)$type)), mode = "all") %>%
-  map(function(.x) .x[grepl("MLG", names(.x))])
-lengths(in_reach) %>%
-  enframe(name = "index", value = "N") %>%
-  ggplot(aes(x = N - 1)) +
-  geom_histogram(binwidth = 1, center = 0, color = "black", fill = "grey50") +
+  map_df(function(.x) as_tibble(list(MLG = list(names(.x[grepl("MLG", names(.x))])), 
+                                     MCG = list(names(.x[!grepl("MLG", names(.x))])), 
+                                     WHO = names(.x[1])
+                                     )
+                                )) %>% 
+  mutate(N = lengths(MLG))
+
+in_reach_summary <- in_reach %>% 
+  group_by(N) %>%
+  summarize(E = diversity_stats(as.integer(table(unlist(MCG))))["E.5"], 
+            MCG = n_distinct(unlist(MCG)), 
+            MLG = n())
+knitr::kable(in_reach_summary)
+ggplot(in_reach_summary, aes(x = N - 1, y = MLG, fill = E)) +
+  geom_segment(aes(yend = 0, xend = N-1)) +
+  geom_point(aes(size = MCG), pch = 21) +
+  # ggrepel::geom_label_repel(aes(label = MCG)) +
+  viridis::scale_fill_viridis(option = "D") +
   theme_bw(base_size = 16, base_family = "Helvetica") +
-  theme(aspect.ratio = 1/2) +
+  theme(aspect.ratio = 1/1.75) +
   theme(axis.text = element_text(color = "black")) +
-  scale_y_continuous(expand = c(0, 0.1)) +
-  scale_x_continuous(expand = c(0, 0.1), breaks = c((0:6)*10), limits = c(-1, 60)) +
+  scale_y_continuous(expand = c(0, 0.1), limits = c(0, 25)) +
+  scale_x_continuous(expand = c(0, 0.1), breaks = c((0:6)*10), limits = c(-2, 60)) +
   theme(panel.grid.major = element_line(colour = "grey20")) +
@@
   theme(panel.border = element_blank()) +
-  xlab("Number of MLG Neigbors")
+  labs(list(fill = expression(E[5]), 
+            size = "MCG"
+            # caption = "size and fill represent\nthe number and distribution\nof MCG"
+            )) +
+  xlab("Maximum Heterothallic Pairings") +
+  ylab("Multilocus Genotypes")
 ```
 
-The maximum here is `r max(lengths(in_reach))`, which represents MLG 25.
+The maximum here is `r in_reach %>% arrange(N) %>% tail(1) %>% pull(N)`, which 
+represents `r in_reach %>% arrange(N) %>% tail(1) %>% pull(WHO)`.
 
-### Interactive Visualizations
+## Interactive Visualizations
 
@@
 t5g
+if (!interactive())
   ggsave(t5g, file = file.path("results", "figures", "publication", "FigureS2B.pdf"), width = 88, height = 88, scale = 3, units = "mm")
@@
 severity_plot
+if (!interactive())
   ggsave(severity_plot, filename = "results/figures/publication/FigureS3.pdf", width = 183, unit = "mm")
@@
 
-## Assessing MCG/MLG community structure/cohesiveness
+# MCG accumulation permutations
+
+On 2017-07-27, Sydney came up with the idea to assess MCG accumulation with 
+increasing genetic distance similar to the genotype accumulation curve. First,
+let's see what the curve would look like if we had one MLG/MCG.
+
+First, we need to set up everything to calculate this quantity
+
+```{r mlg accumulation calculation}
+bd         <- bruvo.dist(dat11, replen = other(dat11)$REPLEN)
+thresholds <- filter_stats(dat11, distance = bd, plot = TRUE)
+thresh     <- c(0, unique(thresholds$farthest$THRESHOLDS) + .Machine$double.eps)
+collapsed  <- map(thresh, . %>% mlg.filter(dat11, threshold = ., distance = bd))
+
+# This function will count the number of original MLGs within each collapsed MLG
+# Note that "original" here can be any MLG source.
+compare_mlg <- function(collapsed, thresh, original){
+  as_tibble(list(MLG = collapsed, y = original, threshold = thresh)) %>%
+    group_by(MLG, threshold) %>%
+    summarize(N = n_distinct(y)) %>%
+    ungroup() %>%
+    mutate(NMLG = n())
+}
+
+# A custom function for stat summary that returns a single point when there is
+# only a single observation
+my_meanse <- function (x, mult = 1) {
+  x <- stats::na.omit(x)
+  se <- mult * sqrt(stats::var(x)/length(x))
+  mean <- mean(x)
+  if (length(x) == 1) return(data.frame(y = mean, ymin = mean, ymax = mean))
+  data.frame(y = mean, ymin = mean - se, ymax = mean + se)
+}
+```
+
+Now, as a demontration, we can see what the curve would look like if all MCGs
+were randmly distributed (e.g. not clumpy) among the MLGs
+
+```{r mlg accumulation noclump, fig.width = 5, fig.height = 3}
+map2_df(collapsed, thresh, compare_mlg, original = sample(87, 366, replace = TRUE)) %>%
+  ggplot(aes(x = threshold, y = N)) +
+  geom_point(alpha = 0.1) +
+  stat_smooth(se = FALSE, color = "black") +
+  stat_summary(pch = 21, fun.data = my_meanse, fill = "white")
+```
+
+Now we can see what it looks like when there is a 1:1 relationship of MCGs to
+MLGs:
+
+```{r mlg accumulation mlg, fig.width = 5, fig.height = 3}
+map2_df(collapsed, thresh, compare_mlg, original = mll(dat11, "original")) %>%
+  ggplot(aes(x = threshold, y = N)) +
+  geom_point(alpha = 0.1) +
+  stat_smooth(se = FALSE, color = "black") +
+  stat_summary(pch = 21, fun.data = my_meanse, fill = "white")
+```
+
+Yet another comparison would be if each sample were a unique MCG:
+
+```{r mlg accumulation uniq, fig.width = 5, fig.height = 3}
+map2_df(collapsed, thresh, compare_mlg, original = seq(366)) %>%
+  ggplot(aes(x = threshold, y = N)) +
+  geom_point(alpha = 0.1) +
+  stat_smooth(se = FALSE, color = "black") +
+  stat_summary(pch = 21, fun.data = my_meanse, fill = "white")
+```
+
+And it's worth examining what happens if there are only 10 MCGs:
+
+```{r mlg accumulation ten, fig.width = 5, fig.height = 3}
+map2_df(collapsed, thresh, compare_mlg, original = sample(10, 366, replace = TRUE)) %>%
+  ggplot(aes(x = threshold, y = N)) +
+  geom_point(alpha = 0.1) +
+  stat_smooth(se = FALSE, color = "black") +
+  stat_summary(pch = 21, fun.data = my_meanse, fill = "white")
+```
+
+And, of course, we should see what our observed data looks like:
+
+```{r mlg accumulation observed, fig.width = 5, fig.height = 3}
+observed <- map2_df(collapsed, thresh, compare_mlg, original = mll(dat11, "custom")) %>%
+  mutate(sample_number = "observed") %>%
+  mutate(source = "observed")
+
+# Revisiting random dist
+map2_df(collapsed, thresh, compare_mlg, original = sample(87, 366, replace = TRUE)) %>%
+  ggplot(aes(x = threshold, y = N)) +
+  geom_point(alpha = 0.1) +
+  stat_smooth(se = FALSE, color = "black") +
+  scale_y_continuous(breaks = c(25, 50, 75)) +
+  stat_summary(pch = 21, fun.data = my_meanse, fill = "white")
+
+observed %>%
+  ggplot(aes(x = threshold, y = N)) +
+  geom_point(alpha = 0.1) +
+  stat_smooth(se = FALSE, color = "black") +
+  scale_y_continuous(breaks = c(25, 50, 75)) +
+  stat_summary(pch = 21, fun.data = my_meanse, fill = "white")
+```
+
+```{r randomizations, cache = TRUE}
+set.seed(5000)
+random   <- map_df(1:20, function(x) map2_df(collapsed, thresh, compare_mlg, original = sample(mll(dat11, "custom"))), .id = "sample_number") %>%
+  mutate(sample_number = sprintf("%3d", as.integer(sample_number))) %>%
+  mutate(source = "random")
+```
+
+```{r visualization, fig.width = 5, fig.height = 3}
+
+ggplot(random, aes(x = threshold, y = N, group = sample_number, color = source, fill = source)) +
+  # geom_point(alpha = 0.01, color = "grey50") +
+  stat_summary(alpha = 0.125, fun.data = my_meanse) +
+  # geom_point(alpha = 0.5, pch = 21, data = observed, fill = "white") +
+  stat_smooth(se = FALSE, data = observed) +
+  stat_summary(pch = 21, fun.data = my_meanse, data = observed) +
+  scale_color_manual(values = c("black", "grey50")) +
+  scale_fill_manual(values = c("white", "grey50")) +
+  theme_bw(base_size = 16, base_family = "Helvetica") +
+  theme(aspect.ratio = 1/2) +
+  theme(axis.text = element_text(color = "black")) +
+  theme(panel.grid.major = element_line(colour = "grey20")) +
+  theme(panel.grid.minor = element_line(colour = "grey50", linetype = 3)) +
+  theme(panel.grid.major.x = element_blank()) +
+  theme(panel.grid.minor.x = element_blank()) +
+  theme(panel.border = element_blank()) +
+  theme(legend.position = "top") +
+  ylab("Number of MCG/MLG") +
+  xlab("Bruvo's Distance")
+```
+
+# Assessing MCG/MLG community structure/cohesiveness
 

### 2017-07-27 doc/RMD/mlg-mcg.Rmd
remake

@@
 
-```{r mlg accumulation calculation}
+```{r mlg_accumulation_calculation}
 bd         <- bruvo.dist(dat11, replen = other(dat11)$REPLEN)
@@
 
-```{r mlg accumulation noclump, fig.width = 5, fig.height = 3}
+```{r mlg_accumulation_noclump, fig.width = 5, fig.height = 3}
 map2_df(collapsed, thresh, compare_mlg, original = sample(87, 366, replace = TRUE)) %>%
@@
 
-```{r mlg accumulation mlg, fig.width = 5, fig.height = 3}
+```{r mlg_accumulation_mlg, fig.width = 5, fig.height = 3}
 map2_df(collapsed, thresh, compare_mlg, original = mll(dat11, "original")) %>%
@@
 
-```{r mlg accumulation uniq, fig.width = 5, fig.height = 3}
+```{r mlg_accumulation_uniq, fig.width = 5, fig.height = 3}
 map2_df(collapsed, thresh, compare_mlg, original = seq(366)) %>%
@@
 
-```{r mlg accumulation ten, fig.width = 5, fig.height = 3}
+```{r mlg_accumulation_ten, fig.width = 5, fig.height = 3}
 map2_df(collapsed, thresh, compare_mlg, original = sample(10, 366, replace = TRUE)) %>%
@@
 
-```{r mlg accumulation observed, fig.width = 5, fig.height = 3}
+```{r mlg_accumulation_observed, fig.width = 5, fig.height = 3}
 observed <- map2_df(collapsed, thresh, compare_mlg, original = mll(dat11, "custom")) %>%

### 2017-07-27 doc/manuscript/manuscript.Rmd
add delet-able items to manuscript

@@
 Community analysis showed 51 communities, 15 of which consisted of a single MLG unconnected with any other community indicating that just 9.09% of the 165 MLGs are unable to cross with any other MLG in this data set.
-Of all the MLGs, MLG 25 had 57 neighbors that shared the same mycelial compatibility group. 
 The three communities with the most members contained 8 of the ten most abundant MCGs. 
 Comparing these communities with Bruvo's genetic distance revealed and average distance of 0.451 among communities and an average distance of 0.437 within communities, showing no significant difference.
+When we assessed the number of potential heterothallic parings between different MLGs in our data set (two different MLGs that are in the same mycelial compatibility group), we found an average of 16.5 potential heterothallic parings per MLG.
+MLG 75, representing just four isolates, had 57 neighbors that shared the same mycelial compatibility group. 
 Overall, there was no clear pattern to the association between MLG and MCGs.
@@
 Of particular interest in this study was testing the association between mycelial compatibility groups (MCGs) and multilocus genotypes (MLGs).
-In the past MCGs have been used as a proxy for clonal lineages [@otto-hanson2011variation; @kohn1995clonal; @carbone1999patterns; @lehner2015analysis]. 
+In the past MCGs have been used as a proxy for clonal lineages [@leslie1993fungal; @otto-hanson2011variation; @kohn1995clonal; @carbone1999patterns; @lehner2015analysis]. 
 When using genetic markers with MCGs, there are three basic relationship patterns, 1) One MLG to one MCG, 2) Several MLGs belong to one MCG, and 3) Several MCGs belong to one MLG, noting that 2 and 3 are not mutually exclusive.
 Similar to what has been observed in other studies, our results show all three patterns (Fig. \@ref(biggraph)) [@aldrich-wolfe2015genetic; @carbone1999patterns; @attanayake2012genetic]. 
-Because we took a bipartite network approach to assessing the relationship between MCG and MLG, we were able to assess just how concordant MCGs were with MLGs.
+In contrast to other studies, we took a network approach to assessing the relationship between MCG and MLG and found no clear causality between MLGs and MCGs.
 

### 2017-07-31 doc/manuscript/manuscript.Rmd
attempt to address methods in manuscript

@@
 
-## Isolate collection and DNA extraction {-}
+## Isolate collection {-}
+
+Several of the isolates used for this study were collected and reported in previous studies and were collected using the same methods [@otto-hanson2011variation].
+Broadly, isolates were collected from two sources, white mold screening nurseries (wmn), or producer fields.
+The white mold screening nurseries were 5m x 10m in size and maintained without applying fungicides to observe natural incidence of white mold.
+The nursery plots were incorporated with a basal dressing of N:P:K = 1:3:2 and side dressing of 0:3:2 during the growing season [@steadman2003identification; @steadman2004identification; @steadman2005identification; @steadman2006identification; @otto2007identification].
+Sampling was carried out by collecting diseased tissue in zig-zag transects across field plots. 
+Because sampling depended on disease incidence, the number of samples isolated varied from year to year.
+Although the nursery locations were the same over sampling years, sampling plots within a location varied for sampling years.
 
@@
 
-The white mold nurseries were 5m x 10m in size and maintained without applying fungicides to observe natural incidence of white mold.
-The nursery plots were incorporated with a basal dressing of N:P:K = 1:3:2 and side dressing of 0:3:2 during the growing season.
-Although the nursery locations were the same over sampling years, sampling plots within a location varied for sampling years.
-Sampling of diseased tissue was carried out on a 'W' or 'V' shape transect, or randomly.
-Because sampling depended on disease incidence, the number of samples isolated varied from year to year.
-In total, we collected 366 isolates that were used in the present study.
+Each isolate was regenerated by activating one or several sclerotia and transferring the leading edge of the mycelial mat according to the method described previously [@otto-hanson2011variation].
+In combination with the 156 isolates described previously, we collected 210 isolates for a total of 366 isolates [@otto-hanson2011variation].
 
@@
 N = Number of Isolates. 
-Key abbreviations: wmn = white mold nursery, unk = unknown.
+Key abbreviations: wmn = white mold screening nursery, unk = unknown.
 
-Each isolate was regenerated by activating one or several sclerotia according to the method described previously [@matheron1989in-vitro].
-Once actively growing, mycelial compatibility group (MCG) and aggressiveness was determined for each isolate.
-MCG was determined as described in @otto-hanson2011variation through co-culturing isolates in pairs on potato-dextrose agar (PDA) plates, where incompatibility of different MCGs resulted in formation of a mycelial barrier of dead tissue (barrage line) [@kohn1990mycelial].
-Since MCGs are considered mutually exclusive, pairing was not required between all isolates in the study.
+## Mycelial Compatibility {-}
 
-> SEE: Is that correct?  Perhaps consult with Becky or Rachana and then elaborate upon the sentence above.
+MCG was determined as described previously through co-culturing pairs of 2-day-old isolates 2.5cm apart on Diana Sermons (DS) Medium [@cubeta2001mycelial]. 
+Incompatibility of different MCGs resulted in formation of a mycelial barrier of dead tissue (referred to as a barrage line) accompanied by the formation of sclerotia on either side of the barrage line, indicating the limits of the isolates' growth [@otto-hanson2011variation; @kohn1990mycelial; @leslie1993fungal].
+Isolates were compared in a pairwise manner for each site and then representatives among sites were compared to determine mycelial compatibility groups by scoring compatible and incompatible interactions [@otto-hanson2011variation].
 
-Aggressiveness of each isolate was assessed using a straw test that rated necrotic lesion size on a scale from 1--9 with increasing aggressiveness [@petzoldt1996straw; @teran2006modified].
+<!-- ![
+Examples of mycelial compatibility results in *S. sclerotiorum*. 
+A) Compatible reaction showing aerial hyphae.
+B) Incompatible reaction showing barrage reaction of dead cells at point of contact.
+](../../results/figures/publication/incompatibility2.JPG){#incompatibility width=25%} -->
 
-> SEE: Rating is not clear from the single sentence above.  On what plant host was it evaluated?  In the greenhouse?  What is a rating of 1 compared to a rating of 9 on the straw test?
+## Aggressiveness {-}
+
+Aggressiveness of each isolate was assessed using a straw test as carried out in @otto-hanson2011variation that rated necrotic lesion size [@petzoldt1996straw; @teran2006modified].
+Briefly, the straw test uses 21-day-old G122 plants as the host in a greenhouse setting. 
+Clear drinking straws cut to 2.5cm and sealed were used to place two mycelial plugs of inoculum on the host plant after removing the everything &gt;2.5cm above the fourth node.
+Measurements were taken after 8 days using the Modified Petzoldt and Dickson scale where a score of 1--9 where 1 is no disease and 9 is plant death [@petzoldt1996straw; @teran2006modified].
+
+## Microsatellite genotyping {-}
 
@@
 > SEE: More details on how tissue was disrupted may be needed here.
-
-## Microsatellite genotyping {-}
 
@@
 Of particular interest in this study was testing the association between mycelial compatibility groups (MCGs) and multilocus genotypes (MLGs).
-In the past MCGs have been used as a proxy for clonal lineages [@leslie1993fungal; @otto-hanson2011variation; @kohn1995clonal; @carbone1999patterns; @lehner2015analysis]. 
+In the past MCGs have been used as a proxy for clonal lineages [@leslie1993fungal; @otto-hanson2011variation; @kohn1990mycelial; @schafer2006optimized; @kohn1995clonal; @carbone1999patterns; @lehner2015genetic].
 When using genetic markers with MCGs, there are three basic relationship patterns, 1) One MLG to one MCG, 2) Several MLGs belong to one MCG, and 3) Several MCGs belong to one MLG, noting that 2 and 3 are not mutually exclusive.
-Similar to what has been observed in other studies, our results show all three patterns (Fig. \@ref(biggraph)) [@aldrich-wolfe2015genetic; @carbone1999patterns; @attanayake2012genetic]. 
-In contrast to other studies, we took a network approach to assessing the relationship between MCG and MLG and found no clear causality between MLGs and MCGs.
+Similar to what has been observed in other studies, our results show all three patterns (Fig.
+\@ref(biggraph)) [@ford1995heterokaryon; @aldrich-wolfe2015genetic; @carbone1999patterns; @attanayake2012genetic].
+
+@jo2008reassessment; @papaioannou2014barrage; @micali2003independence;
 

### 2017-07-31 doc/manuscript/manuscript.Rmd
update MCG discussion

@@
 
-Of particular interest in this study was testing the association between mycelial compatibility groups (MCGs) and multilocus genotypes (MLGs).
-In the past MCGs have been used as a proxy for clonal lineages [@leslie1993fungal; @otto-hanson2011variation; @kohn1990mycelial; @schafer2006optimized; @kohn1995clonal; @carbone1999patterns; @lehner2015genetic].
-When using genetic markers with MCGs, there are three basic relationship patterns, 1) One MLG to one MCG, 2) Several MLGs belong to one MCG, and 3) Several MCGs belong to one MLG, noting that 2 and 3 are not mutually exclusive.
-Similar to what has been observed in other studies, our results show all three patterns (Fig.
-\@ref(biggraph)) [@ford1995heterokaryon; @aldrich-wolfe2015genetic; @carbone1999patterns; @attanayake2012genetic].
+Of interest in this study was testing the association between mycelial compatibility groups (MCGs) and multilocus genotypes (MLGs).
+In the past MCGs have been used as a proxy for clonal lineages [@leslie1993fungal; @otto-hanson2011variation; @kohn1990mycelial; @schafer2006optimized; @kohn1995clonal; @carbone1999patterns]. 
+Our results do not support this hypothesis (Fig. \@ref(biggraph)).
+Over the past few years, researchers have noticed inconsistencies among the relationship between MCGs and MLGs [@aldrich-wolfe2015genetic; @carbone1999patterns; @attanayake2012genetic; @lehner2015genetic].
+Either several MCGs belong to one MLG, which could be explained by insufficient sampling of loci; several MLGs belong to one MCG, which could be explained by clonal expansion; or a mixture of both.
+Some studies have shown a correlation between MCG and MLG [@aldrich-wolfe2015genetic; @carbone1999patterns; @lehner2015genetic], whereas other studies have shown no apparent correlation, even on small spatial scales [@attanayake2012genetic; @attanayake2013sclerotinia].
 
-@jo2008reassessment; @papaioannou2014barrage; @micali2003independence;
+One long-held assumption was that MCGs (as determined via barrage reaction) represent vegetative compatibility groups (VCGs) [@schafer2006optimized; @kohn1990mycelial; @lehner2015genetic], which are known to have a genetic component [@saupe2000molecular; @hall2010evolution; @strom2016genomes].
+While our results appear to contradict this assumption, our protocol for assessing MCGs utilized a different substrate for these reactions and, thus may not accurately reflect the true MCG dynamics. 
+However, it has been demonstrated in several Ascomycetes---including *Neurospora crassa* [@micali2003independence], *Sclerotinia homoecarpa* [@jo2008reassessment], *Verticillium dahliae* [@papaioannou2014barrage], and *S.
+sclerotiorum* [@ford1995heterokaryon]---that barrage reactions are independent from stable anastomosis.
+Thus, the inconsistencies in this study and other studies indicate that researchers studying *S.
+sclerotiorum* should not rely on MCG data derived from barrage reactions as an indicator for genetic diversity.
 

### 2017-08-01 doc/manuscript/manuscript.Rmd
add discussion on dbRDA

@@
 
-## Phenotypic variable assessment {-}
+## Phenotypic Variable Assessment {-}
 
@@
 Assessment of the marginal effects showed that state, host, and year significantly explained genetic variation (P = 0.001), but aggressiveness was only mildly significant (P = 0.045). 
-Much of the variation appears to be driven by isolates from Mexico and 2005 (Fig. \@ref(rda-biplot)); it's worth noting that all Mexican isolates were collected in 2005.
+Much of the variation appears to be driven by isolates from Mexico and 2005 (Fig. \@ref(rda-biplot)).
 Variance partitioning of the independent variables indicated aggressiveness to be the least influential factor with 0.1% contributing to explaining the variation of molecular data whereas the combination of variables accounted for 3.3%.
@@
 
+## Phenotypic Variable Importance
+
+Borrowing a technique often used in the ecological literature, we used dbRDA to elucidate the effect of all variables (Aggressiveness, Region, Source,   Year, and Host) [@legendre1999distance].
+From the initial results, it appears that the most important factors for predicting genetic structure are region and year.
+While the results indicated that Mexico and the year 2005 were the two most important variables, it's worth noting that all Mexican isolates were collected in 2005.
+The results also show that the Vista cultivar explains some of the variance, but this represents six isolates in MI. 
+Aggressiveness and specific source field had little to no effect on genetic diversity. 
+These results are in agreement with studies that examined differentiation based on host [@aldrich-wolfe2015genetic] and severity [@atallah2004high; @attanayake2012genetic; @attanayake2013sclerotinia] reporting little or no correlation of genetic diversity to these variables.
+This indicates that a) breeders should keep in mind regional differences when assessing resistance and b) it is possible that we have not yet measured biologically relevant variables that can predict genetic differentiation, which could include variables such as soil community composition.
+
+## Mycelial Compatibility Groups and Multillocus Genotypes
+
 Of interest in this study was testing the association between mycelial compatibility groups (MCGs) and multilocus genotypes (MLGs).
@@
 Either several MCGs belong to one MLG, which could be explained by insufficient sampling of loci; several MLGs belong to one MCG, which could be explained by clonal expansion; or a mixture of both.
-Some studies have shown a correlation between MCG and MLG [@aldrich-wolfe2015genetic; @carbone1999patterns; @lehner2015genetic], whereas other studies have shown no apparent correlation, even on small spatial scales [@attanayake2012genetic; @attanayake2013sclerotinia].
+Some studies have shown a correlation between MCG and MLG [@aldrich-wolfe2015genetic; @carbone1999patterns; @lehner2015genetic], whereas other studies have shown no apparent correlation, even on small spatial scales [@atallah2004high; @attanayake2012genetic; @attanayake2013sclerotinia].
 
@@
 sclerotiorum* should not rely on MCG data derived from barrage reactions as an indicator for genetic diversity.
+
+## Genotypic Structure
+
+## Population Differentiation
 

### 2017-08-01 doc/manuscript/manuscript.Rmd
fix some little things

@@
 
-## Phenotypic Variable Importance
+## Phenotypic Variable Importance {-}
 
@@
 From the initial results, it appears that the most important factors for predicting genetic structure are region and year.
-While the results indicated that Mexico and the year 2005 were the two most important variables, it's worth noting that all Mexican isolates were collected in 2005.
+While the results indicated that Mexico and the year 2005 were the two most important variables, it's worth noting that all Mexican isolates were collected in 2005 (Fig. \@ref(rda-biplot)).
 The results also show that the Vista cultivar explains some of the variance, but this represents six isolates in MI. 
@@
 
-## Mycelial Compatibility Groups and Multillocus Genotypes
+## Mycelial Compatibility Groups and Multillocus Genotypes {-}
 
-Of interest in this study was testing the association between mycelial compatibility groups (MCGs) and multilocus genotypes (MLGs).
-In the past MCGs have been used as a proxy for clonal lineages [@leslie1993fungal; @otto-hanson2011variation; @kohn1990mycelial; @schafer2006optimized; @kohn1995clonal; @carbone1999patterns]. 
-Our results do not support this hypothesis (Fig. \@ref(biggraph)).
+Of interest in this study was testing the association between multilocus genotypes (MLGs) and mycelial compatibility groups (MCGs), as the latter have historically been used as a proxy for clonal lineages [@leslie1993fungal; @otto-hanson2011variation; @kohn1990mycelial; @schafer2006optimized; @kohn1995clonal; @carbone1999patterns].
+Our results do not support this assumption (Fig. \@ref(biggraph)).
 Over the past few years, researchers have noticed inconsistencies among the relationship between MCGs and MLGs [@aldrich-wolfe2015genetic; @carbone1999patterns; @attanayake2012genetic; @lehner2015genetic].
@@
 
-## Genotypic Structure
+## Genotypic Structure {-}
 
-## Population Differentiation
+## Population Differentiation {-}
+
+## Conclusions {-}
 

### 2017-08-01 doc/manuscript/manuscript.Rmd
add more text to discussion (probably to be deleted)

@@
 Of interest in this study was testing the association between multilocus genotypes (MLGs) and mycelial compatibility groups (MCGs), as the latter have historically been used as a proxy for clonal lineages [@leslie1993fungal; @otto-hanson2011variation; @kohn1990mycelial; @schafer2006optimized; @kohn1995clonal; @carbone1999patterns].
-Our results do not support this assumption (Fig. \@ref(biggraph)).
+Our results do not support this assumption.
+It can be seen in Fig.
+\@ref(barplot) that the most abundant MCG contains several MLGs, but the diversity of those MLGs are low as indicated by the evenness (transparency), which indicates that there is one dominant MLG (MLG 25).
+What is not shown in Fig. \@ref(barplot) is the MLGs that are shared between MCGs. 
+This is illustrated in both table \@ref(tab:mlg-table) and Fig. \@ref(biggraph). 
+It could be argued, however that MLG 25, with its high value of $P_{sex}$ represents different true MLGs across the five MCGs it occupies, but this doesn't account for the overall structure of Fig.
+\@ref(biggraph) where, for example, MLG 75 ($P_{sex} = 1.81e^-4$) is compatible with 58 other genotypes through three MCG when the population structure of *S.
+sclerotiorum* is known to be clonal.
+
 Over the past few years, researchers have noticed inconsistencies among the relationship between MCGs and MLGs [@aldrich-wolfe2015genetic; @carbone1999patterns; @attanayake2012genetic; @lehner2015genetic].
@@
 ## Genotypic Structure {-}
+
+The results show several genotypes shared across populations, except for Mexico and two states that had fewer than three samples each (Fig. \@ref(community-graph)).
+Our network-based approach by treating the genotypes as edges and weighting each edge with the inverse of $P_{sex}$ treated the edges as springs connecting the populations with the strength proportional to the probability of obtaining the same genotype as a clone.
 

### 2017-08-01 doc/RMD/mlg-mcg.Rmd
add more to the genotypic structure section

@@
 }
+
+cat("Community Structure:\n")
+cluster_walktrap(g)
 ```

### 2017-08-01 doc/manuscript/manuscript.Rmd
add more to the genotypic structure section

@@
 Our network-based approach by treating the genotypes as edges and weighting each edge with the inverse of $P_{sex}$ treated the edges as springs connecting the populations with the strength proportional to the probability of obtaining the same genotype as a clone.
+This allowed us to use a graph walking algorithm to see how close the populations were simply based off of the proportion of clones they shared.
+The most abundant genotype was shared across four populations, but its high value of $P_{sex}$ meant that it did not contribute significantly to the overall structure. 
+The graph walking algorithm was able to divide the network into three groups, but had a modularity of 0.17, which indicates that the groups are weakly differentiated.
+This indicates that multilocus genotypes are widespread across the United States, giving more evidence for a clonal population structure.
 

### 2017-08-01 doc/RMD/wmn-differentiation.Rmd
export AMOVA table

@@
 ssc_amova_nm <- poppr.amova(datnomex, ~SourceType/Region, dist = bdnm, quiet = TRUE)
-ssc_amova_nm_test <- randtest(ssc_amova_nm, nrepet = 999)
+ssc_amova_nm_test <- randtest(ssc_amova_nm, nrepet = 9999)
 plot(ssc_amova_nm_test)
@@
   add_column(Hierarchy = c("Between Source", "Between Region within Source", "Within Region"), .before = 1) %>%
+  readr::write_csv(path = file.path("results", "tables", "AMOVA-region.csv"), col_names = TRUE) %>%
   rename(ps = `Phi statistic`) %>%
   mutate(ps = gsub("0\\.00(\\d{1})(\\d{2})", "\\1.\\2e^-3^", ps)) %>%
+  mutate(ps = case_when(Hierarchy == "Between Source" ~ ps, TRUE ~ paste0("**", ps, "**"))) %>%
   rename(`$\\Phi statistic$` = ps) %>%
@@
   huxtable::as_huxtable(add_colnames = TRUE) %>% 
-  huxtable::set_col_width(c(1.3, 0.7, 0.9, 1, 1.3)) %>% 
+  huxtable::set_col_width(c(1.1, 0.6, 0.9, 0.9, 1.4)) %>% 
   huxtable::set_align(huxtable::everywhere, 2:5, "center") %>% 
-  huxtable::print_md(max_width = 95)
+  huxtable::print_md(max_width = 100)
 ```
@@
 wmndapc <- xvalDapc(tab(wmn_compare), pop(wmn_compare), n.rep = 200, n.pca = 1:10)
-wmndapc
+wmndapc$DAPC
 scatter(wmndapc$DAPC)
@@
   theme(legend.position = "top")
+apply(bebug_dapc$posterior, 2, mean)
 ```

### 2017-08-01 doc/RMD/MCG-virulence.Rmd
Add aggressiveness tests to fix #8

@@
 ```{r load_data}
-cols(
+datcols <- cols(
   .default = col_integer(),
@@
 )
-dat <- read_csv(file.path(PROJHOME, "data", "clean_data.csv")) %>%
-  select(MCG, Severity)
+dat <- read_csv(file.path(PROJHOME, "data", "clean_data.csv"), col_types = datcols) %>%
+  select(Severity, MCG, Region, Source)
 dat
@@
   inner_join(dat, by = "MCG") %>%
-  select(-N) %>%
+  select(MCG, Severity) %>%
   mutate(MCG = forcats::fct_inorder(as.character(MCG)))
@@
   geom_point(position = position_jitter(width = 0.25), alpha = 0.5) +
-  scale_y_continuous(limits = c(0, NA)) +
+  scale_y_continuous(limits = c(1, 9), breaks = c(1, 3, 5, 7, 9), expand = c(0, 0.1)) +
   # scale_x_discrete(position = "top") +
@@
 
+# Assessing Virulence by Region
+
+Since this was done in Otto-Hanson *et al.*, it would be a good idea for us to 
+assess this as well.
+
+```{r fig.width = 5, fig.height = 3.75}
+set.seed(2017-06-29)
+region_dat <- dat %>%
+  group_by(Region) %>%
+  filter(n() > 5) %>%
+  mutate(mean_sev = mean(Severity)) %>%
+  mutate(Source = ifelse(Source == "wmn", "wmn", "other")) %>%
+  arrange(desc(mean_sev)) %>%
+  ungroup() %>%
+  mutate(Region = forcats::fct_inorder(Region)) %>%
+  select(Region, Severity, Source)
+ggplot(region_dat, aes(x = Region, y = Severity)) +
+  geom_point(aes(fill = Source), 
+             pch = 21, 
+             position = position_jitter(width = 0.25), 
+             alpha = 0.35) +
+  geom_boxplot(alpha = 0.5) +
+  scale_y_continuous(limits = c(1, 9), breaks = c(1, 3, 5, 7, 9), expand = c(0, 0.1)) +
+  scale_fill_manual(values = rev(c("black", "white"))) +
+  theme_bw(base_size = 16, base_family = "Helvetica") +
+  theme(aspect.ratio = 1/2) +
+  theme(axis.text = element_text(color = "black")) +
+  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
+  theme(panel.grid.major = element_line(colour = "grey20")) +
+  theme(panel.grid.minor = element_line(colour = "grey50")) +
+  theme(panel.grid.major.x = element_blank()) +
+  theme(panel.border = element_blank()) +
+  theme(legend.position = "top") +
+  theme(plot.margin = unit(c(0, 0, 0, 0), "lines")) +
+  labs(list(
+    y = "Aggressiveness",
+    x = "Region"
+    ))
+if (!interactive()){
+  ggsave(file.path("results", "figures", "publication", "aggressiveness.pdf"),
+         width = 88, units = "mm")
+}
+```
+
+```{r anova2, fig.width = 4, fig.height = 8}
+region_dat %>%
+  group_by(Region) %>%
+  summarize(N = n(), 
+            `Min Aggressiveness` = min(Severity),
+            `Max Aggressiveness` = max(Severity),
+            `Average Aggressiveness` = mean(Severity),
+            `Median Aggressiveness` = median(Severity)
+            ) %>%
+  knitr::kable(digits = 2)
+ANOVA <- aov(Severity ~ Region, data = region_dat)
+ANOVA
+summary(ANOVA)
+plot(TukeyHSD(ANOVA, conf.level = 0.95), las = 2)
+grps <- agricolae::HSD.test(ANOVA, "Region", alpha = 0.05)$groups %>% 
+  dplyr::rename(Region = trt) %>%
+  dplyr::rename(Group = M) %>%
+  dplyr::rename(`Mean Aggressiveness` = means)
+
+readr::write_csv(grps, path = "results/tables/Region-aggressiveness.csv", col_names = TRUE)
+
+grps %>%
+  huxtable::as_huxtable(add_colnames = TRUE) %>%
+  huxtable::set_number_format(row = huxtable::everywhere, col = 1, 0) %>% 
+  huxtable::set_col_width(c(0.13, 0.25, 0.08)) %>%
+  huxtable::set_align(row = huxtable::everywhere, col = 2, "right") %>%
+  huxtable::print_md(max_width = 40)
+```
+
+```{r fig.width = 5, fig.height = 3.5}
+library("poppr")
+load(file.path("data", "sclerotinia_16_loci.rda"))
+strat <- strata(dat11) %>% 
+  add_column(MLG = mll(dat11)) %>%
+  add_column(Severity = other(dat11)$meta$Severity)
+top_mlg <- strat %>% 
+  group_by(MLG) %>%
+  summarize(N = n()) %>%
+  arrange(desc(N)) %>%
+  slice(1:10) %>% 
+  inner_join(strat, by = "MLG") %>%
+  select(MLG, Severity) %>%
+  mutate(MLG = forcats::fct_inorder(as.character(MLG)))
+set.seed(2017-06-29)
+ggplot(top_mlg, aes(x = MLG, y = Severity)) +
+  geom_boxplot() +
+  geom_point(position = position_jitter(width = 0.25), alpha = 0.5) +
+  scale_y_continuous(limits = c(1, 9), breaks = c(1, 3, 5, 7, 9), expand = c(0, 0.1)) +
+  # scale_x_discrete(position = "top") +
+  theme_bw(base_size = 16, base_family = "Helvetica") +
+  theme(aspect.ratio = 1/2) +
+  theme(axis.text = element_text(color = "black")) +
+  # theme(axis.ticks.x = element_blank()) +
+  theme(panel.grid.major = element_line(colour = "grey20")) +
+  theme(panel.grid.minor = element_line(colour = "grey50")) +
+  theme(panel.grid.major.x = element_blank()) +
+  theme(panel.border = element_blank()) +
+  labs(list(
+    # title = "Aggressiveness for the top 10 MCGs",
+    y = "Aggressiveness",
+    x = "Mulitlocus Genotype"
+    ))
+```
+
+```{r anova3, fig.width = 4, fig.height = 8}
+top_mlg %>%
+  group_by(MLG) %>%
+  summarize(N = n(), 
+            `Min Aggressiveness` = min(Severity),
+            `Max Aggressiveness` = max(Severity),
+            `Average Aggressiveness` = mean(Severity)
+            ) %>%
+  knitr::kable(digits = 2)
+ANOVA <- aov(Severity ~ MLG, data = top_mlg)
+ANOVA
+summary(ANOVA)
+plot(TukeyHSD(ANOVA, conf.level = 0.95), las = 2)
+grps <- agricolae::HSD.test(ANOVA, "MLG", alpha = 0.05)$groups %>% 
+  dplyr::rename(MLG = trt) %>%
+  dplyr::rename(Group = M) %>%
+  dplyr::rename(`Mean Aggressiveness` = means)
+
+readr::write_csv(grps, path = "results/tables/MLG-aggressiveness.csv", col_names = TRUE)
+
+grps %>%
+  huxtable::as_huxtable(add_colnames = TRUE) %>%
+  huxtable::set_number_format(row = huxtable::everywhere, col = 1, 0) %>% 
+  huxtable::set_col_width(c(0.13, 0.25, 0.08)) %>%
+  huxtable::set_align(row = huxtable::everywhere, col = 2, "right") %>%
+  huxtable::print_md(max_width = 40)
+```
+
 <details>

### 2017-08-01 doc/manuscript/manuscript.Rmd
rearrange manuscript
 - remove "phenotypic"
 - make AMOVA table auto-generating
 - Update aggressiveness assessment (#8)
 - narrow widths for half-page plots (to account for center column)
 - add aggressiveness figure
 - change aggressiveness to be region-focused
 - move MCG distribution figure to supplemental
 - Move MLG-MCG correlation to Variable Assessment
 - Add statements about cultivar DAPC results
 - begin discussion onf population differentiation

@@
 
-## Assessing importance of categorical and phenotypic variables  {-}
+## Assessing Importance of Variables  {-}
 
-### Distance-based redundancy analysis {-}
+### Distance-based Redundancy Analysis {-}
 
@@
 
-### Aggressiveness assessment {-}
+### Aggressiveness Assessment {-}
 
-We used analysis of variance (ANOVA) to assess if MCGs differed in aggressiveness.
-To minimize complications due to small sample sizes, we chose the top 10 MCGs, representing 57% of the isolates collected.
-Between-MCG differences were assessed using Tukey's HSD test with an alpha of 0.05 using the `HSD.test()` function in the package *agricolae* version 1.2.4 [@mendiburu2015agricolae].
+We used analysis of variance (ANOVA) to assess if aggressiveness could be explained by Region, mycelial compatibility group, or multilocus genotype. 
+To minimize complications due to small sample sizes, we chose the top 10 MCGs, representing 56.5% of the isolates collected, the top 10 MLGs representing 26.7% of the isolates, and populations with more than five isolates.
+Between-group differences were assessed using Tukey's HSD test with an alpha of 0.05 using the `HSD.test()` function in the package *agricolae* version 1.2.4 [@mendiburu2015agricolae].
 
@@
 Four regions had fewer than 10 isolates, Australia (6), WI (2), NY (1), ID (1).
+We observed 87 mycelial compatibility groups (MCG), the most abundant of which (MCG 5) represented 73 isolates over 37 multilocus genotypes (MLG) (Fig \@ref(barplot)).
 After clone-correction on the hierarchy of Region/Source/Host/Year, we were left with 318 isolates representing 165 MLGs (Table \@ref(tab:iatab)).
@@
 
-## Phenotypic Variable Assessment {-}
+## Variable Assessment {-}
 
@@
 
-### Mycelial Compatibility Groups {-}
-
-We observed 87 mycelial compatibility groups (MCG), the most abundant of which (MCG 5) represented 73 isolates over 37 multilocus genotypes (MLG) (Fig. \@ref(barplot)). 
-Our assessment of aggressiveness association with MCGs showed a significant effect (P < 0.001) indicating that MCG 44 is, on average, more aggressive than other MCGs (Table \@ref(tab:mcg-aggressiveness)).
-
 <!-- RDA figure -->
@@
 Points represent the 318 clone-corrected samples in ordination space. 
-](../../results/figures/publication/Figure7Z.pdf){#rda-biplot width=50%}
+](../../results/figures/publication/Figure7Z.pdf){#rda-biplot width=45%}
 
@@
 
-<!-- MCG abundance figure -->
+### Aggressiveness {-}
+
+<!-- aggressiveness figure -->
 
 ![
-Barplot showing abundance of each Mycelial Compatibility Group (MCG) in descending order.
-Singletons (46) have been removed for clarity, leaving 41 MCGs.
-White bars represent sample counts and grey bars represent counts of unique multilocus genotypes (MLG).
-The transparency associated with the grey bars represent the evenness of the distribution of the MLGs within a given MCG as shown in the inset histograms.
-](../../results/figures/publication/Figure1Z.pdf){#barplot width=100%}
+Boxplot representing aggressiveness (vertical axis) by population (horizontal axis) arranged in descending order of mean aggressiveness for all populations with N > 5.
+Box boundaries represent 25% and 75% percentiles and whiskers extend to at most 1.5 $\times$ the interquartile range.
+Horizontal line represents median value.
+Points represent individual observations whose colors indicate if they came from white mold screening nurseries (wmn) or not (other).
+](../../results/figures/publication/aggressiveness.pdf){#aggressiveness-boxplot width=45%}
 
 <!-- end figure -->
+
+Our assessment of aggressiveness in association with Region showed a significant effect (P < 1.00e^-4^), with means that ranged from 5.8 (MN) to 4.0 (CA) (Fig.
+\@ref(aggressiveness-boxplot), Table \@ref(tab:region-aggressiveness)).
+MCGs also showed a significant effect (P < 1.00e^-4^), with means that ranged from 6.0 (MCG 44) to 4.6 (MCG 49).
+We additionally found a significant effect for MLGs (P = 7.44e^-4^) for MLGs, with means that ranged from 6.0 (MLG 78) to 4.3 (MLG 140).
+
+### Correlation of Mulitlocus Genotypes and Mycelial Compatibility Groups {-}
+
+Our network-based approach to correlating MLGs with MCGs revealed a large and complex network (Fig. \@ref(biggraph)). 
+Community analysis showed 51 communities, 15 of which consisted of a single MLG unconnected with any other community indicating that just 9.09% of the 165 MLGs are unable to cross with any other MLG in this data set.
+The three communities with the most members contained 8 of the ten most abundant MCGs. 
+Comparing these communities with Bruvo's genetic distance revealed and average distance of 0.451 among communities and an average distance of 0.437 within communities, showing no significant difference.
+When we assessed the number of potential heterothallic parings between different MLGs in our data set (two different MLGs that are in the same mycelial compatibility group), we found an average of 16.5 potential heterothallic parings per MLG.
+MLG 75, representing just four isolates, had 57 neighbors that shared the same mycelial compatibility group. 
+Overall, there was no clear pattern to the association between MLG and MCGs.
+
+<!-- MCG graph figure -->
+
+![
+Graph showing complex associations between Mycelial Compatibility Groups (dotted nodes) and Multilocus Genotypes (full nodes). 
+Node size reflect the number of samples represented by each node (circle).
+Edges (arrows) point from Mycelial Compatibility Group to Multilocus Genotype where the weight (thickness) of the edge represents the number of samples shared.
+Node color represents the community assignment based on the walktrap algorithm with a maximum of four steps [@pons2006computing].
+](../../results/figures/publication/FigureS2.pdf){#biggraph width=100%}
+
+<!-- end fig -->
 
@@
 \@ref(community-graph16)).
-The international community appears to be driven by MLG 4, which is shared between all three populations and has a $P_{sex}$ value of 2.87 x 10^-5^, in contrast to the abundant MLG 25, which has a $P_{sex}$ value of 0.0168.
+The international community appears to be driven by MLG 4, which is shared between all three populations and has a $P_{sex}$ value of 2.87e^-5^, in contrast to the abundant MLG 25, which has a $P_{sex}$ value of 0.0168.
 
@@
 
-Our network-based approach to correlating MLGs with MCGs revealed a large and complex network (Fig. \@ref(biggraph)). 
-Community analysis showed 51 communities, 15 of which consisted of a single MLG unconnected with any other community indicating that just 9.09% of the 165 MLGs are unable to cross with any other MLG in this data set.
-The three communities with the most members contained 8 of the ten most abundant MCGs. 
-Comparing these communities with Bruvo's genetic distance revealed and average distance of 0.451 among communities and an average distance of 0.437 within communities, showing no significant difference.
-When we assessed the number of potential heterothallic parings between different MLGs in our data set (two different MLGs that are in the same mycelial compatibility group), we found an average of 16.5 potential heterothallic parings per MLG.
-MLG 75, representing just four isolates, had 57 neighbors that shared the same mycelial compatibility group. 
-Overall, there was no clear pattern to the association between MLG and MCGs.
-
-<!-- MCG graph figure -->
-
-![
-Graph showing complex associations between Mycelial Compatibility Groups (dotted nodes) and Multilocus Genotypes (full nodes). 
-Node size reflect the number of samples represented by each node (circle).
-Edges (arrows) point from Mycelial Compatibility Group to Multilocus Genotype where the weight (thickness) of the edge represents the number of samples shared.
-Node color represents the community assignment based on the walktrap algorithm with a maximum of four steps [@pons2006computing].
-](../../results/figures/publication/FigureS2.pdf){#biggraph width=100%}
-
-<!-- end fig -->
+\newpage
 
@@
 Because Mexico shared no genotypes with any of the other populations and we suspected that it was contributing to some of the observed variance, we re-ran AMOVA without the Mexican isolates and found similar decreased, but still significant levels of differentiation between regions.
-In contrast, when we compared the three cultivars, Beryl, Bunsi, and G122, we found no evidence of population differentiation (Supplementary Data).
+In contrast, when we compared the three cultivars, Beryl, Bunsi, and G122, we found no significant differentiation (Supplementary Data).
 
--------------------------------------------------------------------------------
-Hierarchy               d.f.        S.S.       % variation    $\Phi$ statistic  
--------------------- ---------- ------------- -------------- ------------------
-Between Source         1 (1)    0.897 (0.753) -0.158 (0.116)   0.136 (0.0965)
-
-Between Region        20 (19)     12 (8.91)    13.8 (9.53)     **0.138 (0.0955)**   
-within Source
-
-Within Region        296 (282)   56.7 (55.1)   86.4 (90.3)   **-1.58e^-3^ (1.16e^-3^)**
-
--------------------------------------------------------------------------------
+```{r table4, echo = FALSE, warning = FALSE, message = FALSE, results = "asis"}
+suppressPackageStartupMessages(library("tidyverse"))
+read_csv("../../results/tables/AMOVA-region.csv", 
+         na = "NA", # spaces read in as spaces
+         col_types = cols(.default = col_character()))  %>%
+  rename(ps = `Phi statistic`) %>%
+  mutate(ps = gsub("0\\.00(\\d{1})(\\d{2})", "\\1.\\2e^-3^", ps)) %>%
+  mutate(ps = case_when(Hierarchy == "Between Source" ~ ps, TRUE ~ paste0("**", ps, "**"))) %>%
+  rename(`$\\Phi statistic$` = ps) %>%
+  rename(`% variation` = `Percent variation`) %>%
+  rename(S.S. = `Sum of Squares`) %>%
+  select(-P) %>%
+  huxtable::as_huxtable(add_colnames = TRUE) %>% 
+  huxtable::set_col_width(c(1.1, 0.6, 0.9, 0.9, 1.4)) %>% 
+  huxtable::set_align(huxtable::everywhere, 2:5, "center") %>% 
+  huxtable::print_md(max_width = 100)
+```
 
@@
 The hierarchy was constructed as Source/Region where source is defined as belonging to a white mold nursery or not.
-Bold $\Phi$ values indicate significant values.
+Bold $\Phi$ values indicate significant values (P < 1.00e^-4^).
 Numbers in parentheses represent AMOVA run without Mexican isolates included.
@@
 
-Discriminant Analysis of Principal Components (DAPC) was performed with the first 21 principal components, representing 88% of the total variance. 
+Discriminant Analysis of Principal Components (DAPC) was performed grouping by Region with the first 21 principal components, representing 88% of the total variance. 
 The first discriminant axis (representing 64% of the discriminatory power) separated the centroid for the Mexico isolates from the rest of the data, indicating strong differentiation (Fig.
@@
 \@ref(DAPC)a).
+DAPC grouping by cultivar used the first 20 principal components, representing 89% of the total variance.
+The first two discriminant axes failed to separate any of the cultivars where the mean posterior assignment probabilities were 33.97% (G122), 35.95% (Beryl), and 30.08% (Bunsi).
 
@@
 
-## Phenotypic Variable Importance {-}
+## Variable Importance {-}
 
@@
 It could be argued, however that MLG 25, with its high value of $P_{sex}$ represents different true MLGs across the five MCGs it occupies, but this doesn't account for the overall structure of Fig.
-\@ref(biggraph) where, for example, MLG 75 ($P_{sex} = 1.81e^-4$) is compatible with 58 other genotypes through three MCG when the population structure of *S.
+\@ref(biggraph) where, for example, MLG 75 ($P_{sex}$ = 1.81e^-4^) is compatible with 58 other genotypes through three MCG when the population structure of *S.
 sclerotiorum* is known to be clonal.
@@
 
-## Population Differentiation {-}
+## Regional Differentiation {-}
+
+The primary interest of this study was to assess if isolates sampled from white mold screening nurseries represent isolates from producer fields within the region [@steadman2003identification; @otto-hanson2011variation].
+According to our AMOVA results, we have evidence for differentiation at the region level, but little to no differentiation within region (Table \@ref(tab:amova)).
 
@@
 suppressPackageStartupMessages(library("tidyverse"))
-read_csv("../../results/tables/MCG-aggressiveness.csv", 
+read_csv("../../results/tables/Region-aggressiveness.csv", 
          na = "NA", # spaces read in as spaces
          col_types = cols(.default = col_character())) %>%
-  dplyr::rename(MCG = trt) %>%
-  dplyr::rename(Group = M) %>%
-  dplyr::rename(`Mean Aggressiveness` = means) %>%
   huxtable::as_huxtable(add_colnames = TRUE) %>%
   huxtable::set_number_format(row = huxtable::everywhere, col = 1, 0) %>% 
-  huxtable::set_col_width(c(0.05, 0.3, 0.08)) %>%
+  huxtable::set_col_width(c(0.13, 0.25, 0.08)) %>%
   huxtable::set_align(row = huxtable::everywhere, col = 2, "right") %>%
-  huxtable::print_md(max_width = 31)
+  huxtable::print_md(max_width = 40)
 ```
 
-Table: (\#tab:mcg-aggressiveness) Mean aggressiveness ratings for the ten most abundant mycelial compatibility groups (MCG) with groupings according to 95% family-wise confidence interval.
-
-\newpage
+Table: (\#tab:region-aggressiveness) Mean aggressiveness ratings for the Regions with greater than 5 samples; groupings according to 95% family-wise confidence interval.
 
@@
 
+<!-- MCG abundance figure -->
+
+![
+Barplot showing abundance of each Mycelial Compatibility Group (MCG) in descending order.
+Singletons (46) have been removed for clarity, leaving 41 MCGs.
+White bars represent sample counts and grey bars represent counts of unique multilocus genotypes (MLG).
+The transparency associated with the grey bars represent the evenness of the distribution of the MLGs within a given MCG as shown in the inset histograms.
+](../../results/figures/publication/Figure1Z.pdf){#barplot width=100%}
+
+<!-- end figure -->
+
+\newpage
+
 <!-- community graph figure -->

### 2017-08-01 doc/RMD/table-1.Rmd
use the correct flags this time :see_no_evil:
change table 1 to be more concise
I changed table 1 so that it highlights the salient features of
our data.

@@
 
-```{r load_packages, warnings = FALSE, messages = FALSE}
+```{r load_packages, warning = FALSE, message = FALSE}
 library("readr")
@@
   mutate(Region = ifelse(nchar(Region) == 2, Region, "-")) %>%    # Adding blanks for international "states"
+  mutate(Source = ifelse(Source == "wmn", "wmn", "producer")) %>% # Changing fields to binary counts
   select(Country, Region, Source, Year, Host) %>%                 # Summarizing the number of isolates

### 2017-08-01 doc/manuscript/manuscript.Rmd
change table 1 to be more concise
I changed table 1 so that it highlights the salient features of
our data.

@@
 
-USA       CO     gree              2007, 2010 Pinto, Yellow                14
-
-                 luc               2007, 2010 Yellow                       14
-
-                 eat               2007, 2010 Yellow                       13
+USA       CO     producer          2007, 2010 Pinto, Yellow                41
 
@@
 
-USA       ID     unk                     2003 GH                            1
+USA       ID     producer                2003 GH                            1
 
@@
                                    2008, 2009 Beryl, Bunsi, cornell,       
-                                              G122, ion, PO7863, WM31
+                                              G122, Orion, PO7863, WM31
 
-                 tuc               2008, 2009 Black, Fuji, Vista            6
-
-                 hur               2008, 2009 BL, Black, SR06233, Vista     5
-
-                 sanc                    2008 Merlot                        3
-
-                 unk               2003, 2009 GH, unk                       2
-
-                 eln                     2009 unk                           1
-
-                 mung                    2009 Zorro                         1
-
-                 Sebewaing               2009 Black                         1
+                 producer    2003, 2008, 2009 BL, Black, Fuji, GH,         19 
+                                              Merlot, SR06233, unk,        
+                                              Vista, Zorro
 
@@
 
-USA       ND     rrv                     2007 unk                          21
-
-                 gfc               2007, 2010 unk                           9
-
-                 tlc                     2010 unk                           9
-
-                 pmc                     2010 unk                           7
+USA       ND     producer          2007, 2010 unk                          53
 
                  wmn                     2005 Beryl, Bunsi, G122            7
-
-                 wlc                     2010 unk                           4
-
-                 stc                     2010 unk                           2
-
-                 nec                     2010 unk                           1
 
@@
 
-                 mtc               2007, 2010 Beryl, Emerson, Orion        12
+                 producer   2003, 2007, 2009, Beryl, Emerson, GH, Orion,   20 
+                                         2010  Pinto, Weihing
 
-                 sbf               2007, 2009 Beryl, Pinto, Weihing         6
-
-                 unk                     2003 GH                            2
-
-USA       NY     unk                     2003 GH                            1
+USA       NY     producer                2003 GH                            1
 
@@
 
-                 corv                    2003 G122                          1
-
-                 unk                     2003 GH                            1
+                 producer                2003 G122, GH                      2
 
 USA       WA     wmn        2003, 2004, 2005, 11A, 37, 38, Beryl, Bunsi,   36 
-                                         2008  cornell, G122, Orion,          
-                                              PO7104, PO7863, WM31
+                                         2008  cornell, G122, Orion, PO7      
+                                              104, PO7863, WM31
 
-                 prsr                    2007 Merlot, Pinto, redkid        22
+                 producer          2003, 2007 GH, Merlot, Pinto, redkid    23
 
-                 unk                     2003 GH                            1
-
-USA       WI     dfor                    2003 GH                            1
-
-                 unk                     2003 GH                            1
+USA       WI     producer                2003 GH                            2
 
@@
 
-                 flds                    2012 unk                           4
+                 producer                2012 unk                           4
 
@@
 
-                 vic                     2004 Beryl                         2
+                 producer                2004 Beryl                         2
 
@@
 N = Number of Isolates. 
-Key abbreviations: wmn = white mold screening nursery, unk = unknown.
+Key abbreviations: wmn = white mold screening nursery, producer = producer field, unk = unknown cultivar.
 

### 2017-08-01 doc/RMD/wmn-differentiation.Rmd
Add test for NE WA, and MI

@@
 ssc_amova_region <- poppr.amova(dat11cc, ~SourceType/Region, dist = bd, quiet = TRUE)
-ssc_amova_region_test <- randtest(ssc_amova_region, nrepet = 999)
+ssc_amova_region_test <- randtest(ssc_amova_region, nrepet = 9999)
 plot(ssc_amova_region_test)
@@
 
+When we remove the Mexican isolates (which only contained white mold nurseries
+and shared no genotypes), we see that indeed, the degree of differentiation
+went down.  Of course, if we look at the distribution of isolates between white
+mold nurseries and regions, we can see things are a bit lopsided:
+
+```{r wmntable}
+table(strata(dat11cc, ~SourceType/Region, combine = FALSE))
+```
+
+The only regions that have at least ten samples in both white mold nurseries and
+production regions are NE, WA, and MI. We can see what happens when we subsample
+to these
+
+```{r AMOVA-newami, cache = TRUE}
+datnewami <- setPop(dat11cc, ~Region) %>% popsub(sublist = c("NE", "WA", "MI"))
+bdnewami     <- bruvo.dist(datnewami, replen = other(datnewami)$REPLEN)
+ssc_amova_newami <- poppr.amova(datnewami, ~SourceType/Region, dist = bdnewami, quiet = TRUE)
+ssc_amova_newami_test <- randtest(ssc_amova_newami, nrepet = 9999)
+plot(ssc_amova_newami_test)
+(ssc_amova_newami_table <- make_amova_table(ssc_amova_newami, ssc_amova_newami_test))
+```
+
 ```{r AMOVA_table, comment = NA}
-make_amova_printable(ssc_amova_table, ssc_amova_nm_table) %>%  
+make_amova_printable(ssc_amova_table, ssc_amova_newami_table) %>%  
   as_tibble() %>%
@@
 ```
-
-When we remove the Mexican isolates (which only contained white mold nurseries
-and shared no genotypes), we see that indeed, the degree of differentiation
-went down.
 

### 2017-08-01 doc/manuscript/manuscript.Rmd
update manuscript with new info re NEWAMI

@@
 The analysis of molecular variance (AMOVA) over the hierarchy of Source and Region showed no significant variation between white mold nurseries and non-white mold nurseries after accounting for variation between regions (Table \@ref(tab:amova)).
-Because Mexico shared no genotypes with any of the other populations and we suspected that it was contributing to some of the observed variance, we re-ran AMOVA without the Mexican isolates and found similar decreased, but still significant levels of differentiation between regions.
+Because we observed that the sample sizes of wmn isolates and producer isolates were lopsided in several populations (Table \@ref(tab:table-1)), we ran the analysis again with regions that had at least ten isolates from both wmn and producer fields, showing similar results (Table \@ref(tab:amova)).
 In contrast, when we compared the three cultivars, Beryl, Bunsi, and G122, we found no significant differentiation (Supplementary Data).
@@
 Bold $\Phi$ values indicate significant values (P < 1.00e^-4^).
-Numbers in parentheses represent AMOVA run without Mexican isolates included.
+Numbers in parentheses represent AMOVA run with the regions NE, WA, and MI.
 S.S.
@@
 The primary interest of this study was to assess if isolates sampled from white mold screening nurseries represent isolates from producer fields within the region [@steadman2003identification; @otto-hanson2011variation].
-According to our AMOVA results, we have evidence for differentiation at the region level, but little to no differentiation within region (Table \@ref(tab:amova)).
+According to our AMOVA results, we have evidence for differentiation at the region level, but little to no differentiation between white mold nursery isolates and production field isolates within region (Table \@ref(tab:amova)).
+This
 

### 2017-08-02 doc/RMD/RDA-analysis.Rmd
update RDA to include MCG

@@
   META   <- select(other(DAT)$meta, -Isolate)
-  STRATA <- select(strata(DAT), -MCG)
+  STRATA <- strata(DAT)
   STRATA <- if (wmn) mutate(STRATA, Source = ifelse(Source == "wmn", "wmn", "other")) else STRATA
@@
   META   <- select(other(CC)$meta, -Isolate)
-  STRATA <- select(strata(CC), -MCG)
+  STRATA <- strata(CC)
   MLL    <- data.frame(MLG = mll(CC))
@@
 # @return a ggplot2 object from the scores 
-plot_dbrda <- function(db, arrows = 10, seed = 2017-06-28){
+plot_dbrda <- function(db, arrows = 10, seed = 2017-06-28, lab = TRUE){
   set.seed(seed)
@@
   SampleCentroids <- rownames_to_column(data.frame(dbsum$sites), var = "isolate_names")
-  labs    <- summary(db, axes = 0)[["cont"]][["importance"]]["Proportion Explained", 1:2]
+  if (lab){
+    labs    <- vegan:::summary.cca(db, axes = 0)[["cont"]][["importance"]]["Proportion Explained", 1:2]
+    xl      <- paste0("Eig 1 (", round(labs[[1]]*100, 2), "% variance explained)")
+    yl      <- paste0("Eig 2 (", round(labs[[2]]*100, 2), "% variance explained)")
+  } else {
+    xl      <- "Eig 1"
+    yl      <- "Eig 2"
+  }
   terms   <- paste0("(", paste(attr(db$terms, "term.labels"), collapse = "|"), ")")
@@
                      data = Arrows) +
-    xlab(paste0("Eig 1 (", round(labs[[1]]*100, 2), "% variance explained)")) +
-    ylab(paste0("Eig 2 (", round(labs[[2]]*100, 2), "% variance explained)"))
+    xlab(xl) +
+    ylab(yl)
 }
@@
   theme(aspect.ratio = 1)
+# if (!interactive()){
   FILE <- file.path(PROJHOME, "results", "figures", "publication", "Figure7Z.pdf")
   ggsave(filename = FILE, width = 88, height = 88, units = "mm", scale = 1.5)
+# }
 plot_dbrda(cap16cc, arrows = 10) +
@@
 ```{r, echo = FALSE, results = "asis"}
-cap11sum <- summary(cap11cc, axes = 0, scaling = "sites")
-cap16sum <- summary(cap16cc, axes = 0, scaling = "sites")
+cap11sum <- vegan:::summary.cca(cap11cc, axes = 0, scaling = "sites")
+cap16sum <- vegan:::summary.cca(cap16cc, axes = 0, scaling = "sites")
 vartab <- function(x){
@@
 dat16raw <- genind2df(dat16cc, usepop = FALSE) %>% mutate_all(funs(as.integer(.)))
-vp11     <- varpart(dat11cc.bruvo, ~Severity + Year + Region + Host, data = ENV11, comm = dat11raw, add = TRUE)
+vp11     <- varpart(dat11cc.bruvo, ~Year + Region + Host + MCG, data = ENV11, comm = dat11raw, add = TRUE)
 plot(vp11, Xnames = c("Full Model", "No Model"))
@@
 ```{r}
-vp11 <- varpart(dat11cc.bruvo, ~Severity, ~Year, ~Region, ~Host, data = ENV11, add = TRUE)
-vp16 <- varpart(dat16cc.bruvo, ~Year, ~Region, ~Host, data = ENV16, add = TRUE)
+vp11 <- varpart(dat11cc.bruvo, ~Year, ~Region, ~Host, ~MCG, data = ENV11, add = TRUE)
+vp16 <- varpart(dat16cc.bruvo, ~Year, ~Region, ~Host, ~MCG, data = ENV16, add = TRUE)
 
-plot(vp11, digits = 2, Xnames = c("Severity", "Year", "Region", "Host"))
+plot(vp11, digits = 2, Xnames = c("Year", "Region", "Host", "MCG"))
 
-plot(vp16, digits = 2, Xnames = c("Year", "Region", "Host"))
+plot(vp16, digits = 2, Xnames = c("Year", "Region", "Host", "MCG"))
 ```

### 2017-08-02 doc/RMD/mlg-mcg.Rmd
update layout for igraph 1.1.2

@@
 set.seed(2017-05-03)
-lay2 <- create_layout(g, layout = "igraph", algorithm = "nicely")
+# set.seed(2017-08-02)
+
+lay2 <- create_layout(g, layout = "igraph", 
+                      algorithm = "fr", 
+                      weights = rep(1, length(E(g)$weight))
+                      # weights = ifelse(E(g)$weight == 1, 1, 1 + E(g)$weight/100)
+                      )
 
@@
 set.seed(2017-05-08)
-top5lay <- create_layout(top5g, layout = "igraph", algorithm = "nicely")
+top5lay <- create_layout(top5g, layout = "igraph", algorithm = "fr", weights = rep(1, length(E(top5g))))
 top5lay$community <- inner_join(data_frame(vertex = top5lay$name), the_communities)$comm

### 2017-08-02 doc/RMD/pop-diff.Rmd
add seed to poppr run

@@
 ```{r geno-div, cache = TRUE, fig.width = 10, fig.height = 10, comment = NA}
+set.seed(2017-08-02)
 ptab <- dat11cc %>% setPop(~Region) %>% poppr(sample = 999, total = FALSE)
@@
   arrange(desc(N), desc(MLG)) %>%
+  write_csv(file.path(PROJHOME, "results/tables/genotype_diversity_table_region.csv"), col_names = TRUE) %>%
   mutate(Ia = case_when(
@@
   huxtable::print_md(max_width = 80)
-
-ptab %>% 
-  arrange(desc(N), desc(MLG)) %>% 
-  write_csv(file.path(PROJHOME, "results/tables/genotype_diversity_table_region.csv"), col_names = TRUE)
 

### 2017-08-02 doc/manuscript/manuscript.Rmd
update versions, auto generate poppr table

@@
 All analyses were performed on this data set unless otherwise mentioned.
-All data processing and analyses were performed in R version 3.4.0 [@R] and are openly available and reproducible at https://github.com/everhartlab/sclerotinia-366/.
+All data processing and analyses were performed in R version 3.4.1 [@R] and are openly available and reproducible at https://github.com/everhartlab/sclerotinia-366/.
 Of the 16 microsatellite loci genotyped, five included compound repeats, which made it challenging to accurately/confidently bin alleles into fragment sizes expected for each locus based on the described repeat motif.
@@
 This was then converted to a bipartite graph where top nodes represented MCGs and bottom nodes represented MLGs.
-The cluster walktrap community detection algorithm as implemented in the `cluster_walktrap()` function in *igraph* version 1.0.1 was used to define clusters of MLGs and MCGs [@csardi2006igraph; @pons2006computing].
+The cluster walktrap community detection algorithm as implemented in the `cluster_walktrap()` function in *igraph* version 1.1.2 was used to define clusters of MLGs and MCGs [@csardi2006igraph; @pons2006computing].
 All communities that had fewer than 10 members were then consolidated into one.
@@
 
-----------------------------------------------------------------------------
-Pop             N        eMLG    H    G  $\lambda$  $E_5$   *h* $I_A$ $\bar{ 
-                                                                       r}_d$ 
---------- ------- ----------- ---- ---- ---------- ------ ----- ----- ------
-WA        58 (56) 9.95 (0.23)  4.0 54.3       0.98   0.98  0.60 0.65*  0.07*
-
-MI        58 (43)  9.3 (0.79)  3.6 29.0       0.97   0.78  0.54 1.25*  0.14*
-
-ND        41 (35) 9.44 (0.73)  3.5 25.9       0.96   0.82  0.54    1*   0.1*
-
-NE        37 (28) 8.93 (0.94)  3.2 17.8       0.94   0.75  0.55 2.22*  0.25*
-
-CO        34 (28) 9.46 (0.67)  3.3 24.1       0.96   0.92  0.56 2.33*  0.27*
-
-France    21 (14)  8.5 (0.85)  2.6 12.6       0.92   0.95  0.48 0.94*  0.11*
-
-CA        18 (15) 9.12 (0.72)  2.7 13.5       0.93   0.94  0.51 0.28~  0.03~
-
-OR        17 (13) 8.52 (0.85)  2.5 10.7       0.91   0.89  0.47 0.92*   0.1*
-
-Mexico     15 (9)  7.1 (0.85)  2.1  7.3       0.86   0.89  0.28 2.57*  0.37*
-
-MN          9 (7)       7 (0)  1.9  6.2       0.84   0.93  0.47 1.34*  0.19*
-
-Australia   6 (6)       6 (0)  1.8  6.0       0.83   1.00  0.48  0.7~  0.12~
-
-WI          2 (2)       2 (0)  0.7  2.0       0.50   1.00  0.27     -      -
-
-NY          1 (1)       1 (0)  0.0  1.0       0.00    NaN   NaN     -      -
-
-ID          1 (1)       1 (0)  0.0  1.0       0.00    NaN   NaN     -      -
-
-----------------------------------------------------------------------------
+```{r table_ia, echo = FALSE, warning = FALSE, message = FALSE, results = "asis"}
+suppressPackageStartupMessages(library("tidyverse"))
+read_csv("../../results/tables/genotype_diversity_table_region.csv", 
+         na = "NA", # spaces read in as spaces
+         col_types = cols(
+         Pop = col_character(),
+         N = col_integer(),
+         MLG = col_integer(),
+         eMLG = col_double(),
+         SE = col_double(),
+         H = col_double(),
+         G = col_double(),
+         lambda = col_double(),
+         E.5 = col_double(),
+         Hexp = col_double(),
+         Ia = col_double(),
+         p.Ia = col_double(),
+         rbarD = col_double(),
+         p.rD = col_double(),
+         File = col_character()
+      )) %>%
+  mutate(Ia = case_when(
+    !is.finite(p.Ia) ~ "-",
+    p.Ia == 0.001 ~ paste0(round(Ia, 2), "*"),
+    p.Ia <= 0.01  ~ paste0(round(Ia, 2), "-"),
+    p.Ia <= 0.05  ~ paste0(round(Ia, 2), "~"),
+    TRUE          ~ as.character(round(Ia, 2))
+  )) %>%
+  mutate(rbarD = case_when(
+    !is.finite(p.rD) ~ "-",
+    p.rD == 0.001 ~ paste0(round(rbarD, 2), "*"),
+    p.rD <= 0.01  ~ paste0(round(rbarD, 2), "-"),
+    p.rD <= 0.05  ~ paste0(round(rbarD, 2), "~"),
+    TRUE          ~ as.character(round(rbarD, 2))
+  )) %>%
+  mutate_if(is.numeric, round, digits = 2) %>%
+  mutate(N = paste(N, MLG, sep = " (") %>% paste0(")") %>% formatC()) %>%
+  mutate(eMLG = paste(eMLG, SE, sep = " (") %>%  paste0(")") %>% formatC()) %>%
+  rename(`*h*` = Hexp) %>%
+  rename(`$E_5$` = E.5) %>%
+  rename(`$I_A$` = Ia) %>%
+  rename(`$\\lambda$` = lambda) %>%
+  rename(`$\\bar{r}_d$` = rbarD) %>%
+  select(-File, -MLG, -SE, -p.rD, -p.Ia) %>% 
+  huxtable::as_huxtable(add_colnames = TRUE) %>%
+  huxtable::set_number_format(huxtable::everywhere, 4:5, 1) %>%
+  huxtable::set_align(huxtable::everywhere, -1, "right") %>%
+  huxtable::set_col_width(c(1.1, 0.9, 1.3, 0.5, 0.5, 1.2, 0.7, 0.6, 0.6, 0.8)) %>%
+  huxtable::print_md(max_width = 80)
+```
 
@@
 $\bar{r}_d$ = standardized index of association. 
-P-value symbols: \* = 0.001, ~ = $\leq 0.05$.
+P-value symbols: \* = 0.001, - = $\leq 0.01$, ~ = $\leq 0.05$.
 

### 2017-08-02 doc/RMD/RDA-analysis.Rmd
add original analysis to RDA to remove collinearity

@@
   theme(aspect.ratio = 1)
-# if (!interactive()){
-  FILE <- file.path(PROJHOME, "results", "figures", "publication", "Figure7Z.pdf")
-  ggsave(filename = FILE, width = 88, height = 88, units = "mm", scale = 1.5)
-# }
 plot_dbrda(cap16cc, arrows = 10) +
@@
 This is all well and good, but what exactly does it actually mean to have only 
-1.93 variance explained (for 11 loci, that is)? I was a bit stumped as well. 
+3.33% variance explained (for 11 loci, that is) on the first axis? I was a bit stumped as well. 
 Luckily, Gavin Simpson has some answers: http://stackoverflow.com/a/22537820/2752888
 
-Basically, we have a situation where we have all of the variables only
-explaining a whopping 16 - 17% of the variance and the rest goes unexplained.
+Basically, we have a situation where we have all of the variables
+explaining 45% of the variance and the rest goes unexplained.
 
@@
 
-Out of both of the models, only Region explains the most variance, almost double
-that of Host or Year and Severity contributes almost nothing.
+Out of both of the models, MCG explains the most variance, but we have an 
+interesting warning pop up indicating that there is multicollinearity within our
+variables. It appears that all of these are multicollinear with MCG. Since dbRDA
+assumes that all the variables are independent, it would be prudent to re-run 
+the analysis without the MCG variables as well
+
+## Removing Source of Multicollinearity
+
+```{r rda_calculation2, cache = TRUE}
+# 11 loci
+ENV112   <- select(ENV11, -MCG)
+cap11cc2 <- choose_dbrda(dat11cc.bruvo, ENV = ENV112, CHOOSER = "ordistep")
+```
 
 ```{r}
-wmn11 <- seppop(dat11, ~Source)$wmn %>% setPop(~Region)
-wmn11_sub <- popsub(wmn11, blacklist = character(0))
-min_span_net <- bruvo.msn(wmn11_sub, replen = other(wmn11)$REPLEN, add = TRUE, loss = TRUE, showplot = FALSE, include.ties = TRUE)
-set.seed(67)
-plot_poppr_msn(wmn11,
-               min_span_net,
-               inds = "none",
-               mlg = TRUE,
-               gadj = 6,
-               nodelab = 100,
-               nodebase = 1.15,
-               palette = funky,
-               cutoff = (0.5/11)*3,
-               quantiles = FALSE,
-               beforecut = TRUE,
-               layfun = igraph::layout_with_lgl)
+cap11cc2
+cap11cc2$anova
+vegan::RsquareAdj(cap11cc2)
 ```
+
+```{r total_anova2, cache = TRUE}
+anova(cap11cc2)
+```
+
+We can also see the marginal effects with `by = "margin"`
+
+```{r marginal_anova2, cache = TRUE}
+anova(cap11cc2, by = "margin")
+```
+
+```{r resultplot2, fig.width = 5, fig.height = 5}
+set.seed(999)
+plot_dbrda(cap11cc2, arrows = 5) + 
+  theme_classic(base_size = 16, base_family = "Helvetica") + 
+  theme(axis.text = element_text(color = "black")) +
+  theme(aspect.ratio = 1)
+if (!interactive()){
+  FILE <- file.path(PROJHOME, "results", "figures", "publication", "Figure7Z.pdf")
+  ggsave(filename = FILE, width = 88, height = 88, units = "mm", scale = 1.5)
+}
+```
+
+What we can see here is that the amount of explained variance went down a bit to
+1.92, and our model now only explains 16% of the variance.
+
+```{r, echo = FALSE, results = "asis"}
+cap11sum2 <- vegan:::summary.cca(cap11cc2, axes = 0, scaling = "sites")
+knitr::kable(vartab(cap11sum2), digits = 3)
+```
+
+```{r}
+vp11     <- varpart(dat11cc.bruvo, ~Severity + Year + Region + Host, data = ENV11, comm = dat11raw, add = TRUE)
+plot(vp11, Xnames = c("Full Model", "No Model"))
+vp11 <- varpart(dat11cc.bruvo, ~Severity, ~Year, ~Region, ~Host, data = ENV11, add = TRUE)
+
+plot(vp11, digits = 2, Xnames = c("Year", "Region", "Host", "MCG"))
+```
+
+# Conclusions
+
+This analysis showed us two things: Severity is not an important predictor for 
+genetic diversity, but Region, Host, Year, and MCG are all predictors of genetic
+diversit. Explanations for this result could be due to a few reasons: MCG, a
+phenotypic variable associated with pre-zygotic barriers has SOME genetic 
+component to it; the MCGs that explained the most variation were also the MCGs
+that were strongly associated with one genotype (i.e. MCG 5 has 19 isolates of
+MLG 25).
 

### 2017-08-02 doc/manuscript/manuscript.Rmd
add original analysis to RDA to remove collinearity

@@
 This method uses constrained ordinations on a distance matrix representing the response variable to delineate relative contribution of any number of independent explanatory variables. 
-We used this method to delineate the phenotypic, geographic, and temporal components in influencing genetic composition of the populations.
+We used this method to delineate the phenotypic (Aggressiveness, MCG), geographic (Region, Host, Location), and temporal (Year) components in influencing genetic composition of the populations.
 <!-- RDA is a multivariate linear regression analysis of independent (explanatory: eg. environmental and phenotypic) variables on dependent (non-explanatory or response) variables. -->
-The distance matrix we used as the response variable was generated using Bruvo's genetic distance as implemented in *poppr* version 2.4.1, which employed a stepwise mutation model for microsatellite data [@bruvo2004simple; @kamvar2014poppr].
+The distance matrix we used as the response variable was generated using Bruvo's genetic distance from clone-corrected data on the hierarchy of Region, Location, Host, Year as implemented in *poppr* version 2.4.1, which employed a stepwise mutation model for microsatellite data [@bruvo2004simple; @kamvar2014poppr].
+Because aggressiveness measures differed with each isolate, aggressiveness first averaged across clone-corrected isolates.
 <!-- In our case, a pairwise Bruvo's [@bruvo2004simple] distance matrix was generated from microsatellite data using *poppr* 2.4.1 [@kamvar2014poppr; @kamvar2015novel] and employed it as the dependent matrix for RDA analysis.
@@
 The dbRDA was performed on clone-corrected isolates (N = 318) after grouping them by location, host, region, and year.
-The forward-backward selection process revealed severity, year, region, and host to be the optimal variables for the reduced model, accounting for 17.6% of the total variation.
+The forward-backward selection process revealed aggressiveness, year, region, and host to be the optimal variables for the reduced model, accounting for 17.6% of the total variation.
 ANOVA revealed that the reduced model was significant with an adjusted R^2^ of 0.033 (P = 0.001).
@@
      Region has a strong influence on both population structure and
-     severity.
+     aggressiveness.
    + North Dakota appeared
@@
 Aggressiveness and specific source field had little to no effect on genetic diversity. 
-These results are in agreement with studies that examined differentiation based on host [@aldrich-wolfe2015genetic] and severity [@atallah2004high; @attanayake2012genetic; @attanayake2013sclerotinia] reporting little or no correlation of genetic diversity to these variables.
+These results are in agreement with studies that examined differentiation based on host [@aldrich-wolfe2015genetic] and aggressiveness [@atallah2004high; @attanayake2012genetic; @attanayake2013sclerotinia] reporting little or no correlation of genetic diversity to these variables.
 This indicates that a) breeders should keep in mind regional differences when assessing resistance and b) it is possible that we have not yet measured biologically relevant variables that can predict genetic differentiation, which could include variables such as soil community composition.

### 2017-08-02 doc/manuscript/manuscript.Rmd
add interpretation of RDA

@@
 The dbRDA was performed on clone-corrected isolates (N = 318) after grouping them by location, host, region, and year.
-The forward-backward selection process revealed aggressiveness, year, region, and host to be the optimal variables for the reduced model, accounting for 17.6% of the total variation.
-ANOVA revealed that the reduced model was significant with an adjusted R^2^ of 0.033 (P = 0.001).
-Assessment of the marginal effects showed that state, host, and year significantly explained genetic variation (P = 0.001), but aggressiveness was only mildly significant (P = 0.045). 
+The forward-backward selection process revealed year, region, host, and aggressiveness to be the optimal variables for the reduced model, accounting for 45% of the total variation.
+ANOVA revealed that the reduced model was significant with an adjusted R^2^ of 0.0674 (P = 0.001).
+Assessment of the marginal effects showed that region, host, year, and aggressiveness significantly explained genetic variation (P = 0.001, P = 0.009 for region).
+It was revealed that there was multicollinearity combining all variables with MCG, so we followed the same procedure, dropping MCG from the list of potential predictors. 
+From these results, we found that aggressiveness, year, region, and host were optimal, accounting for 17.6% of the total variation.
+ANOVA revealed significant effects with an adjusted R^2^ of 0.0325 (P = 0.001).
+While the marginal effect assessment showed that region, host, and year all significantly explained the variation at P = 0.001, and aggressiveness significantly explained the variation at P = 0.045.
 Much of the variation appears to be driven by isolates from Mexico and 2005 (Fig. \@ref(rda-biplot)).
-Variance partitioning of the independent variables indicated aggressiveness to be the least influential factor with 0.1% contributing to explaining the variation of molecular data whereas the combination of variables accounted for 3.3%.
+Variance partitioning of the independent variables without MCG indicated aggressiveness to be the least influential factor with 0.1% contributing to explaining the variation of molecular data whereas the combination of variables accounted for 3.3%.
 
@@
 Borrowing a technique often used in the ecological literature, we used dbRDA to elucidate the effect of all variables (Aggressiveness, Region, Source,   Year, and Host) [@legendre1999distance].
-From the initial results, it appears that the most important factors for predicting genetic structure are region and year.
+From the initial results, it appears that the most important factors for predicting genetic structure are MCG, region, and year.
+When we inspected the biplot of the initial results, we saw that the most important predictors were MCGs 44, 5, and 9. 
+We believe that this is driven by the fact that these particular MCGs have uneven MLG distributions, meaning that they are heavily associated with one particular MLG (Fig.
+\@ref(biggraph), \@ref(barplot)).
 While the results indicated that Mexico and the year 2005 were the two most important variables, it's worth noting that all Mexican isolates were collected in 2005 (Fig. \@ref(rda-biplot)).
@@
 
-## Mycelial Compatibility Groups and Multillocus Genotypes {-}
+## Mycelial Compatibility Groups and Multilocus Genotypes {-}
 

### 2017-08-02 doc/manuscript/manuscript.Rmd
add more things to discussion

@@
 The analysis of molecular variance (AMOVA) over the hierarchy of Source and Region showed no significant variation between white mold nurseries and non-white mold nurseries after accounting for variation between regions (Table \@ref(tab:amova)).
-Because we observed that the sample sizes of wmn isolates and producer isolates were lopsided in several populations (Table \@ref(tab:table-1)), we ran the analysis again with regions that had at least ten isolates from both wmn and producer fields, showing similar results (Table \@ref(tab:amova)).
+Because we observed that the sample sizes of wmn isolates and producer isolates were lopsided in several populations (Table \@ref(tab:isolate-table)), we ran the analysis again with regions that had at least ten isolates from both wmn and producer fields, showing similar results (Table \@ref(tab:amova)).
 In contrast, when we compared the three cultivars, Beryl, Bunsi, and G122, we found no significant differentiation (Supplementary Data).
@@
 
-## Regional Differentiation {-}
+## Population Differentiation {-}
 
@@
 According to our AMOVA results, we have evidence for differentiation at the region level, but little to no differentiation between white mold nursery isolates and production field isolates within region (Table \@ref(tab:amova)).
-This
+This result indicates that breeders may consider regional differences when assessing resistance of new bean lines. 
+The fact that much variation was reduced when assessing populations with equally representative wmn and producer fields indirectly supports this claim.
+
+## Limitations {-}
+
+One of the main limitations of this study is the focus on *P. vulgaris* as a host. 
+It has been shown that *S. sclerotiorum* in the midwestern United States does not have a particular preference for host [@aldrich-wolfe2015genetic]. 
+If the distribution of *S. sclerotiorum* is even across agricultural hosts in the US, then our sample may yet be representative of the genetic pool present in other crops and weedy species.
+
+Another limitation was the microsatellite markers used for this particular study [@sirjusingh2001characterisation].
+The genotype accumulation curve showed no indication of a plateau, indicating that if we had sampled more loci, we would have resolved more multilocus genotypes. 
+While 16 loci showed us similar results and began to show a plateau for the genotype accumulation curve, we were unable to use these results due to our uncertainty in the allele calls for these five extra loci.
+With the availability of an optically-mapped genome [citation-needed], future studies describing the genetic diversity of *S.
+sclerotiorum* should employ techniques such as Genotyping-By-Sequencing [citation-needed], Sequence Capture [citation-needed], or Whole Genome Sequencing [citation-needed].
 

### 2017-08-03 doc/RMD/mlg-mcg.Rmd
update visNetwork stuff

@@
        width = 88*3, height = 88*3.5, units = "mm")
+  tmp <- new.env()
+  assign("mcg_mlg_igraph", g, envir = tmp)
+  assign("mcg_mlg_gg", mcg_mlg_graph, envir = tmp)
+  V(tmp$mcg_mlg_igraph)$size <- osize
+  save(list = ls(envir = tmp), 
+       envir = tmp, 
+       file = file.path(PROJHOME, "data", "mcg-mlg-graph.rda"))
 }
@@
 }
+community_colors <- viridis::viridis(4, direction = -1)[as.integer(factor(the_communities$comm))]
 vg <- g %>%
-  set_vertex_attr("size", value = osize) %>%
+  set_vertex_attr("size", value = osize * 10) %>%
   set_vertex_attr("value", value = osize) %>%
-  set_vertex_attr("color", value = viridis(4)[as.integer(factor(the_communities$comm))]) %>%
-  set_vertex_attr("shape", value = ifelse(V(.)$type == "Multilocus Genotype", "triangle", "circle")) %>%
+  set_vertex_attr("label", value = NULL) %>%
+  set_vertex_attr("color", value = community_colors) %>%
+  set_vertex_attr("shape", value = ifelse(V(.)$type == "Multilocus Genotype", "triangle", "dot")) %>%
   set_edge_attr("width", value = E(.)$weight) %>%
@@
 vg$nodes <- vg$nodes %>% 
-  group_by(id) %>%
-  mutate(color = list(list(background = make_rgb(color, 0.8), 
+  dplyr::group_by(id) %>%
+  dplyr::mutate(color = list(list(background = make_rgb(color, 0.8), 
                       border = make_rgb(rep("black", n()), 0.8),
@@
                                        )
-                      )))
+                      ))) %>%
+  dplyr::mutate(title = paste0("<p>",
+                               "<i>", type, " ", gsub("MLG.", "", id), "</i><br>",
+                               "<b>Community: ", community, "</b><br>",
+                               "<b>N: ", size/10, "</b><br>",
+                               "</p>")
+                )
+vg$edges <- vg$edges %>%
+  dplyr::mutate(title = paste0("<p>",
+                        "<b>Mycelial Compatibility Group: ", from, "</b><br>",
+                        "<b>Multilocus Genotype: ", gsub("MLG.", "", to), "</b><br>",
+                        "<b>N: ", weight, "</b>",
+                        "</p>")
+         )
 vgn <- visNetwork(nodes = vg$nodes, edges = vg$edges,# height = "500px", 
@@
 vgn %>%
-  visIgraphLayout("layout_nicely") %>%
+  # visIgraphLayout("layout_nicely") %>% # activate this for a poseable network
+  visNodes(label = NULL, shadow = TRUE) %>%
   visOptions(highlightNearest = list(enabled = TRUE, hover = TRUE), 

### 2017-08-03 doc/manuscript/manuscript.Rmd
add more discussion on aggressiveness

@@
 Points represent the 318 clone-corrected samples in ordination space. 
-](../../results/figures/publication/Figure7Z.pdf){#rda-biplot width=45%}
+](../../results/figures/publication/Figure7Z.pdf){#rda-biplot width=61%}
 
@@
 Points represent individual observations whose colors indicate if they came from white mold screening nurseries (wmn) or not (other).
-](../../results/figures/publication/aggressiveness.pdf){#aggressiveness-boxplot width=45%}
+](../../results/figures/publication/aggressiveness.pdf){#aggressiveness-boxplot}
 
@@
 According to our AMOVA results, we have evidence for differentiation at the region level, but little to no differentiation between white mold nursery isolates and production field isolates within region (Table \@ref(tab:amova)).
-This result indicates that breeders may consider regional differences when assessing resistance of new bean lines. 
-The fact that much variation was reduced when assessing populations with equally representative wmn and producer fields indirectly supports this claim.
+Additionally, we observed significant differences in aggressiveness based on region (Fig. \@ref(aggressiveness-boxplot), Table \@ref(tab:region-aggressiveness)). 
+These results show a similar pattern to what was found previously with the exception of North Dakota, which increased in mean aggressiveness from 5 to 5.77 [@otto-hanson2011variation].
+This increase was due in part to new data from producer field isolates collected after the previous study.
+Two caveats should come with interpreting these data, however.
+First, these straw tests were performed by a different person for these later isolates, which could suggest a more lenient or strict scoring system. 
+Moreover to this point, many of the ND isolates fell within the 6--7 range, which denotes a physical boundary (disease symptoms around the second node) between intermediate and susceptible [@otto-hanson2011variation].
+Second, these results are all within the intermediate range (4--6) of aggressiveness, so the differences between states may not be significant for a producer.
 
@@
 While 16 loci showed us similar results and began to show a plateau for the genotype accumulation curve, we were unable to use these results due to our uncertainty in the allele calls for these five extra loci.
-With the availability of an optically-mapped genome [citation-needed], future studies describing the genetic diversity of *S.
-sclerotiorum* should employ techniques such as Genotyping-By-Sequencing [citation-needed], Sequence Capture [citation-needed], or Whole Genome Sequencing [citation-needed].
+With the availability of an optically-mapped genome [@derbyshire2017complete], future studies describing the genetic diversity of *S.
+sclerotiorum* should employ techniques such as Genotyping-By-Sequencing [@davey2011genome], Sequence Capture [@grover2012targeted], or Whole Genome Sequencing.
 

### 2017-08-03 doc/RMD/tree.Rmd
add tree

---
title: "tree"
---

```{r setup, include = FALSE}
knitr::opts_knit$set(root.dir = PROJHOME)
knitr::opts_chunk$set(fig.align = "center")
```

## Packages and Data

```{r load_packages, warning = FALSE, message = FALSE}
library('tidyverse')
library('poppr')
library('ggtree')
```

```{r load_data}
load(file.path(PROJHOME, "data", "sclerotinia_16_loci.rda"))
setPop(dat11) <- ~Host/Source/Region/Year
dat11cc <- clonecorrect(dat11, ~Host/Source/Region/Year, keep = 1:4)
dat11cc

# Asserting that nothing messed up with the metadata.
stopifnot(identical(indNames(dat11cc), other(dat11cc)$meta$Isolate))
```

The purpose of this document is simply to calculate a bootstrapped tree for 
Bruvo's distance.

```{r bdtree, cache = TRUE}
set.seed(2017-08-03)
bt <- bruvo.boot(clonecorrect(dat11, strata = NA), 
                 replen = other(dat11)$REPLEN, 
                 sample = 1000, 
                 tree = "nj", 
                 showtree = FALSE)
```

Let's take a look at the results of the bootstrap analysis. Note, that we would
traditionally ignore results < 75.

```{r bsresults}
bt$node.labels
summary(bt$node.labels)
hist(bt$node.labels)
```

It's clear that our results show that NONE of the clades are well supported
(except for the whole tree, which is default), so let's take a look at how the
tree looks. One of the things we want to do is see where the popualtions fit.
Since we created a clone-corrected tree here (for speed), we are going to create
a matrix to tally up the samples from different populations per MLG.

An important thing to keep track of is the fact that ggtree works from rownames
for tip labels, so we have to add them in.

```{r}
data <- bind_cols(strata(dat11), other(dat11)$meta) %>% 
  add_column(MLG = mll(dat11))

otherdf <- data %>% 
  group_by(MLG, Region) %>%
  summarize(N = n()) %>%
  spread(Region, N)
otherdf <- inner_join(data %>% 
                        group_by(MLG) %>% 
                        summarize(N = n(), id = Isolate[1]), 
                      otherdf, 
                      by = "MLG")
df <- otherdf %>% 
  select(-MLG, -N) %>% 
  as.data.frame() %>% 
  column_to_rownames("id")
otherdf
```

Now we can create the tree with the matrix, coloring by number of isolates.

```{r the_tree, fig.width = 7, fig.height = 12}
gbt <- ggtree(bt) + geom_tippoint() + theme_tree2() + xlab("Bruvo's Distance (11 loci)")
gheatmap(gbt, df, colnames = FALSE, width = 0.6) %>% 
  scale_x_ggtree() + 
  viridis::scale_fill_viridis(guide = "legend") +
  scale_y_continuous(expand = c(0, 0)) +
  theme(text = element_text(size = 16, family = "Helvetica")) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, color = "black")) +
  theme(panel.grid.major.x = element_line(colour = "grey50", linetype = 3)) +
  labs(list(fill = "N isolates"))
```

Of course, because of the low bootstrap support, we don't have very much faith
in this tree. One of the promising aspects, however is the fact that all the
isolates from Mexico group together, as was shown in the DAPC. Moreover, we can
see that this is not driven by private alleles:

```{r private allelels}
private_alleles(dat11, locus ~ Region, count.alleles = FALSE)
```

<details>
<summary>Session Information</summary>

```{r, echo = FALSE}
options(width = 100)
devtools::session_info()
```

</details>

### 2017-08-03 doc/manuscript/manuscript.Rmd
add discussion of population differentiation

@@
 According to our AMOVA results, we have evidence for differentiation at the region level, but little to no differentiation between white mold nursery isolates and production field isolates within region (Table \@ref(tab:amova)).
-Additionally, we observed significant differences in aggressiveness based on region (Fig. \@ref(aggressiveness-boxplot), Table \@ref(tab:region-aggressiveness)). 
+When we assessed AMOVA using only populations with equally representative white mold screening nursery isolates and producer field isolates, we see the same pattern, except the $\Phi$ values decrease by nearly half (Table \@ref(tab:amova)).
+When we analyze the AMOVA results in light of the DAPC results (Fig. \@ref(DAPC)), it becomes clear that the patterns of differentiation are largely driven by isolates from Mexico and California. 
+Isolates from these populations had a higher posterior probability (> 0.75) of being reassigned to their own populations than any other (Fig. \@ref(DAPC)a). 
+All other populations in comparison (except France) has reassignment probabilities of < 0.5, which is reflected in the failure of the first two discriminant functions to separate these populations (Fig.
+\@ref(DAPC)b).
+DAPC is based on the principal components of Euclidean distance between genotypes [@jombart2010discriminant]. 
+Unlike Bruvo's distance, Euclidean distance does not take into account the magnitude of the difference between alleles, which could inflate the distance measure in the presence of private alleles [@bruvo2004simple].
+We found no evidence of private alleles in the Mexican and Californian isolates, suggesting that this pattern separating these populations from the others was not an artifact.
+
+It has been proposed within the *S.
+sclerotiorum* literature that isolates from sub-tropical and tropical climates are differentiated or more variable than populations from temperate climates [@lehner2017sclerotinia;@carbone2001multilocus; @attanayake2013sclerotinia].
+This has been attributed to the notion that the fungus has the chance to undergo more reproductive cycles in the warmer climate [@carbone2001multilocus; @attanayake2013sclerotinia]. 
+Populations in sub-tropical regions of China have been found to be more variable, sexually reproducing, and unrelated to populations in temperate regions of the United States [@attanayake2013sclerotinia].
+The differentiation between the Mexican isolates and all other regions gives some evidence for this hypothesis as the only isolates in our study from sub-tropical regions are in the Mexican populations.
+
+## Aggressiveness {-}
+
+We observed significant differences in aggressiveness based on region (Fig. \@ref(aggressiveness-boxplot), Table \@ref(tab:region-aggressiveness)). 
 These results show a similar pattern to what was found previously with the exception of North Dakota, which increased in mean aggressiveness from 5 to 5.77 [@otto-hanson2011variation].
@@
 Moreover to this point, many of the ND isolates fell within the 6--7 range, which denotes a physical boundary (disease symptoms around the second node) between intermediate and susceptible [@otto-hanson2011variation].
-Second, these results are all within the intermediate range (4--6) of aggressiveness, so the differences between states may not be significant for a producer.
+Second, these results are all within the intermediate range (4--6) of aggressiveness, which may not represent a significant biological difference.
 

### 2017-08-03 doc/manuscript/manuscript.Rmd
begin to talk about sex

@@
 \@ref(biggraph), \@ref(barplot)).
+We note these results with caution because of the apparent multicolinearity between MCG and Region, which is a violation of the analysis [@legendre1999distance].
 While the results indicated that Mexico and the year 2005 were the two most important variables, it's worth noting that all Mexican isolates were collected in 2005 (Fig. \@ref(rda-biplot)).
@@
 The differentiation between the Mexican isolates and all other regions gives some evidence for this hypothesis as the only isolates in our study from sub-tropical regions are in the Mexican populations.
+
+We were unable to find evidence for sexual reproduction in any region. 
+While the Australian population had a non-significant value of $\bar{r}_d$---which would suggest that we cannot reject the null hypothesis of random mating---the sample size was insufficient from which to draw conclusions [@agapow2001indices]
 

### 2017-08-03 doc/manuscript/manuscript.Rmd
add milgroom

--- 2017-08-03 doc/manuscript/manuscript.Rmd
@@
 
+We were unable to find evidence for sexual reproduction in any region. 
+While the Australian population had a non-significant value of $\bar{r}_d$---which would suggest that we cannot reject the null hypothesis of random mating---the sample size was insufficient from which to draw conclusions [@agapow2001indices; @milgroom1996recombination].
+
 ## Aggressiveness {-}
@@
 ## Conclusions {-}
+
+eh.
 
@@
 # Supplementary Information {-}
-\newpage
 
@@
 Table: (\#tab:region-aggressiveness) Mean aggressiveness ratings for the Regions with greater than 5 samples; groupings according to 95% family-wise confidence interval.
+
+\newpage
 
@@
 
-\newpage
-
 <!-- community graph figure -->

### 2017-08-14 doc/manuscript/manuscript.Rmd
white mold nurs -> white mold screening nurs

@@
 sclerotiorum* isolates collected between 2003 and 2012 from dry bean cultivars among different geographic locations in the United States, France, Mexico, and Australia.
-Because the white mold nurseries serve as testing grounds for new varieties of dry bean, we wanted to know if the *S.
-sclerotiorum* populations from white mold nurseries were representative of the fields within the same region.
+Because the white mold screening nurseries serve as testing grounds for new varieties of dry bean, we wanted to know if the *S.
+sclerotiorum* populations from white mold screening nurseries were representative of the fields within the same region.
 As these nurseries are not treated with any chemical or cultural control of white mold, we hypothesized that these nurseries would represent the natural population of *S. sclerotiorum*.
-Furthermore, as three cultivars (Beryl, Bunsi, and G122) were grown and tested previously by @otto-hanson2011variation among all white mold nurseries, we wanted to investigate the potential effect of cultivar on genetic diversity.
+Furthermore, as three cultivars (Beryl, Bunsi, and G122) were grown and tested previously by @otto-hanson2011variation among all white mold screening nurseries, we wanted to investigate the potential effect of cultivar on genetic diversity.
 We additionally wanted to assess what categorical or phenotypic variables best predict genetic structure and if there was correlation between multilocus genotype and mycelial compatibility group.
@@
 Sclerotia of *S.
-sclerotiorum* were collected over several years from grower fields and/or white mold nurseries in 11 states of the United States of America, Mexico, France, and Australia (Table \@ref(tab:isolate-table)).
-The white mold nurseries in different regions are indicated by wmn (white mold nursery) and all other isolates came from commercial fields (Table \@ref(tab:isolate-table)).
+sclerotiorum* were collected over several years from grower fields and/or white mold screening nurseries in 11 states of the United States of America, Mexico, France, and Australia (Table \@ref(tab:isolate-table)).
+The white mold screening nurseries in different regions are indicated by wmn (white mold screening nursery) and all other isolates came from commercial fields (Table \@ref(tab:isolate-table)).
 
@@
 
-The analysis of molecular variance (AMOVA) over the hierarchy of Source and Region showed no significant variation between white mold nurseries and non-white mold nurseries after accounting for variation between regions (Table \@ref(tab:amova)).
+The analysis of molecular variance (AMOVA) over the hierarchy of Source and Region showed no significant variation between white mold screening nurseries and non-white mold screening nurseries after accounting for variation between regions (Table \@ref(tab:amova)).
 Because we observed that the sample sizes of wmn isolates and producer isolates were lopsided in several populations (Table \@ref(tab:isolate-table)), we ran the analysis again with regions that had at least ten isolates from both wmn and producer fields, showing similar results (Table \@ref(tab:amova)).
@@
 
-Table: (\#tab:amova) Analysis of Molecular Variance on Bruvo's genetic distance showing little differentiation between White Mold Nurseries and other sources.
-The hierarchy was constructed as Source/Region where source is defined as belonging to a white mold nursery or not.
+Table: (\#tab:amova) Analysis of Molecular Variance on Bruvo's genetic distance showing little differentiation between white mold screening nurseries and other sources.
+The hierarchy was constructed as Source/Region where source is defined as belonging to a white mold screening nursery or not.
 Bold $\Phi$ values indicate significant values (P < 1.00e^-4^).
@@
    + This is an extension of a previous work [@otto-hanson2011variation]
-   + White Mold Nurseries serve as good way to screen potential resistant lines, avoiding environmental effects
+   + white mold screening nurseries serve as good way to screen potential resistant lines, avoiding environmental effects
    + It's unclear how genetic structure can influence
@@
 In this study, we characterized the diversity of *Sclerotinia sclerotiorum* from dry bean fields across the United States.
-Our results suggest that, broadly, populations from white mold nurseries reflect the populations of the surrounding regions, indicating that resistance screening may be successful at a regional level.
+Our results suggest that, broadly, populations from white mold screening nurseries reflect the populations of the surrounding regions, indicating that resistance screening may be successful at a regional level.
 We found significant population differentiation by geographic region, mainly differentiated into three broad North American groups based on shared genotypes and posterior groupings, a Coastal Region, Midwestern Region, and Mexico.
 To date, with 366 isolates, this is the largest population genetic study of *S. sclerotiorum* assessing population structure within managed and unmanaged agricultural environments. 
-These findings indicate that the white mold nurseries can be effective at screening for potential resistant lines within growing regions.
+These findings indicate that the white mold screening nurseries can be effective at screening for potential resistant lines within growing regions.
 
@@
 The primary interest of this study was to assess if isolates sampled from white mold screening nurseries represent isolates from producer fields within the region [@steadman2003identification; @otto-hanson2011variation].
-According to our AMOVA results, we have evidence for differentiation at the region level, but little to no differentiation between white mold nursery isolates and production field isolates within region (Table \@ref(tab:amova)).
+According to our AMOVA results, we have evidence for differentiation at the region level, but little to no differentiation between white mold screening nursery isolates and production field isolates within region (Table \@ref(tab:amova)).
 When we assessed AMOVA using only populations with equally representative white mold screening nursery isolates and producer field isolates, we see the same pattern, except the $\Phi$ values decrease by nearly half (Table \@ref(tab:amova)).
@@
 The same study reported increased ascospores discharge at 20-25$^{\circ}$ C compared to 15$^{\circ}$ C and the survival of *S. sclerotiorum* ascospores was reduced at high temperature and RH.
-Our samples were collected from 2003 to 2012 and except for the white mold nurseries, came from different fields.
+Our samples were collected from 2003 to 2012 and except for the white mold screening nurseries, came from different fields.
 Different agro-ecological conditions prevailed during sampling seasons may have resulted in different rates of sexual outcrossing of the pathogen and hence reduced the chance of extracting a common genotype.
@@
 Our isolates had a total of 165 MLGs and they did not correspond well with 87 MCGs.
-Linkage disequilibrium analysis of populations in white mold nurseries indicated the possibility of sexual outcrossing in some regions (ex.
+Linkage disequilibrium analysis of populations in white mold screening nurseries indicated the possibility of sexual outcrossing in some regions (ex.
 CA, FR, OR and WA) during certain sampling years while most locations indicated clonal reproduction.

### 2017-08-14 doc/manuscript/manuscript.Rmd
update last paragraph of intro

@@
 sclerotiorum* isolates collected between 2003 and 2012 from dry bean cultivars among different geographic locations in the United States, France, Mexico, and Australia.
-Because the white mold screening nurseries serve as testing grounds for new varieties of dry bean, we wanted to know if the *S.
-sclerotiorum* populations from white mold screening nurseries were representative of the fields within the same region.
+We wanted to know if the *S. sclerotiorum* populations from white mold screening nurseries were representative of the fields within the same region. 
 As these nurseries are not treated with any chemical or cultural control of white mold, we hypothesized that these nurseries would represent the natural population of *S. sclerotiorum*.
-Furthermore, as three cultivars (Beryl, Bunsi, and G122) were grown and tested previously by @otto-hanson2011variation among all white mold screening nurseries, we wanted to investigate the potential effect of cultivar on genetic diversity.
-We additionally wanted to assess what categorical or phenotypic variables best predict genetic structure and if there was correlation between multilocus genotype and mycelial compatibility group.
-By investigating these aims, we hope to effectively describe the population dynamics of *S. sclerotiorum* and make available our database of isolates for use in future dry bean breeding efforts.
+Furthermore, we wanted to investigate the potential effect of cultivar on genetic diversity by assessing three dry bean cultivars with different levels of resistance, Beryl (great northern bean, susceptible), Bunsi (navy bean, low susceptibility), and G122 (cranberry bean, partial resistance) [@otto-hanson2011variation].
+We additionally wanted to determine categorical or phenotypic variables that best predicted genetic structure and if there was correlation between multilocus genotype and mycelial compatibility group.
+By investigating these aims, we hope to effectively describe the population structure of *S.
+sclerotiorum* in the USA and make available our database of isolates for use in future dry bean breeding efforts.
 

### 2017-08-14 doc/RMD/mlg-mcg.Rmd
Update figures to fix #9
Ironically, this doesn't explicitly simplify, it's now a bit more complex in content! :joy_cat:

I've also started saving descriptive figure names.
At some point I will burn all the figures down and only the descriptive names will be spared.

@@
   scale_y_continuous(expand = c(0, 2)) +
+  # scale_alpha(breaks = c(0.5, 0.75, 1)) +
   theme_minimal(base_size = 16, base_family = "Helvetica") +
@@
 
-vp1 <- grid::viewport(width = 0.25, height = 0.4, x = 1, y = 0.975, just = c("right", "top"))
-mcg_graph <- mcg_graph + theme(legend.position = c(.763, .75))
+# vp1 <- grid::viewport(width = 0.25, height = 0.4, x = 1, y = 0.975, just = c("right", "top"))
+# mcg_graph <- mcg_graph + theme(legend.position = c(.763, .75))
+# print(mcg_graph)
+# print(even, vp = vp1)
 
-print(mcg_graph)
-print(even, vp = vp1)
-
+mcg_graph_inset <- cowplot::ggdraw() + 
+  cowplot::draw_plot(mcg_graph, 0, 0, 1, 1) + 
+  cowplot::draw_plot(even, 0.75, 0.5675, 0.25, 0.4)
+mcg_graph_inset
 if (!interactive()){
-  pdf(file.path(PROJHOME, "results/figures/publication/Figure1Z.pdf"), width = 10, height = 5)
-  print(mcg_graph)
-  print(even, vp = vp1)
-  dev.off()
+  cowplot::ggsave(mcg_graph_inset, 
+                  filename = file.path(PROJHOME, "results/figures/publication/mcg-barplot.pdf"),
+                  width = 10,
+                  height = 5)
+  # pdf(file.path(PROJHOME, "results/figures/publication/Figure1Z.pdf"), width = 10, height = 5)
+  # print(mcg_graph)
+  # print(even, vp = vp1)
+  # dev.off()
 }
@@
   g <- gdf %>% 
-    select(MCG, MLG, Freq) %>%
+    select(MLG, MCG, Freq) %>%
     rename(weight = Freq) %>%
@@
                      width = weight),
-                 arrow = arrow(length = unit(0.01, "native"))) +
+                 arrow = arrow(length = unit(0.005, "native"))) +
   geom_node_text(aes(label = gsub("MLG.", "", name), 
@@
 if (!interactive()){
-  ggsave(plot = mcg_mlg_graph, filename = file.path(PROJHOME, "results/figures/publication/FigureS2.pdf"), 
+  ggsave(plot = mcg_mlg_graph, filename = file.path(PROJHOME, "results/figures/publication/full-graph.pdf"), 
        width = 88*3, height = 88*3.5, units = "mm")
@@
 }
-
 cat("Community Structure:\n")
 cluster_walktrap(g)
+```
+
+Because showing this big, hairy graph can overload a reader, I'm creating an
+inset graph for later use.
+
+```{r mlgmcggraph2, fig.width = 2, fig.height = 2}
+mcg_mlg_graph2 <- ggraph(lay2) +
+  geom_edge_arc(#aes(width = weight), 
+                start_cap = circle(1.1, "mm"), 
+                end_cap = circle(1.1, "mm"), 
+                curvature = 0.1
+                ) + #arrow = arrow(length = unit(2, "mm"))) +
+  geom_node_point(aes(fill = type), pch = 21, size = 0.75) +
+  coord_fixed() +
+  scale_edge_width(range = c(0.25, 1.5)) +
+  scale_fill_manual(values = c("black", "white")) +
+  theme_graph(base_size = 16, base_family = "Helvetica") +
+  theme(legend.position = "none", legend.direction = "vertical") +
+  # theme(legend.box.spacing = unit(-1, "lines")) +
+  theme(plot.margin = unit(c(0, 0, 0, 0), "lines")) +
+  theme(plot.background = element_rect(fill = "grey98", color = "black"))
+mcg_mlg_graph2
 ```
@@
 So that's a big hairy graph. What happens when we look at a subgraph of the
-top 5 MCGs?
+top 8 MCGs? Note, that these are all MCGs that have > 10 isolates associated.
 
 ```{r top5graph, fig.height = 7, fig.width = 7}
-top5 <- filter(mcgmlg, as.character(MCG) %in% mcgs$MCG[1:5])
+top5 <- filter(mcgmlg, as.character(MCG) %in% (mcgs %>% arrange(desc(Samples)) %>% slice(1:8) %>% pull(MCG)))
 top5g <- make_mcgmlg_graph(top5)
@@
 V(top5g)$size <- sqrt(tosize)/10
-set.seed(2017-05-08)
+# set.seed(2017-05-08)
+set.seed(2017-08-1)
 top5lay <- create_layout(top5g, layout = "igraph", algorithm = "fr", weights = rep(1, length(E(top5g))))
 top5lay$community <- inner_join(data_frame(vertex = top5lay$name), the_communities)$comm
-t5g <- mcg_mlg_graph %+% top5lay + ggtitle("Top 5 Mycelial Compatibility Groups and associated MLGs")
+levels(top5lay$type) <- c("Multilocus Genotype", "Mycelial Compatibility\nGroup")
+t5g <- ggraph(top5lay) +
+  geom_node_circle(aes(r = size, fill = type)) +
+  geom_edge_link(aes(start_cap = circle(node1.size, unit = "native"), 
+                     end_cap = circle(node2.size, unit = "native"), 
+                     width = weight),
+                 arrow = arrow(length = unit(0.01, "native"))) +
+  geom_node_text(aes(label = gsub("MLG.", "", name), 
+                     color = type, 
+                     size = size/10),
+                 show.legend = FALSE) +
+  coord_fixed() +
+  scale_edge_width(range = c(0.25, 1.5)) +
+  scale_fill_manual(values = c("black", "white")) +
+  scale_color_manual(values = c("white", "black")) +
+  theme_graph(base_size = 16, base_family = "Helvetica") +
+  # theme(legend.position = c(0.85, 0.35), legend.direction = "vertical") +
+  theme(legend.position = c(0.5, 0), legend.direction = "horizontal", legend.box = "horizontal") +
+  theme(legend.box.spacing = unit(-1, "lines")) +
+  theme(legend.spacing.x = unit(0, "lines")) +
+  # theme(legend.box.background = element_rect(color = "black")) +
+  theme(legend.text = element_text(size = 10)) +
+  theme(plot.margin = unit(c(0, 0, 0, 0), "lines"))
+t5g$labels$edge_width <- "shared isolates"
+t5g$labels$fill <- ""
 t5g
+
 if (!interactive())
-  ggsave(t5g, file = file.path("results", "figures", "publication", "FigureS2B.pdf"), width = 88, height = 88, scale = 3, units = "mm")
+  ggsave(t5g, file = file.path("results", "figures", "publication", "mcg-mlg-subgraph.pdf"), width = 88, height = 88, scale = 3, units = "mm")
 ```
@@
 Vey nice!
+
+```{r subgraphs, fig.height = 10, fig.width = 8}
+mcg_theme <- theme(legend.text = element_text(size = 10)) +
+  theme(legend.title = element_text(size = 10)) +
+  theme(axis.text.x = element_blank()) +
+  # theme(legend.background = element_blank()) +
+  # theme(legend.box.background = element_blank()) +
+  theme(legend.key.size = unit(0.5, "line")) +
+  theme(legend.position = c(0.9, 0.5)) +
+  theme(legend.box = "vertical") +
+  theme(legend.box.margin = unit(c(0, 0, 0, 0), "lines")) +
+  theme(legend.spacing = unit(0, "lines")) + 
+  theme(aspect.ratio = 0.45) +
+  theme(plot.background = element_rect(color = "black", fill = "grey98"))
+mcg_subgraph_with_context <- cowplot::ggdraw(xlim = c(0, 2), ylim = c(0, 2.25)) +
+  cowplot::draw_plot(t5g, 0, -0.125, 2, 2) +
+  cowplot::draw_plot(mcg_graph + scale_alpha(breaks = c(0.5, 0.75, 1)) + mcg_theme, 0, 1.45, 1.35, 1) +
+  cowplot::draw_plot(mcg_mlg_graph2, x = 1.355, y = 1.633, width = 0.6345, height = 0.6345) +
+  cowplot::draw_plot_label(c("A", "B", "C"), x = c(0, 1.35, 0), y = c(2.25, 2.25, 1.65), size = 25)
+mcg_subgraph_with_context
+if (!interactive()){
+  cowplot::ggsave(plot = mcg_subgraph_with_context,
+                  filename = file.path(PROJHOME, "results", "figures", "publication", "mcg-subgraph-with-context.pdf"),
+                  width = 8, 
+                  height = 10)
+}
+```
 

### 2017-08-14 doc/manuscript/manuscript.Rmd
replace figure 3 in manuscript to further address #9

@@
 Four regions had fewer than 10 isolates, Australia (6), WI (2), NY (1), ID (1).
-We observed 87 mycelial compatibility groups (MCG), the most abundant of which (MCG 5) represented 73 isolates over 37 multilocus genotypes (MLG) (Fig \@ref(barplot)).
+We observed 87 mycelial compatibility groups (MCG), the most abundant of which (MCG 5) represented 73 isolates over 37 multilocus genotypes (MLG) (Fig \@ref(biggraph)A).
 After clone-correction on the hierarchy of Region/Source/Host/Year, we were left with 318 isolates representing 165 MLGs (Table \@ref(tab:iatab)).
@@
 
-Our network-based approach to correlating MLGs with MCGs revealed a large and complex network (Fig. \@ref(biggraph)). 
+Our network-based approach to correlating MLGs with MCGs revealed a large and complex network (Fig. \@ref(biggraph), \@ref(fullgraph)). 
 Community analysis showed 51 communities, 15 of which consisted of a single MLG unconnected with any other community indicating that just 9.09% of the 165 MLGs are unable to cross with any other MLG in this data set.
@@
 ![
-Graph showing complex associations between Mycelial Compatibility Groups (dotted nodes) and Multilocus Genotypes (full nodes). 
+Data show complex associations between Mycelial Compatibility Groups and Multilocus Genotypes.
+**A)** Barplot showing abundance of each Mycelial Compatibility Group (MCG) in descending order.
+Singletons (46) have been removed for clarity, leaving 41 MCGs.
+White bars represent sample counts and grey bars represent counts of unique multilocus genotypes (MLG).
+The transparency associated with the grey bars represent the evenness of the distribution of the MLGs within a given MCG.
+**B)** Full graph-representation of the relationship between MCGs (open circles) and MLGs (filled circles). 
+Details in fig. \@ref(fullgraph). 
+**C)** A subset of **B** representing the 8 most common MCGs and their associated MLGs.
 Node size reflect the number of samples represented by each node (circle).
-Edges (arrows) point from Mycelial Compatibility Group to Multilocus Genotype where the weight (thickness) of the edge represents the number of samples shared.
-Node color represents the community assignment based on the walktrap algorithm with a maximum of four steps [@pons2006computing].
-](../../results/figures/publication/FigureS2.pdf){#biggraph width=100%}
+Edges (arrows) point from Multilocus Genotype to Mycelial Compatibility Group where the weight (thickness) of the edge represents the number of shared isolates.
+Edges extending from MLGs displayed to other MCGs are not shown.
+](../../results/figures/publication/mcg-subgraph-with-context.pdf){#biggraph width=100%}
 
@@
 When we inspected the biplot of the initial results, we saw that the most important predictors were MCGs 44, 5, and 9. 
-We believe that this is driven by the fact that these particular MCGs have uneven MLG distributions, meaning that they are heavily associated with one particular MLG (Fig.
-\@ref(biggraph), \@ref(barplot)).
+We believe that this is driven by the fact that these particular MCGs have uneven MLG distributions, meaning that they are heavily associated with one particular MLG (Fig. \@ref(biggraph)).
 We note these results with caution because of the apparent multicolinearity between MCG and Region, which is a violation of the analysis [@legendre1999distance].
@@
 It can be seen in Fig.
-\@ref(barplot) that the most abundant MCG contains several MLGs, but the diversity of those MLGs are low as indicated by the evenness (transparency), which indicates that there is one dominant MLG (MLG 25).
-What is not shown in Fig. \@ref(barplot) is the MLGs that are shared between MCGs. 
-This is illustrated in both table \@ref(tab:mlg-table) and Fig. \@ref(biggraph). 
+\@ref(biggraph)A that the most abundant MCG contains several MLGs, but the diversity of those MLGs are low as indicated by the evenness (transparency), which indicates that there is one dominant MLG (MLG 25).
+What is not shown in Fig. \@ref(biggraph)A is the MLGs that are shared between MCGs. 
+This is illustrated in both table \@ref(tab:mlg-table) and Fig. \@ref(biggraph)B,C. 
 It could be argued, however that MLG 25, with its high value of $P_{sex}$ represents different true MLGs across the five MCGs it occupies, but this doesn't account for the overall structure of Fig.
-\@ref(biggraph) where, for example, MLG 75 ($P_{sex}$ = 1.81e^-4^) is compatible with 58 other genotypes through three MCG when the population structure of *S.
+\@ref(fullgraph) where, for example, MLG 75 ($P_{sex}$ = 1.81e^-4^) is compatible with 58 other genotypes through three MCG when the population structure of *S.
 sclerotiorum* is known to be clonal.
@@
 
-<!-- MCG abundance figure -->
-
 ![
-Barplot showing abundance of each Mycelial Compatibility Group (MCG) in descending order.
-Singletons (46) have been removed for clarity, leaving 41 MCGs.
-White bars represent sample counts and grey bars represent counts of unique multilocus genotypes (MLG).
-The transparency associated with the grey bars represent the evenness of the distribution of the MLGs within a given MCG as shown in the inset histograms.
-](../../results/figures/publication/Figure1Z.pdf){#barplot width=100%}
-
-<!-- end figure -->
+Graph showing complex associations between Mycelial Compatibility Groups (dotted nodes) and Multilocus Genotypes (full nodes). 
+Node size reflect the number of samples represented by each node (circle). 
+Edges (arrows) point from Mycelial Compatibility Group to Multilocus Genotype where the weight (thickness) of the edge represents the number of samples shared. 
+Node color represents the community assignment based on the walktrap algorithm with a maximum of four steps [@pons2006computing]. 
+](../../results/figures/publication/full-graph.pdf){#fullgraph width=100%}
 

### 2017-08-14 doc/manuscript/manuscript.Rmd
update manuscript up to genetic diversity
I also moved the poppr table to the main text and
futzed with it.

Things I can't address right now:

 - how sclerotia was moved from field to lab
 - whether or not the DS media was in-fact
   modified (I got that from Otto-Hanson)
 - manual scoring of alleles

Things I won't address:

 - Changing years to two-digit years: did you
   forget about Y2K? Four-digit years give more
   information and we have the space.

@@
 Several of the isolates used for this study were collected and reported in previous studies and were collected using the same methods [@otto-hanson2011variation].
-Broadly, isolates were collected from two sources, white mold screening nurseries (wmn), or producer fields.
+Broadly, isolates were collected from two sources, white mold screening nurseries (wmn) or producer fields.
 The white mold screening nurseries were 5m x 10m in size and maintained without applying fungicides to observe natural incidence of white mold.
@@
 
-Sclerotia of *S.
-sclerotiorum* were collected over several years from grower fields and/or white mold screening nurseries in 11 states of the United States of America, Mexico, France, and Australia (Table \@ref(tab:isolate-table)).
-The white mold screening nurseries in different regions are indicated by wmn (white mold screening nursery) and all other isolates came from commercial fields (Table \@ref(tab:isolate-table)).
+Sclerotia of *S. sclerotiorum* were collected over several years from grower fields and/or wmn in 11 states of the United States of America, Mexico, France, and Australia.
+The white mold screening nurseries in different regions are indicated by wmn and all other isolates came from commercial fields (Table \@ref(tab:isolate-table)).
 
@@
 MCG was determined as described previously through co-culturing pairs of 2-day-old isolates 2.5cm apart on Diana Sermons (DS) Medium [@cubeta2001mycelial]. 
-Incompatibility of different MCGs resulted in formation of a mycelial barrier of dead tissue (referred to as a barrage line) accompanied by the formation of sclerotia on either side of the barrage line, indicating the limits of the isolates' growth [@otto-hanson2011variation; @kohn1990mycelial; @leslie1993fungal].
+Incompatibility of different MCGs resulted in formation of a mycelial barrier of dead tissue (termed a barrage line) accompanied by formation of sclerotia on either side of the barrage line, indicating the limits of the isolates' growth [@otto-hanson2011variation; @kohn1990mycelial; @leslie1993fungal].
 Isolates were compared in a pairwise manner for each site and then representatives among sites were compared to determine mycelial compatibility groups by scoring compatible and incompatible interactions [@otto-hanson2011variation].
@@
 
-Aggressiveness of each isolate was assessed using a straw test as carried out in @otto-hanson2011variation that rated necrotic lesion size [@petzoldt1996straw; @teran2006modified].
+Aggressiveness of each isolate was assessed using a straw test as described in @otto-hanson2011variation that rated necrotic lesion size [@petzoldt1996straw; @teran2006modified].
 Briefly, the straw test uses 21-day-old G122 plants as the host in a greenhouse setting. 
-Clear drinking straws cut to 2.5cm and sealed were used to place two mycelial plugs of inoculum on the host plant after removing the everything &gt;2.5cm above the fourth node.
+Clear drinking straws cut to 2.5cm and sealed were used to place two mycelial plugs of inoculum on the host plant after removing all plant growth &gt;2.5cm above the fourth node.
 Measurements were taken after 8 days using the Modified Petzoldt and Dickson scale where a score of 1--9 where 1 is no disease and 9 is plant death [@petzoldt1996straw; @teran2006modified].
@@
 
-DNA was purified from 100 mg mycelium collected from an actively growing culture of each isolate using a Qiagen DNeasy Plant Mini Kit (Qiagen Inc., Valencia, CA) according to the manufacturer's instructions.
+DNA was purified from approximately 100 mg mycelium collected from an actively growing culture of each isolate using a Qiagen DNeasy Plant Mini Kit (Qiagen Inc., Valencia, CA) according to the manufacturer's instructions.
 
@@
 We genotyped *S. sclerotiorum* isolates using 16 microsatellite primer pairs developed previously [@sirjusingh2001characterisation].
-PCR was carried out as described using primers labeled with FAM fluorophore, and resulting PCR products were first resolved in a 1.5% agarose gel stained with ethidium bromide to ensure product was within the expected size range.
+PCR was carried out as described using primers labeled with FAM fluorophore, and resulting amplicons were first resolved in a 1.5% agarose gel stained with ethidium bromide to ensure product was within the expected size range prior to capillary electrophoresis.
 
@@
 
-## Data processing and Analyses {-}
+## Data processing {-}
 
@@
 We assessed the power of our 11 markers by generating a genotype accumulation curve in the R package *poppr* version 2.4.1, looking for evidence of saturation, which would indicate that loci were sufficiently sampled to adequately represent the full set of genotypes [@arnaud2007standardizing; @kamvar2015novel].
-To avoid including samples that were potentially collected from the same plant (which increases the probability of collecting the same sample more than once), data were clone-corrected on the region/source/host/year hierarchy for subsequent analyses.
-The clone-corrected data set included 318 isolates.
+To avoid including isolates potentially collected from the same plant (which increases the probability of collecting sclerotioa from the same point of infection more than once), data were clone-corrected on a hierarchy of region/source/host/year---meaning that one representative of each genotype was retained for each combined regions, source, host, and year---for subsequent analysis.
 
@@
 This method uses constrained ordinations on a distance matrix representing the response variable to delineate relative contribution of any number of independent explanatory variables. 
-We used this method to delineate the phenotypic (Aggressiveness, MCG), geographic (Region, Host, Location), and temporal (Year) components in influencing genetic composition of the populations.
+We used this method to delineate the phenotypic (Aggressiveness, Mycelial Compatibility Group (MCG)), geographic (Region, Host, Location), and temporal (Year) components in predicting genetic composition of the populations.
 <!-- RDA is a multivariate linear regression analysis of independent (explanatory: eg. environmental and phenotypic) variables on dependent (non-explanatory or response) variables. -->
-The distance matrix we used as the response variable was generated using Bruvo's genetic distance from clone-corrected data on the hierarchy of Region, Location, Host, Year as implemented in *poppr* version 2.4.1, which employed a stepwise mutation model for microsatellite data [@bruvo2004simple; @kamvar2014poppr].
+The distance matrix we used as the response variable was generated using Bruvo's genetic distance from clone-corrected data (procedure described above) as implemented in *poppr* version 2.4.1, which employed a stepwise mutation model for microsatellite data [@bruvo2004simple; @kamvar2014poppr].
 Because aggressiveness measures differed with each isolate, aggressiveness first averaged across clone-corrected isolates.
@@
 
-We used analysis of variance (ANOVA) to assess if aggressiveness could be explained by Region, mycelial compatibility group, or multilocus genotype. 
+We used analysis of variance (ANOVA) to assess if aggressiveness could be explained by Region, MCG, or multilocus genotype (MLG). 
 To minimize complications due to small sample sizes, we chose the top 10 MCGs, representing 56.5% of the isolates collected, the top 10 MLGs representing 26.7% of the isolates, and populations with more than five isolates.
@@
 
-We wanted to assess if there was correlation between multilocus genotypes (MLG) and mycelial compatibility groups (MCG). 
-This was done using a graph-based approach with community analysis.
-We first created a contingency table with MLGs and MCGs and converted it to a directed and weighted edgelist where each edge was a connection from an MCG to an MLG weighted by the number of samples shared in the connection.
+We wanted to assess if there was correlation between MLGs and MCGs. 
+In simple terms, this was done using a network-based approach where both MLGs and MCGs were considered nodes and the number of isolates in which they were found together was the strength of the connection between an MLG and and MCG node.
+The network-based approach allowed us to assess how "clumpy" the relationships between MLGs and MCGs were.
+In more specific terms, we first created a contingency table with MLGs and MCGs and converted it to a directed and weighted edgelist where each edge was a connection from an MCG to an MLG weighted by the number of samples shared in the connection.
 This was then converted to a bipartite graph where top nodes represented MCGs and bottom nodes represented MLGs.
@@
 All communities that had fewer than 10 members were then consolidated into one.
-These community definitions were then used to assess the average genetic distance (as defined by Bruvo's distance) within members of the community [@bruvo2004simple].
+Community definitions were used to assess the average genetic distance (as defined by Bruvo's distance) within members of the community [@bruvo2004simple].
 
@@
 ## Genotypic Structure {-}
+
+```{r table_ia, echo = FALSE, warning = FALSE, message = FALSE, results = "asis"}
+suppressPackageStartupMessages(library("tidyverse"))
+read_csv("../../results/tables/genotype_diversity_table_region.csv", 
+         na = "NA", # spaces read in as spaces
+         col_types = cols(
+         Pop = col_character(),
+         N = col_integer(),
+         MLG = col_integer(),
+         eMLG = col_double(),
+         SE = col_double(),
+         H = col_double(),
+         G = col_double(),
+         lambda = col_double(),
+         E.5 = col_double(),
+         Hexp = col_double(),
+         Ia = col_double(),
+         p.Ia = col_double(),
+         rbarD = col_double(),
+         p.rD = col_double(),
+         File = col_character()
+      )) %>%
+  # mutate(Ia = case_when(
+  #   !is.finite(p.Ia) ~ "-",
+  #   p.Ia == 0.001 ~ paste0(round(Ia, 2), "*"),
+  #   p.Ia <= 0.01  ~ paste0(round(Ia, 2), "-"),
+  #   p.Ia <= 0.05  ~ paste0(round(Ia, 2), "~"),
+  #   TRUE          ~ as.character(round(Ia, 2))
+  # )) %>%
+  mutate(rbarD = case_when(
+    !is.finite(p.rD) ~ "-",
+    p.rD == 0.001 ~ paste0(round(rbarD, 2), "*"),
+  #   p.rD <= 0.01  ~ paste0(round(rbarD, 2), "-"),
+  #   p.rD <= 0.05  ~ paste0(round(rbarD, 2), "~"),
+    TRUE          ~ as.character(round(rbarD, 2))
+  )) %>%
+  mutate_if(is.numeric, round, digits = 2) %>%
+  mutate(N = paste(N, MLG, sep = " (") %>% paste0(")") %>% formatC()) %>%
+  mutate(eMLG = paste(eMLG, SE, sep = " (") %>%  paste0(")") %>% formatC()) %>%
+  rename(`*h*` = Hexp) %>%
+  rename(`$E_5$` = E.5) %>%
+  # rename(`$I_A$` = Ia) %>%
+  rename(`$\\lambda$` = lambda) %>%
+  rename(`$\\bar{r}_d$` = rbarD) %>%
+  select(-File, -MLG, -SE, -p.Ia, -p.rD, -Ia) %>% 
+  huxtable::as_huxtable(add_colnames = TRUE) %>%
+  huxtable::set_number_format(huxtable::everywhere, 4:5, 1) %>%
+  huxtable::set_align(huxtable::everywhere, -1, "right") %>%
+  huxtable::set_col_width(c(1.1, 0.9, 1.3, 0.5, 0.5, 1.1, 0.7, 0.6, 0.7)) %>%
+  huxtable::print_md(max_width = 80)
+```
+
+Table: (\#tab:iatab) Genotypic diversity and Linkage Disequilibrium summary for geographic populations arranged by abundance after clone-correction by a hierarchy of Region/Source/Host/Year. 
+Pop = Population, N = number of individuals (number of MLG in parentheses), 
+eMLG = expected number of MLGs based on rarefaction at min(10, N) individuals (standard error in parentheses), 
+H = Shannon-Weiner Index, G = Stoddardt and Taylor's Index, 
+$\lambda$ = Simpson's Index, 
+*h* = Nei's 1978 gene diversity, 
+$E_5$ = Evenness,
+$\bar{r}_d$ = standardized index of association.
+An asterix indicates a significant value of $\bar{r}_d$ after 999 permutations, $P \leq 0.001$.
 
@@
 
-```{r table_ia, echo = FALSE, warning = FALSE, message = FALSE, results = "asis"}
-suppressPackageStartupMessages(library("tidyverse"))
-read_csv("../../results/tables/genotype_diversity_table_region.csv", 
-         na = "NA", # spaces read in as spaces
-         col_types = cols(
-         Pop = col_character(),
-         N = col_integer(),
-         MLG = col_integer(),
-         eMLG = col_double(),
-         SE = col_double(),
-         H = col_double(),
-         G = col_double(),
-         lambda = col_double(),
-         E.5 = col_double(),
-         Hexp = col_double(),
-         Ia = col_double(),
-         p.Ia = col_double(),
-         rbarD = col_double(),
-         p.rD = col_double(),
-         File = col_character()
-      )) %>%
-  mutate(Ia = case_when(
-    !is.finite(p.Ia) ~ "-",
-    p.Ia == 0.001 ~ paste0(round(Ia, 2), "*"),
-    p.Ia <= 0.01  ~ paste0(round(Ia, 2), "-"),
-    p.Ia <= 0.05  ~ paste0(round(Ia, 2), "~"),
-    TRUE          ~ as.character(round(Ia, 2))
-  )) %>%
-  mutate(rbarD = case_when(
-    !is.finite(p.rD) ~ "-",
-    p.rD == 0.001 ~ paste0(round(rbarD, 2), "*"),
-    p.rD <= 0.01  ~ paste0(round(rbarD, 2), "-"),
-    p.rD <= 0.05  ~ paste0(round(rbarD, 2), "~"),
-    TRUE          ~ as.character(round(rbarD, 2))
-  )) %>%
-  mutate_if(is.numeric, round, digits = 2) %>%
-  mutate(N = paste(N, MLG, sep = " (") %>% paste0(")") %>% formatC()) %>%
-  mutate(eMLG = paste(eMLG, SE, sep = " (") %>%  paste0(")") %>% formatC()) %>%
-  rename(`*h*` = Hexp) %>%
-  rename(`$E_5$` = E.5) %>%
-  rename(`$I_A$` = Ia) %>%
-  rename(`$\\lambda$` = lambda) %>%
-  rename(`$\\bar{r}_d$` = rbarD) %>%
-  select(-File, -MLG, -SE, -p.rD, -p.Ia) %>% 
-  huxtable::as_huxtable(add_colnames = TRUE) %>%
-  huxtable::set_number_format(huxtable::everywhere, 4:5, 1) %>%
-  huxtable::set_align(huxtable::everywhere, -1, "right") %>%
-  huxtable::set_col_width(c(1.1, 0.9, 1.3, 0.5, 0.5, 1.2, 0.7, 0.6, 0.6, 0.8)) %>%
-  huxtable::print_md(max_width = 80)
-```
-
-Table: (\#tab:iatab) Genotypic diversity and Linkage Disequilibrium summary for geographic populations arranged by abundance after clone-correction by a hierarchy of Region/Source/Host/Year. 
-Pop = Population, N = number of individuals (number of MLG in parentheses), 
-eMLG = expected number of MLGs based on rarefaction at min(10, N) individuals (standard error in parentheses), 
-H = Shannon-Weiner Index, G = Stoddardt and Taylor's Index, 
-$\lambda$ = Simpson's Index, 
-*h* = Nei's 1978 gene diversity, 
-$E_5$ = Evenness,
-$I_A$ = the index of association, 
-$\bar{r}_d$ = standardized index of association. 
-P-value symbols: \* = 0.001, - = $\leq 0.01$, ~ = $\leq 0.05$.
-
 ![

### 2017-08-15 doc/RMD/MLG-distribution.Rmd
Add graphical legend to mlg distribution plot

--- 2017-07-12 doc/RMD/MLG-distribution.Rmd
@@
     labs(list(
-      title = "Shared haplotypes across regions",
-      fill = "Number of\nhaplotypes",
+     # title = "Shared MLGs across regions",
+      fill = "Number of\nMLGs",
       edge_alpha = "Probability of\nsecond encounter",
-      edge_width = "Populations\nper haplotype",
-      caption = "Outer circle: Number of haplotypes in the region\nInner Circle: Number of private haplotypes in the region"
+      edge_width = "Populations\nper MLG"#,
+      #caption = "Outer circle: Number of MLGs in the region\nInner Circle: Number of private MLGs in the region"
     ))
@@
 
+## Graph legend
+
+```{r}
+g <- graph_from_adjacency_matrix(matrix(0, 2, 2))
+V(g)$size <- c(56, 56 - 32)
+V(g)$label <- c("Number of MLGs in Region", "Number of private MLGs")
+lay <- as.data.frame(matrix(0, 2, 2, dimnames = list(NULL, c("x", "y"))))
+glegend <- create_layout(g, "manual", node.positions = as.data.frame(lay)) %>% ggraph() + 
+  geom_node_circle(aes(r = size, fill = size)) + 
+  coord_fixed() + 
+  viridis::scale_fill_viridis(option = "C", begin = 1 - 32/56, end = 1, guide = "none") + 
+  ggrepel::geom_text_repel(aes(label = label), 
+                           x = c(20, 0) + 25, 
+                           y = c(35, 0), 
+                           nudge_x = 350, 
+                           segment.size = 1, 
+                           arrow = arrow(length = unit(0.15, "native")), 
+                           family = "Helvetica") + 
+  theme_void() +
+  # theme(plot.background = element_rect(color = "black", fill = "grey98")) +
+  xlim(c(-60, 480))
+
+glegend
+```
+
 ### Individual population subgraphs
@@
 
-```{r graph-16-loci, fig.width = 7, fig.height = 7}
+```{r graph-16-loci, fig.width = 7, fig.height = 6.25}
 par(mfrow = c(1, 1))
 (g16o <- cluster_optimal(graph16loc$total))
-plot_mlg_graph(graph16loc$total, alt_layout) + labs(list(subtitle = "(16 loci)"))
-ggsave(filename = "results/figures/publication/FigureS1Z.pdf", width = 88, height = 88, units = "mm", scale = 2)
+gg16 <- plot_mlg_graph(graph16loc$total, alt_layout) #+ labs(list(subtitle = "(16 loci)"))
+gg16
+ggsave(filename = "results/figures/publication/FigureS1Z.pdf", width = 88, height = (6.25/7)*88, units = "mm", scale = 2)
+```
+
+```{r, fig.width = 7, fig.height = 6.25}
+suppressWarnings(
+  gg16cp <- cowplot::ggdraw(xlim = c(0, 1), ylim = c(0, 1)) +
+  cowplot::draw_plot(gg16, x = 0, y = 0.05) +
+  cowplot::draw_plot(glegend, x = -0.125, y = 0, height = 0.1)
+  )
+gg16cp
+cowplot::ggsave(plot = gg16cp, 
+                file = file.path(PROJHOME, "results", "figures", "publication", "mlg-16.pdf"),
+                width = 88,
+                height = (6.25/7)*88,
+                units = "mm",
+                scale = 2)
 ```
@@
 
-```{r graph-11-loci, fig.width = 7, fig.height = 7}
+```{r graph-11-loci, fig.width = 7, fig.height = 6.25}
 (g11o <- cluster_optimal(graph11loc$total))
-plot_mlg_graph(graph11loc$total, alt_layout) + labs(list(subtitle = "(11 loci)"))
-ggsave(filename = "results/figures/publication/Figure3Z.pdf", width = 88, height = 88, units = "mm", scale = 2)
+gg11 <- plot_mlg_graph(graph11loc$total, alt_layout)
+gg11
+ggsave(filename = "results/figures/publication/Figure3Z.pdf", width = 88, height = (6.25/7)*88, units = "mm", scale = 2)
+```
+
+```{r, fig.width = 7, fig.height = 6.25}
+suppressWarnings(
+  gg11cp <- cowplot::ggdraw(xlim = c(0, 1), ylim = c(0, 1)) +
+  cowplot::draw_plot(gg11, x = 0, y = 0.05) +
+  cowplot::draw_plot(glegend, x = -0.125, y = 0, height = 0.1)
+  )
+gg11cp
+cowplot::ggsave(plot = gg11cp, 
+                file = file.path(PROJHOME, "results", "figures", "publication", "mlg-11.pdf"),
+                width = 88,
+                height = (6.25/7)*88,
+                units = "mm",
+                scale = 2)
 ```

### 2017-08-15 doc/RMD/mlg-mcg.Rmd
Add rectangle to mcg plot delimiting 8 MCGs

@@
   geom_col(aes(width = ifelse(Type == "MLGs", 0.5, 0.85)), color = "black", position = "identity") +
+  annotate(geom = "rect", xmin = 0.5, xmax = 8.5, ymin = 0, ymax = 75, color = "black", size = 1.5, alpha = 0, lty = 2) +
   # annotate(geom = "text", x = 18, y =35, 
@@
   theme(legend.box = "horizontal") +
+  theme(legend.position = c(0.77, 0.75)) +
   theme(aspect.ratio = 1/2) +
@@
                   width = 10,
-                  height = 5)
+                  height = 5,
+                  units = "in")
   # pdf(file.path(PROJHOME, "results/figures/publication/Figure1Z.pdf"), width = 10, height = 5)
@@
 ```{r subgraphs, fig.height = 10, fig.width = 8}
+greyground <- theme(plot.background = element_rect(color = "black", fill = "grey98"))
 mcg_theme <- theme(legend.text = element_text(size = 10)) +
@@
   # theme(legend.background = element_blank()) +
-  # theme(legend.box.background = element_blank()) +
-  theme(legend.key.size = unit(0.5, "line")) +
+  theme(legend.box.background = element_rect(fill = "grey98")) +
+  theme(legend.key.size = unit(0.75, "line")) +
+  theme(legend.box.just = "left") +
   theme(legend.position = c(0.9, 0.5)) +
@@
   theme(aspect.ratio = 0.45) +
-  theme(plot.background = element_rect(color = "black", fill = "grey98"))
+  greyground
 mcg_subgraph_with_context <- cowplot::ggdraw(xlim = c(0, 2), ylim = c(0, 2.25)) +
@@
   cowplot::draw_plot(mcg_graph + scale_alpha(breaks = c(0.5, 0.75, 1)) + mcg_theme, 0, 1.45, 1.35, 1) +
-  cowplot::draw_plot(mcg_mlg_graph2, x = 1.355, y = 1.633, width = 0.6345, height = 0.6345) +
+  cowplot::draw_plot(mcg_mlg_graph2 + greyground, x = 1.355, y = 1.633, width = 0.6345, height = 0.6345) +
   cowplot::draw_plot_label(c("A", "B", "C"), x = c(0, 1.35, 0), y = c(2.25, 2.25, 1.65), size = 25)

### 2017-08-15 doc/RMD/pop-diff.Rmd
Fix DAPC plot by increasing pointsize and asp. ratio

@@
 LDS_PLOT <- ggplot(LDS, aes(x = LD1, y = LD2, color = Population)) + 
-  geom_point(aes(fill = Population), alpha = 0.5, pch = 21, color = "black") +
+  geom_point(aes(fill = Population), alpha = 0.5, pch = 21, color = "black", size = 2) +
   geom_segment(aes(x = mean1, y = mean2, xend = LD1, yend = LD2), alpha = 0.5) +
@@
   theme_bw() +  
-  theme(aspect.ratio = 1/1.618) +
+  # theme(aspect.ratio = 1/1.618) +
   theme(legend.position = "bottom") +
@@
   geom_segment(aes(xend = Population, yend = 0)) +
-  geom_point(aes(fill = Source), pch = 21) + 
+  geom_point(aes(fill = Source), pch = 21, size = 2) + 
   facet_wrap(~`Original Population`, ncol = 3, strip.position = "top") +
@@
 ```{r dapc-region-dualplot, fig.width = 13, fig.height = 6}
-cowplot::plot_grid(REGION_PLOT, LDS_PLOT, 
+LEG <- cowplot::get_legend(LDS_PLOT + theme(legend.position = "right") + theme(legend.text = element_text(size = 12)))
+cowplot::plot_grid(REGION_PLOT +
+                     theme(plot.margin = unit(c(0, 0, 0, 0), "lines")), 
+                   LDS_PLOT + 
+                     theme(legend.position = "none") +
+                     theme(plot.margin = unit(c(0, 0, 0, 0), "lines")) +
+                     theme(aspect.ratio = 0.85), 
+                   LEG,
                    nrow = 1, 
-                   rel_widths = c(1, 0.85),
-                   label_size = 16,
-                   labels = "auto")
-ggsave(filename = "results/figures/publication/DAPC.pdf")
+                   rel_widths = c(1, 0.85, 0.25),
+                   label_size = 25,
+                   label_fontfamily = "Helvetica",
+                   labels = c("A", "B", NA),
+                   align = "v",
+                   axis = "t")
+cowplot::ggsave(filename = file.path(PROJHOME, "results/figures/publication/DAPC.pdf"),
+       width = 13,
+       height = 6
+      )
 ```

### 2017-08-15 doc/manuscript/manuscript.Rmd
Address comments up to Discussion

@@
 There were no shared microsatellite multilocus genotypes (MLG) between the two populations.
-Sexual outcrossing was detected in both populations with higher recombination rate in Chinese population than the US population.
+Sexual outcrossing was detected in both populations with higher recombination rate in Chinese population than the USA population.
 Additionally, phenotypic traits such as production of sclerotia, mycelial compatibility group (MCG), growth rate were also significantly different between the two populations.
@@
 To date, no commercial white mold-resistant dry bean cultivars exist. 
-Field trials in some regions of the US have shown the G122 cultivar is partially resistant while the Beryl cultivar is susceptible [@otto-hanson2011variation; @steadman2004identification].
+Field trials in some regions of the USA have shown the G122 cultivar is partially resistant while the Beryl cultivar is susceptible [@otto-hanson2011variation; @steadman2004identification].
 <!-- This puny paragraph needs a lot of work to more fully describe the operation of the wmn -->
@@
 In combination with the 156 isolates described previously, we collected 210 isolates for a total of 366 isolates [@otto-hanson2011variation].
+
+## Mycelial Compatibility {-}
+
+MCG was determined as described previously through co-culturing pairs of 2-day-old isolates 2.5cm apart on Diana Sermons (DS) Medium [@cubeta2001mycelial]. 
+Incompatibility of different MCGs resulted in formation of a mycelial barrier of dead tissue (termed a barrage line) accompanied by formation of sclerotia on either side of the barrage line, indicating the limits of the isolates' growth [@otto-hanson2011variation; @kohn1990mycelial; @leslie1993fungal].
+Isolates were compared in a pairwise manner for each site and then representatives among sites were compared to determine mycelial compatibility groups by scoring compatible and incompatible interactions [@otto-hanson2011variation].
+
+<!-- ![
+Examples of mycelial compatibility results in *S. sclerotiorum*. 
+A) Compatible reaction showing aerial hyphae.
+B) Incompatible reaction showing barrage reaction of dead cells at point of contact.
+](../../results/figures/publication/incompatibility2.JPG){#incompatibility width=25%} -->
+
+## Aggressiveness {-}
+
+Aggressiveness of each isolate was assessed using a straw test as described in @otto-hanson2011variation that rated necrotic lesion size [@petzoldt1996straw; @teran2006modified].
+Briefly, the straw test uses 21-day-old G122 plants as the host in a greenhouse setting. 
+Clear drinking straws cut to 2.5cm and sealed were used to place two mycelial plugs of inoculum on the host plant after removing all plant growth &gt;2.5cm above the fourth node.
+Measurements were taken after 8 days using the Modified Petzoldt and Dickson scale where a score of 1--9 where 1 is no disease and 9 is plant death [@petzoldt1996straw; @teran2006modified].
+
+## Microsatellite genotyping {-}
+
+DNA was purified from approximately 100 mg mycelium collected from an actively growing culture of each isolate using a Qiagen DNeasy Plant Mini Kit (Qiagen Inc., Valencia, CA) according to the manufacturer's instructions.
+
+> SEE: More details on how tissue was disrupted may be needed here.
+
+We genotyped *S. sclerotiorum* isolates using 16 microsatellite primer pairs developed previously [@sirjusingh2001characterisation].
+PCR was carried out as described using primers labeled with FAM fluorophore, and resulting amplicons were first resolved in a 1.5% agarose gel stained with ethidium bromide to ensure product was within the expected size range prior to capillary electrophoresis.
+
+Capillary electrophoresis (fragment analysis) of amplicons, with size standard GeneScan™ 500 LIZ®, was performed using an ABI 3730 genetic analyzer (Life Technologies Corporation, Carlsbad, CA) at the Michigan State University Genomic Sequencing Center.
+Alleles were scored using PeakScanner version 1.0 (Life Technologies Corporation, Carlsbad, CA) and recorded manually in an excel sheet.
+
+## Data processing and Analysis{-}
+
+All analyses were performed on this data set unless otherwise mentioned.
+All data processing and analyses were performed in R version 3.4.1 [@R] and are openly available and reproducible at https://github.com/everhartlab/sclerotinia-366/.
+Of the 16 microsatellite loci genotyped, five included compound repeats, which made it challenging to accurately/confidently bin alleles into fragment sizes expected for each locus based on the described repeat motif.
+<!-- add something about repeat length consistency  -->
+Loci with compound repeats were removed for the reported statistics, but were processed and included in data analyses to ensure the integrity of the results with 11 loci as compared to 16.
+We assessed the power of our 11 markers by generating a genotype accumulation curve in the R package *poppr* version 2.4.1, looking for evidence of saturation, which would indicate that loci were sufficiently sampled to adequately represent the full set of genotypes [@arnaud2007standardizing; @kamvar2015novel].
+To avoid including isolates potentially collected from the same plant (which increases the probability of collecting sclerotioa from the same point of infection more than once), data were clone-corrected on a hierarchy of region/source/host/year---meaning that one representative of each genotype was retained for each combined regions, source, host, and year---for subsequent analysis.
+We assessed genotypic diversity by calculating Stoddart and Taylor's index (*G*) [@stoddart1988genotypic], Shannon's index (*H*) [@shannon2001mathematical], Simpson's index ($\lambda$) [@simpson1949measurement], and evenness ($E_5$, *eMLG*) [@heck1975explicit; @hurlbert1971nonconcept; @pielou1975ecological; @grunwald2003analysis].
+To assess the potential for random mating, we tested for linkage disequilibrium with the index of association, $I_A$ and its standardized version, $\bar{r}_d$ using 999 permutations [@agapow2001indices; @smith1993how; @brown1980multilocus].
+Both genotypic diversity and linkage disequilibrium were calculated in *poppr* version 2.4.1 [@kamvar2014poppr].
+
+## Assessing Importance of Variables  {-}
+
+### Distance-based Redundancy Analysis {-}
+
+A distance-based redundancy analysis (dbRDA) [@legendre1999distance] was performed with the function `capscale()` in the *vegan* package version 2.4-3 [@vegan].
+This method uses constrained ordinations on a distance matrix representing the response variable to delineate relative contribution of any number of independent explanatory variables. 
+We used this method to delineate the phenotypic (Aggressiveness, Mycelial Compatibility Group (MCG)), geographic (Region, Host, Location), and temporal (Year) components in predicting genetic composition of the populations.
+<!-- RDA is a multivariate linear regression analysis of independent (explanatory: eg. environmental and phenotypic) variables on dependent (non-explanatory or response) variables. -->
+The distance matrix we used as the response variable was generated using Bruvo's genetic distance from clone-corrected data (procedure described above) as implemented in *poppr* version 2.4.1, which employed a stepwise mutation model for microsatellite data [@bruvo2004simple; @kamvar2014poppr].
+Because aggressiveness measures differed with each isolate, aggressiveness first averaged across clone-corrected isolates.
+<!-- In our case, a pairwise Bruvo's [@bruvo2004simple] distance matrix was generated from microsatellite data using *poppr* 2.4.1 [@kamvar2014poppr; @kamvar2015novel] and employed it as the dependent matrix for RDA analysis.
+-->
+To identify explanatory variable(s) correlated with genetic variation, a forward-backward selection process was applied with the vegan function `ordistep()` to identify the set of variables that explained more variation than predicted with the global model containing all variables.
+An ANOVA was then performed to test for significance of the reduced model and marginal effects using 999 permutations.
+The `varpart()` function of vegan was used to determine variation partitioning of explanatory variables.
+
+### Aggressiveness Assessment {-}
+
+We used analysis of variance (ANOVA) to assess if aggressiveness could be explained by Region, MCG, or multilocus genotype (MLG). 
+To minimize complications due to small sample sizes, we chose the top 10 MCGs, representing 56.5% of the isolates collected, the top 10 MLGs representing 26.7% of the isolates, and populations with more than five isolates.
+Between-group differences were assessed using Tukey's HSD test with an alpha of 0.05 using the `HSD.test()` function in the package *agricolae* version 1.2.4 [@mendiburu2015agricolae].
+
+### Correlating Multilocus Genotypes with Mycelial Compatibility Groups {-}
+
+We wanted to assess if there was correlation between MLGs and MCGs. 
+In simple terms, this was done using a network-based approach where both MLGs and MCGs were considered nodes and the number of isolates in which they were found together was the strength of the connection between an MLG and and MCG node.
+The network-based approach allowed us to assess how "clumpy" the relationships between MLGs and MCGs were.
+In more specific terms, we first created a contingency table with MLGs and MCGs and converted it to a directed and weighted edgelist where each edge was a connection from an MCG to an MLG weighted by the number of samples shared in the connection.
+This was then converted to a bipartite graph where top nodes represented MCGs and bottom nodes represented MLGs.
+The cluster walktrap community detection algorithm as implemented in the `cluster_walktrap()` function in *igraph* version 1.1.2 was used to define clusters of MLGs and MCGs [@csardi2006igraph; @pons2006computing].
+All communities that had fewer than 10 members were then consolidated into one.
+Community definitions were used to assess the average genetic distance (as defined by Bruvo's distance) within members of the community [@bruvo2004simple].
+
+## Genetic Diversity {-}
+
+### Population Differentiation {-}
+
+We used analysis of molecular variance (AMOVA) with Bruvo's genetic distance in *poppr* version 2.4.1 to test for differentiation associated with wmn by quantifying the differentiation conferred by Source and Region [@excoffier1992analysis; @bruvo2004simple; @kamvar2014poppr].
+Source in this case was defined as a population belonging to a wmn or not.
+
+To identify specific Regions that were more differentiated than others, we used discriminant analysis of principal components (DAPC) as implemented in *adegenet* version 2.1.0, assessing the per-sample posterior group assignment probability [@jombart2008adegenet].
+This method decomposes the genetic data into principal components, and then uses a subset of these as the inputs for discriminant analysis, which attempts to minimize within-group variation and maximize among-group variation [@jombart2010discriminant].
+To avoid over-fitting data, the optimal number of principal components was selected by using the *adegenet* function `xvalDapc()`. 
+This function implements a cross-validation procedure to iterate over an increasing number of principal components on a subset (90%) of the data, trying to find the minimum number of principal components that maximizes the rate of successful group reassignment.
+To assess if cultivar had an influence on genetic diversity between wmn, we first subset the clone-corrected data to contain only samples from wmn and from the cultivars Beryl, Bunsi, and G122 and tested differentiation using AMOVA and DAPC as described above.
+
+### Analysis of Shared Multilocus Genotypes {-}
+
+We wanted to evaluate patterns of connectivity between shared multilocus genotypes across geographic regions.
+We first tabulated the multilocus genotypes shared between at least two populations (defined as states or countries) with the *poppr* function `mlg.crosspop()` [@kamvar2014poppr].
+From these data, we constructed a graph with populations as nodes and shared genotypes as edges (connections) between nodes using the R packages *igraph* version 1.0.1 [@csardi2006igraph], *dplyr* version 0.7.1 [@dplyr], and *purrr* version 0.2.2.2 [@purrr].
+Each node was weighted by the fraction of shared MLGs.
+Each edge represented a single MLG, but because a single MLG could be present in more than one population, that MLG would be represented by (*n*\*(*n*-1))/2 edges where n represents the number of populations crossed.
+Edges were weighted by $1-P_{sex}$, where $P_{sex}$ is the probability of encountering the same genotype via two independent meiotic events [@arnaud2007standardizing; @parks1993study].
+We then used the `cluster_optimal()` function from *igraph* to define communities [@csardi2006igraph].
+The graph was plotted using the R packages *ggplot2* version 2.2.1 [@ggplot2] and *ggraph* version 1.0.0 [@ggraph].
+To ensure that we captured the same community signal, we additionally performed this analysis including the five polymorphic markers described above.
+
+# Results {-}
+
+A total of 366 isolates were collected from 2003 to 2012 (except 2006 and 2011) from diseased dry bean plants in Australia, Mexico, France, and 11 states in the USA (Table \@ref(tab:isolate-table)).
+Over 50% of the isolates came from four states, MI (62), ND (60), WA (59), NE (47). 
+Four regions had fewer than 10 isolates, Australia (6), WI (2), NY (1), ID (1).
+We observed 87 MCGs, the most abundant of which ('MCG 5') was represented by 73 isolates over 37 MLGs (Fig \@ref(biggraph)A,C).
+After clone-correction on the hierarchy of Region/Source/Host/Year, we were left with 318 isolates representing 165 MLGs (Table \@ref(tab:iatab)).
+Analysis of the genotype accumulation curve showed no clear plateau (Supplemental Information), indicating that we would likely obtain more multilocus genotypes if we were to genotype more loci.
+The number of observed alleles per locus ranged from two to 10 with an average of 6.27 (Table \@ref(tab:locus-stats)).
 
@@
 
-## Mycelial Compatibility {-}
-
-MCG was determined as described previously through co-culturing pairs of 2-day-old isolates 2.5cm apart on Diana Sermons (DS) Medium [@cubeta2001mycelial]. 
-Incompatibility of different MCGs resulted in formation of a mycelial barrier of dead tissue (termed a barrage line) accompanied by formation of sclerotia on either side of the barrage line, indicating the limits of the isolates' growth [@otto-hanson2011variation; @kohn1990mycelial; @leslie1993fungal].
-Isolates were compared in a pairwise manner for each site and then representatives among sites were compared to determine mycelial compatibility groups by scoring compatible and incompatible interactions [@otto-hanson2011variation].
-
-<!-- ![
-Examples of mycelial compatibility results in *S. sclerotiorum*. 
-A) Compatible reaction showing aerial hyphae.
-B) Incompatible reaction showing barrage reaction of dead cells at point of contact.
-](../../results/figures/publication/incompatibility2.JPG){#incompatibility width=25%} -->
-
-## Aggressiveness {-}
-
-Aggressiveness of each isolate was assessed using a straw test as described in @otto-hanson2011variation that rated necrotic lesion size [@petzoldt1996straw; @teran2006modified].
-Briefly, the straw test uses 21-day-old G122 plants as the host in a greenhouse setting. 
-Clear drinking straws cut to 2.5cm and sealed were used to place two mycelial plugs of inoculum on the host plant after removing all plant growth &gt;2.5cm above the fourth node.
-Measurements were taken after 8 days using the Modified Petzoldt and Dickson scale where a score of 1--9 where 1 is no disease and 9 is plant death [@petzoldt1996straw; @teran2006modified].
-
-## Microsatellite genotyping {-}
-
-DNA was purified from approximately 100 mg mycelium collected from an actively growing culture of each isolate using a Qiagen DNeasy Plant Mini Kit (Qiagen Inc., Valencia, CA) according to the manufacturer's instructions.
-
-> SEE: More details on how tissue was disrupted may be needed here.
-
-We genotyped *S. sclerotiorum* isolates using 16 microsatellite primer pairs developed previously [@sirjusingh2001characterisation].
-PCR was carried out as described using primers labeled with FAM fluorophore, and resulting amplicons were first resolved in a 1.5% agarose gel stained with ethidium bromide to ensure product was within the expected size range prior to capillary electrophoresis.
-
-Capillary electrophoresis (fragment analysis) of amplicons, with size standard GeneScan™ 500 LIZ®, was performed using an ABI 3730 genetic analyzer (Life Technologies Corporation, Carlsbad, CA) at the Michigan State University Genomic Sequencing Center.
-Alleles were scored using PeakScanner version 1.0 (Life Technologies Corporation, Carlsbad, CA) and recorded manually in an excel sheet.
-
-## Data processing {-}
-
-All analyses were performed on this data set unless otherwise mentioned.
-All data processing and analyses were performed in R version 3.4.1 [@R] and are openly available and reproducible at https://github.com/everhartlab/sclerotinia-366/.
-Of the 16 microsatellite loci genotyped, five included compound repeats, which made it challenging to accurately/confidently bin alleles into fragment sizes expected for each locus based on the described repeat motif.
-<!-- add something about repeat length consistency  -->
-Loci with compound repeats were removed for the reported statistics, but were processed and included in data analyses to ensure the integrity of the results with 11 loci as compared to 16.
-We assessed the power of our 11 markers by generating a genotype accumulation curve in the R package *poppr* version 2.4.1, looking for evidence of saturation, which would indicate that loci were sufficiently sampled to adequately represent the full set of genotypes [@arnaud2007standardizing; @kamvar2015novel].
-To avoid including isolates potentially collected from the same plant (which increases the probability of collecting sclerotioa from the same point of infection more than once), data were clone-corrected on a hierarchy of region/source/host/year---meaning that one representative of each genotype was retained for each combined regions, source, host, and year---for subsequent analysis.
-
-## Assessing Importance of Variables  {-}
-
-### Distance-based Redundancy Analysis {-}
-
-A distance-based redundancy analysis (dbRDA) [@legendre1999distance] was performed with the function `capscale()` in the *vegan* package version 2.4-3 [@vegan].
-This method uses constrained ordinations on a distance matrix representing the response variable to delineate relative contribution of any number of independent explanatory variables. 
-We used this method to delineate the phenotypic (Aggressiveness, Mycelial Compatibility Group (MCG)), geographic (Region, Host, Location), and temporal (Year) components in predicting genetic composition of the populations.
-<!-- RDA is a multivariate linear regression analysis of independent (explanatory: eg. environmental and phenotypic) variables on dependent (non-explanatory or response) variables. -->
-The distance matrix we used as the response variable was generated using Bruvo's genetic distance from clone-corrected data (procedure described above) as implemented in *poppr* version 2.4.1, which employed a stepwise mutation model for microsatellite data [@bruvo2004simple; @kamvar2014poppr].
-Because aggressiveness measures differed with each isolate, aggressiveness first averaged across clone-corrected isolates.
-<!-- In our case, a pairwise Bruvo's [@bruvo2004simple] distance matrix was generated from microsatellite data using *poppr* 2.4.1 [@kamvar2014poppr; @kamvar2015novel] and employed it as the dependent matrix for RDA analysis.
--->
-To identify explanatory variable(s) correlated with genetic variation, a forward-backward selection process was applied with the vegan function `ordistep()` to identify the set of variables that explained more variation than predicted with the global model containing all variables.
-An ANOVA was then performed to test for significance of the reduced model and marginal effects using 999 permutations.
-The `varpart()` function of vegan was used to determine variation partitioning of explanatory variables.
-
-### Aggressiveness Assessment {-}
-
-We used analysis of variance (ANOVA) to assess if aggressiveness could be explained by Region, MCG, or multilocus genotype (MLG). 
-To minimize complications due to small sample sizes, we chose the top 10 MCGs, representing 56.5% of the isolates collected, the top 10 MLGs representing 26.7% of the isolates, and populations with more than five isolates.
-Between-group differences were assessed using Tukey's HSD test with an alpha of 0.05 using the `HSD.test()` function in the package *agricolae* version 1.2.4 [@mendiburu2015agricolae].
-
-### Correlating Multilocus Genotypes with Mycelial Compatibility Groups {-}
-
-We wanted to assess if there was correlation between MLGs and MCGs. 
-In simple terms, this was done using a network-based approach where both MLGs and MCGs were considered nodes and the number of isolates in which they were found together was the strength of the connection between an MLG and and MCG node.
-The network-based approach allowed us to assess how "clumpy" the relationships between MLGs and MCGs were.
-In more specific terms, we first created a contingency table with MLGs and MCGs and converted it to a directed and weighted edgelist where each edge was a connection from an MCG to an MLG weighted by the number of samples shared in the connection.
-This was then converted to a bipartite graph where top nodes represented MCGs and bottom nodes represented MLGs.
-The cluster walktrap community detection algorithm as implemented in the `cluster_walktrap()` function in *igraph* version 1.1.2 was used to define clusters of MLGs and MCGs [@csardi2006igraph; @pons2006computing].
-All communities that had fewer than 10 members were then consolidated into one.
-Community definitions were used to assess the average genetic distance (as defined by Bruvo's distance) within members of the community [@bruvo2004simple].
-
-## Genetic Diversity {-}
-
-### Population Differentiation {-}
-
-We used analysis of molecular variance (AMOVA) with Bruvo's genetic distance in *poppr* version 2.4.1 to assess differentiation associated with wmn by quantifying the differentiation conferred by source and region [@excoffier1992analysis; @bruvo2004simple; @kamvar2014poppr].
-Source in this case was defined as a population belonging to a wmn or not and region defined as state.
-
-To identify specific regions that were more differentiated than others, we used discriminant analysis of principal components (DAPC) as implemented in *adegenet* version 2.1.0, assessing the per-sample posterior group assignment probability [@jombart2008adegenet].
-This method decomposes the genetic data into principal components, and then uses a subset of these as the inputs for discriminant analysis, which attempts to minimize within-group variation and maximize among-group variation [@jombart2010discriminant].
-To avoid over-fitting the data, the optimal number of principal components was selected by using the *adegenet* function `xvalDapc()`. 
-This function implements a cross-validation procedure to iterate over an increasing number of principal components on a subset (90%) of the data, trying to find the minimum number of principal components that maximizes the rate of successful group reassignment.
-We additionally assessed genotypic diversity by calculating Stoddart and Taylor's index (*G*) [@stoddart1988genotypic], Shannon's index (*H*) [@shannon2001mathematical], Simpson's index ($\lambda$) [@simpson1949measurement], and evenness ($E_5$, *eMLG*) [@heck1975explicit; @hurlbert1971nonconcept; @pielou1975ecological; @grunwald2003analysis].
-To assess the potential for random mating, we tested for linkage disequilibrium with the index of association, $I_A$ and its standardized version, $\bar{r}_d$ using 999 permutations [@agapow2001indices; @smith1993how; @brown1980multilocus].
-Both genotypic diversity and linkage disequilibrium were calculated in *poppr* version 2.4.1 [@kamvar2014poppr].
-
-### Cultivar Differentiation {-}
-
-To assess if cultivar had an influence on genetic diversity across wmn, we first subset the clone-corrected data to contain only samples from wmn and from the cultivars Beryl, Bunsi, and G122 and tested differentiation using AMOVA and DAPC as described above.
-
-### Shared Multilocus Genotype Community Analysis {-}
-
-We wanted to evaluate patterns of connectivity between shared multilocus genotypes across geographic regions.
-We first tabulated the multilocus genotypes shared between at least two populations (defined as states or countries) with the *poppr* function `mlg.crosspop()` [@kamvar2014poppr].
-From these data, we constructed a graph with populations as nodes and shared genotypes as edges (connections) between nodes using the R packages *igraph* version 1.0.1 [@csardi2006igraph], *dplyr* version 0.7.1 [@dplyr], and *purrr* version 0.2.2.2 [@purrr].
-Each node was weighted by the fraction of shared MLGs.
-Each edge represented a single MLG, but because a single MLG could cross populations, that MLG could be represented by (n\*(n-1))/2 edges where n represented the number of populations it crossed.
-Edges were weighted by $1-P_{sex}$, where $P_{sex}$ is the probability of encountering the same genotype via two independent meiotic events [@arnaud2007standardizing; @parks1993study].
-We then used the `cluster_optimal()` function from *igraph* to define communities [@csardi2006igraph].
-The graph was plotted using the R packages *ggplot2* version 2.2.1 [@ggplot2] and *ggraph* version 1.0.0 [@ggraph].
-To ensure that we captured the same community signal, we additionally performed this analysis including the 5 polymorphic markers described above.
-
-# Results {-}
-
-A total of 366 isolates were collected from 2003 to 2012 (except 2006 and 2011) from diseased dry bean plants in 11 states in the US, Australia, Mexico, and France (Table \@ref(tab:isolate-table)).
-Over 50% of the isolates came from four states, MI (62), ND (60), WA (59), NE (47). 
-Four regions had fewer than 10 isolates, Australia (6), WI (2), NY (1), ID (1).
-We observed 87 mycelial compatibility groups (MCG), the most abundant of which (MCG 5) represented 73 isolates over 37 multilocus genotypes (MLG) (Fig \@ref(biggraph)A).
-After clone-correction on the hierarchy of Region/Source/Host/Year, we were left with 318 isolates representing 165 MLGs (Table \@ref(tab:iatab)).
-Analysis of the genotype accumulation curve showed no clear plateau, indicating that we would likely obtain more multilocus genotypes if we were to collect more loci.
-The number of observed alleles per locus ranged from 2 to 10 with an average of 6.27 (Table \@ref(tab:locus-stats)).
-
 --------------------------------------------
@@
 
-The dbRDA was performed on clone-corrected isolates (N = 318) after grouping them by location, host, region, and year.
-The forward-backward selection process revealed year, region, host, and aggressiveness to be the optimal variables for the reduced model, accounting for 45% of the total variation.
-ANOVA revealed that the reduced model was significant with an adjusted R^2^ of 0.0674 (P = 0.001).
-Assessment of the marginal effects showed that region, host, year, and aggressiveness significantly explained genetic variation (P = 0.001, P = 0.009 for region).
-It was revealed that there was multicollinearity combining all variables with MCG, so we followed the same procedure, dropping MCG from the list of potential predictors. 
+The forward-backward selection process of the dbRDA models on clone-corrected data revealed Year, Region, Host, and Aggressiveness to be the optimal variables for the reduced model, accounting for 45% of the total variation.
+ANOVA revealed that the reduced model was significant with an adjusted R^2^ of 0.0674 (*P* = 0.001).
+Assessment of the marginal effects showed that region, host, year, and aggressiveness significantly explained genetic variation (*P* = 0.001, *P* = 0.009 for region).
+It was revealed that there was multicollinearity when MCG was combined with any other variable, so we followed the same procedure, dropping MCG from the list of potential predictors. 
 From these results, we found that aggressiveness, year, region, and host were optimal, accounting for 17.6% of the total variation.
-ANOVA revealed significant effects with an adjusted R^2^ of 0.0325 (P = 0.001).
-While the marginal effect assessment showed that region, host, and year all significantly explained the variation at P = 0.001, and aggressiveness significantly explained the variation at P = 0.045.
+ANOVA revealed significant effects with an adjusted R^2^ of 0.0325 (*P* = 0.001).
+While the marginal effect assessment showed that region, host, and year all significantly explained the variation at *P* = 0.001, and aggressiveness significantly explained the variation at *P* = 0.045.
 Much of the variation appears to be driven by isolates from Mexico and 2005 (Fig. \@ref(rda-biplot)).
@@
 The length of the arrows are directly proportional to the strength of the correlation between explanatory and molecular variables. 
-Points represent the 318 clone-corrected samples in ordination space. 
+Open circles represent the 318 clone-corrected samples in ordination space. 
 ](../../results/figures/publication/Figure7Z.pdf){#rda-biplot width=61%}
@@
 Horizontal line represents median value.
-Points represent individual observations whose colors indicate if they came from white mold screening nurseries (wmn) or not (other).
+Circles represent individual observations where filled circles indicate isolates from white mold screening nurseries (wmn).
 ](../../results/figures/publication/aggressiveness.pdf){#aggressiveness-boxplot}
@@
 MCGs also showed a significant effect (P < 1.00e^-4^), with means that ranged from 6.0 (MCG 44) to 4.6 (MCG 49).
-We additionally found a significant effect for MLGs (P = 7.44e^-4^) for MLGs, with means that ranged from 6.0 (MLG 78) to 4.3 (MLG 140).
+We additionally found a significant effect for MLGs (*P* = 7.44e^-4^) for MLGs, with means that ranged from 6.0 (MLG 78) to 4.3 (MLG 140).
 
@@
 
-Our network-based approach to correlating MLGs with MCGs revealed a large and complex network (Fig. \@ref(biggraph), \@ref(fullgraph)). 
-Community analysis showed 51 communities, 15 of which consisted of a single MLG unconnected with any other community indicating that just 9.09% of the 165 MLGs are unable to cross with any other MLG in this data set.
-The three communities with the most members contained 8 of the ten most abundant MCGs. 
-Comparing these communities with Bruvo's genetic distance revealed and average distance of 0.451 among communities and an average distance of 0.437 within communities, showing no significant difference.
-When we assessed the number of potential heterothallic parings between different MLGs in our data set (two different MLGs that are in the same mycelial compatibility group), we found an average of 16.5 potential heterothallic parings per MLG.
-MLG 75, representing just four isolates, had 57 neighbors that shared the same mycelial compatibility group. 
+Our network-based approach to correlating MLGs with MCGs revealed a large and complex network (Fig. \@ref(biggraph)). 
+Community analysis showed 51 communities, 15 of which consisted of a single MLG unconnected with any other community indicating that just 9.09% of the 165 MLGs are unable to cross with any other MLG in this data set (Fig.
+\@ref(fullgraph)).
+The three communities with the most members contained eight of the 10 most abundant MCGs. 
+Comparing these communities with Bruvo's genetic distance revealed an average distance of 0.451 among communities and an average distance of 0.437 within communities, showing no significant difference.
+When we assessed the number of times two different MLGs that are in the same MCG, considering these as potential heterothallic pairings that could result in sexual recombination, we found an average of 16.5 potential heterothallic parings per MLG.
+Representing just four isolates, 'MLG 75' had 57 neighbors that shared the same mycelial compatibility group (Fig \@ref(biggraph), \@ref(fullgraph)). 
 Overall, there was no clear pattern to the association between MLG and MCGs.
@@
 Data show complex associations between Mycelial Compatibility Groups and Multilocus Genotypes.
-**A)** Barplot showing abundance of each Mycelial Compatibility Group (MCG) in descending order.
-Singletons (46) have been removed for clarity, leaving 41 MCGs.
+**A)** Barplot of Mycelial Compatibility Group (MCG) abundance in descending order.
+Singletons (46) were truncated, leaving 41 MCGs.
 White bars represent sample counts and grey bars represent counts of unique multilocus genotypes (MLG).
-The transparency associated with the grey bars represent the evenness of the distribution of the MLGs within a given MCG.
+The transparency of the bars represent the evenness of the distribution of the MLGs within a given MCG.
+A dashed box surrounds the 8 most common MCGs representing >51% of the data.
 **B)** Full graph-representation of the relationship between MCGs (open circles) and MLGs (filled circles).
 Details in fig. \@ref(fullgraph). 
-**C)** A subset of **B** representing the 8 most common MCGs and their associated MLGs.
-Node size reflect the number of samples represented by each node (circle).
-Edges (arrows) point from Multilocus Genotype to Mycelial Compatibility Group where the weight (thickness) of the edge represents the number of shared isolates.
+**C)** A subset of **B** representing the 8 most common MCGs and their associated MLGs (dashed box in **A**).
+Filled nodes represent MLGs and open nodes represent MCGs.
+Numbers inside nodes are the MLG/MCG label. 
+Node (circle) area scaled to the number of samples represented (range: 1--73).
+Edges (arrows) point from MLG to MCG where the weight (thickness) of the edge represents the number of shared isolates (range: 1--19).
 Edges extending from MLGs displayed to other MCGs are not shown.
@@
 
-## Genotypic Structure {-}
+## Structure of Shared Multilocus Genotypes {-}
 
@@
 
-We observed a total of 165 multilocus genotypes (MLG) (215 with 16 loci).
-The most abundant MLG represented 27 isolates (Table \@ref(tab:mlg-table)) across five regions (NE, MI, WA, CO, and ND). 
-The genotypes were relatively evenly distributed within regions with moderate to high diversity (Table \@ref(tab:iatab)). 
-Of the 165 MLGs, 76 (46%) were found across all geographic regions except WI (2), ID (1), and Mexico (18) (Fig. \@ref(community-graph)).
+We observed a total of 165 MLGs (215 with 16 loci).
+The most abundant MLG was represented by 27 isolates (Table \@ref(tab:mlg-table)) from five regions (NE, MI, WA, CO, and ND). 
+Within regions, the genotypes were relatively evenly distributed with moderate to high diversity (Table \@ref(tab:iatab)). 
+Of the 165 MLGs, 76 (46%) were found in at least two geographic regions except those found in WI (2), ID (1), and Mexico (18) (Fig. \@ref(community-graph)).
 Community analysis of the MLGs shared between populations revealed 4 communities with a modularity of 0.17 (Fig.
-\@ref(community-graph)): A coastal community (CA, OR, WA, and NY), a mid-west community (CO, NE, ND, MI), and an international community (MN, France, Australia).
+\@ref(community-graph)): A coastal community (CA, OR, WA, and NY), a midwest community (CO, NE, ND, MI), and an international community (MN, France, Australia).
 Analysis with 16 loci resulted in the removal of the NY node because it no longer shared a genotype with OR, but the same overall community structure was present with a modularity of 0.2 (Fig.
@@
 ![
-Network of populations (nodes/circles) and their shared multilocus haplotypes (MLG) (edges/lines) genotyped over 11 loci. 
+Network of populations (nodes/circles) and their shared multilocus genotypes (MLG) (edges/lines) genotyped over 11 loci. 
 Each node is labeled with **name (number of MLGs shared/number of MLGs total).** 
-The shade and area of the nodes are proportional to the number of unique MLGs within the node and the inner nodes are proportional to the number of private MLGs to the region. 
+The shade and area of the nodes are proportional to the number of unique MLGs within the node and the inner nodes are proportional to the number of private MLGs to the region (bottom legend). 
 Each edge represents a single MLG where its thickness represents the number of populations that share the MLG and the shade represents the value of $P_{sex}$, or the probability of encountering that MLG from two independent meiotic events.
-](../../results/figures/publication/Figure3Z.pdf){#community-graph width=100%}
+](../../results/figures/publication/mlg-11.pdf){#community-graph width=100%}
 
@@
 
-The analysis of molecular variance (AMOVA) over the hierarchy of Source and Region showed no significant variation between white mold screening nurseries and non-white mold screening nurseries after accounting for variation between regions (Table \@ref(tab:amova)).
+The analysis of molecular variance (AMOVA) over the hierarchy of Source and Region showed no significant variation between wmn and non-wmn after accounting for variation between regions (Table \@ref(tab:amova)).
 Because we observed that the sample sizes of wmn isolates and producer isolates were lopsided in several populations (Table \@ref(tab:isolate-table)), we ran the analysis again with regions that had at least ten isolates from both wmn and producer fields, showing similar results (Table \@ref(tab:amova)).
@@
 
-Table: (\#tab:amova) Analysis of Molecular Variance on Bruvo's genetic distance showing little differentiation between white mold screening nurseries and other sources.
-The hierarchy was constructed as Source/Region where source is defined as belonging to a white mold screening nursery or not.
+Table: (\#tab:amova) Analysis of Molecular Variance on Bruvo's genetic distance showing little differentiation between white mold screening nurseries (wmn) and other sources.
+The hierarchy was constructed as Source/Region where source is defined as belonging to a wmn or not.
 Bold $\Phi$ values indicate significant values (P < 1.00e^-4^).
@@
 \@ref(DAPC)b).
-The second discriminant axis separated the centroid for the Californian isolates.
+The second discriminant axis separated the centroid for the California isolates.
 The mean population assignment probabilities for all populations with n > 10 showed that only isolates from Mexico, CA, and France had > 50% probabilities of being reassigned to their source populations (Fig.
@@
 DAPC grouping by cultivar used the first 20 principal components, representing 89% of the total variance.
-The first two discriminant axes failed to separate any of the cultivars where the mean posterior assignment probabilities were 33.97% (G122), 35.95% (Beryl), and 30.08% (Bunsi).
+The first two discriminant axes failed to separate any of the cultivars where the mean posterior assignment probabilities were 34.0% (G122), 36.0% (Beryl), and 30.1% (Bunsi).
 
@@
 Discriminant Analysis of Principal Components on regions showing that Mexico is differentiated from other populations. 
-**(a)** Mean population assignment (horizontal axis) probability (vertical axis) for all populations with N > 10 (facets).
+**A)** Mean population assignment (horizontal axis) probability (vertical axis) for all populations with N > 10 (facets).
 Numbers next to source populations indicate population size.
 All values sum to one.
-**(b)** Scatter plot of first two components from DAPC.
+**B)** Scatter plot of first two components from DAPC.
 Points represent observed individuals connected to the population centroids with ellipses representing a 66% confidence interval for a normal distribution. 
@@
 In this study, we characterized the diversity of *Sclerotinia sclerotiorum* from dry bean fields across the United States.
-Our results suggest that, broadly, populations from white mold screening nurseries reflect the populations of the surrounding regions, indicating that resistance screening may be successful at a regional level.
+Our results suggest that, broadly, populations from white mold screening nurseries reflect the populations of the surrounding regions, indicating that resistance screening may be successful at a within regions.
 We found significant population differentiation by geographic region, mainly differentiated into three broad North American groups based on shared genotypes and posterior groupings, a Coastal Region, Midwestern Region, and Mexico.
-To date, with 366 isolates, this is the largest population genetic study of *S. sclerotiorum* assessing population structure within managed and unmanaged agricultural environments. 
-These findings indicate that the white mold screening nurseries can be effective at screening for potential resistant lines within growing regions.
+To date, with 366 isolates, this is the largest single population genetic study of *S. sclerotiorum* assessing population structure within managed and unmanaged agricultural environments. 
+These findings indicate that the wmns can be effective at screening for potential resistant lines within growing regions.
 
@@
 It has been shown that *S. sclerotiorum* in the midwestern United States does not have a particular preference for host [@aldrich-wolfe2015genetic]. 
-If the distribution of *S. sclerotiorum* is even across agricultural hosts in the US, then our sample may yet be representative of the genetic pool present in other crops and weedy species.
+If the distribution of *S. sclerotiorum* is even across agricultural hosts in the USA, then our sample may yet be representative of the genetic pool present in other crops and weedy species.
 
@@
 ![
-Graph showing complex associations between Mycelial Compatibility Groups (dotted nodes) and Multilocus Genotypes (full nodes). 
+Graph showing complex associations between Mycelial Compatibility Groups (MCG) (dotted nodes) and Multilocus Genotypes (MLG) (full nodes) where the number in each node represents the MLG/MCG assignment.
 Node size reflect the number of samples represented by each node (circle). 
-Edges (arrows) point from Mycelial Compatibility Group to Multilocus Genotype where the weight (thickness) of the edge represents the number of samples shared. 
+Edges (arrows) point from MLG to MCG where the weight (thickness) of the edge represents the number of samples shared. 
 Node color represents the community assignment based on the walktrap algorithm with a maximum of four steps [@pons2006computing]. 
@@
 ![
-Network of populations (nodes/circles) and their shared multilocus haplotypes (MLG) (edges/lines) genotyped over 16 loci. 
+Network of populations (nodes/circles) and their shared multilocus genotypes (MLG) (edges/lines) genotyped over 16 loci. 
 Each node is labeled with **name (number of MLGs shared/number of MLGs total).** 
-The shade and area of the nodes are proportional to the number of unique MLGs within the node and the inner nodes are proportional to the number of private MLGs to the region. 
+The shade and area of the nodes are proportional to the number of unique MLGs within the node and the inner nodes are proportional to the number of private MLGs to the region (bottom legend). 
 Each edge represents a single MLG where its thickness represents the number of populations that share the MLG and the shade represents the value of $P_{sex}$, or the probability of encountering that MLG from two independent meiotic events.
-](../../results/figures/publication/FigureS1Z.pdf){#community-graph16 width=100%}
+](../../results/figures/publication/mlg-16.pdf){#community-graph16 width=100%}
 

### 2017-08-15 doc/manuscript/manuscript.Rmd
Add McCoy as Author; update DNA isolation protocol

@@
     email: rjhala2@unl.edu
+    affiliation: 1
+  - name: Serena McCoy
+    email: ¯\_(ツ)_/¯
     affiliation: 1
@@
 
-Each isolate was regenerated by activating one or several sclerotia and transferring the leading edge of the mycelial mat according to the method described previously [@otto-hanson2011variation].
+After collection, the isolates were stored in petri plates lined with filter paper, then stored at $20^{\circ}$F or –$4^{\circ}$C. 
+Sclerotia were triple rinsed with (i) 50% Clorox bleach/50% double-distilled (dd)H2O solution for 3 min, (ii) ddH2O rinse for 3 min, and (iii) final ddH2O rinse for 3 min, then plated on water agar plates (16g of Bacto agar per liter of ddH2O), with four to five sclerotia of each isolate separated on each plate and stored on the bench top at room temperature for 5 to 6 days An 8-mm plug from a 5- or 6-day-old culture was transferred from the advancing margin of the mycelia onto a plate of Difco potato dextrose agar (PDA at 39 g/liter of ddH2O) [@otto-hanson2011variation].
 In combination with the 156 isolates described previously, we collected 210 isolates for a total of 366 isolates [@otto-hanson2011variation].
@@
 
+ <!-- Preceeding the steps in that, we grew the isolate(s) on PDA, and transferred plugs to Potato Dextrose Broth.
+We grew the culture in broth until there was significant mycelial growth, but before the mycelial mat became 'solid' (about 4-5 days).
+We collected the mycelia using a buchner funnel, removed the agar plugs, lyophilized the mycelia, then pulverized it by hand in whirl-pak bags.
+The lyophilized mycelia was then used in the procedure as printed.
+Lyophilzed mycelia was stored in microcentrifuge tubes in the -20 freezer until needed for extractions.
+-->
+
+Prior to DNA extraction, isolates were grown on PDA and plugs were subsequently transferred to Potato Dextrose Broth (PDB) where they were grown until there was significant mycelial growth, but before the mycelial mat became solidified (4--5 days).
+Mycelial was collected in a buchner funnel, agar plugs removed, lyophilizedm and pulverized manually in whirl-pak bags. 
+Lyophilized mycelia was then stored in microcentrifuge tubes at -$20^{\circ}$C until needed for DNA extraction.
+
 DNA was purified from approximately 100 mg mycelium collected from an actively growing culture of each isolate using a Qiagen DNeasy Plant Mini Kit (Qiagen Inc., Valencia, CA) according to the manufacturer's instructions.
 
-> SEE: More details on how tissue was disrupted may be needed here.
+<!-- Fungal DNA mini-prep Updated from Jeff Rollins’ protocol           Rev. Nov 2012
+
+Sclerotinia spp. Phenol/Chloroform extraction-
+
+1) Spin 2.0 ml Phase-lock gel microcentrifuge tubes @14,000 rpm for 30 sec.- 2 for each sample to be extracted
+
+2) Place 600 ul LETS buffer in 2.0 ml microcentrifuge (round-bottomed)  tube
+
+3) Add ~25 mg of lyophilized mycelia; mix by inverting (shaking, flicking) tube.
+
+4) Add an equal volume (600ul) of buffer-saturated Phenol.  Invert to mix.
+
+5) Add 300ul chloroform to tubes- invert to mix.
+
+6) Pour suspension into Phase-lock  gel tubes.
+
+7) Spin Phase-lock tubes @12,000 RPM for 8-10 min. to separate phases.
+
+8) Add another volume (300ul per tube) of chloroform, invert gently to mix.  
+    Spin again @ 12,000 RPM for 8 min.
+
+9) Pour off top (aqueous) layer into micro-centrifuge tube.
+
+10) Add an equal volume 2-propanol to volume in each tube.  Mix by gently inverting. Nucleic acids will be visible as transluscent strands in tubes.  
+(Optional- Place in freezer (-20OC) overnight.)
+
+**Turn on heat block to ‘low’ setting**
+
+11) Spin tubes @ 12,000 RPM for 14 min.  White or loose semi-translucent pellets should be visible in tube.
+
+12) Rinse pellet with ice-cold 70% EtOH, pouring off alcohol carefully- watch out for ‘loose’ pellets (re-spin if necessary).  Invert on paper towel for ~ 20 min to dry.
+
+13) Resuspend pellet in 450 ul TE, then add 10 ul RNAase to each tube- invert or flick to dissolve pellet.  Incubate tubes on heat block at 37OC for 30 min.
+
+14) Add 450 ul Phenol to each tube.  Invert tubes to mix.
+
+16) Add 450 ul chloroform to each tube, invert to mix.
+
+15) Transfer tube contents to Phase-lock  tubes, then spin @ 14,000 RPM for 10 min. to separate phases
+
+16) Add 300 ul chloroform to each tube, invert to mix.
+
+17) Spin @ 14,000 RPM for 8-10 min. to again separate phases.
+
+18) Using wide-bore pipette tip, transfer aqueous phase to 1.5 micro-centrifuge tube- recover up to 450ul (not more) to keep total volume within capacity of micro-centrifuge tube.
+
+19) Add 0.1 volume (45ul) 3M NaOAc to each tube.  Add two volumes ice-cold absolute EtOH (fill tube).  Incubate on ice for 30 min.- translucent material should be visible in tubes.
+(Optional- Place in freezer (-20OC) overnight.)
+
+20) Centrifuge at 14,000 RPM for 10 min.
+
+21) Pour off EtOH, being careful not to dislodge pellet.
+Rinse in cold 70% EtOH, then dry pellet on paper towels for at least 15 min.
+Pellets may appear as white “pellets” in bottom of tube, or may appear as translucent smears on side of tube.
+
+22) Resuspend pellets in 200ul TE.  (Place in 4OC refrig for ~24 hrs to allow pellet to rehydrate). -->
 
@@
 It can be seen in Fig.
-\@ref(biggraph)A that the most abundant MCG contains several MLGs, but the diversity of those MLGs are low as indicated by the evenness (transparency), which indicates that there is one dominant MLG (MLG 25).
+\@ref(biggraph)A that the most abundant MCG contains several MLGs, but the diversity of those MLGs are low as indicated by the evenness (transparency), which indicates that there is one dominant MLG ('MLG 25').
 What is not shown in Fig. \@ref(biggraph)A is the MLGs that are shared between MCGs. 
 This is illustrated in both table \@ref(tab:mlg-table) and Fig. \@ref(biggraph)B,C. 
-It could be argued, however that MLG 25, with its high value of $P_{sex}$ represents different true MLGs across the five MCGs it occupies, but this doesn't account for the overall structure of Fig.
-\@ref(fullgraph) where, for example, MLG 75 ($P_{sex}$ = 1.81e^-4^) is compatible with 58 other genotypes through three MCG when the population structure of *S.
+It could be argued, however that 'MLG 25', with its high value of $P_{sex}$ represents different true MLGs across the five MCGs it occupies, but this doesn't account for the overall structure of Fig.
+\@ref(fullgraph) where, for example, 'MLG 75' ($P_{sex}$ = 1.81e^-4^) is compatible with 58 other genotypes through three MCG when the population structure of *S.
 sclerotiorum* is known to be clonal.
@@
 The most abundant genotype was shared across four populations, but its high value of $P_{sex}$ meant that it did not contribute significantly to the overall structure. 
-The graph walking algorithm was able to divide the network into three groups, but had a modularity of 0.17, which indicates that the groups are weakly differentiated.
+The graph walking algorithm was able to divide the network into three groups, but had a modularity of 0.17, which indicates that the groups are only weakly differentiated.
 This indicates that multilocus genotypes are widespread across the United States, giving more evidence for a clonal population structure.

### 2017-08-15 doc/manuscript/manuscript.Rmd
Begin protocol translation

@@
 
-DNA was purified from approximately 100 mg mycelium collected from an actively growing culture of each isolate using a Qiagen DNeasy Plant Mini Kit (Qiagen Inc., Valencia, CA) according to the manufacturer's instructions.
+DNA was purified using a phenol-chloroform extraction method.
+First, approximately 25 mg of lyophilized mycelia was suspended in 600$\mu$l of LETS Buffer (0.1 M LiCl, 0.01 M Na2EDTA, 0.01 M Tris-Cl (pH 7.4), 0.2% SDS) in a 2.0ml microcentrifuge tube.
+An equal amount of buffer-saturated phenol (600$\mu$l) was added and mixed by inversion, followed by 300$\mu$l chloroform, mixed by inversion. 
+The suspension was then transferred into a 2.0ml Phase-lock gel tube (VWR, Randor, PA) and centrifuged for 8--10 minutes at 12,000 RPM.
+Another aloquat of 300$\mu$l was added and centrifuged for another 8--10 minutes at 12,000 RPM and the aqueous phase was subsequently transferred into a microcentrifuge tube with an equal volume of 2-propanol added and mixed via gentle inversion.
+
+<!-- using a Qiagen DNeasy Plant Mini Kit (Qiagen Inc., Valencia, CA) according to the manufacturer's instructions. -->
 

### 2017-08-16 doc/RMD/wmn-differentiation.Rmd
Correct mistake with AMOVA table.
Due to an oversight back in 2014/2015, I had
accidentally flipped the Phi values for the AMOVA
table, rendering the results silly.

This still needs to be fixed in text.

@@
                          Pval    = rev(amt$pvalue), 
-                         Phi     = am$statphi$Phi[-tot]))
+                         Phi     = rev(am$statphi$Phi[-tot])))
   res <- as.matrix(res)
@@
   add_column(Hierarchy = c("Between Source", "Between Region within Source", "Within Region"), .before = 1) %>%
-  readr::write_csv(path = file.path("results", "tables", "AMOVA-region.csv"), col_names = TRUE) %>%
+  readr::write_csv(path = file.path(PROJHOME, "results", "tables", "AMOVA-region.csv"), col_names = TRUE) %>%
   rename(ps = `Phi statistic`) %>%
@@
   huxtable::as_huxtable(add_colnames = TRUE) %>% 
-  huxtable::set_col_width(c(1.1, 0.6, 0.9, 0.9, 1.4)) %>% 
+  huxtable::set_col_width(c(1.1, 0.6, 0.8, 0.8, 1.1)) %>% 
   huxtable::set_align(huxtable::everywhere, 2:5, "center") %>% 
-  huxtable::print_md(max_width = 100)
+  huxtable::print_md(max_width = 90)
 ```

### 2017-08-16 doc/manuscript/manuscript.Rmd
I have added a by-year component.
I realized that I never analyzed the data across
years, which I should have done from the beginning.

This brought us a new AMOVA table that makes more
sense and some new questions regarding WA in 2008.

@@
 
-After collection, the isolates were stored in petri plates lined with filter paper, then stored at $20^{\circ}$F or –$4^{\circ}$C. 
+After collection, the isolates were stored in petri plates lined with filter paper, then stored at $20^{\circ}$F or -$4^{\circ}$C. 
 Sclerotia were triple rinsed with (i) 50% Clorox bleach/50% double-distilled (dd)H2O solution for 3 min, (ii) ddH2O rinse for 3 min, and (iii) final ddH2O rinse for 3 min, then plated on water agar plates (16g of Bacto agar per liter of ddH2O), with four to five sclerotia of each isolate separated on each plate and stored on the bench top at room temperature for 5 to 6 days An 8-mm plug from a 5- or 6-day-old culture was transferred from the advancing margin of the mycelia onto a plate of Difco potato dextrose agar (PDA at 39 g/liter of ddH2O) [@otto-hanson2011variation].
@@
 suppressPackageStartupMessages(library("tidyverse"))
-read_csv("../../results/tables/AMOVA-region.csv", 
+read_csv("../../results/tables/AMOVA-year.csv", 
          na = "NA", # spaces read in as spaces
@@
   rename(ps = `Phi statistic`) %>%
+  # mutate_if(is.numeric, format, digits = 3) %>%
+  mutate(ps = signif(as.numeric(ps), digits = 3)) %>%
   mutate(ps = gsub("0\\.00(\\d{1})(\\d{2})", "\\1.\\2e^-3^", ps)) %>%
-  mutate(ps = case_when(Hierarchy == "Between Source" ~ ps, TRUE ~ paste0("**", ps, "**"))) %>%
-  rename(`$\\Phi statistic$` = ps) %>%
+  mutate(ps = case_when(P > 0.05 ~ ps, TRUE ~ paste0("**", ps, "**"))) %>%
+  rename(`$\\Phi$ statistic` = ps) %>%
   rename(`% variation` = `Percent variation`) %>%
   rename(S.S. = `Sum of Squares`) %>%
-  select(-P) %>%
+  rename(`*P*` = P) %>%
+  select(c(1:4, 6, 5)) %>%
   huxtable::as_huxtable(add_colnames = TRUE) %>% 
-  huxtable::set_col_width(c(1.1, 0.6, 0.9, 0.9, 1.4)) %>% 
-  huxtable::set_align(huxtable::everywhere, 2:5, "center") %>% 
-  huxtable::print_md(max_width = 100)
+  huxtable::set_col_width(c(0.8, 0.15, 0.15, 0.35, 0.5, 0.15)) %>% 
+  huxtable::set_align(huxtable::everywhere, 2:6, "right") %>% 
+  huxtable::set_number_format(huxtable::everywhere, 2, 0) %>%
+  huxtable::set_number_format(huxtable::everywhere, 6, 3) %>%
+  huxtable::set_number_format(huxtable::everywhere, 5, 4) %>%
+  huxtable::print_md(max_width = 80)
 ```
@@
 The hierarchy was constructed as Source/Region where source is defined as belonging to a wmn or not.
-Bold $\Phi$ values indicate significant values (P < 1.00e^-4^).
-Numbers in parentheses represent AMOVA run with the regions NE, WA, and MI.
+Bold $\Phi$ values indicate significant values (*P* < 0.05).
 S.S.

### 2017-08-16 doc/RMD/by-year.Rmd
I have added a by-year component.
I realized that I never analyzed the data across
years, which I should have done from the beginning.

This brought us a new AMOVA table that makes more
sense and some new questions regarding WA in 2008.

---
title: "Differentiation By Year"
output: 
  html_notebook:
    toc: true
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = PROJHOME)
```

# Purpose

Because we have collections of isolates from the same WMN across different years,
we can investigate how populations change over the years. I'm attempting to be
careful with this analysis because, much like the white mold screening nursery
analysis, there are small sample sizes, which could bias the results.

# Data

```{r packages}
library("poppr")
library("tidyverse")
library("cowplot")
make_amova_table <- function(am, amt, samples = "Region"){
  tot <- nrow(am$results)
  res <- data.frame(list(am$results[-tot, c("Df", "Sum Sq")], 
                         Percent = am$componentsofcovariance[-tot, 2],
                         Pval    = rev(amt$pvalue), 
                         Phi     = rev(am$statphi$Phi[-tot])))
  res <- as.matrix(res)
  colnames(res) <- c("d.f.", "Sum of Squares", "Percent variation", "P", 
                     "Phi statistic")
  names(dimnames(res)) <- c("levels", "statistic")
  rownames(res) <- gsub("samples", samples, rownames(res))
  return(res)
}

make_amova_printable <- function(amtab, amtabcc){
  am_array <- array(dim      = c(dim(amtab), 2),
                    dimnames = c(dimnames(amtab), 
                                 list(c("full", "clone-corrected"))))
  am_array[, , 1] <- amtab
  am_array[, , 2] <- amtabcc
  tabfun <- function(x){
    x <- paste0(paste0(signif(x, 3), collapse = " ("), ")")
    return(x)
  }
  res <- apply(am_array, c(1, 2), tabfun)
  return(res)
}
```

```{r data}
load("data/sclerotinia_16_loci.rda")
dat11cc <- dat11 %>%
  clonecorrect(~Region/Source/Host/Year) %>% 
  setPop(~Region/Year)

addStrata(dat11cc) <- strata(dat11cc) %>% 
  mutate(SourceType = forcats::fct_inorder(ifelse(Source == "wmn", "wmn", "other"))) %>%
  select(SourceType)
wmncc   <- dat11cc %>% 
  setPop(~Source) %>% 
  popsub("wmn") %>% 
  setPop(~Region/Year)
table(strata(dat11cc) %>% select(Region, Year))
table(strata(wmncc) %>% select(Region, Year))
```

# Analysis of Molecular Variance

First, we should see what AMOVA tells us about population structure

```{r wmn_amova, cache = TRUE}
bdwmn <- bruvo.dist(wmncc, replen = other(wmncc)$REPLEN)
(wmn.amova <- poppr.amova(wmncc, ~Region/Year, dist = bdwmn))
(wmn.amova.test <- randtest(wmn.amova, nrepet = 999))
make_amova_table(wmn.amova, wmn.amova.test, samples = "Year")
```
```{r, full_amova, cache = TRUE}
bdfull <- bruvo.dist(dat11cc, replen = other(dat11cc)$REPLEN)
(full.amova.2 <- poppr.amova(dat11cc, ~Region/Year, dist = bdfull))
(full.amova.2.test <- randtest(full.amova.2, nrepet = 999))
(full.amova.3 <- poppr.amova(dat11cc, ~Region/SourceType/Year, dist = bdfull))
(full.amova.3.test <- randtest(full.amova.3, nrepet = 999))
make_amova_table(full.amova.2, full.amova.2.test, samples = "Year")
make_amova_table(full.amova.3, full.amova.3.test, samples = "Year")
```

```{r AMOVA_table, comment = NA}
make_amova_table(full.amova.3, full.amova.3.test, samples = "Year") %>%  
  as_tibble() %>%
  add_column(Hierarchy = c("Between Region", "Between Source within Region", "Between Year within Source", "Within Year"), .before = 1) %>%
  readr::write_csv(path = file.path(PROJHOME, "results", "tables", "AMOVA-year.csv"), col_names = TRUE) %>%
  rename(ps = `Phi statistic`) %>%
  mutate_if(is.numeric, format, digits = 3) %>%
  # mutate(ps = signif(ps, digits = 3)) %>%
  mutate(ps = gsub("0\\.00(\\d{1})(\\d{2})", "\\1.\\2e^-3^", ps)) %>%
  # mutate(ps = case_when(P > 0.05 ~ ps, TRUE ~ paste0("**", ps, "**"))) %>%
  rename(`$\\Phi$ statistic` = ps) %>%
  rename(`% variation` = `Percent variation`) %>%
  rename(S.S. = `Sum of Squares`) %>%
  rename(`*P*` = P) %>%
  select(c(1:4, 6, 5)) %>%
  huxtable::as_huxtable(add_colnames = TRUE) %>% 
  huxtable::set_col_width(c(0.8, 0.15, 0.15, 0.35, 0.5, 0.15)) %>% 
  huxtable::set_align(huxtable::everywhere, 2:6, "right") %>% 
  huxtable::set_number_format(huxtable::everywhere, 2, 0) %>%
  huxtable::set_number_format(huxtable::everywhere, 6, 3) %>%
  huxtable::set_number_format(huxtable::everywhere, 5, 4) %>%
  huxtable::print_md(max_width = 80)
```

# DAPC

Now that we have a result for the AMOVA showing a signficant difference in years, we can use DAPC to see where this difference is.

```{r dapc}
setPop(dat11cc) <- ~Region/Year
regyear.dapc <- dapc(dat11cc, n.pca = 20, n.da = 20)
```

```{r dapc_plot, fig.width = 7, fig.height = 7}
RYD <- bind_cols(Population = regyear.dapc$grp, strata(dat11cc), as.data.frame(regyear.dapc$ind.coord)) %>%
  as_tibble()
RYD_pop <- RYD %>% 
  group_by(Population) %>% 
  summarize_if(is.numeric, mean, na.rm = TRUE) %>%
  rename_all(function(x) gsub("LD", "mean", x))
RYD <- full_join(RYD, RYD_pop, by = "Population")
RYD_PLOT <- ggplot(RYD, aes(x = LD1, y = LD2, color = Year)) + 
  geom_text(aes(label = Year), alpha = 0.5) +
  geom_segment(aes(x = mean1, y = mean2, xend = LD1, yend = LD2), alpha = 0.5) +
  stat_ellipse(type = "norm", level = 0.66, alpha = 0.75) + 
  facet_wrap(~Region) +
  theme_bw() +  
  theme(aspect.ratio = 0.8) +
  theme(legend.position = "bottom") +
  theme(axis.text = element_blank()) + 
  theme(axis.title = element_blank()) + 
  theme(axis.ticks = element_blank()) + 
  viridis::scale_color_viridis(discrete = TRUE, direction = -1) +
  viridis::scale_fill_viridis(discrete = TRUE, direction = -1) +
  scale_y_continuous(breaks = 0) + 
  scale_x_continuous(breaks = 0) + 
  theme(panel.background = element_rect(fill = "grey95")) +
  theme(panel.grid.major = element_line(color = "black"))
RYD_PLOT
par(mfrow = c(2, 1))
loadingplot(regyear.dapc$var.contr, axis = 1)
loadingplot(regyear.dapc$var.contr, axis = 2)
par(mfrow = c(1, 1))
```

The loading plot shows us the variables that are important for the first axis of
separation, mainly:

----------------
 Locus    Allele
-------- -------
7-2(F)       174
8-3(H)       256
9-2(F)       372
55-4(F)      157
110-4(H)     370
----------------

There's an important thing going on in Washington where it appears that the
2008 population is separated from the main population.

If we look at the tables above, we can see that there are only 12 samples from 
2008, all of whicha are in white mold screening nurseries. We also notice that
the coördinates for these isolates in the first two discriminant components are
overlapping with CA isolates. I suspect that there may be duplicate genotypes:

```{r crosspop, results = "asis"}
mlg.crosspop(dat11cc, ~Region/Year, quiet = TRUE, df = TRUE) %>%
  inner_join(filter(., Population == "WA_2008") %>% select(MLG), by = "MLG") %>%
  spread(Population, Count, fill = 0) %>%
  knitr::kable()
```

Well! It appears that ALL of the isolates in WA in 2008 have buddies in 2005 in
CA and France. We can take a look at how the allele frequencies are responding.

```{r fig.width = 10, fig.height = 12}
dat11ccloc <- dat11cc[loc = c("7-2(F)", "8-3(H)", "9-2(F)", "55-4(F)", "110-4(H)")]
loc <- map(seploc(dat11ccloc), . %>% 
             tab() %>% 
             as.data.frame() %>% 
             bind_cols(strata(dat11cc)) %>% 
             select(-MCG) %>%
             group_by(Region, Year) %>% 
             mutate(N = n()) %>%
             filter(N > 5) %>%
             summarize_if(is.numeric, mean, na.rm = TRUE) %>%
             ungroup() %>%
             mutate(Year = as.integer(as.character(Year))) %>%
             gather(allele, frequency, -Region, -Year, -N) %>%
             mutate(allele = as.integer(gsub("^.+?\\.([[:alnum:]]+)$", "\\1", allele)))
             # ,
             # .id = "Locus"
           )

ggloc <- function(dat, REG, years = 2003:2008){
  map(dat, . %>% 
    filter(Region == REG) %>%
    {ggplot(., aes(x = Year, y = frequency, color = allele, group = allele)) +
    geom_line() + 
    geom_point(aes(size = N), show.legend = FALSE) +
    # geom_label(aes(label = allele), size = 1, color = "black") +
    scale_x_continuous(breaks = years) +
    scale_y_continuous(limits = c(0, 1)) +
    theme_bw(base_size = 16, base_family = "Helvetica") +
    theme(legend.position = "top") +
    theme(aspect.ratio = 1/2) +
    theme(axis.text = element_text(color = "black")) +
    theme(axis.ticks.y = element_blank()) +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
    # theme(axis.text.x.top = element_text(angle = 90, vjust = 0.5, hjust = 0)) +
    theme(panel.grid.minor.x = element_line(linetype = 0, color = "grey50")) +
    theme(panel.grid.major.x = element_line(linetype = 0, color = "grey50")) +
    theme(panel.grid.major = element_line(colour = "grey20")) +
    theme(panel.grid.minor = element_line(linetype = 3, colour = "grey50")) +
    # theme(panel.spacing.y = unit(0, "line")) +
    theme(panel.background = element_rect(color = NA, fill = "grey96")) +
    theme(strip.text = element_text(face = "bold", hjust = 0.05)) +
    theme(panel.border = element_blank()) +
    scale_size(range = c(3, 6)) +
    scale_color_viridis(option = "C", end = 0.9, 
                        guide = "legend",
                        breaks = sort(unique(.$allele)))
  })
}
ggl <- ggloc(loc, "WA")

cowplot::plot_grid(plotlist = ggl, labels = names(ggl), ncol = 2)
```

We can see that the allele frequencies seem to have a drastic change in the 
allele frequencies from 2007 to 2008. What happens if we look at a stable region
like NE?

```{r NEplot, fig.width = 10, fig.height = 10}
gglNE <- ggloc(loc, "NE", 2003:2012)
cowplot::plot_grid(plotlist = gglNE, labels = names(gglNE), ncol = 2)
```
```{r MIplot, fig.width = 10, fig.height = 10}
gglMI <- ggloc(loc, "MI", 2003:2012)
cowplot::plot_grid(plotlist = gglMI, labels = names(gglMI), ncol = 2)
```

It appears that MI is fairly stable, NE has some weird things going on with 2007
and 2008... I'm wondering if this may be a labeling issue.

<details>
<summary>Session Information</summary>

```{r, echo = FALSE}
options(width = 100)
devtools::session_info()
```

</details>

### 2017-08-17 doc/manuscript/manuscript.Rmd
update methods.

@@
 
- <!-- Preceeding the steps in that, we grew the isolate(s) on PDA, and transferred plugs to Potato Dextrose Broth.
-We grew the culture in broth until there was significant mycelial growth, but before the mycelial mat became 'solid' (about 4-5 days).
-We collected the mycelia using a buchner funnel, removed the agar plugs, lyophilized the mycelia, then pulverized it by hand in whirl-pak bags.
-The lyophilized mycelia was then used in the procedure as printed.
-Lyophilzed mycelia was stored in microcentrifuge tubes in the -20 freezer until needed for extractions.
--->
-
 Prior to DNA extraction, isolates were grown on PDA and plugs were subsequently transferred to Potato Dextrose Broth (PDB) where they were grown until there was significant mycelial growth, but before the mycelial mat became solidified (4--5 days).
-Mycelial was collected in a buchner funnel, agar plugs removed, lyophilizedm and pulverized manually in whirl-pak bags. 
+Mycelial was collected in a buchner funnel, agar plugs removed, lyophilized and pulverized manually in whirl-pak bags. 
 Lyophilized mycelia was then stored in microcentrifuge tubes at -$20^{\circ}$C until needed for DNA extraction.
-
-DNA was purified using a phenol-chloroform extraction method.
-First, approximately 25 mg of lyophilized mycelia was suspended in 600$\mu$l of LETS Buffer (0.1 M LiCl, 0.01 M Na2EDTA, 0.01 M Tris-Cl (pH 7.4), 0.2% SDS) in a 2.0ml microcentrifuge tube.
-An equal amount of buffer-saturated phenol (600$\mu$l) was added and mixed by inversion, followed by 300$\mu$l chloroform, mixed by inversion. 
-The suspension was then transferred into a 2.0ml Phase-lock gel tube (VWR, Randor, PA) and centrifuged for 8--10 minutes at 12,000 RPM.
-Another aloquat of 300$\mu$l was added and centrifuged for another 8--10 minutes at 12,000 RPM and the aqueous phase was subsequently transferred into a microcentrifuge tube with an equal volume of 2-propanol added and mixed via gentle inversion.
-
-<!-- using a Qiagen DNeasy Plant Mini Kit (Qiagen Inc., Valencia, CA) according to the manufacturer's instructions. -->
-
-<!-- Fungal DNA mini-prep Updated from Jeff Rollins’ protocol           Rev. Nov 2012
-
-Sclerotinia spp. Phenol/Chloroform extraction-
-
-1) Spin 2.0 ml Phase-lock gel microcentrifuge tubes @14,000 rpm for 30 sec.- 2 for each sample to be extracted
-
-2) Place 600 ul LETS buffer in 2.0 ml microcentrifuge (round-bottomed)  tube
-
-3) Add ~25 mg of lyophilized mycelia; mix by inverting (shaking, flicking) tube.
-
-4) Add an equal volume (600ul) of buffer-saturated Phenol.  Invert to mix.
-
-5) Add 300ul chloroform to tubes- invert to mix.
-
-6) Pour suspension into Phase-lock  gel tubes.
-
-7) Spin Phase-lock tubes @12,000 RPM for 8-10 min. to separate phases.
-
-8) Add another volume (300ul per tube) of chloroform, invert gently to mix.  
-    Spin again @ 12,000 RPM for 8 min.
-
-9) Pour off top (aqueous) layer into micro-centrifuge tube.
-
-10) Add an equal volume 2-propanol to volume in each tube.  Mix by gently inverting. Nucleic acids will be visible as transluscent strands in tubes.  
-(Optional- Place in freezer (-20OC) overnight.)
-
-**Turn on heat block to ‘low’ setting**
-
-11) Spin tubes @ 12,000 RPM for 14 min.  White or loose semi-translucent pellets should be visible in tube.
-
-12) Rinse pellet with ice-cold 70% EtOH, pouring off alcohol carefully- watch out for ‘loose’ pellets (re-spin if necessary).  Invert on paper towel for ~ 20 min to dry.
-
-13) Resuspend pellet in 450 ul TE, then add 10 ul RNAase to each tube- invert or flick to dissolve pellet.  Incubate tubes on heat block at 37OC for 30 min.
-
-14) Add 450 ul Phenol to each tube.  Invert tubes to mix.
-
-16) Add 450 ul chloroform to each tube, invert to mix.
-
-15) Transfer tube contents to Phase-lock  tubes, then spin @ 14,000 RPM for 10 min. to separate phases
-
-16) Add 300 ul chloroform to each tube, invert to mix.
-
-17) Spin @ 14,000 RPM for 8-10 min. to again separate phases.
-
-18) Using wide-bore pipette tip, transfer aqueous phase to 1.5 micro-centrifuge tube- recover up to 450ul (not more) to keep total volume within capacity of micro-centrifuge tube.
-
-19) Add 0.1 volume (45ul) 3M NaOAc to each tube.  Add two volumes ice-cold absolute EtOH (fill tube).  Incubate on ice for 30 min.- translucent material should be visible in tubes.
-(Optional- Place in freezer (-20OC) overnight.)
-
-20) Centrifuge at 14,000 RPM for 10 min.
-
-21) Pour off EtOH, being careful not to dislodge pellet.
-Rinse in cold 70% EtOH, then dry pellet on paper towels for at least 15 min.
-Pellets may appear as white “pellets” in bottom of tube, or may appear as translucent smears on side of tube.
-
-22) Resuspend pellets in 200ul TE.  (Place in 4OC refrig for ~24 hrs to allow pellet to rehydrate). -->
+DNA from 25mg of pulverized mycelia was purified using a phenol-chloroform extraction method followed by alcohol precipitation, suspending the DNA in 200$\mu$l TE [@sambrook1989molecular].
+Suspended DNA was stored at $4^{\circ}$C until genotyping.
 

### 2017-08-17 doc/RMD/by-year.Rmd
functionalize ggscatter plot

@@
 
-```{r dapc_plot, fig.width = 7, fig.height = 7}
-RYD <- bind_cols(Population = regyear.dapc$grp, strata(dat11cc), as.data.frame(regyear.dapc$ind.coord)) %>%
+```{r dapc_plot, fig.width = 7, fig.height = 7, warning = FALSE, message = FALSE}
+# A ggplot2 version of scatter.dapc
+#
+# @param DAPC an object of class "dapc" derived from [adegenet::dapc]
+# @param STRATA a data frame defining the population strata (see [adegenet::strata])
+# @param color the variable in `STRATA` defining the color palette for the plot
+# @param filter a "quosure" containing a filtering method to be passed to [dplyr::filter]
+#
+# @return
+# @export
+#
+# @examples
+# library("adegenet")
+# data(microbov)
+# strata(microbov) <- as.data.frame(other(microbov))
+# setPop(microbov) <- ~spe/breed
+# mscat <- ggscatter(dapc(microbov, n.pca = 20, n.da = 40), strata(microbov),
+#                    color = "breed",
+#                    filter = quo(breed %in% c("Borgou", "Zebu", "Montbeliard", "Salers")))
+# mscat + facet_wrap(~coun)
+# mscat + facet_wrap(~spe)
+# mscat + facet_wrap(~breed)
+ggscatter <- function(DAPC, STRATA, color = "Year", filter = NULL){
+  RYD <- bind_cols(Population = DAPC$grp, STRATA, as.data.frame(DAPC$ind.coord)) %>%
     as_tibble()
@@
   RYD <- full_join(RYD, RYD_pop, by = "Population")
-RYD_PLOT <- ggplot(RYD, aes(x = LD1, y = LD2, color = Year)) + 
-  geom_text(aes(label = Year), alpha = 0.5) +
+  yminor <- pretty(RYD$LD2)
+  xminor <- pretty(RYD$LD1)
+  RYD <- if (!is.null(filter)) filter(RYD, !!filter) else RYD
+  RYD_PLOT <- ggplot(RYD, aes_string(x = "LD1", y = "LD2", color = color)) + 
+    geom_text(aes_string(label = color), alpha = 0.75) +
     geom_segment(aes(x = mean1, y = mean2, xend = LD1, yend = LD2), alpha = 0.5) +
     stat_ellipse(type = "norm", level = 0.66, alpha = 0.75) + 
-  facet_wrap(~Region) +
-  theme_bw() +  
+    theme_bw(base_size = 16, base_family = "Helvetica") +  
     theme(aspect.ratio = 0.8) +
@@
     viridis::scale_fill_viridis(discrete = TRUE, direction = -1) +
-  scale_y_continuous(breaks = 0) + 
-  scale_x_continuous(breaks = 0) + 
+    scale_y_continuous(breaks = 0, minor_breaks = yminor) + 
+    scale_x_continuous(breaks = 0, minor_breaks = xminor) + 
     theme(panel.background = element_rect(fill = "grey95")) +
-  theme(panel.grid.major = element_line(color = "black"))
+    theme(panel.grid.major = element_line(color = "black")) +
+    theme(panel.grid.minor = element_line(color = "white")) 
   RYD_PLOT
+}
+
 par(mfrow = c(2, 1))
@@
 par(mfrow = c(1, 1))
+
+ggscatter(regyear.dapc, strata(dat11cc)) +
+  facet_wrap(~Region) +
+  theme(legend.position = c(0.75, 0.1)) +
+  guides(color = guide_legend(nrow = 4)) +
+  theme(legend.direction = "horizontal")
 ```
@@
 2008 population is separated from the main population.
+
+```{r dapc_plot_newamica, fig.width = 5, fig.height = 4, warning = FALSE, message = FALSE}
+ggscatter(regyear.dapc, strata(dat11cc), filter = quo(Region %in% c("WA", "CA", "NE", "MI"))) +
+  facet_wrap(~Region, nrow = 2) +
+  theme(legend.position = "right") +
+  theme(legend.key = element_rect(fill = "grey95"))
+```
 
@@
   map(dat, . %>% 
-    filter(Region == REG) %>%
-    {ggplot(., aes(x = Year, y = frequency, color = allele, group = allele)) +
-    geom_line() + 
-    geom_point(aes(size = N), show.legend = FALSE) +
-    # geom_label(aes(label = allele), size = 1, color = "black") +
+    filter(Region == REG) %>% {
+      ggplot(., aes(x = Year, y = frequency, color = allele, group = allele)) +
+        geom_text(aes(label = paste0("(", N, ")")), y = 0.99, color = "grey20", vjust = 1) +
+        geom_line(alpha = 0.75) +
+        geom_point(aes(size = N), show.legend = FALSE, alpha = 0.75) +
         scale_x_continuous(breaks = years) +
@@
         # theme(panel.spacing.y = unit(0, "line")) +
-    theme(panel.background = element_rect(color = NA, fill = "grey96")) +
+        theme(panel.background = element_rect(color = NA, fill = "grey98")) +
         theme(strip.text = element_text(face = "bold", hjust = 0.05)) +
         theme(panel.border = element_blank()) +
-    scale_size(range = c(3, 6)) +
-    scale_color_viridis(option = "C", end = 0.9, 
+        scale_size(range = c(3, 6), limits = c(6, 25)) +
+        viridis::scale_color_viridis(option = "C", end = 0.9, 
                                      guide = "legend",
@@
 }
-ggl <- ggloc(loc, "WA")
+ggl <- suppressMessages(ggloc(loc, "WA"))
 
@@
 ```{r NEplot, fig.width = 10, fig.height = 10}
-gglNE <- ggloc(loc, "NE", 2003:2012)
+gglNE <- suppressMessages(ggloc(loc, "NE", 2003:2012))
 cowplot::plot_grid(plotlist = gglNE, labels = names(gglNE), ncol = 2)
@@
 ```{r MIplot, fig.width = 10, fig.height = 10}
-gglMI <- ggloc(loc, "MI", 2003:2012)
+gglMI <- suppressMessages(ggloc(loc, "MI", 2003:2012))
 cowplot::plot_grid(plotlist = gglMI, labels = names(gglMI), ncol = 2)

### 2017-08-17 doc/RMD/RDA-analysis.Rmd
fix plot names in RDA
update figures and rerun nearly everything

@@
 
-```{r}
+```{r varpart1}
 dat11raw <- genind2df(dat11cc, usepop = FALSE) %>% mutate_all(funs(as.integer(.)))
@@
   theme(aspect.ratio = 1)
-if (!interactive()){
+if (!interactive()) {
   FILE <- file.path(PROJHOME, "results", "figures", "publication", "Figure7Z.pdf")
-  ggsave(filename = FILE, width = 88, height = 88, units = "mm", scale = 1.5)
+  ggsave(filename = FILE, width = 88, height = 88, units = "mm", scale = 1.25)
 }
@@
 
-plot(vp11, digits = 2, Xnames = c("Year", "Region", "Host", "MCG"))
+plot(vp11, digits = 2, Xnames = c("Severity", "Year", "Region", "Host"))
 ```

### 2017-08-17 doc/RMD/by-year.Rmd
update figures and rerun nearly everything

@@
     theme(axis.ticks = element_blank()) + 
-    viridis::scale_color_viridis(discrete = TRUE, direction = -1) +
-    viridis::scale_fill_viridis(discrete = TRUE, direction = -1) +
+    viridis::scale_color_viridis(discrete = TRUE, option = "C", direction = -1) +
+    viridis::scale_fill_viridis(discrete = TRUE,  option = "C", direction = -1) +
     scale_y_continuous(breaks = 0, minor_breaks = yminor) + 
@@
     theme(panel.background = element_rect(fill = "grey95")) +
-    theme(panel.grid.major = element_line(color = "black")) +
+    theme(panel.grid.major = element_line(color = "grey20")) +
     theme(panel.grid.minor = element_line(color = "white")) 
@@
 
-ggscatter(regyear.dapc, strata(dat11cc)) +
+gg_region_year <- ggscatter(regyear.dapc, strata(dat11cc)) +
   facet_wrap(~Region) +
@@
   guides(color = guide_legend(nrow = 4)) +
-  theme(legend.direction = "horizontal")
+  theme(legend.direction = "horizontal") +
+  theme(strip.background = element_rect(color = NA, fill = "grey90")) +
+  theme(strip.text = element_text(face = "bold", hjust = 0.05)) +
+  theme(panel.border = element_blank())
+gg_region_year
+if (!interactive()) {
+  ggsave(filename = file.path(PROJHOME, "results/figures/publication/dapc_region_year.pdf"),
+         plot = gg_region_year,
+         width = 7,
+         height = 7)
+}
+
 ```
@@
 ```{r dapc_plot_newamica, fig.width = 5, fig.height = 4, warning = FALSE, message = FALSE}
-ggscatter(regyear.dapc, strata(dat11cc), filter = quo(Region %in% c("WA", "CA", "NE", "MI"))) +
+gg_region_year_micanewa <- ggscatter(regyear.dapc, strata(dat11cc), filter = quo(Region %in% c("WA", "CA", "NE", "MI"))) +
   facet_wrap(~Region, nrow = 2) +
   theme(legend.position = "right") +
+  theme(legend.justification = "bottom") +
+  theme(legend.box.margin = unit(c(0, 0, 0, 0), "lines")) +
+  theme(legend.margin = unit(c(0, 0, 0, 0), "lines")) +
+  theme(strip.background = element_rect(color = NA, fill = "grey90")) +
+  theme(strip.text = element_text(face = "bold", hjust = 0.05)) +
+  theme(panel.border = element_blank()) +
   theme(legend.key = element_rect(fill = "grey95"))
+gg_region_year_micanewa
+if (!interactive()) {
+    ggsave(filename = file.path(PROJHOME, "results/figures/publication/dapc_region_year_micanewa.pdf"),
+       plot = gg_region_year_micanewa,
+       width = 88,
+       height = 0.8*88,
+       units = "mm",
+       scale = 1.25
+       )
+}
 ```

### 2017-08-17 doc/manuscript/manuscript.Rmd
update figures and rerun nearly everything

@@
 
-We used analysis of molecular variance (AMOVA) with Bruvo's genetic distance in *poppr* version 2.4.1 to test for differentiation associated with wmn by quantifying the differentiation conferred by Source and Region [@excoffier1992analysis; @bruvo2004simple; @kamvar2014poppr].
+We used analysis of molecular variance (AMOVA) with Bruvo's genetic distance in *poppr* version 2.4.1 to test for differentiation associated with wmn by quantifying the differentiation conferred by Region, Source, and Year [@excoffier1992analysis; @bruvo2004simple; @kamvar2014poppr].
 Source in this case was defined as a population belonging to a wmn or not.
@@
 To assess if cultivar had an influence on genetic diversity between wmn, we first subset the clone-corrected data to contain only samples from wmn and from the cultivars Beryl, Bunsi, and G122 and tested differentiation using AMOVA and DAPC as described above.
+We additionally assessed population stability over time by calculating DAPC over the combined groups of Region and Year as described above.
 
@@
 
-The analysis of molecular variance (AMOVA) over the hierarchy of Source and Region showed no significant variation between wmn and non-wmn after accounting for variation between regions (Table \@ref(tab:amova)).
-Because we observed that the sample sizes of wmn isolates and producer isolates were lopsided in several populations (Table \@ref(tab:isolate-table)), we ran the analysis again with regions that had at least ten isolates from both wmn and producer fields, showing similar results (Table \@ref(tab:amova)).
+The analysis of molecular variance (AMOVA) over the hierarchy of Region, Source, and Year showed significant variation between regions and years, but no significant variation between wmn and non-wmn (Table \@ref(tab:amova)).
 In contrast, when we compared the three cultivars, Beryl, Bunsi, and G122, we found no significant differentiation (Supplementary Data).
@@
 The first two discriminant axes failed to separate any of the cultivars where the mean posterior assignment probabilities were 34.0% (G122), 36.0% (Beryl), and 30.1% (Bunsi).
+DAPC grouping by Region and Year used the first 20 principal components, representing 87.7% of the total variance.
+The Midwestern populations (NE, MI, CO, ND) did not appear to have any variation across time in contrast to WA, which showed a shift in population structure in the last year of sampling, 2008 (\@ref(DAPC-RY)).
+Further analysis of this population revealed that all 12 isolates in WA circa 2008 originated in a wmn and nine of these genotypes were shared with CA, and three were shared with France (Fig.
+\@ref(community-graph), \@ref(community-graph16)).
 
@@
 Points represent observed individuals connected to the population centroids with ellipses representing a 66% confidence interval for a normal distribution. 
-The center of each component is represented as grid lines.
+The center of each component is represented as black grid lines.
 ](../../results/figures/publication/DAPC.pdf){#DAPC width=100%}
+
+<!-- end figure -->
+
+<!-- DAPC figure 2: Electric Boogaloo-->
+
+![
+Scatter plot of Discriminant Analysis of Principal Components on Regions and Years showing temporal variation in WA and CA populations as compared to Midwestern Regions, NE and MI. 
+Points (text labels) represent observed individuals connected to the population centroids with ellipses representing a 66% confidence interval for a normal distribution. 
+The center of each component is represented as black grid lines.
+A more detailed view can be seen in Fig. \@ref(DAPC-RY-FULL).
+](../../results/figures/publication/dapc_region_year_micanewa.pdf){#DAPC-RY width=61%}
 
@@
 The primary interest of this study was to assess if isolates sampled from white mold screening nurseries represent isolates from producer fields within the region [@steadman2003identification; @otto-hanson2011variation].
-According to our AMOVA results, we have evidence for differentiation at the region level, but little to no differentiation between white mold screening nursery isolates and production field isolates within region (Table \@ref(tab:amova)).
-When we assessed AMOVA using only populations with equally representative white mold screening nursery isolates and producer field isolates, we see the same pattern, except the $\Phi$ values decrease by nearly half (Table \@ref(tab:amova)).
-When we analyze the AMOVA results in light of the DAPC results (Fig. \@ref(DAPC)), it becomes clear that the patterns of differentiation are largely driven by isolates from Mexico and California. 
-Isolates from these populations had a higher posterior probability (> 0.75) of being reassigned to their own populations than any other (Fig. \@ref(DAPC)a). 
+According to our AMOVA results, we have evidence for differentiation at the Region and Year, but little to no differentiation between wmn isolates and production field isolates (Table \@ref(tab:amova)).
+When we analyze the AMOVA results in light of the DAPC results (Fig. \@ref(DAPC)), it becomes clear that the regional patterns of differentiation are largely driven by isolates from Mexico and CA.
+Isolates from these Regions had a higher posterior probability (> 0.75) of being reassigned to their own populations than any other (Fig. \@ref(DAPC)a). 
 All other populations in comparison (except France) has reassignment probabilities of < 0.5, which is reflected in the failure of the first two discriminant functions to separate these populations (Fig.
 \@ref(DAPC)b).
+
+Despite the evidence that Mexico and CA contributing to a large amount of the population differentiation, Region like WA still had a large amount of internal variation (Supplementary Information).
+The two distinct clusters for the WA Region showed that the 2008 population appeared different and, under further investigation, we found that the genotypes from this population were shared between CA and France (Fig.
+\@ref(DAPC-RY), \@ref(DAPC-RY-FULL), \@ref(community-graph)).
+From the DAPC, we extracted 5 loci that contributed >10% of the observed variation to either of the first two discriminant axes and found that the alleles with an allele frequency of >50% in 2007 were reduced to < 10% in 2008, replaced by an allele that was > 2 mutational steps away (Supplementary Information).
+This could be due to several reasons which include mis-labeling of source material or introduction of migrant sclerotia in plant material.
+Except for the WA Region, populations that were sampled across several years appeared to be relatively stable over time with overlapping distributions in the DAPC (i.e.
+NE and MI, Fig.
+\@ref(DAPC-RY)).
 DAPC is based on the principal components of Euclidean distance between genotypes [@jombart2010discriminant]. 
@@
 
+<!-- DAPC figure 3: The figure strikes back-->
+
+![
+Scatter plot of Discriminant Analysis of Principal Components on Regions and Years showing temporal variation across all Regions. 
+Points (text labels) represent observed individuals connected to the population centroids with ellipses representing a 66% confidence interval for a normal distribution. 
+The center of each component is represented as black grid lines.
+](../../results/figures/publication/dapc_region_year.pdf){#DAPC-RY-FULL width=100%}
+
+<!-- end figure -->
+
 \newpage

### 2017-08-17 doc/RMD/locus-stats.Rmd
Add repeat motif to locus table stats
I also removed the average values because it only
made sense for three of the numbers. Now all but
table 1 are reproducible.

@@
 ```{r locus_table, comment = NA}
-makerange <- . %>% as.integer() %>% range() %>% paste(collapse = "-")
+makerange <- . %>% as.integer() %>% range() %>% paste(collapse = "--")
 
 ranges <- map_chr(alleles(dat11), makerange)
-ranges <- c(ranges, alleles(dat11) %>% unlist(use.names = FALSE) %>% makerange)
+# ranges <- c(ranges, alleles(dat11) %>% unlist(use.names = FALSE) %>% makerange)
 locus_table(dat11, information = FALSE) %>% 
@@
   rownames_to_column("Locus") %>%
+  slice(-n()) %>%
   mutate(Locus = gsub("\\([FH]\\)", "", Locus)) %>%
   add_column(Range = ranges) %>%
-  select(Locus, Range, allele, Hexp, Evenness) %>%
+  add_column(Repeats = other(dat11)$REPLEN[locNames(dat11)]) %>%
+  select(Locus, Range, Repeats, allele, Hexp, Evenness) %>%
+  readr::write_csv(file.path(PROJHOME, "results/tables/locus-table.csv"), col_names = TRUE) %>%
   mutate(allele = round(allele, 2)) %>%
-  rename(`*h*` = Hexp, `No. alleles` = allele) %>%
+  mutate(Repeats = case_when(
+    Repeats <= 2 ~ "di-",
+    Repeats <= 3 ~ "tri-",
+    Repeats <= 4 ~ "tetra-",
+    Repeats <= 5 ~ "penta-",
+    Repeats <= 6 ~ "hexa-"
+  )) %>%
+  rename(`*h*` = Hexp) %>%
+  rename(`No. alleles` = allele) %>%
+  rename(`Repeat Motif` = Repeats) %>%
   huxtable::as_huxtable(add_colnames = TRUE) %>%
-  huxtable::set_number_format(1:12, 3, 0) %>%
-  huxtable::set_number_format(huxtable::everywhere, 4:5, 2) %>%
-  huxtable::set_align(huxtable::everywhere, 3:5, "right") %>%
-  huxtable::set_col_width(c(0.06, 0.07, 0.12, 0.06, 0.09)) %>%
-  huxtable::print_md(max_width = 47)
+  huxtable::set_number_format(huxtable::everywhere, 3:4, 0) %>%
+  huxtable::set_number_format(huxtable::everywhere, 5:6, 2) %>%
+  huxtable::set_align(huxtable::everywhere, 2:6, "right") %>%
+  huxtable::set_col_width(c(0.06, 0.07, 0.11, 0.1, 0.05, 0.075)) %>%
+  huxtable::print_md(max_width = 61)
 ```

### 2017-08-17 doc/manuscript/manuscript.Rmd
Add repeat motif to locus table stats
I also removed the average values because it only
made sense for three of the numbers. Now all but
table 1 are reproducible.

@@
 
---------------------------------------------
-Locus  Range    No. alleles    *h*  Evenness 
------- ------- ------------ ------ ---------
-5-2    318-324            4   0.45      0.62
+```{r , echo = FALSE, warning = FALSE, message = FALSE, results = "asis"}
+suppressPackageStartupMessages(library("tidyverse"))
+loc_columns <- cols(
+  Locus = col_character(),
+  Range = col_character(),
+  Repeats = col_double(),
+  allele = col_integer(),
+  Hexp = col_double(),
+  Evenness = col_double()
+)
+loc_table <- readr::read_csv("../../results/tables/locus-table.csv", 
+                             col_types = loc_columns,
+                             col_names = TRUE)
+loc_table %>%
+  mutate(allele = round(allele, 2)) %>%
+  mutate(Repeats = case_when(
+    Repeats <= 2 ~ "di-",
+    Repeats <= 3 ~ "tri-",
+    Repeats <= 4 ~ "tetra-",
+    Repeats <= 5 ~ "penta-",
+    Repeats <= 6 ~ "hexa-"
+  )) %>%
+  rename(`*h*` = Hexp) %>%
+  rename(`No. alleles` = allele) %>%
+  rename(`Repeat Motif` = Repeats) %>%
+  huxtable::as_huxtable(add_colnames = TRUE) %>%
+  huxtable::set_number_format(huxtable::everywhere, 3:4, 0) %>%
+  huxtable::set_number_format(huxtable::everywhere, 5:6, 2) %>%
+  huxtable::set_align(huxtable::everywhere, 2:6, "right") %>%
+  huxtable::set_col_width(c(0.06, 0.07, 0.11, 0.1, 0.05, 0.075)) %>%
+  huxtable::print_md(max_width = 61)
 
-6-2    483-495            3   0.64      0.95
+Amean <- mean(loc_table$allele)
+Hmean <- mean(loc_table$Hexp)
+Emean <- mean(loc_table$Evenness)
+```
 
-7-2    158-174            7   0.73      0.76
-
-8-3    244-270            7   0.74      0.79
-
-9-2    360-382            9   0.35      0.41
-
-12-2   214-222            5   0.58      0.78
-
-17-3   342-363            7   0.55      0.53
-
-20-3   280-282            2   0.05      0.42
-
-55-4   153-216           10   0.72      0.66
-
-110-4  370-386            5   0.76      0.91
-
-114-4  339-416           10   0.83      0.80
-
-mean   153-495         6.27   0.58      0.69
-
---------------------------------------------
-
-Table: (\#tab:locus-stats) Allelic diversity on full data set at loci used in this study. *h* = Nei's Gene Diversity [@nei1978estimation].
+Table: (\#tab:locus-stats) Allelic diversity on full data set at loci used in this study.
+*h* = Nei's Gene Diversity [@nei1978estimation].
+Average *h* = `r signif(Hmean, 3)`, average Evenness = `r signif(Emean, 3)`, average no.
+alleles = `r signif(Amean, 3)`
 

### 2017-08-17 doc/manuscript/manuscript.Rmd
fix caps in Table 1

@@
 USA       MI     wmn        2003, 2004, 2005, 11A, 37, 38, B07104,         43 
-                                   2008, 2009 Beryl, Bunsi, cornell,       
+                                   2008, 2009 Beryl, Bunsi, Cornell,       
                                               G122, Orion, PO7863, WM31
@@
 USA       WA     wmn        2003, 2004, 2005, 11A, 37, 38, Beryl, Bunsi,   36 
-                                         2008  cornell, G122, Orion, PO7      
+                                         2008  Cornell, G122, Orion, PO7      
                                               104, PO7863, WM31
 
-                 producer          2003, 2007 GH, Merlot, Pinto, redkid    23
+                 producer          2003, 2007 GH, Merlot, Pinto, Redkid    23
 

### 2017-08-17 doc/RMD/by-year.Rmd
adjust supplementary plots

@@
 
-```{r NEplot, fig.width = 10, fig.height = 10}
+```{r NEplot, fig.width = 10, fig.height = 12}
 gglNE <- suppressMessages(ggloc(loc, "NE", 2003:2012))
@@
 ```
-```{r MIplot, fig.width = 10, fig.height = 10}
+```{r MIplot, fig.width = 10, fig.height = 12}
 gglMI <- suppressMessages(ggloc(loc, "MI", 2003:2012))

### 2017-08-17 doc/RMD/by-year.Rmd
use xvalDapc for Year analysis

@@
 
-```{r dapc}
+```{r dapc, cache = TRUE}
 setPop(dat11cc) <- ~Region/Year
-regyear.dapc <- dapc(dat11cc, n.pca = 20, n.da = 20)
+set.seed(2017-08-18)
+regyear.dapc <- xvalDapc(tab(dat11cc), pop(dat11cc), n.pca.max = 20, n.rep = 1000)$DAPC
+regyear.dapc
 ```
@@
 
+quart <- quantile(regyear.dapc$var.contr, 0.95)
 par(mfrow = c(2, 1))
-loadingplot(regyear.dapc$var.contr, axis = 1)
-loadingplot(regyear.dapc$var.contr, axis = 2)
+ax1 <- loadingplot(regyear.dapc$var.contr, axis = 1, threshold = quart)$var.names %>%
+  strsplit("\\.") %>% 
+  map_chr(1) %>%
+  unique()
+ax2 <- loadingplot(regyear.dapc$var.contr, axis = 2, threshold = quart)$var.names %>%
+  strsplit("\\.") %>% 
+  map_chr(1) %>%
+  unique()
 par(mfrow = c(1, 1))
@@
 The loading plot shows us the variables that are important for the first axis of
-separation, mainly:
-
-----------------
- Locus    Allele
--------- -------
-7-2(F)       174
-8-3(H)       256
-9-2(F)       372
-55-4(F)      157
-110-4(H)     370
-----------------
+separation.
 
@@
 
-```{r fig.width = 10, fig.height = 12}
-dat11ccloc <- dat11cc[loc = c("7-2(F)", "8-3(H)", "9-2(F)", "55-4(F)", "110-4(H)")]
+```{r fig.width = 12, fig.height = 14}
+dat11ccloc <- dat11cc[loc = c(ax1, ax2)]
 loc <- map(seploc(dat11ccloc), . %>% 
@@
 
-```{r NEplot, fig.width = 10, fig.height = 12}
+```{r NEplot, fig.width = 12, fig.height = 14}
 gglNE <- suppressMessages(ggloc(loc, "NE", 2003:2012))

### 2017-08-17 doc/RMD/pop-diff.Rmd
Add automatic DAPC reporting in manuscript

@@
 ```{r dapc-region, fig.width = 9, fig.height = 8, cache = TRUE}
+region.dapc
 scatter(region.dapc, 

### 2017-08-17 doc/manuscript/manuscript.Rmd
Add automatic DAPC reporting in manuscript

@@
 knitr::opts_chunk$set(message = FALSE, warning = FALSE)
+```
+
+```{r initiate, echo = FALSE, message = FALSE, warning = FALSE}
+library("poppr")
+load("../../data/sclerotinia_16_loci.rda")
+dat11cc <- clonecorrect(dat11, ~Region/Source/Host/Year)
+
+dapc_region      <- dapc(setPop(dat11cc, ~Region), n.da = 13, n.pca = 20)
+dapc_region_year <- dapc(setPop(dat11cc, ~Region/Year), n.da = 12, n.pca = 22)
+dapc_host <- dat11cc %>% 
+  setPop(~Source) %>%
+  popsub("wmn") %>%
+  setPop(~Host) %>% 
+  popsub(c("Beryl", "Bunsi", "G122")) %>%
+  dapc(n.da = 2, n.pca = 20)
+
+prettify <- function(x, i) signif((1 - x[i])*100, 3)
+paxis <- function(d, i = 1 ){
+  res <- cumsum(d$eig)/sum(d$eig)
+  prettify(res, i)
+}
+passp <- function(d, i = 1){
+  res <- apply(d$posterior, 2, mean, na.rm = TRUE)
+  prettify(1 - res, i)
+}
 ```
@@
 
-Discriminant Analysis of Principal Components (DAPC) was performed grouping by Region with the first 21 principal components, representing 88% of the total variance. 
-The first discriminant axis (representing 64% of the discriminatory power) separated the centroid for the Mexico isolates from the rest of the data, indicating strong differentiation (Fig.
+Discriminant Analysis of Principal Components (DAPC) was performed grouping by Region with the first `r dapc_region$n.pca` principal components, representing `r signif(dapc_region$var*100, 3)`% of the total variance.
+The first discriminant axis (representing `r paxis(dapc_region)`% of the discriminatory power) separated the centroid for the Mexico isolates from the rest of the data, indicating strong differentiation (Fig.
 \@ref(DAPC)b).
-The second discriminant axis separated the centroid for the California isolates.
+The second discriminant axis, representing `r paxis(dapc_region, 2)`% of the discriminatory power, separated the centroid for the California isolates.
 The mean population assignment probabilities for all populations with n > 10 showed that only isolates from Mexico, CA, and France had > 50% probabilities of being reassigned to their source populations (Fig.
@@
 DAPC grouping by cultivar used the first 20 principal components, representing 89% of the total variance.
-The first two discriminant axes failed to separate any of the cultivars where the mean posterior assignment probabilities were 34.0% (G122), 36.0% (Beryl), and 30.1% (Bunsi).
-DAPC grouping by Region and Year used the first 20 principal components, representing 87.7% of the total variance.
-The Midwestern populations (NE, MI, CO, ND) did not appear to have any variation across time in contrast to WA, which showed a shift in population structure in the last year of sampling, 2008 (\@ref(DAPC-RY)).
+The first two discriminant axes failed to separate any of the cultivars where the mean posterior assignment probabilities were `r passp(dapc_host, "G122")`% (G122), `r passp(dapc_host, "Beryl")`% (Beryl), and `r passp(dapc_host, "Bunsi")`% (Bunsi).
+DAPC grouping by Region and Year used the first `r dapc_region_year$n.pca` principal components, representing `r signif(dapc_region_year$var*100, 3)`% of the total variance.
+The Midwestern populations (NE, MI, CO, ND) did not appear to have any variation across time in contrast to WA, which showed a shift in population structure in the last year of sampling, 2008 (Fig.
+\@ref(DAPC-RY)).
 Further analysis of this population revealed that all 12 isolates in WA circa 2008 originated in a wmn and nine of these genotypes were shared with CA, and three were shared with France (Fig.

### 2017-08-18 doc/RMD/locus-stats.Rmd
Clean out a lot of old results
There were extra figures and tables littering the
results directory due to various things including
misspelling. These have been cleared.

@@
 load(file.path(PROJHOME, "data", "sclerotinia_16_loci.rda"))
+dir.create(file.path(PROJHOME, "results", "tables"))
 ```

### 2017-08-18 doc/manuscript/manuscript.Rmd
Clean out a lot of old results
There were extra figures and tables littering the
results directory due to various things including
misspelling. These have been cleared.

@@
 
-prettify <- function(x, i) signif((1 - x[i])*100, 3)
-paxis <- function(d, i = 1 ){
+# convert fractions to percents with 3 sig figs
+prettify <- function(x, i) signif((x[i])*100, 3)
+
+# Get proportion of variance per discriminant axis
+paxis <- function(d, i = 1){
   res <- cumsum(d$eig)/sum(d$eig)
-  prettify(res, i)
+  prettify(1 - res, i)
 }
+
+# Get proportion of assignment probability for a given population
 passp <- function(d, i = 1){
   res <- apply(d$posterior, 2, mean, na.rm = TRUE)
-  prettify(1 - res, i)
+  prettify(res, i)
 }

### 2017-08-18 doc/RMD/RDA-analysis.Rmd
change size of points in RDA plot

@@
                color = "grey45", 
-               size = 2.5, 
+               size = 3, 
                pch = 21)+

### 2017-08-18 doc/manuscript/manuscript.Rmd
fix duplicated word
Add acknowledgements and contributions

@@
 
-\newpage
-
 ## Population Differentiation {-}
@@
 
+## Acknowledgements {-}
+
+The authors would like to thank Rebecca Higgens for providing valuable insights
+into the historical context of the data collection and curation. Funding for
+this research was provided by the North Central Soybean Research Program
+(#639K623), a Layman Award (#2446), and start-up funds from the University of
+Nebraska-Lincoln.
+
+## Conflicts of Interest {-}
+
+The authors declare no conflict of interest.
+
+## Author Contributions {-}
+
+<!-- Conceptualization: SEE. -->
+<!-- Formal analysis: BSA SEE. -->
+<!-- Funding acquisition: SEE. -->
+<!-- Investigation: BSA. -->
+<!-- Resources: SEE. -->
+<!-- Supervision: BSA. -->
+<!-- Visualization: BSA SEE. -->
+<!-- Writing – original draft: BSA. -->
+<!-- Writing – review & editing: SEE. -->
+
 <!-- 
@@
 
+\newpage
 # Supplementary Information {-}
@@
 Table: (\#tab:region-aggressiveness) Mean aggressiveness ratings for the Regions with greater than 5 samples; groupings according to 95% family-wise confidence interval.
-
-\newpage
 

### 2017-08-18 doc/manuscript/manuscript.Rmd
Updated acknowledgements

@@
 this research was provided by the North Central Soybean Research Program
-(#639K623), a Layman Award (#2446), and start-up funds from the University of
-Nebraska-Lincoln.
+(#639K623) to SEE, a Layman Award (#2446) to SEE, USDA-ARS National Sclerotinia
+Initiative (#58-5442-2-209) to JRS/SEE, and start-up funds from the University of
+Nebraska-Lincoln to SEE.
 

### 2017-08-18 doc/manuscript/manuscript.Rmd
update abstract; remove old commented text

@@
 
-<!-- 
-Our study was conducted to characterize *S.
-sclerotiorum* isolates from different geographic areas to determine important phenotypic variables influencing genotypes; to ascertain reproductive behavior of the pathogen; to test whether different sampling years or locations have any effect on genotype/sexuality; and to identify abundant and unique genotypes across isolates and geographies.
-The dbRDA revealed region, year, and host as the most influential phenotypic variables on genotypes.
-However, further molecular analysis of pathogen isolates from selected hosts did not show any significant genetic difference among populations.
-According to dbRDA, aggressiveness and location of isolates had little or no effect.
-Previous studies on *S.
-sclerotiorum* from alfalfa soil [@attanayake2012genetic] and potato [@atallah2004high], and canola [@attanayake2013sclerotinia] plants reported little or no correlation of genotypes on pa25thogen aggressiveness.
-This is important information for resistant breeding since genotypes of *S.
-sclerotiorum* isolates seem to depend to a certain degree on different levels of region and also on environmental conditions of the sampling year.
-For instance our analysis indicated that MX, 2009, and 2005 were highly influential predictors of *S. sclerotiorum* genotypes (Fig. 7).
-Conversely, France and Australia, which are geographically distant to North America, were not within the top 10 predictors of genetic composition of isolates (Fig. 7).
-According to dbRDA, Mexican region imposed the highest impact on microsatellite loci of *S. sclerotiorum* isolates.
-We used Bruvo's distance for calculating genetic distances of different isolates since it incorporates stepwise mutation model and applicable to microsatellite analysis.
-
-The 366 isolates of our study belonged to 87 MCGs and 165 MLGs.
-MCGs with multiple isolates were represented by several to many microsatellite genotypes and demonstrated lack of similarity between MCGs and microsatellite genotypes.
-For instance, MCG 5 with the highest number of isolates (N = 73) represented 37 MLGs when characterized by 11 microsatellites.
-The most common genotype among these 73 isolates was MLG 25 with 19 isolates.
-Similarly, second most common MCG with 44 isolates was represented by 19 genotypes.
-Some studies on *S.
-sclerotiorum* found a strong association between MCG and multilocus genotypes [@aldrich-wolfe2015genetic] while others reported no evidence for such consistent association [@atallah2004high; @attanayake2012genetic] or either reported an association for a part of the sample [@kohn1995clonal].
-@attanayake2013sclerotinia carried out a genetic diversity study of *S. sclerotiorum* populations from canola in China and USA.
-They reported the MCGs were in accordance with the microsatellite MLGs in the U.S. population but not in the Chinese population.
-A study conducted on Canadian field populations of cultivated canola revealed MCGs and RFLP fingerprinting of *S.
-sclerotiorum* isolates were associated but MCGs of Norwegian isolates from wild perennial plant Ranunculua ficaria did not show such an association [@kohn1995clonal].
-@aldrich-wolfe2015genetic characterized 145 isolates from four different host crops in north central region of the USA.
-Isolates belonging to same MCG either shared identical genotypes or differed at no more than 2 of the 12 microsatellite loci.
-They reported only 14% of MCGs contained isolates with different genotypes.
-Different results were reported by @attanayake2012genetic, who characterized 40 *S. sclerotiorum* isolates from 1 m^2^ area of alfalfa seed production soil using microsatellite markers.
-There were 15 MCGs with seven representing just a single isolate.
-Though single-isolate MCGs contained distinct multilocus genotype, they also found several instances where MCGs and microsatellite MLGs did not similarly correspond.
-Some isolates of the same multilocus genotype belonged to different MCGs.
-Our results agree with this type of relationship between MCGs and MLGs.
-Similarly, @atallah2004high did not find any strong correlation between microsatellite haplotypes and MCGs in 167 *S. sclerotiorum* isolates from four potato fields in the Columbia basin of WA state.
-These 167 isolates belonged to 82 MCGs with 64 single-isolate MCGs.
-We came across 43 single-isolate MCGs among 366 isolates from dry bean cultivars.
-
-Our study did not find any multilocus genotype/s common across all or majority of sampling geographies.
-<!-~ ZNK: I think this statement is a bit misleading.
-Yes, we didn't find any individual MLGs, but 46% of these are present across at least two populations.
-~->
-The most widespread and common genotype (MLG 25, Supplemental Table 1) was found only in five states of the USA with majority of isolates confined to ND.
-However, we did find some MLGs of France and Australian isolates to be present in the USA samples.
-Some studies on *S. sclerotiorum* did not detect any shared MLGs between geographically distant populations.
-@attanayake2013sclerotinia were unable to detect any shared microsatellite MLGs in *S. sclerotiorum* populations from infected canola fields in China and USA.
-Isolate collections were limited to a single county in China whereas USA isolates were from ND state only.
-This may have limited the chance of discovering genotypes common to both regions.
-@atallah2004high examined the population structure of *S.
-sclerotiorum* in different potato fields of Columbia basin and reported only 15 isolates shared the most common microsatellite haplotype in a sample of 167.
-Similarly, several other studies did not detect common fingerprints of *S.
-sclerotiorum* isolates in geographically distant populations: Australia and North America [@ekins1999genetic];  California, Washington, and Ontario [@malvarez2007new]; Canada, North Carolina, and Louisiana [@cubeta1997clonality].
-
-Contrary to above results, @aldrich-wolfe2015genetic who did a survey on *S.
-sclerotiorum* from different host crops in the North Central states of the United States found a single microsatellite haplotype to be prevalent throughout the region.
-A study in the UK that examined this pathogen from 12 populations in different crop plants and weed species *Ranunculus acris* (meadow buttercup) also reported one microsatellite haplotype was widely distributed across hosts and areas [@clarkson2013population].
-Reason for our isolates from dry bean not to show a widespread MLG may be due to extensive sexual outcrossing and movement sclerotia by human activities.
-Though conditions of outcrossing for *S.
-sclerotiorum* is not well studied, it is believed weather and agricultural conditions play a role in production and release of ascospores of this pathogen [@huang1991induction; @sexton2006population].
-One study found release of spores from apothecia declined over 36- 168 H at low RH of 65-75% but not at high RH of 90-95% [@clarkson2003ascospore].
-The same study reported increased ascospores discharge at 20-25$^{\circ}$ C compared to 15$^{\circ}$ C and the survival of *S. sclerotiorum* ascospores was reduced at high temperature and RH.
-Our samples were collected from 2003 to 2012 and except for the white mold screening nurseries, came from different fields.
-Different agro-ecological conditions prevailed during sampling seasons may have resulted in different rates of sexual outcrossing of the pathogen and hence reduced the chance of extracting a common genotype.
-When conditions are favorable for ascospore production, release, and survival, co-infection of a single site on a plant by multiple ascospores is possible [@sexton2006population].
-This may eventually result in facilitating outcrossing and recombination.
-
-Interestingly, all Mexican genotypes were unique to that region.
-On the other hand CA genotypes were present in the West Coast states of OR and WA which may be due to movement of sclerotia with planting material.
-<!-~ ZNK: Do we have some sort of evidence for this, or is this pure speculation?
-~->
-<!-~ 
-I'm not sure what the point of this paragraph is.
-I believe I can put the discussion of the graph results here. 
-~->
-Surprisingly, France and Australian genotypes were also isolated in the USA.
-MLG 4 (see supplemental Table 1) was recorded from France, Australia and MN while France and USA isolates had four common MLGs.
-Similarly, five MLGs were recorded common to both Australia and USA isolates.
-However, there were six unique genotypes in clone-corrected France population (N= 6/14) and three genotypes in CA population (N = 3/15).
-MLGs from Australian region, our samples were limited to 6, which is not a sufficient number to make inferences.
-Our results indicated that different continents were not a defining factor for microsatellite genotypes of *S. sclerotiorum* isolated from dry bean.
-This may be due to involuntary movement of sclerotia by human activities and probably because of recent speciation of *S. sclerotiorum*. 
-A phylogenetic study on Sclerotinaceae showed that most of the sclerotial lineages including different species of Sclerotinia had evolved relatively recently from a common ancestor [@carbone1993ribosomal].
-Few multilocus genotypes were unique to each sampling state of the USA but most of them were common among several states (Supplemental Table 1).
-For instance, Michigan state from which we had the largest collection of isolates (N = 62) consisted of 43 MLGs.
-Of these, 32 MLGs were present in at least one of the 5 states of NE, ND, WA, CO, MN; 4 MLGs were present in France isolates; and 11 were unique to MI sites (see Supplemental Table 1 for unique MLGs in different regions).
-When selecting *S. sclerotiorum* isolates for screening dry bean germplasm for disease resistance, plant breeders may find above information useful.
-A study on genetic diversity of *S.
-sclerotiorum* from multiple host crops in the North central region of the USA revealed there was little evidence for genetic isolation of this pathogen by host crop or by distance [@aldrich-wolfe2015genetic].
-Though our isolates didn't come from different crops, we had isolates from many resistant and susceptible dry bean cultivars, which also didn't show genetic differentiation according to AMOVA.
-It seems Mexican pathogen isolates are different from the rest and warrants further investigation using a larger population.
-Some France genotypes were common in USA isolates but at the same time 43% (6/14) of the France genotypes were unique to that region indicating white mold resistant cultivars produced in the USA might not work in that region.
-<!-~ ZNK: I'm not entirely sold on this conclusion.
-~->
-Geographic distance between different states of the USA seem not a barrier for gene transfer of *S.
-sclerotiorum* isolates though unique genotypes did prevail at a low frequency in each state.
-<!-~ ZNK: I don't know that this adds anything.
-~->
-The genetic diversity of regions with >10 isolates had similar values (0.43 - 0.59) except for Mexican isolates.
-*h* of MX was 0.26 and indicated less exposure of sampling sites to outside inocula. <!-~ ZNK: Does this mean that there's heavy inbreeding here? ~->
-Test for linkage disequilibrium showed evidence of clonal reproduction for MX isolates (Ia = 2.61, P.Ia = 0.001, $\bar{r}_d$ = 0.37, P.$\bar{r}_d$  = 0.001).
-This may have been another reason for low gene diversity.
-
-DAPC analysis by region followed by AMOVA corroborated the uniqueness of Mexican genotypes by grouping them away from the rest of the isolates (Fig. 8).
-Other populations based on geography were unable to group in separate clusters except for the Californian population.
-We tested whether genotypes of *S. sclerotiorum* populations were influenced by different dry bean host cultivars and year of sampling by DAPC and AMOVA.
-There was no host specificity but year of sampling did show a significant impact on genotypes.
-Other studies also reported host identity is no barrier for gene flow of this pathogen [@aldrich-wolfe2015genetic; @clarkson2013population].
-@atallah2004high revealed statistically significant impact of the collection date of *S. sclerotiorum* isolates from four potato fields in the Columbia Basin on molecular variance.
-But there was no impact of the year or location of collection.
-We also did not detect location as a significant predictor of molecular variance but year had a significant impact.
-Our samples were collected on a single day in each year from each field and therefore did not warrant for an analysis of date of sample collection. <!-~ BSA: Is this true, coauthors? ~->
-
-We performed DAPC without assigning isolates to different populations based on sampling information to see the how isolates fall into clusters according to their genetic relatedness.
-We decided the most appropriate number as six since the DAPC results had 87% (N = 13/15) of Mexican isolates in a single cluster (Fig. 9 cluster 3).
-When number of clusters was assigned less than that MX isolates did not group separately while higher number of cluster assignment resulted in breaking down of other clusters in Fig.
-9 into smaller groups.
-DAPC analysis without population assignments revealed that in general geography is not a barrier for gene flow since isolates of different geographies from USA, Australia and France clustered together indicating genetic relatedness and possible sexual recombination within each region.
-For instance France isolates were present in cluster 1, 2, 4 and 5.
-As mentioned above only exception was the Mexican isolates.
-@njambere2014genotypic studied *S. trifoliorum*, a close relative *S. sclerotiorum* and reported four cryptic genetic populations by Bayesian clustering analysis of microsatellite loci.
-Those clusters had no correlation to geographic locations of samples.
-Similarly, a cluster analysis of microsatellite data of *S.
-sclerotiorum* collected from 12 geographically distant host populations in  England and Wales resulted only two main clusters pooling isolates from different populations together [@clarkson2013population].
-However, some other studies found *S.
-sclerotiorum* isolates from distant locations did differ significantly when molecularly characterized [@attanayake2013sclerotinia; @cubeta1997clonality; @ekins1999genetic @malvarez2007new].
-
-There were isolates that shared the same MLG belonging to different MCGs.
-For instance MLG 25 consisted of 27 isolates in 5 MCGs.
-This phenomenon is suggestive of a recombination by sexual outcrossing [@njambere2014genotypic; @sexton2006population].
-In general high level of MLG diversity; and lack of association of MCG and microsatellite genotype are evidence of *S.
-sclerotiorum* undergoing sexual outcrossing and movement of isolates due to agricultural practices [@njambere2014genotypic].
-Our isolates had a total of 165 MLGs and they did not correspond well with 87 MCGs.
-Linkage disequilibrium analysis of populations in white mold screening nurseries indicated the possibility of sexual outcrossing in some regions (ex.
-CA, FR, OR and WA) during certain sampling years while most locations indicated clonal reproduction.
-For example wmn of WA was sampled in 2003, 2004, 2005, and 2008 but only 2008 isolates had evidence of outcrossing.
-Similar to our results many studies have reported primarily clonal reproduction of *S.
-sclerotiorum* populations with evidence of occasional outcrossing:  canola and sunflower pathogen populations in north central region of the USA [@aldrich-wolfe2015genetic]; cabbage population in North Carolina [@cubeta1997clonality]; Canadian canola and North Carolina cabbage populations [@kohli1998random]; and 12 populations from five crops in England and Wales [@clarkson2013population].
-Contrary to above, one study from an experimental cabbage field plot in Louisiana, found a population of 16 isolates of *S. sclerotiorum* were strictly clonal [@cubeta1997clonality].
-@atallah2004high studied the genetic structure of *S. sclerotiorum* in four potato fields in the Columbia basin and reported evidence of potential outcrossing of populations in two fields.
-When all isolates were combined, there was significant linkage disequilibrium and hence the population was highly clonal.
-Our study supports this phenomenon since linkage disequilibrium analysis revealed evidence of sexual outcrossing only in CA when populations were based on region.
-However, when populations were sub divided according each year/region, we could detect evidence of sexual out crossing in some populations.
-One reason for this is the overshadowing of low frequency genotypes of heterothallic recombination by a few successful genotypes of clonal nature [@atallah2004high].
-
-*Sclerotinia sclerotiorum* isolates from dry bean represented high microsatellite diversity and lack of association MCGs and microsatellite MLGs.
-Analysis of molecular variance of populations based on region and sampling year showed low level of differentiation with majority of variance detected within populations.
-These features are indicative of sexual outcrossing and perhaps human mediated migration of *S. sclerotiorum* in dry bean.
-However, we also detected clones or identical microsatellite MLGs within and among populations indicating asexual reproduction.
-Further, linkage disequilibrium and index of association calculations suggested *S. sclerotiorum* populations are largely clonal.
-Although, there was low impact of region and year on population differentiation, we did not detect host cultivar as a barrier to gene flow of the pathogen.
-Our study did not find a widespread genotype among sampled regions from dry beans but some Australian and France MLGs were found in the USA isolates.
-The analysis of populations from same field over different years for linkage disequilibrium did not provide evidence for sexual outcrossing.
-However, analysis of molecular variance did show populations were significantly different between some years (eg.
-MI 2003 and 2005 populations; WA 2008 compared to both 2003 and 2005).
-These results indicate some level of heterothallic recombination is taking place in these fields in addition to natural mutations.
-An understanding of sexual behavior and ability of form new genotypes by *S. sclerotiorum* in dry bean cropping systems is useful for formulating disease management strategies. -->
-
 \newpage

### 2017-08-18 doc/manuscript/manuscript.Rmd
Add conclusions

@@
 
-eh.
+This study represents the largest genetic analysis of *S.
+sclerotiorum* from the USA to date, giving us a unique insight to continent-wide population structure and relationships between phenotypic and genotypic variables.
+Based on our analysis of the relationships between MCG and MLG, we found no clear evidence that the two are directly related, suggesting that MCG do not necessarily represent vegetative compatibility groups and thus should not be used as a proxy for identifying clones.
+Populations in wmn appear to show no significant differentiation when compared to their production field counterparts, suggesting that the wmn populations of *S.
+sclerotiorum* may be considered representative of the surrounding regions.
+While we found no direct relationship between genotype and severity, it is evident that there is a gradient of severity by region, further supporting the need for screening in multiple locations.
 
@@
 
-The authors would like to thank Rebecca Higgens for providing valuable insights
-into the historical context of the data collection and curation. Funding for
-this research was provided by the North Central Soybean Research Program
-(#639K623) to SEE, a Layman Award (#2446) to SEE, USDA-ARS National Sclerotinia
-Initiative (#58-5442-2-209) to JRS/SEE, and start-up funds from the University of
-Nebraska-Lincoln to SEE.
+The authors would like to thank Rebecca Higgens for providing valuable insights into the historical context of the data collection and curation.
+Funding for this research was provided by the North Central Soybean Research Program (#639K623) to SEE, a Layman Award (#2446) to SEE, USDA-ARS National Sclerotinia Initiative (#58-5442-2-209) to JRS/SEE, and start-up funds from the University of Nebraska-Lincoln to SEE.
 
@@
 <!-- Visualization: BSA SEE. -->
-<!-- Writing – original draft: BSA. -->
-<!-- Writing – review & editing: SEE. -->
+<!-- Writing - original draft: BSA. -->
+<!-- Writing - review & editing: SEE. -->
 

### 2017-08-18 doc/RMD/RDA-analysis.Rmd
Update manuscript to auto-generate results
I also added a comment on data availability and
a comment on the low value of rd in CA

@@
 ```{r marginal_anova, cache = TRUE}
-anova(cap11cc, by = "margin")
-anova(cap16cc, by = "margin")
+set.seed(2017-08-17)
+(cap11anova <- anova(cap11cc, by = "margin"))
+set.seed(2017-08-17)
+(cap16anova <- anova(cap16cc, by = "margin"))
 ```
@@
 ```{r total_anova2, cache = TRUE}
+set.seed(2017-08-17)
 anova(cap11cc2)
@@
 ```{r marginal_anova2, cache = TRUE}
-anova(cap11cc2, by = "margin")
+set.seed(2017-08-17)
+(cap11anova2 <- anova(cap11cc2, by = "margin"))
 ```
@@
 
+## Saving Data
+
+These data are important for reporting in the manuscript and expensive to 
+generate, so I will save them here
+
+```{r save_data}
+save(cap11cc, cap11cc2, 
+     cap11sum, cap11sum2, 
+     cap11anova, cap11anova2,
+     file = file.path("data", "RDA_results.rda"))
+```
+
 # Conclusions

### 2017-08-18 doc/manuscript/manuscript.Rmd
Update manuscript to auto-generate results
I also added a comment on data availability and
a comment on the low value of rd in CA

@@
 load("../../data/sclerotinia_16_loci.rda")
+load("../../data/RDA_results.rda")
 dat11cc <- clonecorrect(dat11, ~Region/Source/Host/Year)
@@
 }
+
+vartab <- function(x){
+  tot <- x$tot.chi
+  con <- x$constr.chi
+  uco <- x$unconst.chi
+  data.frame(list(Inertia = c(tot, con, uco),
+                  Proportion = c(tot, con, uco)/tot))
+}
+RDA2R2 <- prettify(vegan::RsquareAdj(cap11cc2)$adj.r.squared)/100
+RDA1R2 <- prettify(vegan::RsquareAdj(cap11cc)$adj.r.squared)/100
+
+anova1res <- setNames(cap11anova$`Pr(>F)`, rownames(cap11anova))
+anova2res <- setNames(cap11anova2$`Pr(>F)`, rownames(cap11anova2))
 ```
@@
 All analyses were performed on this data set unless otherwise mentioned.
-All data processing and analyses were performed in R version 3.4.1 [@R] and are openly available and reproducible at https://github.com/everhartlab/sclerotinia-366/.
+All data processing and analyses were performed in `r substr(R.version.string, 1, 15)` [@R] and are openly available and reproducible at https://github.com/everhartlab/sclerotinia-366/.
 Of the 16 microsatellite loci genotyped, five included compound repeats, which made it challenging to accurately/confidently bin alleles into fragment sizes expected for each locus based on the described repeat motif.
@@
 Loci with compound repeats were removed for the reported statistics, but were processed and included in data analyses to ensure the integrity of the results with 11 loci as compared to 16.
-We assessed the power of our 11 markers by generating a genotype accumulation curve in the R package *poppr* version 2.4.1, looking for evidence of saturation, which would indicate that loci were sufficiently sampled to adequately represent the full set of genotypes [@arnaud2007standardizing; @kamvar2015novel].
+We assessed the power of our 11 markers by generating a genotype accumulation curve in the R package *poppr* version `r packageVersion("poppr")`, looking for evidence of saturation, which would indicate that loci were sufficiently sampled to adequately represent the full set of genotypes [@arnaud2007standardizing; @kamvar2015novel].
 To avoid including isolates potentially collected from the same plant (which increases the probability of collecting sclerotioa from the same point of infection more than once), data were clone-corrected on a hierarchy of region/source/host/year---meaning that one representative of each genotype was retained for each combined regions, source, host, and year---for subsequent analysis.
@@
 To assess the potential for random mating, we tested for linkage disequilibrium with the index of association, $I_A$ and its standardized version, $\bar{r}_d$ using 999 permutations [@agapow2001indices; @smith1993how; @brown1980multilocus].
-Both genotypic diversity and linkage disequilibrium were calculated in *poppr* version 2.4.1 [@kamvar2014poppr].
+Both genotypic diversity and linkage disequilibrium were calculated in *poppr* [@kamvar2014poppr].
 
@@
 
-A distance-based redundancy analysis (dbRDA) [@legendre1999distance] was performed with the function `capscale()` in the *vegan* package version 2.4-3 [@vegan].
+A distance-based redundancy analysis (dbRDA) [@legendre1999distance] was performed with the function `capscale()` in the *vegan* package version `r packageVersion("vegan")` [@vegan].
 This method uses constrained ordinations on a distance matrix representing the response variable to delineate relative contribution of any number of independent explanatory variables. 
 We used this method to delineate the phenotypic (Aggressiveness, Mycelial Compatibility Group (MCG)), geographic (Region, Host, Location), and temporal (Year) components in predicting genetic composition of the populations.
-<!-- RDA is a multivariate linear regression analysis of independent (explanatory: eg. environmental and phenotypic) variables on dependent (non-explanatory or response) variables. -->
-The distance matrix we used as the response variable was generated using Bruvo's genetic distance from clone-corrected data (procedure described above) as implemented in *poppr* version 2.4.1, which employed a stepwise mutation model for microsatellite data [@bruvo2004simple; @kamvar2014poppr].
+The distance matrix we used as the response variable was generated using Bruvo's genetic distance from clone-corrected data (procedure described above) as implemented in *poppr*, which employed a stepwise mutation model for microsatellite data [@bruvo2004simple; @kamvar2014poppr].
 Because aggressiveness measures differed with each isolate, aggressiveness first averaged across clone-corrected isolates.
-<!-- In our case, a pairwise Bruvo's [@bruvo2004simple] distance matrix was generated from microsatellite data using *poppr* 2.4.1 [@kamvar2014poppr; @kamvar2015novel] and employed it as the dependent matrix for RDA analysis.
--->
 To identify explanatory variable(s) correlated with genetic variation, a forward-backward selection process was applied with the vegan function `ordistep()` to identify the set of variables that explained more variation than predicted with the global model containing all variables.
@@
 To minimize complications due to small sample sizes, we chose the top 10 MCGs, representing 56.5% of the isolates collected, the top 10 MLGs representing 26.7% of the isolates, and populations with more than five isolates.
-Between-group differences were assessed using Tukey's HSD test with an alpha of 0.05 using the `HSD.test()` function in the package *agricolae* version 1.2.4 [@mendiburu2015agricolae].
+Between-group differences were assessed using Tukey's HSD test with an alpha of 0.05 using the `HSD.test()` function in the package *agricolae* version `r packageVersion("agricolae")` [@mendiburu2015agricolae].
 
@@
 This was then converted to a bipartite graph where top nodes represented MCGs and bottom nodes represented MLGs.
-The cluster walktrap community detection algorithm as implemented in the `cluster_walktrap()` function in *igraph* version 1.1.2 was used to define clusters of MLGs and MCGs [@csardi2006igraph; @pons2006computing].
+The cluster walktrap community detection algorithm as implemented in the `cluster_walktrap()` function in *igraph* version `r packageVersion("igraph")` was used to define clusters of MLGs and MCGs [@csardi2006igraph; @pons2006computing].
 All communities that had fewer than 10 members were then consolidated into one.
@@
 
-We used analysis of molecular variance (AMOVA) with Bruvo's genetic distance in *poppr* version 2.4.1 to test for differentiation associated with wmn by quantifying the differentiation conferred by Region, Source, and Year [@excoffier1992analysis; @bruvo2004simple; @kamvar2014poppr].
+We used analysis of molecular variance (AMOVA) with Bruvo's genetic distance in *poppr* to test for differentiation associated with wmn by quantifying the differentiation conferred by Region, Source, and Year [@excoffier1992analysis; @bruvo2004simple; @kamvar2014poppr].
 Source in this case was defined as a population belonging to a wmn or not.
 
-To identify specific Regions that were more differentiated than others, we used discriminant analysis of principal components (DAPC) as implemented in *adegenet* version 2.1.0, assessing the per-sample posterior group assignment probability [@jombart2008adegenet].
+To identify specific Regions that were more differentiated than others, we used discriminant analysis of principal components (DAPC) as implemented in *adegenet* version `r packageVersion("adegenet")`, assessing the per-sample posterior group assignment probability [@jombart2008adegenet].
 This method decomposes the genetic data into principal components, and then uses a subset of these as the inputs for discriminant analysis, which attempts to minimize within-group variation and maximize among-group variation [@jombart2010discriminant].
@@
 We first tabulated the multilocus genotypes shared between at least two populations (defined as states or countries) with the *poppr* function `mlg.crosspop()` [@kamvar2014poppr].
-From these data, we constructed a graph with populations as nodes and shared genotypes as edges (connections) between nodes using the R packages *igraph* version 1.0.1 [@csardi2006igraph], *dplyr* version 0.7.1 [@dplyr], and *purrr* version 0.2.2.2 [@purrr].
+From these data, we constructed a graph with populations as nodes and shared genotypes as edges (connections) between nodes using the R packages *igraph* [@csardi2006igraph], *dplyr* version `r packageVersion("dplyr")` [@dplyr], and *purrr* version `r packageVersion("purrr")` [@purrr].
 Each node was weighted by the fraction of shared MLGs.
@@
 We then used the `cluster_optimal()` function from *igraph* to define communities [@csardi2006igraph].
-The graph was plotted using the R packages *ggplot2* version 2.2.1 [@ggplot2] and *ggraph* version 1.0.0 [@ggraph].
+The graph was plotted using the R packages *ggplot2* version `r packageVersion("ggplot2")` [@ggplot2] and *ggraph* `r packageVersion("ggraph")` [@ggraph].
 To ensure that we captured the same community signal, we additionally performed this analysis including the five polymorphic markers described above.
@@
 
-The forward-backward selection process of the dbRDA models on clone-corrected data revealed Year, Region, Host, and Aggressiveness to be the optimal variables for the reduced model, accounting for 45% of the total variation.
-ANOVA revealed that the reduced model was significant with an adjusted R^2^ of 0.0674 (*P* = 0.001).
-Assessment of the marginal effects showed that region, host, year, and aggressiveness significantly explained genetic variation (*P* = 0.001, *P* = 0.009 for region).
+The forward-backward selection process of the dbRDA models on clone-corrected data revealed Year, Region, Host, and Aggressiveness to be the optimal variables for the reduced model, accounting for `r prettify(vartab(cap11sum)[2, "Proportion"])`% of the total variation.
+ANOVA revealed that the reduced model was significant with an adjusted R^2^ of `r RDA1R2` (*P* = 0.001).
+Assessment of the marginal effects showed that region, host, year, and aggressiveness significantly explained genetic variation (*P* = `r anova1res[["Year"]]`, *P* = `r anova1res[["Region"]]` for region).
 It was revealed that there was multicollinearity when MCG was combined with any other variable, so we followed the same procedure, dropping MCG from the list of potential predictors. 
-From these results, we found that aggressiveness, year, region, and host were optimal, accounting for 17.6% of the total variation.
-ANOVA revealed significant effects with an adjusted R^2^ of 0.0325 (*P* = 0.001).
-While the marginal effect assessment showed that region, host, and year all significantly explained the variation at *P* = 0.001, and aggressiveness significantly explained the variation at *P* = 0.045.
+From these results, we found that aggressiveness, year, region, and host were optimal, accounting for `r prettify(vartab(cap11sum2)[2, "Proportion"])`% of the total variation.
+ANOVA revealed significant effects with an adjusted R^2^ of `r RDA2R2` (*P* = 0.001).
+While the marginal effect assessment showed that region, host, and year all significantly explained the variation at *P* = `r anova2res[["Year"]]`, and aggressiveness significantly explained the variation at *P* = `r anova2res[["Severity"]]`.
 Much of the variation appears to be driven by isolates from Mexico and 2005 (Fig. \@ref(rda-biplot)).
@@
 \@ref(aggressiveness-boxplot), Table \@ref(tab:region-aggressiveness)).
-MCGs also showed a significant effect (P < 1.00e^-4^), with means that ranged from 6.0 (MCG 44) to 4.6 (MCG 49).
-We additionally found a significant effect for MLGs (*P* = 7.44e^-4^) for MLGs, with means that ranged from 6.0 (MLG 78) to 4.3 (MLG 140).
+MCGs also showed a significant effect (P < 1.00e^-4^), with means that ranged from 6.0 ('MCG 44') to 4.6 ('MCG 49').
+We additionally found a significant effect for MLGs (*P* = 7.44e^-4^) for MLGs, with means that ranged from 6.0 ('MLG 78') to 4.3 ('MLG 140').
 
@@
 The transparency of the bars represent the evenness of the distribution of the MLGs within a given MCG.
-A dashed box surrounds the 8 most common MCGs representing >51% of the data.
+A dashed box surrounds the 8 most common MCGs representing > 51% of the data.
 **B)** Full graph-representation of the relationship between MCGs (open circles) and MLGs (filled circles).
@@
 The analysis of molecular variance (AMOVA) over the hierarchy of Region, Source, and Year showed significant variation between regions and years, but no significant variation between wmn and non-wmn (Table \@ref(tab:amova)).
-In contrast, when we compared the three cultivars, Beryl, Bunsi, and G122, we found no significant differentiation (Supplementary Data).
+In contrast, when we compared the three cultivars, Beryl, Bunsi, and G122, we found no significant differentiation (Supplementary Information).
 
@@
 
-Discriminant Analysis of Principal Components (DAPC) was performed grouping by Region with the first `r dapc_region$n.pca` principal components, representing `r signif(dapc_region$var*100, 3)`% of the total variance.
+Discriminant Analysis of Principal Components (DAPC) was performed grouping by Region with the first `r dapc_region$n.pca` principal components, representing `r prettify(dapc_region$var)`% of the total variance.
 The first discriminant axis (representing `r paxis(dapc_region)`% of the discriminatory power) separated the centroid for the Mexico isolates from the rest of the data, indicating strong differentiation (Fig.
@@
 The first two discriminant axes failed to separate any of the cultivars where the mean posterior assignment probabilities were `r passp(dapc_host, "G122")`% (G122), `r passp(dapc_host, "Beryl")`% (Beryl), and `r passp(dapc_host, "Bunsi")`% (Bunsi).
-DAPC grouping by Region and Year used the first `r dapc_region_year$n.pca` principal components, representing `r signif(dapc_region_year$var*100, 3)`% of the total variance.
+DAPC grouping by Region and Year used the first `r dapc_region_year$n.pca` principal components, representing `r prettify(dapc_region_year$var)`% of the total variance.
 The Midwestern populations (NE, MI, CO, ND) did not appear to have any variation across time in contrast to WA, which showed a shift in population structure in the last year of sampling, 2008 (Fig.
@@
 
-<!-- 
- - Big Picture reminder of importance
-   + This is an extension of a previous work [@otto-hanson2011variation]
-   + white mold screening nurseries serve as good way to screen potential resistant lines, avoiding environmental effects
-   + It's unclear how genetic structure can influence
- - Critical Analysis of Major Findings
-   + This is the largest genetic analysis of *S. sclerotiorum* to date
-   + Similar to the findings of @otto-hanson2011variation, we found that
-     Region has a strong influence on both population structure and
-     aggressiveness.
-   + North Dakota appeared
- - Additional Findings and how they fit into the literature
- - Limitations
- - Future Directions
- - Overall Conclusion and Impact
--->
-
 In this study, we characterized the diversity of *Sclerotinia sclerotiorum* from dry bean fields across the United States.
@@
 To date, with 366 isolates, this is the largest single population genetic study of *S. sclerotiorum* assessing population structure within managed and unmanaged agricultural environments. 
-These findings indicate that the wmns can be effective at screening for potential resistant lines within growing regions.
+These findings indicate that the wmn can be effective at screening for potential resistant lines within growing regions.
 
@@
 From the initial results, it appears that the most important factors for predicting genetic structure are MCG, region, and year.
-When we inspected the biplot of the initial results, we saw that the most important predictors were MCGs 44, 5, and 9. 
+When we inspected the biplot of the initial results, we saw that the most important predictors were 'MCG 44', 'MCG 5', and 'MCG 9'. 
 We believe that this is driven by the fact that these particular MCGs have uneven MLG distributions, meaning that they are heavily associated with one particular MLG (Fig. \@ref(biggraph)).
 We note these results with caution because of the apparent multicolinearity between MCG and Region, which is a violation of the analysis [@legendre1999distance].
-While the results indicated that Mexico and the year 2005 were the two most important variables, it's worth noting that all Mexican isolates were collected in 2005 (Fig. \@ref(rda-biplot)).
+While the results indicated that Mexico and the year 2005 were the two most important variables, it's worth noting that all Mexico isolates were collected in 2005 (Fig. \@ref(rda-biplot)).
 The results also show that the Vista cultivar explains some of the variance, but this represents six isolates in MI. 
@@
 When we analyze the AMOVA results in light of the DAPC results (Fig. \@ref(DAPC)), it becomes clear that the regional patterns of differentiation are largely driven by isolates from Mexico and CA.
-Isolates from these Regions had a higher posterior probability (> 0.75) of being reassigned to their own populations than any other (Fig. \@ref(DAPC)a). 
+Isolates from these Regions had a higher posterior probability (> 0.75) of being reassigned to their own populations than any other (Fig. \@ref(DAPC)A). 
 All other populations in comparison (except France) has reassignment probabilities of < 0.5, which is reflected in the failure of the first two discriminant functions to separate these populations (Fig.
-\@ref(DAPC)b).
+\@ref(DAPC)B).
 
-Despite the evidence that Mexico and CA contributing to a large amount of the population differentiation, Region like WA still had a large amount of internal variation (Supplementary Information).
+Despite the evidence that Mexico and CA contributing to a large amount of the population differentiation, Regions like WA still had a large amount of internal variation (Supplementary Information).
 The two distinct clusters for the WA Region showed that the 2008 population appeared different and, under further investigation, we found that the genotypes from this population were shared between CA and France (Fig.
-\@ref(DAPC-RY), \@ref(DAPC-RY-FULL), \@ref(community-graph)).
-From the DAPC, we extracted 5 loci that contributed >10% of the observed variation to either of the first two discriminant axes and found that the alleles with an allele frequency of >50% in 2007 were reduced to < 10% in 2008, replaced by an allele that was > 2 mutational steps away (Supplementary Information).
+\@ref(community-graph), \@ref(DAPC-RY), \@ref(DAPC-RY-FULL)).
+From the DAPC, we extracted 6 loci that had alleles contributing more than the upper 95% quantile of the principal component loadings to either of the first two discriminant axes and found that the alleles with an allele frequency of > 50% in 2007 were reduced to < 10% in 2008, replaced by an allele that was > 2 mutational steps away (Supplementary Information).
 This could be due to several reasons which include mis-labeling of source material or introduction of migrant sclerotia in plant material.
@@
 Unlike Bruvo's distance, Euclidean distance does not take into account the magnitude of the difference between alleles, which could inflate the distance measure in the presence of private alleles [@bruvo2004simple].
-We found no evidence of private alleles in the Mexican and Californian isolates, suggesting that this pattern separating these populations from the others was not an artifact.
+We found no evidence of private alleles in the Mexico and CA isolates, suggesting that this pattern separating these populations from the others was not an artifact.
 
@@
 This has been attributed to the notion that the fungus has the chance to undergo more reproductive cycles in the warmer climate [@carbone2001multilocus; @attanayake2013sclerotinia]. 
-Populations in sub-tropical regions of China have been found to be more variable, sexually reproducing, and unrelated to populations in temperate regions of the United States [@attanayake2013sclerotinia].
-The differentiation between the Mexican isolates and all other regions gives some evidence for this hypothesis as the only isolates in our study from sub-tropical regions are in the Mexican populations.
+Populations in sub-tropical regions of China have been found to be more variable, sexually reproducing, and unrelated to populations in temperate regions of the USA [@attanayake2013sclerotinia].
+The differentiation between the Mexico isolates and all other regions gives some evidence for this hypothesis as the only isolates in our study from sub-tropical regions are in Mexico.
 
-We were unable to find evidence for sexual reproduction in any region. 
-While the Australian population had a non-significant value of $\bar{r}_d$---which would suggest that we cannot reject the null hypothesis of random mating---the sample size was insufficient from which to draw conclusions [@agapow2001indices; @milgroom1996recombination].
+We were unable to find evidence for sexual reproduction in any region except for CA and Australia. 
+While the Australia population had a non-significant value of $\bar{r}_d$---which would suggest that we cannot reject the null hypothesis of random mating---the sample size was insufficient from which to draw conclusions [@agapow2001indices; @milgroom1996recombination].
+The low value of $\bar{r}_d$ in the CA population may represent sexual reproduction, but this could also be an artifact of sampling two different populations, which is known to reduce the value of $\bar{r}_d$ [@prugnolle2010apparent].
 
@@
 While we found no direct relationship between genotype and severity, it is evident that there is a gradient of severity by region, further supporting the need for screening in multiple locations.
+
+## Data Availability {-}
+
+All scripts, data, and resources used to generate the results presented in this publication (including supplementary information) are fully reproducible and available at https://github.com/everhartlab/sclerotinia-366/ and archived at ZENODO.
 

### 2017-08-21 doc/manuscript/manuscript.Rmd
Attempt at expanding introduction

@@
 
-*Sclerotinia sclerotiorum* (Lib.) de Bary is an ascomycete plant pathogen found worldwide.
-This is a necrotrophic pathogen that is primarily homothallic (self-fertilization) [@bolton2006sclerotinia; @sexton2006population] that causes disease on more than 400 plant species belonging to 75 families [@boland1994index].
-These include crops of major economic importance such as sunflower (*Helianthus spp.*), soybean (*Glycine max* L.), canola (*Brassica napa* L., *Brassica campestris* L.), and dry bean (*Phaseolus vulgaris* L.) [@bolton2006sclerotinia].
-On dry bean, *S. sclerotiorum* is the causal agent of White Mold, a devastating disease that can be a yield-limiting in temperate climates [@steadman1983white].
+*Sclerotinia sclerotiorum* (Lib.) de Bary is an ascomycete plant pathogen with a worldwide distribution.
+This is a necrotrophic pathogen that is primarily homothallic (self-fertilization) and has the ability to survive fore more than five years in soil using melanized survival structures called sclerotia [@bolton2006sclerotinia; @sexton2006population].
+It causes disease on more than 400 plant species belonging to 75 families [@boland1994index] including crops of major economic importance such as sunflower (*Helianthus spp.*), soybean (*Glycine max* L.), canola (*Brassica napa* L., *Brassica campestris* L.), and dry bean (*Phaseolus vulgaris* L.) [@bolton2006sclerotinia].
+On dry bean, *S. sclerotiorum* is the causal agent of white mold, a devastating disease that can be a yield-limiting in temperate climates [@steadman1983white].
+
+Currently, there are no known resistant cultivars of dry bean commercially available [@otto-hanson2011variation].
+Organized breeding efforts have been using a common-garden approach with white mold screening nurseries in dry bean production areas across the United States with additional sites in Mexico, France, and Australia [@steadman2003identification; @steadman2004identification; @steadman2005identification; @steadman2006identification; @otto2007identification].
+These white mold screening nurseries use no chemical or cultural treatments against *S. sclerotiorum* and employ standardized protocols for screening new cultivars for resistance to white mold. 
+It has previously been shown that aggressiveness (the severity of disease symptoms on the host) is significantly different across white mold screening nursery sites in separate geographic regions [@otto-hanson2011variation].
+The genetic structure and mode of reproduction in these populations, however, is currently unknown.
+Understanding genetic relationships and reproduction behavior of *S.
+sclerotiorum* populations is beneficial for breeders seeking to develop new resistant cultivars for worldwide deployment [@mcdonald2002pathogen; @milgroom1996recombination].
+
+Most populations of *S.
+sclerotiorum* are predominantly clonal and have a large degree of population fragmentation [@attanayake2012genetic; @carbone2001multilocus; @cubeta1997clonality; @ekins2011population; @kohli1998random; @kohli1995clonal].
+Some populations are sexual [@atallah2004high; @attanayake2013sclerotinia; @sexton2004microsatellite]....
+
+Nearly all population genetic studies of *S.
+sclerotiorum* employ a macroscopic assay to determine mycelial compatibility, the ability for fungal hyphae from different colonies to appear to grow together without forming a barrier of dead cells between them (known as a barrage line) [@leslie1993fungal; @sirjusingh2001characterisation].
+Mycelial compatibility has been used as a proxy for vegetative compatibility, a fungal trait controlled by several independent genes controlling the ability for two hyphae to fuse and grow as a single unit [@leslie1993fungal; @schafer2006optimized].
+Because of the genetic connection to vegetative compatibility, two isolates that are mycelially compatible have been considered to be clones, but correlation with genetic markers such as microsatellites have shown mixed results [@attanayake2012genetic; @lehner2017independently; @leslie1993fungal].
 
@@
 
-Previous studies of *S.
+<!-- Previous studies of *S.
 sclerotiorum* population biology in North America and Australia have reported clonal propagation of the pathogen [@cubeta1997clonality; @ekins2011population; @kohli1998random; @kohli1995clonal] with evidence of outcrossing observed in some areas [@atallah2004high; @kohli1998random; @malvarez2007new; @sexton2004microsatellite].
-One such study examined *S. sclerotiorum* population structure using microsatellite markers from potato fields in the Columbia Basin of Washington state.
-Due to high crop diversity and high presence of the pathogen, it was important to understand the potential for outcrossing in order to develop disease management strategies.
-Molecular analysis revealed 92% of the isolate variability was found within subpopulations [@atallah2004high].
-However, there was evidence of linkage equilibrium of microsatellite loci in 2 locations and illustrated the possibility of outcrossing in this region.
+-->
+<!-- One such study examined *S. sclerotiorum* population structure using microsatellite markers from potato fields in the Columbia Basin of Washington state. -->
+<!-- Due to high crop diversity and high presence of the pathogen, it was important to understand the potential for outcrossing in order to develop disease management strategies. -->
+<!-- Molecular analysis revealed 92% of the isolate variability was found within subpopulations [@atallah2004high]. -->
+<!-- However, there was evidence of linkage equilibrium of microsatellite loci in 2 locations and illustrated the possibility of outcrossing in this region.   -->
 
@@
 <!-- SEE: Avoid sentences that start with the authors/reference. -->
-@attanayake2013sclerotinia studied population differentiation of *S. sclerotiorum* isolates infecting canola from China and the USA.
-There were no shared microsatellite multilocus genotypes (MLG) between the two populations.
-Sexual outcrossing was detected in both populations with higher recombination rate in Chinese population than the USA population.
-Additionally, phenotypic traits such as production of sclerotia, mycelial compatibility group (MCG), growth rate were also significantly different between the two populations.
-Genetic diversity of *S. sclerotiorum* from soy bean, canola, sunflower, and dry bean in the North Central region of the USA was studied by [@aldrich-wolfe2015genetic].
-Microsatellite analysis revealed little host- specificity or genetic-isolation of *S.
+<!-- @attanayake2013sclerotinia studied population differentiation of *S. sclerotiorum* isolates infecting canola from China and the USA. -->
+<!-- There were no shared microsatellite multilocus genotypes (MLG) between the two populations. -->
+<!-- Sexual outcrossing was detected in both populations with higher recombination rate in Chinese population than the USA population. -->
+<!-- Additionally, phenotypic traits such as production of sclerotia, mycelial compatibility group (MCG), growth rate were also significantly different between the two populations. -->
+<!-- Genetic diversity of *S. sclerotiorum* from soy bean, canola, sunflower, and dry bean in the North Central region of the USA was studied by [@aldrich-wolfe2015genetic]. -->
+<!-- Microsatellite analysis revealed little host- specificity or genetic-isolation of *S.
 sclerotiorum* isolates among host crops as 93% of the observed genetic variation was explained by within-host differences.
-Isolates from all host crops were primarily reproducing clonally while there were also evidence of genetic recombination of isolates from canola and sunflower.
+-->
+<!-- Isolates from all host crops were primarily reproducing clonally while there were also evidence of genetic recombination of isolates from canola and sunflower. -->
 
@@
 Canopy architecture and growth pattern of some cultivars enable improved disease avoidance [@kolkman2002agronomic]. -->
-To date, no commercial white mold-resistant dry bean cultivars exist. 
-Field trials in some regions of the USA have shown the G122 cultivar is partially resistant while the Beryl cultivar is susceptible [@otto-hanson2011variation; @steadman2004identification].
+<!-- To date, no commercial white mold-resistant dry bean cultivars exist.  -->
+<!-- Field trials in some regions of the USA have shown the G122 cultivar is partially resistant while the Beryl cultivar is susceptible [@otto-hanson2011variation; @steadman2004identification]. -->
 <!-- This puny paragraph needs a lot of work to more fully describe the operation of the wmn -->
-
-Understanding genetic relationships and reproduction behavior of *S.
-sclerotiorum* populations from locations worldwide is beneficial for plant breeders seeking to develop white mold resistance cultivars for deployment worldwide.
-This information is also helpful in making disease management decisions and developing disease resistance screening methods.
 

### 2017-08-21 doc/manuscript/manuscript.Rmd
Update introduction to include info about popgen

@@
 The genetic structure and mode of reproduction in these populations, however, is currently unknown.
+
 Understanding genetic relationships and reproduction behavior of *S.
 sclerotiorum* populations is beneficial for breeders seeking to develop new resistant cultivars for worldwide deployment [@mcdonald2002pathogen; @milgroom1996recombination].
-
+In particular, genetically diverse populations with high rates of sexual reproduction are more likely to overcome host resistance. 
 Most populations of *S.
-sclerotiorum* are predominantly clonal and have a large degree of population fragmentation [@attanayake2012genetic; @carbone2001multilocus; @cubeta1997clonality; @ekins2011population; @kohli1998random; @kohli1995clonal].
-Some populations are sexual [@atallah2004high; @attanayake2013sclerotinia; @sexton2004microsatellite]....
+sclerotiorum* are predominantly clonal with low genetic diversity and have a large degree of population fragmentation [@attanayake2012genetic; @carbone2001multilocus; @cubeta1997clonality; @ekins2011population; @kohli1998random; @kohli1995clonal].
+Some studies, however have found populations that show signatures of sexual reproduction [@atallah2004high; @attanayake2013sclerotinia; @sexton2004microsatellite; @aldrich-wolfe2015genetic].
 
@@
 Because of the genetic connection to vegetative compatibility, two isolates that are mycelially compatible have been considered to be clones, but correlation with genetic markers such as microsatellites have shown mixed results [@attanayake2012genetic; @lehner2017independently; @leslie1993fungal].
-
-<!-- ZNK: TMI -->
-<!-- Sclerotia are capable of surviving more than five years in soil [@steadman1983white] and germinate carpogenically or myceliogenically to infect the plant when environmental conditions are favorable.
-A carpogenically germinated sclerotia gives rise to apothecia in which ascospores are formed.
-Ascospores infect senescent plant tissues that, when supplemented with cool and humid micro-climate within the plant canopy, leads to mycelial infection of the host plant.
-Consequently, flowering, which coincide with canopy closure is a critical factor of disease initiation since sensing flower parts fallen on to leaves, petioles or stems provide the required energy source for plant infection.
-Myceliogenically germinated sclerotia result in formation of hyphae, which also have potential for directly plant infection, though this is considered rare and less important within the disease cycle of the pathogen [@abawi1975source].
-Diseased bean fields initially show scattered wilted leaves.
-Symptoms on individual plants are water-soaked lesions on leaves, pods and stems, which expand over time and show signs of white fungal mycelia.
-Lesions on stems and branches become necrotic as they age and dead tissues appear bleached.
-Sclerotia are often produced on dead or dying plants once conditions become unfavorable for fungal growth. -->
-
-<!-- Previous studies of *S.
-sclerotiorum* population biology in North America and Australia have reported clonal propagation of the pathogen [@cubeta1997clonality; @ekins2011population; @kohli1998random; @kohli1995clonal] with evidence of outcrossing observed in some areas [@atallah2004high; @kohli1998random; @malvarez2007new; @sexton2004microsatellite].
--->
-<!-- One such study examined *S. sclerotiorum* population structure using microsatellite markers from potato fields in the Columbia Basin of Washington state. -->
-<!-- Due to high crop diversity and high presence of the pathogen, it was important to understand the potential for outcrossing in order to develop disease management strategies. -->
-<!-- Molecular analysis revealed 92% of the isolate variability was found within subpopulations [@atallah2004high]. -->
-<!-- However, there was evidence of linkage equilibrium of microsatellite loci in 2 locations and illustrated the possibility of outcrossing in this region.   -->
-
-<!-- ZNK: This is more discussion-material -->
-<!-- A population study from Australian canola fields showed evidence of sexual outcrossing of *S. sclerotiorum* isolates [@sexton2004microsatellite].
-Another study characterized three populations of *S. sclerotiorum* from lettuce in California (CA) and Ontario (ON), and pea/lentil in Washington (WA) [@malvarez2007new].
-While all three populations were  clonal, the index of association ($I_A$) based on DNA fingerprint was closer to zero for CA than it was for the other populations.
-Populations from Australian sunflower with evidence of clonal reproduction were reported, and despite populations sampled from fields separated 50 to 400 km and from both head and basal rots, all showed similar gene diversity, with authors concluding isolates representing different geographies belong to one large genetic population of *S.
-sclerotiorum* [@ekins2011population].
--->
-
-<!-- SEE: Avoid sentences that start with the authors/reference. -->
-<!-- @attanayake2013sclerotinia studied population differentiation of *S. sclerotiorum* isolates infecting canola from China and the USA. -->
-<!-- There were no shared microsatellite multilocus genotypes (MLG) between the two populations. -->
-<!-- Sexual outcrossing was detected in both populations with higher recombination rate in Chinese population than the USA population. -->
-<!-- Additionally, phenotypic traits such as production of sclerotia, mycelial compatibility group (MCG), growth rate were also significantly different between the two populations. -->
-<!-- Genetic diversity of *S. sclerotiorum* from soy bean, canola, sunflower, and dry bean in the North Central region of the USA was studied by [@aldrich-wolfe2015genetic]. -->
-<!-- Microsatellite analysis revealed little host- specificity or genetic-isolation of *S.
-sclerotiorum* isolates among host crops as 93% of the observed genetic variation was explained by within-host differences.
--->
-<!-- Isolates from all host crops were primarily reproducing clonally while there were also evidence of genetic recombination of isolates from canola and sunflower. -->
-
-<!-- Most of the previous studies on *S. sclerotiorum* had been conducted in the USA, Canada and Australia but not in Europe. 
-In those studies, different dry bean cultivars had been rarely used as  hosts to investigate any host-pathogen relationships or specificity.
-No white mold resistant dry bean cultivars exist.
-Canopy architecture and growth pattern of some cultivars enable improved disease avoidance [@kolkman2002agronomic]. -->
-<!-- To date, no commercial white mold-resistant dry bean cultivars exist.  -->
-<!-- Field trials in some regions of the USA have shown the G122 cultivar is partially resistant while the Beryl cultivar is susceptible [@otto-hanson2011variation; @steadman2004identification]. -->
-<!-- This puny paragraph needs a lot of work to more fully describe the operation of the wmn -->
-
-<!-- Microsatellite markers (simple sequence repeats or SSR) are codominant, polymorphic and also able to detect multiple alleles per locus [@atallah2004high; @sirjusingh2001characterisation].
-Reproducibility of microsatellites is higher compared other fingerprinting techniques such as random amplified polymorphic DNA (RAPD) [@atallah2004high].
-High mutagenic rate and multiallelic nature of microsatellites render suitability for inferring population structure of organisms and their potential for outcrossing.
-In the case of plant pathogens like *S.
-sclerotiorum* that is predominantly clonal and occasionally outcrossing, microsatellites have been used successfully to ascertain their ability to reproduce by sexual outcrossing [@atallah2004high; @sexton2004microsatellite] and also to detect differences among clonally propagated individuals.
--->
 

### 2017-08-21 doc/manuscript/manuscript.Rmd
Rearrange figures and tables

@@
 library("poppr")
+library("tidyverse")
 load("../../data/sclerotinia_16_loci.rda")
@@
 dat11cc <- clonecorrect(dat11, ~Region/Source/Host/Year)
+pop_size <- table(pop(dat11))
 
-dapc_region      <- dapc(setPop(dat11cc, ~Region), n.da = 13, n.pca = 20)
-dapc_region_year <- dapc(setPop(dat11cc, ~Region/Year), n.da = 12, n.pca = 22)
-dapc_host        <- dat11cc %>% 
-  setPop(~Source) %>%
-  popsub("wmn") %>%
-  setPop(~Host) %>% 
-  popsub(c("Beryl", "Bunsi", "G122")) %>%
-  dapc(n.da = 2, n.pca = 20)
+# Utility Functions -------------------------------------------------------
 
@@
 }
+
+# Locus Table -------------------------------------------------------------
+
+loc_columns <- cols(
+  Locus = col_character(),
+  Range = col_character(),
+  Repeats = col_double(),
+  allele = col_integer(),
+  Hexp = col_double(),
+  Evenness = col_double()
+)
+loc_table <- readr::read_csv("../../results/tables/locus-table.csv", 
+                             col_types = loc_columns,
+                             col_names = TRUE)
+Amean <- mean(loc_table$allele)
+Hmean <- mean(loc_table$Hexp)
+Emean <- mean(loc_table$Evenness)
+
+# DAPC data ---------------------------------------------------------------
+
+dapc_region      <- dapc(setPop(dat11cc, ~Region), n.da = 13, n.pca = 20)
+dapc_region_year <- dapc(setPop(dat11cc, ~Region/Year), n.da = 12, n.pca = 22)
+dapc_host        <- dat11cc %>% 
+  setPop(~Source) %>%
+  popsub("wmn") %>%
+  setPop(~Host) %>% 
+  popsub(c("Beryl", "Bunsi", "G122")) %>%
+  dapc(n.da = 2, n.pca = 20)
+
+# RDA results -------------------------------------------------------------
+
 RDA2R2    <- prettify(vegan::RsquareAdj(cap11cc2)$adj.r.squared)/100
 RDA1R2    <- prettify(vegan::RsquareAdj(cap11cc)$adj.r.squared)/100
-
 anova1res <- setNames(cap11anova$`Pr(>F)`, rownames(cap11anova))
@@
 A total of 366 isolates were collected from 2003 to 2012 (except 2006 and 2011) from diseased dry bean plants in Australia, Mexico, France, and 11 states in the USA (Table \@ref(tab:isolate-table)).
-Over 50% of the isolates came from four states, MI (62), ND (60), WA (59), NE (47). 
-Four regions had fewer than 10 isolates, Australia (6), WI (2), NY (1), ID (1).
+Over 50% of the isolates came from four states, MI (`r pop_size[["MI"]]`), ND (`r pop_size[["ND"]]`), WA (`r pop_size[["WA"]]`), NE (`r pop_size[["NE"]]`). 
+Four regions had fewer than 10 isolates, Australia (`r pop_size[["Australia"]]`), WI (`r pop_size[["WI"]]`), NY (`r pop_size[["NY"]]`), ID (`r pop_size[["ID"]]`).
 We observed 87 MCGs, the most abundant of which ('MCG 5') was represented by 73 isolates over 37 MLGs (Fig \@ref(biggraph)A,C).
-After clone-correction on the hierarchy of Region/Source/Host/Year, we were left with 318 isolates representing 165 MLGs (Table \@ref(tab:iatab)).
 Analysis of the genotype accumulation curve showed no clear plateau (Supplemental Information), indicating that we would likely obtain more multilocus genotypes if we were to genotype more loci.
-The number of observed alleles per locus ranged from two to 10 with an average of 6.27 (Table \@ref(tab:locus-stats)).
-
------------------------------------------------------------------------------
-Country   State  Field Code              Year Host                          N 
---------- ------ ---------- ----------------- -------------------------- ----
-USA       CA     wmn               2004, 2005 Beryl, Bunsi, G122           18
-
-USA       CO     producer          2007, 2010 Pinto, Yellow                41
-
-                 wmn                     2003 GH                            1
-
-USA       ID     producer                2003 GH                            1
-
-USA       MI     wmn        2003, 2004, 2005, 11A, 37, 38, B07104,         43 
-                                   2008, 2009 Beryl, Bunsi, Cornell,       
-                                              G122, Orion, PO7863, WM31
-
-                 producer    2003, 2008, 2009 BL, Black, Fuji, GH,         19 
-                                              Merlot, SR06233, unk,        
-                                              Vista, Zorro
-
-USA       MN     wmn               2003, 2004 Beryl, Bunsi, G122           11
-
-USA       ND     producer          2007, 2010 unk                          53
-
-                 wmn                     2005 Beryl, Bunsi, G122            7
-
-USA       NE     wmn        2004, 2005, 2008, Beryl, Bunsi, G122,          27 
-                                         2010 PO7683, unk
-
-                 producer   2003, 2007, 2009, Beryl, Emerson, GH, Orion,   20 
-                                         2010  Pinto, Weihing
-
-USA       NY     producer                2003 GH                            1
-
-USA       OR     wmn               2003, 2004 Beryl, Bunsi, G122           15
-
-                 producer                2003 G122, GH                      2
-
-USA       WA     wmn        2003, 2004, 2005, 11A, 37, 38, Beryl, Bunsi,   36 
-                                         2008  Cornell, G122, Orion, PO7      
-                                              104, PO7863, WM31
-
-                 producer          2003, 2007 GH, Merlot, Pinto, Redkid    23
-
-USA       WI     producer                2003 GH                            2
-
-Mexico    -      wmn                     2005 Beryl, Bunsi, G122           18
-
-France    -      wmn               2004, 2005 Beryl, Bunsi, G122           18
-
-                 producer                2012 unk                           4
-
-Australia -      wmn                     2004 Beryl, Bunsi, G122            4
-
-                 producer                2004 Beryl                         2
-
------------------------------------------------------------------------------
-
-Table: (\#tab:isolate-table) Description of *Sclerotinia sclerotiorum* isolates used in this study.
-N = Number of Isolates. 
-Key abbreviations: wmn = white mold screening nursery, producer = producer field, unk = unknown cultivar.
+The number of observed alleles per locus ranged from two to 10 with an average of `r signif(Amean, 3)` (Table \@ref(tab:locus-stats)).
+After clone-correction on the hierarchy of Region/Source/Host/Year, we were left with `r nInd(dat11cc)` isolates representing `r nmll(dat11cc)` MLGs (Table \@ref(tab:iatab)).
 
 ```{r , echo = FALSE, warning = FALSE, message = FALSE, results = "asis"}
-suppressPackageStartupMessages(library("tidyverse"))
-loc_columns <- cols(
-  Locus = col_character(),
-  Range = col_character(),
-  Repeats = col_double(),
-  allele = col_integer(),
-  Hexp = col_double(),
-  Evenness = col_double()
-)
-loc_table <- readr::read_csv("../../results/tables/locus-table.csv", 
-                             col_types = loc_columns,
-                             col_names = TRUE)
 loc_table %>%
@@
   huxtable::print_md(max_width = 61)
-
-Amean <- mean(loc_table$allele)
-Hmean <- mean(loc_table$Hexp)
-Emean <- mean(loc_table$Evenness)
 ```
@@
 alleles = `r signif(Amean, 3)`
-
-## Variable Assessment {-}
-
-### Variable Importance {-}
-
-The forward-backward selection process of the dbRDA models on clone-corrected data revealed Year, Region, Host, and Aggressiveness to be the optimal variables for the reduced model, accounting for `r prettify(vartab(cap11sum)[2, "Proportion"])`% of the total variation.
-ANOVA revealed that the reduced model was significant with an adjusted R^2^ of `r RDA1R2` (*P* = 0.001).
-Assessment of the marginal effects showed that region, host, year, and aggressiveness significantly explained genetic variation (*P* = `r anova1res[["Year"]]`, *P* = `r anova1res[["Region"]]` for region).
-It was revealed that there was multicollinearity when MCG was combined with any other variable, so we followed the same procedure, dropping MCG from the list of potential predictors. 
-From these results, we found that aggressiveness, year, region, and host were optimal, accounting for `r prettify(vartab(cap11sum2)[2, "Proportion"])`% of the total variation.
-ANOVA revealed significant effects with an adjusted R^2^ of `r RDA2R2` (*P* = 0.001).
-While the marginal effect assessment showed that region, host, and year all significantly explained the variation at *P* = `r anova2res[["Year"]]`, and aggressiveness significantly explained the variation at *P* = `r anova2res[["Severity"]]`.
-Much of the variation appears to be driven by isolates from Mexico and 2005 (Fig. \@ref(rda-biplot)).
-Variance partitioning of the independent variables without MCG indicated aggressiveness to be the least influential factor with 0.1% contributing to explaining the variation of molecular data whereas the combination of variables accounted for 3.3%.
-
-<!-- RDA figure -->
-
-![
-Biplot showing five most influential explanatory variables (arrows) overlayed on the first two eigenvectors of distance based redundancy analysis of *S. sclerotiorum* isolates. 
-The length of the arrows are directly proportional to the strength of the correlation between explanatory and molecular variables. 
-Open circles represent the 318 clone-corrected samples in ordination space. 
-](../../results/figures/publication/Figure7Z.pdf){#rda-biplot width=61%}
-
-<!-- end figure -->
-
-### Aggressiveness {-}
-
-<!-- aggressiveness figure -->
-
-![
-Boxplot representing aggressiveness (vertical axis) by population (horizontal axis) arranged in descending order of mean aggressiveness for all populations with N > 5.
-Box boundaries represent 25% and 75% percentiles and whiskers extend to at most 1.5 $\times$ the interquartile range.
-Horizontal line represents median value.
-Circles represent individual observations where filled circles indicate isolates from white mold screening nurseries (wmn).
-](../../results/figures/publication/aggressiveness.pdf){#aggressiveness-boxplot}
-
-<!-- end figure -->
-
-Our assessment of aggressiveness in association with Region showed a significant effect (P < 1.00e^-4^), with means that ranged from 5.8 (MN) to 4.0 (CA) (Fig.
-\@ref(aggressiveness-boxplot), Table \@ref(tab:region-aggressiveness)).
-MCGs also showed a significant effect (P < 1.00e^-4^), with means that ranged from 6.0 ('MCG 44') to 4.6 ('MCG 49').
-We additionally found a significant effect for MLGs (*P* = 7.44e^-4^) for MLGs, with means that ranged from 6.0 ('MLG 78') to 4.3 ('MLG 140').
-
-### Correlation of Mulitlocus Genotypes and Mycelial Compatibility Groups {-}
-
-Our network-based approach to correlating MLGs with MCGs revealed a large and complex network (Fig. \@ref(biggraph)). 
-Community analysis showed 51 communities, 15 of which consisted of a single MLG unconnected with any other community indicating that just 9.09% of the 165 MLGs are unable to cross with any other MLG in this data set (Fig.
-\@ref(fullgraph)).
-The three communities with the most members contained eight of the 10 most abundant MCGs. 
-Comparing these communities with Bruvo's genetic distance revealed an average distance of 0.451 among communities and an average distance of 0.437 within communities, showing no significant difference.
-When we assessed the number of times two different MLGs that are in the same MCG, considering these as potential heterothallic pairings that could result in sexual recombination, we found an average of 16.5 potential heterothallic parings per MLG.
-Representing just four isolates, 'MLG 75' had 57 neighbors that shared the same mycelial compatibility group (Fig \@ref(biggraph), \@ref(fullgraph)). 
-Overall, there was no clear pattern to the association between MLG and MCGs.
-
-<!-- MCG graph figure -->
-
-![
-Data show complex associations between Mycelial Compatibility Groups and Multilocus Genotypes.
-**A)** Barplot of Mycelial Compatibility Group (MCG) abundance in descending order.
-Singletons (46) were truncated, leaving 41 MCGs.
-White bars represent sample counts and grey bars represent counts of unique multilocus genotypes (MLG).
-The transparency of the bars represent the evenness of the distribution of the MLGs within a given MCG.
-A dashed box surrounds the 8 most common MCGs representing > 51% of the data.
-**B)** Full graph-representation of the relationship between MCGs (open circles) and MLGs (filled circles).
-Details in fig. \@ref(fullgraph). 
-**C)** A subset of **B** representing the 8 most common MCGs and their associated MLGs (dashed box in **A**).
-Filled nodes represent MLGs and open nodes represent MCGs.
-Numbers inside nodes are the MLG/MCG label. 
-Node (circle) area scaled to the number of samples represented (range: 1--73).
-Edges (arrows) point from MLG to MCG where the weight (thickness) of the edge represents the number of shared isolates (range: 1--19).
-Edges extending from MLGs displayed to other MCGs are not shown.
-](../../results/figures/publication/mcg-subgraph-with-context.pdf){#biggraph width=100%}
-
-<!-- end fig -->
-
-## Structure of Shared Multilocus Genotypes {-}
 
@@
 
+<!-- MCG graph figure -->
+
+![
+Data show complex associations between Mycelial Compatibility Groups and Multilocus Genotypes.
+**A)** Barplot of Mycelial Compatibility Group (MCG) abundance in descending order.
+Singletons (46) were truncated, leaving 41 MCGs.
+White bars represent sample counts and grey bars represent counts of unique multilocus genotypes (MLG).
+The transparency of the bars represent the evenness of the distribution of the MLGs within a given MCG.
+A dashed box surrounds the 8 most common MCGs representing > 51% of the data.
+**B)** Full graph-representation of the relationship between MCGs (open circles) and MLGs (filled circles).
+Details in fig. \@ref(fullgraph). 
+**C)** A subset of **B** representing the 8 most common MCGs and their associated MLGs (dashed box in **A**).
+Filled nodes represent MLGs and open nodes represent MCGs.
+Numbers inside nodes are the MLG/MCG label. 
+Node (circle) area scaled to the number of samples represented (range: 1--73).
+Edges (arrows) point from MLG to MCG where the weight (thickness) of the edge represents the number of shared isolates (range: 1--19).
+Edges extending from MLGs displayed to other MCGs are not shown.
+](../../results/figures/publication/mcg-subgraph-with-context.pdf){#biggraph width=100%}
+
+<!-- end fig -->
+
+## Variable Assessment {-}
+
+### Variable Importance {-}
+
+The forward-backward selection process of the dbRDA models on clone-corrected data revealed Year, Region, Host, and Aggressiveness to be the optimal variables for the reduced model, accounting for `r prettify(vartab(cap11sum)[2, "Proportion"])`% of the total variation.
+ANOVA revealed that the reduced model was significant with an adjusted R^2^ of `r RDA1R2` (*P* = 0.001).
+Assessment of the marginal effects showed that region, host, year, and aggressiveness significantly explained genetic variation (*P* = `r anova1res[["Year"]]`, *P* = `r anova1res[["Region"]]` for region).
+It was revealed that there was multicollinearity when MCG was combined with any other variable, so we followed the same procedure, dropping MCG from the list of potential predictors. 
+From these results, we found that aggressiveness, year, region, and host were optimal, accounting for `r prettify(vartab(cap11sum2)[2, "Proportion"])`% of the total variation.
+ANOVA revealed significant effects with an adjusted R^2^ of `r RDA2R2` (*P* = 0.001).
+While the marginal effect assessment showed that region, host, and year all significantly explained the variation at *P* = `r anova2res[["Year"]]`, and aggressiveness significantly explained the variation at *P* = `r anova2res[["Severity"]]`.
+Much of the variation appears to be driven by isolates from Mexico and 2005 (Fig. \@ref(rda-biplot)).
+Variance partitioning of the independent variables without MCG indicated aggressiveness to be the least influential factor with 0.1% contributing to explaining the variation of molecular data whereas the combination of variables accounted for 3.3%.
+
+<!-- RDA figure -->
+
+![
+Biplot showing five most influential explanatory variables (arrows) overlayed on the first two eigenvectors of distance based redundancy analysis of *S. sclerotiorum* isolates. 
+The length of the arrows are directly proportional to the strength of the correlation between explanatory and molecular variables. 
+Open circles represent the 318 clone-corrected samples in ordination space. 
+](../../results/figures/publication/Figure7Z.pdf){#rda-biplot width=61%}
+
+<!-- end figure -->
+
+### Aggressiveness {-}
+
+<!-- aggressiveness figure -->
+
+![
+Boxplot representing aggressiveness (vertical axis) by population (horizontal axis) arranged in descending order of mean aggressiveness for all populations with N > 5.
+Box boundaries represent 25% and 75% percentiles and whiskers extend to at most 1.5 $\times$ the interquartile range.
+Horizontal line represents median value.
+Circles represent individual observations where filled circles indicate isolates from white mold screening nurseries (wmn).
+](../../results/figures/publication/aggressiveness.pdf){#aggressiveness-boxplot}
+
+<!-- end figure -->
+
+Our assessment of aggressiveness in association with Region showed a significant effect (P < 1.00e^-4^), with means that ranged from 5.8 (MN) to 4.0 (CA) (Fig.
+\@ref(aggressiveness-boxplot), Table \@ref(tab:region-aggressiveness)).
+MCGs also showed a significant effect (P < 1.00e^-4^), with means that ranged from 6.0 ('MCG 44') to 4.6 ('MCG 49').
+We additionally found a significant effect for MLGs (*P* = 7.44e^-4^) for MLGs, with means that ranged from 6.0 ('MLG 78') to 4.3 ('MLG 140').
+
+### Correlation of Mulitlocus Genotypes and Mycelial Compatibility Groups {-}
+
+Our network-based approach to correlating MLGs with MCGs revealed a large and complex network (Fig. \@ref(biggraph)). 
+Community analysis showed 51 communities, 15 of which consisted of a single MLG unconnected with any other community indicating that just 9.09% of the 165 MLGs are unable to cross with any other MLG in this data set (Fig.
+\@ref(fullgraph)).
+The three communities with the most members contained eight of the 10 most abundant MCGs. 
+Comparing these communities with Bruvo's genetic distance revealed an average distance of 0.451 among communities and an average distance of 0.437 within communities, showing no significant difference.
+When we assessed the number of times two different MLGs that are in the same MCG, considering these as potential heterothallic pairings that could result in sexual recombination, we found an average of 16.5 potential heterothallic parings per MLG.
+Representing just four isolates, 'MLG 75' had 57 neighbors that shared the same mycelial compatibility group (Fig \@ref(biggraph), \@ref(fullgraph)). 
+Overall, there was no clear pattern to the association between MLG and MCGs.
+
+## Structure of Shared Multilocus Genotypes {-}
+
 We observed a total of 165 MLGs (215 with 16 loci).
@@
 
+<!-- community graph figure -->
+
+![
+Network of populations (nodes/circles) and their shared multilocus genotypes (MLG) (edges/lines) genotyped over 11 loci. 
+Each node is labeled with **name (number of MLGs shared/number of MLGs total).** 
+The shade and area of the nodes are proportional to the number of unique MLGs within the node and the inner nodes are proportional to the number of private MLGs to the region (bottom legend). 
+Each edge represents a single MLG where its thickness represents the number of populations that share the MLG and the shade represents the value of $P_{sex}$, or the probability of encountering that MLG from two independent meiotic events.
+](../../results/figures/publication/mlg-11.pdf){#community-graph width=75%}
+
+<!-- end fig -->
+
 ```{r table3, echo = FALSE, warning = FALSE, message = FALSE, results = "asis"}
-suppressPackageStartupMessages(library("tidyverse"))
 read_csv("../../results/tables/mlg_table.csv", 
@@
 
-<!-- community graph figure -->
-
-![
-Network of populations (nodes/circles) and their shared multilocus genotypes (MLG) (edges/lines) genotyped over 11 loci. 
-Each node is labeled with **name (number of MLGs shared/number of MLGs total).** 
-The shade and area of the nodes are proportional to the number of unique MLGs within the node and the inner nodes are proportional to the number of private MLGs to the region (bottom legend). 
-Each edge represents a single MLG where its thickness represents the number of populations that share the MLG and the shade represents the value of $P_{sex}$, or the probability of encountering that MLG from two independent meiotic events.
-](../../results/figures/publication/mlg-11.pdf){#community-graph width=100%}
-
-<!-- end fig -->
-
 ## Population Differentiation {-}
@@
 ```{r table4, echo = FALSE, warning = FALSE, message = FALSE, results = "asis"}
-suppressPackageStartupMessages(library("tidyverse"))
 read_csv("../../results/tables/AMOVA-year.csv", 
@@
 \@ref(DAPC)a).
-DAPC grouping by cultivar used the first 20 principal components, representing 89% of the total variance.
-The first two discriminant axes failed to separate any of the cultivars where the mean posterior assignment probabilities were `r passp(dapc_host, "G122")`% (G122), `r passp(dapc_host, "Beryl")`% (Beryl), and `r passp(dapc_host, "Bunsi")`% (Bunsi).
-DAPC grouping by Region and Year used the first `r dapc_region_year$n.pca` principal components, representing `r prettify(dapc_region_year$var)`% of the total variance.
-The Midwestern populations (NE, MI, CO, ND) did not appear to have any variation across time in contrast to WA, which showed a shift in population structure in the last year of sampling, 2008 (Fig.
-\@ref(DAPC-RY)).
-Further analysis of this population revealed that all 12 isolates in WA circa 2008 originated in a wmn and nine of these genotypes were shared with CA, and three were shared with France (Fig.
-\@ref(community-graph), \@ref(community-graph16)).
 
@@
 <!-- end figure -->
+
+DAPC grouping by cultivar used the first 20 principal components, representing 89% of the total variance.
+The first two discriminant axes (representing `r sum(paxis(dapc_host, 1:2))`% of the discriminatory power) failed to separate any of the cultivars where the mean posterior assignment probabilities were `r passp(dapc_host, "G122")`% (G122), `r passp(dapc_host, "Beryl")`% (Beryl), and `r passp(dapc_host, "Bunsi")`% (Bunsi).
+DAPC grouping by Region and Year used the first `r dapc_region_year$n.pca` principal components, representing `r prettify(dapc_region_year$var)`% of the total variance.
+The Midwestern populations (NE, MI, CO, ND) did not appear to have any variation across time in contrast to WA, which showed a shift in population structure in the last year of sampling, 2008 (Fig.
+\@ref(DAPC-RY)).
+Further analysis of this population revealed that all 12 isolates in WA circa 2008 originated in a wmn and nine of these genotypes were shared with CA, and three were shared with France (Fig.
+\@ref(community-graph), \@ref(community-graph16)).
 
@@
 
-<!-- Conceptualization: SEE. -->
-<!-- Formal analysis: BSA SEE. -->
-<!-- Funding acquisition: SEE. -->
-<!-- Investigation: BSA. -->
-<!-- Resources: SEE. -->
-<!-- Supervision: BSA. -->
-<!-- Visualization: BSA SEE. -->
-<!-- Writing - original draft: BSA. -->
-<!-- Writing - review & editing: SEE. -->
+ - **Zhian N. Kamvar** analyzed the data, contributed analysis tools, wrote the paper, prepared figures and tables, reviewed drafts of the paper.
 
-\newpage
+ - **Bimal Sajeewa Amadarasa** ...
+
+ - **Rachana Jhala** ...
+
+ - **Serena McCoy** ...
+
+ - **James Steadman** ...
+
+ - **Sydney E. Everhart** ...
+
 # Supplementary Information {-}
@@
 \renewcommand{\thefigure}{S\arabic{figure}}
+
+-----------------------------------------------------------------------------
+Country   State  Field Code              Year Host                          N 
+--------- ------ ---------- ----------------- -------------------------- ----
+USA       CA     wmn               2004, 2005 Beryl, Bunsi, G122           18
+
+USA       CO     producer          2007, 2010 Pinto, Yellow                41
+
+                 wmn                     2003 GH                            1
+
+USA       ID     producer                2003 GH                            1
+
+USA       MI     wmn        2003, 2004, 2005, 11A, 37, 38, B07104,         43 
+                                   2008, 2009 Beryl, Bunsi, Cornell,       
+                                              G122, Orion, PO7863, WM31
+
+                 producer    2003, 2008, 2009 BL, Black, Fuji, GH,         19 
+                                              Merlot, SR06233, unk,        
+                                              Vista, Zorro
+
+USA       MN     wmn               2003, 2004 Beryl, Bunsi, G122           11
+
+USA       ND     producer          2007, 2010 unk                          53
+
+                 wmn                     2005 Beryl, Bunsi, G122            7
+
+USA       NE     wmn        2004, 2005, 2008, Beryl, Bunsi, G122,          27 
+                                         2010 PO7683, unk
+
+                 producer   2003, 2007, 2009, Beryl, Emerson, GH, Orion,   20 
+                                         2010  Pinto, Weihing
+
+USA       NY     producer                2003 GH                            1
+
+USA       OR     wmn               2003, 2004 Beryl, Bunsi, G122           15
+
+                 producer                2003 G122, GH                      2
+
+USA       WA     wmn        2003, 2004, 2005, 11A, 37, 38, Beryl, Bunsi,   36 
+                                         2008  Cornell, G122, Orion, PO7      
+                                              104, PO7863, WM31
+
+                 producer          2003, 2007 GH, Merlot, Pinto, Redkid    23
+
+USA       WI     producer                2003 GH                            2
+
+Mexico    -      wmn                     2005 Beryl, Bunsi, G122           18
+
+France    -      wmn               2004, 2005 Beryl, Bunsi, G122           18
+
+                 producer                2012 unk                           4
+
+Australia -      wmn                     2004 Beryl, Bunsi, G122            4
+
+                 producer                2004 Beryl                         2
+
+-----------------------------------------------------------------------------
+
+Table: (\#tab:isolate-table) Description of *Sclerotinia sclerotiorum* isolates used in this study.
+N = Number of Isolates. 
+Key abbreviations: wmn = white mold screening nursery, producer = producer field, unk = unknown cultivar.
 

### 2017-08-22 doc/RMD/mlg-mcg.Rmd
update mlg-mcg to save the heterothallic pairings

@@
 in_reach <- ego(g, 2, nodes = which(grepl("Genotype", V(g)$type)), mode = "all") %>%
-  map_df(function(.x) as_tibble(list(MLG = list(names(.x[grepl("MLG", names(.x))])), 
+  map_df(function(.x){
+    as_tibble(
+      list(
+        MLG = list(names(.x[grepl("MLG", names(.x))][-1])), # -1 to remove self
         MCG = list(names(.x[!grepl("MLG", names(.x))])), 
@@
         )
-                                )) %>% 
+      )
+    }) %>% 
   mutate(N = lengths(MLG))
@@
 knitr::kable(in_reach_summary)
-ggplot(in_reach_summary, aes(x = N - 1, y = MLG, fill = E)) +
-  geom_segment(aes(yend = 0, xend = N-1)) +
+in_reach_plot <- ggplot(in_reach_summary, aes(x = N, y = MLG, fill = E)) +
+  geom_segment(aes(yend = 0, xend = N)) +
   geom_point(aes(size = MCG), pch = 21) +
@@
             )) +
-  xlab("Maximum Heterothallic Pairings") +
+  xlab("Potential Heterothallic Pairings") +
   ylab("Multilocus Genotypes")
+in_reach_plot
+if (!interactive()){
+  save(in_reach, in_reach_summary, in_reach_plot, file = file.path(PROJHOME, "data/heterothallic-pairings.rda"))
+}
 ```

### 2017-08-22 doc/RMD/MCG-virulence.Rmd
Clean out a lot of old results
There were extra figures and tables littering the
results directory due to various things including
misspelling. These have been cleared.
modify MCG-virulence to use write_csv in pipe

@@
 library("tidyverse")
+dir.create(file.path(PROJHOME, "results", "tables"))
 ```
@@
 grps <- agricolae::HSD.test(ANOVA, "MCG", alpha = 0.05)$groups
-readr::write_csv(grps, path = "results/tables/MCG-aggressiveness.csv", col_names = TRUE)
 
@@
   dplyr::rename(`Mean Aggressiveness` = means) %>%
+  readr::write_csv(path = "results/tables/MCG-aggressiveness.csv", col_names = TRUE) %>%
   huxtable::as_huxtable(add_colnames = TRUE) %>%
@@
   dplyr::rename(Group = M) %>%
-  dplyr::rename(`Mean Aggressiveness` = means)
-
-readr::write_csv(grps, path = "results/tables/Region-aggressiveness.csv", col_names = TRUE)
-
-grps %>%
+  dplyr::rename(`Mean Aggressiveness` = means) %>%
+  readr::write_csv(path = "results/tables/Region-aggressiveness.csv", col_names = TRUE) %>%
   huxtable::as_huxtable(add_colnames = TRUE) %>%
@@
   dplyr::rename(Group = M) %>%
-  dplyr::rename(`Mean Aggressiveness` = means)
-
-readr::write_csv(grps, path = "results/tables/MLG-aggressiveness.csv", col_names = TRUE)
-
-grps %>%
+  dplyr::rename(`Mean Aggressiveness` = means) %>%
+  readr::write_csv(path = "results/tables/MLG-aggressiveness.csv", col_names = TRUE) %>%
   huxtable::as_huxtable(add_colnames = TRUE) %>%

### 2017-08-22 doc/manuscript/manuscript.Rmd
Edits to manuscript
 - edited for clarity
 - automate some results
 - add two supplementary tables

@@
 load("../../data/RDA_results.rda")
+load("../../data/mcg-mlg-graph.rda")
+load("../../data/heterothallic-pairings.rda")
 dat11cc <- clonecorrect(dat11, ~Region/Source/Host/Year)
@@
 
+# Tabulate variance partitions
 vartab <- function(x){
@@
 }
+
+# Take a character vector and create a list in english (WITH OXFORD COMMA): 
+# e.g. "Juice, Eggs, and Toast"
+andlist <- function(x) paste( paste(x[-length(x)], collapse = ", "), x[length(x)], sep = ", and " )
 
@@
 RDA1R2    <- prettify(vegan::RsquareAdj(cap11cc)$adj.r.squared)/100
-anova1res <- setNames(cap11anova$`Pr(>F)`, rownames(cap11anova))
-anova2res <- setNames(cap11anova2$`Pr(>F)`, rownames(cap11anova2))
+anova1res <- setNames(cap11anova$`Pr(>F)`, rownames(cap11anova)) %>% na.omit()
+anova2res <- setNames(cap11anova2$`Pr(>F)`, rownames(cap11anova2)) %>% na.omit()
+RDAVAR1   <- andlist(names(cap11cc$terminfo$ordered))
+RDAVAR2   <- andlist(gsub("Severity", "Aggressiveness", names(cap11cc2$terminfo$ordered)))
+
+# MLG/MCG comparison
+
+big_genotype <- in_reach %>% filter(N == max(N))
 ```
@@
 
-Sclerotia of *S. sclerotiorum* were collected over several years from grower fields and/or wmn in 11 states of the United States of America, Mexico, France, and Australia.
-The white mold screening nurseries in different regions are indicated by wmn and all other isolates came from commercial fields (Table \@ref(tab:isolate-table)).
-
+Sclerotia of *S.
+sclerotiorum* were collected over several years from grower fields and/or wmn in 11 states of the United States of America, Mexico, France, and Australia (Table \@ref(tab:isolate-table)).
 After collection, the isolates were stored in petri plates lined with filter paper, then stored at $20^{\circ}$F or -$4^{\circ}$C. 
@@
 All analyses were performed on this data set unless otherwise mentioned.
-All data processing and analyses were performed in `r substr(R.version.string, 1, 15)` [@R] and are openly available and reproducible at https://github.com/everhartlab/sclerotinia-366/.
+All data processing and analyses were performed in `r R.version.string %>% strsplit("\\(") %>% map_chr(1)`[@R] and are openly available and reproducible at https://github.com/everhartlab/sclerotinia-366/.
 Of the 16 microsatellite loci genotyped, five included compound repeats, which made it challenging to accurately/confidently bin alleles into fragment sizes expected for each locus based on the described repeat motif.
@@
 We assessed the power of our 11 markers by generating a genotype accumulation curve in the R package *poppr* version `r packageVersion("poppr")`, looking for evidence of saturation, which would indicate that loci were sufficiently sampled to adequately represent the full set of genotypes [@arnaud2007standardizing; @kamvar2015novel].
-To avoid including isolates potentially collected from the same plant (which increases the probability of collecting sclerotioa from the same point of infection more than once), data were clone-corrected on a hierarchy of region/source/host/year---meaning that one representative of each genotype was retained for each combined regions, source, host, and year---for subsequent analysis.
+To avoid including isolates potentially collected from the same plant (which increases the probability of collecting sclerotioa from the same point of infection more than once), data were clone-corrected on a hierarchy of Region/Source/Host/Year---meaning that one representative of each genotype was retained for each nested combination of Region, Source, Host, and Year---for subsequent analysis.
 We assessed genotypic diversity by calculating Stoddart and Taylor's index (*G*) [@stoddart1988genotypic], Shannon's index (*H*) [@shannon2001mathematical], Simpson's index ($\lambda$) [@simpson1949measurement], and evenness ($E_5$, *eMLG*) [@heck1975explicit; @hurlbert1971nonconcept; @pielou1975ecological; @grunwald2003analysis].
@@
 Because aggressiveness measures differed with each isolate, aggressiveness first averaged across clone-corrected isolates.
-To identify explanatory variable(s) correlated with genetic variation, a forward-backward selection process was applied with the vegan function `ordistep()` to identify the set of variables that explained more variation than predicted with the global model containing all variables.
-An ANOVA was then performed to test for significance of the reduced model and marginal effects using 999 permutations.
+To identify explanatory variable(s) correlated with genetic variation, a forward-backward selection process was applied with the vegan function `ordistep()`.
+An analysis of variance (ANOVA) was then performed to test for significance of the reduced model and marginal effects using 999 permutations.
 The `varpart()` function of vegan was used to determine variation partitioning of explanatory variables.
@@
 
-We used analysis of variance (ANOVA) to assess if aggressiveness could be explained by Region, MCG, or multilocus genotype (MLG). 
+We used ANOVA to assess if aggressiveness could be explained by Region, MCG, or multilocus genotype (MLG). 
 To minimize complications due to small sample sizes, we chose the top 10 MCGs, representing 56.5% of the isolates collected, the top 10 MLGs representing 26.7% of the isolates, and populations with more than five isolates.
@@
 
-The forward-backward selection process of the dbRDA models on clone-corrected data revealed Year, Region, Host, and Aggressiveness to be the optimal variables for the reduced model, accounting for `r prettify(vartab(cap11sum)[2, "Proportion"])`% of the total variation.
+The forward-backward selection process of the dbRDA models on clone-corrected data revealed `r RDAVAR1` to be the optimal variables for the reduced model, accounting for `r prettify(vartab(cap11sum)[2, "Proportion"])`% of the total variation.
 ANOVA revealed that the reduced model was significant with an adjusted R^2^ of `r RDA1R2` (*P* = 0.001).
-Assessment of the marginal effects showed that region, host, year, and aggressiveness significantly explained genetic variation (*P* = `r anova1res[["Year"]]`, *P* = `r anova1res[["Region"]]` for region).
+Assessment of the marginal effects showed that all of these varaibles significantly explained genetic variation (*P* = `r anova1res[["Year"]]`, *P* = `r anova1res[["Region"]]` for region).
 It was revealed that there was multicollinearity when MCG was combined with any other variable, so we followed the same procedure, dropping MCG from the list of potential predictors. 
-From these results, we found that aggressiveness, year, region, and host were optimal, accounting for `r prettify(vartab(cap11sum2)[2, "Proportion"])`% of the total variation.
+From these results, we found that `r RDAVAR2` were optimal, accounting for `r prettify(vartab(cap11sum2)[2, "Proportion"])`% of the total variation.
 ANOVA revealed significant effects with an adjusted R^2^ of `r RDA2R2` (*P* = 0.001).
-While the marginal effect assessment showed that region, host, and year all significantly explained the variation at *P* = `r anova2res[["Year"]]`, and aggressiveness significantly explained the variation at *P* = `r anova2res[["Severity"]]`.
+While the marginal effect assessment showed that `r andlist(names(anova2res)[anova2res <= 0.01])` all significantly explained the variation at *P* = `r anova2res[["Year"]]`, and Aggressiveness significantly explained the variation at *P* = `r anova2res[["Severity"]]`.
 Much of the variation appears to be driven by isolates from Mexico and 2005 (Fig. \@ref(rda-biplot)).
@@
 \@ref(aggressiveness-boxplot), Table \@ref(tab:region-aggressiveness)).
-MCGs also showed a significant effect (P < 1.00e^-4^), with means that ranged from 6.0 ('MCG 44') to 4.6 ('MCG 49').
-We additionally found a significant effect for MLGs (*P* = 7.44e^-4^) for MLGs, with means that ranged from 6.0 ('MLG 78') to 4.3 ('MLG 140').
+MCGs also showed a significant effect (P < 1.00e^-4^), with means that ranged from 6.0 ('MCG 44') to 4.6 ('MCG 49') (Table \@ref(tab:MCG-aggressiveness)).
+We additionally found a significant effect for MLGs (*P* = 7.44e^-4^) for MLGs, with means that ranged from 6.0 ('MLG 78') to 4.3 ('MLG 140') (Table \@ref(tab:MLG-aggressiveness)).
 
@@
 Our network-based approach to correlating MLGs with MCGs revealed a large and complex network (Fig. \@ref(biggraph)). 
-Community analysis showed 51 communities, 15 of which consisted of a single MLG unconnected with any other community indicating that just 9.09% of the 165 MLGs are unable to cross with any other MLG in this data set (Fig.
+Community analysis showed 51 communities, `r sum(in_reach$N == 0)` of which consisted of a single MLG unconnected with any other community indicating that just `r prettify(mean(in_reach$N == 0))`% of the 165 MLGs are unable to cross with any other MLG in this data set (Fig.
 \@ref(fullgraph)).
@@
 Comparing these communities with Bruvo's genetic distance revealed an average distance of 0.451 among communities and an average distance of 0.437 within communities, showing no significant difference.
-When we assessed the number of times two different MLGs that are in the same MCG, considering these as potential heterothallic pairings that could result in sexual recombination, we found an average of 16.5 potential heterothallic parings per MLG.
-Representing just four isolates, 'MLG 75' had 57 neighbors that shared the same mycelial compatibility group (Fig \@ref(biggraph), \@ref(fullgraph)). 
+When we assessed the number of times two different MLGs that are in the same MCG, considering these as potential heterothallic pairings that could result in sexual recombination, we found an average of `r signif(mean(in_reach$N), 3)` potential heterothallic parings per MLG.
+Representing just four isolates, 'MLG 75' had `r big_genotype$N` neighbors that shared the same mycelial compatibility group (Fig \@ref(biggraph), \@ref(fullgraph)). 
 Overall, there was no clear pattern to the association between MLG and MCGs.
@@
 
-Borrowing a technique often used in the ecological literature, we used dbRDA to elucidate the effect of all variables (Aggressiveness, Region, Source,   Year, and Host) [@legendre1999distance].
+Borrowing a technique often used in the ecological literature, we used dbRDA to elucidate the effect of all variables (`r andlist(c(names(strata(dat11)), "Aggressiveness"))`) [@legendre1999distance].
 From the initial results, it appears that the most important factors for predicting genetic structure are MCG, region, and year.
@@
 Aggressiveness and specific source field had little to no effect on genetic diversity. 
-These results are in agreement with studies that examined differentiation based on host [@aldrich-wolfe2015genetic] and aggressiveness [@atallah2004high; @attanayake2012genetic; @attanayake2013sclerotinia] reporting little or no correlation of genetic diversity to these variables.
+These results are in agreement with studies that examined differentiation based on Host [@aldrich-wolfe2015genetic] and Aggressiveness [@atallah2004high; @attanayake2012genetic; @attanayake2013sclerotinia] reporting little or no correlation of genetic diversity to these variables.
 This indicates that a) breeders should keep in mind regional differences when assessing resistance and b) it is possible that we have not yet measured biologically relevant variables that can predict genetic differentiation, which could include variables such as soil community composition.
@@
 It could be argued, however that 'MLG 25', with its high value of $P_{sex}$ represents different true MLGs across the five MCGs it occupies, but this doesn't account for the overall structure of Fig.
-\@ref(fullgraph) where, for example, 'MLG 75' ($P_{sex}$ = 1.81e^-4^) is compatible with 58 other genotypes through three MCG when the population structure of *S.
+\@ref(fullgraph) where, for example, 'MLG 75' ($P_{sex}$ = 1.81e^-4^) is compatible with `r big_genotype$N` other genotypes through three MCG when the population structure of *S.
 sclerotiorum* is known to be clonal.
@@
 
+```{r MCG-ag, echo = FALSE, warning = FALSE, message = FALSE, results = "asis"}
+read_csv("../../results/tables/MCG-aggressiveness.csv", 
+         na = "NA", # spaces read in as spaces
+         col_types = cols(.default = col_character())) %>%
+  huxtable::as_huxtable(add_colnames = TRUE) %>%
+  huxtable::set_number_format(row = huxtable::everywhere, col = 1, 0) %>% 
+  huxtable::set_col_width(c(0.13, 0.25, 0.08)) %>%
+  huxtable::set_align(row = huxtable::everywhere, col = 2, "right") %>%
+  huxtable::print_md(max_width = 40)
+```
+
+Table: (\#tab:MCG-aggressiveness) Mean aggressiveness ratings for the first 10 MCG; groupings according to 95% family-wise confidence interval.
+
+```{r MLG-ag, echo = FALSE, warning = FALSE, message = FALSE, results = "asis"}
+read_csv("../../results/tables/MLG-aggressiveness.csv", 
+         na = "NA", # spaces read in as spaces
+         col_types = cols(.default = col_character())) %>%
+  huxtable::as_huxtable(add_colnames = TRUE) %>%
+  huxtable::set_number_format(row = huxtable::everywhere, col = 1, 0) %>% 
+  huxtable::set_col_width(c(0.13, 0.25, 0.08)) %>%
+  huxtable::set_align(row = huxtable::everywhere, col = 2, "right") %>%
+  huxtable::print_md(max_width = 40)
+```
+
+Table: (\#tab:MLG-aggressiveness) Mean aggressiveness ratings for the first 10 MLG; groupings according to 95% family-wise confidence interval.
+
 ![

### 2017-08-23 doc/RMD/RDA-analysis.Rmd
vastly reduce the size of RDA data
Turns out that saving environments is really friggen
expensive.

@@
 ```{r save_data}
-save(cap11cc, cap11cc2, 
-     cap11sum, cap11sum2, 
-     cap11anova, cap11anova2,
-     file = file.path("data", "RDA_results.rda"))
+caplist1 <- list(
+  anova = cap11anova,
+  R2    = RsquareAdj(cap11cc),
+  components = cap11sum[-1]
+)
+caplist2 <- list(
+  anova = cap11anova2,
+  R2    = RsquareAdj(cap11cc2),
+  components = cap11sum2[-1]
+)
+saveRDS(caplist1, file.path(PROJHOME, "data", "caplist1.rds")) 
+saveRDS(caplist2, file.path(PROJHOME, "data", "caplist2.rds")) 
 ```

### 2017-08-23 doc/manuscript/manuscript.Rmd
vastly reduce the size of RDA data
Turns out that saving environments is really friggen
expensive.

@@
 load("../../data/sclerotinia_16_loci.rda")
-load("../../data/RDA_results.rda")
 load("../../data/mcg-mlg-graph.rda")
 load("../../data/heterothallic-pairings.rda")
+caplist1 <- read_rds("../../data/caplist1.rds")
+caplist2 <- read_rds("../../data/caplist2.rds")
 dat11cc  <- clonecorrect(dat11, ~Region/Source/Host/Year)
@@
 
-RDA2R2    <- prettify(vegan::RsquareAdj(cap11cc2)$adj.r.squared)/100
-RDA1R2    <- prettify(vegan::RsquareAdj(cap11cc)$adj.r.squared)/100
-anova1res <- setNames(cap11anova$`Pr(>F)`, rownames(cap11anova)) %>% na.omit()
-anova2res <- setNames(cap11anova2$`Pr(>F)`, rownames(cap11anova2)) %>% na.omit()
-RDAVAR1   <- andlist(names(cap11cc$terminfo$ordered))
-RDAVAR2   <- andlist(gsub("Severity", "Aggressiveness", names(cap11cc2$terminfo$ordered)))
+RDA2R2    <- prettify(caplist1$R2$adj.r.squared)/100
+RDA1R2    <- prettify(caplist2$R2$adj.r.squared)/100
+anova1res <- setNames(caplist1$anova$`Pr(>F)`, rownames(caplist1$anova)) %>% na.omit()
+anova2res <- setNames(caplist2$anova$`Pr(>F)`, rownames(caplist2$anova)) %>% na.omit()
+RDAVAR1   <- andlist(names(anova1res))
+RDAVAR2   <- andlist(gsub("Severity", "Aggressiveness", names(anova2res)))
 
@@
 
-The forward-backward selection process of the dbRDA models on clone-corrected data revealed `r RDAVAR1` to be the optimal variables for the reduced model, accounting for `r prettify(vartab(cap11sum)[2, "Proportion"])`% of the total variation.
+The forward-backward selection process of the dbRDA models on clone-corrected data revealed `r RDAVAR1` to be the optimal variables for the reduced model, accounting for `r prettify(vartab(caplist1$components)[2, "Proportion"])`% of the total variation.
 ANOVA revealed that the reduced model was significant with an adjusted R^2^ of `r RDA1R2` (*P* = 0.001).
@@
 It was revealed that there was multicollinearity when MCG was combined with any other variable, so we followed the same procedure, dropping MCG from the list of potential predictors. 
-From these results, we found that `r RDAVAR2` were optimal, accounting for `r prettify(vartab(cap11sum2)[2, "Proportion"])`% of the total variation.
+From these results, we found that `r RDAVAR2` were optimal, accounting for `r prettify(vartab(caplist2$components)[2, "Proportion"])`% of the total variation.
 ANOVA revealed significant effects with an adjusted R^2^ of `r RDA2R2` (*P* = 0.001).

### 2017-08-23 doc/RMD/MLG-distribution.Rmd
change MLG to MLH

@@
      # title = "Shared MLGs across regions",
-      fill = "Number of\nMLGs",
+      fill = "Number of\nMLHs",
       edge_alpha = "Probability of\nsecond encounter",
-      edge_width = "Populations\nper MLG"#,
+      edge_width = "Populations\nper MLH"#,
       #caption = "Outer circle: Number of MLGs in the region\nInner Circle: Number of private MLGs in the region"
@@
 V(g)$size <- c(56, 56 - 32)
-V(g)$label <- c("Number of MLGs in Region", "Number of private MLGs")
+V(g)$label <- c("Number of MLHs in Region", "Number of private MLHs")
 lay <- as.data.frame(matrix(0, 2, 2, dimnames = list(NULL, c("x", "y"))))

### 2017-08-23 doc/RMD/mlg-mcg.Rmd
change MLG to MLH

@@
   rename(Type = type) %>%
+  mutate(Type = ifelse(Type == "MLGs", "MLHs", Type)) %>%
   mutate(MCG = forcats::fct_inorder(MCG, ordered = TRUE)) %>%
   ggplot(aes(x = MCG, y = count, group = Type, fill = Type, alpha = Evenness)) +
-  geom_col(aes(width = ifelse(Type == "MLGs", 0.5, 0.85)), color = "black", position = "identity") +
+  geom_col(aes(width = ifelse(Type == "MLHs", 0.5, 0.85)), color = "black", position = "identity") +
   annotate(geom = "rect", xmin = 0.5, xmax = 8.5, ymin = 0, ymax = 75, color = "black", size = 1.5, alpha = 0, lty = 2) +
@@
   labs(list(
-    alpha = "MLG\nEvenness"
+    alpha = "MLH\nEvenness"
   ))
@@
   xlab("Potential Heterothallic Pairings") +
-  ylab("Multilocus Genotypes")
+  ylab("Multilocus Haplotypes")
 in_reach_plot
@@
 top5lay$community <- inner_join(data_frame(vertex = top5lay$name), the_communities)$comm
-levels(top5lay$type) <- c("Multilocus Genotype", "Mycelial Compatibility\nGroup")
+levels(top5lay$type) <- c("Multilocus Haplotype", "Mycelial Compatibility\nGroup")
 t5g <- ggraph(top5lay) +
@@
     title = "Severity by MCG and Region",
-    fill = "Multilocus Genotype",
+    fill = "Multilocus Haplotype",
     subtitle = "Five most abundant multilocus genotypes shown"

### 2017-08-23 doc/manuscript/manuscript.Rmd
change MLG to MLH

@@
 dapc_region      <- dapc(setPop(dat11cc, ~Region), n.da = 13, n.pca = 20)
-dapc_region_year <- dapc(setPop(dat11cc, ~Region/Year), n.da = 12, n.pca = 22)
+dapc_region_year <- dapc(setPop(dat11cc, ~Region/Year), n.da = 15, n.pca = 15)
 dapc_host        <- dat11cc %>% 
@@
 
-big_genotype <- in_reach %>% filter(N == max(N))
+big_haplotype <- in_reach %>% filter(N == max(N))
 ```
@@
  - What are we missing? A detailed analysis of *S. sclerotiorum* from diverse
-   regions that may potentially identify genotypes or MCGs that may be good
+   regions that may potentially identify haplotypes or MCGs that may be good
    tools for testing breeding efforts.
@@
 Furthermore, we wanted to investigate the potential effect of cultivar on genetic diversity by assessing three dry bean cultivars with different levels of resistance, Beryl (great northern bean, susceptible), Bunsi (navy bean, low susceptibility), and G122 (cranberry bean, partial resistance) [@otto-hanson2011variation].
-We additionally wanted to determine categorical or phenotypic variables that best predicted genetic structure and if there was correlation between multilocus genotype and mycelial compatibility group.
+We additionally wanted to determine categorical or phenotypic variables that best predicted genetic structure and if there was correlation between multilocus haplotype and mycelial compatibility group.
 By investigating these aims, we hope to effectively describe the population structure of *S.
@@
 Loci with compound repeats were removed for the reported statistics, but were processed and included in data analyses to ensure the integrity of the results with 11 loci as compared to 16.
-We assessed the power of our 11 markers by generating a genotype accumulation curve in the R package *poppr* version `r packageVersion("poppr")`, looking for evidence of saturation, which would indicate that loci were sufficiently sampled to adequately represent the full set of genotypes [@arnaud2007standardizing; @kamvar2015novel].
-To avoid including isolates potentially collected from the same plant (which increases the probability of collecting sclerotioa from the same point of infection more than once), data were clone-corrected on a hierarchy of Region/Source/Host/Year---meaning that one representative of each genotype was retained for each nested combination of Region, Source, Host, and Year---for subsequent analysis.
+We assessed the power of our 11 markers by generating a genotype accumulation curve in the R package *poppr* version `r packageVersion("poppr")`, looking for evidence of saturation, which would indicate that loci were sufficiently sampled to adequately represent the full set of haplotypes [@arnaud2007standardizing; @kamvar2015novel].
+To avoid including isolates potentially collected from the same plant (which increases the probability of collecting sclerotioa from the same point of infection more than once), data were clone-corrected on a hierarchy of Region/Source/Host/Year---meaning that one representative of each haplotype was retained for each nested combination of Region, Source, Host, and Year---for subsequent analysis.
 We assessed genotypic diversity by calculating Stoddart and Taylor's index (*G*) [@stoddart1988genotypic], Shannon's index (*H*) [@shannon2001mathematical], Simpson's index ($\lambda$) [@simpson1949measurement], and evenness ($E_5$, *eMLG*) [@heck1975explicit; @hurlbert1971nonconcept; @pielou1975ecological; @grunwald2003analysis].
@@
 
-We used ANOVA to assess if aggressiveness could be explained by Region, MCG, or multilocus genotype (MLG). 
-To minimize complications due to small sample sizes, we chose the top 10 MCGs, representing 56.5% of the isolates collected, the top 10 MLGs representing 26.7% of the isolates, and populations with more than five isolates.
+We used ANOVA to assess if aggressiveness could be explained by Region, MCG, or multilocus haplotype (MLH). 
+To minimize complications due to small sample sizes, we chose the top 10 MCGs, representing 56.5% of the isolates collected, the top 10 MLHs representing 26.7% of the isolates, and populations with more than five isolates.
 Between-group differences were assessed using Tukey's HSD test with an alpha of 0.05 using the `HSD.test()` function in the package *agricolae* version `r packageVersion("agricolae")` [@mendiburu2015agricolae].
 
-### Correlating Multilocus Genotypes with Mycelial Compatibility Groups {-}
+### Correlating Multilocus Haplotypes with Mycelial Compatibility Groups {-}
 
-We wanted to assess if there was correlation between MLGs and MCGs. 
-In simple terms, this was done using a network-based approach where both MLGs and MCGs were considered nodes and the number of isolates in which they were found together was the strength of the connection between an MLG and and MCG node.
-The network-based approach allowed us to assess how "clumpy" the relationships between MLGs and MCGs were.
-In more specific terms, we first created a contingency table with MLGs and MCGs and converted it to a directed and weighted edgelist where each edge was a connection from an MCG to an MLG weighted by the number of samples shared in the connection.
-This was then converted to a bipartite graph where top nodes represented MCGs and bottom nodes represented MLGs.
-The cluster walktrap community detection algorithm as implemented in the `cluster_walktrap()` function in *igraph* version `r packageVersion("igraph")` was used to define clusters of MLGs and MCGs [@csardi2006igraph; @pons2006computing].
+We wanted to assess if there was correlation between MLHs and MCGs. 
+In simple terms, this was done using a network-based approach where both MLHs and MCGs were considered nodes and the number of isolates in which they were found together was the strength of the connection between an MLH and and MCG node.
+The network-based approach allowed us to assess how "clumpy" the relationships between MLHs and MCGs were.
+In more specific terms, we first created a contingency table with MLHs and MCGs and converted it to a directed and weighted edgelist where each edge was a connection from an MCG to an MLH weighted by the number of samples shared in the connection.
+This was then converted to a bipartite graph where top nodes represented MCGs and bottom nodes represented MLHs.
+The cluster walktrap community detection algorithm as implemented in the `cluster_walktrap()` function in *igraph* version `r packageVersion("igraph")` was used to define clusters of MLHs and MCGs [@csardi2006igraph; @pons2006computing].
 All communities that had fewer than 10 members were then consolidated into one.
@@
 
-### Analysis of Shared Multilocus Genotypes {-}
+### Analysis of Shared Multilocus Haplotypes {-}
 
-We wanted to evaluate patterns of connectivity between shared multilocus genotypes across geographic regions.
-We first tabulated the multilocus genotypes shared between at least two populations (defined as states or countries) with the *poppr* function `mlg.crosspop()` [@kamvar2014poppr].
-From these data, we constructed a graph with populations as nodes and shared genotypes as edges (connections) between nodes using the R packages *igraph* [@csardi2006igraph], *dplyr* version `r packageVersion("dplyr")` [@dplyr], and *purrr* version `r packageVersion("purrr")` [@purrr].
-Each node was weighted by the fraction of shared MLGs.
-Each edge represented a single MLG, but because a single MLG could be present in more than one population, that MLG would be represented by (*n*\*(*n*-1))/2 edges where n represents the number of populations crossed.
-Edges were weighted by $1-P_{sex}$, where $P_{sex}$ is the probability of encountering the same genotype via two independent meiotic events [@arnaud2007standardizing; @parks1993study].
+We wanted to evaluate patterns of connectivity between shared multilocus haplotypes across geographic regions.
+We first tabulated the multilocus haplotypes shared between at least two populations (defined as states or countries) with the *poppr* function `mlg.crosspop()` [@kamvar2014poppr].
+From these data, we constructed a graph with populations as nodes and shared haplotypes as edges (connections) between nodes using the R packages *igraph* [@csardi2006igraph], *dplyr* version `r packageVersion("dplyr")` [@dplyr], and *purrr* version `r packageVersion("purrr")` [@purrr].
+Each node was weighted by the fraction of shared MLHs.
+Each edge represented a single MLG, but because a single MLH could be present in more than one population, that MLH would be represented by (*n*\*(*n*-1))/2 edges where n represents the number of populations crossed.
+Edges were weighted by $1-P_{sex}$, where $P_{sex}$ is the probability of encountering the same haplotype via two independent meiotic events [@arnaud2007standardizing; @parks1993study].
 We then used the `cluster_optimal()` function from *igraph* to define communities [@csardi2006igraph].
@@
 Four regions had fewer than 10 isolates, Australia (`r pop_size[["Australia"]]`), WI (`r pop_size[["WI"]]`), NY (`r pop_size[["NY"]]`), ID (`r pop_size[["ID"]]`).
-We observed 87 MCGs, the most abundant of which ('MCG 5') was represented by 73 isolates over 37 MLGs (Fig \@ref(biggraph)A,C).
-Analysis of the genotype accumulation curve showed no clear plateau (Supplemental Information), indicating that we would likely obtain more multilocus genotypes if we were to genotype more loci.
+We observed 87 MCGs, the most abundant of which ('MCG 5') was represented by 73 isolates over 37 MLHs (Fig \@ref(biggraph)A,C).
+Analysis of the haplotype accumulation curve showed no clear plateau (Supplemental Information), indicating that we would likely obtain more multilocus haplotypes if we were to haplotype more loci.
 The number of observed alleles per locus ranged from two to 10 with an average of `r signif(Amean, 3)` (Table \@ref(tab:locus-stats)).
-After clone-correction on the hierarchy of Region/Source/Host/Year, we were left with `r nInd(dat11cc)` isolates representing `r nmll(dat11cc)` MLGs (Table \@ref(tab:iatab)).
+After clone-correction on the hierarchy of Region/Source/Host/Year, we were left with `r nInd(dat11cc)` isolates representing `r nmll(dat11cc)` MLHs (Table \@ref(tab:iatab)).
 
@@
   rename(`$\\bar{r}_d$` = rbarD) %>%
+  rename(eMLH = eMLG) %>%
   select(-File, -MLG, -SE, -p.Ia, -p.rD, -Ia) %>% 
@@
 Table: (\#tab:iatab) Genotypic diversity and Linkage Disequilibrium summary for geographic populations arranged by abundance after clone-correction by a hierarchy of Region/Source/Host/Year. 
-Pop = Population, N = number of individuals (number of MLG in parentheses), 
-eMLG = expected number of MLGs based on rarefaction at min(10, N) individuals (standard error in parentheses), 
+Pop = Population, N = number of individuals (number of MLH in parentheses), 
+eMLH = expected number of MLHs based on rarefaction at min(10, N) individuals (standard error in parentheses), 
 H = Shannon-Weiner Index, G = Stoddardt and Taylor's Index, 
@@
 ![
-Data show complex associations between Mycelial Compatibility Groups and Multilocus Genotypes.
+Data show complex associations between Mycelial Compatibility Groups and Multilocus Haplotypes.
 **A)** Barplot of Mycelial Compatibility Group (MCG) abundance in descending order.
 Singletons (46) were truncated, leaving 41 MCGs.
-White bars represent sample counts and grey bars represent counts of unique multilocus genotypes (MLG).
-The transparency of the bars represent the evenness of the distribution of the MLGs within a given MCG.
+White bars represent sample counts and grey bars represent counts of unique multilocus haplotypes (MLH).
+The transparency of the bars represent the evenness of the distribution of the MLHs within a given MCG.
 A dashed box surrounds the 8 most common MCGs representing > 51% of the data.
-**B)** Full graph-representation of the relationship between MCGs (open circles) and MLGs (filled circles).
+**B)** Full graph-representation of the relationship between MCGs (open circles) and MLHs (filled circles).
 Details in fig. \@ref(fullgraph). 
-**C)** A subset of **B** representing the 8 most common MCGs and their associated MLGs (dashed box in **A**).
-Filled nodes represent MLGs and open nodes represent MCGs.
+**C)** A subset of **B** representing the 8 most common MCGs and their associated MLHs (dashed box in **A**).
+Filled nodes represent MLHs and open nodes represent MCGs.
 Numbers inside nodes are the MLG/MCG label. 
 Node (circle) area scaled to the number of samples represented (range: 1--73).
-Edges (arrows) point from MLG to MCG where the weight (thickness) of the edge represents the number of shared isolates (range: 1--19).
-Edges extending from MLGs displayed to other MCGs are not shown.
+Edges (arrows) point from MLH to MCG where the weight (thickness) of the edge represents the number of shared isolates (range: 1--19).
+Edges extending from MLHs displayed to other MCGs are not shown.
 ](../../results/figures/publication/mcg-subgraph-with-context.pdf){#biggraph width=100%}
@@
 MCGs also showed a significant effect (P < 1.00e^-4^), with means that ranged from 6.0 ('MCG 44') to 4.6 ('MCG 49') (Table \@ref(tab:MCG-aggressiveness)).
-We additionally found a significant effect for MLGs (*P* = 7.44e^-4^) for MLGs, with means that ranged from 6.0 ('MLG 78') to 4.3 ('MLG 140') (Table \@ref(tab:MLG-aggressiveness)).
+We additionally found a significant effect for MLHs (*P* = 7.44e^-4^) for MLHs, with means that ranged from 6.0 ('MLH 78') to 4.3 ('MLH 140') (Table \@ref(tab:MLG-aggressiveness)).
 
-### Correlation of Mulitlocus Genotypes and Mycelial Compatibility Groups {-}
+### Correlation of Mulitlocus Haplotypes and Mycelial Compatibility Groups {-}
 
-Our network-based approach to correlating MLGs with MCGs revealed a large and complex network (Fig. \@ref(biggraph)). 
-Community analysis showed 51 communities, `r sum(in_reach$N == 0)` of which consisted of a single MLG unconnected with any other community indicating that just `r prettify(mean(in_reach$N == 0))`% of the 165 MLGs are unable to cross with any other MLG in this data set (Fig.
+Our network-based approach to correlating MLHs with MCGs revealed a large and complex network (Fig. \@ref(biggraph)). 
+Community analysis showed 51 communities, `r sum(in_reach$N == 0)` of which consisted of a single MLH unconnected with any other community indicating that just `r prettify(mean(in_reach$N == 0))`% of the 165 MLHs are unable to cross with any other MLH in this data set (Fig.
 \@ref(fullgraph)).
@@
 Comparing these communities with Bruvo's genetic distance revealed an average distance of 0.451 among communities and an average distance of 0.437 within communities, showing no significant difference.
-When we assessed the number of times two different MLGs that are in the same MCG, considering these as potential heterothallic pairings that could result in sexual recombination, we found an average of `r signif(mean(in_reach$N), 3)` potential heterothallic parings per MLG.
-Representing just four isolates, 'MLG 75' had `r big_genotype$N` neighbors that shared the same mycelial compatibility group (Fig \@ref(biggraph), \@ref(fullgraph)). 
-Overall, there was no clear pattern to the association between MLG and MCGs.
+When we assessed the number of times two different MLHs that are in the same MCG, considering these as potential heterothallic pairings that could result in sexual recombination, we found an average of `r signif(mean(in_reach$N), 3)` potential heterothallic parings per MLG.
+Representing just four isolates, 'MLH 75' had `r big_haplotype$N` neighbors that shared the same mycelial compatibility group (Fig \@ref(biggraph), \@ref(fullgraph)). 
+Overall, there was no clear pattern to the association between MLH and MCGs.
 
-## Structure of Shared Multilocus Genotypes {-}
+## Structure of Shared Multilocus Haplotypes {-}
 
-We observed a total of 165 MLGs (215 with 16 loci).
-The most abundant MLG was represented by 27 isolates (Table \@ref(tab:mlg-table)) from five regions (NE, MI, WA, CO, and ND). 
-Within regions, the genotypes were relatively evenly distributed with moderate to high diversity (Table \@ref(tab:iatab)). 
-Of the 165 MLGs, 76 (46%) were found in at least two geographic regions except those found in WI (2), ID (1), and Mexico (18) (Fig. \@ref(community-graph)).
-Community analysis of the MLGs shared between populations revealed 4 communities with a modularity of 0.17 (Fig.
+We observed a total of 165 MLHs (215 with 16 loci).
+The most abundant MLH was represented by 27 isolates (Table \@ref(tab:mlg-table)) from five regions (NE, MI, WA, CO, and ND). 
+Within regions, the haplotypes were relatively evenly distributed with moderate to high diversity (Table \@ref(tab:iatab)). 
+Of the 165 MLHs, 76 (46%) were found in at least two geographic regions except those found in WI (2), ID (1), and Mexico (18) (Fig. \@ref(community-graph)).
+Community analysis of the MLHs shared between populations revealed 4 communities with a modularity of 0.17 (Fig.
 \@ref(community-graph)): A coastal community (CA, OR, WA, and NY), a midwest community (CO, NE, ND, MI), and an international community (MN, France, Australia).
-Analysis with 16 loci resulted in the removal of the NY node because it no longer shared a genotype with OR, but the same overall community structure was present with a modularity of 0.2 (Fig.
+Analysis with 16 loci resulted in the removal of the NY node because it no longer shared a haplotype with OR, but the same overall community structure was present with a modularity of 0.2 (Fig.
 \@ref(community-graph16)).
-The international community appears to be driven by MLG 4, which is shared between all three populations and has a $P_{sex}$ value of 2.87e^-5^, in contrast to the abundant MLG 25, which has a $P_{sex}$ value of 0.0168.
+The international community appears to be driven by MLH 4, which is shared between all three populations and has a $P_{sex}$ value of 2.87e^-5^, in contrast to the abundant MLH 25, which has a $P_{sex}$ value of 0.0168.
 
@@
 ![
-Network of populations (nodes/circles) and their shared multilocus genotypes (MLG) (edges/lines) genotyped over 11 loci. 
-Each node is labeled with **name (number of MLGs shared/number of MLGs total).** 
-The shade and area of the nodes are proportional to the number of unique MLGs within the node and the inner nodes are proportional to the number of private MLGs to the region (bottom legend). 
-Each edge represents a single MLG where its thickness represents the number of populations that share the MLG and the shade represents the value of $P_{sex}$, or the probability of encountering that MLG from two independent meiotic events.
+Network of populations (nodes/circles) and their shared multilocus haplotypes (MLH) (edges/lines) haplotyped over 11 loci. 
+Each node is labeled with **name (number of MLHs shared/number of MLHs total).** 
+The shade and area of the nodes are proportional to the number of unique MLHs within the node and the inner nodes are proportional to the number of private MLHs to the region (bottom legend). 
+Each edge represents a single MLH where its thickness represents the number of populations that share the MLH and the shade represents the value of $P_{sex}$, or the probability of encountering that MLH from two independent meiotic events.
 ](../../results/figures/publication/mlg-11.pdf){#community-graph width=75%}
@@
 
-Table: (\#tab:mlg-table) The 5 most abundant Multilocus Genotypes (MLG) with
+Table: (\#tab:mlg-table) The 5 most abundant Multilocus Haplotypes (MLH) with
 the probability of second encounter ($P_{sex}$), Mycelial Compatibility Groups (MCG), and Regions with sample sizes in parentheses.
@@
 \@ref(DAPC-RY)).
-Further analysis of this population revealed that all 12 isolates in WA circa 2008 originated in a wmn and nine of these genotypes were shared with CA, and three were shared with France (Fig.
+Further analysis of this population revealed that all 12 isolates in WA circa 2008 originated in a wmn and nine of these haplotypes were shared with CA, and three were shared with France (Fig.
 \@ref(community-graph), \@ref(community-graph16)).
@@
 Our results suggest that, broadly, populations from white mold screening nurseries reflect the populations of the surrounding regions, indicating that resistance screening may be successful at a within regions.
-We found significant population differentiation by geographic region, mainly differentiated into three broad North American groups based on shared genotypes and posterior groupings, a Coastal Region, Midwestern Region, and Mexico.
+We found significant population differentiation by geographic region, mainly differentiated into three broad North American groups based on shared haplotypes and posterior groupings, a Coastal Region, Midwestern Region, and Mexico.
 To date, with 366 isolates, this is the largest single population genetic study of *S. sclerotiorum* assessing population structure within managed and unmanaged agricultural environments. 
@@
 When we inspected the biplot of the initial results, we saw that the most important predictors were 'MCG 44', 'MCG 5', and 'MCG 9'. 
-We believe that this is driven by the fact that these particular MCGs have uneven MLG distributions, meaning that they are heavily associated with one particular MLG (Fig. \@ref(biggraph)).
+We believe that this is driven by the fact that these particular MCGs have uneven MLH distributions, meaning that they are heavily associated with one particular MLH (Fig. \@ref(biggraph)).
 We note these results with caution because of the apparent multicolinearity between MCG and Region, which is a violation of the analysis [@legendre1999distance].
@@
 
-## Mycelial Compatibility Groups and Multilocus Genotypes {-}
+## Mycelial Compatibility Groups and Multilocus Haplotypes {-}
 
-Of interest in this study was testing the association between multilocus genotypes (MLGs) and mycelial compatibility groups (MCGs), as the latter have historically been used as a proxy for clonal lineages [@leslie1993fungal; @otto-hanson2011variation; @kohn1990mycelial; @schafer2006optimized; @kohn1995clonal; @carbone1999patterns].
+Of interest in this study was testing the association between multilocus haplotypes (MLHs) and mycelial compatibility groups (MCGs), as the latter have historically been used as a proxy for clonal lineages [@leslie1993fungal; @otto-hanson2011variation; @kohn1990mycelial; @schafer2006optimized; @kohn1995clonal; @carbone1999patterns].
 Our results do not support this assumption.
 It can be seen in Fig.
-\@ref(biggraph)A that the most abundant MCG contains several MLGs, but the diversity of those MLGs are low as indicated by the evenness (transparency), which indicates that there is one dominant MLG ('MLG 25').
-What is not shown in Fig. \@ref(biggraph)A is the MLGs that are shared between MCGs. 
+\@ref(biggraph)A that the most abundant MCG contains several MLHs, but the diversity of those MLHs are low as indicated by the evenness (transparency), which indicates that there is one dominant MLH ('MLH 25').
+What is not shown in Fig. \@ref(biggraph)A is the MLHs that are shared between MCGs. 
 This is illustrated in both table \@ref(tab:mlg-table) and Fig. \@ref(biggraph)B,C. 
-It could be argued, however that 'MLG 25', with its high value of $P_{sex}$ represents different true MLGs across the five MCGs it occupies, but this doesn't account for the overall structure of Fig.
-\@ref(fullgraph) where, for example, 'MLG 75' ($P_{sex}$ = 1.81e^-4^) is compatible with `r big_genotype$N` other genotypes through three MCG when the population structure of *S.
+It could be argued, however that 'MLH 25', with its high value of $P_{sex}$ represents different true MLHs across the five MCGs it occupies, but this doesn't account for the overall structure of Fig.
+\@ref(fullgraph) where, for example, 'MLH 75' ($P_{sex}$ = 1.81e^-4^) is compatible with `r big_haplotype$N` other haplotypes through three MCG when the population structure of *S.
 sclerotiorum* is known to be clonal.
 
-Over the past few years, researchers have noticed inconsistencies among the relationship between MCGs and MLGs [@aldrich-wolfe2015genetic; @carbone1999patterns; @attanayake2012genetic; @lehner2015genetic].
-Either several MCGs belong to one MLG, which could be explained by insufficient sampling of loci; several MLGs belong to one MCG, which could be explained by clonal expansion; or a mixture of both.
-Some studies have shown a correlation between MCG and MLG [@aldrich-wolfe2015genetic; @carbone1999patterns; @lehner2015genetic], whereas other studies have shown no apparent correlation, even on small spatial scales [@atallah2004high; @attanayake2012genetic; @attanayake2013sclerotinia].
+Over the past few years, researchers have noticed inconsistencies among the relationship between MCGs and MLHs [@aldrich-wolfe2015genetic; @carbone1999patterns; @attanayake2012genetic; @lehner2015genetic].
+Either several MCGs belong to one MLG, which could be explained by insufficient sampling of loci; several MLHs belong to one MCG, which could be explained by clonal expansion; or a mixture of both.
+Some studies have shown a correlation between MCG and MLH [@aldrich-wolfe2015genetic; @carbone1999patterns; @lehner2015genetic], whereas other studies have shown no apparent correlation, even on small spatial scales [@atallah2004high; @attanayake2012genetic; @attanayake2013sclerotinia].
 
@@
 
-The results show several genotypes shared across populations, except for Mexico and two states that had fewer than three samples each (Fig. \@ref(community-graph)).
-Our network-based approach by treating the genotypes as edges and weighting each edge with the inverse of $P_{sex}$ treated the edges as springs connecting the populations with the strength proportional to the probability of obtaining the same genotype as a clone.
+The results show several haplotypes shared across populations, except for Mexico and two states that had fewer than three samples each (Fig. \@ref(community-graph)).
+Our network-based approach by treating the haplotypes as edges and weighting each edge with the inverse of $P_{sex}$ treated the edges as springs connecting the populations with the strength proportional to the probability of obtaining the same haplotype as a clone.
 This allowed us to use a graph walking algorithm to see how close the populations were simply based off of the proportion of clones they shared.
-The most abundant genotype was shared across four populations, but its high value of $P_{sex}$ meant that it did not contribute significantly to the overall structure. 
+The most abundant haplotype was shared across four populations, but its high value of $P_{sex}$ meant that it did not contribute significantly to the overall structure. 
 The graph walking algorithm was able to divide the network into three groups, but had a modularity of 0.17, which indicates that the groups are only weakly differentiated.
-This indicates that multilocus genotypes are widespread across the United States, giving more evidence for a clonal population structure.
+This indicates that multilocus haplotypes are widespread across the United States, giving more evidence for a clonal population structure.
 
@@
 Despite the evidence that Mexico and CA contributing to a large amount of the population differentiation, Regions like WA still had a large amount of internal variation (Supplementary Information).
-The two distinct clusters for the WA Region showed that the 2008 population appeared different and, under further investigation, we found that the genotypes from this population were shared between CA and France (Fig.
+The two distinct clusters for the WA Region showed that the 2008 population appeared different and, under further investigation, we found that the haplotypes from this population were shared between CA and France (Fig.
 \@ref(community-graph), \@ref(DAPC-RY), \@ref(DAPC-RY-FULL)).
@@
 \@ref(DAPC-RY)).
-DAPC is based on the principal components of Euclidean distance between genotypes [@jombart2010discriminant]. 
+DAPC is based on the principal components of Euclidean distance between haplotypes [@jombart2010discriminant]. 
 Unlike Bruvo's distance, Euclidean distance does not take into account the magnitude of the difference between alleles, which could inflate the distance measure in the presence of private alleles [@bruvo2004simple].
@@
 Another limitation was the microsatellite markers used for this particular study [@sirjusingh2001characterisation].
-The genotype accumulation curve showed no indication of a plateau, indicating that if we had sampled more loci, we would have resolved more multilocus genotypes. 
-While 16 loci showed us similar results and began to show a plateau for the genotype accumulation curve, we were unable to use these results due to our uncertainty in the allele calls for these five extra loci.
+The haplotype accumulation curve showed no indication of a plateau, indicating that if we had sampled more loci, we would have resolved more multilocus haplotypes. 
+While 16 loci showed us similar results and began to show a plateau for the haplotype accumulation curve, we were unable to use these results due to our uncertainty in the allele calls for these five extra loci.
 With the availability of an optically-mapped genome [@derbyshire2017complete], future studies describing the genetic diversity of *S.
@@
 sclerotiorum* may be considered representative of the surrounding regions.
-While we found no direct relationship between genotype and severity, it is evident that there is a gradient of severity by region, further supporting the need for screening in multiple locations.
+While we found no direct relationship between haplotype and severity, it is evident that there is a gradient of severity by region, further supporting the need for screening in multiple locations.
 
@@
 ![
-Graph showing complex associations between Mycelial Compatibility Groups (MCG) (dotted nodes) and Multilocus Genotypes (MLG) (full nodes) where the number in each node represents the MLG/MCG assignment.
+Graph showing complex associations between Mycelial Compatibility Groups (MCG) (dotted nodes) and Multilocus Haplotypes (MLH) (full nodes) where the number in each node represents the MLG/MCG assignment.
 Node size reflect the number of samples represented by each node (circle). 
-Edges (arrows) point from MLG to MCG where the weight (thickness) of the edge represents the number of samples shared. 
+Edges (arrows) point from MLH to MCG where the weight (thickness) of the edge represents the number of samples shared. 
 Node color represents the community assignment based on the walktrap algorithm with a maximum of four steps [@pons2006computing]. 
@@
 ![
-Network of populations (nodes/circles) and their shared multilocus genotypes (MLG) (edges/lines) genotyped over 16 loci. 
-Each node is labeled with **name (number of MLGs shared/number of MLGs total).** 
-The shade and area of the nodes are proportional to the number of unique MLGs within the node and the inner nodes are proportional to the number of private MLGs to the region (bottom legend). 
-Each edge represents a single MLG where its thickness represents the number of populations that share the MLG and the shade represents the value of $P_{sex}$, or the probability of encountering that MLG from two independent meiotic events.
+Network of populations (nodes/circles) and their shared multilocus haplotypes (MLH) (edges/lines) haplotyped over 16 loci. 
+Each node is labeled with **name (number of MLHs shared/number of MLHs total).** 
+The shade and area of the nodes are proportional to the number of unique MLHs within the node and the inner nodes are proportional to the number of private MLHs to the region (bottom legend). 
+Each edge represents a single MLH where its thickness represents the number of populations that share the MLH and the shade represents the value of $P_{sex}$, or the probability of encountering that MLH from two independent meiotic events.
 ](../../results/figures/publication/mlg-16.pdf){#community-graph16 width=100%}

### 2017-08-23 doc/manuscript/manuscript.Rmd
minor tweaks, now no MLG exist!

@@
 To avoid including isolates potentially collected from the same plant (which increases the probability of collecting sclerotioa from the same point of infection more than once), data were clone-corrected on a hierarchy of Region/Source/Host/Year---meaning that one representative of each haplotype was retained for each nested combination of Region, Source, Host, and Year---for subsequent analysis.
-We assessed genotypic diversity by calculating Stoddart and Taylor's index (*G*) [@stoddart1988genotypic], Shannon's index (*H*) [@shannon2001mathematical], Simpson's index ($\lambda$) [@simpson1949measurement], and evenness ($E_5$, *eMLG*) [@heck1975explicit; @hurlbert1971nonconcept; @pielou1975ecological; @grunwald2003analysis].
+We assessed genotypic diversity by calculating Stoddart and Taylor's index (*G*) [@stoddart1988genotypic], Shannon's index (*H*) [@shannon2001mathematical], Simpson's index ($\lambda$) [@simpson1949measurement], and evenness ($E_5$, *eMLH*) [@heck1975explicit; @hurlbert1971nonconcept; @pielou1975ecological; @grunwald2003analysis].
 To assess the potential for random mating, we tested for linkage disequilibrium with the index of association, $I_A$ and its standardized version, $\bar{r}_d$ using 999 permutations [@agapow2001indices; @smith1993how; @brown1980multilocus].
@@
 Each node was weighted by the fraction of shared MLHs.
-Each edge represented a single MLG, but because a single MLH could be present in more than one population, that MLH would be represented by (*n*\*(*n*-1))/2 edges where n represents the number of populations crossed.
+Each edge represented a single MLH, but because a single MLH could be present in more than one population, that MLH would be represented by (*n*\*(*n*-1))/2 edges where n represents the number of populations crossed.
 Edges were weighted by $1-P_{sex}$, where $P_{sex}$ is the probability of encountering the same haplotype via two independent meiotic events [@arnaud2007standardizing; @parks1993study].
@@
 Filled nodes represent MLHs and open nodes represent MCGs.
-Numbers inside nodes are the MLG/MCG label. 
+Numbers inside nodes are the MLH/MCG label. 
 Node (circle) area scaled to the number of samples represented (range: 1--73).
@@
 Comparing these communities with Bruvo's genetic distance revealed an average distance of 0.451 among communities and an average distance of 0.437 within communities, showing no significant difference.
-When we assessed the number of times two different MLHs that are in the same MCG, considering these as potential heterothallic pairings that could result in sexual recombination, we found an average of `r signif(mean(in_reach$N), 3)` potential heterothallic parings per MLG.
+When we assessed the number of times two different MLHs that are in the same MCG, considering these as potential heterothallic pairings that could result in sexual recombination, we found an average of `r signif(mean(in_reach$N), 3)` potential heterothallic parings per MLH.
 Representing just four isolates, 'MLH 75' had `r big_haplotype$N` neighbors that shared the same mycelial compatibility group (Fig \@ref(biggraph), \@ref(fullgraph)). 
@@
   rename(`$P_{sex}$` = Psex) %>% 
+  rename(MLH = MLG) %>%
   knitr::kable(format = "pandoc")
@@
 Over the past few years, researchers have noticed inconsistencies among the relationship between MCGs and MLHs [@aldrich-wolfe2015genetic; @carbone1999patterns; @attanayake2012genetic; @lehner2015genetic].
-Either several MCGs belong to one MLG, which could be explained by insufficient sampling of loci; several MLHs belong to one MCG, which could be explained by clonal expansion; or a mixture of both.
+Either several MCGs belong to one MLH, which could be explained by insufficient sampling of loci; several MLHs belong to one MCG, which could be explained by clonal expansion; or a mixture of both.
 Some studies have shown a correlation between MCG and MLH [@aldrich-wolfe2015genetic; @carbone1999patterns; @lehner2015genetic], whereas other studies have shown no apparent correlation, even on small spatial scales [@atallah2004high; @attanayake2012genetic; @attanayake2013sclerotinia].
@@
 sclerotiorum* from the USA to date, giving us a unique insight to continent-wide population structure and relationships between phenotypic and genotypic variables.
-Based on our analysis of the relationships between MCG and MLG, we found no clear evidence that the two are directly related, suggesting that MCG do not necessarily represent vegetative compatibility groups and thus should not be used as a proxy for identifying clones.
+Based on our analysis of the relationships between MCG and MLH, we found no clear evidence that the two are directly related, suggesting that MCG do not necessarily represent vegetative compatibility groups and thus should not be used as a proxy for identifying clones.
 Populations in wmn appear to show no significant differentiation when compared to their production field counterparts, suggesting that the wmn populations of *S.
@@
          col_types = cols(.default = col_character())) %>%
+  rename(MLH = MLG) %>%
   huxtable::as_huxtable(add_colnames = TRUE) %>%
@@
 
-Table: (\#tab:MLG-aggressiveness) Mean aggressiveness ratings for the first 10 MLG; groupings according to 95% family-wise confidence interval.
+Table: (\#tab:MLG-aggressiveness) Mean aggressiveness ratings for the first 10 MLH; groupings according to 95% family-wise confidence interval.
 
 ![
-Graph showing complex associations between Mycelial Compatibility Groups (MCG) (dotted nodes) and Multilocus Haplotypes (MLH) (full nodes) where the number in each node represents the MLG/MCG assignment.
+Graph showing complex associations between Mycelial Compatibility Groups (MCG) (dotted nodes) and Multilocus Haplotypes (MLH) (full nodes) where the number in each node represents the MLH/MCG assignment.
 Node size reflect the number of samples represented by each node (circle). 

### 2017-08-25 doc/RMD/by-year.Rmd
update colors on by-year

@@
 set.seed(2017-08-18)
-regyear.dapc <- xvalDapc(tab(dat11cc), pop(dat11cc), n.pca.max = 20, n.rep = 1000)$DAPC
+regyear.dapc <- xvalDapc(tab(dat11cc), pop(dat11cc), n.pca = 4:20, n.rep = 1000)$DAPC
 regyear.dapc
@@
 
+yearscale <- viridis::viridis(nlevels(strata(dat11cc)$Year), option = "C")
+names(yearscale) <- sort(levels(strata(dat11cc)$Year))
 gg_region_year <- ggscatter(regyear.dapc, strata(dat11cc)) +
@@
   guides(color = guide_legend(nrow = 4)) +
+  scale_color_manual(values =  yearscale, breaks = names(yearscale)) +
   theme(legend.direction = "horizontal") +
@@
   theme(legend.justification = "bottom") +
+  scale_color_manual(values =  yearscale, breaks = names(yearscale)) +
   theme(legend.box.margin = unit(c(0, 0, 0, 0), "lines")) +
@@
 ```
-```{r MIplot, fig.width = 10, fig.height = 12}
+```{r MIplot, fig.width = 12, fig.height = 14}
 gglMI <- suppressMessages(ggloc(loc, "MI", 2003:2012))

### 2017-08-28 doc/manuscript/manuscript.Rmd
reran RDA with new version of vegan; added WA text
The new version of vegan apparently treats the marginal anova a
bit differently, but I see no change in the results which is :+1:

We got a response back from Phil Miklas RE the WA 2008 population:

Dear Zhian,

 The 2008 trial was conducted in a different section of the field.
 In 2002 and 2003 the pea pathologist added sclerotia to pea plots
 to induce white mold (it was unsuccessful).
 The sclerotia came from the seed millings from a seed company.
 I am trying to find out which company and crop (peas?), and will
 pass that information along

@@
 
-Despite the evidence that Mexico and CA contributing to a large amount of the population differentiation, Regions like WA still had a large amount of internal variation (Supplementary Information).
-The two distinct clusters for the WA Region showed that the 2008 population appeared different and, under further investigation, we found that the genotypes from this population were shared between CA and France (Fig.
+Despite the evidence that Mexico and CA contributing to a large amount of the population differentiation, Regions like WA still had a large amount of internal variation.
+The two distinct clusters for the WA Region showed that the 2008 population appeared different and, under further investigation, we found that the haplotypes from this population were shared between CA and France (Fig.
 \@ref(community-graph), \@ref(DAPC-RY), \@ref(DAPC-RY-FULL)).
+The isolates from WA in 2003--2005, and 2008 all came from the same wmn.
+All of the isolates in 2003--2005 came from the same field location that were cropped with dry bean since 2002 (Miklas, Phil Pers. comm.).
+While white mold pressure had been sufficient for infection in previous years, sclerotia from the central bean company was added to the innoculum load. 
+The 2008 isolates, on the other hand, came from a field that was previously cropped with brassica, sundgrass, peas, and potatoes.
+
 From the DAPC, we extracted 6 loci that had alleles contributing more than the upper 95% quantile of the principal component loadings to either of the first two discriminant axes and found that the alleles with an allele frequency of > 50% in 2007 were reduced to < 10% in 2008, replaced by an allele that was > 2 mutational steps away (Supplementary Information).

### 2017-08-28 doc/manuscript/manuscript.Rmd
Attempt to address WA 2008
I'm feeling a bit doughy, so this will definitely
need to be rewritten.

@@
 The isolates from WA in 2003--2005, and 2008 all came from the same wmn.
-All of the isolates in 2003--2005 came from the same field location that were cropped with dry bean since 2002 (Miklas, Phil Pers. comm.).
-While white mold pressure had been sufficient for infection in previous years, sclerotia from the central bean company was added to the innoculum load. 
-The 2008 isolates, on the other hand, came from a field that was previously cropped with brassica, sundgrass, peas, and potatoes.
+All of the isolates in 2003--2005 came from the same NE field location that were cropped with dry bean since 2002 (Miklas, Phil Pers. comm.).
+The 2008 isolates, on the other hand, came from a SE field that was previously cropped with brassica, sundgrass, peas, beans, and potatoes.
+Both of these fields were inoculated with sclerotia in 2002, the NE field with sclerotia provided by a commercial bean producer and the SE field with sclerotia from peas (although this was thought to be unsuccessful).
+Despite this information, it is still unclear what has contributed to the differentiation of the 2008 population from WA.
+While the evidence my suggest host as being a factor, previous studies have shown no significant differentiation across host species [@aldrich-wolfe2015genetic].
 

### 2017-08-29 doc/manuscript/manuscript.Rmd
Add Serena's Email

@@
   - name: Serena McCoy
-    email: ¯\_(ツ)_/¯
+    email: serena.mccoy@icloud.com
     affiliation: 1

### 2017-08-29 doc/manuscript/manuscript.Rmd
update discussion

@@
 Currently, there are no known resistant cultivars of dry bean commercially available [@otto-hanson2011variation].
-Organized breeding efforts have been using a common-garden approach with white mold screening nurseries in dry bean production areas across the United States with additional sites in Mexico, France, and Australia [@steadman2003identification; @steadman2004identification; @steadman2005identification; @steadman2006identification; @otto2007identification].
+Organized breeding efforts have been using a common-garden approach with white mold screening nurseries in dry bean production areas across the United States with additional sites in Mexico, France, and Australia [@steadman2003identification; @steadman2004identification; @steadman2005identification; @steadman2006identification; @otto2007identification; @otto2008identification; @mccoy2009use].
 These white mold screening nurseries use no chemical or cultural treatments against *S. sclerotiorum* and employ standardized protocols for screening new cultivars for resistance to white mold. 
@@
 The white mold screening nurseries were 5m x 10m in size and maintained without applying fungicides to observe natural incidence of white mold.
-The nursery plots were incorporated with a basal dressing of N:P:K = 1:3:2 and side dressing of 0:3:2 during the growing season [@steadman2003identification; @steadman2004identification; @steadman2005identification; @steadman2006identification; @otto2007identification].
+The nursery plots were incorporated with a basal dressing of N:P:K = 1:3:2 and side dressing of 0:3:2 during the growing season [@steadman2003identification; @steadman2004identification; @steadman2005identification; @steadman2006identification; @otto2007identification; @otto2008identification; @mccoy2009use].
 Sampling was carried out by collecting diseased tissue in zig-zag transects across field plots. 
@@
 Our results suggest that, broadly, populations from white mold screening nurseries reflect the populations of the surrounding regions, indicating that resistance screening may be successful at a within regions.
-We found significant population differentiation by geographic region, mainly differentiated into three broad North American groups based on shared genotypes and posterior groupings, a Coastal Region, Midwestern Region, and Mexico.
+We found significant population differentiation by geographic region and year, mainly differentiated into three broad North American groups based on shared genotypes and posterior groupings, a Coastal Region, Midwestern Region, and Mexico.
 To date, with 366 isolates, this is the largest single population genetic study of *S. sclerotiorum* assessing population structure within managed and unmanaged agricultural environments. 
@@
 One long-held assumption was that MCGs (as determined via barrage reaction) represent vegetative compatibility groups (VCGs) [@schafer2006optimized; @kohn1990mycelial; @lehner2015genetic], which are known to have a genetic component [@saupe2000molecular; @hall2010evolution; @strom2016genomes].
-While our results appear to contradict this assumption, our protocol for assessing MCGs utilized a different substrate for these reactions and, thus may not accurately reflect the true MCG dynamics. 
-However, it has been demonstrated in several Ascomycetes---including *Neurospora crassa* [@micali2003independence], *Sclerotinia homoecarpa* [@jo2008reassessment], *Verticillium dahliae* [@papaioannou2014barrage], and *S.
+While our protocol for assessing MCGs utilized a different substrate for the MCG reactions, the patterns we observe are not dissimilar from what have previously been reported in the literature.
+It has been demonstrated in several Ascomycetes---including *Neurospora crassa* [@micali2003independence], *Sclerotinia homoecarpa* [@jo2008reassessment], *Verticillium dahliae* [@papaioannou2014barrage], and *S.
 sclerotiorum* [@ford1995heterokaryon]---that barrage reactions are independent from stable anastomosis.

### 2017-08-29 doc/manuscript/manuscript.Rmd
update text about wmn

@@
 
-Despite the evidence that Mexico and CA contributing to a large amount of the population differentiation, Regions like WA still had a large amount of internal variation.
-The two distinct clusters for the WA Region showed that the 2008 population appeared different and, under further investigation, we found that the haplotypes from this population were shared between CA and France (Fig.
+These patterns may reflect the degree of innoculum exchange among white mold screening nurseries.
+One possible explanation for these overlapping patterns of uncertainty could be that sclerotia from the source fields (where breeders would bulk their seed) were being transplanted to test fields.
+Because records of these fields were not kept for these purposes, it was not possible to test this hypothesis.
+Another explanation could be that the breeders were all obtaining innoculum from the same source.
+As the purpose of the wmn was to test the resistance of new bean varieties, the breeders needed to ensure that there was enough innoculum pressure in the field to induce disease. 
+To acheive this, it was not uncommon for breeders to innoculate their fields with sclerotia from anyone who had it. 
+These hypotheses, however, do not explain why regional fields are not significantly differentiated and share genotypes with wmn.
+
+Despite the evidence that Mexico and CA contributed to much of the population differentiation, Regions like WA still had a large amount of internal variation.
+The two distinct clusters for the WA Region showed that the 2008 population appeared differentiated and, under further investigation, we found that all the haplotypes from this year were shared between CA and France (Fig.
 \@ref(community-graph), \@ref(DAPC-RY), \@ref(DAPC-RY-FULL)).
-The isolates from WA in 2003--2005, and 2008 all came from the same wmn.
-All of the isolates in 2003--2005 came from the same NE field location that were cropped with dry bean since 2002 (Miklas, Phil Pers. comm.).
-The 2008 isolates, on the other hand, came from a SE field that was previously cropped with brassica, sundgrass, peas, beans, and potatoes.
+All of the isolates from WA in 2003--2005, and 2008 came from the same wmn; those in 2003--2005 came a NE field location cropped with dry bean since 2002, and those in 2008 from a SE field that was previously cropped with brassica, sundgrass, peas, beans, and potatoes (Miklas, Phil Pers.
+comm.).
 Both of these fields were inoculated with sclerotia in 2002, the NE field with sclerotia provided by a commercial bean producer and the SE field with sclerotia from peas (although this was thought to be unsuccessful).
@@
 
-From the DAPC, we extracted 6 loci that had alleles contributing more than the upper 95% quantile of the principal component loadings to either of the first two discriminant axes and found that the alleles with an allele frequency of > 50% in 2007 were reduced to < 10% in 2008, replaced by an allele that was > 2 mutational steps away (Supplementary Information).
-This could be due to several reasons which include mis-labeling of source material or introduction of migrant sclerotia in plant material.
+<!-- From the DAPC, we extracted 6 loci that had alleles contributing more than the upper 95% quantile of the principal component loadings to either of the first two discriminant axes and found that the alleles with an allele frequency of > 50% in 2007 were reduced to < 10% in 2008, replaced by an allele that was > 2 mutational steps away (Supplementary Information).
+-->
+<!-- This could be due to several reasons which include mis-labeling of source material or introduction of migrant sclerotia in plant material. -->
+
 Except for the WA Region, populations that were sampled across several years appeared to be relatively stable over time with overlapping distributions in the DAPC (i.e.

### 2017-08-29 doc/RMD/by-year.Rmd
update some figures for results

@@
                          Pval    = rev(amt$pvalue), 
+                         Sigma   = am$componentsofcovariance[-tot, 1],
                          Phi     = rev(am$statphi$Phi[-tot])))
@@
   rename(`$\\Phi$ statistic` = ps) %>%
+  rename(`$\\sigma^2$` = Sigma) %>%
   rename(`% variation` = `Percent variation`) %>%
@@
   rename(`*P*` = P) %>%
-  select(c(1:4, 6, 5)) %>%
+  select(c(1:3, 6, 4, 7, 5)) %>%
   huxtable::as_huxtable(add_colnames = TRUE) %>% 
-  huxtable::set_col_width(c(0.8, 0.15, 0.15, 0.35, 0.5, 0.15)) %>% 
-  huxtable::set_align(huxtable::everywhere, 2:6, "right") %>% 
+  huxtable::set_col_width(c(0.7, 0.1, 0.12, 0.27, 0.31, 0.42, 0.15)) %>% 
+  huxtable::set_align(huxtable::everywhere, 2:7, "right") %>% 
   huxtable::set_number_format(huxtable::everywhere, 2, 0) %>%
-  huxtable::set_number_format(huxtable::everywhere, 6, 3) %>%
-  huxtable::set_number_format(huxtable::everywhere, 5, 4) %>%
-  huxtable::print_md(max_width = 80)
+  huxtable::set_number_format(huxtable::everywhere, c(4, 6), 3) %>%
+  huxtable::set_number_format(huxtable::everywhere, 7, 4) %>%
+  huxtable::print_md(max_width = 93)
 ```

### 2017-08-29 doc/RMD/wmn-differentiation.Rmd
update some figures for results

@@
                          Pval    = rev(amt$pvalue), 
+                         # Sigma   = rev(amt$)
                          Phi     = rev(am$statphi$Phi[-tot])))

### 2017-08-29 doc/manuscript/manuscript.Rmd
update some figures for results

@@
   rename(ps = `Phi statistic`) %>%
+  select(-Sigma) %>%
   # mutate_if(is.numeric, format, digits = 3) %>%

### 2017-08-30 doc/manuscript/manuscript.Rmd
Add comment on potential pathway for spread

@@
 The graph walking algorithm was able to divide the network into three groups, but had a modularity of 0.17, which indicates that the groups are only weakly differentiated.
-This indicates that multilocus genotypes are widespread across the United States, giving more evidence for a clonal population structure.
+
+The widespread nature of multilocus genotypes in both wmn and production fields with relatively small values of $P_{sex}$ may indicate the spread of inoculum between regions. 
+It has previously been shown that infections can be transmitted through seed [@botelho2013performance], but seedborne transmission is thought to be of insignificant epidemiological importance [@strausbaugh2003management].
+Thus, we hypothesize that shared genotypes between populations may arise due to transmission events of seed or sclerotia.
+However, because we observe population structure based on Region, we speculate that these transmission events are rare.
 
@@
 
-These patterns may reflect the degree of innoculum exchange among white mold screening nurseries.
-One possible explanation for these overlapping patterns of uncertainty could be that sclerotia from the source fields (where breeders would bulk their seed) were being transplanted to test fields.
-Because records of these fields were not kept for these purposes, it was not possible to test this hypothesis.
-Another explanation could be that the breeders were all obtaining innoculum from the same source.
-As the purpose of the wmn was to test the resistance of new bean varieties, the breeders needed to ensure that there was enough innoculum pressure in the field to induce disease. 
-To acheive this, it was not uncommon for breeders to innoculate their fields with sclerotia from anyone who had it. 
-These hypotheses, however, do not explain why regional fields are not significantly differentiated and share genotypes with wmn.
-
 Despite the evidence that Mexico and CA contributed to much of the population differentiation, Regions like WA still had a large amount of internal variation.
@@
 
-<!-- From the DAPC, we extracted 6 loci that had alleles contributing more than the upper 95% quantile of the principal component loadings to either of the first two discriminant axes and found that the alleles with an allele frequency of > 50% in 2007 were reduced to < 10% in 2008, replaced by an allele that was > 2 mutational steps away (Supplementary Information).
--->
-<!-- This could be due to several reasons which include mis-labeling of source material or introduction of migrant sclerotia in plant material. -->
-
-Except for the WA Region, populations that were sampled across several years appeared to be relatively stable over time with overlapping distributions in the DAPC (i.e.
+With the exception of the WA Region, populations that were sampled across several years appeared to be relatively stable over time with overlapping distributions in the DAPC (i.e.
 NE and MI, Fig.

### 2017-08-30 doc/manuscript/manuscript.Rmd
change to pandoc-native citation engine

@@
 bibliography: ssc_bibliography.bib
+csl: apa.csl
 output: 
   rticles::peerj_article:
-    citation_package: natbib
+    citation_package: none
     base_format: bookdown::pdf_document2 # for using \@ref()
@@
 
+\newpage
+
 # Supplementary Information {-}
@@
 \newpage
+
+# References {-}

### 2017-08-30 doc/manuscript/manuscript.Rmd
add osf link

@@
 
-All scripts, data, and resources used to generate the results presented in this publication (including supplementary information) are fully reproducible and available at https://github.com/everhartlab/sclerotinia-366/ and archived at ZENODO.
+All scripts, data, and resources used to generate the results presented in this publication (including supplementary information) are fully reproducible and available at The Open Science Framework https://osf.io/ejb5y.
 

### 2017-09-06 doc/manuscript/manuscript.Rmd
update some text

@@
 Currently, there are no known resistant cultivars of dry bean commercially available [@otto-hanson2011variation].
-Organized breeding efforts have been using a common-garden approach with white mold screening nurseries in dry bean production areas across the United States with additional sites in Mexico, France, and Australia [@steadman2003identification; @steadman2004identification; @steadman2005identification; @steadman2006identification; @otto2007identification; @otto2008identification; @mccoy2009use].
+Organized breeding efforts have used a common-garden approach with white mold screening nurseries in dry bean production areas across the United States with additional sites in Mexico, France, and Australia [@steadman2003identification; @steadman2004identification; @steadman2005identification; @steadman2006identification; @otto2007identification; @otto2008identification; @mccoy2009use].
 These white mold screening nurseries use no chemical or cultural treatments against *S. sclerotiorum* and employ standardized protocols for screening new cultivars for resistance to white mold. 
@@
 Mycelial compatibility has been used as a proxy for vegetative compatibility, a fungal trait controlled by several independent genes controlling the ability for two hyphae to fuse and grow as a single unit [@leslie1993fungal; @schafer2006optimized].
-Because of the genetic connection to vegetative compatibility, two isolates that are mycelially compatible have been considered to be clones, but correlation with genetic markers such as microsatellites have shown mixed results [@attanayake2012genetic; @lehner2017independently; @leslie1993fungal].
+Because of the genetic connection to vegetative compatibility, two isolates that are mycelially compatible were considered clones [@leslie1993fungal]; but correlation with genetic markers, such as microsatellites, have shown mixed results [@attanayake2012genetic; @lehner2017independently].
 

### 2017-09-06 doc/manuscript/manuscript.Rmd

@@
 Loci with compound repeats were removed for the reported statistics, but were processed and included in data analyses to ensure the integrity of the results with 11 loci as compared to 16.
-We assessed the power of our 11 markers by generating a genotype accumulation curve in the R package *poppr* version `r packageVersion("poppr")`, looking for evidence of saturation, which would indicate that loci were sufficiently sampled to adequately represent the full set of genotypes [@arnaud2007standardizing; @kamvar2015novel].
-To avoid including isolates potentially collected from the same plant (which increases the probability of collecting sclerotioa from the same point of infection more than once), data were clone-corrected on a hierarchy of Region/Source/Host/Year---meaning that one representative of each genotype was retained for each nested combination of Region, Source, Host, and Year---for subsequent analysis.
-We assessed genotypic diversity by calculating Stoddart and Taylor's index (*G*) [@stoddart1988genotypic], Shannon's index (*H*) [@shannon2001mathematical], Simpson's index ($\lambda$) [@simpson1949measurement], and evenness ($E_5$, *eMLG*) [@heck1975explicit; @hurlbert1971nonconcept; @pielou1975ecological; @grunwald2003analysis].
+We assessed the power of our 11 markers by generating a genotype accumulation curve in the R package *poppr* version `r packageVersion("poppr")`, looking for evidence of saturation, which would indicate that loci were sufficiently sampled to adequately represent the full set of haplotypes [@arnaud2007standardizing; @kamvar2015novel].
+To avoid including isolates potentially collected from the same plant (which increases the probability of collecting sclerotioa from the same point of infection more than once), data were clone-corrected on a hierarchy of Region/Source/Host/Year---meaning that one representative of each haplotype was retained for each nested combination of Region, Source, Host, and Year---for subsequent analysis.
+We assessed genotypic diversity by calculating Stoddart and Taylor's index (*G*) [@stoddart1988genotypic], Shannon's index (*H*) [@shannon2001mathematical], Simpson's index ($\lambda$) [@simpson1949measurement], and evenness ($E_5$, *eMLH*) [@heck1975explicit; @hurlbert1971nonconcept; @pielou1975ecological; @grunwald2003analysis].
 To assess the potential for random mating, we tested for linkage disequilibrium with the index of association, $I_A$ and its standardized version, $\bar{r}_d$ using 999 permutations [@agapow2001indices; @smith1993how; @brown1980multilocus].
@@
 
-### Analysis of Shared Multilocus Genotypes {-}
+### Analysis of Shared Multilocus Haplotypes {-}
 
-We wanted to evaluate patterns of connectivity between shared multilocus genotypes across geographic regions.
-We first tabulated the multilocus genotypes shared between at least two populations (defined as states or countries) with the *poppr* function `mlg.crosspop()` [@kamvar2014poppr].
-From these data, we constructed a graph with populations as nodes and shared genotypes as edges (connections) between nodes using the R packages *igraph* [@csardi2006igraph], *dplyr* version `r packageVersion("dplyr")` [@dplyr], and *purrr* version `r packageVersion("purrr")` [@purrr].
-Each node was weighted by the fraction of shared MLGs.
-Each edge represented a single MLG, but because a single MLG could be present in more than one population, that MLG would be represented by (*n*\*(*n*-1))/2 edges where n represents the number of populations crossed.
-Edges were weighted by $1-P_{sex}$, where $P_{sex}$ is the probability of encountering the same genotype via two independent meiotic events [@arnaud2007standardizing; @parks1993study].
+We wanted to evaluate patterns of connectivity between shared multilocus haplotypes across geographic regions.
+We first tabulated the multilocus haplotypes shared between at least two populations (defined as states or countries) with the *poppr* function `mlg.crosspop()` [@kamvar2014poppr].
+From these data, we constructed a graph with populations as nodes and shared haplotypes as edges (connections) between nodes using the R packages *igraph* [@csardi2006igraph], *dplyr* version `r packageVersion("dplyr")` [@dplyr], and *purrr* version `r packageVersion("purrr")` [@purrr].
+Each node was weighted by the fraction of shared MLHs.
+Each edge represented a single MLH, but because a single MLH could be present in more than one population, that MLH would be represented by (*n*\*(*n*-1))/2 edges where n represents the number of populations crossed.
+Edges were weighted by $1-P_{sex}$, where $P_{sex}$ is the probability of encountering the same haplotype via two independent meiotic events [@arnaud2007standardizing; @parks1993study].
 We then used the `cluster_optimal()` function from *igraph* to define communities [@csardi2006igraph].
@@
 ![
-Data show complex associations between Mycelial Compatibility Groups and Multilocus Genotypes.
+Data show complex associations between Mycelial Compatibility Groups and Multilocus Haplotypes.
 **A)** Barplot of Mycelial Compatibility Group (MCG) abundance in descending order.
 Singletons (46) were truncated, leaving 41 MCGs.
-White bars represent sample counts and grey bars represent counts of unique multilocus genotypes (MLG).
-The transparency of the bars represent the evenness of the distribution of the MLGs within a given MCG.
+White bars represent sample counts and grey bars represent counts of unique multilocus haplotypes (MLH).
+The transparency of the bars represent the evenness of the distribution of the MLHs within a given MCG.
 A dashed box surrounds the 8 most common MCGs representing > 51% of the data.
-**B)** Full graph-representation of the relationship between MCGs (open circles) and MLGs (filled circles).
+**B)** Full graph-representation of the relationship between MCGs (open circles) and MLHs (filled circles).
 Details in fig. \@ref(fullgraph). 
-**C)** A subset of **B** representing the 8 most common MCGs and their associated MLGs (dashed box in **A**).
-Filled nodes represent MLGs and open nodes represent MCGs.
-Numbers inside nodes are the MLG/MCG label. 
+**C)** A subset of **B** representing the 8 most common MCGs and their associated MLHs (dashed box in **A**).
+Filled nodes represent MLHs and open nodes represent MCGs.
+Numbers inside nodes are the MLH/MCG label. 
 Node (circle) area scaled to the number of samples represented (range: 1--73).
-Edges (arrows) point from MLG to MCG where the weight (thickness) of the edge represents the number of shared isolates (range: 1--19).
-Edges extending from MLGs displayed to other MCGs are not shown.
+Edges (arrows) point from MLH to MCG where the weight (thickness) of the edge represents the number of shared isolates (range: 1--19).
+Edges extending from MLHs displayed to other MCGs are not shown.
 ](../../results/figures/publication/mcg-subgraph-with-context.pdf){#biggraph width=100%}
@@
 Comparing these communities with Bruvo's genetic distance revealed an average distance of 0.451 among communities and an average distance of 0.437 within communities, showing no significant difference.
-When we assessed the number of times two different MLGs that are in the same MCG, considering these as potential heterothallic pairings that could result in sexual recombination, we found an average of `r signif(mean(in_reach$N), 3)` potential heterothallic parings per MLG.
-Representing just four isolates, 'MLG 75' had `r big_genotype$N` neighbors that shared the same mycelial compatibility group (Fig \@ref(biggraph), \@ref(fullgraph)). 
-Overall, there was no clear pattern to the association between MLG and MCGs.
+When we assessed the number of times two different MLHs that are in the same MCG, considering these as potential heterothallic pairings that could result in sexual recombination, we found an average of `r signif(mean(in_reach$N), 3)` potential heterothallic parings per MLH.
+Representing just four isolates, 'MLH 75' had `r big_haplotype$N` neighbors that shared the same mycelial compatibility group (Fig \@ref(biggraph), \@ref(fullgraph)). 
+Overall, there was no clear pattern to the association between MLH and MCGs.
 
-## Structure of Shared Multilocus Genotypes {-}
+## Structure of Shared Multilocus Haplotypes {-}
 
-We observed a total of 165 MLGs (215 with 16 loci).
-The most abundant MLG was represented by 27 isolates (Table \@ref(tab:mlg-table)) from five regions (NE, MI, WA, CO, and ND). 
-Within regions, the genotypes were relatively evenly distributed with moderate to high diversity (Table \@ref(tab:iatab)). 
-Of the 165 MLGs, 76 (46%) were found in at least two geographic regions except those found in WI (2), ID (1), and Mexico (18) (Fig. \@ref(community-graph)).
-Community analysis of the MLGs shared between populations revealed 4 communities with a modularity of 0.17 (Fig.
+We observed a total of 165 MLHs (215 with 16 loci).
+The most abundant MLH was represented by 27 isolates (Table \@ref(tab:mlg-table)) from five regions (NE, MI, WA, CO, and ND). 
+Within regions, the haplotypes were relatively evenly distributed with moderate to high diversity (Table \@ref(tab:iatab)). 
+Of the 165 MLHs, 76 (46%) were found in at least two geographic regions except those found in WI (2), ID (1), and Mexico (18) (Fig. \@ref(community-graph)).
+Community analysis of the MLHs shared between populations revealed 4 communities with a modularity of 0.17 (Fig.
 \@ref(community-graph)): A coastal community (CA, OR, WA, and NY), a midwest community (CO, NE, ND, MI), and an international community (MN, France, Australia).
-Analysis with 16 loci resulted in the removal of the NY node because it no longer shared a genotype with OR, but the same overall community structure was present with a modularity of 0.2 (Fig.
+Analysis with 16 loci resulted in the removal of the NY node because it no longer shared a haplotype with OR, but the same overall community structure was present with a modularity of 0.2 (Fig.
 \@ref(community-graph16)).
-The international community appears to be driven by MLG 4, which is shared between all three populations and has a $P_{sex}$ value of 2.87e^-5^, in contrast to the abundant MLG 25, which has a $P_{sex}$ value of 0.0168.
+The international community appears to be driven by MLH 4, which is shared between all three populations and has a $P_{sex}$ value of 2.87e^-5^, in contrast to the abundant MLH 25, which has a $P_{sex}$ value of 0.0168.
 
@@
 Our results suggest that, broadly, populations from white mold screening nurseries reflect the populations of the surrounding regions, indicating that resistance screening may be successful at a within regions.
-We found significant population differentiation by geographic region and year, mainly differentiated into three broad North American groups based on shared genotypes and posterior groupings, a Coastal Region, Midwestern Region, and Mexico.
+We found significant population differentiation by geographic region and year, mainly differentiated into three broad North American groups based on shared haplotypes and posterior groupings, a Coastal Region, Midwestern Region, and Mexico.
 To date, with 366 isolates, this is the largest single population genetic study of *S. sclerotiorum* assessing population structure within managed and unmanaged agricultural environments. 
@@
 
-## Mycelial Compatibility Groups and Multilocus Genotypes {-}
+## Mycelial Compatibility Groups and Multilocus Haplotypes {-}
 
-Of interest in this study was testing the association between multilocus genotypes (MLGs) and mycelial compatibility groups (MCGs), as the latter have historically been used as a proxy for clonal lineages [@leslie1993fungal; @otto-hanson2011variation; @kohn1990mycelial; @schafer2006optimized; @kohn1995clonal; @carbone1999patterns].
+Of interest in this study was testing the association between multilocus haplotypes (MLHs) and mycelial compatibility groups (MCGs), as the latter have historically been used as a proxy for clonal lineages [@leslie1993fungal; @otto-hanson2011variation; @kohn1990mycelial; @schafer2006optimized; @kohn1995clonal; @carbone1999patterns].
 Our results do not support this assumption.
 It can be seen in Fig.
-\@ref(biggraph)A that the most abundant MCG contains several MLGs, but the diversity of those MLGs are low as indicated by the evenness (transparency), which indicates that there is one dominant MLG ('MLG 25').
-What is not shown in Fig. \@ref(biggraph)A is the MLGs that are shared between MCGs. 
+\@ref(biggraph)A that the most abundant MCG contains several MLHs, but the diversity of those MLHs are low as indicated by the evenness (transparency), which indicates that there is one dominant MLH ('MLH 25').
+What is not shown in Fig. \@ref(biggraph)A is the MLHs that are shared between MCGs. 
 This is illustrated in both table \@ref(tab:mlg-table) and Fig. \@ref(biggraph)B,C. 
-It could be argued, however that 'MLG 25', with its high value of $P_{sex}$ represents different true MLGs across the five MCGs it occupies, but this doesn't account for the overall structure of Fig.
-\@ref(fullgraph) where, for example, 'MLG 75' ($P_{sex}$ = 1.81e^-4^) is compatible with `r big_genotype$N` other genotypes through three MCG when the population structure of *S.
+It could be argued, however that 'MLH 25', with its high value of $P_{sex}$ represents different true MLHs across the five MCGs it occupies, but this doesn't account for the overall structure of Fig.
+\@ref(fullgraph) where, for example, 'MLH 75' ($P_{sex}$ = 1.81e^-4^) is compatible with `r big_haplotype$N` other haplotypes through three MCG when the population structure of *S.
 sclerotiorum* is known to be clonal.
 
-Over the past few years, researchers have noticed inconsistencies among the relationship between MCGs and MLGs [@aldrich-wolfe2015genetic; @carbone1999patterns; @attanayake2012genetic; @lehner2015genetic].
-Either several MCGs belong to one MLG, which could be explained by insufficient sampling of loci; several MLGs belong to one MCG, which could be explained by clonal expansion; or a mixture of both.
-Some studies have shown a correlation between MCG and MLG [@aldrich-wolfe2015genetic; @carbone1999patterns; @lehner2015genetic], whereas other studies have shown no apparent correlation, even on small spatial scales [@atallah2004high; @attanayake2012genetic; @attanayake2013sclerotinia].
+Over the past few years, researchers have noticed inconsistencies among the relationship between MCGs and MLHs [@aldrich-wolfe2015genetic; @carbone1999patterns; @attanayake2012genetic; @lehner2015genetic].
+Either several MCGs belong to one MLH, which could be explained by insufficient sampling of loci; several MLHs belong to one MCG, which could be explained by clonal expansion; or a mixture of both.
+Some studies have shown a correlation between MCG and MLH [@aldrich-wolfe2015genetic; @carbone1999patterns; @lehner2015genetic], whereas other studies have shown no apparent correlation, even on small spatial scales [@atallah2004high; @attanayake2012genetic; @attanayake2013sclerotinia].
 
@@
 
-The results show several genotypes shared across populations, except for Mexico and two states that had fewer than three samples each (Fig. \@ref(community-graph)).
-Our network-based approach by treating the genotypes as edges and weighting each edge with the inverse of $P_{sex}$ treated the edges as springs connecting the populations with the strength proportional to the probability of obtaining the same genotype as a clone.
+The results show several haplotypes shared across populations, except for Mexico and two states that had fewer than three samples each (Fig. \@ref(community-graph)).
+Our network-based approach by treating the haplotypes as edges and weighting each edge with the inverse of $P_{sex}$ treated the edges as springs connecting the populations with the strength proportional to the probability of obtaining the same haplotype as a clone.
 This allowed us to use a graph walking algorithm to see how close the populations were simply based off of the proportion of clones they shared.
-The most abundant genotype was shared across four populations, but its high value of $P_{sex}$ meant that it did not contribute significantly to the overall structure. 
+The most abundant haplotype was shared across four populations, but its high value of $P_{sex}$ meant that it did not contribute significantly to the overall structure. 
 The graph walking algorithm was able to divide the network into three groups, but had a modularity of 0.17, which indicates that the groups are only weakly differentiated.
 
-The widespread nature of multilocus genotypes in both wmn and production fields with relatively small values of $P_{sex}$ may indicate the spread of inoculum between regions. 
+The widespread nature of multilocus haplotypes in both wmn and production fields with relatively small values of $P_{sex}$ may indicate the spread of inoculum between regions. 
 It has previously been shown that infections can be transmitted through seed [@botelho2013performance], but seedborne transmission is thought to be of insignificant epidemiological importance [@strausbaugh2003management].
-Thus, we hypothesize that shared genotypes between populations may arise due to transmission events of seed or sclerotia.
+Thus, we hypothesize that shared haplotypes between populations may arise due to transmission events of seed or sclerotia.
 However, because we observe population structure based on Region, we speculate that these transmission events are rare.
@@
 sclerotiorum* from the USA to date, giving us a unique insight to continent-wide population structure and relationships between phenotypic and genotypic variables.
-Based on our analysis of the relationships between MCG and MLG, we found no clear evidence that the two are directly related, suggesting that MCG do not necessarily represent vegetative compatibility groups and thus should not be used as a proxy for identifying clones.
+Based on our analysis of the relationships between MCG and MLH, we found no clear evidence that the two are directly related, suggesting that MCG do not necessarily represent vegetative compatibility groups and thus should not be used as a proxy for identifying clones.
 Populations in wmn appear to show no significant differentiation when compared to their production field counterparts, suggesting that the wmn populations of *S.
 sclerotiorum* may be considered representative of the surrounding regions.
-While we found no direct relationship between genotype and severity, it is evident that there is a gradient of severity by region, further supporting the need for screening in multiple locations.
+While we found no direct relationship between haplotype and severity, it is evident that there is a gradient of severity by region, further supporting the need for screening in multiple locations.
 
@@
 
-Table: (\#tab:MLG-aggressiveness) Mean aggressiveness ratings for the first 10 MLG; groupings according to 95% family-wise confidence interval.
+Table: (\#tab:MLG-aggressiveness) Mean aggressiveness ratings for the first 10 MLH; groupings according to 95% family-wise confidence interval.
 
 ![
-Graph showing complex associations between Mycelial Compatibility Groups (MCG) (dotted nodes) and Multilocus Genotypes (MLG) (full nodes) where the number in each node represents the MLG/MCG assignment.
+Graph showing complex associations between Mycelial Compatibility Groups (MCG) (dotted nodes) and Multilocus Haplotypes (MLH) (full nodes) where the number in each node represents the MLH/MCG assignment.
 Node size reflect the number of samples represented by each node (circle). 
-Edges (arrows) point from MLG to MCG where the weight (thickness) of the edge represents the number of samples shared. 
+Edges (arrows) point from MLH to MCG where the weight (thickness) of the edge represents the number of samples shared. 
 Node color represents the community assignment based on the walktrap algorithm with a maximum of four steps [@pons2006computing]. 

### 2017-09-06 doc/manuscript/manuscript.Rmd
update introduction

@@
 In this study, we analyze and characterize the genetic and phenotypic diversity of 366 *S.
-sclerotiorum* isolates collected between 2003 and 2012 from dry bean cultivars among different geographic locations in the United States, France, Mexico, and Australia.
+sclerotiorum* isolates collected over 10 years between 2003 and 2012 from dry bean cultivars among different geographic locations in the United States, France, Mexico, and Australia.
 We wanted to know if the *S. sclerotiorum* populations from white mold screening nurseries were representative of the fields within the same region. 
-As these nurseries are not treated with any chemical or cultural control of white mold, we hypothesized that these nurseries would represent the natural population of *S. sclerotiorum*.
+As these nurseries were not treated with any chemical or cultural control of white mold, we hypothesized that these nurseries would represent the natural population of *S. sclerotiorum*.
 Furthermore, we wanted to investigate the potential effect of cultivar on genetic diversity by assessing three dry bean cultivars with different levels of resistance, Beryl (great northern bean, susceptible), Bunsi (navy bean, low susceptibility), and G122 (cranberry bean, partial resistance) [@otto-hanson2011variation].
 We additionally wanted to determine categorical or phenotypic variables that best predicted genetic structure and if there was correlation between multilocus haplotype and mycelial compatibility group.
+Knowing what variables predict genetic structure can help direct breeding efforts.
 By investigating these aims, we hope to effectively describe the population structure of *S.

### 2017-09-06 doc/manuscript/manuscript.Rmd
updte collection

@@
 
-Several of the isolates used for this study were collected and reported in previous studies and were collected using the same methods [@otto-hanson2011variation].
-Broadly, isolates were collected from two sources, white mold screening nurseries (wmn) or producer fields.
-The white mold screening nurseries were 5m x 10m in size and maintained without applying fungicides to observe natural incidence of white mold.
+Several of the isolates used for this study were collected as reported in previous studies using the same methods [@otto-hanson2011variation].
+Broadly, isolates were collected from two sources: white mold screening nurseries (wmn) or producer fields.
+<!-- SEE: Insect or weed control? -->
+White mold screening nurseries were 5m x 10m in size and maintained without application of fungicides to observe natural incidence of white mold.
 The nursery plots were incorporated with a basal dressing of N:P:K = 1:3:2 and side dressing of 0:3:2 during the growing season [@steadman2003identification; @steadman2004identification; @steadman2005identification; @steadman2006identification; @otto2007identification; @otto2008identification; @mccoy2009use].
-Sampling was carried out by collecting diseased tissue in zig-zag transects across field plots. 
+
+Sampling was carried out by collecting sclerotia from diseased tissue in zig-zag transects across field plots. 
 Because sampling depended on disease incidence, the number of samples isolated varied from year to year.
@@
 sclerotiorum* were collected over several years from grower fields and/or wmn in 11 states of the United States of America, Mexico, France, and Australia (Table \@ref(tab:isolate-table)).
-After collection, the isolates were stored in petri plates lined with filter paper, then stored at $20^{\circ}$F or -$4^{\circ}$C. 
-Sclerotia were triple rinsed with (i) 50% Clorox bleach/50% double-distilled (dd)H2O solution for 3 min, (ii) ddH2O rinse for 3 min, and (iii) final ddH2O rinse for 3 min, then plated on water agar plates (16g of Bacto agar per liter of ddH2O), with four to five sclerotia of each isolate separated on each plate and stored on the bench top at room temperature for 5 to 6 days An 8-mm plug from a 5- or 6-day-old culture was transferred from the advancing margin of the mycelia onto a plate of Difco potato dextrose agar (PDA at 39 g/liter of ddH2O) [@otto-hanson2011variation].
+After collection, sclerotia were stored in Petri plates lined with filter paper, then stored at $20~^{\circ}$F or -$4~^{\circ}$C. 
+Sclerotia were surface-sterilized with 50% sodium hypochlorite (at least 6% NaOCL, The Clorox Company, Oakland, CA) solution for 3 min, and double rinsed with ddH$_2$O 3 min.
+The sterilized sclerotia were then plated on water agar plates (16g of Bacto agar per liter of ddH$_2$O), with four to five sclerotia of each isolate separated on each plate and stored on the counter top at room temperature for 5 to 6 days.
+An 8-mm plug from a 5- or 6-day-old culture was transferred from the advancing margin of the mycelia onto a plate of Difco potato dextrose agar (PDA at 39 g/liter of ddH$_2$O) [@otto-hanson2011variation].
 In combination with the 156 isolates described previously, we collected 210 isolates for a total of 366 isolates [@otto-hanson2011variation].
@@
 Mycelial was collected in a buchner funnel, agar plugs removed, lyophilized and pulverized manually in whirl-pak bags. 
-Lyophilized mycelia was then stored in microcentrifuge tubes at -$20^{\circ}$C until needed for DNA extraction.
+Lyophilized mycelia was then stored in microcentrifuge tubes at -$20~^{\circ}$C until needed for DNA extraction.
 DNA from 25mg of pulverized mycelia was purified using a phenol-chloroform extraction method followed by alcohol precipitation, suspending the DNA in 200$\mu$l TE [@sambrook1989molecular].
-Suspended DNA was stored at $4^{\circ}$C until genotyping.
+Suspended DNA was stored at $4~^{\circ}$C until genotyping.
 

### 2017-09-06 doc/manuscript/manuscript.Rmd
update methods to data processing

@@
 Sclerotia were surface-sterilized with 50% sodium hypochlorite (at least 6% NaOCL, The Clorox Company, Oakland, CA) solution for 3 min, and double rinsed with ddH$_2$O 3 min.
-The sterilized sclerotia were then plated on water agar plates (16g of Bacto agar per liter of ddH$_2$O), with four to five sclerotia of each isolate separated on each plate and stored on the counter top at room temperature for 5 to 6 days.
+The sterilized sclerotia were then plated on water agar plates (16g of Bacto agar per liter of ddH$_2$O, BD Diagnostic Systems, Sparks, MD), with four to five sclerotia of each isolate separated on each plate and stored on the counter top at room temperature for 5 to 6 days.
 An 8-mm plug from a 5- or 6-day-old culture was transferred from the advancing margin of the mycelia onto a plate of Difco potato dextrose agar (PDA at 39 g/liter of ddH$_2$O) [@otto-hanson2011variation].
@@
 
-MCG was determined as described previously through co-culturing pairs of 2-day-old isolates 2.5cm apart on Diana Sermons (DS) Medium [@cubeta2001mycelial]. 
+MCG was determined as described previously through co-culturing pairs of 2-day-old isolates 2.5 cm apart on Diana Sermons (DS) Medium [@cubeta2001mycelial]. 
 Incompatibility of different MCGs resulted in formation of a mycelial barrier of dead tissue (termed a barrage line) accompanied by formation of sclerotia on either side of the barrage line, indicating the limits of the isolates' growth [@otto-hanson2011variation; @kohn1990mycelial; @leslie1993fungal].
@@
 Briefly, the straw test uses 21-day-old G122 plants as the host in a greenhouse setting. 
-Clear drinking straws cut to 2.5cm and sealed were used to place two mycelial plugs of inoculum on the host plant after removing all plant growth &gt;2.5cm above the fourth node.
-Measurements were taken after 8 days using the Modified Petzoldt and Dickson scale where a score of 1--9 where 1 is no disease and 9 is plant death [@petzoldt1996straw; @teran2006modified].
+Clear drinking straws cut to 2.5 cm and sealed were used to place two mycelial plugs of inoculum on the host plant after removing all plant growth &gt;2.5 cm above the fourth node.
+Measurements were taken 8 days later using the Modified Petzoldt and Dickson scale of 1--9, where 1 is no disease and 9 is plant death [@petzoldt1996straw; @teran2006modified].
 
@@
 Prior to DNA extraction, isolates were grown on PDA and plugs were subsequently transferred to Potato Dextrose Broth (PDB) where they were grown until there was significant mycelial growth, but before the mycelial mat became solidified (4--5 days).
-Mycelial was collected in a buchner funnel, agar plugs removed, lyophilized and pulverized manually in whirl-pak bags. 
+Mycelia for each plate was collected in a buchner funnel, agar plugs removed, lyophilized and pulverized manually in Whirl-pak® HDPE sampling bags (Sigma-Aldrich, St. Louis, MO). 
 Lyophilized mycelia was then stored in microcentrifuge tubes at -$20~^{\circ}$C until needed for DNA extraction.
-DNA from 25mg of pulverized mycelia was purified using a phenol-chloroform extraction method followed by alcohol precipitation, suspending the DNA in 200$\mu$l TE [@sambrook1989molecular].
+DNA from 25mg of pulverized mycelia was purified using a phenol-chloroform extraction method followed by alcohol precipitation and evaporation, suspending the DNA in 200$\mu$l TE [@sambrook1989molecular].
 Suspended DNA was stored at $4~^{\circ}$C until genotyping.
 
-We genotyped *S. sclerotiorum* isolates using 16 microsatellite primer pairs developed previously [@sirjusingh2001characterisation].
-PCR was carried out as described using primers labeled with FAM fluorophore, and resulting amplicons were first resolved in a 1.5% agarose gel stained with ethidium bromide to ensure product was within the expected size range prior to capillary electrophoresis.
-
-Capillary electrophoresis (fragment analysis) of amplicons, with size standard GeneScan™ 500 LIZ®, was performed using an ABI 3730 genetic analyzer (Life Technologies Corporation, Carlsbad, CA) at the Michigan State University Genomic Sequencing Center.
+We genotyped each *S. sclerotiorum* isolate using 16 microsatellite primer pairs developed previously [@sirjusingh2001characterisation].
+PCR was carried out as described previously, using primers labeled with FAM fluorophore.
+Resulting amplicons were first resolved in a 1.5% agarose gel stained with ethidium bromide to ensure product was within the expected size range prior to capillary electrophoresis.
+Capillary electrophoresis (fragment analysis) of amplicons, with size standard GeneScan™ 500 LIZ®, was performed using an ABI 3730 genetic analyzer (Life Technologies Corporation, Carlsbad, CA) at the Michigan State University Genomic Sequencing Center (East Lansing, MI).
 Alleles were scored using PeakScanner version 1.0 (Life Technologies Corporation, Carlsbad, CA) and recorded manually in an excel sheet.
@@
 
-All analyses were performed on this data set unless otherwise mentioned.
+<!-- All analyses were performed on this data set unless otherwise mentioned. -->
 All data processing and analyses were performed in `r R.version.string %>% strsplit("\\(") %>% map_chr(1)`[@R] and are openly available and reproducible at https://github.com/everhartlab/sclerotinia-366/.
 Of the 16 microsatellite loci genotyped, five included compound repeats, which made it challenging to accurately/confidently bin alleles into fragment sizes expected for each locus based on the described repeat motif.
-<!-- add something about repeat length consistency  -->
-Loci with compound repeats were removed for the reported statistics, but were processed and included in data analyses to ensure the integrity of the results with 11 loci as compared to 16.
+Loci with compound repeats were removed for the reported statistics.
+To ensure the integrity of the results we additionally processed these loci and included them in concurrent analyses.
 We assessed the power of our 11 markers by generating a genotype accumulation curve in the R package *poppr* version `r packageVersion("poppr")`, looking for evidence of saturation, which would indicate that loci were sufficiently sampled to adequately represent the full set of haplotypes [@arnaud2007standardizing; @kamvar2015novel].
-To avoid including isolates potentially collected from the same plant (which increases the probability of collecting sclerotioa from the same point of infection more than once), data were clone-corrected on a hierarchy of Region/Source/Host/Year---meaning that one representative of each haplotype was retained for each nested combination of Region, Source, Host, and Year---for subsequent analysis.
-We assessed genotypic diversity by calculating Stoddart and Taylor's index (*G*) [@stoddart1988genotypic], Shannon's index (*H*) [@shannon2001mathematical], Simpson's index ($\lambda$) [@simpson1949measurement], and evenness ($E_5$, *eMLH*) [@heck1975explicit; @hurlbert1971nonconcept; @pielou1975ecological; @grunwald2003analysis].
+To avoid including isolates potentially collected from the same plant (which increases the probability of collecting sclerotia from the same point of infection more than once), data were clone-corrected on a hierarchy of Region/Source/Host/Year---meaning that duplicated genotypes were reduced to a single observation when they were collected in the same year from the same host cultivar located in the same source field (wmn or producer)---for subsequent analysis.
+We assessed haplotype diversity by calculating Stoddart and Taylor's index (*G*) [@stoddart1988genotypic], Shannon's index (*H*) [@shannon2001mathematical], Simpson's index ($\lambda$) [@simpson1949measurement], and evenness ($E_5$, *eMLH*) [@heck1975explicit; @hurlbert1971nonconcept; @pielou1975ecological; @grunwald2003analysis].
 To assess the potential for random mating, we tested for linkage disequilibrium with the index of association, $I_A$ and its standardized version, $\bar{r}_d$ using 999 permutations [@agapow2001indices; @smith1993how; @brown1980multilocus].
-Both genotypic diversity and linkage disequilibrium were calculated in *poppr* [@kamvar2014poppr].
+Both haplotype diversity and linkage disequilibrium were calculated in *poppr* [@kamvar2014poppr].
 

### 2017-09-06 doc/manuscript/manuscript.Rmd
update to ANOVA and network analysis

@@
 
-RDA2R2    <- prettify(caplist1$R2$adj.r.squared)/100
-RDA1R2    <- prettify(caplist2$R2$adj.r.squared)/100
+RDA1R2    <- prettify(caplist1$R2$adj.r.squared)/100
+RDA2R2    <- prettify(caplist2$R2$adj.r.squared)/100
 anova1res <- setNames(caplist1$anova$`Pr(>F)`, rownames(caplist1$anova)) %>% na.omit()
@@
 To avoid including isolates potentially collected from the same plant (which increases the probability of collecting sclerotia from the same point of infection more than once), data were clone-corrected on a hierarchy of Region/Source/Host/Year---meaning that duplicated genotypes were reduced to a single observation when they were collected in the same year from the same host cultivar located in the same source field (wmn or producer)---for subsequent analysis.
-We assessed haplotype diversity by calculating Stoddart and Taylor's index (*G*) [@stoddart1988genotypic], Shannon's index (*H*) [@shannon2001mathematical], Simpson's index ($\lambda$) [@simpson1949measurement], and evenness ($E_5$, *eMLH*) [@heck1975explicit; @hurlbert1971nonconcept; @pielou1975ecological; @grunwald2003analysis].
+We assessed haplotype diversity by calculating Stoddart and Taylor's index (*G*) [@stoddart1988genotypic], Shannon's index (*H*) [@shannon2001mathematical], Simpson's index ($\lambda$) [@simpson1949measurement], evenness ($E_5$), and the expected number of multilocus haplotypes (*eMLH*) [@heck1975explicit; @hurlbert1971nonconcept; @pielou1975ecological; @grunwald2003analysis].
 To assess the potential for random mating, we tested for linkage disequilibrium with the index of association, $I_A$ and its standardized version, $\bar{r}_d$ using 999 permutations [@agapow2001indices; @smith1993how; @brown1980multilocus].
@@
 The distance matrix we used as the response variable was generated using Bruvo's genetic distance from clone-corrected data (procedure described above) as implemented in *poppr*, which employed a stepwise mutation model for microsatellite data [@bruvo2004simple; @kamvar2014poppr].
-Because aggressiveness measures differed with each isolate, aggressiveness first averaged across clone-corrected isolates.
+Because aggressiveness measures differed between isolates that were reduced to a single observation during clone-correction, aggressiveness was first averaged across clone-corrected isolates.
 To identify explanatory variable(s) correlated with genetic variation, a forward-backward selection process was applied with the vegan function `ordistep()`.
@@
 
-We used ANOVA to assess if aggressiveness could be explained by Region, MCG, or multilocus haplotype (MLH). 
-To minimize complications due to small sample sizes, we chose the top 10 MCGs, representing 56.5% of the isolates collected, the top 10 MLHs representing 26.7% of the isolates, and populations with more than five isolates.
-Between-group differences were assessed using Tukey's HSD test with an alpha of 0.05 using the `HSD.test()` function in the package *agricolae* version `r packageVersion("agricolae")` [@mendiburu2015agricolae].
+We used ANOVA to assess if aggressiveness was significantly different with respect to Region, MCG, or multilocus haplotype (MLH). 
+To minimize complications due to small sample sizes, we chose the top 10 MCGs, representing 56.5% of the isolates collected, the 10 most abundant MLHs representing 26.7% of the isolates, and populations with more than five isolates.
+If ANOVA results were significantly different at $\alpha$ = 0.05, pairwise differences were assessed using Tukey's HSD test ($\alpha$ = 0.05) using the `HSD.test()` function in the package *agricolae* version `r packageVersion("agricolae")` [@mendiburu2015agricolae].
 
@@
 We wanted to assess if there was correlation between MLHs and MCGs. 
-In simple terms, this was done using a network-based approach where both MLHs and MCGs were considered nodes and the number of isolates in which they were found together was the strength of the connection between an MLH and and MCG node.
+This was performed using a network-based approach where both MLHs and MCGs were considered nodes and the number of isolates in which they were found together was the strength of the connection between an MLH and and MCG node.
 The network-based approach allowed us to assess how "clumpy" the relationships between MLHs and MCGs were.
-In more specific terms, we first created a contingency table with MLHs and MCGs and converted it to a directed and weighted edgelist where each edge was a connection from an MCG to an MLH weighted by the number of samples shared in the connection.
-This was then converted to a bipartite graph where top nodes represented MCGs and bottom nodes represented MLHs.
-The cluster walktrap community detection algorithm as implemented in the `cluster_walktrap()` function in *igraph* version `r packageVersion("igraph")` was used to define clusters of MLHs and MCGs [@csardi2006igraph; @pons2006computing].
-All communities that had fewer than 10 members were then consolidated into one.
+To construct the network, a contingency table was created with MLHs and MCGs and converted to a directed and weighted edgelist where each edge represented a connection from an MCG to an MLH, weighted by the number of samples shared in the connection.
+This was then converted to a bipartite graph where top nodes represented MLHs and bottom nodes represented MCGs.
+To identify clusters of MLHs and MCGs within the network, we used the cluster walktrap community detection algorithm as implemented in the `cluster_walktrap()` function in *igraph* version `r packageVersion("igraph")` [@csardi2006igraph; @pons2006computing].
+This algorithm attempts to define clusters of nodes by starting at a random nodes and performing short, random "walks" along the edges between nodes, assuming that these walks would stay within clusters.
+For this analysis, we set the number of steps within a walk to four and allowed the algorithm to use the edge weights in determining the path.
+All of the resulting communities that had fewer than 10 members were then consolidated into one.
 Community definitions were used to assess the average genetic distance (as defined by Bruvo's distance) within members of the community [@bruvo2004simple].

### 2017-09-06 doc/manuscript/manuscript.Rmd
finish edits to M&M

@@
 
-We used analysis of molecular variance (AMOVA) with Bruvo's genetic distance in *poppr* to test for differentiation associated with wmn by quantifying the differentiation conferred by Region, Source, and Year [@excoffier1992analysis; @bruvo2004simple; @kamvar2014poppr].
-Source in this case was defined as a population belonging to a wmn or not.
-
-To identify specific Regions that were more differentiated than others, we used discriminant analysis of principal components (DAPC) as implemented in *adegenet* version `r packageVersion("adegenet")`, assessing the per-sample posterior group assignment probability [@jombart2008adegenet].
+We used analysis of molecular variance (AMOVA) with Bruvo's genetic distance in *poppr* to test for differentiation between populations in wmn and producer fields from the same region and collected in the same year [@excoffier1992analysis; @bruvo2004simple; @kamvar2014poppr].
+To identify Regions with greater differentiation, we used discriminant analysis of principal components (DAPC) as implemented in *adegenet* version `r packageVersion("adegenet")`, assessing the per-sample posterior group assignment probability [@jombart2008adegenet].
 This method decomposes the genetic data into principal components, and then uses a subset of these as the inputs for discriminant analysis, which attempts to minimize within-group variation and maximize among-group variation [@jombart2010discriminant].
@@
 Each node was weighted by the fraction of shared MLHs.
-Each edge represented a single MLH, but because a single MLH could be present in more than one population, that MLH would be represented by (*n*\*(*n*-1))/2 edges where n represents the number of populations crossed.
+Each edge represented a single MLH, but because a single MLH could be present in more than one population, that MLH would have a number of edges equivalent to the total number of possible connections, calculated as (*n*\*(*n*-1))/2 edges where n represents the number of populations crossed.
 Edges were weighted by $1-P_{sex}$, where $P_{sex}$ is the probability of encountering the same haplotype via two independent meiotic events [@arnaud2007standardizing; @parks1993study].
-We then used the `cluster_optimal()` function from *igraph* to define communities [@csardi2006igraph].
+This weighting scheme would thus strengthen the connection of edges that represented genotypes with a low probability of being produced via sexual reproduction. 
+We then identified communities (among the Regions) in the graph using the `cluster_optimal()` function from *igraph* [@csardi2006igraph].
 The graph was plotted using the R packages *ggplot2* version `r packageVersion("ggplot2")` [@ggplot2] and *ggraph* `r packageVersion("ggraph")` [@ggraph].

### 2017-09-06 doc/manuscript/manuscript.Rmd
Finished with results

@@
 big_haplotype <- in_reach %>% filter(N == max(N))
+
+# Aggressiveness tables --------------------------------------------------
+agghux <- . %>%   huxtable::as_huxtable(add_colnames = TRUE) %>%
+  huxtable::set_number_format(row = huxtable::everywhere, col = 1, 0) %>% 
+  huxtable::set_col_width(c(0.13, 0.25, 0.08)) %>%
+  huxtable::set_align(row = huxtable::everywhere, col = 2, "right") %>%
+  huxtable::print_md(max_width = 40)
+
+Regionagg <- read_csv("../../results/tables/Region-aggressiveness.csv", 
+         na = "NA", # spaces read in as spaces
+         col_types = cols(.default = col_character()))
+
+MCGagg <- read_csv("../../results/tables/MCG-aggressiveness.csv", 
+         na = "NA", # spaces read in as spaces
+         col_types = cols(.default = col_character()))
+MLHagg <- read_csv("../../results/tables/MLG-aggressiveness.csv", 
+         na = "NA", # spaces read in as spaces
+         col_types = cols(.default = col_character())) %>%
+  rename(MLH = MLG)
+
+Raggmean <- mean(as.numeric(Regionagg[[2]]))
+Maggmean <- mean(as.numeric(MCGagg[[2]]))
+Haggmean <- mean(as.numeric(MLHagg[[2]]))
 ```
@@
 A total of 366 isolates were collected from 2003 to 2012 (except 2006 and 2011) from diseased dry bean plants in Australia, Mexico, France, and 11 states in the USA (Table \@ref(tab:isolate-table)).
+We observed a total of 165 MLHs (215 with 16 loci).
 Over 50% of the isolates came from four states, MI (`r pop_size[["MI"]]`), ND (`r pop_size[["ND"]]`), WA (`r pop_size[["WA"]]`), NE (`r pop_size[["NE"]]`). 
@@
 We observed 87 MCGs, the most abundant of which ('MCG 5') was represented by 73 isolates over 37 MLHs (Fig \@ref(biggraph)A,C).
-Analysis of the haplotype accumulation curve showed no clear plateau (Supplemental Information), indicating that we would likely obtain more multilocus haplotypes if we were to haplotype more loci.
+
 The number of observed alleles per locus ranged from two to 10 with an average of `r signif(Amean, 3)` (Table \@ref(tab:locus-stats)).
+Locus `r locus_table$Locus[locus_table$allele==2]`, which contained only 2 alleles, showed low values of both *h* (`r signif(locus_table$Hexp[locus_table$allele==2], 3)`) and evenness (`r signif(locus_table$Evenness[locus_table$allele==2], 3)`), indicating that there was one dominant allele present.
+Analysis of the haplotype accumulation curve showed no clear plateau (Supplemental Information), indicating that we would likely obtain more multilocus haplotypes if we were to genotype more loci.
+
 After clone-correction on the hierarchy of Region/Source/Host/Year, we were left with `r nInd(dat11cc)` isolates representing `r nmll(dat11cc)` MLHs (Table \@ref(tab:iatab)).
+The results show that most populations exhibited relatively high amounts of genotypic diversity, with the exception of MI and NE, which exhibited lower-than average $E_5$ values.
+Mexico had the lowest average value of gene diversity, *h* and the lowest *eMLH*, suggesting low overall genetic diversity. 
+Nearly all populations showed evidence for linkage (Table \@ref(tab:iatab)), which serves as evidence for clonal reproduction or other forms of non-random mating. 
+The only exceptions are CA and Australia. Both of these populations showed only moderate significance (P < 0.05) with $\bar{r}_d$ values of 0.03 and 0.12, respectively.
 
@@
 Pop = Population, N = number of individuals (number of MLH in parentheses), 
-eMLH = expected number of MLHs based on rarefaction at min(10, N) individuals (standard error in parentheses), 
+eMLH = expected number of MLHs based on rarefaction at 10 individuals (standard error in parentheses), 
 H = Shannon-Weiner Index, G = Stoddardt and Taylor's Index, 
@@
 ![
-Data show complex associations between Mycelial Compatibility Groups and Multilocus Haplotypes.
+Associations between Mycelial Compatibility Groups and Multilocus Haplotypes.
 **A)** Barplot of Mycelial Compatibility Group (MCG) abundance in descending order.
@@
 The transparency of the bars represent the evenness of the distribution of the MLHs within a given MCG.
-A dashed box surrounds the 8 most common MCGs representing > 51% of the data.
+A dashed box surrounds the eight most common MCGs representing > 51% of the data.
 **B)** Full graph-representation of the relationship between MCGs (open circles) and MLHs (filled circles).
-Details in fig. \@ref(fullgraph). 
+Details in Fig. \@ref(fullgraph). 
 **C)** A subset of **B** representing the 8 most common MCGs and their associated MLHs (dashed box in **A**).
@@
 The forward-backward selection process of the dbRDA models on clone-corrected data revealed `r RDAVAR1` to be the optimal variables for the reduced model, accounting for `r prettify(vartab(caplist1$components)[2, "Proportion"])`% of the total variation.
-ANOVA revealed that the reduced model was significant with an adjusted R^2^ of `r RDA1R2` (*P* = 0.001).
-Assessment of the marginal effects showed that all of these varaibles significantly explained genetic variation (*P* = `r anova1res[["Year"]]`, *P* = `r anova1res[["Region"]]` for region).
-It was revealed that there was multicollinearity when MCG was combined with any other variable, so we followed the same procedure, dropping MCG from the list of potential predictors. 
-From these results, we found that `r RDAVAR2` were optimal, accounting for `r prettify(vartab(caplist2$components)[2, "Proportion"])`% of the total variation.
+ANOVA showed that the reduced model was significant with an adjusted R^2^ of `r RDA1R2` (*P* = 0.001).
+Assessment of the marginal effects showed that all varaibles significantly explained genetic variation (*P* = `r anova1res[["Year"]]`, *P* = `r anova1res[["Region"]]` for Region).
+We found that there was multicollinearity when MCG was combined with any other variable, so repeated the analysis, dropping MCG from the list of potential predictors. 
+From these results, `r RDAVAR2` were found to be optimal, accounting for `r prettify(vartab(caplist2$components)[2, "Proportion"])`% of the total variation.
 ANOVA revealed significant effects with an adjusted R^2^ of `r RDA2R2` (*P* = 0.001).
-While the marginal effect assessment showed that `r andlist(names(anova2res)[anova2res <= 0.01])` all significantly explained the variation at *P* = `r anova2res[["Year"]]`, and Aggressiveness significantly explained the variation at *P* = `r anova2res[["Severity"]]`.
-Much of the variation appears to be driven by isolates from Mexico and 2005 (Fig. \@ref(rda-biplot)).
-Variance partitioning of the independent variables without MCG indicated aggressiveness to be the least influential factor with 0.1% contributing to explaining the variation of molecular data whereas the combination of variables accounted for 3.3%.
+While the marginal effect assessment revealed that `r andlist(names(anova2res)[anova2res <= 0.01])` significantly explained variation at *P* = `r anova2res[["Year"]]`, and Aggressiveness significantly explained variation at *P* = `r anova2res[["Severity"]]`.
+Much of the variation appeared to be driven by isolates from Mexico and 2005 (Fig. \@ref(rda-biplot)).
+Variance partitioning of the independent variables without MCG indicated aggressiveness to be the least influential factor with 0.1% contributing to explaining the variation of molecular data, whereas the combination of variables accounted for 3.3%.
 
@@
 ![
-Biplot showing five most influential explanatory variables (arrows) overlayed on the first two eigenvectors of distance based redundancy analysis of *S. sclerotiorum* isolates. 
+Biplot showing five most influential explanatory variables (arrows) overlayed on the first two eigenvectors of distance based redundancy analysis of *Sclerotinia sclerotiorum* isolates. 
 The length of the arrows are directly proportional to the strength of the correlation between explanatory and molecular variables. 
-Open circles represent the 318 clone-corrected samples in ordination space. 
+Open circles represent the 318 clone-corrected haplotypes in ordination space. 
 ](../../results/figures/publication/Figure7Z.pdf){#rda-biplot width=61%}
@@
 ![
-Boxplot representing aggressiveness (vertical axis) by population (horizontal axis) arranged in descending order of mean aggressiveness for all populations with N > 5.
-Box boundaries represent 25% and 75% percentiles and whiskers extend to at most 1.5 $\times$ the interquartile range.
+Boxplot of aggressiveness by population arranged in descending order of mean aggressiveness for all populations with N > 5.
+Box boundaries represent 25th and 75th percentiles and whiskers extend to at most 1.5 $\times$ the interquartile range.
 Horizontal line represents median value.
-Circles represent individual observations where filled circles indicate isolates from white mold screening nurseries (wmn).
+Circles represent individual isolates where filled circles are isolates from white mold screening nurseries (wmn) and open circles are isolates from producer fields.
 ](../../results/figures/publication/aggressiveness.pdf){#aggressiveness-boxplot}
@@
 
-Our assessment of aggressiveness in association with Region showed a significant effect (P < 1.00e^-4^), with means that ranged from 5.8 (MN) to 4.0 (CA) (Fig.
+Agressiveness of the isolates ranged from `r paste(range(other(dat11)$meta$Severity), collapse = " to ")` with a mean of `r signif(mean(other(dat11)$meta$Severity), 3)` and median of `r signif(median(other(dat11)$meta$Severity), 3)`.
+The group mean averages were `r sprintf("%0.2f, %0.2f, and %0.2f", Raggmean, Maggmean, Haggmean)` for Region, MCG, and MLH, respectively.
+Our assessment of aggressiveness in association with Region showed a significant effect (*P* < 1.00e^-4^), with means that ranged from 5.8 (MN) to 4.0 (CA) (Fig.
 \@ref(aggressiveness-boxplot), Table \@ref(tab:region-aggressiveness)).
-MCGs also showed a significant effect (P < 1.00e^-4^), with means that ranged from 6.0 ('MCG 44') to 4.6 ('MCG 49') (Table \@ref(tab:MCG-aggressiveness)).
+MCGs also showed a significant effect (*P* < 1.00e^-4^), with means that ranged from 6.0 ('MCG 44') to 4.6 ('MCG 49'; Table \@ref(tab:MCG-aggressiveness)).
 We additionally found a significant effect for MLHs (*P* = 7.44e^-4^) for MLHs, with means that ranged from 6.0 ('MLH 78') to 4.3 ('MLH 140') (Table \@ref(tab:MLG-aggressiveness)).
@@
 
+In our analysis, we found `r nmll(dat11)` MLHs with `r sum(table(mll(dat11)) == 1)` singletons and `r nPop(setPop(dat11, ~MCG))` MCGs with `r sum(table(strata(dat11)$MCG) == 1)` singletons (Fig.
+\@ref(biggraph)A,B) where the eight most abundant MCGs represented > 51% of the data over 11 Regions, and all years except for 2012.
 Our network-based approach to correlating MLHs with MCGs revealed a large and complex network (Fig. \@ref(biggraph)). 
@@
 The three communities with the most members contained eight of the 10 most abundant MCGs. 
-Comparing these communities with Bruvo's genetic distance revealed an average distance of 0.451 among communities and an average distance of 0.437 within communities, showing no significant difference.
+Comparing these communities with Bruvo's genetic distance showed an average distance of 0.451 among communities and an average distance of 0.437 within communities, which were not significantly different.
 When we assessed the number of times two different MLHs that are in the same MCG, considering these as potential heterothallic pairings that could result in sexual recombination, we found an average of `r signif(mean(in_reach$N), 3)` potential heterothallic parings per MLH.
@@
 
-We observed a total of 165 MLHs (215 with 16 loci).
-The most abundant MLH was represented by 27 isolates (Table \@ref(tab:mlg-table)) from five regions (NE, MI, WA, CO, and ND). 
-Within regions, the haplotypes were relatively evenly distributed with moderate to high diversity (Table \@ref(tab:iatab)). 
-Of the 165 MLHs, 76 (46%) were found in at least two geographic regions except those found in WI (2), ID (1), and Mexico (18) (Fig. \@ref(community-graph)).
-Community analysis of the MLHs shared between populations revealed 4 communities with a modularity of 0.17 (Fig.
-\@ref(community-graph)): A coastal community (CA, OR, WA, and NY), a midwest community (CO, NE, ND, MI), and an international community (MN, France, Australia).
-Analysis with 16 loci resulted in the removal of the NY node because it no longer shared a haplotype with OR, but the same overall community structure was present with a modularity of 0.2 (Fig.
+The most abundant MLH was represented by 27 isolates (Table \@ref(tab:mlg-table)) from five Regions (NE, MI, WA, CO, and ND). 
+Within Regions, haplotypes were relatively evenly distributed with moderate to high diversity (Table \@ref(tab:iatab)). 
+Of the 165 MLHs, 76 (46%) were found in at least two Regions, except those found in WI (2), ID (1), and Mexico (18) (Fig. \@ref(community-graph)).
+
+We had performed an analysis on a network where the connections represented shared MLHs across populations, weighted by $1 - P_{sex}$ (Fig. \@ref(community-graph)). 
+Community analysis of the MLHs shared between populations revealed 4 communities with a modularity of 0.17: A coastal community (CA, OR, WA, and NY), a midwest community (CO, NE, ND, MI), and an international community (MN, France, Australia).
+Although analysis with 16 loci resulted in the removal of the NY node because it no longer shared a haplotype with OR, the same overall community structure was present with a modularity of 0.2 (Fig.
 \@ref(community-graph16)).
-The international community appears to be driven by MLH 4, which is shared between all three populations and has a $P_{sex}$ value of 2.87e^-5^, in contrast to the abundant MLH 25, which has a $P_{sex}$ value of 0.0168.
+Relative to the US, the international community appears to be driven by MLH 4, which is shared between all three populations and has a $P_{sex}$ value of 2.87e^-5^, in contrast to the abundant MLH 25, which has a $P_{sex}$ value of 0.0168.
 
@@
 
-Table: (\#tab:mlg-table) The 5 most abundant Multilocus Haplotypes (MLH) with
+Table: (\#tab:mlg-table) The five most abundant Multilocus Haplotypes (MLH) with
 the probability of second encounter ($P_{sex}$), Mycelial Compatibility Groups (MCG), and Regions with sample sizes in parentheses.
@@
 
-The analysis of molecular variance (AMOVA) over the hierarchy of Region, Source, and Year showed significant variation between regions and years, but no significant variation between wmn and non-wmn (Table \@ref(tab:amova)).
+The analysis of molecular variance (AMOVA) for clone-corrected samples over the hierarchy of Region, Source, and Year showed significant variation between regions and years, but no significant variation between wmn and non-wmn (Table \@ref(tab:amova)).
 In contrast, when we compared the three cultivars, Beryl, Bunsi, and G122, we found no significant differentiation (Supplementary Information).
@@
 
-Table: (\#tab:amova) Analysis of Molecular Variance on Bruvo's genetic distance showing little differentiation between white mold screening nurseries (wmn) and other sources.
-The hierarchy was constructed as Source/Region where source is defined as belonging to a wmn or not.
-Bold $\Phi$ values indicate significant values (*P* < 0.05).
+Table: (\#tab:amova) Comparison of populations in the white mold screening nurseries (wmn) and producer fields using an analysis of molecular variance on Bruvo's genetic distance showing little differentiation between white mold screening nurseries (wmn) and other sources.
+The hierarchy was constructed as Source/Region where source is defined as belonging to a wmn or producer field.
+Bold $\Phi$ values indicate significant difference (*P* < 0.05).
 S.S.
@@
 \@ref(DAPC)b).
-The second discriminant axis, representing `r paxis(dapc_region, 2)`% of the discriminatory power, separated the centroid for the California isolates.
+The second discriminant axis, representing `r paxis(dapc_region, 2)`% of the discriminatory power, separated the centroid for the CA isolates.
 The mean population assignment probabilities for all populations with n > 10 showed that only isolates from Mexico, CA, and France had > 50% probabilities of being reassigned to their source populations (Fig.
@@
 DAPC grouping by Region and Year used the first `r dapc_region_year$n.pca` principal components, representing `r prettify(dapc_region_year$var)`% of the total variance.
-The Midwestern populations (NE, MI, CO, ND) did not appear to have any variation across time in contrast to WA, which showed a shift in population structure in the last year of sampling, 2008 (Fig.
+The North Central US populations (NE, MI, CO, ND) did not appear to have any variation across time in contrast to WA, which showed a shift in population structure in the last year of sampling, 2008 (Fig.
 \@ref(DAPC-RY)).
-Further analysis of this population revealed that all 12 isolates in WA circa 2008 originated in a wmn and nine of these haplotypes were shared with CA, and three were shared with France (Fig.
+Further analysis of this population revealed that all 12 isolates in WA circa 2008 originated in a wmn; nine haplotypes were shared with CA, and three were shared with France (Fig.
 \@ref(community-graph), \@ref(community-graph16)).
@@
 ![
-Scatter plot of Discriminant Analysis of Principal Components on Regions and Years showing temporal variation in WA and CA populations as compared to Midwestern Regions, NE and MI. 
+Scatter plot of Discriminant Analysis of Principal Components on Regions and Years showing non-differentiated temporal variation NE and MI and temporal variation in WA and CA. 
 Points (text labels) represent observed individuals connected to the population centroids with ellipses representing a 66% confidence interval for a normal distribution. 
 The center of each component is represented as black grid lines.
-A more detailed view can be seen in Fig. \@ref(DAPC-RY-FULL).
+A more detailed view is shown in Fig. \@ref(DAPC-RY-FULL).
 ](../../results/figures/publication/dapc_region_year_micanewa.pdf){#DAPC-RY width=61%}
@@
 <!-- end figure -->
-
-### Linkage Analysis {-}
-
-Nearly all populations showed evidence for linkage (Table \@ref(tab:iatab)), which serves as evidence for clonal reproduction or other forms of non-random mating. 
-The only exceptions are CA and Australia. Both of these populations showed only moderate significance (P < 0.05) with $\bar{r}_d$ values of 0.03 and 0.12, respectively.
 
@@
 ```{r table2, echo = FALSE, warning = FALSE, message = FALSE, results = "asis"}
-suppressPackageStartupMessages(library("tidyverse"))
-read_csv("../../results/tables/Region-aggressiveness.csv", 
-         na = "NA", # spaces read in as spaces
-         col_types = cols(.default = col_character())) %>%
-  huxtable::as_huxtable(add_colnames = TRUE) %>%
-  huxtable::set_number_format(row = huxtable::everywhere, col = 1, 0) %>% 
-  huxtable::set_col_width(c(0.13, 0.25, 0.08)) %>%
-  huxtable::set_align(row = huxtable::everywhere, col = 2, "right") %>%
-  huxtable::print_md(max_width = 40)
+Regionagg %>% agghux
 ```
@@
 ```{r MCG-ag, echo = FALSE, warning = FALSE, message = FALSE, results = "asis"}
-read_csv("../../results/tables/MCG-aggressiveness.csv", 
-         na = "NA", # spaces read in as spaces
-         col_types = cols(.default = col_character())) %>%
-  huxtable::as_huxtable(add_colnames = TRUE) %>%
-  huxtable::set_number_format(row = huxtable::everywhere, col = 1, 0) %>% 
-  huxtable::set_col_width(c(0.13, 0.25, 0.08)) %>%
-  huxtable::set_align(row = huxtable::everywhere, col = 2, "right") %>%
-  huxtable::print_md(max_width = 40)
+MCGagg %>% agghux
 ```
@@
 ```{r MLG-ag, echo = FALSE, warning = FALSE, message = FALSE, results = "asis"}
-read_csv("../../results/tables/MLG-aggressiveness.csv", 
-         na = "NA", # spaces read in as spaces
-         col_types = cols(.default = col_character())) %>%
-  rename(MLH = MLG) %>%
-  huxtable::as_huxtable(add_colnames = TRUE) %>%
-  huxtable::set_number_format(row = huxtable::everywhere, col = 1, 0) %>% 
-  huxtable::set_col_width(c(0.13, 0.25, 0.08)) %>%
-  huxtable::set_align(row = huxtable::everywhere, col = 2, "right") %>%
-  huxtable::print_md(max_width = 40)
+MLHagg %>% agghux
 ```

### 2017-09-06 doc/manuscript/manuscript.Rmd
5 -> five

@@
 
-Table: (\#tab:region-aggressiveness) Mean aggressiveness ratings for the Regions with greater than 5 samples; groupings according to 95% family-wise confidence interval.
+Table: (\#tab:region-aggressiveness) Mean aggressiveness ratings for Regions with more than five samples; groupings according to 95% family-wise confidence interval.
 

### 2017-09-06 doc/RMD/MCG-virulence.Rmd
update MCG virulence; add ggforce
APPARENTLY agricolae has change such that HSD.test
has completely different output, messing up my flow
>:(

Oh, I also am using geom_sina and crossbars for the
figure.

@@
 grps %>% 
-  dplyr::rename(MCG = trt) %>%
-  dplyr::rename(Group = M) %>%
-  dplyr::rename(`Mean Aggressiveness` = means) %>%
+  tibble::rownames_to_column("MCG") %>%
+  dplyr::rename(Group = groups) %>%
+  dplyr::rename(`Mean Aggressiveness` = Severity) %>%
   readr::write_csv(path = "results/tables/MCG-aggressiveness.csv", col_names = TRUE) %>%
@@
   mutate(mean_sev = mean(Severity)) %>%
-  mutate(Source = ifelse(Source == "wmn", "wmn", "other")) %>%
+  mutate(Source = ifelse(Source == "wmn", "wmn", "producer field")) %>%
   arrange(desc(mean_sev)) %>%
@@
   select(Region, Severity, Source)
+
+plot.median <- function(x) {
+  m <- mean(x, na.rm = TRUE)
+  c(y = m, ymin = m, ymax = m)
+}
+
 ggplot(region_dat, aes(x = Region, y = Severity)) +
-  geom_point(aes(fill = Source), 
+  ggforce::geom_sina(aes(fill = Source), 
              pch = 21, 
-             position = position_jitter(width = 0.25), 
-             alpha = 0.35) +
-  geom_boxplot(alpha = 0.5) +
+             alpha = 0.4) +
+  # geom_point(aes(fill = Source), 
+  #            pch = 21, 
+  #            position = position_jitter(width = 0.25), 
+  #            alpha = 0.35) +
+  # geom_boxplot(alpha = 0.5) +
+  stat_summary(fun.data = "plot.median", geom = "errorbar", colour = "black", width = 0.6, size = 2, alpha = 0.5) +
+  stat_summary(fun.data = "plot.median", geom = "errorbar", colour = "white", width = 0.5, size = 1, alpha = 1) +
   scale_y_continuous(limits = c(1, 9), breaks = c(1, 3, 5, 7, 9), expand = c(0, 0.1)) +
@@
 grps <- agricolae::HSD.test(ANOVA, "Region", alpha = 0.05)$groups %>% 
-  dplyr::rename(Region = trt) %>%
-  dplyr::rename(Group = M) %>%
-  dplyr::rename(`Mean Aggressiveness` = means) %>%
+  tibble::rownames_to_column("Region") %>%
+  dplyr::rename(Group = groups) %>%
+  dplyr::rename(`Mean Aggressiveness` = Severity) %>%
   readr::write_csv(path = "results/tables/Region-aggressiveness.csv", col_names = TRUE) %>%
@@
 grps <- agricolae::HSD.test(ANOVA, "MLG", alpha = 0.05)$groups %>% 
-  dplyr::rename(MLG = trt) %>%
-  dplyr::rename(Group = M) %>%
-  dplyr::rename(`Mean Aggressiveness` = means) %>%
+  tibble::rownames_to_column("MLG") %>%
+  dplyr::rename(Group = groups) %>%
+  dplyr::rename(`Mean Aggressiveness` = Severity) %>%
   readr::write_csv(path = "results/tables/MLG-aggressiveness.csv", col_names = TRUE) %>%

### 2017-09-06 doc/RMD/mlg-mcg.Rmd
update figure 1

@@
   geom_col(aes(width = ifelse(Type == "MLHs", 0.5, 0.85)), color = "black", position = "identity") +
-  annotate(geom = "rect", xmin = 0.5, xmax = 8.5, ymin = 0, ymax = 75, color = "black", size = 1.5, alpha = 0, lty = 2) +
+  annotate(geom = "rect", xmin = 0.5, xmax = 8.5, ymin = 0, ymax = 75, color = "black", size = 1.125, alpha = 0, lty = 2) +
   # annotate(geom = "text", x = 18, y =35, 
   #          label = sprintf("Mean Evenness: %.3f", mean(mcgs$Evenness, na.rm = TRUE))) +
-  scale_fill_manual(values = c("black", "white")) +
+  scale_fill_manual(values = c("black", "white"), labels = c("MLHs", "Isolates")) +
   scale_y_continuous(expand = c(0, 2)) +
@@
                  arrow = arrow(length = unit(0.01, "native"))) +
-  geom_node_text(aes(label = gsub("MLG.", "", name), 
+  geom_node_text(aes(label = gsub("MLG.", "", name), # ifelse(size > 0.1, gsub("MLG.", "", name), ""), 
                      color = type, 
@@
   theme(plot.margin = unit(c(0, 0, 0, 0), "lines"))
-t5g$labels$edge_width <- "shared isolates"
+t5g$labels$edge_width <- "Shared Isolates"
 t5g$labels$fill <- ""

### 2017-09-06 doc/manuscript/manuscript.Rmd
fix locus_table -> loc_table

@@
 The number of observed alleles per locus ranged from two to 10 with an average of `r signif(Amean, 3)` (Table \@ref(tab:locus-stats)).
-Locus `r locus_table$Locus[locus_table$allele==2]`, which contained only 2 alleles, showed low values of both *h* (`r signif(locus_table$Hexp[locus_table$allele==2], 3)`) and evenness (`r signif(locus_table$Evenness[locus_table$allele==2], 3)`), indicating that there was one dominant allele present.
+Locus `r loc_table$Locus[loc_table$allele==2]`, which contained only 2 alleles, showed low values of both *h* (`r signif(loc_table$Hexp[loc_table$allele==2], 3)`) and evenness (`r signif(loc_table$Evenness[loc_table$allele==2], 3)`), indicating that there was one dominant allele present.
 Analysis of the haplotype accumulation curve showed no clear plateau (Supplemental Information), indicating that we would likely obtain more multilocus haplotypes if we were to genotype more loci.

### 2017-09-06 doc/RMD/MCG-virulence.Rmd
update graph to use a better binwidth

@@
 
-plot.median <- function(x) {
+plot.mean <- function(x) {
   m <- mean(x, na.rm = TRUE)
@@
 
+set.seed(2017-09-06)
 ggplot(region_dat, aes(x = Region, y = Severity)) +
@@
                      pch = 21, 
+                     binwidth = 0.1, 
                      alpha = 0.4) +
@@
   # geom_boxplot(alpha = 0.5) +
-  stat_summary(fun.data = "plot.median", geom = "errorbar", colour = "black", width = 0.6, size = 2, alpha = 0.5) +
-  stat_summary(fun.data = "plot.median", geom = "errorbar", colour = "white", width = 0.5, size = 1, alpha = 1) +
+  stat_summary(fun.data = plot.mean, geom = "errorbar", colour = "black", width = 0.6, size = 2, alpha = 0.5) +
+  stat_summary(fun.data = plot.mean, geom = "errorbar", colour = "white", width = 0.5, size = 1, alpha = 1) +
   scale_y_continuous(limits = c(1, 9), breaks = c(1, 3, 5, 7, 9), expand = c(0, 0.1)) +

### 2017-09-06 doc/RMD/MLG-distribution.Rmd
remove "arms" from labels in MLG-distributions

@@
     geom_node_circle(aes(r = drop(scale(size, center = FALSE))/10, fill = size))
-  outg <- if (label) outg + geom_node_label(aes(label = label), repel = TRUE, parse = TRUE, label.size = 0.75, nudge_x = x_nudge) else outg
+  outg <- if (label) outg + geom_node_label(aes(label = label), repel = TRUE, parse = TRUE, label.size = 0.75, nudge_x = x_nudge, segment.size = 0) else outg
   outg + 

### 2017-09-06 doc/RMD/MLG-distribution.Rmd
update figures

--- 2017-08-23 doc/RMD/MLG-distribution.Rmd
@@
     geom_node_circle(aes(r = drop(scale(size, center = FALSE))/10, fill = size))
-  outg <- if (label) outg + geom_node_label(aes(label = label), repel = TRUE, parse = TRUE, label.size = 0.75, nudge_x = x_nudge) else outg
+  outg <- if (label) outg + geom_node_label(aes(label = label), repel = TRUE, parse = TRUE, label.size = 0.75, nudge_x = x_nudge, segment.size = 0, segment.alpha = 0) else outg
   outg + 

### 2017-09-06 doc/RMD/by-year.Rmd
update figures

@@
   colnames(res) <- c("d.f.", "Sum of Squares", "Percent variation", "P", 
-                     "Phi statistic")
+                     "Sigma", "Phi statistic")
   names(dimnames(res)) <- c("levels", "statistic")
@@
 
-yearscale <- viridis::viridis(nlevels(strata(dat11cc)$Year), option = "C")
+yearscale <- viridis::viridis(nlevels(strata(dat11cc)$Year), option = "C", end = 0.9)
 names(yearscale) <- sort(levels(strata(dat11cc)$Year))

### 2017-09-06 doc/RMD/pop-diff.Rmd
update figures

@@
   # scale_x_discrete(position = "top") +
-  scale_y_continuous(limits = c(0, 1), minor_breaks = c(seq(0, 1, by = 0.125)), breaks = c(0, 0.5, 1)) +
+  scale_y_continuous(limits = c(0, 1), minor_breaks = c(seq(0, 1, by = 0.125)), breaks = c(0, 0.5, 1), position = "right") +
   theme_bw(base_size = 16, base_family = "Helvetica") +
@@
 LEG <- cowplot::get_legend(LDS_PLOT + theme(legend.position = "right") + theme(legend.text = element_text(size = 12)))
-cowplot::plot_grid(REGION_PLOT +
-                     theme(plot.margin = unit(c(0, 0, 0, 0), "lines")), 
+cowplot::plot_grid(LEG,
                    LDS_PLOT + 
@@
                      theme(plot.margin = unit(c(0, 0, 0, 0), "lines")) +
-                     theme(aspect.ratio = 0.85), 
-                   LEG,
+                     theme(panel.border = element_rect(color = "grey50")) +
+                     theme(aspect.ratio = 0.935), 
+                   REGION_PLOT +
+                     theme(plot.margin = unit(c(0, 1, 0, 0), "lines")), 
                    nrow = 1, 
-                   rel_widths = c(1, 0.85, 0.25),
+                   rel_widths = c(0.25, 0.75, 1),
                    label_size = 25,
                    label_fontfamily = "Helvetica",
-                   labels = c("A", "B", NA),
+                   labels = c(NA, "A", "B"),
                    align = "v",

### 2017-09-06 doc/manuscript/manuscript.Rmd
update figures

@@
 Discriminant Analysis of Principal Components on regions showing that Mexico is differentiated from other populations. 
-**A)** Mean population assignment (horizontal axis) probability (vertical axis) for all populations with N > 10 (facets).
+**A)** Scatter plot of first two components from DAPC.
+Points represent observed individuals connected to the population centroids with ellipses representing a 66% confidence interval for a normal distribution. 
+The center of each component is represented as black grid lines.
+**B)** Mean population assignment probability from the DAPC for all populations with N > 10 (facets).
+Populations represented along the horizontal axis and probability of assignment on the vertical.
 Numbers next to source populations indicate population size.
 All values sum to one.
-**B)** Scatter plot of first two components from DAPC.
-Points represent observed individuals connected to the population centroids with ellipses representing a 66% confidence interval for a normal distribution. 
-The center of each component is represented as black grid lines.
 ](../../results/figures/publication/DAPC.pdf){#DAPC width=100%}

### 2017-09-07 doc/manuscript/manuscript.Rmd
Add SEE notes and reformat discussion

@@
 **C)** A subset of **B** representing the 8 most common MCGs and their associated MLHs (dashed box in **A**).
-Filled nodes represent MLHs and open nodes represent MCGs.
-Numbers inside nodes are the MLH/MCG label. 
-Node (circle) area scaled to the number of samples represented (range: 1--73).
+Filled nodes (circles) represent MLHs and open nodes represent MCGs.
+Node area scaled to the number of samples represented (range: 1--73).
+Numbers inside nodes are the MLH/MCG label (if n > 1). 
 Edges (arrows) point from MLH to MCG where the weight (thickness) of the edge represents the number of shared isolates (range: 1--19).
@@
 
-### Variable Importance {-}
+### Variable Contributions {-}
+
+<!-- RDA figure -->
+
+![
+Biplot showing five most influential explanatory variables (arrows) overlayed on the first two eigenvectors of distance based redundancy analysis of *Sclerotinia sclerotiorum* isolates. 
+The length of the arrows are directly proportional to the strength of the correlation between explanatory and molecular variables. 
+Open circles represent the 318 clone-corrected haplotypes in ordination space. 
+](../../results/figures/publication/Figure7Z.pdf){#rda-biplot width=61%}
+
+<!-- end figure -->
+
+<!-- aggressiveness figure -->
+
+![
+Strip Plot of aggressiveness by population arranged in descending order of mean aggressiveness for all populations with N > 5.
+White bars represent mean value.
+Circles represent individual isolates where filled circles are isolates from white mold screening nurseries (wmn) and open circles are isolates from producer fields.
+](../../results/figures/publication/aggressiveness.pdf){#aggressiveness-boxplot}
+
+<!-- end figure -->
 
@@
 
-<!-- RDA figure -->
-
-![
-Biplot showing five most influential explanatory variables (arrows) overlayed on the first two eigenvectors of distance based redundancy analysis of *Sclerotinia sclerotiorum* isolates. 
-The length of the arrows are directly proportional to the strength of the correlation between explanatory and molecular variables. 
-Open circles represent the 318 clone-corrected haplotypes in ordination space. 
-](../../results/figures/publication/Figure7Z.pdf){#rda-biplot width=61%}
-
-<!-- end figure -->
-
 ### Aggressiveness {-}
-
-<!-- aggressiveness figure -->
-
-![
-Boxplot of aggressiveness by population arranged in descending order of mean aggressiveness for all populations with N > 5.
-Box boundaries represent 25th and 75th percentiles and whiskers extend to at most 1.5 $\times$ the interquartile range.
-Horizontal line represents median value.
-Circles represent individual isolates where filled circles are isolates from white mold screening nurseries (wmn) and open circles are isolates from producer fields.
-](../../results/figures/publication/aggressiveness.pdf){#aggressiveness-boxplot}
-
-<!-- end figure -->
 
@@
 \@ref(biggraph)A,B) where the eight most abundant MCGs represented > 51% of the data over 11 Regions, and all years except for 2012.
-Our network-based approach to correlating MLHs with MCGs revealed a large and complex network (Fig. \@ref(biggraph)). 
+Our network-based approach to correlating MLHs with MCGs revealed a large and complex network (Fig. \@ref(biggraph), Table \@ref(tab:mlg-table)). 
 Community analysis showed 51 communities, `r sum(in_reach$N == 0)` of which consisted of a single MLH unconnected with any other community indicating that just `r prettify(mean(in_reach$N == 0))`% of the 165 MLHs are unable to cross with any other MLH in this data set (Fig.
@@
 
+```{r table3, echo = FALSE, warning = FALSE, message = FALSE, results = "asis"}
+read_csv("../../results/tables/mlg_table.csv", 
+         na = "NA", # spaces read in as spaces
+         col_types = cols(.default = col_character())) %>%
+  rename(`$P_{sex}$` = Psex) %>% 
+  rename(MLH = MLG) %>%
+  knitr::kable(format = "pandoc")
+```
+
+Table: (\#tab:mlg-table) The five most abundant Multilocus Haplotypes (MLH) with
+the probability of second encounter ($P_{sex}$), Mycelial Compatibility Groups (MCG), and Regions with sample sizes in parentheses.
+
 ## Structure of Shared Multilocus Haplotypes {-}
@@
 
-We had performed an analysis on a network where the connections represented shared MLHs across populations, weighted by $1 - P_{sex}$ (Fig. \@ref(community-graph)). 
+We had performed an analysis on a network where the connections represented shared MLHs across populations, weighted by $1 - P_{sex}$ (Fig. \@ref(community-graph), Table \@ref(tab:mlg-table)). 
 Community analysis of the MLHs shared between populations revealed 4 communities with a modularity of 0.17: A coastal community (CA, OR, WA, and NY), a midwest community (CO, NE, ND, MI), and an international community (MN, France, Australia).
@@
 Each edge represents a single MLH where its thickness represents the number of populations that share the MLH and the shade represents the value of $P_{sex}$, or the probability of encountering that MLH from two independent meiotic events.
-](../../results/figures/publication/mlg-11.pdf){#community-graph width=75%}
+](../../results/figures/publication/mlg-11.pdf){#community-graph width=61%}
 
 <!-- end fig -->
-
-```{r table3, echo = FALSE, warning = FALSE, message = FALSE, results = "asis"}
-read_csv("../../results/tables/mlg_table.csv", 
-         na = "NA", # spaces read in as spaces
-         col_types = cols(.default = col_character())) %>%
-  rename(`$P_{sex}$` = Psex) %>% 
-  rename(MLH = MLG) %>%
-  knitr::kable(format = "pandoc")
-```
-
-Table: (\#tab:mlg-table) The five most abundant Multilocus Haplotypes (MLH) with
-the probability of second encounter ($P_{sex}$), Mycelial Compatibility Groups (MCG), and Regions with sample sizes in parentheses.
 
@@
 
+\newpage
+
 ### Discriminant Analysis of Principal Components {-}
@@
 \@ref(DAPC)a).
+
+DAPC grouping by cultivar used the first 20 principal components, representing 89% of the total variance.
+The first two discriminant axes (representing `r sum(paxis(dapc_host, 1:2))`% of the discriminatory power) failed to separate any of the cultivars where the mean posterior assignment probabilities were `r passp(dapc_host, "G122")`% (G122), `r passp(dapc_host, "Beryl")`% (Beryl), and `r passp(dapc_host, "Bunsi")`% (Bunsi).
+DAPC grouping by Region and Year used the first `r dapc_region_year$n.pca` principal components, representing `r prettify(dapc_region_year$var)`% of the total variance.
+The North Central US populations (NE, MI, CO, ND) did not appear to have any variation across time in contrast to WA, which showed a shift in population structure in the last year of sampling, 2008 (Fig.
+\@ref(DAPC-RY)).
+Further analysis of this population revealed that all 12 isolates in WA circa 2008 originated in a wmn; nine haplotypes were shared with CA, and three were shared with France (Fig.
+\@ref(community-graph), \@ref(community-graph16)).
 
@@
 
-DAPC grouping by cultivar used the first 20 principal components, representing 89% of the total variance.
-The first two discriminant axes (representing `r sum(paxis(dapc_host, 1:2))`% of the discriminatory power) failed to separate any of the cultivars where the mean posterior assignment probabilities were `r passp(dapc_host, "G122")`% (G122), `r passp(dapc_host, "Beryl")`% (Beryl), and `r passp(dapc_host, "Bunsi")`% (Bunsi).
-DAPC grouping by Region and Year used the first `r dapc_region_year$n.pca` principal components, representing `r prettify(dapc_region_year$var)`% of the total variance.
-The North Central US populations (NE, MI, CO, ND) did not appear to have any variation across time in contrast to WA, which showed a shift in population structure in the last year of sampling, 2008 (Fig.
-\@ref(DAPC-RY)).
-Further analysis of this population revealed that all 12 isolates in WA circa 2008 originated in a wmn; nine haplotypes were shared with CA, and three were shared with France (Fig.
-\@ref(community-graph), \@ref(community-graph16)).
-
 <!-- DAPC figure 2: Electric Boogaloo-->
@@
 
+<!-- 
+SEE:
+
+The foundation of the discussion is well built.
+Missing, however, is a synthesis of the information within each paragraph,
+connecting it from one section to the next. I suggest you move the paragraph on
+MCGs to the last (prior to limitations) and then delete all the sub-headings
+(keep limitations and conclusions sub-headers).
+
+I would also like to see more direct discussion of how the pouplation genetic
+can be interpreted epidemiologically. For some guidance here, I suggest
+revisiting Milgroom and Peever 2003 (10.1094/PDIS.2003.87.6.608).
+
+I also suggest supporting epidemiological interpretations with conclusions that
+other *S. sclerotiorum* population have shown. You've done an excellent job 
+using the broader *S. sclerotiorum* literature to support your claims without
+re-hashing the results. I think picking a couple of examples to highlight to
+compare and contrast with the current results would be helpful to readers.
+-->
+
 In this study, we characterized the diversity of *Sclerotinia sclerotiorum* from dry bean fields across the United States.
@@
 
-## Variable Importance {-}
+<!-- ## Variable Contributions {-} -->
 
@@
 The results also show that the Vista cultivar explains some of the variance, but this represents six isolates in MI. 
-Aggressiveness and specific source field had little to no effect on genetic diversity. 
+Aggressiveness and specific source field had little to no effect on prediction of genetic diversity. 
 These results are in agreement with studies that examined differentiation based on Host [@aldrich-wolfe2015genetic] and Aggressiveness [@atallah2004high; @attanayake2012genetic; @attanayake2013sclerotinia] reporting little or no correlation of genetic diversity to these variables.
@@
 
-## Mycelial Compatibility Groups and Multilocus Haplotypes {-}
-
-Of interest in this study was testing the association between multilocus haplotypes (MLHs) and mycelial compatibility groups (MCGs), as the latter have historically been used as a proxy for clonal lineages [@leslie1993fungal; @otto-hanson2011variation; @kohn1990mycelial; @schafer2006optimized; @kohn1995clonal; @carbone1999patterns].
-Our results do not support this assumption.
-It can be seen in Fig.
-\@ref(biggraph)A that the most abundant MCG contains several MLHs, but the diversity of those MLHs are low as indicated by the evenness (transparency), which indicates that there is one dominant MLH ('MLH 25').
-What is not shown in Fig. \@ref(biggraph)A is the MLHs that are shared between MCGs. 
-This is illustrated in both table \@ref(tab:mlg-table) and Fig. \@ref(biggraph)B,C. 
-It could be argued, however that 'MLH 25', with its high value of $P_{sex}$ represents different true MLHs across the five MCGs it occupies, but this doesn't account for the overall structure of Fig.
-\@ref(fullgraph) where, for example, 'MLH 75' ($P_{sex}$ = 1.81e^-4^) is compatible with `r big_haplotype$N` other haplotypes through three MCG when the population structure of *S.
-sclerotiorum* is known to be clonal.
-
-Over the past few years, researchers have noticed inconsistencies among the relationship between MCGs and MLHs [@aldrich-wolfe2015genetic; @carbone1999patterns; @attanayake2012genetic; @lehner2015genetic].
-Either several MCGs belong to one MLH, which could be explained by insufficient sampling of loci; several MLHs belong to one MCG, which could be explained by clonal expansion; or a mixture of both.
-Some studies have shown a correlation between MCG and MLH [@aldrich-wolfe2015genetic; @carbone1999patterns; @lehner2015genetic], whereas other studies have shown no apparent correlation, even on small spatial scales [@atallah2004high; @attanayake2012genetic; @attanayake2013sclerotinia].
-
-One long-held assumption was that MCGs (as determined via barrage reaction) represent vegetative compatibility groups (VCGs) [@schafer2006optimized; @kohn1990mycelial; @lehner2015genetic], which are known to have a genetic component [@saupe2000molecular; @hall2010evolution; @strom2016genomes].
-While our protocol for assessing MCGs utilized a different substrate for the MCG reactions, the patterns we observe are not dissimilar from what have previously been reported in the literature.
-It has been demonstrated in several Ascomycetes---including *Neurospora crassa* [@micali2003independence], *Sclerotinia homoecarpa* [@jo2008reassessment], *Verticillium dahliae* [@papaioannou2014barrage], and *S.
-sclerotiorum* [@ford1995heterokaryon]---that barrage reactions are independent from stable anastomosis.
-Thus, the inconsistencies in this study and other studies indicate that researchers studying *S.
-sclerotiorum* should not rely on MCG data derived from barrage reactions as an indicator for genetic diversity.
-
-## Genotypic Structure {-}
+<!-- ## Genotypic Structure {-} -->
 
@@
 
-## Population Differentiation {-}
+<!-- ## Population Differentiation {-} -->
 
@@
 According to our AMOVA results, we have evidence for differentiation at the Region and Year, but little to no differentiation between wmn isolates and production field isolates (Table \@ref(tab:amova)).
+This lack of differentiation, however, may reflect the breeder practice of inoculating screening plots with sclerotia collected from sources within the region.
 When we analyze the AMOVA results in light of the DAPC results (Fig. \@ref(DAPC)), it becomes clear that the regional patterns of differentiation are largely driven by isolates from Mexico and CA.
@@
 \@ref(community-graph), \@ref(DAPC-RY), \@ref(DAPC-RY-FULL)).
-All of the isolates from WA in 2003--2005, and 2008 came from the same wmn; those in 2003--2005 came a NE field location cropped with dry bean since 2002, and those in 2008 from a SE field that was previously cropped with brassica, sundgrass, peas, beans, and potatoes (Miklas, Phil Pers.
+All of the isolates from WA in 2003--2005, and 2008 came from the same wmn; within the wmn, those in 2003--2005 came a Northeastern field location cropped with dry bean since 2002, and those in 2008 from a Southeastern field that was previously cropped with brassica, sundgrass, peas, beans, and potatoes (Miklas, Phil Pers.
 comm.).
-Both of these fields were inoculated with sclerotia in 2002, the NE field with sclerotia provided by a commercial bean producer and the SE field with sclerotia from peas (although this was thought to be unsuccessful).
+Both of these fields were inoculated with sclerotia in 2002, the Northeastern field with sclerotia provided by a commercial bean producer and the Southeastern field with sclerotia from peas (although this was thought to be unsuccessful).
 Despite this information, it is still unclear what has contributed to the differentiation of the 2008 population from WA.
@@
 It has been proposed within the *S.
-sclerotiorum* literature that isolates from sub-tropical and tropical climates are differentiated or more variable than populations from temperate climates [@lehner2017sclerotinia;@carbone2001multilocus; @attanayake2013sclerotinia].
+sclerotiorum* literature that isolates from sub-tropical and tropical climates are differentiated or more variable than populations from temperate climates [@carbone2001multilocus; @attanayake2013sclerotinia; @lehner2017sclerotinia;].
 This has been attributed to the notion that the fungus has the chance to undergo more reproductive cycles in the warmer climate [@carbone2001multilocus; @attanayake2013sclerotinia]. 
@@
 We were unable to find evidence for sexual reproduction in any region except for CA and Australia. 
-While the Australia population had a non-significant value of $\bar{r}_d$---which would suggest that we cannot reject the null hypothesis of random mating---the sample size was insufficient from which to draw conclusions [@agapow2001indices; @milgroom1996recombination].
+While the Australia population had a non-significant value of $\bar{r}_d$---which would suggest that we cannot reject the null hypothesis of random mating---the sample size was insufficient from which to draw conclusions [@milgroom1996recombination; @agapow2001indices].
 The low value of $\bar{r}_d$ in the CA population may represent sexual reproduction, but this could also be an artifact of sampling two different populations, which is known to reduce the value of $\bar{r}_d$ [@prugnolle2010apparent].
 
-## Aggressiveness {-}
+<!-- ## Aggressiveness {-} -->
 
@@
 This increase was due in part to new data from producer field isolates collected after the previous study.
-Two caveats should come with interpreting these data, however.
-First, these straw tests were performed by a different person for these later isolates, which could suggest a more lenient or strict scoring system. 
-Moreover to this point, many of the ND isolates fell within the 6--7 range, which denotes a physical boundary (disease symptoms around the second node) between intermediate and susceptible [@otto-hanson2011variation].
-Second, these results are all within the intermediate range (4--6) of aggressiveness, which may not represent a significant biological difference.
+These straw tests were performed by a different person for these later isolates, which could suggest a more lenient or strict scoring system.
+However, when we examined the within-region differences, we found no significant effect by individual. 
+Many of the ND isolates fell within the 6--7 range, which denotes a physical boundary (disease symptoms around the second node) between intermediate and susceptible [@otto-hanson2011variation]. 
+Thus, we observed a shift in aggressiveness without a significant shift in genotypic structure, which may indicate that aggressiveness may be controlled by environmental factors as opposed to genetics.
+
+<!-- ## Mycelial Compatibility Groups and Multilocus Haplotypes {-} -->
+
+Mycelial Compatibility Groups have historically been used as a proxy for clonal lineages, and thus, of interest in this study was testing the association between multilocus haplotypes (MLHs) and mycelial compatibility groups (MCGs) [@leslie1993fungal; @otto-hanson2011variation; @kohn1990mycelial; @schafer2006optimized; @kohn1995clonal; @carbone1999patterns].
+Our results, however, do not support this assumption.
+It can be seen in Fig.
+\@ref(biggraph)A that the most abundant MCG contains several MLHs, but the diversity of those MLHs are low as indicated by the evenness (transparency), which indicates that there is one dominant MLH ('MLH 25').
+What is not shown in Fig. \@ref(biggraph)A is the MLHs that are shared between MCGs. 
+This is illustrated in both Table \@ref(tab:mlg-table) and Fig. \@ref(biggraph)B,C. 
+It could be argued, however that 'MLH 25', with its high value of $P_{sex}$ represents different true MLHs across the five MCGs it occupies, but this doesn't account for the overall structure of Fig.
+\@ref(fullgraph) where, for example, 'MLH 75' ($P_{sex}$ = 1.81e^-4^) is compatible with `r big_haplotype$N` other haplotypes through three MCG when the population structure of *S.
+sclerotiorum* is known to be clonal.
+
+Over the past few years, researchers have noticed inconsistencies among the relationship between MCGs and MLHs [@aldrich-wolfe2015genetic; @carbone1999patterns; @attanayake2012genetic; @lehner2015genetic].
+Either several MCGs belong to one MLH, which could be explained by insufficient sampling of loci; several MLHs belong to one MCG, which could be explained by clonal expansion; or a mixture of both.
+Some studies have shown a correlation between MCG and MLH [@aldrich-wolfe2015genetic; @carbone1999patterns; @lehner2015genetic], whereas other studies have shown no apparent correlation, even on small spatial scales [@atallah2004high; @attanayake2012genetic; @attanayake2013sclerotinia].
+
+<!-- 
+SEE:
+
+Perhaps add some discussion of independently evolving vegetatively compatible
+lineages. 
+-->
+
+One long-held assumption was that MCGs (as determined via barrage reaction) represent vegetative compatibility groups (VCGs) [@schafer2006optimized; @kohn1990mycelial; @lehner2015genetic], which are known to have a genetic component [@saupe2000molecular; @hall2010evolution; @strom2016genomes].
+While our protocol for assessing MCGs utilized a different substrate for the MCG reactions, the patterns we observe are not dissimilar from what have previously been reported in the literature.
+It has been demonstrated in several Ascomycetes---including *Neurospora crassa* [@micali2003independence], *Sclerotinia homoecarpa* [@jo2008reassessment], *Verticillium dahliae* [@papaioannou2014barrage], and *S.
+sclerotiorum* [@ford1995heterokaryon]---that barrage reactions are independent from stable anastomosis.
+Thus, the inconsistencies in this study and other studies indicate that researchers studying *S.
+sclerotiorum* should not rely on MCG data derived from barrage reactions as an indicator for genetic diversity.
 
@@
 sclerotiorum* from the USA to date, giving us a unique insight to continent-wide population structure and relationships between phenotypic and genotypic variables.
-Based on our analysis of the relationships between MCG and MLH, we found no clear evidence that the two are directly related, suggesting that MCG do not necessarily represent vegetative compatibility groups and thus should not be used as a proxy for identifying clones.
 Populations in wmn appear to show no significant differentiation when compared to their production field counterparts, suggesting that the wmn populations of *S.
@@
 While we found no direct relationship between haplotype and severity, it is evident that there is a gradient of severity by region, further supporting the need for screening in multiple locations.
+Based on our analysis of the relationships between MCG and MLH, we found no clear evidence that the two are directly related, suggesting that MCG do not necessarily represent vegetative compatibility groups and thus should not be used as a proxy for identifying clones.
 
@@
 
-All scripts, data, and resources used to generate the results presented in this publication (including supplementary information) are fully reproducible and available at The Open Science Framework https://osf.io/ejb5y.
+All scripts, data, and resources used to generate the results presented in this publication (including Supplementary Information) are fully reproducible and available at The Open Science Framework https://osf.io/ejb5y.
 
@@
 
-Table: (\#tab:MCG-aggressiveness) Mean aggressiveness ratings for the first 10 MCG; groupings according to 95% family-wise confidence interval.
+Table: (\#tab:MCG-aggressiveness) Mean aggressiveness ratings for the 10 most abundant MCG; groupings according to 95% family-wise confidence interval.
 
@@
 
-Table: (\#tab:MLG-aggressiveness) Mean aggressiveness ratings for the first 10 MLH; groupings according to 95% family-wise confidence interval.
+Table: (\#tab:MLG-aggressiveness) Mean aggressiveness ratings for the 10 MLH most abundant; groupings according to 95% family-wise confidence interval.
 

### 2017-09-07 doc/RMD/mlg-mcg.Rmd
remove labels from nodes where n = 1

@@
                  arrow = arrow(length = unit(0.01, "native"))) +
-  geom_node_text(aes(label = gsub("MLG.", "", name), # ifelse(size > 0.1, gsub("MLG.", "", name), ""), 
+  geom_node_text(aes(label = ifelse(size > 0.1, gsub("MLG.", "", name), ""), 
                      color = type, 

### 2017-09-07 doc/RMD/MCG-virulence.Rmd
Add test for assessor

@@
 dat <- read_csv(file.path(PROJHOME, "data", "clean_data.csv"), col_types = datcols) %>%
-  select(Severity, MCG, Region, Source)
+  select(Severity, MCG, Region, Source, Year)
 dat
@@
 
-## Visualizing distributions
+# Assessing virulence by MCG
 
@@
 
+## Testing for differences by assessor
+
+The straw test, until the end of 2007, was performed by Lindsey Otto-Hanson. 
+After that, these were performed by Serena McCoy. The Steadman lab was careful
+to train their members consistently in these practices, so the results from this
+test should be equivalent, but we want to ensure that there are no hidden biases
+between the two. To do this, we will test for differences within region.
+
+```{r assessor-diff}
+set.seed(2017-06-29)
+assessor_dat <- dat %>%
+  group_by(Region, Year) %>%
+  filter(n() > 5) %>%
+  mutate(mean_sev = mean(Severity)) %>%
+  mutate(Source = ifelse(Source == "wmn", "wmn", "producer field")) %>%
+  arrange(desc(mean_sev)) %>%
+  ungroup() %>%
+  mutate(Region = forcats::fct_inorder(Region)) %>%
+  mutate(Assessor = factor(ifelse(Year <= 2007, "Otto-Hanson", "McCoy"))) %>% 
+  select(Region, Assessor, Severity, Source)
+res <- aov(Severity ~ Region + Assessor, data = assessor_dat)
+res
+summary(res)
+```
+
+# Assessing virulence by multilocus genotype
+
 ```{r fig.width = 5, fig.height = 3.5}

### 2017-09-07 doc/RMD/pop-diff.Rmd
attempt to adjust legend width in DAPC fig

@@
 ```{r dapc-region-dualplot, fig.width = 13, fig.height = 6}
-LEG <- cowplot::get_legend(LDS_PLOT + theme(legend.position = "right") + theme(legend.text = element_text(size = 12)))
+LEG <- cowplot::get_legend(LDS_PLOT + 
+                             theme(legend.position = "right") + 
+                             theme(legend.margin = margin(t = 0, r = 0, b = 0, l = 0, unit = "lines")) +
+                             theme(legend.text = element_text(size = 12)))
 cowplot::plot_grid(LEG,

### 2017-09-07 doc/manuscript/manuscript.Rmd
update discussion

@@
 To date, with 366 isolates, this is the largest single population genetic study of *S. sclerotiorum* assessing population structure within managed and unmanaged agricultural environments. 
-These findings indicate that the wmn can be effective at screening for potential resistant lines within growing regions.
+These findings indicate that the white mold screening nurseries can be effective at screening for potential resistant lines within growing regions.
 
@@
 
+We found that te best predictors of genetic structure are Region and Year, supporting the hypothesis that *S. sclerotiorum* populations are spatially structured [@carbone2001multilocus].
 Borrowing a technique often used in the ecological literature, we used dbRDA to elucidate the effect of all variables (`r andlist(c(names(strata(dat11)), "Aggressiveness"))`) [@legendre1999distance].
-From the initial results, it appears that the most important factors for predicting genetic structure are MCG, region, and year.
+From the initial results, it appeared that the most important factors for predicting genetic structure were MCG, region, and year.
 When we inspected the biplot of the initial results, we saw that the most important predictors were 'MCG 44', 'MCG 5', and 'MCG 9'. 
-We believe that this is driven by the fact that these particular MCGs have uneven MLH distributions, meaning that they are heavily associated with one particular MLH (Fig. \@ref(biggraph)).
+We believe that this was driven by the fact that these particular MCGs have uneven MLH distributions, meaning that they are heavily associated with one particular MLH (Fig. \@ref(biggraph)).
 We note these results with caution because of the apparent multicolinearity between MCG and Region, which is a violation of the analysis [@legendre1999distance].
 While the results indicated that Mexico and the year 2005 were the two most important variables, it's worth noting that all Mexico isolates were collected in 2005 (Fig. \@ref(rda-biplot)).
-The results also show that the Vista cultivar explains some of the variance, but this represents six isolates in MI. 
-Aggressiveness and specific source field had little to no effect on prediction of genetic diversity. 
+The results also show that the Vista cultivar explains some of the variance, but this represents six isolates in MI, and thus we cannot draw broad conclusions from this axis. 
+Aggressiveness and source field had little to no effect on prediction of genetic diversity. 
 These results are in agreement with studies that examined differentiation based on Host [@aldrich-wolfe2015genetic] and Aggressiveness [@atallah2004high; @attanayake2012genetic; @attanayake2013sclerotinia] reporting little or no correlation of genetic diversity to these variables.
@@
 
-<!-- ## Genotypic Structure {-} -->
+<!-- ## Aggressiveness {-} -->
 
-The results show several haplotypes shared across populations, except for Mexico and two states that had fewer than three samples each (Fig. \@ref(community-graph)).
-Our network-based approach by treating the haplotypes as edges and weighting each edge with the inverse of $P_{sex}$ treated the edges as springs connecting the populations with the strength proportional to the probability of obtaining the same haplotype as a clone.
-This allowed us to use a graph walking algorithm to see how close the populations were simply based off of the proportion of clones they shared.
-The most abundant haplotype was shared across four populations, but its high value of $P_{sex}$ meant that it did not contribute significantly to the overall structure. 
-The graph walking algorithm was able to divide the network into three groups, but had a modularity of 0.17, which indicates that the groups are only weakly differentiated.
-
-The widespread nature of multilocus haplotypes in both wmn and production fields with relatively small values of $P_{sex}$ may indicate the spread of inoculum between regions. 
-It has previously been shown that infections can be transmitted through seed [@botelho2013performance], but seedborne transmission is thought to be of insignificant epidemiological importance [@strausbaugh2003management].
-Thus, we hypothesize that shared haplotypes between populations may arise due to transmission events of seed or sclerotia.
-However, because we observe population structure based on Region, we speculate that these transmission events are rare.
+While aggressiveness was not shown to predict genetic structure, it is an important factor in breeding efforts, and we observed significant differences in aggressiveness based on region (Fig.
+\@ref(aggressiveness-boxplot), Table \@ref(tab:region-aggressiveness)).
+These results show a similar pattern to what was found previously in @otto-hanson2011variation with the exception of North Dakota, which increased in mean aggressiveness from 5 to 5.77.
+This increase was due in part to new data from producer field isolates collected after the previous study.
+These straw tests were performed by a different person for these later isolates, which could suggest a more lenient or strict scoring system.
+However, when we examined the within-region differences, we found no significant effect by individual. 
+Many of the ND isolates fell within the 6--7 range, which denotes a physical boundary (disease symptoms around the second node) between intermediate and susceptible [@otto-hanson2011variation]. 
+Thus, we observed a shift in aggressiveness without a significant shift in genotypic structure, which may indicate that aggressiveness may be controlled by environmental factors as opposed to genetic profile.
 
@@
 We found no evidence of private alleles in the Mexico and CA isolates, suggesting that this pattern separating these populations from the others was not an artifact.
+We believe that the differences we observe from Mexico may reflect differences in climate.
 
+Many of the isolates in our study were from temperate climates and the only isolates representing a sub-tropical climate were from Mexico.
 It has been proposed within the *S.
@@
 This has been attributed to the notion that the fungus has the chance to undergo more reproductive cycles in the warmer climate [@carbone2001multilocus; @attanayake2013sclerotinia]. 
-Populations in sub-tropical regions of China have been found to be more variable, sexually reproducing, and unrelated to populations in temperate regions of the USA [@attanayake2013sclerotinia].
-The differentiation between the Mexico isolates and all other regions gives some evidence for this hypothesis as the only isolates in our study from sub-tropical regions are in Mexico.
+The strongest evidence to date supporting this hypothesis is from @attanayake2013sclerotinia, showing that populations in sub-tropical regions of China have been found to be more variable, sexually reproducing, and unrelated to populations in temperate regions of the USA.
+This result however, may be driven more by geography and agricultural practice as opposed to climate.
 
-We were unable to find evidence for sexual reproduction in any region except for CA and Australia. 
+<!-- ## Genotypic Structure {-} -->
+
+The results from our shared haplotype analysis show several populations with at least one haplotype between them, except for Mexico and two states that had fewer than three samples each (Fig.
+\@ref(community-graph)).
+Our network-based approach by treating the haplotypes as edges and weighting each edge with the inverse of $P_{sex}$ treated the edges as springs connecting the populations with the strength proportional to the probability of obtaining the same haplotype as a clone.
+This allowed us to use a graph walking algorithm to see how close the populations were simply based off of the proportion of clones they shared.
+The most abundant haplotype was shared across four populations, but its high value of $P_{sex}$ meant that it did not contribute significantly to the overall structure. 
+The graph walking algorithm was able to divide the network into three groups, but had a modularity of 0.17, which indicates that the groups are only weakly differentiated.
+
+The widespread nature of multilocus haplotypes in both wmn and production fields with relatively small values of $P_{sex}$ may indicate the spread of inoculum between regions. 
+While seedborne transmission is thought to be of insignificant epidemiological importance [@strausbaugh2003management], it has since been shown that *S.
+sclerotiorum* infections can be transmitted through seed [@botelho2013performance].
+Thus, we hypothesize that shared haplotypes between populations may arise due to transmission events of seed or sclerotia.
+This could explain the fact that we see shared genotypes with low $P_{sex}$ values shared between France, Australia, and the US.
+While we speculate that these transmission events are rare due to the genetic structuring by Region, these results suggest that seedborne infections may indeed reflect a source of inoculum.
+This may, in turn increase the risk of introducing new sources of genetic variation through potential outcrossing events.
+
+When we tested for sexual reproduction, we were unable to find evidence for it in any region except for CA and Australia. 
 While the Australia population had a non-significant value of $\bar{r}_d$---which would suggest that we cannot reject the null hypothesis of random mating---the sample size was insufficient from which to draw conclusions [@milgroom1996recombination; @agapow2001indices].
-The low value of $\bar{r}_d$ in the CA population may represent sexual reproduction, but this could also be an artifact of sampling two different populations, which is known to reduce the value of $\bar{r}_d$ [@prugnolle2010apparent].
-
-<!-- ## Aggressiveness {-} -->
-
-We observed significant differences in aggressiveness based on region (Fig. \@ref(aggressiveness-boxplot), Table \@ref(tab:region-aggressiveness)). 
-These results show a similar pattern to what was found previously with the exception of North Dakota, which increased in mean aggressiveness from 5 to 5.77 [@otto-hanson2011variation].
-This increase was due in part to new data from producer field isolates collected after the previous study.
-These straw tests were performed by a different person for these later isolates, which could suggest a more lenient or strict scoring system.
-However, when we examined the within-region differences, we found no significant effect by individual. 
-Many of the ND isolates fell within the 6--7 range, which denotes a physical boundary (disease symptoms around the second node) between intermediate and susceptible [@otto-hanson2011variation]. 
-Thus, we observed a shift in aggressiveness without a significant shift in genotypic structure, which may indicate that aggressiveness may be controlled by environmental factors as opposed to genetics.
+The low value of $\bar{r}_d$ in the CA population may represent sexual reproduction, but we can see in Figure \@ref(DAPC-RY) that there is differentiation by year. 
+Thus, this could also be an artifact of sampling two different populations, which is known to reduce the value of $\bar{r}_d$ [@prugnolle2010apparent].
 
@@
 
-Mycelial Compatibility Groups have historically been used as a proxy for clonal lineages, and thus, of interest in this study was testing the association between multilocus haplotypes (MLHs) and mycelial compatibility groups (MCGs) [@leslie1993fungal; @otto-hanson2011variation; @kohn1990mycelial; @schafer2006optimized; @kohn1995clonal; @carbone1999patterns].
+The previous study of the white mold screening nursery populations used MCGs to assess genotypic diversity [@otto-hanson2011variation].
+Historically, MCGs have been used as a proxy for clonal lineages, and thus, of interest in this study was testing the association between multilocus haplotypes (MLHs) and mycelial compatibility groups (MCGs) [@leslie1993fungal; @otto-hanson2011variation; @kohn1990mycelial; @schafer2006optimized; @kohn1995clonal; @carbone1999patterns].
 Our results, however, do not support this assumption.
@@
 Some studies have shown a correlation between MCG and MLH [@aldrich-wolfe2015genetic; @carbone1999patterns; @lehner2015genetic], whereas other studies have shown no apparent correlation, even on small spatial scales [@atallah2004high; @attanayake2012genetic; @attanayake2013sclerotinia].
-
-<!-- 
-SEE:
-
-Perhaps add some discussion of independently evolving vegetatively compatible
-lineages. 
--->
 

### 2017-09-18 doc/RMD/RDA-analysis.Rmd
rerun RDA analysis
removed lable segments as requested by Syd.

@@
 # @param db a capscale object
+# @param arrows the number of arrows to plot
+# @param seed a random seed to control plotting of labels
+# @param lab a switch indicating if percent variance should be displayed on the
+#   axes (default: TRUE)
+# @param segal for the arrow labels, the segment alpha channel. Default to 1,
+#   but can be set to 0, indicating that the labels should disappear.
 # @return a ggplot2 object from the scores 
-plot_dbrda <- function(db, arrows = 10, seed = 2017-06-28, lab = TRUE){
+plot_dbrda <- function(db, arrows = 10, seed = 2017-06-28, lab = TRUE, segal = 1){
   set.seed(seed)
@@
     labs    <- vegan:::summary.cca(db, axes = 0)[["cont"]][["importance"]]["Proportion Explained", 1:2]
-    xl      <- paste0("Eig 1 (", round(labs[[1]]*100, 2), "% variance explained)")
-    yl      <- paste0("Eig 2 (", round(labs[[2]]*100, 2), "% variance explained)")
+    xl      <- paste0("Eig. 1 (", round(labs[[1]]*100, 2), "% variance explained)")
+    yl      <- paste0("Eig. 2 (", round(labs[[2]]*100, 2), "% variance explained)")
   } else {
@@
                      segment.color = "grey25",
+                     segment.alpha = segal,
                      data = Arrows) +
@@
 set.seed(999)
-plot_dbrda(cap11cc2, arrows = 5) + 
+plot_dbrda(cap11cc2, arrows = 5, segal = 0) + 
   theme_classic(base_size = 16, base_family = "Helvetica") + 

### 2017-09-18 doc/manuscript/manuscript.Rmd
fix typos and other issues in manuscript

@@
 
-<!-- 
-ZNK: The introduction needs a lot of work. Here's an outline of what I think
-     needs to go in here.
-Outline:
-
- - *S. sclerotiorum* basic biology
- - What have genetic studies and phenotypic studies found
- - What are we missing? A detailed analysis of *S. sclerotiorum* from diverse
-   regions that may potentially identify haplotypes or MCGs that may be good
-   tools for testing breeding efforts.
- - Our hypotheses and specific questions.
- -->
-
 *Sclerotinia sclerotiorum* (Lib.) de Bary is an ascomycete plant pathogen with a worldwide distribution.
-This is a necrotrophic pathogen that is primarily homothallic (self-fertilization) and has the ability to survive fore more than five years in soil using melanized survival structures called sclerotia [@bolton2006sclerotinia; @sexton2006population].
+This is a necrotrophic pathogen that is primarily homothallic (self-fertilization) and has the ability to survive for more than five years in soil using melanized survival structures called sclerotia [@bolton2006sclerotinia; @sexton2006population].
 It causes disease on more than 400 plant species belonging to 75 families [@boland1994index] including crops of major economic importance such as sunflower (*Helianthus spp.*), soybean (*Glycine max* L.), canola (*Brassica napa* L., *Brassica campestris* L.), and dry bean (*Phaseolus vulgaris* L.) [@bolton2006sclerotinia].
@@
 Mycelial compatibility has been used as a proxy for vegetative compatibility, a fungal trait controlled by several independent genes controlling the ability for two hyphae to fuse and grow as a single unit [@leslie1993fungal; @schafer2006optimized].
-Because of the genetic connection to vegetative compatibility, two isolates that are mycelially compatible were considered clones [@leslie1993fungal]; but correlation with genetic markers, such as microsatellites, have shown mixed results [@attanayake2012genetic; @lehner2017independently].
+Because of the genetic connection to vegetative compatibility, two isolates that are mycelially compatible were considered clones [@leslie1993fungal]; but correlation with genetic markers, such as microsatellites, have shown mixed results [@ford1995heterokaryon; @micali2003independence; @jo2008reassessment; @attanayake2012genetic; @papaioannou2014barrage; @lehner2017independently].
 
@@
 Broadly, isolates were collected from two sources: white mold screening nurseries (wmn) or producer fields.
-<!-- SEE: Insect or weed control? -->
 White mold screening nurseries were 5m x 10m in size and maintained without application of fungicides to observe natural incidence of white mold.
-The nursery plots were incorporated with a basal dressing of N:P:K = 1:3:2 and side dressing of 0:3:2 during the growing season [@steadman2003identification; @steadman2004identification; @steadman2005identification; @steadman2006identification; @otto2007identification; @otto2008identification; @mccoy2009use].
+The nursery plots were incorporated with a basal dressing of N:P:K = 1:3:2 and side dressing of 0:3:2 during the growing season [@steadman2003identification].
 
@@
 After collection, sclerotia were stored in Petri plates lined with filter paper, then stored at $20~^{\circ}$F or -$4~^{\circ}$C. 
-Sclerotia were surface-sterilized with 50% sodium hypochlorite (at least 6% NaOCL, The Clorox Company, Oakland, CA) solution for 3 min, and double rinsed with ddH$_2$O 3 min.
-The sterilized sclerotia were then plated on water agar plates (16g of Bacto agar per liter of ddH$_2$O, BD Diagnostic Systems, Sparks, MD), with four to five sclerotia of each isolate separated on each plate and stored on the counter top at room temperature for 5 to 6 days.
+Sclerotia were surface-sterilized with 50% Clorox bleach (at least 6% NaOCL, The Clorox Company, Oakland, CA) solution for 3 min, and double rinsed with ddH$_2$O 3 min.
+The sterilized sclerotia were then placed on water agar plates (16g of Bacto agar per liter of ddH$_2$O, BD Diagnostic Systems, Sparks, MD), with four to five sclerotia of each isolate separated on each plate and stored on the counter top at room temperature for 5 to 6 days.
 An 8-mm plug from a 5- or 6-day-old culture was transferred from the advancing margin of the mycelia onto a plate of Difco potato dextrose agar (PDA at 39 g/liter of ddH$_2$O) [@otto-hanson2011variation].
@@
 MCG was determined as described previously through co-culturing pairs of 2-day-old isolates 2.5 cm apart on Diana Sermons (DS) Medium [@cubeta2001mycelial]. 
-Incompatibility of different MCGs resulted in formation of a mycelial barrier of dead tissue (termed a barrage line) accompanied by formation of sclerotia on either side of the barrage line, indicating the limits of the isolates' growth [@otto-hanson2011variation; @kohn1990mycelial; @leslie1993fungal].
+Incompatibility of different MCGs resulted in formation of a barrage line accompanied by formation of sclerotia on either side of the barrage line, indicating the limits of the isolates' growth [@otto-hanson2011variation; @kohn1990mycelial; @leslie1993fungal].
 Isolates were compared in a pairwise manner for each site and then representatives among sites were compared to determine mycelial compatibility groups by scoring compatible and incompatible interactions [@otto-hanson2011variation].
-
-<!-- ![
-Examples of mycelial compatibility results in *S. sclerotiorum*. 
-A) Compatible reaction showing aerial hyphae.
-B) Incompatible reaction showing barrage reaction of dead cells at point of contact.
-](../../results/figures/publication/incompatibility2.JPG){#incompatibility width=25%} -->
+No MCGs were compatible with any other MCG.
 
@@
 Clear drinking straws cut to 2.5 cm and sealed were used to place two mycelial plugs of inoculum on the host plant after removing all plant growth &gt;2.5 cm above the fourth node.
-Measurements were taken 8 days later using the Modified Petzoldt and Dickson scale of 1--9, where 1 is no disease and 9 is plant death [@petzoldt1996straw; @teran2006modified].
+Measurements of the necrotic lesion were taken 8 days later using the Modified Petzoldt and Dickson scale of 1--9, where 1 is no disease and 9 is plant death [@petzoldt1996straw; @teran2006modified].
 
@@
 Capillary electrophoresis (fragment analysis) of amplicons, with size standard GeneScan™ 500 LIZ®, was performed using an ABI 3730 genetic analyzer (Life Technologies Corporation, Carlsbad, CA) at the Michigan State University Genomic Sequencing Center (East Lansing, MI).
-Alleles were scored using PeakScanner version 1.0 (Life Technologies Corporation, Carlsbad, CA) and recorded manually in an excel sheet.
+Alleles were scored using PeakScanner version 1.0 (Life Technologies Corporation, Carlsbad, CA) and recorded manually in a spreadsheet.
 
@@
 
-<!-- All analyses were performed on this data set unless otherwise mentioned. -->
 All data processing and analyses were performed in `r R.version.string %>% strsplit("\\(") %>% map_chr(1)`[@R] and are openly available and reproducible at https://github.com/everhartlab/sclerotinia-366/.
@@
 This was performed using a network-based approach where both MLHs and MCGs were considered nodes and the number of isolates in which they were found together was the strength of the connection between an MLH and and MCG node.
-The network-based approach allowed us to assess how "clumpy" the relationships between MLHs and MCGs were.
+The network-based approach allowed us to assess the associations between MLHs and MCGs.
 To construct the network, a contingency table was created with MLHs and MCGs and converted to a directed and weighted edgelist where each edge represented a connection from an MCG to an MLH, weighted by the number of samples shared in the connection.
@@
 Locus `r loc_table$Locus[loc_table$allele==2]`, which contained only 2 alleles, showed low values of both *h* (`r signif(loc_table$Hexp[loc_table$allele==2], 3)`) and evenness (`r signif(loc_table$Evenness[loc_table$allele==2], 3)`), indicating that there was one dominant allele present.
-Analysis of the haplotype accumulation curve showed no clear plateau (Supplemental Information), indicating that we would likely obtain more multilocus haplotypes if we were to genotype more loci.
+Analysis of the haplotype accumulation curve showed no clear plateau for 11 or 16 loci (Supplemental Information), indicating that we would likely obtain more multilocus haplotypes if we were to genotype more loci.
 
@@
 The transparency of the bars represent the evenness of the distribution of the MLHs within a given MCG.
-A dashed box surrounds the eight most common MCGs representing > 51% of the data.
+A dashed box surrounds the eight most common MCGs representing >&nbsp;51% of the data.
 **B)** Full graph-representation of the relationship between MCGs (open circles) and MLHs (filled circles).
@@
 ![
-Network of populations (nodes/circles) and their shared multilocus haplotypes (MLH) (edges/lines) haplotyped over 11 loci. 
+Network of populations (nodes/circles) and their shared multilocus haplotypes (MLH) (edges/lines) genotyped over 11 loci. 
 Each node is labeled with **name (number of MLHs shared/number of MLHs total).** 
@@
 
-The analysis of molecular variance (AMOVA) for clone-corrected samples over the hierarchy of Region, Source, and Year showed significant variation between regions and years, but no significant variation between wmn and non-wmn (Table \@ref(tab:amova)).
+The analysis of molecular variance (AMOVA) for clone-corrected samples over the hierarchy of Region, Source, and Year showed significant variation between Regions and Years, but no significant variation between wmn and producer fields (Table \@ref(tab:amova)).
 In contrast, when we compared the three cultivars, Beryl, Bunsi, and G122, we found no significant differentiation (Supplementary Information).
@@
 
-Table: (\#tab:amova) Comparison of populations in the white mold screening nurseries (wmn) and producer fields using an analysis of molecular variance on Bruvo's genetic distance showing little differentiation between white mold screening nurseries (wmn) and other sources.
+Table: (\#tab:amova) Comparison of populations in the white mold screening nurseries (wmn) and producer fields using an analysis of molecular variance on Bruvo's genetic distance showing no apparent differentiation between wmn and other sources.
 The hierarchy was constructed as Source/Region where source is defined as belonging to a wmn or producer field.
@@
 DAPC grouping by Region and Year used the first `r dapc_region_year$n.pca` principal components, representing `r prettify(dapc_region_year$var)`% of the total variance.
-The North Central US populations (NE, MI, CO, ND) did not appear to have any variation across time in contrast to WA, which showed a shift in population structure in the last year of sampling, 2008 (Fig.
+The North Central USA populations (NE, MI, CO, ND) did not appear to have any variation across time in contrast to WA, which showed a shift in population structure in the last year of sampling, 2008 (Fig.
 \@ref(DAPC-RY)).
@@
 
-<!-- 
-SEE:
-
-The foundation of the discussion is well built.
-Missing, however, is a synthesis of the information within each paragraph,
-connecting it from one section to the next. I suggest you move the paragraph on
-MCGs to the last (prior to limitations) and then delete all the sub-headings
-(keep limitations and conclusions sub-headers).
-
-I would also like to see more direct discussion of how the pouplation genetic
-can be interpreted epidemiologically. For some guidance here, I suggest
-revisiting Milgroom and Peever 2003 (10.1094/PDIS.2003.87.6.608).
-
-I also suggest supporting epidemiological interpretations with conclusions that
-other *S. sclerotiorum* population have shown. You've done an excellent job 
-using the broader *S. sclerotiorum* literature to support your claims without
-re-hashing the results. I think picking a couple of examples to highlight to
-compare and contrast with the current results would be helpful to readers.
--->
-
 In this study, we characterized the diversity of *Sclerotinia sclerotiorum* from dry bean fields across the United States.
-Our results suggest that, broadly, populations from white mold screening nurseries reflect the populations of the surrounding regions, indicating that resistance screening may be successful at a within regions.
+Our results suggest that, broadly, populations from white mold screening nurseries reflect the populations of the surrounding regions, indicating that resistance screening may be successful within regions.
 We found significant population differentiation by geographic region and year, mainly differentiated into three broad North American groups based on shared haplotypes and posterior groupings, a Coastal Region, Midwestern Region, and Mexico.
@@
 
-We found that te best predictors of genetic structure are Region and Year, supporting the hypothesis that *S. sclerotiorum* populations are spatially structured [@carbone2001multilocus].
+We found that the best predictors of genetic structure are Region and Year, supporting the hypothesis that *S. sclerotiorum* populations are spatially structured [@carbone2001multilocus].
 Borrowing a technique often used in the ecological literature, we used dbRDA to elucidate the effect of all variables (`r andlist(c(names(strata(dat11)), "Aggressiveness"))`) [@legendre1999distance].
@@
 
-While aggressiveness was not shown to predict genetic structure, it is an important factor in breeding efforts, and we observed significant differences in aggressiveness based on region (Fig.
+While aggressiveness was not shown to predict genetic structure, it is an important factor in breeding efforts, and we observed significant differences in aggressiveness based on Region (Fig.
 \@ref(aggressiveness-boxplot), Table \@ref(tab:region-aggressiveness)).
@@
 This is illustrated in both Table \@ref(tab:mlg-table) and Fig. \@ref(biggraph)B,C. 
-It could be argued, however that 'MLH 25', with its high value of $P_{sex}$ represents different true MLHs across the five MCGs it occupies, but this doesn't account for the overall structure of Fig.
+It could be argued, however that 'MLH 25', with its high value of $P_{sex}$ represents different true MLHs across the five MCGs it occupies, but this does not account for the overall structure of Fig.
 \@ref(fullgraph) where, for example, 'MLH 75' ($P_{sex}$ = 1.81e^-4^) is compatible with `r big_haplotype$N` other haplotypes through three MCG when the population structure of *S.

### 2017-09-18 doc/RMD/mlg-mcg.Rmd
converting to joyplots to ridgeline plots
See here for details:
http://serialmentor.com/blog/2017/9/15/goodbye-joyplots

@@
 
-if (require("ggjoy")){
+if (require("ggridges")){
   ggplot(distplot, aes(x = dist, y = COMMS, group = COMMS, height = ..density..)) + 
-    geom_joy(aes(lty = comparison, fill = mean)) + 
-    theme_joy() + 
+    geom_density_ridges(aes(lty = comparison, fill = mean)) + 
+    theme_ridges() + 
     theme(axis.text.y = element_text(hjust = 0)) + 

### 2017-09-18 doc/manuscript/manuscript.tex
clean results and rerun analyses

%% Submissions for peer-review must enable line-numbering
%% using the lineno option in the \documentclass command.
%%
%% Preprints and camera-ready submissions do not need
%% line numbers, and should have this option removed.
%%
%% Please note that the line numbering option requires
%% version 1.1 or newer of the wlpeerj.cls file, and
%% the corresponding author info requires v1.2

\documentclass[fleqn,10pt,lineno]{wlpeerj} % for journal submissions

% ZNK -- Adding headers for pandoc

\setlength{\emergencystretch}{3em}
\providecommand{\tightlist}{
\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{lipsum}
\usepackage[unicode=true]{hyperref}
\usepackage{longtable}

\usepackage{indentfirst}

\title{Population structure and phenotypic variation of \emph{Sclerotinia
sclerotiorum} from dry bean in the United States}

\author[1]{Zhian N. Kamvar}

\author[1]{Bimal Sajeewa Amadarasa}

\author[1]{Rachana Jhala}

\author[1]{Serena McCoy}

\author[1]{James Steadman}

\author[1]{Sydney E. Everhart}

\corrauthor[1]{Sydney E. Everhart}{\href{mailto:everhart@unl.edu}{\nolinkurl{everhart@unl.edu}}}

\affil[1]{Department of Plant Pathology, University of Nebraska-Lincoln, Lincoln
68583}

%
% \author[1]{First Author}
% \author[2]{Second Author}
% \affil[1]{Address of first author}
% \affil[2]{Address of second author}
% \corrauthor[1]{First Author}{f.author@email.com}

%

\begin{abstract}
The ascomycete pathogen \emph{Sclerotinia sclerotiorum} is a
necrotrophic pathogen on over 400 known host plants, and is the causal
agent of white mold on dry bean. Currently, there are no known cultivars
of dry bean with complete resistance to white mold. For more than 20
years, bean breeders have been using white mold screening nurseries with
natural populations of \emph{S. sclerotiorum} to screen new cultivars
for resistance. It is thus important to know if the genetic diversity in
populations of \emph{S. sclerotiorum} within these nurseries a) reflect
the genetic diversity of the populations in the surrounding region and
b) are stable over time. Furthermore, previous studies have investigated
the correlation between mycelial compatibility groups (MCG) and
multilocus haplotypes (MLH), but none have formally tested these
patterns. We genotyped 366 isolates of \emph{S. sclerotiorum} from
producer fields and white mold screening nurseries surveyed over 10
years in 2003--2012 representing 11 states, Mexico, France, and
Australia at 11 microsatellite loci resulting in 165 MLHs. Populations
were loosely structured over space and time based on analysis of
molecular variance and discriminant analysis of principal components,
but not by cultivar, aggressiveness, or field source. Of all the regions
tested, only Mexico (n=18) shared no MLHs with any other region. Using a
bipartite network-based approach, we found no evidence that the MCGs
accurately represent MLHs. Our study suggests that breeders should
continue to test dry bean lines in several white mold screening
nurseries across the US to account for both the phenotypic and genotypic
variation that exists across regions.
% Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text. Dummy abstract text.
\end{abstract}

\usepackage{amsthm}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}{Lemma}[section]
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{proposition}{Proposition}[section]
\theoremstyle{definition}
\newtheorem{example}{Example}[section]
\theoremstyle{definition}
\newtheorem{exercise}{Exercise}[section]
\theoremstyle{remark}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\begin{document}

\flushbottom
\maketitle
\thispagestyle{empty}

\section*{Introduction}\label{introduction}
\addcontentsline{toc}{section}{Introduction}

\emph{Sclerotinia sclerotiorum} (Lib.) de Bary is an ascomycete plant
pathogen with a worldwide distribution. This is a necrotrophic pathogen
that is primarily homothallic (self-fertilization) and has the ability
to survive for more than five years in soil using melanized survival
structures called sclerotia (Bolton et al., 2006; Sexton et al., 2006).
It causes disease on more than 400 plant species belonging to 75
families (Boland \& Hall, 1994) including crops of major economic
importance such as sunflower (\emph{Helianthus spp.}), soybean
(\emph{Glycine max} L.), canola (\emph{Brassica napa} L., \emph{Brassica
campestris} L.), and dry bean (\emph{Phaseolus vulgaris} L.) (Bolton et
al., 2006). On dry bean, \emph{S. sclerotiorum} is the causal agent of
white mold, a devastating disease that can be a yield-limiting in
temperate climates (Steadman, 1983).

Currently, there are no known resistant cultivars of dry bean
commercially available (Otto-Hanson et al., 2011). Organized breeding
efforts have used a common-garden approach with white mold screening
nurseries in dry bean production areas across the United States with
additional sites in Mexico, France, and Australia (Steadman et al.,
2003, 2004, 2005, 2006; Otto-Hanson \& Steadman, 2007, 2008; McCoy \&
Steadman, 2009). These white mold screening nurseries use no chemical or
cultural treatments against \emph{S. sclerotiorum} and employ
standardized protocols for screening new cultivars for resistance to
white mold. It has previously been shown that aggressiveness (the
severity of disease symptoms on the host) is significantly different
across white mold screening nursery sites in separate geographic regions
(Otto-Hanson et al., 2011). The genetic structure and mode of
reproduction in these populations, however, is currently unknown.

Understanding genetic relationships and reproduction behavior of
\emph{S. sclerotiorum} populations is beneficial for breeders seeking to
develop new resistant cultivars for worldwide deployment (Milgroom,
1996; McDonald \& Linde, 2002). In particular, genetically diverse
populations with high rates of sexual reproduction are more likely to
overcome host resistance. Most populations of \emph{S. sclerotiorum} are
predominantly clonal with low genetic diversity and have a large degree
of population fragmentation (Kohli et al., 1995; Cubeta et al., 1997;
Kohli \& Kohn, 1998; Carbone \& Kohn, 2001; Ekins et al., 2011;
Attanayake et al., 2012). Some studies, however have found populations
that show signatures of sexual reproduction (Atallah et al., 2004;
Sexton \& Howlett, 2004; Attanayake et al., 2013; Aldrich-Wolfe et al.,
2015).

Nearly all population genetic studies of \emph{S. sclerotiorum} employ a
macroscopic assay to determine mycelial compatibility, the ability for
fungal hyphae from different colonies to appear to grow together without
forming a barrier of dead cells between them (known as a barrage line)
(Leslie, 1993; Sirjusingh \& Kohn, 2001). Mycelial compatibility has
been used as a proxy for vegetative compatibility, a fungal trait
controlled by several independent genes controlling the ability for two
hyphae to fuse and grow as a single unit (Leslie, 1993; Schafer \& Kohn,
2006). Because of the genetic connection to vegetative compatibility,
two isolates that are mycelially compatible were considered clones
(Leslie, 1993); but correlation with genetic markers, such as
microsatellites, have shown mixed results (Ford et al., 1995; Micali \&
Smith, 2003; Jo et al., 2008; Attanayake et al., 2012; Papaioannou \&
Typas, 2014; Lehner et al., 2017).

In this study, we analyze and characterize the genetic and phenotypic
diversity of 366 \emph{S. sclerotiorum} isolates collected over 10 years
between 2003 and 2012 from dry bean cultivars among different geographic
locations in the United States, France, Mexico, and Australia. We wanted
to know if the \emph{S. sclerotiorum} populations from white mold
screening nurseries were representative of the fields within the same
region. As these nurseries were not treated with any chemical or
cultural control of white mold, we hypothesized that these nurseries
would represent the natural population of \emph{S. sclerotiorum}.
Furthermore, we wanted to investigate the potential effect of cultivar
on genetic diversity by assessing three dry bean cultivars with
different levels of resistance, Beryl (great northern bean,
susceptible), Bunsi (navy bean, low susceptibility), and G122 (cranberry
bean, partial resistance) (Otto-Hanson et al., 2011). We additionally
wanted to determine categorical or phenotypic variables that best
predicted genetic structure and if there was correlation between
multilocus haplotype and mycelial compatibility group. Knowing what
variables predict genetic structure can help direct breeding efforts. By
investigating these aims, we hope to effectively describe the population
structure of \emph{S. sclerotiorum} in the USA and make available our
database of isolates for use in future dry bean breeding efforts.

\section*{Materials and Methods}\label{materials-and-methods}
\addcontentsline{toc}{section}{Materials and Methods}

\subsection*{Isolate collection}\label{isolate-collection}
\addcontentsline{toc}{subsection}{Isolate collection}

Several of the isolates used for this study were collected as reported
in previous studies using the same methods (Otto-Hanson et al., 2011).
Broadly, isolates were collected from two sources: white mold screening
nurseries (wmn) or producer fields. White mold screening nurseries were
5m x 10m in size and maintained without application of fungicides to
observe natural incidence of white mold. The nursery plots were
incorporated with a basal dressing of N:P:K = 1:3:2 and side dressing of
0:3:2 during the growing season (Steadman et al., 2003).

Sampling was carried out by collecting sclerotia from diseased tissue in
zig-zag transects across field plots. Because sampling depended on
disease incidence, the number of samples isolated varied from year to
year. Although the nursery locations were the same over sampling years,
sampling plots within a location varied for sampling years.

Sclerotia of \emph{S. sclerotiorum} were collected over several years
from grower fields and/or wmn in 11 states of the United States of
America, Mexico, France, and Australia (Table \ref{tab:isolate-table}).
After collection, sclerotia were stored in Petri plates lined with
filter paper, then stored at \(20~^{\circ}\)F or -\(4~^{\circ}\)C.
Sclerotia were surface-sterilized with 50\% Clorox bleach (at least 6\%
NaOCL, The Clorox Company, Oakland, CA) solution for 3 min, and double
rinsed with ddH\(_2\)O 3 min. The sterilized sclerotia were then placed
on water agar plates (16g of Bacto agar per liter of ddH\(_2\)O, BD
Diagnostic Systems, Sparks, MD), with four to five sclerotia of each
isolate separated on each plate and stored on the counter top at room
temperature for 5 to 6 days. An 8-mm plug from a 5- or 6-day-old culture
was transferred from the advancing margin of the mycelia onto a plate of
Difco potato dextrose agar (PDA at 39 g/liter of ddH\(_2\)O)
(Otto-Hanson et al., 2011). In combination with the 156 isolates
described previously, we collected 210 isolates for a total of 366
isolates (Otto-Hanson et al., 2011).

\subsection*{Mycelial Compatibility}\label{mycelial-compatibility}
\addcontentsline{toc}{subsection}{Mycelial Compatibility}

MCG was determined as described previously through co-culturing pairs of
2-day-old isolates 2.5 cm apart on Diana Sermons (DS) Medium (Cubeta et
al., 2001). Incompatibility of different MCGs resulted in formation of a
barrage line accompanied by formation of sclerotia on either side of the
barrage line, indicating the limits of the isolates' growth (Kohn et
al., 1990; Leslie, 1993; Otto-Hanson et al., 2011). Isolates were
compared in a pairwise manner for each site and then representatives
among sites were compared to determine mycelial compatibility groups by
scoring compatible and incompatible interactions (Otto-Hanson et al.,
2011). No MCGs were compatible with any other MCG.

\subsection*{Aggressiveness}\label{aggressiveness}
\addcontentsline{toc}{subsection}{Aggressiveness}

Aggressiveness of each isolate was assessed using a straw test as
described in Otto-Hanson et al. (2011) that rated necrotic lesion size
(Petzoldt \& Dickson, 1996; Teran et al., 2006). Briefly, the straw test
uses 21-day-old G122 plants as the host in a greenhouse setting. Clear
drinking straws cut to 2.5 cm and sealed were used to place two mycelial
plugs of inoculum on the host plant after removing all plant growth
\textgreater{}2.5 cm above the fourth node. Measurements of the necrotic
lesion were taken 8 days later using the Modified Petzoldt and Dickson
scale of 1--9, where 1 is no disease and 9 is plant death (Petzoldt \&
Dickson, 1996; Teran et al., 2006).

\subsection*{Microsatellite genotyping}\label{microsatellite-genotyping}
\addcontentsline{toc}{subsection}{Microsatellite genotyping}

Prior to DNA extraction, isolates were grown on PDA and plugs were
subsequently transferred to Potato Dextrose Broth (PDB) where they were
grown until there was significant mycelial growth, but before the
mycelial mat became solidified (4--5 days). Mycelia for each plate was
collected in a buchner funnel, agar plugs removed, lyophilized and
pulverized manually in Whirl-pak® HDPE sampling bags (Sigma-Aldrich,
St.~Louis, MO). Lyophilized mycelia was then stored in microcentrifuge
tubes at -\(20~^{\circ}\)C until needed for DNA extraction. DNA from
25mg of pulverized mycelia was purified using a phenol-chloroform
extraction method followed by alcohol precipitation and evaporation,
suspending the DNA in 200\(\mu\)l TE (Sambrook et al., 1989). Suspended
DNA was stored at \(4~^{\circ}\)C until genotyping.

We genotyped each \emph{S. sclerotiorum} isolate using 16 microsatellite
primer pairs developed previously (Sirjusingh \& Kohn, 2001). PCR was
carried out as described previously, using primers labeled with FAM
fluorophore. Resulting amplicons were first resolved in a 1.5\% agarose
gel stained with ethidium bromide to ensure product was within the
expected size range prior to capillary electrophoresis. Capillary
electrophoresis (fragment analysis) of amplicons, with size standard
GeneScan™ 500 LIZ®, was performed using an ABI 3730 genetic analyzer
(Life Technologies Corporation, Carlsbad, CA) at the Michigan State
University Genomic Sequencing Center (East Lansing, MI). Alleles were
scored using PeakScanner version 1.0 (Life Technologies Corporation,
Carlsbad, CA) and recorded manually in a spreadsheet.

\subsection*{Data processing and
Analysis}\label{data-processing-and-analysis}
\addcontentsline{toc}{subsection}{Data processing and Analysis}

All data processing and analyses were performed in R version 3.4.1 (R
Core Team, 2017) and are openly available and reproducible at
\url{https://github.com/everhartlab/sclerotinia-366/}. Of the 16
microsatellite loci genotyped, five included compound repeats, which
made it challenging to accurately/confidently bin alleles into fragment
sizes expected for each locus based on the described repeat motif. Loci
with compound repeats were removed for the reported statistics. To
ensure the integrity of the results we additionally processed these loci
and included them in concurrent analyses. We assessed the power of our
11 markers by generating a genotype accumulation curve in the R package
\emph{poppr} version 2.5.0, looking for evidence of saturation, which
would indicate that loci were sufficiently sampled to adequately
represent the full set of haplotypes (Arnaud-Hanod et al., 2007; Kamvar
et al., 2015). To avoid including isolates potentially collected from
the same plant (which increases the probability of collecting sclerotia
from the same point of infection more than once), data were
clone-corrected on a hierarchy of Region/Source/Host/Year---meaning that
duplicated genotypes were reduced to a single observation when they were
collected in the same year from the same host cultivar located in the
same source field (wmn or producer)---for subsequent analysis. We
assessed haplotype diversity by calculating Stoddart and Taylor's index
(\emph{G}) (Stoddart \& Taylor, 1988), Shannon's index (\emph{H})
(Shannon, 1948), Simpson's index (\(\lambda\)) (Simpson, 1949), evenness
(\(E_5\)), and the expected number of multilocus haplotypes
(\emph{eMLH}) (Hurlbert, 1971; Heck et al., 1975; Pielou, 1975; Grünwald
et al., 2003). To assess the potential for random mating, we tested for
linkage disequilibrium with the index of association, \(I_A\) and its
standardized version, \(\bar{r}_d\) using 999 permutations (Brown et
al., 1980; Smith et al., 1993; Agapow \& Burt, 2001). Both haplotype
diversity and linkage disequilibrium were calculated in \emph{poppr}
(Kamvar et al., 2014).

\subsection*{Assessing Importance of
Variables}\label{assessing-importance-of-variables}
\addcontentsline{toc}{subsection}{Assessing Importance of Variables}

\subsubsection*{Distance-based Redundancy
Analysis}\label{distance-based-redundancy-analysis}
\addcontentsline{toc}{subsubsection}{Distance-based Redundancy Analysis}

A distance-based redundancy analysis (dbRDA) (Legendre \& Anderson,
1999) was performed with the function \texttt{capscale()} in the
\emph{vegan} package version 2.4.4 (Oksanen et al., 2017). This method
uses constrained ordinations on a distance matrix representing the
response variable to delineate relative contribution of any number of
independent explanatory variables. We used this method to delineate the
phenotypic (Aggressiveness, Mycelial Compatibility Group (MCG)),
geographic (Region, Host, Location), and temporal (Year) components in
predicting genetic composition of the populations. The distance matrix
we used as the response variable was generated using Bruvo's genetic
distance from clone-corrected data (procedure described above) as
implemented in \emph{poppr}, which employed a stepwise mutation model
for microsatellite data (Bruvo et al., 2004; Kamvar et al., 2014).
Because aggressiveness measures differed between isolates that were
reduced to a single observation during clone-correction, aggressiveness
was first averaged across clone-corrected isolates. To identify
explanatory variable(s) correlated with genetic variation, a
forward-backward selection process was applied with the vegan function
\texttt{ordistep()}. An analysis of variance (ANOVA) was then performed
to test for significance of the reduced model and marginal effects using
999 permutations. The \texttt{varpart()} function of vegan was used to
determine variation partitioning of explanatory variables.

\subsubsection*{Aggressiveness
Assessment}\label{aggressiveness-assessment}
\addcontentsline{toc}{subsubsection}{Aggressiveness Assessment}

We used ANOVA to assess if aggressiveness was significantly different
with respect to Region, MCG, or multilocus haplotype (MLH). To minimize
complications due to small sample sizes, we chose the top 10 MCGs,
representing 56.5\% of the isolates collected, the 10 most abundant MLHs
representing 26.7\% of the isolates, and populations with more than five
isolates. If ANOVA results were significantly different at \(\alpha\) =
0.05, pairwise differences were assessed using Tukey's HSD test
(\(\alpha\) = 0.05) using the \texttt{HSD.test()} function in the
package \emph{agricolae} version 1.2.8 (Mendiburu \& Simon, 2015).

\subsubsection*{Correlating Multilocus Haplotypes with Mycelial
Compatibility
Groups}\label{correlating-multilocus-haplotypes-with-mycelial-compatibility-groups}
\addcontentsline{toc}{subsubsection}{Correlating Multilocus Haplotypes
with Mycelial Compatibility Groups}

We wanted to assess if there was correlation between MLHs and MCGs. This
was performed using a network-based approach where both MLHs and MCGs
were considered nodes and the number of isolates in which they were
found together was the strength of the connection between an MLH and and
MCG node. The network-based approach allowed us to assess the
associations between MLHs and MCGs. To construct the network, a
contingency table was created with MLHs and MCGs and converted to a
directed and weighted edgelist where each edge represented a connection
from an MCG to an MLH, weighted by the number of samples shared in the
connection. This was then converted to a bipartite graph where top nodes
represented MLHs and bottom nodes represented MCGs. To identify clusters
of MLHs and MCGs within the network, we used the cluster walktrap
community detection algorithm as implemented in the
\texttt{cluster\_walktrap()} function in \emph{igraph} version 1.1.2
(Csardi \& Nepusz, 2006; Pons \& Latapy, 2006). This algorithm attempts
to define clusters of nodes by starting at a random nodes and performing
short, random ``walks'' along the edges between nodes, assuming that
these walks would stay within clusters. For this analysis, we set the
number of steps within a walk to four and allowed the algorithm to use
the edge weights in determining the path. All of the resulting
communities that had fewer than 10 members were then consolidated into
one. Community definitions were used to assess the average genetic
distance (as defined by Bruvo's distance) within members of the
community (Bruvo et al., 2004).

\subsection*{Genetic Diversity}\label{genetic-diversity}
\addcontentsline{toc}{subsection}{Genetic Diversity}

\subsubsection*{Population
Differentiation}\label{population-differentiation}
\addcontentsline{toc}{subsubsection}{Population Differentiation}

We used analysis of molecular variance (AMOVA) with Bruvo's genetic
distance in \emph{poppr} to test for differentiation between populations
in wmn and producer fields from the same region and collected in the
same year (Excoffier et al., 1992; Bruvo et al., 2004; Kamvar et al.,
2014). To identify Regions with greater differentiation, we used
discriminant analysis of principal components (DAPC) as implemented in
\emph{adegenet} version 2.1.0, assessing the per-sample posterior group
assignment probability (Jombart, 2008). This method decomposes the
genetic data into principal components, and then uses a subset of these
as the inputs for discriminant analysis, which attempts to minimize
within-group variation and maximize among-group variation (Jombart et
al., 2010). To avoid over-fitting data, the optimal number of principal
components was selected by using the \emph{adegenet} function
\texttt{xvalDapc()}. This function implements a cross-validation
procedure to iterate over an increasing number of principal components
on a subset (90\%) of the data, trying to find the minimum number of
principal components that maximizes the rate of successful group
reassignment. To assess if cultivar had an influence on genetic
diversity between wmn, we first subset the clone-corrected data to
contain only samples from wmn and from the cultivars Beryl, Bunsi, and
G122 and tested differentiation using AMOVA and DAPC as described above.
We additionally assessed population stability over time by calculating
DAPC over the combined groups of Region and Year as described above.

\subsubsection*{Analysis of Shared Multilocus
Haplotypes}\label{analysis-of-shared-multilocus-haplotypes}
\addcontentsline{toc}{subsubsection}{Analysis of Shared Multilocus
Haplotypes}

We wanted to evaluate patterns of connectivity between shared multilocus
haplotypes across geographic regions. We first tabulated the multilocus
haplotypes shared between at least two populations (defined as states or
countries) with the \emph{poppr} function \texttt{mlg.crosspop()}
(Kamvar et al., 2014). From these data, we constructed a graph with
populations as nodes and shared haplotypes as edges (connections)
between nodes using the R packages \emph{igraph} (Csardi \& Nepusz,
2006), \emph{dplyr} version 0.7.3 (Wickham et al., 2017), and
\emph{purrr} version 0.2.3 (Henry \& Wickham, 2017). Each node was
weighted by the fraction of shared MLHs. Each edge represented a single
MLH, but because a single MLH could be present in more than one
population, that MLH would have a number of edges equivalent to the
total number of possible connections, calculated as
(\emph{n}*(\emph{n}-1))/2 edges where n represents the number of
populations crossed. Edges were weighted by \(1-P_{sex}\), where
\(P_{sex}\) is the probability of encountering the same haplotype via
two independent meiotic events (Parks \& Werth, 1993; Arnaud-Hanod et
al., 2007). This weighting scheme would thus strengthen the connection
of edges that represented genotypes with a low probability of being
produced via sexual reproduction. We then identified communities (among
the Regions) in the graph using the \texttt{cluster\_optimal()} function
from \emph{igraph} (Csardi \& Nepusz, 2006). The graph was plotted using
the R packages \emph{ggplot2} version 2.2.1 (Wickham, 2009) and
\emph{ggraph} 1.0.0 (Pedersen, 2017). To ensure that we captured the
same community signal, we additionally performed this analysis including
the five polymorphic markers described above.

\section*{Results}\label{results}
\addcontentsline{toc}{section}{Results}

A total of 366 isolates were collected from 2003 to 2012 (except 2006
and 2011) from diseased dry bean plants in Australia, Mexico, France,
and 11 states in the USA (Table \ref{tab:isolate-table}). We observed a
total of 165 MLHs (215 with 16 loci). Over 50\% of the isolates came
from four states, MI (62), ND (60), WA (59), NE (47). Four regions had
fewer than 10 isolates, Australia (6), WI (2), NY (1), ID (1). We
observed 87 MCGs, the most abundant of which (`MCG 5') was represented
by 73 isolates over 37 MLHs (Fig \ref{biggraph}A,C).

The number of observed alleles per locus ranged from two to 10 with an
average of 6.27 (Table \ref{tab:locus-stats}). Locus 20-3, which
contained only 2 alleles, showed low values of both \emph{h} (0.0533)
and evenness (0.42), indicating that there was one dominant allele
present. Analysis of the haplotype accumulation curve showed no clear
plateau for 11 or 16 loci (Supplemental Information), indicating that we
would likely obtain more multilocus haplotypes if we were to genotype
more loci.

After clone-correction on the hierarchy of Region/Source/Host/Year, we
were left with 318 isolates representing 165 MLHs (Table
\ref{tab:iatab}). The results show that most populations exhibited
relatively high amounts of genotypic diversity, with the exception of MI
and NE, which exhibited lower-than average \(E_5\) values. Mexico had
the lowest average value of gene diversity, \emph{h} and the lowest
\emph{eMLH}, suggesting low overall genetic diversity. Nearly all
populations showed evidence for linkage (Table \ref{tab:iatab}), which
serves as evidence for clonal reproduction or other forms of non-random
mating. The only exceptions are CA and Australia. Both of these
populations showed only moderate significance (P \textless{} 0.05) with
\(\bar{r}_d\) values of 0.03 and 0.12, respectively.

\begin{longtable}[]{@{}lrrrrr@{}}
\caption{\label{tab:locus-stats} Allelic diversity on full data set at loci
used in this study. \emph{h} = Nei's Gene Diversity (Nei, 1978). Average
\emph{h} = 0.583, average Evenness = 0.693, average no. alleles =
6.27}\tabularnewline
\toprule
\begin{minipage}[b]{0.09\columnwidth}\raggedright\strut
Locus\strut
\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedleft\strut
Range\strut
\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\raggedleft\strut
Repeat Motif\strut
\end{minipage} & \begin{minipage}[b]{0.15\columnwidth}\raggedleft\strut
No. alleles\strut
\end{minipage} & \begin{minipage}[b]{0.08\columnwidth}\raggedleft\strut
\emph{h}\strut
\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedleft\strut
Evenness\strut
\end{minipage}\tabularnewline
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{0.09\columnwidth}\raggedright\strut
Locus\strut
\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedleft\strut
Range\strut
\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\raggedleft\strut
Repeat Motif\strut
\end{minipage} & \begin{minipage}[b]{0.15\columnwidth}\raggedleft\strut
No. alleles\strut
\end{minipage} & \begin{minipage}[b]{0.08\columnwidth}\raggedleft\strut
\emph{h}\strut
\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedleft\strut
Evenness\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
5-2\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
318--324\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedleft\strut
di-\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedleft\strut
4\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
0.45\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
0.62\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
6-2\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
483--495\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedleft\strut
hexa-\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedleft\strut
3\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
0.64\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
0.95\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
7-2\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
158--174\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedleft\strut
di-\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedleft\strut
7\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
0.73\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
0.76\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
8-3\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
244--270\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedleft\strut
di-\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedleft\strut
7\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
0.74\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
0.79\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
9-2\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
360--382\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedleft\strut
di-\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedleft\strut
9\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
0.35\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
0.41\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
12-2\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
214--222\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedleft\strut
di-\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedleft\strut
5\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
0.58\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
0.78\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
17-3\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
342--363\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedleft\strut
tri-\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedleft\strut
7\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
0.55\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
0.53\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
20-3\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
280--282\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedleft\strut
di-\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedleft\strut
2\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
0.05\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
0.42\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
55-4\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
153--216\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedleft\strut
tetra-\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedleft\strut
10\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
0.72\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
0.66\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
110-4\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
370--386\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedleft\strut
tetra-\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedleft\strut
5\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
0.76\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
0.91\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
114-4\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
339--416\strut
\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedleft\strut
tetra-\strut
\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedleft\strut
10\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
0.83\strut
\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
0.80\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

\begin{longtable}[]{@{}lrrrrrrrl@{}}
\caption{\label{tab:iatab} Genotypic diversity and Linkage Disequilibrium
summary for geographic populations arranged by abundance after
clone-correction by a hierarchy of Region/Source/Host/Year. Pop =
Population, N = number of individuals (number of MLH in parentheses),
eMLH = expected number of MLHs based on rarefaction at 10 individuals
(standard error in parentheses), H = Shannon-Weiner Index, G = Stoddardt
and Taylor's Index, \(\lambda\) = Simpson's Index, \emph{h} = Nei's 1978
gene diversity, \(E_5\) = Evenness, \(\bar{r}_d\) = standardized index
of association. An asterix indicates a significant value of
\(\bar{r}_d\) after 999 permutations, \(P \leq 0.001\).}\tabularnewline
\toprule
\begin{minipage}[b]{0.11\columnwidth}\raggedright\strut
Pop\strut
\end{minipage} & \begin{minipage}[b]{0.09\columnwidth}\raggedleft\strut
N\strut
\end{minipage} & \begin{minipage}[b]{0.14\columnwidth}\raggedleft\strut
eMLH\strut
\end{minipage} & \begin{minipage}[b]{0.05\columnwidth}\raggedleft\strut
H\strut
\end{minipage} & \begin{minipage}[b]{0.05\columnwidth}\raggedleft\strut
G\strut
\end{minipage} & \begin{minipage}[b]{0.11\columnwidth}\raggedleft\strut
\(\lambda\)\strut
\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\raggedleft\strut
\(E_5\)\strut
\end{minipage} & \begin{minipage}[b]{0.06\columnwidth}\raggedleft\strut
\emph{h}\strut
\end{minipage} & \begin{minipage}[b]{0.06\columnwidth}\raggedright\strut
\(\bar{ r}_d\)\strut
\end{minipage}\tabularnewline
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{0.11\columnwidth}\raggedright\strut
Pop\strut
\end{minipage} & \begin{minipage}[b]{0.09\columnwidth}\raggedleft\strut
N\strut
\end{minipage} & \begin{minipage}[b]{0.14\columnwidth}\raggedleft\strut
eMLH\strut
\end{minipage} & \begin{minipage}[b]{0.05\columnwidth}\raggedleft\strut
H\strut
\end{minipage} & \begin{minipage}[b]{0.05\columnwidth}\raggedleft\strut
G\strut
\end{minipage} & \begin{minipage}[b]{0.11\columnwidth}\raggedleft\strut
\(\lambda\)\strut
\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\raggedleft\strut
\(E_5\)\strut
\end{minipage} & \begin{minipage}[b]{0.06\columnwidth}\raggedleft\strut
\emph{h}\strut
\end{minipage} & \begin{minipage}[b]{0.06\columnwidth}\raggedright\strut
\(\bar{ r}_d\)\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
WA\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedleft\strut
58 (56)\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedleft\strut
9.95 (0.23)\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
4.0\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
54.3\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedleft\strut
0.98\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedleft\strut
0.98\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
0.60\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
0.07*\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
MI\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedleft\strut
58 (43)\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedleft\strut
9.3 (0.79)\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
3.6\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
29.0\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedleft\strut
0.97\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedleft\strut
0.78\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
0.54\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
0.14*\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
ND\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedleft\strut
41 (35)\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedleft\strut
9.44 (0.73)\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
3.5\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
25.9\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedleft\strut
0.96\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedleft\strut
0.82\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
0.54\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
0.1*\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
NE\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedleft\strut
37 (28)\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedleft\strut
8.93 (0.94)\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
3.2\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
17.8\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedleft\strut
0.94\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedleft\strut
0.75\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
0.55\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
0.25*\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
CO\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedleft\strut
34 (28)\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedleft\strut
9.46 (0.67)\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
3.3\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
24.1\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedleft\strut
0.96\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedleft\strut
0.92\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
0.56\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
0.27*\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
France\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedleft\strut
21 (14)\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedleft\strut
8.5 (0.85)\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
2.6\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
12.6\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedleft\strut
0.92\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedleft\strut
0.95\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
0.48\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
0.11*\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
CA\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedleft\strut
18 (15)\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedleft\strut
9.12 (0.72)\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
2.7\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
13.5\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedleft\strut
0.93\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedleft\strut
0.94\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
0.51\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
0.03\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
OR\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedleft\strut
17 (13)\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedleft\strut
8.52 (0.85)\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
2.5\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
10.7\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedleft\strut
0.91\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedleft\strut
0.89\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
0.47\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
0.1*\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
Mexico\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedleft\strut
15 (9)\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedleft\strut
7.1 (0.85)\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
2.1\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
7.3\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedleft\strut
0.86\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedleft\strut
0.89\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
0.28\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
0.37*\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
MN\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedleft\strut
9 (7)\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedleft\strut
7 (0)\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
1.9\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
6.2\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedleft\strut
0.84\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedleft\strut
0.93\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
0.47\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
0.19*\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
Australia\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedleft\strut
6 (6)\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedleft\strut
6 (0)\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
1.8\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
6.0\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedleft\strut
0.83\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedleft\strut
1.00\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
0.48\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
0.12\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
WI\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedleft\strut
2 (2)\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedleft\strut
2 (0)\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
0.7\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
2.0\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedleft\strut
0.50\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedleft\strut
1.00\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
0.27\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
-\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
NY\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedleft\strut
1 (1)\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedleft\strut
1 (0)\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
0.0\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
1.0\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedleft\strut
0.00\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedleft\strut
NaN\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
NaN\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
-\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
ID\strut
\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedleft\strut
1 (1)\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedleft\strut
1 (0)\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
0.0\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
1.0\strut
\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedleft\strut
0.00\strut
\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedleft\strut
NaN\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
NaN\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
-\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

\begin{figure}
\centering
\includegraphics[width=1.00000\textwidth]{../../results/figures/publication/mcg-subgraph-with-context.pdf}
\caption{Associations between Mycelial Compatibility Groups and
Multilocus Haplotypes. \textbf{A)} Barplot of Mycelial Compatibility
Group (MCG) abundance in descending order. Singletons (46) were
truncated, leaving 41 MCGs. White bars represent sample counts and grey
bars represent counts of unique multilocus haplotypes (MLH). The
transparency of the bars represent the evenness of the distribution of
the MLHs within a given MCG. A dashed box surrounds the eight most
common MCGs representing \textgreater{}~51\% of the data. \textbf{B)}
Full graph-representation of the relationship between MCGs (open
circles) and MLHs (filled circles). Details in Fig. \ref{fullgraph}.
\textbf{C)} A subset of \textbf{B} representing the 8 most common MCGs
and their associated MLHs (dashed box in \textbf{A}). Filled nodes
(circles) represent MLHs and open nodes represent MCGs. Node area scaled
to the number of samples represented (range: 1--73). Numbers inside
nodes are the MLH/MCG label (if n \textgreater{} 1). Edges (arrows)
point from MLH to MCG where the weight (thickness) of the edge
represents the number of shared isolates (range: 1--19). Edges extending
from MLHs displayed to other MCGs are not shown.}\label{biggraph}
\end{figure}

\subsection*{Variable Assessment}\label{variable-assessment}
\addcontentsline{toc}{subsection}{Variable Assessment}

\subsubsection*{Variable Contributions}\label{variable-contributions}
\addcontentsline{toc}{subsubsection}{Variable Contributions}

\begin{figure}
\centering
\includegraphics[width=0.61000\textwidth]{../../results/figures/publication/Figure7Z.pdf}
\caption{Biplot showing five most influential explanatory variables
(arrows) overlayed on the first two eigenvectors of distance based
redundancy analysis of \emph{Sclerotinia sclerotiorum} isolates. The
length of the arrows are directly proportional to the strength of the
correlation between explanatory and molecular variables. Open circles
represent the 318 clone-corrected haplotypes in ordination
space.}\label{rda-biplot}
\end{figure}

\begin{figure}
\centering
\includegraphics{../../results/figures/publication/aggressiveness.pdf}
\caption{Strip Plot of aggressiveness by population arranged in
descending order of mean aggressiveness for all populations with N
\textgreater{} 5. White bars represent mean value. Circles represent
individual isolates where filled circles are isolates from white mold
screening nurseries (wmn) and open circles are isolates from producer
fields.}\label{aggressiveness-boxplot}
\end{figure}

The forward-backward selection process of the dbRDA models on
clone-corrected data revealed Year, Region, Host, and MCG to be the
optimal variables for the reduced model, accounting for 45\% of the
total variation. ANOVA showed that the reduced model was significant
with an adjusted R\textsuperscript{2} of 0.0675 (\emph{P} = 0.001).
Assessment of the marginal effects showed that all varaibles
significantly explained genetic variation (\emph{P} = 0.001, \emph{P} =
0.007 for Region). We found that there was multicollinearity when MCG
was combined with any other variable, so repeated the analysis, dropping
MCG from the list of potential predictors. From these results, Year,
Region, Host, and Aggressiveness were found to be optimal, accounting
for 17.6\% of the total variation. ANOVA revealed significant effects
with an adjusted R\textsuperscript{2} of 0.0325 (\emph{P} = 0.001).
While the marginal effect assessment revealed that Year, Region, and
Host significantly explained variation at \emph{P} = 0.001, and
Aggressiveness significantly explained variation at \emph{P} = 0.039.
Much of the variation appeared to be driven by isolates from Mexico and
2005 (Fig. \ref{rda-biplot}). Variance partitioning of the independent
variables without MCG indicated aggressiveness to be the least
influential factor with 0.1\% contributing to explaining the variation
of molecular data, whereas the combination of variables accounted for
3.3\%.

\subsubsection*{Aggressiveness}\label{aggressiveness-1}
\addcontentsline{toc}{subsubsection}{Aggressiveness}

Agressiveness of the isolates ranged from 1.4 to 7.9 with a mean of 5.02
and median of 4.85. The group mean averages were 4.88, 5.13, and 5.19
for Region, MCG, and MLH, respectively. Our assessment of aggressiveness
in association with Region showed a significant effect (\emph{P}
\textless{} 1.00e\textsuperscript{-4}), with means that ranged from 5.8
(MN) to 4.0 (CA) (Fig. \ref{aggressiveness-boxplot}, Table
\ref{tab:region-aggressiveness}). MCGs also showed a significant effect
(\emph{P} \textless{} 1.00e\textsuperscript{-4}), with means that ranged
from 6.0 (`MCG 44') to 4.6 (`MCG 49'; Table
\ref{tab:MCG-aggressiveness}). We additionally found a significant
effect for MLHs (\emph{P} = 7.44e\textsuperscript{-4}) for MLHs, with
means that ranged from 6.0 (`MLH 78') to 4.3 (`MLH 140') (Table
\ref{tab:MLG-aggressiveness}).

\subsubsection*{Correlation of Mulitlocus Haplotypes and Mycelial
Compatibility
Groups}\label{correlation-of-mulitlocus-haplotypes-and-mycelial-compatibility-groups}
\addcontentsline{toc}{subsubsection}{Correlation of Mulitlocus
Haplotypes and Mycelial Compatibility Groups}

In our analysis, we found 165 MLHs with 70 singletons and 87 MCGs with
43 singletons (Fig. \ref{biggraph}A,B) where the eight most abundant
MCGs represented \textgreater{} 51\% of the data over 11 Regions, and
all years except for 2012. Our network-based approach to correlating
MLHs with MCGs revealed a large and complex network (Fig.
\ref{biggraph}, Table \ref{tab:mlg-table}). Community analysis showed 51
communities, 15 of which consisted of a single MLH unconnected with any
other community indicating that just 9.09\% of the 165 MLHs are unable
to cross with any other MLH in this data set (Fig. \ref{fullgraph}). The
three communities with the most members contained eight of the 10 most
abundant MCGs. Comparing these communities with Bruvo's genetic distance
showed an average distance of 0.451 among communities and an average
distance of 0.437 within communities, which were not significantly
different. When we assessed the number of times two different MLHs that
are in the same MCG, considering these as potential heterothallic
pairings that could result in sexual recombination, we found an average
of 14.3 potential heterothallic parings per MLH. Representing just four
isolates, `MLH 75' had 57 neighbors that shared the same mycelial
compatibility group (Fig \ref{biggraph}, \ref{fullgraph}). Overall,
there was no clear pattern to the association between MLH and MCGs.

\begin{longtable}[]{@{}llll@{}}
\caption{\label{tab:mlg-table} The five most abundant Multilocus Haplotypes
(MLH) with the probability of second encounter (\(P_{sex}\)), Mycelial
Compatibility Groups (MCG), and Regions with sample sizes in
parentheses.}\tabularnewline
\toprule
MLH & \(P_{sex}\) & MCG & Region\tabularnewline
\midrule
\endfirsthead
\toprule
MLH & \(P_{sex}\) & MCG & Region\tabularnewline
\midrule
\endhead
25 & 0.016824 & 5 & ND (15), CO (2), MI (2)\tabularnewline
& & 13 & ND (3)\tabularnewline
& & 60 & ND (2), WA (1)\tabularnewline
& & 1 & NE (1)\tabularnewline
& & 4 & MI (1)\tabularnewline
163 & 0.049932 & 45 & CO (5), ND (2), NE (1)\tabularnewline
& & 5 & MI (7)\tabularnewline
65 & 0.000071 & 44 & NE (10)\tabularnewline
& & 5 & MI (1)\tabularnewline
140 & 0.000155 & 8 & CO (5)\tabularnewline
& & 5 & MI (3)\tabularnewline
& & 20 & MI (2)\tabularnewline
66 & 0.000016 & 9 & NE (4), CO (2), MI (2)\tabularnewline
\bottomrule
\end{longtable}

\subsection*{Structure of Shared Multilocus
Haplotypes}\label{structure-of-shared-multilocus-haplotypes}
\addcontentsline{toc}{subsection}{Structure of Shared Multilocus
Haplotypes}

The most abundant MLH was represented by 27 isolates (Table
\ref{tab:mlg-table}) from five Regions (NE, MI, WA, CO, and ND). Within
Regions, haplotypes were relatively evenly distributed with moderate to
high diversity (Table \ref{tab:iatab}). Of the 165 MLHs, 76 (46\%) were
found in at least two Regions, except those found in WI (2), ID (1), and
Mexico (18) (Fig. \ref{community-graph}).

We had performed an analysis on a network where the connections
represented shared MLHs across populations, weighted by \(1 - P_{sex}\)
(Fig. \ref{community-graph}, Table \ref{tab:mlg-table}). Community
analysis of the MLHs shared between populations revealed 4 communities
with a modularity of 0.17: A coastal community (CA, OR, WA, and NY), a
midwest community (CO, NE, ND, MI), and an international community (MN,
France, Australia). Although analysis with 16 loci resulted in the
removal of the NY node because it no longer shared a haplotype with OR,
the same overall community structure was present with a modularity of
0.2 (Fig. \ref{community-graph16}). Relative to the US, the
international community appears to be driven by MLH 4, which is shared
between all three populations and has a \(P_{sex}\) value of
2.87e\textsuperscript{-5}, in contrast to the abundant MLH 25, which has
a \(P_{sex}\) value of 0.0168.

\begin{figure}
\centering
\includegraphics[width=0.61000\textwidth]{../../results/figures/publication/mlg-11.pdf}
\caption{Network of populations (nodes/circles) and their shared
multilocus haplotypes (MLH) (edges/lines) genotyped over 11 loci. Each
node is labeled with \textbf{name (number of MLHs shared/number of MLHs
total).} The shade and area of the nodes are proportional to the number
of unique MLHs within the node and the inner nodes are proportional to
the number of private MLHs to the region (bottom legend). Each edge
represents a single MLH where its thickness represents the number of
populations that share the MLH and the shade represents the value of
\(P_{sex}\), or the probability of encountering that MLH from two
independent meiotic events.}\label{community-graph}
\end{figure}

\subsection*{Population
Differentiation}\label{population-differentiation-1}
\addcontentsline{toc}{subsection}{Population Differentiation}

\subsubsection*{Analysis of Molecular
Variance}\label{analysis-of-molecular-variance}
\addcontentsline{toc}{subsubsection}{Analysis of Molecular Variance}

The analysis of molecular variance (AMOVA) for clone-corrected samples
over the hierarchy of Region, Source, and Year showed significant
variation between Regions and Years, but no significant variation
between wmn and producer fields (Table \ref{tab:amova}). In contrast,
when we compared the three cultivars, Beryl, Bunsi, and G122, we found
no significant differentiation (Supplementary Information).

\begin{longtable}[]{@{}lrrrrr@{}}
\caption{\label{tab:amova} Comparison of populations in the white mold
screening nurseries (wmn) and producer fields using an analysis of
molecular variance on Bruvo's genetic distance showing no apparent
differentiation between wmn and other sources. The hierarchy was
constructed as Source/Region where source is defined as belonging to a
wmn or producer field. Bold \(\Phi\) values indicate significant
difference (\emph{P} \textless{} 0.05). S.S. = Sum of Squares, d.f. =
degrees of freedom.}\tabularnewline
\toprule
\begin{minipage}[b]{0.31\columnwidth}\raggedright\strut
Hierarchy\strut
\end{minipage} & \begin{minipage}[b]{0.06\columnwidth}\raggedleft\strut
d.f.\strut
\end{minipage} & \begin{minipage}[b]{0.06\columnwidth}\raggedleft\strut
S.S.\strut
\end{minipage} & \begin{minipage}[b]{0.14\columnwidth}\raggedleft\strut
\% variation\strut
\end{minipage} & \begin{minipage}[b]{0.19\columnwidth}\raggedleft\strut
\(\Phi\) statistic\strut
\end{minipage} & \begin{minipage}[b]{0.05\columnwidth}\raggedleft\strut
\emph{P}\strut
\end{minipage}\tabularnewline
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{0.31\columnwidth}\raggedright\strut
Hierarchy\strut
\end{minipage} & \begin{minipage}[b]{0.06\columnwidth}\raggedleft\strut
d.f.\strut
\end{minipage} & \begin{minipage}[b]{0.06\columnwidth}\raggedleft\strut
S.S.\strut
\end{minipage} & \begin{minipage}[b]{0.14\columnwidth}\raggedleft\strut
\% variation\strut
\end{minipage} & \begin{minipage}[b]{0.19\columnwidth}\raggedleft\strut
\(\Phi\) statistic\strut
\end{minipage} & \begin{minipage}[b]{0.05\columnwidth}\raggedleft\strut
\emph{P}\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.31\columnwidth}\raggedright\strut
Between Region\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
13\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
10.19\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedleft\strut
8.45\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
\textbf{0.0845}\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
0.032\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.31\columnwidth}\raggedright\strut
Between Source within Region\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
8\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
2.74\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedleft\strut
-2.29\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-0.0250\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
0.498\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.31\columnwidth}\raggedright\strut
Between Year within Source\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
22\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
9.37\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedleft\strut
16.28\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
\textbf{0.173}\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
0.001\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.31\columnwidth}\raggedright\strut
Within Year\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
274\strut
\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
47.30\strut
\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedleft\strut
77.56\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
\textbf{0.224}\strut
\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
0.001\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

\newpage

\subsubsection*{Discriminant Analysis of Principal
Components}\label{discriminant-analysis-of-principal-components}
\addcontentsline{toc}{subsubsection}{Discriminant Analysis of Principal
Components}

Discriminant Analysis of Principal Components (DAPC) was performed
grouping by Region with the first 20 principal components, representing
87.1\% of the total variance. The first discriminant axis (representing
35\% of the discriminatory power) separated the centroid for the Mexico
isolates from the rest of the data, indicating strong differentiation
(Fig. \ref{DAPC}b). The second discriminant axis, representing 24.1\% of
the discriminatory power, separated the centroid for the CA isolates.
The mean population assignment probabilities for all populations with n
\textgreater{} 10 showed that only isolates from Mexico, CA, and France
had \textgreater{} 50\% probabilities of being reassigned to their
source populations (Fig. \ref{DAPC}a).

DAPC grouping by cultivar used the first 20 principal components,
representing 89\% of the total variance. The first two discriminant axes
(representing 38.5\% of the discriminatory power) failed to separate any
of the cultivars where the mean posterior assignment probabilities were
34\% (G122), 35.9\% (Beryl), and 30.1\% (Bunsi). DAPC grouping by Region
and Year used the first 15 principal components, representing 80.3\% of
the total variance. The North Central USA populations (NE, MI, CO, ND)
did not appear to have any variation across time in contrast to WA,
which showed a shift in population structure in the last year of
sampling, 2008 (Fig. \ref{DAPC-RY}). Further analysis of this population
revealed that all 12 isolates in WA circa 2008 originated in a wmn; nine
haplotypes were shared with CA, and three were shared with France (Fig.
\ref{community-graph}, \ref{community-graph16}).

\begin{figure}
\centering
\includegraphics[width=1.00000\textwidth]{../../results/figures/publication/DAPC.pdf}
\caption{Discriminant Analysis of Principal Components on regions
showing that Mexico is differentiated from other populations.
\textbf{A)} Scatter plot of first two components from DAPC. Points
represent observed individuals connected to the population centroids
with ellipses representing a 66\% confidence interval for a normal
distribution. The center of each component is represented as black grid
lines. \textbf{B)} Mean population assignment probability from the DAPC
for all populations with N \textgreater{} 10 (facets). Populations
represented along the horizontal axis and probability of assignment on
the vertical. Numbers next to source populations indicate population
size. All values sum to one.}\label{DAPC}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.61000\textwidth]{../../results/figures/publication/dapc_region_year_micanewa.pdf}
\caption{Scatter plot of Discriminant Analysis of Principal Components
on Regions and Years showing non-differentiated temporal variation NE
and MI and temporal variation in WA and CA. Points (text labels)
represent observed individuals connected to the population centroids
with ellipses representing a 66\% confidence interval for a normal
distribution. The center of each component is represented as black grid
lines. A more detailed view is shown in Fig.
\ref{DAPC-RY-FULL}.}\label{DAPC-RY}
\end{figure}

\section*{Discussion}\label{discussion}
\addcontentsline{toc}{section}{Discussion}

In this study, we characterized the diversity of \emph{Sclerotinia
sclerotiorum} from dry bean fields across the United States. Our results
suggest that, broadly, populations from white mold screening nurseries
reflect the populations of the surrounding regions, indicating that
resistance screening may be successful within regions. We found
significant population differentiation by geographic region and year,
mainly differentiated into three broad North American groups based on
shared haplotypes and posterior groupings, a Coastal Region, Midwestern
Region, and Mexico. To date, with 366 isolates, this is the largest
single population genetic study of \emph{S. sclerotiorum} assessing
population structure within managed and unmanaged agricultural
environments. These findings indicate that the white mold screening
nurseries can be effective at screening for potential resistant lines
within growing regions.

We found that the best predictors of genetic structure are Region and
Year, supporting the hypothesis that \emph{S. sclerotiorum} populations
are spatially structured (Carbone \& Kohn, 2001). Borrowing a technique
often used in the ecological literature, we used dbRDA to elucidate the
effect of all variables (MCG, Region, Source, Year, Host, and
Aggressiveness) (Legendre \& Anderson, 1999). From the initial results,
it appeared that the most important factors for predicting genetic
structure were MCG, region, and year. When we inspected the biplot of
the initial results, we saw that the most important predictors were `MCG
44', `MCG 5', and `MCG 9'. We believe that this was driven by the fact
that these particular MCGs have uneven MLH distributions, meaning that
they are heavily associated with one particular MLH (Fig.
\ref{biggraph}). We note these results with caution because of the
apparent multicolinearity between MCG and Region, which is a violation
of the analysis (Legendre \& Anderson, 1999). While the results
indicated that Mexico and the year 2005 were the two most important
variables, it's worth noting that all Mexico isolates were collected in
2005 (Fig. \ref{rda-biplot}). The results also show that the Vista
cultivar explains some of the variance, but this represents six isolates
in MI, and thus we cannot draw broad conclusions from this axis.
Aggressiveness and source field had little to no effect on prediction of
genetic diversity. These results are in agreement with studies that
examined differentiation based on Host (Aldrich-Wolfe et al., 2015) and
Aggressiveness (Atallah et al., 2004; Attanayake et al., 2012, 2013)
reporting little or no correlation of genetic diversity to these
variables. This indicates that a) breeders should keep in mind regional
differences when assessing resistance and b) it is possible that we have
not yet measured biologically relevant variables that can predict
genetic differentiation, which could include variables such as soil
community composition.

While aggressiveness was not shown to predict genetic structure, it is
an important factor in breeding efforts, and we observed significant
differences in aggressiveness based on Region (Fig.
\ref{aggressiveness-boxplot}, Table \ref{tab:region-aggressiveness}).
These results show a similar pattern to what was found previously in
Otto-Hanson et al. (2011) with the exception of North Dakota, which
increased in mean aggressiveness from 5 to 5.77. This increase was due
in part to new data from producer field isolates collected after the
previous study. These straw tests were performed by a different person
for these later isolates, which could suggest a more lenient or strict
scoring system. However, when we examined the within-region differences,
we found no significant effect by individual. Many of the ND isolates
fell within the 6--7 range, which denotes a physical boundary (disease
symptoms around the second node) between intermediate and susceptible
(Otto-Hanson et al., 2011). Thus, we observed a shift in aggressiveness
without a significant shift in genotypic structure, which may indicate
that aggressiveness may be controlled by environmental factors as
opposed to genetic profile.

The primary interest of this study was to assess if isolates sampled
from white mold screening nurseries represent isolates from producer
fields within the region (Steadman et al., 2003; Otto-Hanson et al.,
2011). According to our AMOVA results, we have evidence for
differentiation at the Region and Year, but little to no differentiation
between wmn isolates and production field isolates (Table
\ref{tab:amova}). This lack of differentiation, however, may reflect the
breeder practice of inoculating screening plots with sclerotia collected
from sources within the region. When we analyze the AMOVA results in
light of the DAPC results (Fig. \ref{DAPC}), it becomes clear that the
regional patterns of differentiation are largely driven by isolates from
Mexico and CA. Isolates from these Regions had a higher posterior
probability (\textgreater{} 0.75) of being reassigned to their own
populations than any other (Fig. \ref{DAPC}A). All other populations in
comparison (except France) has reassignment probabilities of \textless{}
0.5, which is reflected in the failure of the first two discriminant
functions to separate these populations (Fig. \ref{DAPC}B).

Despite the evidence that Mexico and CA contributed to much of the
population differentiation, Regions like WA still had a large amount of
internal variation. The two distinct clusters for the WA Region showed
that the 2008 population appeared differentiated and, under further
investigation, we found that all the haplotypes from this year were
shared between CA and France (Fig. \ref{community-graph}, \ref{DAPC-RY},
\ref{DAPC-RY-FULL}). All of the isolates from WA in 2003--2005, and 2008
came from the same wmn; within the wmn, those in 2003--2005 came a
Northeastern field location cropped with dry bean since 2002, and those
in 2008 from a Southeastern field that was previously cropped with
brassica, sundgrass, peas, beans, and potatoes (Miklas, Phil Pers.
comm.). Both of these fields were inoculated with sclerotia in 2002, the
Northeastern field with sclerotia provided by a commercial bean producer
and the Southeastern field with sclerotia from peas (although this was
thought to be unsuccessful). Despite this information, it is still
unclear what has contributed to the differentiation of the 2008
population from WA. While the evidence my suggest host as being a
factor, previous studies have shown no significant differentiation
across host species (Aldrich-Wolfe et al., 2015).

With the exception of the WA Region, populations that were sampled
across several years appeared to be relatively stable over time with
overlapping distributions in the DAPC (i.e.~NE and MI, Fig.
\ref{DAPC-RY}). DAPC is based on the principal components of Euclidean
distance between genotypes (Jombart et al., 2010). Unlike Bruvo's
distance, Euclidean distance does not take into account the magnitude of
the difference between alleles, which could inflate the distance measure
in the presence of private alleles (Bruvo et al., 2004). We found no
evidence of private alleles in the Mexico and CA isolates, suggesting
that this pattern separating these populations from the others was not
an artifact. We believe that the differences we observe from Mexico may
reflect differences in climate.

Many of the isolates in our study were from temperate climates and the
only isolates representing a sub-tropical climate were from Mexico. It
has been proposed within the \emph{S. sclerotiorum} literature that
isolates from sub-tropical and tropical climates are differentiated or
more variable than populations from temperate climates {[}Carbone \&
Kohn (2001); Attanayake et al. (2013); Lehner \& Mizubuti (2017);{]}.
This has been attributed to the notion that the fungus has the chance to
undergo more reproductive cycles in the warmer climate (Carbone \& Kohn,
2001; Attanayake et al., 2013). The strongest evidence to date
supporting this hypothesis is from Attanayake et al. (2013), showing
that populations in sub-tropical regions of China have been found to be
more variable, sexually reproducing, and unrelated to populations in
temperate regions of the USA. This result however, may be driven more by
geography and agricultural practice as opposed to climate.

The results from our shared haplotype analysis show several populations
with at least one haplotype between them, except for Mexico and two
states that had fewer than three samples each (Fig.
\ref{community-graph}). Our network-based approach by treating the
haplotypes as edges and weighting each edge with the inverse of
\(P_{sex}\) treated the edges as springs connecting the populations with
the strength proportional to the probability of obtaining the same
haplotype as a clone. This allowed us to use a graph walking algorithm
to see how close the populations were simply based off of the proportion
of clones they shared. The most abundant haplotype was shared across
four populations, but its high value of \(P_{sex}\) meant that it did
not contribute significantly to the overall structure. The graph walking
algorithm was able to divide the network into three groups, but had a
modularity of 0.17, which indicates that the groups are only weakly
differentiated.

The widespread nature of multilocus haplotypes in both wmn and
production fields with relatively small values of \(P_{sex}\) may
indicate the spread of inoculum between regions. While seedborne
transmission is thought to be of insignificant epidemiological
importance (Strausbaugh \& Forster, 2003), it has since been shown that
\emph{S. sclerotiorum} infections can be transmitted through seed
(Botelho et al., 2013). Thus, we hypothesize that shared haplotypes
between populations may arise due to transmission events of seed or
sclerotia. This could explain the fact that we see shared genotypes with
low \(P_{sex}\) values shared between France, Australia, and the US.
While we speculate that these transmission events are rare due to the
genetic structuring by Region, these results suggest that seedborne
infections may indeed reflect a source of inoculum. This may, in turn
increase the risk of introducing new sources of genetic variation
through potential outcrossing events.

When we tested for sexual reproduction, we were unable to find evidence
for it in any region except for CA and Australia. While the Australia
population had a non-significant value of \(\bar{r}_d\)---which would
suggest that we cannot reject the null hypothesis of random mating---the
sample size was insufficient from which to draw conclusions (Milgroom,
1996; Agapow \& Burt, 2001). The low value of \(\bar{r}_d\) in the CA
population may represent sexual reproduction, but we can see in Figure
\ref{DAPC-RY} that there is differentiation by year. Thus, this could
also be an artifact of sampling two different populations, which is
known to reduce the value of \(\bar{r}_d\) (Prugnolle \& de Meeûs,
2010).

The previous study of the white mold screening nursery populations used
MCGs to assess genotypic diversity (Otto-Hanson et al., 2011).
Historically, MCGs have been used as a proxy for clonal lineages, and
thus, of interest in this study was testing the association between
multilocus haplotypes (MLHs) and mycelial compatibility groups (MCGs)
(Kohn et al., 1990; Leslie, 1993; Kohn, 1995; Carbone et al., 1999;
Schafer \& Kohn, 2006; Otto-Hanson et al., 2011). Our results, however,
do not support this assumption. It can be seen in Fig. \ref{biggraph}A
that the most abundant MCG contains several MLHs, but the diversity of
those MLHs are low as indicated by the evenness (transparency), which
indicates that there is one dominant MLH (`MLH 25'). What is not shown
in Fig. \ref{biggraph}A is the MLHs that are shared between MCGs. This
is illustrated in both Table \ref{tab:mlg-table} and Fig.
\ref{biggraph}B,C. It could be argued, however that `MLH 25', with its
high value of \(P_{sex}\) represents different true MLHs across the five
MCGs it occupies, but this does not account for the overall structure of
Fig. \ref{fullgraph} where, for example, `MLH 75' (\(P_{sex}\) =
1.81e\textsuperscript{-4}) is compatible with 57 other haplotypes
through three MCG when the population structure of \emph{S.
sclerotiorum} is known to be clonal.

Over the past few years, researchers have noticed inconsistencies among
the relationship between MCGs and MLHs (Carbone et al., 1999; Attanayake
et al., 2012; Aldrich-Wolfe et al., 2015; Lehner et al., 2015). Either
several MCGs belong to one MLH, which could be explained by insufficient
sampling of loci; several MLHs belong to one MCG, which could be
explained by clonal expansion; or a mixture of both. Some studies have
shown a correlation between MCG and MLH (Carbone et al., 1999;
Aldrich-Wolfe et al., 2015; Lehner et al., 2015), whereas other studies
have shown no apparent correlation, even on small spatial scales
(Atallah et al., 2004; Attanayake et al., 2012, 2013).

One long-held assumption was that MCGs (as determined via barrage
reaction) represent vegetative compatibility groups (VCGs) (Kohn et al.,
1990; Schafer \& Kohn, 2006; Lehner et al., 2015), which are known to
have a genetic component (Saupe, 2000; Hall et al., 2010; Strom \&
Bushley, 2016). While our protocol for assessing MCGs utilized a
different substrate for the MCG reactions, the patterns we observe are
not dissimilar from what have previously been reported in the
literature. It has been demonstrated in several Ascomycetes---including
\emph{Neurospora crassa} (Micali \& Smith, 2003), \emph{Sclerotinia
homoecarpa} (Jo et al., 2008), \emph{Verticillium dahliae} (Papaioannou
\& Typas, 2014), and \emph{S. sclerotiorum} (Ford et al., 1995)---that
barrage reactions are independent from stable anastomosis. Thus, the
inconsistencies in this study and other studies indicate that
researchers studying \emph{S. sclerotiorum} should not rely on MCG data
derived from barrage reactions as an indicator for genetic diversity.

\subsection*{Limitations}\label{limitations}
\addcontentsline{toc}{subsection}{Limitations}

One of the main limitations of this study is the focus on \emph{P.
vulgaris} as a host. It has been shown that \emph{S. sclerotiorum} in
the midwestern United States does not have a particular preference for
host (Aldrich-Wolfe et al., 2015). If the distribution of \emph{S.
sclerotiorum} is even across agricultural hosts in the USA, then our
sample may yet be representative of the genetic pool present in other
crops and weedy species.

Another limitation was the microsatellite markers used for this
particular study (Sirjusingh \& Kohn, 2001). The haplotype accumulation
curve showed no indication of a plateau, indicating that if we had
sampled more loci, we would have resolved more multilocus haplotypes.
While 16 loci showed us similar results and began to show a plateau for
the haplotype accumulation curve, we were unable to use these results
due to our uncertainty in the allele calls for these five extra loci.
With the availability of an optically-mapped genome (Derbyshire et al.,
2017), future studies describing the genetic diversity of \emph{S.
sclerotiorum} should employ techniques such as Genotyping-By-Sequencing
(Davey et al., 2011), Sequence Capture (Grover et al., 2012), or Whole
Genome Sequencing.

\subsection*{Conclusions}\label{conclusions}
\addcontentsline{toc}{subsection}{Conclusions}

This study represents the largest genetic analysis of \emph{S.
sclerotiorum} from the USA to date, giving us a unique insight to
continent-wide population structure and relationships between phenotypic
and genotypic variables. Populations in wmn appear to show no
significant differentiation when compared to their production field
counterparts, suggesting that the wmn populations of \emph{S.
sclerotiorum} may be considered representative of the surrounding
regions. While we found no direct relationship between haplotype and
severity, it is evident that there is a gradient of severity by region,
further supporting the need for screening in multiple locations. Based
on our analysis of the relationships between MCG and MLH, we found no
clear evidence that the two are directly related, suggesting that MCG do
not necessarily represent vegetative compatibility groups and thus
should not be used as a proxy for identifying clones.

\subsection*{Data Availability}\label{data-availability}
\addcontentsline{toc}{subsection}{Data Availability}

All scripts, data, and resources used to generate the results presented
in this publication (including Supplementary Information) are fully
reproducible and available at The Open Science Framework
\url{https://osf.io/ejb5y}.

\subsection*{Acknowledgements}\label{acknowledgements}
\addcontentsline{toc}{subsection}{Acknowledgements}

The authors would like to thank Rebecca Higgens for providing valuable
insights into the historical context of the data collection and
curation. Funding for this research was provided by the North Central
Soybean Research Program (\#639K623) to SEE, a Layman Award (\#2446) to
SEE, USDA-ARS National Sclerotinia Initiative (\#58-5442-2-209) to
JRS/SEE, and start-up funds from the University of Nebraska-Lincoln to
SEE.

\subsection*{Conflicts of Interest}\label{conflicts-of-interest}
\addcontentsline{toc}{subsection}{Conflicts of Interest}

The authors declare no conflict of interest.

\subsection*{Author Contributions}\label{author-contributions}
\addcontentsline{toc}{subsection}{Author Contributions}

\begin{itemize}
\item
  \textbf{Zhian N. Kamvar} analyzed the data, contributed analysis
  tools, wrote the paper, prepared figures and tables, reviewed drafts
  of the paper.
\item
  \textbf{Bimal Sajeewa Amadarasa} \ldots{}
\item
  \textbf{Rachana Jhala} \ldots{}
\item
  \textbf{Serena McCoy} \ldots{}
\item
  \textbf{James Steadman} \ldots{}
\item
  \textbf{Sydney E. Everhart} \ldots{}
\end{itemize}

\newpage

\section*{Supplementary Information}\label{supplementary-information}
\addcontentsline{toc}{section}{Supplementary Information}

\setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}}
\setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}

\begin{longtable}[]{@{}lllrlr@{}}
\caption{\label{tab:isolate-table} Description of \emph{Sclerotinia
sclerotiorum} isolates used in this study. N = Number of Isolates. Key
abbreviations: wmn = white mold screening nursery, producer = producer
field, unk = unknown cultivar.}\tabularnewline
\toprule
\begin{minipage}[b]{0.11\columnwidth}\raggedright\strut
Country\strut
\end{minipage} & \begin{minipage}[b]{0.08\columnwidth}\raggedright\strut
State\strut
\end{minipage} & \begin{minipage}[b]{0.12\columnwidth}\raggedright\strut
Field Code\strut
\end{minipage} & \begin{minipage}[b]{0.19\columnwidth}\raggedleft\strut
Year\strut
\end{minipage} & \begin{minipage}[b]{0.29\columnwidth}\raggedright\strut
Host\strut
\end{minipage} & \begin{minipage}[b]{0.04\columnwidth}\raggedleft\strut
N\strut
\end{minipage}\tabularnewline
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{0.11\columnwidth}\raggedright\strut
Country\strut
\end{minipage} & \begin{minipage}[b]{0.08\columnwidth}\raggedright\strut
State\strut
\end{minipage} & \begin{minipage}[b]{0.12\columnwidth}\raggedright\strut
Field Code\strut
\end{minipage} & \begin{minipage}[b]{0.19\columnwidth}\raggedleft\strut
Year\strut
\end{minipage} & \begin{minipage}[b]{0.29\columnwidth}\raggedright\strut
Host\strut
\end{minipage} & \begin{minipage}[b]{0.04\columnwidth}\raggedleft\strut
N\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
USA\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
CA\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
wmn\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
2004, 2005\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
Beryl, Bunsi, G122\strut
\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
18\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
USA\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
CO\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
producer\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
2007, 2010\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
Pinto, Yellow\strut
\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
41\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
wmn\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
2003\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
GH\strut
\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
1\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
USA\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
ID\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
producer\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
2003\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
GH\strut
\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
1\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
USA\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
MI\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
wmn\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
2003, 2004, 2005, 2008, 2009\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
11A, 37, 38, B07104, Beryl, Bunsi, Cornell, G122, Orion, PO7863,
WM31\strut
\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
43\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
producer\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
2003, 2008, 2009\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
BL, Black, Fuji, GH, Merlot, SR06233, unk, Vista, Zorro\strut
\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
19\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
USA\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
MN\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
wmn\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
2003, 2004\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
Beryl, Bunsi, G122\strut
\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
11\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
USA\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
ND\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
producer\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
2007, 2010\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
unk\strut
\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
53\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
wmn\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
2005\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
Beryl, Bunsi, G122\strut
\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
7\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
USA\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
NE\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
wmn\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
2004, 2005, 2008, 2010\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
Beryl, Bunsi, G122, PO7683, unk\strut
\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
27\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
producer\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
2003, 2007, 2009, 2010\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
Beryl, Emerson, GH, Orion, Pinto, Weihing\strut
\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
20\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
USA\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
NY\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
producer\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
2003\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
GH\strut
\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
1\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
USA\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
OR\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
wmn\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
2003, 2004\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
Beryl, Bunsi, G122\strut
\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
15\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
producer\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
2003\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
G122, GH\strut
\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
2\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
USA\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
WA\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
wmn\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
2003, 2004, 2005, 2008\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
11A, 37, 38, Beryl, Bunsi, Cornell, G122, Orion, PO7 104, PO7863,
WM31\strut
\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
36\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
producer\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
2003, 2007\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
GH, Merlot, Pinto, Redkid\strut
\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
23\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
USA\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
WI\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
producer\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
2003\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
GH\strut
\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
2\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
Mexico\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
wmn\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
2005\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
Beryl, Bunsi, G122\strut
\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
18\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
France\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
wmn\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
2004, 2005\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
Beryl, Bunsi, G122\strut
\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
18\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
producer\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
2012\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
unk\strut
\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
4\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
Australia\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
wmn\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
2004\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
Beryl, Bunsi, G122\strut
\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
4\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
\strut
\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
producer\strut
\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
2004\strut
\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
Beryl\strut
\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
2\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

\begin{longtable}[]{@{}lrl@{}}
\caption{\label{tab:region-aggressiveness} Mean aggressiveness ratings for
Regions with more than five samples; groupings according to 95\%
family-wise confidence interval.}\tabularnewline
\toprule
\begin{minipage}[b]{0.14\columnwidth}\raggedright\strut
Region\strut
\end{minipage} & \begin{minipage}[b]{0.27\columnwidth}\raggedleft\strut
Mean Aggressiveness\strut
\end{minipage} & \begin{minipage}[b]{0.08\columnwidth}\raggedright\strut
Group\strut
\end{minipage}\tabularnewline
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{0.14\columnwidth}\raggedright\strut
Region\strut
\end{minipage} & \begin{minipage}[b]{0.27\columnwidth}\raggedleft\strut
Mean Aggressiveness\strut
\end{minipage} & \begin{minipage}[b]{0.08\columnwidth}\raggedright\strut
Group\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
MN\strut
\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
5.84\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
a\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
ND\strut
\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
5.77\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
a\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
NE\strut
\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
5.29\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
ab\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
MI\strut
\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
5.13\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
abc\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
OR\strut
\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
4.84\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
abcd\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
CO\strut
\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
4.72\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
bcd\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
WA\strut
\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
4.67\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
cd\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
France\strut
\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
4.66\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
cd\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Mexico\strut
\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
4.58\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
cd\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
Australia\strut
\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
4.12\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
cd\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
CA\strut
\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
4.01\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
d\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

\begin{longtable}[]{@{}lrl@{}}
\caption{\label{tab:MCG-aggressiveness} Mean aggressiveness ratings for the
10 most abundant MCG; groupings according to 95\% family-wise confidence
interval.}\tabularnewline
\toprule
\begin{minipage}[b]{0.14\columnwidth}\raggedright\strut
MCG\strut
\end{minipage} & \begin{minipage}[b]{0.27\columnwidth}\raggedleft\strut
Mean Aggressiveness\strut
\end{minipage} & \begin{minipage}[b]{0.08\columnwidth}\raggedright\strut
Group\strut
\end{minipage}\tabularnewline
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{0.14\columnwidth}\raggedright\strut
MCG\strut
\end{minipage} & \begin{minipage}[b]{0.27\columnwidth}\raggedleft\strut
Mean Aggressiveness\strut
\end{minipage} & \begin{minipage}[b]{0.08\columnwidth}\raggedright\strut
Group\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
44\strut
\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
6.03\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
a\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
3\strut
\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
5.50\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
ab\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
5\strut
\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
5.40\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
b\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
2\strut
\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
5.25\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
b\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
9\strut
\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
5.11\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
b\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
1\strut
\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
4.95\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
b\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
45\strut
\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
4.88\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
b\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
4\strut
\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
4.87\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
b\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
53\strut
\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
4.69\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
b\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
49\strut
\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
4.60\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
b\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

\begin{longtable}[]{@{}lrl@{}}
\caption{\label{tab:MLG-aggressiveness} Mean aggressiveness ratings for the
10 MLH most abundant; groupings according to 95\% family-wise confidence
interval.}\tabularnewline
\toprule
\begin{minipage}[b]{0.14\columnwidth}\raggedright\strut
MLH\strut
\end{minipage} & \begin{minipage}[b]{0.27\columnwidth}\raggedleft\strut
Mean Aggressiveness\strut
\end{minipage} & \begin{minipage}[b]{0.08\columnwidth}\raggedright\strut
Group\strut
\end{minipage}\tabularnewline
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{0.14\columnwidth}\raggedright\strut
MLH\strut
\end{minipage} & \begin{minipage}[b]{0.27\columnwidth}\raggedleft\strut
Mean Aggressiveness\strut
\end{minipage} & \begin{minipage}[b]{0.08\columnwidth}\raggedright\strut
Group\strut
\end{minipage}\tabularnewline
\midrule
\endhead
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
78\strut
\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
6.07\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
a\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
65\strut
\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
5.94\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
a\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
9\strut
\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
5.67\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
ab\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
25\strut
\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
5.41\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
ab\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
66\strut
\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
5.30\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
ab\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
104\strut
\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
5.22\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
ab\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
160\strut
\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
4.80\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
ab\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
163\strut
\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
4.80\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
ab\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
165\strut
\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
4.34\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
b\strut
\end{minipage}\tabularnewline
\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
140\strut
\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
4.31\strut
\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
b\strut
\end{minipage}\tabularnewline
\bottomrule
\end{longtable}

\begin{figure}
\centering
\includegraphics[width=1.00000\textwidth]{../../results/figures/publication/full-graph.pdf}
\caption{Graph showing complex associations between Mycelial
Compatibility Groups (MCG) (dotted nodes) and Multilocus Haplotypes
(MLH) (full nodes) where the number in each node represents the MLH/MCG
assignment. Node size reflect the number of samples represented by each
node (circle). Edges (arrows) point from MLH to MCG where the weight
(thickness) of the edge represents the number of samples shared. Node
color represents the community assignment based on the walktrap
algorithm with a maximum of four steps (Pons \& Latapy,
2006).}\label{fullgraph}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=1.00000\textwidth]{../../results/figures/publication/mlg-16.pdf}
\caption{Network of populations (nodes/circles) and their shared
multilocus haplotypes (MLH) (edges/lines) haplotyped over 16 loci. Each
node is labeled with \textbf{name (number of MLHs shared/number of MLHs
total).} The shade and area of the nodes are proportional to the number
of unique MLHs within the node and the inner nodes are proportional to
the number of private MLHs to the region (bottom legend). Each edge
represents a single MLH where its thickness represents the number of
populations that share the MLH and the shade represents the value of
\(P_{sex}\), or the probability of encountering that MLH from two
independent meiotic events.}\label{community-graph16}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=1.00000\textwidth]{../../results/figures/publication/dapc_region_year.pdf}
\caption{Scatter plot of Discriminant Analysis of Principal Components
on Regions and Years showing temporal variation across all Regions.
Points (text labels) represent observed individuals connected to the
population centroids with ellipses representing a 66\% confidence
interval for a normal distribution. The center of each component is
represented as black grid lines.}\label{DAPC-RY-FULL}
\end{figure}

\newpage

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\hypertarget{ref-agapow2001indices}{}
Agapow, P., \& Burt, A. (2001). Indices of multilocus linkage
disequilibrium. \emph{Molecular Ecology Notes}, \emph{1}, 101--102.

\hypertarget{ref-aldrich-wolfe2015genetic}{}
Aldrich-Wolfe, L., Travers, S., \& Nelson, B. D. (2015). Genetic
variation of \emph{Sclerotinia sclerotiorum} from multiple crops in the
north central United States. \emph{PLOS ONE}, \emph{10}(9), e0139188.
\url{https://doi.org/10.1371/journal.pone.0139188}

\hypertarget{ref-arnaud2007standardizing}{}
Arnaud-Hanod, S., Duarte, C. M., Alberto, F., \& Serrão, E. A. (2007).
Standardizing methods to address clonality in population studies.
\emph{Molecular Ecology}, \emph{16}(24), 5115--5139.
\url{https://doi.org/10.1111/j.1365-294X.2007.03535.x}

\hypertarget{ref-atallah2004high}{}
Atallah, Z. K., Larget, B., Chen, X., \& Johnson, D. A. (2004). High
genetic diversity, phenotypic uniformity, and evidence of outcrossing in
\emph{Sclerotinia sclerotiorum} in the Columbia Basin of Washington
State. \emph{Phytopathology}, \emph{94}, 737--742.

\hypertarget{ref-attanayake2013sclerotinia}{}
Attanayake, R. N., Carter, P. A., Jiang, D., del Río-Mendoza, L., \&
Chen, W. (2013). \emph{Sclerotinia sclerotiorum} populations infecting
canola from China and the United States are genetically and
phenotypically distinct. \emph{Phytopathology}, \emph{103}(7), 750--761.
\url{https://doi.org/10.1094/phyto-07-12-0159-r}

\hypertarget{ref-attanayake2012genetic}{}
Attanayake, R., Porter, L., Johnson, D., \& Chen, W. (2012). Genetic and
phenotypic diversity and random association of DNA markers of isolates
of the fungal plant pathogen \emph{Sclerotinia sclerotiorum} from soil
on a fine geographic scale. \emph{Soil Biology and Biochemistry},
\emph{55}, 28--36. \url{https://doi.org/10.1016/j.soilbio.2012.06.002}

\hypertarget{ref-boland1994index}{}
Boland, G., \& Hall, R. (1994). Index of plant hosts of
\emph{Sclerotinia sclerotiorum}. \emph{Canadian Journal of Plant
Pathology}, \emph{16}(2), 93--108.
\url{https://doi.org/10.1080/07060669409500766}

\hypertarget{ref-bolton2006sclerotinia}{}
Bolton, M. D., Thomma, B. P. H. J., \& Nelson, B. D. (2006).
\emph{Sclerotinia sclerotiorum} (Lib.) de Bary: Biology and molecular
traits of a cosmopolitan pathogen. \emph{Molecular Plant Pathology},
\emph{7}(1), 1--16.
\url{https://doi.org/10.1111/j.1364-3703.2005.00316.x}

\hypertarget{ref-botelho2013performance}{}
Botelho, L. d., Zancan, W. L. A., Cruz Machado, J. da, \& Barrocas, E.
N. (2013). Performance of common bean seeds infected by the fungus
\emph{Sclerotinia sclerotiorum}. \emph{Journal of Seed Science},
\emph{35}(2), 153--160.
\url{https://doi.org/10.1590/s2317-15372013000200003}

\hypertarget{ref-brown1980multilocus}{}
Brown, A. H. D., Feldman, M. W., \& Nevo, E. (1980). Multilocus
structure of natural populations of \emph{Hordeum Spontaneum}.
\emph{Genetics}, \emph{96}(2), 523--536. Retrieved from
\url{http://www.genetics.org/content/96/2/523}

\hypertarget{ref-bruvo2004simple}{}
Bruvo, R., Michiels, N. K., D'Souza, T. G., \& Schulenburg, H. (2004). A
simple method for the calculation of microsatellite genotype distances
irrespective of ploidy level. \emph{Molecular Ecology}, \emph{13}(7),
2101--2106.

\hypertarget{ref-carbone2001multilocus}{}
Carbone, I., \& Kohn, L. M. (2001). Multilocus nested haplotype networks
extended with DNA fingerprints show common origin and fine-scale,
ongoing genetic divergence in a wild microbial metapopulation.
\emph{Molecular Ecology}, \emph{10}(10), 2409--2422.
\url{https://doi.org/10.1046/j.0962-1083.2001.01380.x}

\hypertarget{ref-carbone1999patterns}{}
Carbone, I., Anderson, J. B., \& Kohn, L. M. (1999). Patterns of descent
in clonal lineages and their multilocus fingerprints are resolved with
combined gene genealogies. \emph{Evolution}, \emph{53}(1), 11--21.
\url{https://doi.org/10.1111/j.1558-5646.1999.tb05329.x}

\hypertarget{ref-csardi2006igraph}{}
Csardi, G., \& Nepusz, T. (2006). The igraph software package for
complex network research. \emph{InterJournal}, \emph{Complex Systems},
1695. Retrieved from \url{http://igraph.org}

\hypertarget{ref-cubeta1997clonality}{}
Cubeta, M. A., Cody, B. R., Kohli, Y., \& Kohn, L. M. (1997). Clonality
in \emph{Sclerotinia sclerotiorum} on infected cabbage in eastern North
Carolina. \emph{Phytopathology}, \emph{87}, 1000--1004.

\hypertarget{ref-cubeta2001mycelial}{}
Cubeta, M., Sermons, D., \& Cody, B. (2001). Mycelial interactions of
\emph{Sclerotinia minor}. \emph{Phytopathology}, \emph{91}(6S), S19.
\url{https://doi.org/10.1094/phyto.2001.91.6.s1}

\hypertarget{ref-davey2011genome}{}
Davey, J. W., Hohenlohe, P. A., Etter, P. D., Boone, J. Q., Catchen, J.
M., \& Blaxter, M. L. (2011). Genome-wide genetic marker discovery and
genotyping using next-generation sequencing. \emph{Nature Reviews
Genetics}, \emph{12}(7), 499--510. \url{https://doi.org/10.1038/nrg3012}

\hypertarget{ref-derbyshire2017complete}{}
Derbyshire, M., Denton-Giles, M., Hegedus, D., Seifbarghy, S., Rollins,
J., van Kan, J., Seidl, M. F., Faino, L., Mbengue, M., Navaud, O.,
Raffaele, S., Hammond-Kosack, K., Heard, S., \& Oliver, R. (2017). The
complete genome sequence of the phytopathogenic fungus \emph{Sclerotinia
sclerotiorum} reveals insights into the genome architecture of broad
host range pathogens. \emph{Genome Biology and Evolution}, \emph{9}(3),
593--618. \url{https://doi.org/10.1093/gbe/evx030}

\hypertarget{ref-ekins2011population}{}
Ekins, M. G., Hayden, H. L., Aitken, E. A. B., \& Goulter, K. C. (2011).
Population structure of \emph{Sclerotinia sclerotiorum} on sunflower in
Australia. \emph{Australasian Plant Pathology}, \emph{40}, 99--108.

\hypertarget{ref-excoffier1992analysis}{}
Excoffier, L., Smouse, P. E., \& Quattro, J. M. (1992). Analysis of
molecular variance inferred from metric distances among DNA haplotypes:
Application to human mitochondrial DNA restriction data.
\emph{Genetics}, \emph{131}(2), 479--91.

\hypertarget{ref-ford1995heterokaryon}{}
Ford, E., Miller, R., Gray, H., \& Sherwood, J. (1995). Heterokaryon
formation and vegetative compatibility in \emph{Sclerotinia
sclerotiorum}. \emph{Mycological Research}, \emph{99}(2), 241--247.

\hypertarget{ref-grover2012targeted}{}
Grover, C. E., Salmon, A., \& Wendel, J. F. (2012). Targeted sequence
capture as a powerful tool for evolutionary analysis. \emph{American
Journal of Botany}, \emph{99}(2), 312--319.
\url{https://doi.org/10.3732/ajb.1100323}

\hypertarget{ref-grunwald2003analysis}{}
Grünwald, N. J., Goodwin, S. B., Milgroom, M. G., \& Fry, W. E. (2003).
Analysis of genotypic diversity data for populations of microorganisms.
\emph{Phytopathology}, \emph{93}(6), 738--746.
\url{https://doi.org/10.1094/phyto.2003.93.6.738}

\hypertarget{ref-hall2010evolution}{}
Hall, C., Welch, J., Kowbel, D. J., \& Glass, N. L. (2010). Evolution
and diversity of a fungal self/nonself recognition locus. \emph{PLoS
ONE}, \emph{5}(11), e14055.
\url{https://doi.org/10.1371/journal.pone.0014055}

\hypertarget{ref-heck1975explicit}{}
Heck, K. L., van Belle, G., \& Simberloff, D. (1975). Explicit
calculation of the rarefaction diversity measurement and the
determination of sufficient sample size. \emph{Ecology}, \emph{56}(6),
1459--1461. \url{https://doi.org/10.2307/1934716}

\hypertarget{ref-purrr}{}
Henry, L., \& Wickham, H. (2017). \emph{purrr: Functional programming
tools}. Retrieved from \url{https://CRAN.R-project.org/package=purrr}

\hypertarget{ref-hurlbert1971nonconcept}{}
Hurlbert, S. H. (1971). The nonconcept of species diversity: A critique
and alternative parameters. \emph{Ecology}, \emph{52}(4), 577--586.
\url{https://doi.org/10.2307/1934145}

\hypertarget{ref-jo2008reassessment}{}
Jo, Y.-K., Chang, S. W., Rees, J., \& Jung, G. (2008). Reassessment of
vegetative compatibility of \emph{Sclerotinia homoeocarpa} using
nitrate-nonutilizing mutants. \emph{Phytopathology}, \emph{98}(1),
108--114. \url{https://doi.org/10.1094/phyto-98-1-0108}

\hypertarget{ref-jombart2008adegenet}{}
Jombart, T. (2008). adegenet: A R package for the multivariate analysis
of genetic markers. \emph{Bioinformatics}, \emph{24}(11), 1403--1405.
\url{https://doi.org/10.1093/bioinformatics/btn129}

\hypertarget{ref-jombart2010discriminant}{}
Jombart, T., Devillard, S., \& Balloux, F. (2010). Discriminant analysis
of principal components: A new method for the analysis of genetically
structured populations. \emph{BMC Genetics}, \emph{11:94}.
\url{https://doi.org/10.1186/1471--2156--11--94}

\hypertarget{ref-kamvar2015novel}{}
Kamvar, Z. N., Brooks, J. C., \& Grünwald, N. J. (2015). Novel R tools
for analysis of genome-wide population genetic data with emphasis on
clonality. \emph{Frontiers in Genetics}, \emph{6}.
\url{https://doi.org/10.3389/fgene.2015.00208}

\hypertarget{ref-kamvar2014poppr}{}
Kamvar, Z. N., Tabima, J. F., \& Grünwald, N. J. (2014). Poppr: An R
package for genetic analysis of populations with clonal, partially
clonal, and/or sexual reproduction. \emph{PeerJ}, \emph{2}, e281.
\url{https://doi.org/10.7717/peerj.281}

\hypertarget{ref-kohli1998random}{}
Kohli, Y., \& Kohn, L. M. (1998). Random association among alleles in
clonal populations of \emph{Sclerotinia sclerotiorum}. \emph{Fungal
Genetics and Biology}, \emph{23}, 139--149.

\hypertarget{ref-kohli1995clonal}{}
Kohli, Y., Brunner, L. J., Yoell, H., Milgroom, M. G., Anderson, J. B.,
Morrall, R. A. A., \& Kohn, L. M. (1995). Clonal dispersal and spatial
mixing in populations of the plant pathogenic fungus, \emph{Sclerotinia
sclerotiorum}. \emph{Molecular Ecology}, \emph{4}, 69--77.

\hypertarget{ref-kohn1995clonal}{}
Kohn, L. M. (1995). The clonal dynamic in wild and agricultural
plant-pathogen populations. \emph{Canadian Journal of Botany},
\emph{73}(S1), 1231--1240. \url{https://doi.org/10.1139/b95-383}

\hypertarget{ref-kohn1990mycelial}{}
Kohn, L. M., Carbone, I., \& Anderson, J. B. (1990). Mycelial
interactions in \emph{Sclerotinia sclerotiorum}. \emph{Experimental
Mycology}, \emph{14}, 255--267.

\hypertarget{ref-legendre1999distance}{}
Legendre, P., \& Anderson, M. J. (1999). Distance-based redundancy
analysis: Testing multispecies responses in multifactorial ecological
experiments. \emph{Ecological Monographs}, \emph{69}, 1--24.

\hypertarget{ref-lehner2017sclerotinia}{}
Lehner, M. S., \& Mizubuti, E. S. G. (2017). Are \emph{Sclerotinia
sclerotiorum} populations from the tropics more variable than those from
subtropical and temperate zones? \emph{Tropical Plant Pathology},
\emph{42}(2), 61--69. \url{https://doi.org/10.1007/s40858-016-0125-1}

\hypertarget{ref-lehner2015genetic}{}
Lehner, M. S., Júnior, T. J. P., Júnior, B. T. H., Teixeira, H., Vieira,
R. F., Carneiro, J. E. S., \& Mizubuti, E. S. G. (2015). Low genetic
variability in \emph{Sclerotinia sclerotiorum} populations from common
bean fields in Minas Gerais State, Brazil, at regional, local and
micro-scales. \emph{Plant Pathology}, \emph{64}(4), 921--931.
\url{https://doi.org/10.1111/ppa.12322}

\hypertarget{ref-lehner2017independently}{}
Lehner, M. S., Paula Júnior, T. J. de, Del Ponte, E. M., Mizubuti, E.
S., \& Pethybridge, S. J. (2017). Independently founded populations of
\emph{Sclerotinia sclerotiorum} from a tropical and a temperate region
have similar genetic structure. \emph{PloS One}, \emph{12}(3), e0173915.
\url{https://doi.org/10.1371/journal.pone.0173915}

\hypertarget{ref-leslie1993fungal}{}
Leslie, J. (1993). Fungal vegetative compatibility. \emph{Annual Review
of Phytopathology}, \emph{31}, 127--150. Review.
\url{https://doi.org/10.1146/annurev.py.31.090193.001015}

\hypertarget{ref-mccoy2009use}{}
McCoy, S., \& Steadman, J. R. (2009). Use of multi-site screening to
identify partial resistance to white mold in common bean in 2008.
\emph{Bean Improvement Cooperative Annual Report}, 86--87. Retrieved
from \url{https://naldc.nal.usda.gov/naldc/catalog.xhtml?id=IND44207142}

\hypertarget{ref-mcdonald2002pathogen}{}
McDonald, B. A., \& Linde, C. (2002). Pathogen population genetics,
evolutionary potential, and durable resistance. \emph{Annual Review of
Phytopathology}, \emph{40}(1), 349--379.
\url{https://doi.org/10.1146/annurev.phyto.40.120501.101443}

\hypertarget{ref-mendiburu2015agricolae}{}
Mendiburu, F. D., \& Simon, R. (2015). Agricolae - ten years of an open
source statistical tool for experiments in breeding, agriculture and
biology. \url{https://doi.org/10.7287/peerj.preprints.1404v1}

\hypertarget{ref-micali2003independence}{}
Micali, C. O., \& Smith, M. L. (2003). On the independence of barrage
formation and heterokaryon incompatibility in \emph{Neurospora crassa}.
\emph{Fungal Genetics and Biology}, \emph{38}(2), 209--219.
\url{https://doi.org/10.1016/s1087-1845(02)00533-9}

\hypertarget{ref-milgroom1996recombination}{}
Milgroom, M. G. (1996). Recombination and the multilocus structure of
fungal populations. \emph{Annual Review of Phytopathology},
\emph{34}(1), 457--477.

\hypertarget{ref-nei1978estimation}{}
Nei, M. (1978). Estimation of average heterozygosity and genetic
distance from a small number of individuals. \emph{Genetics}, \emph{89},
583--590.

\hypertarget{ref-vegan}{}
Oksanen, J., Blanchet, F. G., Friendly, M., Kindt, R., Legendre, P.,
McGlinn, D., Minchin, P. R., O'Hara, R. B., Simpson, G. L., Solymos, P.,
Stevens, M. H. H., Szoecs, E., \& Wagner, H. (2017). \emph{Vegan:
Community ecology package}. Retrieved from
\url{https://CRAN.R-project.org/package=vegan}

\hypertarget{ref-otto2007identification}{}
Otto-Hanson, L., \& Steadman, J. R. (2007). Identification of partial
resistance to \emph{Sclerotinia sclerotiorum} in common bean at multiple
locations in 2006. \emph{Bean Improvement Cooperative Annual Report},
133--134. Retrieved from
\url{https://naldc.nal.usda.gov/naldc/catalog.xhtml?id=IND43940892}

\hypertarget{ref-otto2008identification}{}
Otto-Hanson, L., \& Steadman, J. R. (2008). Identification of partial
resistance to \emph{Sclerotinia sclerotiorum} in common bean at multiple
locations in 2007. \emph{Bean Improvement Cooperative Annual Report},
214--215. Retrieved from
\url{https://naldc.nal.usda.gov/naldc/catalog.xhtml?id=IND44063230}

\hypertarget{ref-otto-hanson2011variation}{}
Otto-Hanson, L., Steadman, J. R., Higgins, R., \& Eskridge, K. M.
(2011). Variation in \emph{Sclerotinia sclerotiorum} bean isolates from
multisite resistance screening locations. \emph{Plant Disease},
\emph{95}(11), 1370--1377. \url{https://doi.org/10.1094/pdis-11-10-0865}

\hypertarget{ref-papaioannou2014barrage}{}
Papaioannou, I. A., \& Typas, M. A. (2014). Barrage formation is
independent from heterokaryon incompatibility in \emph{Verticillium
dahliae}. \emph{European Journal of Plant Pathology}, \emph{141}(1),
71--82. \url{https://doi.org/10.1007/s10658-014-0525-3}

\hypertarget{ref-parks1993study}{}
Parks, J. C., \& Werth, C. R. (1993). A study of spatial features of
clones in a population of bracken fern, \emph{Pteridium aquilinum}
(Dennstaedtiaceae). \emph{American Journal of Botany}, \emph{80}(5),
537. \url{https://doi.org/10.2307/2445369}

\hypertarget{ref-ggraph}{}
Pedersen, T. L. (2017). \emph{ggraph: An implementation of grammar of
graphics for graphs and networks}. Retrieved from
\url{https://CRAN.R-project.org/package=ggraph}

\hypertarget{ref-petzoldt1996straw}{}
Petzoldt, R., \& Dickson, M. H. (1996). Straw test for resistance to
white mold in beans. \emph{Bean Improvement Cooperative Annual Report},
\emph{39}, 142--143. Retrieved from
\url{https://naldc.nal.usda.gov/naldc/catalog.xhtml?id=IND20562675}

\hypertarget{ref-pielou1975ecological}{}
Pielou, E. (1975). \emph{Ecological Diversity}. New York: Wiley \& Sons.

\hypertarget{ref-pons2006computing}{}
Pons, P., \& Latapy, M. (2006). Computing communities in large networks
using random walks. \emph{Journal of Graph Algorithms and Applications},
\emph{10}(2), 191--218. \url{https://doi.org/10.7155/jgaa.00124}

\hypertarget{ref-prugnolle2010apparent}{}
Prugnolle, F., \& de Meeûs, T. (2010). Apparent high recombination rates
in clonal parasitic organisms due to inappropriate sampling design.
\emph{Heredity}, \emph{104}(2), 135--140.
\url{https://doi.org/10.1038/hdy.2009.128}

\hypertarget{ref-R}{}
R Core Team. (2017). \emph{R: A language and environment for statistical
computing}. Vienna, Austria: R Foundation for Statistical Computing.
Retrieved from \url{https://www.R-project.org/}

\hypertarget{ref-sambrook1989molecular}{}
Sambrook, J., Fritsch, E. F., Maniatis, T., \& others. (1989).
\emph{Molecular Cloning: A Laboratory Manual.} Cold spring harbor
laboratory press.

\hypertarget{ref-saupe2000molecular}{}
Saupe, S. J. (2000). Molecular genetics of heterokaryon incompatibility
in filamentous ascomycetes. \emph{Microbiology and Molecular Biology
Reviews}, \emph{64}(3), 489--502.
\url{https://doi.org/10.1128/mmbr.64.3.489-502.2000}

\hypertarget{ref-schafer2006optimized}{}
Schafer, M. R., \& Kohn, L. M. (2006). An optimized method for mycelial
compatibility testing in \emph{Sclerotinia sclerotiorum}.
\emph{Mycologia}, \emph{98}(4), 593--597.
\url{https://doi.org/10.1080/15572536.2006.11832662}

\hypertarget{ref-sexton2004microsatellite}{}
Sexton, A. C., \& Howlett, B. J. (2004). Microsatellite markers reveal
genetic differentiation among populations of \emph{Sclerotinia
sclerotiorum} from Australian canola fields. \emph{Current Genetics},
\emph{46}(6), 357--365. \url{https://doi.org/10.1007/s00294-004-0543-3}

\hypertarget{ref-sexton2006population}{}
Sexton, A. C., Whitten, A. R., \& Howlett, B. J. (2006). Population
structure of \emph{Sclerotinia sclerotiorum} in an Australian canola
field at flowering and stem-infection stages of the disease cycle.
\emph{Genome}, \emph{49}(11), 1408--1415.
\url{https://doi.org/10.1139/g06-101}

\hypertarget{ref-shannon2001mathematical}{}
Shannon, C. E. (1948). A mathematical theory of communication. \emph{ACM
SIGMOBILE Mobile Computing and Communications Review}, \emph{5}(1),
3--55.

\hypertarget{ref-simpson1949measurement}{}
Simpson, E. H. (1949). Measurement of diversity. \emph{Nature},
\emph{163}(4148), 688--688. \url{https://doi.org/10.1038/163688a0}

\hypertarget{ref-sirjusingh2001characterisation}{}
Sirjusingh, C., \& Kohn, L. M. (2001). Characterisation of
microsatellites in the fungal plant pathogen, \emph{Sclerotinia
sclerotiorum}. \emph{Molecular Ecology Notes}, \emph{1}(4), 267--269.
\url{https://doi.org/10.1046/j.1471-8278.2001.00102.x}

\hypertarget{ref-smith1993how}{}
Smith, J. M., Smith, N. H., O'Rourke, M., \& Spratt, B. G. (1993). How
clonal are bacteria? \emph{Proceedings of the National Academy of
Sciences}, \emph{90}(10), 4384--4388.
\url{https://doi.org/10.1073/pnas.90.10.4384}

\hypertarget{ref-steadman1983white}{}
Steadman, J. R. (1983). White mold - a serious yield-limiting disease of
bean. \emph{Plant Disease}, \emph{67}, 346--350.

\hypertarget{ref-steadman2003identification}{}
Steadman, J. R., Eskridge, K., \& Powers, K. (2003). Identification of
partial resistance to \emph{Sclerotinia sclerotiorum} in common bean at
multiple locations. \emph{Bean Improvement Cooperative Annual Report},
225--226. Retrieved from
\url{https://naldc.nal.usda.gov/naldc/catalog.xhtml?id=IND43757287}

\hypertarget{ref-steadman2006identification}{}
Steadman, J. R., Otto-Hanson, L., \& Breathnach, J. (2006).
Identification of partial resistance to \emph{Sclerotinia sclerotiorum}
in common bean at multiple locations in 2005. \emph{Bean Improvement
Cooperative Annual Report}, 223--224. Retrieved from
\url{https://naldc.nal.usda.gov/naldc/catalog.xhtml?id=IND43805570}

\hypertarget{ref-steadman2004identification}{}
Steadman, J. R., Otto-Hanson, L., \& Powers, K. (2004). Identification
of partial resistance to \emph{Sclerotinia sclerotiorum} in common bean
at multiple locations. \emph{Bean Improvement Cooperative Annual
Report}, \emph{47}, 281--282. Retrieved from
\url{https://naldc.nal.usda.gov/naldc/catalog.xhtml?id=IND43758354}

\hypertarget{ref-steadman2005identification}{}
Steadman, J. R., Otto-Hanson, L., \& Powers, K. (2005). Identification
of partial resistance to \emph{Sclerotinia sclerotiorum} in common bean
at multiple locations in 2004. \emph{Bean Improvement Cooperative Annual
Report}, 124--125. Retrieved from
\url{https://naldc.nal.usda.gov/naldc/catalog.xhtml?id=IND43759243}

\hypertarget{ref-stoddart1988genotypic}{}
Stoddart, J. A., \& Taylor, J. F. (1988). Genotypic diversity:
Estimation and prediction in samples. \emph{Genetics}, \emph{118}(4),
705--11.

\hypertarget{ref-strausbaugh2003management}{}
Strausbaugh, C., \& Forster, R. (2003). Management of white mold of
beans. \emph{Pacific Northwest Extension}. Retrieved from
\url{http://http://www.extension.uidaho.edu/publishing/pdf/PNW/PNW0568.pdf}

\hypertarget{ref-strom2016genomes}{}
Strom, N. B., \& Bushley, K. E. (2016). Two genomes are better than one:
History, genetics, and biotechnological applications of fungal
heterokaryons. \emph{Fungal Biology and Biotechnology}, \emph{3}(1).
\url{https://doi.org/10.1186/s40694-016-0022-x}

\hypertarget{ref-teran2006modified}{}
Teran, H., Lema, M., Schwartz, H. F., Duncan, R., Gilbeitson, R., \&
Singh, S. P. (2006). Modified Petzoldt and Dickson scale for white mold
rating of common bean. \emph{Bean Improvement Cooperative Annual
Report}, \emph{49}, 115--116. Retrieved from
\url{https://naldc.nal.usda.gov/naldc/catalog.xhtml?id=IND43805401}

\hypertarget{ref-ggplot2}{}
Wickham, H. (2009). \emph{ggplot2: Elegant graphics for data analysis}.
Springer-Verlag New York. Retrieved from \url{http://ggplot2.org}

\hypertarget{ref-dplyr}{}
Wickham, H., Francois, R., Henry, L., \& Müller, K. (2017). \emph{dplyr:
A grammar of data manipulation}. Retrieved from
\url{https://CRAN.R-project.org/package=dplyr}

\end{document}

### 2017-09-19 doc/RMD/data-comparison.Rmd
update data-comparison to use data within project

@@
 ```{r}
-excel_sheets("../Analysis4 ForManu/A1_Copy of binned-genotypes_SE.xlsx")
+excel_sheets(file.path(PROJHOME, "data/raw/2017-02-16_copy-of-binned-genotypes_SE.xlsx"))
 ```
@@
 ```{r sydney_data}
-syd <- read_excel("../Analysis4 ForManu/A1_Copy of binned-genotypes_SE.xlsx", 
+syd <- read_excel(file.path(PROJHOME, "data/raw/2017-02-16_copy-of-binned-genotypes_SE.xlsx"),
                 sheet = "GenAlexBinned", skip = 1) %>%
@@
 )
-saj <- read_csv("../Analysis4 ForManu/A2_Copy4 EUR_AUS_forManu.csv", skip = 2, col_types = column_specification) %>%
+saj <- read_csv(file.path(PROJHOME, "data/raw/2017-02-16_binned-genotypes-genalex_SA.csv"), skip = 2, col_types = column_specification) %>%
   select(-1) %>%
@@
 ```{r read_excel}
-ex <- readxl::read_excel("../Analysis4 ForManu/A1_Copy of binned-genotypes_SE.xlsx", sheet = "GenAlex", skip = 1) %>%
+ex <- readxl::read_excel(file.path(PROJHOME, "data/raw/2017-02-16_copy-of-binned-genotypes_SE.xlsx"), sheet = "GenAlex", skip = 1) %>%
   select(-1) %>%                # removing first column, which is empty

### 2017-09-19 doc/manuscript/manuscript.Rmd
update manuscript to add funding and dapc
THe DAPC update is due to the updated xvalDapc()
function, but it only increases the variance
explained by 1%

@@
 
-dapc_region      <- dapc(setPop(dat11cc, ~Region), n.da = 13, n.pca = 20)
+dapc_region      <- dapc(setPop(dat11cc, ~Region), n.da = 13, n.pca = 21)
 dapc_region_year <- dapc(setPop(dat11cc, ~Region/Year), n.da = 15, n.pca = 15)
@@
 
+## Funding {-}
+
+Funding for this research was provided by the North Central Soybean Research Program (#639K623) to SEE, a Layman Award (#2446) to SEE, USDA-ARS National Sclerotinia Initiative (#58-5442-2-209) to JRS/SEE, and start-up funds from the University of Nebraska-Lincoln to SEE.
+The funders had no role in study design, data collection and analysis, decision to publish, or preparation of the manuscript.
+
 ## Acknowledgements {-}
@@
 The authors would like to thank Rebecca Higgens for providing valuable insights into the historical context of the data collection and curation.
-Funding for this research was provided by the North Central Soybean Research Program (#639K623) to SEE, a Layman Award (#2446) to SEE, USDA-ARS National Sclerotinia Initiative (#58-5442-2-209) to JRS/SEE, and start-up funds from the University of Nebraska-Lincoln to SEE.
 

### 2017-09-19 doc/manuscript/manuscript.tex
run with docker container

@@
 
-\begin{figure}
+\begin{figure}[htbp]
 \centering
@@
 \end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
-0.032\strut
+0.029\strut
 \end{minipage}\tabularnewline
@@
 \end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
-0.498\strut
+0.486\strut
 \end{minipage}\tabularnewline
@@
 Discriminant Analysis of Principal Components (DAPC) was performed
-grouping by Region with the first 20 principal components, representing
-87.1\% of the total variance. The first discriminant axis (representing
-35\% of the discriminatory power) separated the centroid for the Mexico
-isolates from the rest of the data, indicating strong differentiation
-(Fig. \ref{DAPC}b). The second discriminant axis, representing 24.1\% of
-the discriminatory power, separated the centroid for the CA isolates.
-The mean population assignment probabilities for all populations with n
-\textgreater{} 10 showed that only isolates from Mexico, CA, and France
-had \textgreater{} 50\% probabilities of being reassigned to their
-source populations (Fig. \ref{DAPC}a).
+grouping by Region with the first 21 principal components, representing
+88.1\% of the total variance. The first discriminant axis (representing
+36.1\% of the discriminatory power) separated the centroid for the
+Mexico isolates from the rest of the data, indicating strong
+differentiation (Fig. \ref{DAPC}b). The second discriminant axis,
+representing 25.3\% of the discriminatory power, separated the centroid
+for the CA isolates. The mean population assignment probabilities for
+all populations with n \textgreater{} 10 showed that only isolates from
+Mexico, CA, and France had \textgreater{} 50\% probabilities of being
+reassigned to their source populations (Fig. \ref{DAPC}a).
 
@@
 
+\subsection*{Funding}\label{funding}
+\addcontentsline{toc}{subsection}{Funding}
+
+Funding for this research was provided by the North Central Soybean
+Research Program (\#639K623) to SEE, a Layman Award (\#2446) to SEE,
+USDA-ARS National Sclerotinia Initiative (\#58-5442-2-209) to JRS/SEE,
+and start-up funds from the University of Nebraska-Lincoln to SEE. The
+funders had no role in study design, data collection and analysis,
+decision to publish, or preparation of the manuscript.
+
 \subsection*{Acknowledgements}\label{acknowledgements}
@@
 insights into the historical context of the data collection and
-curation. Funding for this research was provided by the North Central
-Soybean Research Program (\#639K623) to SEE, a Layman Award (\#2446) to
-SEE, USDA-ARS National Sclerotinia Initiative (\#58-5442-2-209) to
-JRS/SEE, and start-up funds from the University of Nebraska-Lincoln to
-SEE.
+curation.
 
@@
 Leslie, J. (1993). Fungal vegetative compatibility. \emph{Annual Review
-of Phytopathology}, \emph{31}, 127--150. Review.
+Of Phytopathology}, \emph{31}, 127--150. Review.
 \url{https://doi.org/10.1146/annurev.py.31.090193.001015}

### 2017-09-20 doc/manuscript/manuscript.Rmd
fix spelling of Sajeewa's name :no_mouth:

@@
     affiliation: 1
-  - name: Bimal Sajeewa Amadarasa
-    email: bamadarasa2@unl.edu
+  - name: Bimal Sajeewa Amaradasa
+    email: bamaradasa2@unl.edu
     affiliation: 1
@@
 
- - **Bimal Sajeewa Amadarasa** ...
+ - **Bimal Sajeewa Amaradasa** ...
 

### 2017-09-20 doc/manuscript/manuscript.Rmd
update Sajeewa's current address

@@
   - name: Bimal Sajeewa Amaradasa
-    email: bamaradasa2@unl.edu
+    email: bamaradasa@ufl.edu
     affiliation: 1
@@
   - code: 1
-    address: Department of Plant Pathology, University of Nebraska-Lincoln, Lincoln 68583
+    address: Department of Plant Pathology, University of Nebraska-Lincoln, Lincoln, NE 68583
 abstract: >
   `r paste(readLines("abstract.md"), collapse = "\n  ")`
-header-includes: >
+header-includes: |
   \usepackage{indentfirst}
+
+  \newaff{second}{Plant Pathology Department, University of Florida, Gainsville, FL 32611}
+
 bibliography: ssc_bibliography.bib

### 2017-09-27 doc/RMD/table-1.Rmd
update table S1 to work with new huxtable

@@
   set_align(everywhere, col = 6, value = "right") %>%     # 
-  set_wrap(TRUE) %>%                                      # Wrapping columns
-  set_number_format(value = 0)                            # Print numbers without decimals
-col_width(dt) <- c(0.055, 0.035, 0.06, 0.1, 0.151, 0.025) # Specify width for each column
-print_md(dt)                                              # print in pandoc markdown format
+  # set_wrap(TRUE) %>%                                      # Wrapping columns
+  set_number_format(value = 0) %>%                        # Print numbers without decimals
+  set_col_width(c(0.055, 0.035, 0.06, 0.1, 0.151, 0.025)) %>% # Specify width for each column
+  print_md(max_width = 80)                                # print in pandoc markdown format
 ```

### 2017-09-27 doc/manuscript/manuscript.Rmd
Address co-author comments.

@@
 
+# Genotype Diversity Table ------------------------------------------------
+
+gdt <- read_csv("../../results/tables/genotype_diversity_table_region.csv", 
+         na = "NA", # spaces read in as spaces
+         col_types = cols(
+         Pop = col_character(),
+         N = col_integer(),
+         MLG = col_integer(),
+         eMLG = col_double(),
+         SE = col_double(),
+         H = col_double(),
+         G = col_double(),
+         lambda = col_double(),
+         E.5 = col_double(),
+         Hexp = col_double(),
+         Ia = col_double(),
+         p.Ia = col_double(),
+         rbarD = col_double(),
+         p.rD = col_double(),
+         File = col_character()
+      ))
+gdtrd <- setNames(signif(gdt$p.rD, 3), gdt$Pop)
+
 # DAPC data ---------------------------------------------------------------
@@
 It causes disease on more than 400 plant species belonging to 75 families [@boland1994index] including crops of major economic importance such as sunflower (*Helianthus spp.*), soybean (*Glycine max* L.), canola (*Brassica napa* L., *Brassica campestris* L.), and dry bean (*Phaseolus vulgaris* L.) [@bolton2006sclerotinia].
-On dry bean, *S. sclerotiorum* is the causal agent of white mold, a devastating disease that can be a yield-limiting in temperate climates [@steadman1983white].
+On dry bean, *S. sclerotiorum* is the causal agent of white mold, a devastating disease that can be yield-limiting in temperate climates [@steadman1983white].
 
-Currently, there are no known resistant cultivars of dry bean commercially available [@otto-hanson2011variation].
+Currently, there are no commercially available resistant cultivars of dry bean [@otto-hanson2011variation].
 Organized breeding efforts have used a common-garden approach with white mold screening nurseries in dry bean production areas across the United States with additional sites in Mexico, France, and Australia [@steadman2003identification; @steadman2004identification; @steadman2005identification; @steadman2006identification; @otto2007identification; @otto2008identification; @mccoy2009use].
@@
 In this study, we analyze and characterize the genetic and phenotypic diversity of 366 *S.
-sclerotiorum* isolates collected over 10 years between 2003 and 2012 from dry bean cultivars among different geographic locations in the United States, France, Mexico, and Australia.
+sclerotiorum* isolates collected between 2003 and 2012 from dry bean cultivars among different geographic locations in the United States, France, Mexico, and Australia.
 We wanted to know if the *S. sclerotiorum* populations from white mold screening nurseries were representative of the fields within the same region. 
 As these nurseries were not treated with any chemical or cultural control of white mold, we hypothesized that these nurseries would represent the natural population of *S. sclerotiorum*.
-Furthermore, we wanted to investigate the potential effect of cultivar on genetic diversity by assessing three dry bean cultivars with different levels of resistance, Beryl (great northern bean, susceptible), Bunsi (navy bean, low susceptibility), and G122 (cranberry bean, partial resistance) [@otto-hanson2011variation].
+Furthermore, we wanted to investigate the potential effect of cultivar on genetic diversity of the pathogen by assessing three dry bean cultivars with different levels of resistance, Beryl (great northern bean, susceptible), Bunsi (navy bean, low susceptibility), and G122 (cranberry bean, partial resistance) [@otto-hanson2011variation].
 We additionally wanted to determine categorical or phenotypic variables that best predicted genetic structure and if there was correlation between multilocus haplotype and mycelial compatibility group.
@@
 White mold screening nurseries were 5m x 10m in size and maintained without application of fungicides to observe natural incidence of white mold.
-The nursery plots were incorporated with a basal dressing of N:P:K = 1:3:2 and side dressing of 0:3:2 during the growing season [@steadman2003identification].
+The early nursery plots were incorporated with a basal dressing of N:P:K = 1:3:2 and side dressing of 0:3:2 during the growing season [@steadman2003identification].
 
@@
 After collection, sclerotia were stored in Petri plates lined with filter paper, then stored at $20~^{\circ}$F or -$4~^{\circ}$C. 
-Sclerotia were surface-sterilized with 50% Clorox bleach (at least 6% NaOCL, The Clorox Company, Oakland, CA) solution for 3 min, and double rinsed with ddH$_2$O 3 min.
+Sclerotia were surface-sterilized with 50% Clorox bleach (at least 6% NaOCL, The Clorox Company, Oakland, CA) solution for 3 min, and double rinsed with ddH$_2$O for 3 min.
 The sterilized sclerotia were then placed on water agar plates (16g of Bacto agar per liter of ddH$_2$O, BD Diagnostic Systems, Sparks, MD), with four to five sclerotia of each isolate separated on each plate and stored on the counter top at room temperature for 5 to 6 days.
@@
 Briefly, the straw test uses 21-day-old G122 plants as the host in a greenhouse setting. 
-Clear drinking straws cut to 2.5 cm and sealed were used to place two mycelial plugs of inoculum on the host plant after removing all plant growth &gt;2.5 cm above the fourth node.
+Clear drinking straws cut to 2.5 cm and sealed were used to place two mycelial plugs of inoculum on the host plant after removing plant growth beyond 2.5 cm above the fourth node.
 Measurements of the necrotic lesion were taken 8 days later using the Modified Petzoldt and Dickson scale of 1--9, where 1 is no disease and 9 is plant death [@petzoldt1996straw; @teran2006modified].
@@
 Prior to DNA extraction, isolates were grown on PDA and plugs were subsequently transferred to Potato Dextrose Broth (PDB) where they were grown until there was significant mycelial growth, but before the mycelial mat became solidified (4--5 days).
-Mycelia for each plate was collected in a buchner funnel, agar plugs removed, lyophilized and pulverized manually in Whirl-pak® HDPE sampling bags (Sigma-Aldrich, St. Louis, MO). 
+Each mycelial mat was collected in a filtered Büchner funnel, agar plugs removed, lyophilized and pulverized manually in Whirl-pak® HDPE sampling bags (Sigma-Aldrich, St. Louis, MO). 
 Lyophilized mycelia was then stored in microcentrifuge tubes at -$20~^{\circ}$C until needed for DNA extraction.
@@
 Nearly all populations showed evidence for linkage (Table \@ref(tab:iatab)), which serves as evidence for clonal reproduction or other forms of non-random mating. 
-The only exceptions are CA and Australia. Both of these populations showed only moderate significance (P < 0.05) with $\bar{r}_d$ values of 0.03 and 0.12, respectively.
+The only exceptions are CA (*P* = `r gdtrd[["CA"]]`) and Australia (*P* = `r gdtrd[["Australia"]]`).
+Both of these populations showed only moderate significance with $\bar{r}_d$ values of 0.03 and 0.12, respectively.
 
@@
 ```{r table_ia, echo = FALSE, warning = FALSE, message = FALSE, results = "asis"}
-suppressPackageStartupMessages(library("tidyverse"))
-read_csv("../../results/tables/genotype_diversity_table_region.csv", 
-         na = "NA", # spaces read in as spaces
-         col_types = cols(
-         Pop = col_character(),
-         N = col_integer(),
-         MLG = col_integer(),
-         eMLG = col_double(),
-         SE = col_double(),
-         H = col_double(),
-         G = col_double(),
-         lambda = col_double(),
-         E.5 = col_double(),
-         Hexp = col_double(),
-         Ia = col_double(),
-         p.Ia = col_double(),
-         rbarD = col_double(),
-         p.rD = col_double(),
-         File = col_character()
-      )) %>%
+gdt %>%
   # mutate(Ia = case_when(
@@
 ANOVA showed that the reduced model was significant with an adjusted R^2^ of `r RDA1R2` (*P* = 0.001).
-Assessment of the marginal effects showed that all varaibles significantly explained genetic variation (*P* = `r anova1res[["Year"]]`, *P* = `r anova1res[["Region"]]` for Region).
+Assessment of the marginal effects showed that all varaibles significantly explained genetic variation ($P \leq `r anova1res[["Region"]]`$).
 We found that there was multicollinearity when MCG was combined with any other variable, so repeated the analysis, dropping MCG from the list of potential predictors. 
@@
 MCGs also showed a significant effect (*P* < 1.00e^-4^), with means that ranged from 6.0 ('MCG 44') to 4.6 ('MCG 49'; Table \@ref(tab:MCG-aggressiveness)).
-We additionally found a significant effect for MLHs (*P* = 7.44e^-4^) for MLHs, with means that ranged from 6.0 ('MLH 78') to 4.3 ('MLH 140') (Table \@ref(tab:MLG-aggressiveness)).
+We additionally found a significant effect for MLHs (*P* = 7.44e^-4^), with means that ranged from 6.0 ('MLH 78') to 4.3 ('MLH 140') (Table \@ref(tab:MLG-aggressiveness)).
 
@@
 Despite this information, it is still unclear what has contributed to the differentiation of the 2008 population from WA.
-While the evidence my suggest host as being a factor, previous studies have shown no significant differentiation across host species [@aldrich-wolfe2015genetic].
+While the evidence may suggest host as being a factor, previous studies have shown no significant differentiation across host species [@aldrich-wolfe2015genetic].
 
@@
 One long-held assumption was that MCGs (as determined via barrage reaction) represent vegetative compatibility groups (VCGs) [@schafer2006optimized; @kohn1990mycelial; @lehner2015genetic], which are known to have a genetic component [@saupe2000molecular; @hall2010evolution; @strom2016genomes].
-While our protocol for assessing MCGs utilized a different substrate for the MCG reactions, the patterns we observe are not dissimilar from what have previously been reported in the literature.
+While our protocol for assessing MCGs utilized Diana Sermons Medium [@cubeta2001mycelial] as compared to Patterson's Medium or Potato Dextrose Agar [@schafer2006optimized] for the MCG reactions, the patterns we observe are not dissimilar from what have previously been reported in the literature.
 It has been demonstrated in several Ascomycetes---including *Neurospora crassa* [@micali2003independence], *Sclerotinia homoecarpa* [@jo2008reassessment], *Verticillium dahliae* [@papaioannou2014barrage], and *S.
@@
 While we found no direct relationship between haplotype and severity, it is evident that there is a gradient of severity by region, further supporting the need for screening in multiple locations.
-Based on our analysis of the relationships between MCG and MLH, we found no clear evidence that the two are directly related, suggesting that MCG do not necessarily represent vegetative compatibility groups and thus should not be used as a proxy for identifying clones.
+Based on our analysis of the relationships between MCG and MLH, we found no clear evidence that the two are directly related, suggesting that MCG does not necessarily represent vegetative compatibility groups and thus should not be used as a proxy for identifying clones.
 
@@
 
-The authors would like to thank Rebecca Higgens for providing valuable insights into the historical context of the data collection and curation.
+The authors would like to thank Rebecca Higgens for technical support in generating the data for the MCG assessment, aggressiveness ratings, and genotyping; and for providing valuable insights into the historical context of the data collection and curation.
 
@@
 
- - **Zhian N. Kamvar** analyzed the data, contributed analysis tools, wrote the paper, prepared figures and tables, reviewed drafts of the paper.
+ - **Zhian N. Kamvar** analyzed the data, contributed analysis tools, wrote the paper, prepared figures and tables, edited and reviewed drafts of the paper.
 
- - **Bimal Sajeewa Amaradasa** ...
+ - **Bimal Sajeewa Amaradasa** analyzed the data, contributed analysis tools, wrote drafts of the paper, edited and reviewed drafts of the paper.
 
- - **Rachana Jhala** ...
+ - **Rachana Jhala** Carried out experiments (MCG assessment, aggressiveness ratings, genotyping), edited and reviewed drafts of the paper.
 
- - **Serena McCoy** ...
+ - **Serena McCoy** Carried out experiments (MCG assessment, aggressiveness ratings, genotyping), edited and reviewed drafts of the paper.
 
- - **James Steadman** ...
+ - **James Steadman** Conceived and designed experiments, organized network of white mold screening nurseries, provided *S. sclerotiorum* isolates, edited and reviewed drafts of the paper.
 
- - **Sydney E. Everhart** ...
+ - **Sydney E. Everhart** supervised data analysis, analyzed the data, contributed analysis tools, wrote the paper, edited and reviewed drafts of the paper.
 

### 2017-09-29 doc/manuscript/manuscript.tex
fix spelling of Sajeewa's name :no_mouth:
update pdf manuscript

@@
 
-\author[1]{Bimal Sajeewa Amadarasa}
+\author[1]{Bimal Sajeewa Amaradasa}
 
@@
 
-\affil[1]{Department of Plant Pathology, University of Nebraska-Lincoln, Lincoln
-68583}
+\affil[1]{Department of Plant Pathology, University of Nebraska-Lincoln, Lincoln,
+NE 68583}
 
@@
 al., 2006). On dry bean, \emph{S. sclerotiorum} is the causal agent of
-white mold, a devastating disease that can be a yield-limiting in
+white mold, a devastating disease that can be yield-limiting in
 temperate climates (Steadman, 1983).
 
-Currently, there are no known resistant cultivars of dry bean
-commercially available (Otto-Hanson et al., 2011). Organized breeding
-efforts have used a common-garden approach with white mold screening
-nurseries in dry bean production areas across the United States with
-additional sites in Mexico, France, and Australia (Steadman et al.,
-2003, 2004, 2005, 2006; Otto-Hanson \& Steadman, 2007, 2008; McCoy \&
-Steadman, 2009). These white mold screening nurseries use no chemical or
-cultural treatments against \emph{S. sclerotiorum} and employ
-standardized protocols for screening new cultivars for resistance to
-white mold. It has previously been shown that aggressiveness (the
-severity of disease symptoms on the host) is significantly different
-across white mold screening nursery sites in separate geographic regions
-(Otto-Hanson et al., 2011). The genetic structure and mode of
-reproduction in these populations, however, is currently unknown.
+Currently, there are no commercially available resistant cultivars of
+dry bean (Otto-Hanson et al., 2011). Organized breeding efforts have
+used a common-garden approach with white mold screening nurseries in dry
+bean production areas across the United States with additional sites in
+Mexico, France, and Australia (Steadman et al., 2003, 2004, 2005, 2006;
+Otto-Hanson \& Steadman, 2007, 2008; McCoy \& Steadman, 2009). These
+white mold screening nurseries use no chemical or cultural treatments
+against \emph{S. sclerotiorum} and employ standardized protocols for
+screening new cultivars for resistance to white mold. It has previously
+been shown that aggressiveness (the severity of disease symptoms on the
+host) is significantly different across white mold screening nursery
+sites in separate geographic regions (Otto-Hanson et al., 2011). The
+genetic structure and mode of reproduction in these populations,
+however, is currently unknown.
 
@@
 In this study, we analyze and characterize the genetic and phenotypic
-diversity of 366 \emph{S. sclerotiorum} isolates collected over 10 years
-between 2003 and 2012 from dry bean cultivars among different geographic
-locations in the United States, France, Mexico, and Australia. We wanted
-to know if the \emph{S. sclerotiorum} populations from white mold
-screening nurseries were representative of the fields within the same
-region. As these nurseries were not treated with any chemical or
-cultural control of white mold, we hypothesized that these nurseries
-would represent the natural population of \emph{S. sclerotiorum}.
-Furthermore, we wanted to investigate the potential effect of cultivar
-on genetic diversity by assessing three dry bean cultivars with
-different levels of resistance, Beryl (great northern bean,
-susceptible), Bunsi (navy bean, low susceptibility), and G122 (cranberry
-bean, partial resistance) (Otto-Hanson et al., 2011). We additionally
-wanted to determine categorical or phenotypic variables that best
-predicted genetic structure and if there was correlation between
-multilocus haplotype and mycelial compatibility group. Knowing what
-variables predict genetic structure can help direct breeding efforts. By
-investigating these aims, we hope to effectively describe the population
-structure of \emph{S. sclerotiorum} in the USA and make available our
-database of isolates for use in future dry bean breeding efforts.
+diversity of 366 \emph{S. sclerotiorum} isolates collected between 2003
+and 2012 from dry bean cultivars among different geographic locations in
+the United States, France, Mexico, and Australia. We wanted to know if
+the \emph{S. sclerotiorum} populations from white mold screening
+nurseries were representative of the fields within the same region. As
+these nurseries were not treated with any chemical or cultural control
+of white mold, we hypothesized that these nurseries would represent the
+natural population of \emph{S. sclerotiorum}. Furthermore, we wanted to
+investigate the potential effect of cultivar on genetic diversity of the
+pathogen by assessing three dry bean cultivars with different levels of
+resistance, Beryl (great northern bean, susceptible), Bunsi (navy bean,
+low susceptibility), and G122 (cranberry bean, partial resistance)
+(Otto-Hanson et al., 2011). We additionally wanted to determine
+categorical or phenotypic variables that best predicted genetic
+structure and if there was correlation between multilocus haplotype and
+mycelial compatibility group. Knowing what variables predict genetic
+structure can help direct breeding efforts. By investigating these aims,
+we hope to effectively describe the population structure of \emph{S.
+sclerotiorum} in the USA and make available our database of isolates for
+use in future dry bean breeding efforts.
 
@@
 5m x 10m in size and maintained without application of fungicides to
-observe natural incidence of white mold. The nursery plots were
+observe natural incidence of white mold. The early nursery plots were
 incorporated with a basal dressing of N:P:K = 1:3:2 and side dressing of
@@
 NaOCL, The Clorox Company, Oakland, CA) solution for 3 min, and double
-rinsed with ddH\(_2\)O 3 min. The sterilized sclerotia were then placed
-on water agar plates (16g of Bacto agar per liter of ddH\(_2\)O, BD
-Diagnostic Systems, Sparks, MD), with four to five sclerotia of each
+rinsed with ddH\(_2\)O for 3 min. The sterilized sclerotia were then
+placed on water agar plates (16g of Bacto agar per liter of ddH\(_2\)O,
+BD Diagnostic Systems, Sparks, MD), with four to five sclerotia of each
 isolate separated on each plate and stored on the counter top at room
@@
 drinking straws cut to 2.5 cm and sealed were used to place two mycelial
-plugs of inoculum on the host plant after removing all plant growth
-\textgreater{}2.5 cm above the fourth node. Measurements of the necrotic
-lesion were taken 8 days later using the Modified Petzoldt and Dickson
-scale of 1--9, where 1 is no disease and 9 is plant death (Petzoldt \&
-Dickson, 1996; Teran et al., 2006).
+plugs of inoculum on the host plant after removing plant growth beyond
+2.5 cm above the fourth node. Measurements of the necrotic lesion were
+taken 8 days later using the Modified Petzoldt and Dickson scale of
+1--9, where 1 is no disease and 9 is plant death (Petzoldt \& Dickson,
+1996; Teran et al., 2006).
 
@@
 grown until there was significant mycelial growth, but before the
-mycelial mat became solidified (4--5 days). Mycelia for each plate was
-collected in a buchner funnel, agar plugs removed, lyophilized and
-pulverized manually in Whirl-pak® HDPE sampling bags (Sigma-Aldrich,
+mycelial mat became solidified (4--5 days). Each mycelial mat was
+collected in a filtered Büchner funnel, agar plugs removed, lyophilized
+and pulverized manually in Whirl-pak® HDPE sampling bags (Sigma-Aldrich,
 St.~Louis, MO). Lyophilized mycelia was then stored in microcentrifuge
@@
 serves as evidence for clonal reproduction or other forms of non-random
-mating. The only exceptions are CA and Australia. Both of these
-populations showed only moderate significance (P \textless{} 0.05) with
-\(\bar{r}_d\) values of 0.03 and 0.12, respectively.
-
-\begin{longtable}[]{@{}lrrrrr@{}}
+mating. The only exceptions are CA (\emph{P} = 0.043) and Australia
+(\emph{P} = 0.052). Both of these populations showed only moderate
+significance with \(\bar{r}_d\) values of 0.03 and 0.12, respectively.
+
+\begin{longtable}[]{@{}rrllrl@{}}
 \caption{\label{tab:locus-stats} Allelic diversity on full data set at loci
@@
 \toprule
-\begin{minipage}[b]{0.09\columnwidth}\raggedright\strut
+\begin{minipage}[b]{0.08\columnwidth}\raggedleft\strut
 Locus\strut
@@
 Range\strut
-\end{minipage} & \begin{minipage}[b]{0.16\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[b]{0.15\columnwidth}\raggedright\strut
 Repeat Motif\strut
-\end{minipage} & \begin{minipage}[b]{0.15\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[b]{0.14\columnwidth}\raggedright\strut
 No. alleles\strut
-\end{minipage} & \begin{minipage}[b]{0.08\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[b]{0.06\columnwidth}\raggedleft\strut
 \emph{h}\strut
-\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
 Evenness\strut
@@
 \endhead
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
 5-2\strut
@@
 318--324\strut
-\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright\strut
 di-\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
 4\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
 0.45\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
 0.62\strut
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
 6-2\strut
@@
 483--495\strut
-\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright\strut
 hexa-\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
 3\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
 0.64\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
 0.95\strut
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
 7-2\strut
@@
 158--174\strut
-\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright\strut
 di-\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
 7\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
 0.73\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
 0.76\strut
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
 8-3\strut
@@
 244--270\strut
-\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright\strut
 di-\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
 7\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
 0.74\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
 0.79\strut
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
 9-2\strut
@@
 360--382\strut
-\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright\strut
 di-\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
 9\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
 0.35\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
 0.41\strut
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
 12-2\strut
@@
 214--222\strut
-\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright\strut
 di-\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
 5\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
 0.58\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
 0.78\strut
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
 17-3\strut
@@
 342--363\strut
-\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright\strut
 tri-\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
 7\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
 0.55\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
 0.53\strut
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
 20-3\strut
@@
 280--282\strut
-\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright\strut
 di-\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
 2\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
 0.05\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
 0.42\strut
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
 55-4\strut
@@
 153--216\strut
-\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright\strut
 tetra-\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
 10\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
 0.72\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
 0.66\strut
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
 110-4\strut
@@
 370--386\strut
-\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright\strut
 tetra-\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
 5\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
 0.76\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
 0.91\strut
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.09\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
 114-4\strut
@@
 339--416\strut
-\end{minipage} & \begin{minipage}[t]{0.16\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright\strut
 tetra-\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
 10\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
 0.83\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
 0.80\strut
@@
 
-\begin{longtable}[]{@{}lrrrrrrrl@{}}
+\begin{longtable}[]{@{}crrrrllrl@{}}
 \caption{\label{tab:iatab} Genotypic diversity and Linkage Disequilibrium
@@
 \toprule
-\begin{minipage}[b]{0.11\columnwidth}\raggedright\strut
+\begin{minipage}[b]{0.12\columnwidth}\centering\strut
 Pop\strut
-\end{minipage} & \begin{minipage}[b]{0.09\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[b]{0.08\columnwidth}\raggedleft\strut
 N\strut
-\end{minipage} & \begin{minipage}[b]{0.14\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\raggedleft\strut
 eMLH\strut
-\end{minipage} & \begin{minipage}[b]{0.05\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[b]{0.04\columnwidth}\raggedleft\strut
 H\strut
@@
 G\strut
-\end{minipage} & \begin{minipage}[b]{0.11\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[b]{0.11\columnwidth}\raggedright\strut
 \(\lambda\)\strut
-\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[b]{0.06\columnwidth}\raggedright\strut
 \(E_5\)\strut
@@
 \emph{h}\strut
-\end{minipage} & \begin{minipage}[b]{0.06\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\raggedright\strut
 \(\bar{ r}_d\)\strut
@@
 \endhead
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.12\columnwidth}\centering\strut
 WA\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
 58 (56)\strut
-\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedleft\strut
 9.95 (0.23)\strut
-\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
 4.0\strut
@@
 54.3\strut
-\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
 0.98\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
 0.98\strut
@@
 0.60\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
 0.07*\strut
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.12\columnwidth}\centering\strut
 MI\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
 58 (43)\strut
-\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedleft\strut
 9.3 (0.79)\strut
-\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
 3.6\strut
@@
 29.0\strut
-\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
 0.97\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
 0.78\strut
@@
 0.54\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
 0.14*\strut
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.12\columnwidth}\centering\strut
 ND\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
 41 (35)\strut
-\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedleft\strut
 9.44 (0.73)\strut
-\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
 3.5\strut
@@
 25.9\strut
-\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
 0.96\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
 0.82\strut
@@
 0.54\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
 0.1*\strut
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.12\columnwidth}\centering\strut
 NE\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
 37 (28)\strut
-\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedleft\strut
 8.93 (0.94)\strut
-\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
 3.2\strut
@@
 17.8\strut
-\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
 0.94\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
 0.75\strut
@@
 0.55\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
 0.25*\strut
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.12\columnwidth}\centering\strut
 CO\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
 34 (28)\strut
-\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedleft\strut
 9.46 (0.67)\strut
-\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
 3.3\strut
@@
 24.1\strut
-\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
 0.96\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
 0.92\strut
@@
 0.56\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
 0.27*\strut
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.12\columnwidth}\centering\strut
 France\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
 21 (14)\strut
-\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedleft\strut
 8.5 (0.85)\strut
-\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
 2.6\strut
@@
 12.6\strut
-\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
 0.92\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
 0.95\strut
@@
 0.48\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
 0.11*\strut
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.12\columnwidth}\centering\strut
 CA\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
 18 (15)\strut
-\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedleft\strut
 9.12 (0.72)\strut
-\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
 2.7\strut
@@
 13.5\strut
-\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
 0.93\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
 0.94\strut
@@
 0.51\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
 0.03\strut
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.12\columnwidth}\centering\strut
 OR\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
 17 (13)\strut
-\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedleft\strut
 8.52 (0.85)\strut
-\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
 2.5\strut
@@
 10.7\strut
-\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
 0.91\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
 0.89\strut
@@
 0.47\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
 0.1*\strut
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.12\columnwidth}\centering\strut
 Mexico\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
 15 (9)\strut
-\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedleft\strut
 7.1 (0.85)\strut
-\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
 2.1\strut
@@
 7.3\strut
-\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
 0.86\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
 0.89\strut
@@
 0.28\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
 0.37*\strut
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.12\columnwidth}\centering\strut
 MN\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
 9 (7)\strut
-\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedleft\strut
 7 (0)\strut
-\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
 1.9\strut
@@
 6.2\strut
-\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
 0.84\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
 0.93\strut
@@
 0.47\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
 0.19*\strut
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.12\columnwidth}\centering\strut
 Australia\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
 6 (6)\strut
-\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedleft\strut
 6 (0)\strut
-\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
 1.8\strut
@@
 6.0\strut
-\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
 0.83\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
 1.00\strut
@@
 0.48\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
 0.12\strut
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.12\columnwidth}\centering\strut
 WI\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
 2 (2)\strut
-\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedleft\strut
 2 (0)\strut
-\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
 0.7\strut
@@
 2.0\strut
-\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
 0.50\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
 1.00\strut
@@
 0.27\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
 -\strut
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.12\columnwidth}\centering\strut
 NY\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
 1 (1)\strut
-\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedleft\strut
 1 (0)\strut
-\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
 0.0\strut
@@
 1.0\strut
-\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
 0.00\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
 NaN\strut
@@
 NaN\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
 -\strut
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.12\columnwidth}\centering\strut
 ID\strut
-\end{minipage} & \begin{minipage}[t]{0.09\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
 1 (1)\strut
-\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedleft\strut
 1 (0)\strut
-\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
 0.0\strut
@@
 NaN\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
 -\strut
@@
 
-\begin{figure}[htbp]
+\begin{figure}
 \centering
@@
 Assessment of the marginal effects showed that all varaibles
-significantly explained genetic variation (\emph{P} = 0.001, \emph{P} =
-0.007 for Region). We found that there was multicollinearity when MCG
-was combined with any other variable, so repeated the analysis, dropping
-MCG from the list of potential predictors. From these results, Year,
-Region, Host, and Aggressiveness were found to be optimal, accounting
-for 17.6\% of the total variation. ANOVA revealed significant effects
-with an adjusted R\textsuperscript{2} of 0.0325 (\emph{P} = 0.001).
-While the marginal effect assessment revealed that Year, Region, and
-Host significantly explained variation at \emph{P} = 0.001, and
-Aggressiveness significantly explained variation at \emph{P} = 0.039.
-Much of the variation appeared to be driven by isolates from Mexico and
-2005 (Fig. \ref{rda-biplot}). Variance partitioning of the independent
-variables without MCG indicated aggressiveness to be the least
-influential factor with 0.1\% contributing to explaining the variation
-of molecular data, whereas the combination of variables accounted for
-3.3\%.
+significantly explained genetic variation (\(P \leq 0.007\)). We found
+that there was multicollinearity when MCG was combined with any other
+variable, so repeated the analysis, dropping MCG from the list of
+potential predictors. From these results, Year, Region, Host, and
+Aggressiveness were found to be optimal, accounting for 17.6\% of the
+total variation. ANOVA revealed significant effects with an adjusted
+R\textsuperscript{2} of 0.0325 (\emph{P} = 0.001). While the marginal
+effect assessment revealed that Year, Region, and Host significantly
+explained variation at \emph{P} = 0.001, and Aggressiveness
+significantly explained variation at \emph{P} = 0.039. Much of the
+variation appeared to be driven by isolates from Mexico and 2005 (Fig.
+\ref{rda-biplot}). Variance partitioning of the independent variables
+without MCG indicated aggressiveness to be the least influential factor
+with 0.1\% contributing to explaining the variation of molecular data,
+whereas the combination of variables accounted for 3.3\%.
 
@@
 \ref{tab:MCG-aggressiveness}). We additionally found a significant
-effect for MLHs (\emph{P} = 7.44e\textsuperscript{-4}) for MLHs, with
-means that ranged from 6.0 (`MLH 78') to 4.3 (`MLH 140') (Table
+effect for MLHs (\emph{P} = 7.44e\textsuperscript{-4}), with means that
+ranged from 6.0 (`MLH 78') to 4.3 (`MLH 140') (Table
 \ref{tab:MLG-aggressiveness}).
@@
 
-\begin{longtable}[]{@{}lrrrrr@{}}
+\begin{longtable}[]{@{}clrllc@{}}
 \caption{\label{tab:amova} Comparison of populations in the white mold
@@
 \toprule
-\begin{minipage}[b]{0.31\columnwidth}\raggedright\strut
+\begin{minipage}[b]{0.32\columnwidth}\centering\strut
 Hierarchy\strut
-\end{minipage} & \begin{minipage}[b]{0.06\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[b]{0.06\columnwidth}\raggedright\strut
 d.f.\strut
-\end{minipage} & \begin{minipage}[b]{0.06\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\raggedleft\strut
 S.S.\strut
-\end{minipage} & \begin{minipage}[b]{0.14\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\raggedright\strut
 \% variation\strut
-\end{minipage} & \begin{minipage}[b]{0.19\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[b]{0.19\columnwidth}\raggedright\strut
 \(\Phi\) statistic\strut
-\end{minipage} & \begin{minipage}[b]{0.05\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\centering\strut
 \emph{P}\strut
@@
 \endhead
-\begin{minipage}[t]{0.31\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.32\columnwidth}\centering\strut
 Between Region\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
 13\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedleft\strut
 10.19\strut
-\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
 8.45\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
 \textbf{0.0845}\strut
-\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering\strut
 0.029\strut
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.31\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.32\columnwidth}\centering\strut
 Between Source within Region\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
 8\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedleft\strut
 2.74\strut
-\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
 -2.29\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
 -0.0250\strut
-\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering\strut
 0.486\strut
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.31\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.32\columnwidth}\centering\strut
 Between Year within Source\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
 22\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedleft\strut
 9.37\strut
-\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
 16.28\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
 \textbf{0.173}\strut
-\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering\strut
 0.001\strut
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.31\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.32\columnwidth}\centering\strut
 Within Year\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
 274\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedleft\strut
 47.30\strut
-\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
 77.56\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
 \textbf{0.224}\strut
-\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering\strut
 0.001\strut
@@
 unclear what has contributed to the differentiation of the 2008
-population from WA. While the evidence my suggest host as being a
+population from WA. While the evidence may suggest host as being a
 factor, previous studies have shown no significant differentiation
@@
 have a genetic component (Saupe, 2000; Hall et al., 2010; Strom \&
-Bushley, 2016). While our protocol for assessing MCGs utilized a
-different substrate for the MCG reactions, the patterns we observe are
-not dissimilar from what have previously been reported in the
-literature. It has been demonstrated in several Ascomycetes---including
-\emph{Neurospora crassa} (Micali \& Smith, 2003), \emph{Sclerotinia
-homoecarpa} (Jo et al., 2008), \emph{Verticillium dahliae} (Papaioannou
-\& Typas, 2014), and \emph{S. sclerotiorum} (Ford et al., 1995)---that
-barrage reactions are independent from stable anastomosis. Thus, the
-inconsistencies in this study and other studies indicate that
-researchers studying \emph{S. sclerotiorum} should not rely on MCG data
-derived from barrage reactions as an indicator for genetic diversity.
+Bushley, 2016). While our protocol for assessing MCGs utilized Diana
+Sermons Medium (Cubeta et al., 2001) as compared to Patterson's Medium
+or Potato Dextrose Agar (Schafer \& Kohn, 2006) for the MCG reactions,
+the patterns we observe are not dissimilar from what have previously
+been reported in the literature. It has been demonstrated in several
+Ascomycetes---including \emph{Neurospora crassa} (Micali \& Smith,
+2003), \emph{Sclerotinia homoecarpa} (Jo et al., 2008),
+\emph{Verticillium dahliae} (Papaioannou \& Typas, 2014), and \emph{S.
+sclerotiorum} (Ford et al., 1995)---that barrage reactions are
+independent from stable anastomosis. Thus, the inconsistencies in this
+study and other studies indicate that researchers studying \emph{S.
+sclerotiorum} should not rely on MCG data derived from barrage reactions
+as an indicator for genetic diversity.
 
@@
 on our analysis of the relationships between MCG and MLH, we found no
-clear evidence that the two are directly related, suggesting that MCG do
-not necessarily represent vegetative compatibility groups and thus
+clear evidence that the two are directly related, suggesting that MCG
+does not necessarily represent vegetative compatibility groups and thus
 should not be used as a proxy for identifying clones.
@@
 
-The authors would like to thank Rebecca Higgens for providing valuable
-insights into the historical context of the data collection and
-curation.
+The authors would like to thank Rebecca Higgens for technical support in
+generating the data for the MCG assessment, aggressiveness ratings, and
+genotyping; and for providing valuable insights into the historical
+context of the data collection and curation.
 
@@
   \textbf{Zhian N. Kamvar} analyzed the data, contributed analysis
-  tools, wrote the paper, prepared figures and tables, reviewed drafts
+  tools, wrote the paper, prepared figures and tables, edited and
+  reviewed drafts of the paper.
+\item
+  \textbf{Bimal Sajeewa Amaradasa} analyzed the data, contributed
+  analysis tools, wrote drafts of the paper, edited and reviewed drafts
   of the paper.
 \item
-  \textbf{Bimal Sajeewa Amadarasa} \ldots{}
+  \textbf{Rachana Jhala} Carried out experiments (MCG assessment,
+  aggressiveness ratings, genotyping), edited and reviewed drafts of the
+  paper.
 \item
-  \textbf{Rachana Jhala} \ldots{}
+  \textbf{Serena McCoy} Carried out experiments (MCG assessment,
+  aggressiveness ratings, genotyping), edited and reviewed drafts of the
+  paper.
 \item
-  \textbf{Serena McCoy} \ldots{}
+  \textbf{James Steadman} Conceived and designed experiments, organized
+  network of white mold screening nurseries, provided \emph{S.
+  sclerotiorum} isolates, edited and reviewed drafts of the paper.
 \item
-  \textbf{James Steadman} \ldots{}
-\item
-  \textbf{Sydney E. Everhart} \ldots{}
+  \textbf{Sydney E. Everhart} supervised data analysis, analyzed the
+  data, contributed analysis tools, wrote the paper, edited and reviewed
+  drafts of the paper.
 \end{itemize}
@@
 
-\begin{longtable}[]{@{}lrl@{}}
+\begin{longtable}[]{@{}cll@{}}
 \caption{\label{tab:region-aggressiveness} Mean aggressiveness ratings for
@@
 \toprule
-\begin{minipage}[b]{0.14\columnwidth}\raggedright\strut
+\begin{minipage}[b]{0.14\columnwidth}\centering\strut
 Region\strut
-\end{minipage} & \begin{minipage}[b]{0.27\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[b]{0.25\columnwidth}\raggedright\strut
 Mean Aggressiveness\strut
@@
 \endhead
-\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.14\columnwidth}\centering\strut
 MN\strut
-\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
 5.84\strut
@@
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.14\columnwidth}\centering\strut
 ND\strut
-\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
 5.77\strut
@@
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.14\columnwidth}\centering\strut
 NE\strut
-\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
 5.29\strut
@@
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.14\columnwidth}\centering\strut
 MI\strut
-\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
 5.13\strut
@@
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.14\columnwidth}\centering\strut
 OR\strut
-\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
 4.84\strut
@@
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.14\columnwidth}\centering\strut
 CO\strut
-\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
 4.72\strut
@@
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.14\columnwidth}\centering\strut
 WA\strut
-\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
 4.67\strut
@@
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.14\columnwidth}\centering\strut
 France\strut
-\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
 4.66\strut
@@
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.14\columnwidth}\centering\strut
 Mexico\strut
-\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
 4.58\strut
@@
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.14\columnwidth}\centering\strut
 Australia\strut
-\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
 4.12\strut
@@
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.14\columnwidth}\centering\strut
 CA\strut
-\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
 4.01\strut
@@
 
-\begin{longtable}[]{@{}lrl@{}}
+\begin{longtable}[]{@{}cll@{}}
 \caption{\label{tab:MCG-aggressiveness} Mean aggressiveness ratings for the
@@
 \toprule
-\begin{minipage}[b]{0.14\columnwidth}\raggedright\strut
+\begin{minipage}[b]{0.10\columnwidth}\centering\strut
 MCG\strut
-\end{minipage} & \begin{minipage}[b]{0.27\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[b]{0.25\columnwidth}\raggedright\strut
 Mean Aggressiveness\strut
@@
 \endhead
-\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.10\columnwidth}\centering\strut
 44\strut
-\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
 6.03\strut
@@
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.10\columnwidth}\centering\strut
 3\strut
-\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
 5.50\strut
@@
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.10\columnwidth}\centering\strut
 5\strut
-\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
 5.40\strut
@@
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.10\columnwidth}\centering\strut
 2\strut
-\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
 5.25\strut
@@
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.10\columnwidth}\centering\strut
 9\strut
-\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
 5.11\strut
@@
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.10\columnwidth}\centering\strut
 1\strut
-\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
 4.95\strut
@@
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.10\columnwidth}\centering\strut
 45\strut
-\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
 4.88\strut
@@
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.10\columnwidth}\centering\strut
 4\strut
-\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
 4.87\strut
@@
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.10\columnwidth}\centering\strut
 53\strut
-\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
 4.69\strut
@@
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.10\columnwidth}\centering\strut
 49\strut
-\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
 4.60\strut
@@
 
-\begin{longtable}[]{@{}lrl@{}}
+\begin{longtable}[]{@{}cll@{}}
 \caption{\label{tab:MLG-aggressiveness} Mean aggressiveness ratings for the
@@
 \toprule
-\begin{minipage}[b]{0.14\columnwidth}\raggedright\strut
+\begin{minipage}[b]{0.10\columnwidth}\centering\strut
 MLH\strut
-\end{minipage} & \begin{minipage}[b]{0.27\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[b]{0.25\columnwidth}\raggedright\strut
 Mean Aggressiveness\strut
@@
 \endhead
-\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.10\columnwidth}\centering\strut
 78\strut
-\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
 6.07\strut
@@
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.10\columnwidth}\centering\strut
 65\strut
-\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
 5.94\strut
@@
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.10\columnwidth}\centering\strut
 9\strut
-\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
 5.67\strut
@@
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.10\columnwidth}\centering\strut
 25\strut
-\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
 5.41\strut
@@
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.10\columnwidth}\centering\strut
 66\strut
-\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
 5.30\strut
@@
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.10\columnwidth}\centering\strut
 104\strut
-\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
 5.22\strut
@@
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.10\columnwidth}\centering\strut
 160\strut
-\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
 4.80\strut
@@
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.10\columnwidth}\centering\strut
 163\strut
-\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
 4.80\strut
@@
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.10\columnwidth}\centering\strut
 165\strut
-\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
 4.34\strut
@@
 \end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
+\begin{minipage}[t]{0.10\columnwidth}\centering\strut
 140\strut
-\end{minipage} & \begin{minipage}[t]{0.27\columnwidth}\raggedleft\strut
+\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
 4.31\strut
@@
 Leslie, J. (1993). Fungal vegetative compatibility. \emph{Annual Review
-Of Phytopathology}, \emph{31}, 127--150. Review.
+of Phytopathology}, \emph{31}, 127--150. Review.
 \url{https://doi.org/10.1146/annurev.py.31.090193.001015}

### 2017-10-02 README.Rmd
Create Rmd file for README

---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# Analysis of 366 *S. sclerotiorum* isolates

[![Last-changedate](https://img.shields.io/badge/last%20change-`r gsub('-', '--', Sys.Date())`-brightgreen.svg)](https://github.com/everhartlab/sclerotinia-366/commits/master) [![minimal R version](https://img.shields.io/badge/R%3E%3D-3.4.1-brightgreen.svg)](https://cran.r-project.org/) [![Licence](https://img.shields.io/github/license/mashape/apistatus.svg)](http://choosealicense.com/licenses/mit/)  [![Circle CI](https://circleci.com/gh/everhartlab/sclerotinia-366.svg?style=shield&circle-token=:circle-token)](https://circleci.com/gh/everhartlab/sclerotinia-366)

This repository contains data, code, and a manuscript for analysis of 366 isolates of *Sclerotinia sclerotiorum* from the US and various countries around the world.

> Population structure and phenotypic variation of *Sclerotinia sclerotiorum*
> from dry bean in the United States
>
> Z. N. Kamvar, B. S. Amaradasa, R. Jhala, S. McCoy, J. Steadman, and S. E. Everhart

# TOC

The analyses are arranged in the following order according to the [Makefile]:

 1. [table-1.md]
 1. [MCG-virulence.md]
 1. [locus-stats.md]
 1. [MLG-distribution.md]
 1. [mlg-mcg.md]
 1. [RDA-analysis.md]
 1. [pop-diff.md]
 1. [tree.md]
 1. [wmn-differentiation.md]
 1. [by-year.md]

# Reproducing the analysis

## Locally

This project is controlled via a [Makefile] which means that everything (analyses, tables, figures, the paper itself) is controlled via one command:

```
make
```

This will bootstrap the installation (warning: it will update packages),
process the data, perform the analyses, and compile the paper.

Required software:

 - GNU Make (If you're on Windows, you can use MinGW: http://www.mingw.org/)
 - [R (version 3.4.1 or greater)](https://r-project.org)
 - [LaTeX](https://www.latex-project.org/get)
 - [pandoc](http://pandoc.org/) (Note: pandoc ships with Rstudio)
 - [devtools](https://github.com/hadley/devtools#readme)

## Docker

This repository contains a [Dockerfile](Dockerfile), which specifies the
instructions to build a [docker](https://www.docker.com/) container. This is
designed to capture the complete development environment of the analysis so
that it can be accurately reproduced. The image is ~2.71Gb, so be sure that you
have enough memory on your computer to run it.

To Install Docker, go here: https://docs.docker.com/engine/installation/#desktop. Once you have downloaded docker, you can either pull the container or build it. Pulling is by far the quickest way to do this.
The docker container is located at https://hub.docker.com/r/zkamvar/sclerotinia-366/

> Note: both ways assume that you are in the analysis directory

### Pulling the container

```
docker run --rm -it -v $(pwd):/analysis zkamvar/sclerotinia-366:latest bash
```

### Building the container

If you don't want to pull from docker hub, you can build the container.

```sh
docker build .
```

This will bootstrap the environment to build the container. Once it's finished
you should see:

```
Successfully built 8e1e4cd82e19
```

where `8e1e4cd82e19` will be replaced by your hash. Now that things are built,
you can run the analysis in the image with:

```
docker run -it -v $(pwd):/analysis 8e1e4cd82e19 bash
```

### Running the Analysis

Once you are in the container, you can run the analysis, which is mapped to
`analysis/`. The `make clean` command will wipe out all derivative files and
the `make` command will generate everything. Note that this took almost 2 hours
to run on my machine due to several bootstrapping processes.

```
cd analysis/
make clean
make
```

[Makefile]: Makefile
[table-1.md]: results/table-1.md
[MCG-virulence.md]: results/MCG-virulence.md
[locus-stats.md]: results/locus-stats.md
[MLG-distribution.md]: results/MLG-distribution.md
[mlg-mcg.md]: results/mlg-mcg.md
[RDA-analysis.md]: results/RDA-analysis.md
[pop-diff.md]: results/pop-diff.md
[tree.md]: results/tree.md
[wmn-differentiation.md]: results/wmn-differentiation.md
[by-year.md]: results/by-year.md

## Packages Used

```{r}
options(width = 100)
imports <- packageDescription("WorldSclerotinia")$Imports
imports <- strsplit(imports, "[^A-z]*,\n")[[1]]
for (i in imports) suppressPackageStartupMessages(library(i, character.only = TRUE))
devtools::session_info()
```

### 2017-10-02 doc/RMD/pop-diff.Rmd
remove parallelization for reproducibility
The parallelization here did not give consistent
results, ranging from 20 to 24 PCs. By setting it to
one thread, we retain the reproducibility.

@@
 region.xval <- xvalDapc(tab(dat11cc), strata(dat11cc)$Region, n.pca = 15:25, 
-                        n.rep = 100, parallel = "multicore", ncpus = 4)
+                        n.rep = 100)
 region.dapc <- region.xval$DAPC

### 2017-10-02 doc/manuscript/manuscript.tex
commiting most recent version of manuscript [skip ci]

@@
 between nodes using the R packages \emph{igraph} (Csardi \& Nepusz,
-2006), \emph{dplyr} version 0.7.3 (Wickham et al., 2017), and
+2006), \emph{dplyr} version 0.7.4 (Wickham et al., 2017), and
 \emph{purrr} version 0.2.3 (Henry \& Wickham, 2017). Each node was

### 2017-10-03 README.Rmd
add preprint citation

@@
 
-> Population structure and phenotypic variation of *Sclerotinia sclerotiorum*
-> from dry bean in the United States
->
-> Z. N. Kamvar, B. S. Amaradasa, R. Jhala, S. McCoy, J. Steadman, and S. E. Everhart
+# Citation
+
+## Manuscript
+
+> Kamvar ZN, Amaradasa BS, Jhala R, McCoy S, Steadman JR, Everhart SE.
+(2017) Population structure and phenotypic variation of *Sclerotinia sclerotiorum* from dry bean in the United States.
+*PeerJ Preprints* 5:e3311v1 <https://doi.org/10.7287/peerj.preprints.3311v1>
+
+## Data and Code
+
+> Kamvar, Z.
+N., Amaradasa, B.
+S., Jhala, R., McCoy, S., Steadman, J., & Everhart, S.
+E.
+(2017, October 3).
+Population structure and phenotypic variation of *Sclerotinia sclerotiorum* from dry bean in the United States.
+<http://doi.org/10.17605/OSF.IO/EJB5Y>
 
@@
 
-### Pulling the container
+### RStudio Server
 
 ```
-docker run --rm -it -v $(pwd):/analysis zkamvar/sclerotinia-366:latest bash
+docker run --rm -dp 8787:8787 zkamvar/sclerotinia-366:latest
 ```
 
-### Building the container
+This will spin up the Docker container on your machine and expose it to port 8787.
+You can open your browser and type `localhost:8787`, and an instance of Rstudio server will appear.
+Sign in with the username: rstudio and password: rstudio.
+
+Since the files in `/analysis` are write-protected, if you wanted to explore, you should copy the directory to your current working space:
+
+- in the R console type: `system("cp -R /analysis .")`. 
+- open `/analysis` and double click on znk_analysis.Rproj
+
+From here you can re-run the analyses to your heart's content.
+
+### Building the container locally
 
@@
 ```sh
-docker build .
+docker build -t sclerotinia-366 .
 ```
 
-This will bootstrap the environment to build the container. Once it's finished
-you should see:
+Now that things are built, you can run the analysis in the image with:
 
 ```
-Successfully built 8e1e4cd82e19
-```
-
-where `8e1e4cd82e19` will be replaced by your hash. Now that things are built,
-you can run the analysis in the image with:
-
-```
-docker run -it -v $(pwd):/analysis 8e1e4cd82e19 bash
+docker run -it sclerotinia-366 bash
 ```

### 2017-10-07 doc/manuscript/manuscript.Rmd
update cls file to include current aff
this is for lack of a better way to do this atm
Typo
Change NaOCL to NaOCl

@@
   \usepackage{indentfirst}
-
-  \newaff{second}{Plant Pathology Department, University of Florida, Gainsville, FL 32611}
 
@@
 
-# Genotype Diversity Table ------------------------------------------------
+# Genotype Diversity Table -----------------------------------------------
 
@@
 After collection, sclerotia were stored in Petri plates lined with filter paper, then stored at $20~^{\circ}$F or -$4~^{\circ}$C. 
-Sclerotia were surface-sterilized with 50% Clorox bleach (at least 6% NaOCL, The Clorox Company, Oakland, CA) solution for 3 min, and double rinsed with ddH$_2$O for 3 min.
+Sclerotia were surface-sterilized with 50% Clorox bleach (at least 6% NaOCl, The Clorox Company, Oakland, CA) solution for 3 min, and double rinsed with ddH$_2$O for 3 min.
 The sterilized sclerotia were then placed on water agar plates (16g of Bacto agar per liter of ddH$_2$O, BD Diagnostic Systems, Sparks, MD), with four to five sclerotia of each isolate separated on each plate and stored on the counter top at room temperature for 5 to 6 days.

### 2017-10-31 doc/manuscript/manuscript.Rmd
update title; add citation to first sentence; address #15

@@
 ---
-title: Population structure and phenotypic variation of *Sclerotinia sclerotiorum* from dry bean in the United States
+title: Population structure and phenotypic variation of *Sclerotinia sclerotiorum* from dry bean (*Phaseolus vulgaris*) in the United States
 preprint: false
@@
   - name: Bimal Sajeewa Amaradasa
-    email: bamaradasa@ufl.edu
+    email: b.amaradasa@ufl.edu
     affiliation: 1
@@
 
-*Sclerotinia sclerotiorum* (Lib.) de Bary is an ascomycete plant pathogen with a worldwide distribution.
+*Sclerotinia sclerotiorum* (Lib.) de Bary is an ascomycete plant pathogen with a worldwide distribution [@bolton2006sclerotinia].
 This is a necrotrophic pathogen that is primarily homothallic (self-fertilization) and has the ability to survive for more than five years in soil using melanized survival structures called sclerotia [@bolton2006sclerotinia; @sexton2006population].

### 2017-10-31 doc/manuscript/manuscript.Rmd
address L58--67 to a degree, but probably need more finess
alphabetize state names to address #15

@@
 Currently, there are no commercially available resistant cultivars of dry bean [@otto-hanson2011variation].
-Organized breeding efforts have used a common-garden approach with white mold screening nurseries in dry bean production areas across the United States with additional sites in Mexico, France, and Australia [@steadman2003identification; @steadman2004identification; @steadman2005identification; @steadman2006identification; @otto2007identification; @otto2008identification; @mccoy2009use].
+Organized breeding efforts have used a common-garden approach with white mold screening nurseries in dry bean production areas across the United States with additional sites in Australia, France, and Mexico [@steadman2003identification; @steadman2004identification; @steadman2005identification; @steadman2006identification; @otto2007identification; @otto2008identification; @mccoy2009use].
 These white mold screening nurseries use no chemical or cultural treatments against *S. sclerotiorum* and employ standardized protocols for screening new cultivars for resistance to white mold. 
@@
 Because of the genetic connection to vegetative compatibility, two isolates that are mycelially compatible were considered clones [@leslie1993fungal]; but correlation with genetic markers, such as microsatellites, have shown mixed results [@ford1995heterokaryon; @micali2003independence; @jo2008reassessment; @attanayake2012genetic; @papaioannou2014barrage; @lehner2017independently].
+However, a question remains about the relationship between mycelial compatibility groups and clonal genotypes.
 
 In this study, we analyze and characterize the genetic and phenotypic diversity of 366 *S.
-sclerotiorum* isolates collected between 2003 and 2012 from dry bean cultivars among different geographic locations in the United States, France, Mexico, and Australia.
+sclerotiorum* isolates collected between 2003 and 2012 from dry bean cultivars among different geographic locations in the Australia, France, Mexico, and the United States.
 We wanted to know if the *S. sclerotiorum* populations from white mold screening nurseries were representative of the fields within the same region. 
@@
 
-Sclerotia of *S.
-sclerotiorum* were collected over several years from grower fields and/or wmn in 11 states of the United States of America, Mexico, France, and Australia (Table \@ref(tab:isolate-table)).
+Sclerotia of *S. sclerotiorum* were collected over several years from grower fields and/or wmn in 11 states of the Australia, France, Mexico, and the United States (Table \@ref(tab:isolate-table)). 
 After collection, sclerotia were stored in Petri plates lined with filter paper, then stored at $20~^{\circ}$F or -$4~^{\circ}$C. 
@@
 
-A total of 366 isolates were collected from 2003 to 2012 (except 2006 and 2011) from diseased dry bean plants in Australia, Mexico, France, and 11 states in the USA (Table \@ref(tab:isolate-table)).
+A total of 366 isolates were collected from 2003 to 2012 (except 2006 and 2011) from diseased dry bean plants in 11 states in the United States, as well as Australia, France, and Mexico (Table \@ref(tab:isolate-table)).
 We observed a total of 165 MLHs (215 with 16 loci).
@@
 We had performed an analysis on a network where the connections represented shared MLHs across populations, weighted by $1 - P_{sex}$ (Fig. \@ref(community-graph), Table \@ref(tab:mlg-table)). 
-Community analysis of the MLHs shared between populations revealed 4 communities with a modularity of 0.17: A coastal community (CA, OR, WA, and NY), a midwest community (CO, NE, ND, MI), and an international community (MN, France, Australia).
+Community analysis of the MLHs shared between populations revealed 4 communities with a modularity of 0.17: A coastal community (CA, OR, WA, and NY), a midwest community (CO, ND, NE, MI), and an international community (Australia, France, MN).
 Although analysis with 16 loci resulted in the removal of the NY node because it no longer shared a haplotype with OR, the same overall community structure was present with a modularity of 0.2 (Fig.
@@
 Thus, we hypothesize that shared haplotypes between populations may arise due to transmission events of seed or sclerotia.
-This could explain the fact that we see shared genotypes with low $P_{sex}$ values shared between France, Australia, and the US.
+This could explain the fact that we see shared genotypes with low $P_{sex}$ values shared between Australia, France, and the United States.
 While we speculate that these transmission events are rare due to the genetic structuring by Region, these results suggest that seedborne infections may indeed reflect a source of inoculum.
@@
 
-When we tested for sexual reproduction, we were unable to find evidence for it in any region except for CA and Australia. 
+When we tested for sexual reproduction, we were unable to find evidence for it in any region except for Australia and CA. 
 While the Australia population had a non-significant value of $\bar{r}_d$---which would suggest that we cannot reject the null hypothesis of random mating---the sample size was insufficient from which to draw conclusions [@milgroom1996recombination; @agapow2001indices].

### 2017-10-31 doc/manuscript/manuscript.Rmd
remove all hope for #15
address many of the simple line changes for #15

@@
 Knowing what variables predict genetic structure can help direct breeding efforts.
-By investigating these aims, we hope to effectively describe the population structure of *S.
+By investigating these aims, we will effectively describe the population structure of *S.
 sclerotiorum* in the USA and make available our database of isolates for use in future dry bean breeding efforts.
@@
 
-Several of the isolates used for this study were collected as reported in previous studies using the same methods [@otto-hanson2011variation].
+Several (156) of the isolates used for this study were collected as reported in previous studies using the same methods [@otto-hanson2011variation].
 Broadly, isolates were collected from two sources: white mold screening nurseries (wmn) or producer fields.
@@
 
-We used ANOVA to assess if aggressiveness was significantly different with respect to Region, MCG, or multilocus haplotype (MLH). 
+We used ANOVA to assess if aggressiveness (determined via straw test on a scale of 1--9) was significantly different with respect to Region, MCG, or multilocus haplotype (MLH). 
 To minimize complications due to small sample sizes, we chose the top 10 MCGs, representing 56.5% of the isolates collected, the 10 most abundant MLHs representing 26.7% of the isolates, and populations with more than five isolates.
@@
 
-A total of 366 isolates were collected from 2003 to 2012 (except 2006 and 2011) from diseased dry bean plants in 11 states in the United States, as well as Australia, France, and Mexico (Table \@ref(tab:isolate-table)).
-We observed a total of 165 MLHs (215 with 16 loci).
+A total of 366 isolates were collected from 2003 to 2012 (except 2006 and 2011) from diseased dry bean plants in 11 states in the United States as well as Australia, France, and Mexico (Table \@ref(tab:isolate-table)).
+With the 11 loci used in the analyses (Table \@ref(tab:locus-stats)), we observed a total of 165 MLHs (215 with 16 loci).
 Over 50% of the isolates came from four states, MI (`r pop_size[["MI"]]`), ND (`r pop_size[["ND"]]`), WA (`r pop_size[["WA"]]`), NE (`r pop_size[["NE"]]`). 
@@
 
-After clone-correction on the hierarchy of Region/Source/Host/Year, we were left with `r nInd(dat11cc)` isolates representing `r nmll(dat11cc)` MLHs (Table \@ref(tab:iatab)).
+After clone-correction on the hierarchy of Region/Source/Host/Year, a total of `r nInd(dat11) - nInd(dat11cc)` isolates were removed from the data set, resulting in `r nInd(dat11cc)` isolates representing `r nmll(dat11cc)` MLHs that were used in subsequent analyses (Table \@ref(tab:iatab)).
 The results show that most populations exhibited relatively high amounts of genotypic diversity, with the exception of MI and NE, which exhibited lower-than average $E_5$ values.
@@
 ![
-Strip Plot of aggressiveness by population arranged in descending order of mean aggressiveness for all populations with N > 5.
+Strip plot of aggressiveness by population arranged in descending order of mean aggressiveness for all populations with N > 5.
 White bars represent mean value.
@@
 ANOVA showed that the reduced model was significant with an adjusted R^2^ of `r RDA1R2` (*P* = 0.001).
-Assessment of the marginal effects showed that all varaibles significantly explained genetic variation ($P \leq `r anova1res[["Region"]]`$).
+Assessment of the marginal effects showed that all varibles significantly explained genetic variation ($P \leq `r anova1res[["Region"]]`$).
 We found that there was multicollinearity when MCG was combined with any other variable, so repeated the analysis, dropping MCG from the list of potential predictors. 
@@
 
-Discriminant Analysis of Principal Components (DAPC) was performed grouping by Region with the first `r dapc_region$n.pca` principal components, representing `r prettify(dapc_region$var)`% of the total variance.
+Discriminant Analysis of Principal Components (DAPC) was performed by grouping Region with the first `r dapc_region$n.pca` principal components, representing `r prettify(dapc_region$var)`% of the total variance.
 The first discriminant axis (representing `r paxis(dapc_region)`% of the discriminatory power) separated the centroid for the Mexico isolates from the rest of the data, indicating strong differentiation (Fig.

### 2017-10-31 doc/RMD/MLG-distribution.Rmd
add code to export genotype curves

@@
 genotype_curve(dat, sample = 1000, quiet = TRUE)
-genotype_curve(dat11, quiet = TRUE)
+gt16 <- last_plot() + 
+  theme_bw(base_size = 16, base_family = "Helvetica") +
+  stat_smooth() +
+  ggtitle("Haplotype accumulation curve for 16 loci") +
+  ylab("Number of multilocus haplotypes")
+genotype_curve(dat11, sample = 1000, quiet = TRUE)
+gt11 <- last_plot() + 
+  theme_bw(base_size = 16, base_family = "Helvetica") +
+  stat_smooth() +
+  ggtitle("Haplotype accumulation curve for 11 loci") +
+  ylab("Number of multilocus haplotypes")
+```
+
+```{r saveplots}
+if (!interactive()){
+  ggsave(gt16,
+    filename = "results/figures/publication/genotype-curve16.pdf", 
+    width = 88*1.66, 
+    height = 88, 
+    units = "mm", 
+    scale = 1)
+  ggsave(gt11,
+    filename = "results/figures/publication/genotype-curve11.pdf", 
+    width = 88*1.66, 
+    height = 88, 
+    units = "mm", 
+    scale = 1)
+}
 ```

### 2017-11-01 doc/manuscript/manuscript.Rmd
update locus table

@@
 ```{r , echo = FALSE, warning = FALSE, message = FALSE, results = "asis"}
+#    Locus    Range Repeats allele      Hexp  Evenness
+#  1   5-2 318--324 2.00000      4 0.4522344 0.6164147
+#  2   6-2 483--495 5.99999      3 0.6445842 0.9493734
+#  3   7-2 158--174 2.00000      7 0.7290665 0.7635293
+#  4   8-3 244--270 2.00000      7 0.7418669 0.7892240
+#  5   9-2 360--382 2.00000      9 0.3475110 0.4055573
+#  6  12-2 214--222 2.00000      5 0.5804776 0.7791017
+#  7  17-3 342--363 3.00000      7 0.5527510 0.5264941
+#  8  20-3 280--282 2.00000      2 0.0532974 0.4204995
+#  9  55-4 153--216 4.00000     10 0.7231380 0.6564025
+# 10 110-4 370--386 3.99999      5 0.7555955 0.9147773
+# 11 114-4 339--416 4.00000     10 0.8306310 0.8012051
+
 loc_table %>%
@@
   mutate(Repeats = case_when(
-    Repeats <= 2 ~ "di-",
-    Repeats <= 3 ~ "tri-",
-    Repeats <= 4 ~ "tetra-",
-    Repeats <= 5 ~ "penta-",
-    Repeats <= 6 ~ "hexa-"
+    Locus == "5-2"   ~ "(GT)",
+    Locus == "6-2"   ~ "(TTTTTC)(TTTTTG)(TTTTTC)", # "(TTTTTC)~2~(TTTTTG)~2~(TTTTTC)",
+    Locus == "7-2"   ~ "(GA)",
+    Locus == "8-3"   ~ "(CA)",
+    Locus == "9-2"   ~ "(CA)(CT)", # "(CA)~9~(CT)~9~",
+    Locus == "12-2"  ~ "(CA)",
+    Locus == "17-3"  ~ "(TTA)",
+    Locus == "20-3"  ~ "(GT)GG(GT)", # "(GT)~7~GG(GT)~5~",
+    Locus == "55-4"  ~ "(TACA)",
+    Locus == "110-4" ~ "(TATG)",
+    Locus == "114-4" ~ "(TAGA)"
   )) %>%
@@
   huxtable::set_align(huxtable::everywhere, 2:6, "right") %>%
-  huxtable::set_col_width(c(0.06, 0.07, 0.11, 0.1, 0.05, 0.075)) %>%
-  huxtable::print_md(max_width = 61)
+  huxtable::set_col_width(c(0.06, 0.07, 0.25, 0.1, 0.05, 0.075)) %>%
+  huxtable::print_md(max_width = 90)
 ```

### 2017-11-01 doc/RMD/mlg-mcg.Rmd
update visualization a bit

@@
   scale_linetype_manual(values = c(1, 3)) +
+  # scale_alpha_manual(values = c(1, 0.4)) +
   scale_edge_width(range = c(0.25, 1.5)) +
@@
   arrange(desc(nobs)) # arrange by number of samples and reorder factors 
-Severity <- filter(strat, MCG %in% mcgs$MCG[1:5]) %>%
+Severity <- filter(strat, MCG %in% mcgs$MCG[1:8]) %>%
   group_by(MCG) %>% 
@@
              pch = 21) +
-  scale_fill_viridis(discrete = TRUE, direction = -1) +
+  scale_fill_manual(values = c(viridis(5, direction = -1), NA)) +
   theme_bw() +

### 2017-11-02 doc/manuscript/manuscript.Rmd
update visualization a bit
fix P-value format and add citation for protocols
These address #15
I also added Reviewer 2's comment on the introduction

@@
 
+<!-- 
+Reviewer 2:
+I would expand introduction a bit to describe the symptoms in greater 
+details, maybe even provide some information regarding economic damage (if 
+available) to nursery production. I was wondering what cultivars previous 
+studies focused on – the authors provided 3 cultivars at the end of 
+introduction. Maybe adding another paragraph on differences among 
+cultivars – they chose 3 with different levels of resistance but how was 
+that selected (assuming some preliminary data)? For someone not familiar 
+with the host, I would like to have more details on both host and pathogen. 
+-->
 *Sclerotinia sclerotiorum* (Lib.) de Bary is an ascomycete plant pathogen with a worldwide distribution [@bolton2006sclerotinia].
@@
 Organized breeding efforts have used a common-garden approach with white mold screening nurseries in dry bean production areas across the United States with additional sites in Australia, France, and Mexico [@steadman2003identification; @steadman2004identification; @steadman2005identification; @steadman2006identification; @otto2007identification; @otto2008identification; @mccoy2009use].
-These white mold screening nurseries use no chemical or cultural treatments against *S. sclerotiorum* and employ standardized protocols for screening new cultivars for resistance to white mold. 
+These white mold screening nurseries use no chemical or cultural treatments against *S.
+sclerotiorum* and employ standardized protocols for screening new cultivars for resistance to white mold [@steadman2003identification; @otto-hanson2011variation].
 It has previously been shown that aggressiveness (the severity of disease symptoms on the host) is significantly different across white mold screening nursery sites in separate geographic regions [@otto-hanson2011variation].
@@
 
-## Data processing and Analysis{-}
+## Data processing and Analysis {-}
 
@@
   huxtable::set_col_width(c(0.06, 0.07, 0.25, 0.1, 0.05, 0.075)) %>%
-  huxtable::print_md(max_width = 90)
+  huxtable::print_md(max_width = 100)
 ```
@@
 \@ref(aggressiveness-boxplot), Table \@ref(tab:region-aggressiveness)).
-MCGs also showed a significant effect (*P* < 1.00e^-4^), with means that ranged from 6.0 ('MCG 44') to 4.6 ('MCG 49'; Table \@ref(tab:MCG-aggressiveness)).
-We additionally found a significant effect for MLHs (*P* = 7.44e^-4^), with means that ranged from 6.0 ('MLH 78') to 4.3 ('MLH 140') (Table \@ref(tab:MLG-aggressiveness)).
+MCGs also showed a significant effect (*P* < 0.001), with means that ranged from 6.0 ('MCG 44') to 4.6 ('MCG 49'; Table \@ref(tab:MCG-aggressiveness)).
+We additionally found a significant effect for MLHs (*P* < 0.001), with means that ranged from 6.0 ('MLH 78') to 4.3 ('MLH 140') (Table \@ref(tab:MLG-aggressiveness)).
 

### 2017-11-02 doc/RMD/MLG-distribution.Rmd
rescale nodes of Fig 4 and S2 to fix #16
This also addresses #15, although it was not explicitly requested.

@@
     geom_edge_fan(aes(alpha = 1 - weight, width = width + 1)) +
-    geom_node_circle(aes(r = drop(scale(size, center = FALSE))/10, fill = size))
+    geom_node_circle(aes(r = sqrt(size)/30, fill = size))
   outg <- if (label) outg + geom_node_label(aes(label = label), repel = TRUE, parse = TRUE, label.size = 0.75, nudge_x = x_nudge, segment.size = 0, segment.alpha = 0) else outg
@@
 g <- graph_from_adjacency_matrix(matrix(0, 2, 2))
-V(g)$size <- c(56, 56 - 32)
+V(g)$size <- c(43, 43 - 32)
 V(g)$label <- c("Number of MLHs in Region", "Number of private MLHs")
@@
 glegend <- create_layout(g, "manual", node.positions = as.data.frame(lay)) %>% ggraph() + 
-  geom_node_circle(aes(r = size, fill = size)) + 
+  geom_node_circle(aes(r = sqrt(size)*5, fill = size)) + 
   coord_fixed() + 
-  viridis::scale_fill_viridis(option = "C", begin = 1 - 32/56, end = 1, guide = "none") + 
+  viridis::scale_fill_viridis(option = "C", begin = 1 - 32/43, end = 43/56, guide = "none") + 
   ggrepel::geom_text_repel(aes(label = label), 
-                           x = c(20, 0) + 25, 
-                           y = c(35, 0), 
-                           nudge_x = 350, 
+                           x = c(6, -2) + 20, 
+                           y = c(20, -6), 
+                           nudge_x = 150, 
                            segment.size = 1, 
@@
   cowplot::draw_plot(gg16, x = 0, y = 0.05) +
-  cowplot::draw_plot(glegend, x = -0.125, y = 0, height = 0.1)
+  cowplot::draw_plot(glegend, x = -0.0125, y = 0, height = 0.1)
   )
@@
   cowplot::draw_plot(gg11, x = 0, y = 0.05) +
-  cowplot::draw_plot(glegend, x = -0.125, y = 0, height = 0.1)
+  cowplot::draw_plot(glegend, x = -0.0125, y = 0, height = 0.1)
   )
@@
 
-```{r region_dapc}
+```{r region_dapc, cache = TRUE}
 set.seed(2017-07-12)

### 2017-11-02 doc/RMD/wmn-differentiation.Rmd
add DA screeplot to show absolutely nothing

@@
         scree.pca = TRUE, 
-        scree.da = FALSE,
+        scree.da = TRUE,
         col = viridis::plasma(3, end = 0.85),

### 2017-11-02 doc/manuscript/manuscript.Rmd
add link to SOM to address #15

@@
 The analysis of molecular variance (AMOVA) for clone-corrected samples over the hierarchy of Region, Source, and Year showed significant variation between Regions and Years, but no significant variation between wmn and producer fields (Table \@ref(tab:amova)).
-In contrast, when we compared the three cultivars, Beryl, Bunsi, and G122, we found no significant differentiation (Supplementary Information).
+In contrast, when we compared the three cultivars, Beryl, Bunsi, and G122, we found no significant differentiation (Supplementary Information[^1]).
+
+[^1]: https://github.com/everhartlab/sclerotinia-366/blob/v1.0/results/wmn-differentiation.md#host-differentiation
 

### 2017-11-02 doc/manuscript/manuscript.Rmd
add more footnotes to Supplementary Info to address #15

--- 2017-11-02 doc/manuscript/manuscript.Rmd
@@
 Locus `r loc_table$Locus[loc_table$allele==2]`, which contained only 2 alleles, showed low values of both *h* (`r signif(loc_table$Hexp[loc_table$allele==2], 3)`) and evenness (`r signif(loc_table$Evenness[loc_table$allele==2], 3)`), indicating that there was one dominant allele present.
-Analysis of the haplotype accumulation curve showed no clear plateau for 11 or 16 loci (Supplemental Information), indicating that we would likely obtain more multilocus haplotypes if we were to genotype more loci.
+Analysis of the haplotype accumulation curve showed no clear plateau for 11 or 16 loci (Supplementary Information[^1]), indicating that we would likely obtain more multilocus haplotypes if we were to genotype more loci.
+
+[^1]: https://github.com/everhartlab/sclerotinia-366/blob/v1.0/results/MLG-distribution.md#loading-data-and-setting-strata
 
@@
 The analysis of molecular variance (AMOVA) for clone-corrected samples over the hierarchy of Region, Source, and Year showed significant variation between Regions and Years, but no significant variation between wmn and producer fields (Table \@ref(tab:amova)).
-In contrast, when we compared the three cultivars, Beryl, Bunsi, and G122, we found no significant differentiation (Supplementary Information).
+In contrast, when we compared the three cultivars, Beryl, Bunsi, and G122, we found no significant differentiation (Supplementary Information[^2]).
+
+[^2]: https://github.com/everhartlab/sclerotinia-366/blob/v1.0/results/wmn-differentiation.md#host-differentiation
 

### 2017-11-02 doc/RMD/MLG-distribution.Rmd
remove saving of genotype curves
Since I pointed to the resource in a footnote, it represents nothing
more than redundancy.

@@
   ylab("Number of multilocus haplotypes")
-```
-
-```{r saveplots}
-if (!interactive()){
-  ggsave(gt16,
-    filename = "results/figures/publication/genotype-curve16.pdf", 
-    width = 88*1.66, 
-    height = 88, 
-    units = "mm", 
-    scale = 1)
-  ggsave(gt11,
-    filename = "results/figures/publication/genotype-curve11.pdf", 
-    width = 88*1.66, 
-    height = 88, 
-    units = "mm", 
-    scale = 1)
-}
 ```

### 2017-11-02 doc/manuscript/manuscript.Rmd
attempt to find solution to #17

--- 2017-11-02 doc/manuscript/manuscript.Rmd
@@
 Locus `r loc_table$Locus[loc_table$allele==2]`, which contained only 2 alleles, showed low values of both *h* (`r signif(loc_table$Hexp[loc_table$allele==2], 3)`) and evenness (`r signif(loc_table$Evenness[loc_table$allele==2], 3)`), indicating that there was one dominant allele present.
-Analysis of the haplotype accumulation curve showed no clear plateau for 11 or 16 loci (Supplemental Information), indicating that we would likely obtain more multilocus haplotypes if we were to genotype more loci.
+Analysis of the haplotype accumulation curve showed no clear plateau for 11 or 16 loci (See section on 'Loading Data and Setting Strata' in the MLG-distribution.md[^1] file in the supplemental files [@kamvar2017data]), indicating that we would likely obtain more multilocus haplotypes if we were to genotype more loci.
+
+[^1]: Direct link: https://github.com/everhartlab/sclerotinia-366/blob/master/results/MLG-distribution.md#loading-data-and-setting-strata
 
@@
 The analysis of molecular variance (AMOVA) for clone-corrected samples over the hierarchy of Region, Source, and Year showed significant variation between Regions and Years, but no significant variation between wmn and producer fields (Table \@ref(tab:amova)).
-In contrast, when we compared the three cultivars, Beryl, Bunsi, and G122, we found no significant differentiation (Supplementary Information).
+In contrast, when we compared the three cultivars, Beryl, Bunsi, and G122, we found no significant differentiation (See section on 'Host Differentiation' in the wmn-differentiation.md[^2] file in the supplemental files [@kamvar2017data]).
+
+[^2]: Direct link: https://github.com/everhartlab/sclerotinia-366/blob/master/results/wmn-differentiation.md#host-differentiation
 
@@
 
-All scripts, data, and resources used to generate the results presented in this publication (including Supplementary Information) are fully reproducible and available at The Open Science Framework https://osf.io/ejb5y.
+All scripts, data, and resources used to generate the results presented in this publication (including Supplementary Information) are fully reproducible and available at The Open Science Framework https://osf.io/ejb5y [@kamvar2017data].
 

### 2017-11-03 doc/manuscript/manuscript.Rmd
attempt to address #15, Line 251
There's still something weird about attempting to describe a
genotypic diversity table. It feels laundry-list-y

@@
 After clone-correction on the hierarchy of Region/Source/Host/Year, a total of `r nInd(dat11) - nInd(dat11cc)` isolates were removed from the data set, resulting in `r nInd(dat11cc)` isolates representing `r nmll(dat11cc)` MLHs that were used in subsequent analyses (Table \@ref(tab:iatab)).
-The results show that most populations exhibited relatively high amounts of genotypic diversity, with the exception of MI and NE, which exhibited lower-than average $E_5$ values.
-Mexico had the lowest average value of gene diversity, *h* and the lowest *eMLH*, suggesting low overall genetic diversity. 
+<!-- 
+The fact that genotypic diversity in poppr uses H, G and lambda should be clarified here.
+It should reflect the results interpretation as well.
+According to Table 2, MI and NE subpopulations exhibit high genotypic diversity when compared to other subpopulations (3.6, 29 and 0.97 for H, G and lambda, respectively for MI and 3.2, 17.8 and 0.94 for NE).
+If this was based on evenness, that should be explained as well.
+It can be confusing stated like this.
+When you shift focus onto diversity (h), the story changes – so please clarify this section to reflect that
+-->
+The results showed that, in terms of genotypic diversity (*H*, *G*, and $\lambda$), WA was the most diverse population with both *G* (`r signif(setNames(gdt$G, gdt$Pop)["WA"], 3)`) and $e^H$ (`r signif(setNames(exp(gdt$H), gdt$Pop)["WA"], 3)`) being close to the observed number of MLGs (`r setNames(gdt$MLG, gdt$Pop)["WA"]`).
+This indicated that there are few duplicated genotypes in WA (Table \@ref(tab:iatab)).
+A more useful metric to compare populations, however, is $E_5$, which scales from 0 to 1 where 1 indicates all unique genotypes [@grunwald2003analysis]. 
+Evaluating by $E_5$ shows that both MI and NE exhibit lower than average values, indicating that there are over-represented genotypes in the popualtions (table \@ref(tab:iatab)). 
+When we look at Mexico, we can observed that it had relatively high values of $E_5$ and genotypic diversity, but low richness, as measured by *eMLG*.
+Moreover, Mexico had the lowest value for *h*, which is a measure of allelic diversity.
+<!-- The results show that most populations exhibited relatively high amounts of genotypic diversity, with the exception of MI and NE, which have which exhibited lower-than average $E_5$ values. -->
+<!-- Mexico had the lowest average value of gene diversity, *h* and the lowest *eMLH*, suggesting low overall genetic diversity.  -->
 Nearly all populations showed evidence for linkage (Table \@ref(tab:iatab)), which serves as evidence for clonal reproduction or other forms of non-random mating. 

### 2017-11-03 doc/manuscript/manuscript.Rmd
add note on additional phenotypes for #15
add information about the chromosomal coverage

@@
 With the 11 loci used in the analyses (Table \@ref(tab:locus-stats)), we observed a total of 165 MLHs (215 with 16 loci).
+These 11 loci represented 7 chromosomes in the *S. sclerotiorum* genome with a minimum distance of 55Kbp between two loci on the same chromosome.
 Over 50% of the isolates came from four states, MI (`r pop_size[["MI"]]`), ND (`r pop_size[["ND"]]`), WA (`r pop_size[["WA"]]`), NE (`r pop_size[["NE"]]`). 
@@
 If the distribution of *S. sclerotiorum* is even across agricultural hosts in the USA, then our sample may yet be representative of the genetic pool present in other crops and weedy species. 
+Additionally, while we found no signficant association between genotype and aggressiveness, it is important to note that the straw test is only one measure of aggressiveness. 
+Additional phenotypes for aggressiveness should be evaluated for future studieds.
 

### 2017-11-03 doc/RMD/mlg-mcg.Rmd
add new supplementary figure for #15

@@
 
-There are a whole buttload of MLGs for those 5 MCGs. What are the severity 
+There are a whole boatload of MLGs for those 8 MCGs. What are the severity 
 ratings for those? Again, we take take a look at these AND simultaneously query
@@
 
-```{r severity, fig.width = 7, fig.height = 4}
+```{r severity, fig.width = 7, fig.height = 4.5}
 count_group <- . %>% 
@@
   mutate(MCG = forcats::fct_inorder(factor(MCG), ordered = TRUE)) %>%
-  mutate(MLG = ifelse(MLG %in% mlgs$MLG[1:5], paste("MLG", MLG), "Other"))
+  mutate(MLG = ifelse(MLG %in% mlgs$MLG[1:5], paste("MLH", MLG), "Other"))
 
@@
   scale_fill_manual(values = c(viridis(5, direction = -1), NA)) +
-  theme_bw() +
+  theme_bw(base_size = 16, base_family = "Helvetica") +
   theme(legend.position = "bottom") +
-  theme(aspect.ratio = 0.6) +
-  facet_wrap(~Region, nrow = 2) +
-  ylim(c(3.5, 8)) +
+  theme(aspect.ratio = 0.5) +
+  facet_wrap(~Region, nrow = 3) +
+  ylim(c(1, 9)) +
   labs(list(
-    title = "Severity by MCG and Region",
-    fill = "Multilocus Haplotype",
-    subtitle = "Five most abundant multilocus genotypes shown"
-  ))
+    y = "Aggressiveness",
+    #title = "Aggressiveness by MCG and Region",
+    fill = "Multilocus Haplotype"
+    #subtitle = "Five most abundant multilocus haplotypes shown"
+  )) +
+  theme(legend.position = c(0.75, 0.1)) +
+  theme(legend.title.align = 0.5) +
+  guides(fill = guide_legend(nrow = 2))
 severity_plot
 if (!interactive())
-  ggsave(severity_plot, filename = "results/figures/publication/FigureS3.pdf", width = 183, unit = "mm")
+  ggsave(severity_plot, filename = "results/figures/publication/RMM-aggressiveness.pdf", width = 183, height = 183*(4.5/7), unit = "mm")
 ```

### 2017-11-03 doc/manuscript/manuscript.Rmd
add new supplementary figure for #15

@@
 The group mean averages were `r sprintf("%0.2f, %0.2f, and %0.2f", Raggmean, Maggmean, Haggmean)` for Region, MCG, and MLH, respectively.
+A strip plot showing the distribution of severity across these three variables simultaneously can be seen in figure \@ref(mcg-mlh-region-aggressiveness).
 Our assessment of aggressiveness in association with Region showed a significant effect (*P* < 1.00e^-4^), with means that ranged from 5.8 (MN) to 4.0 (CA) (Fig.
@@
 
+<!-- Strip plot of Aggressiveness for MLH, MCG, and Region -->
+
+![
+Strip plot of aggressiveness for the eight most abundant MCGs partitioned by region.
+Filled circles indicate one of the five most abundant MLHs and open circles indicate a MLH of lesser abundance.
+](../../results/figures/publication/RMM-aggressiveness.pdf){#mcg-mlh-region-aggressiveness width=100%}
+
+\newpage
+
 ![

### 2017-11-06 doc/manuscript/manuscript.Rmd
add note about MX population
This addresses #15. I still am not exactly sure how to
treat this issue. I think some time away from this document may
help with it.
add figure for MCG delination

@@
 
+<!-- Include discussion on symptoms and economic damage leading in to cultivar discussion? -->
+
 Currently, there are no commercially available resistant cultivars of dry bean [@otto-hanson2011variation].
@@
 Nearly all population genetic studies of *S.
-sclerotiorum* employ a macroscopic assay to determine mycelial compatibility, the ability for fungal hyphae from different colonies to appear to grow together without forming a barrier of dead cells between them (known as a barrage line) [@leslie1993fungal; @sirjusingh2001characterisation].
-Mycelial compatibility has been used as a proxy for vegetative compatibility, a fungal trait controlled by several independent genes controlling the ability for two hyphae to fuse and grow as a single unit [@leslie1993fungal; @schafer2006optimized].
+sclerotiorum* employ a macroscopic assay to determine mycelial compatibility, the ability for fungal hyphae from different colonies to appear to grow together without forming a barrier of dead cells between them (known as a barrage line, Fig.
+\@ref(mcg-fig)B) [@leslie1993fungal; @sirjusingh2001characterisation].
+Mycelial compatibility has been used as a proxy for vegetative compatibility, a fungal trait controlled by several independent genes controlling the ability for two hyphae to fuse and grow as a single unit (Fig.
+\@ref(mcg-fig)A) [@leslie1993fungal; @schafer2006optimized].
 Because of the genetic connection to vegetative compatibility, two isolates that are mycelially compatible were considered clones [@leslie1993fungal]; but correlation with genetic markers, such as microsatellites, have shown mixed results [@ford1995heterokaryon; @micali2003independence; @jo2008reassessment; @attanayake2012genetic; @papaioannou2014barrage; @lehner2017independently].
@@
 This may, in turn increase the risk of introducing new sources of genetic variation through potential outcrossing events.
+<!-- I was also interested in more interpretation/analysis as to why the Mexican population did not have MLH shared with any other regions, the absence of clarity on this is done at a loss. -->
 
@@
 
+![
+Example of MCG test plates showing (A) a compatible reaction with mycelia from two strains overgrowing each other and (B) an incompatible reaction with a barrage line of dead tissue forming between the two strains.
+Photo Credit: Rebecca Higgens.
+](images/mcg-trimmed.png){#mcg-fig width=100%}
+
 <!-- Strip plot of Aggressiveness for MLH, MCG, and Region -->
@@
 ](../../results/figures/publication/RMM-aggressiveness.pdf){#mcg-mlh-region-aggressiveness width=100%}
-
-\newpage
 

### 2017-11-06 doc/manuscript/manuscript.Rmd
attempt to add discussion to intro for #15

@@
 It causes disease on more than 400 plant species belonging to 75 families [@boland1994index] including crops of major economic importance such as sunflower (*Helianthus spp.*), soybean (*Glycine max* L.), canola (*Brassica napa* L., *Brassica campestris* L.), and dry bean (*Phaseolus vulgaris* L.) [@bolton2006sclerotinia].
+
 On dry bean, *S. sclerotiorum* is the causal agent of white mold, a devastating disease that can be yield-limiting in temperate climates [@steadman1983white]. 
-
-<!-- Include discussion on symptoms and economic damage leading in to cultivar discussion? -->
+All above-ground tissues (flowers, stems, leaves, pods) are susceptible to infection, first appearing as wet lesions with white mycelial tufts, and then bleaching as the tissue senesces [@steadman1983white; @bolton2006sclerotinia].
+For many years, white mold has been the most serious dry bean disease in the Northwestern United States [@otto-hanson2011variation; @knodel2012dry; @knodel2015dry; @knodel2016dry].
+The impact of white mold on the dry bean industry in the Northwestern United States alone has been estimated at a loss of 140 kg/ha with just 10% disease incidence [@ramasubramaniam2008estimates].
 
@@
 sclerotiorum* and employ standardized protocols for screening new cultivars for resistance to white mold [@steadman2003identification; @otto-hanson2011variation].
+These protocols included three established cultivars used for comparison in the trials: Beryl (great northern bean, susceptible), Bunsi (aka Ex Rico, navy bean, low susceptibility), and G122 (cranberry bean, partial resistance) [@tu1982tolerance; @steadman2005identification; @otto-hanson2011variation].
 It has previously been shown that aggressiveness (the severity of disease symptoms on the host) is significantly different across white mold screening nursery sites in separate geographic regions [@otto-hanson2011variation].
@@
 
-MCG was determined as described previously through co-culturing pairs of 2-day-old isolates 2.5 cm apart on Diana Sermons (DS) Medium [@cubeta2001mycelial]. 
+MCG was determined as described previously through co-culturing pairs of 2-day-old isolates 2.5 cm apart on Diana Sermons (DS) Medium (Fig. \@ref(mcg-fig)) [@cubeta2001mycelial]. 
 Incompatibility of different MCGs resulted in formation of a barrage line accompanied by formation of sclerotia on either side of the barrage line, indicating the limits of the isolates' growth [@otto-hanson2011variation; @kohn1990mycelial; @leslie1993fungal].

### 2017-11-07 doc/manuscript/manuscript.Rmd
DAPC does not operate on Euclidian distances
WTH was I thinking when I wrote that?!?! :flushed:

@@
 \@ref(DAPC-RY)).
-DAPC is based on the principal components of Euclidean distance between genotypes [@jombart2010discriminant]. 
-Unlike Bruvo's distance, Euclidean distance does not take into account the magnitude of the difference between alleles, which could inflate the distance measure in the presence of private alleles [@bruvo2004simple].
+DAPC is based on the principal components of allele counts [@jombart2010discriminant]. 
+Unlike Bruvo's distance, this does not take into account the magnitude of the difference between alleles, which could inflate the distance measure in the presence of private alleles [@bruvo2004simple].
 We found no evidence of private alleles in the Mexico and CA isolates, suggesting that this pattern separating these populations from the others was not an artifact.

### 2017-11-07 doc/manuscript/manuscript.Rmd
fix error in calculation of axis contribution
Yet another thing that didn't make sense. Why would two out of two
discriminant axes only represent 38.5% of the discriminatory power?
(See line 323). I fixed the calculation of the percentage of axis
contribution by not taking the inverse of the cumulative sum.
address Mexico discussion for #15

@@
 paxis <- function(d, i = 1){
-  res <- cumsum(d$eig)/sum(d$eig)
-  prettify(1 - res, i)
+  res <- d$eig/sum(d$eig)
+  prettify(res, i)
 }
@@
 Haggmean <- mean(as.numeric(MLHagg[[2]]))
+
+# Additional Mexico analyses
+# 
+# We have no private alleles in Mexico, but the DAPC is being driven by near-private
+# alleles. We can confirm this by inspecting both the DAPC loadings and inspecting
+# these loci
+
+first_axis <- dapc_region$var.contr[, 1]
+q95 <- quantile(first_axis, 0.95)
+poptab <- dat11 %>% 
+  genind2genpop(quiet = TRUE) %>% 
+  tab()
+near_private <- apply(poptab, 2, function(i) names(i)[which((i/sum(i)) > 0.75)])
+near_private <- unlist(near_private[lengths(near_private) == 1])
+near_private <- near_private[near_private == "Mexico"]
+fa_alleles   <- poptab[, first_axis > q95]
+mx_percent   <- t(fa_alleles[, names(near_private)])/colSums(fa_alleles)
+mx_percent   <- prettify(mx_percent[, "Mexico", drop = TRUE], TRUE)
+mx_loci      <- substr(names(mx_percent), 1, 3)
+mx_alleles   <- substr(names(mx_percent), 8, 11)
+mx_allele_string <- glue::glue("alleles {mx_alleles[1]}, {mx_alleles[2]}, and {mx_alleles[3]} in loci {mx_loci[1]}, {mx_loci[2]}, and {mx_loci[3]}")
 ```
@@
 Unlike Bruvo's distance, this does not take into account the magnitude of the difference between alleles, which could inflate the distance measure in the presence of private alleles [@bruvo2004simple].
-We found no evidence of private alleles in the Mexico and CA isolates, suggesting that this pattern separating these populations from the others was not an artifact.
+While we found no evidence of private alleles in the Mexico and CA isolates, we did find that the alleles driving the first axis in figure \@ref(DAPC)A (`r mx_allele_string`, respectively) were overrepresented in Mexico (where >75% of the alleles came from the region).
+However, all three of these alleles i) conform to the expected stepwise mutation model [@bruvo2004simple] and ii) are at or near the extremes of the total range (except for `r glue::glue("allele {mx_alleles[3]} at locus {mx_loci[3]}")`).
+Moreover, the fact that we find three alleles at three independent loci segregating for Mexico suggests that the pattern separating these populations from the others was not an artifact. 
 We believe that the differences we observe from Mexico may reflect differences in climate.
@@
 It has been proposed within the *S.
-sclerotiorum* literature that isolates from sub-tropical and tropical climates are differentiated or more variable than populations from temperate climates [@carbone2001multilocus; @attanayake2013sclerotinia; @lehner2017sclerotinia;].
+sclerotiorum* literature that isolates from sub-tropical and tropical climates are differentiated or more variable than populations from temperate climates [@carbone2001multilocus; @attanayake2013sclerotinia; @lehner2017sclerotinia].
 This has been attributed to the notion that the fungus has the chance to undergo more reproductive cycles in the warmer climate [@carbone2001multilocus; @attanayake2013sclerotinia]. 

### 2017-11-07 doc/RMD/compare-aldrich-wolfe.Rmd
add aldrich-wolfe comparison for #15

---
title: Re-analysis of *Sclerotinia sclerotiorum* data
author: Zhian N. Kamvar
output: 
  html_notebook:
    toc: true
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = PROJHOME)
```

# Introduction

In the PLoS paper "Genetic Variation of *Sclerotinia sclerotiorum* from
Multiple Crops in the North Central United States" (DOI:
10.1371/journal.pone.0139188), Aldrich-Wolfe *et al.* show that there is more
genetic differentiation within host than among hosts in *S. sclerotiorum*.

It is of interest here to determine if the data from the present study is similar
to the data in their study.

# Packages

I'm going to use *poppr* to analyze the data, *dplyr* to merge the data, and
*tidyr* to unnest individuals.

```{r packages, warning = FALSE, message = FALSE}
library("poppr")
library("tidyverse")
```

# Data Preparation

Here, I'm going to re-analyze their data. They did not provide an easily
accessible format, so I was forced to copy the data from the paper.

> A note on cleaning, The following data was corrected from soybean to sunflower:
> 
> ```
> 3;3a;2008;North Dakota;Helianthus annuus;soybean;4;129,134,180,189
> ```

```{r read data}
haplotypes <- "Haplotype	7_2	8_3	110_4	55_4	13_2	23_4	7_3	5_2	17_3	12_2	92_4	106_4
1	188	268	382	189	311	407	225	332	359	238	388	572
2a	188	270	397	189	322	407	220	332	359	232	388	572
2b	188	270	397	189	322	407	220	332	359	232	388	576
3a	186	268	390	189	332	407	225	332	353	238	388	577
3b	186	268	394	189	332	407	225	332	353	238	388	577
4	186	268	382	189	311	407	222	332	353	234	388	572
6a	186	268	397	189	332	407	222	332	353	234	388	568
6b	186	268	397	189	332	407	222	332	353	234	388	593
7	188	270	390	189	322	407	225	332	353	238	390	577
8a	188	270	390	189	322	407	220	332	359	232	390	581
8b	188	270	390	189	322	407	220	332	359	232	390	585
8c	188	270	390	189	322	407	220	332	365	232	390	605
9	188	268	390	189	332	407	225	334	359	238	391	568
11	188	268	393	189	332	407	220	332	359	232	391	554
12a	176	262	386	205	332	405	220	332	365	232	388	574
12b	176	262	386	205	332	405	220	332	365	232	388	578
13	188	272	382	193	322	407	222	332	359	234	388	565
14	188	272	393	189	332	407	225	332	361	238	390	550
15	186	272	382	173	322	407	222	332	359	234	388	554
16	186	270	397	189	332	407	222	332	353	234	388	585
17	188	270	397	189	332	407	222	332	353	234	388	574
18	186	268	382	240	332	407	225	332	359	238	388	585
19	186	270	397	181	332	407	220	334	359	232	391	550
20	188	270	382	181	322	407	222	332	353	234	391	543
21	186	270	397	189	322	407	220	332	359	232	388	572
23	186	268	397	173	311	405	220	332	359	232	388	589
24	188	270	397	189	322	407	220	332	359	232	388	572
25a	176	262	386	232	332	405	220	332	375	232	388	598
25b	176	262	386	232	332	405	220	332	375	232	388	589
28a	186	270	390	189	338	407	225	332	359	238	388	572
28b	186	270	390	189	342	407	225	332	359	238	388	572
30	186	270	382	181	322	407	225	332	353	238	388	562
31	188	268	390	181	322	407	225	334	353	238	388	572
33	188	270	390	193	322	407	220	332	359	232	390	585
34	186	268	397	177	347	407	225	332	359	238	388	554
36	186	270	394	193	322	407	222	332	353	234	388	558
37	188	268	382	193	311	407	225	332	361	238	390	572
38	186	268	382	181	311	407	222	332	359	234	390	581
43	186	270	390	185	332	407	222	332	353	234	391	585
46	188	268	382	193	322	407	225	332	359	238	391	546
53	188	268	397	181	374	407	225	332	359	238	388	572
57	176	262	386	185	332	405	220	332	367	232	388	535
60	188	268	397	181	332	407	220	334	359	232	391	568
61	188	268	397	181	332	407	225	332	359	238	388	554
62	186	268	382	173	311	405	220	332	371	232	390	577
64	186	270	397	185	332	407	222	332	353	234	388	562
66	188	270	390	189	332	407	220	332	353	232	390	581
69	188	270	382	189	311	407	225	332	378	238	395	574
71	176	262	386	228	332	405	220	332	367	232	388	535
73	186	268	382	181	311	407	220	332	353	232	390	581
77	188	268	397	181	322	407	222	332	353	234	388	565
"

haplotypes <- readr::read_table2(haplotypes)

replens <- c(
"5_2" = 2,
"7_2" = 2,
"7_3" = 2,
"8_3" = 2,
"12_2" = 2,
"13_2" = 5,
"17_3" = 3,
"23_4" = 2,
"55_4" = 4,
"92_4" = 2,
"106_4" = 4,
"110_4" = 4
)

isolates <- "MCG;Haplotype;Year;State/Province;Host;Host common name;Number of isolates;Isolate designation
1;1;2007;North Dakota;Brassica napus;canola;1;100
2;2a;2008;Minnesota;Phaseolus vulgaris;dry bean;1;137
2;2a;2008;Nebraska;Glycine max;soybean;1;191
2;2a;2008;North Dakota;Brassica napus;canola;2;175,181
2;2a;2008;North Dakota;Phaseolus vulgaris;dry bean;2;101,122
2;2a;2008;North Dakota;Glycine max;soybean;1;132
2;2a;2008;South Dakota;Glycine max;soybean;1;201
2;2a;2007;North Dakota;Brassica napus;canola;1;800
2;2a;1996;Colorado;Phaseolus vulgaris;dry bean;1;196
2;2a;2008;Manitoba, Canada;Phaseolus vulgaris;dry bean;1;103
3;3a;2008;North Dakota;Brassica napus;canola;1;159
3;3a;2008;North Dakota;Phaseolus vulgaris;dry bean;1;102
3;3a;2007;North Dakota;Brassica napus;canola;2;700,900
3;3a;-;Minnesota;Daucus carota;carrot;1;205
3;3a;2008;North Dakota;Helianthus annuus;sunflower;4;129,134,180,189
3;3b;2008;North Dakota;Phaseolus vulgaris;dry bean;1;156
3;3b;2008;North Dakota;Glycine max;soybean;1;240
3;3b;2008;North Dakota;Helianthus annuus;sunflower;1;188
4;4;2008;North Dakota;Phaseolus vulgaris;dry bean;1;104
5;2a;2008;North Dakota;Phaseolus vulgaris;dry bean;1;105
5;2a;2008;North Dakota;Glycine max;soybean;1;116
6;6a;2008;Minnesota;Helianthus annuus;sunflower;2;136,139
6;6a;2008;North Dakota;Phaseolus vulgaris;dry bean;1;106
6;6b;2008;North Dakota;Glycine max;soybean;1;176
7;7;2008;North Dakota;Glycine max;soybean;1;107
8;8a;2008;Iowa;Glycine max;soybean;1;213
8;8a;2008;North Dakota;Brassica napus;canola;2;167,186
8;8a;2008;North Dakota;Phaseolus vulgaris;dry bean;4;109,115,119,133
8;8b;2008;North Dakota;Phaseolus vulgaris;dry bean;1;120
8;8b;2007;Montana;Carthamus tinctorius;safflower;1;244
8;8b;2006;Minnesota;Glycine max;soybean;1;203
8;8c;2008;North Dakota;Glycine max;soybean;1;239
8;8c;2008;North Dakota;Helianthus annuus;sunflower;1;183
9;9;2008;Illinois;Glycine max;soybean;1;148
9;9;2008;Indiana;Glycine max;soybean;2;206,207
9;9;2008;Iowa;Glycine max;soybean;6;211,212,214,215,216,217
9;9;2008;Minnesota;Glycine max;soybean;4;145,146,197,204
9;9;2008;Minnesota;Helianthus annuus;sunflower;1;141
9;9;2008;Nebraska;Glycine max;soybean;1;153
9;9;2008;North Dakota;Brassica napus;canola;1;150
9;9;2008;North Dakota;Phaseolus vulgaris;dry bean;7;108,117,124,131,151,157,184
9;9;2008;North Dakota;Glycine max;soybean;5;113,114,125,135,178
9;9;2008;North Dakota;Helianthus annuus;sunflower;2;126,128
9;9;2007;Illinois;Glycine max;soybean;1;147
9;9;2007;North Dakota;Brassica napus;canola;1;110
9;9;2004;Missouri;Glycine max;soybean;1;234
9;9;2004;Wisconsin;Nicotiana tabacum;tobacco;1;233
9;9;2003;Wisconsin;Nicotiana tabacum;tobacco;1;232
9;9;2002;Ohio;Glycine max;soybean;1;210
9;9;2002;Wisconsin;Glycine max;soybean;1;226
9;9;2000;Wisconsin;Glycine max;soybean;4;220,221,223,224
11;11;1987;Colorado;Solanum tuberosum;potato;1;198
12;12a;2008;Minnesota;Glycine max;soybean;1;140
12;12a;2008;North Dakota;Helianthus annuus;sunflower;1;123
12;12b;2008;North Dakota;Phaseolus vulgaris;dry bean;1;118
12;12b;2008;North Dakota;Glycine max;soybean;1;127
12;12b;2006;Colorado;Helianthus annuus;sunflower;1;193
13;13;2008;North Dakota;Brassica napus;canola;1;168
14;14;2007;North Dakota;Brassica napus;canola;1;111
15;15;2008;Minnesota;Brassica napus;canola;1;190
15;15;2008;North Dakota;Brassica napus;canola;2;171,177
15;15;2007;North Dakota;Brassica napus;canola;1;112
15;15;1992;Colorado;Brassica napus;canola;1;192
16;16;2008;North Dakota;Glycine max;soybean;1;130
17;17;2008;North Dakota;Brassica napus;canola;1;185
18;18;2008;Minnesota;Phaseolus vulgaris;dry bean;1;138
19;19;2008;Minnesota;Glycine max;soybean;1;142
19;19;2003;Wisconsin;Glycine max;soybean;1;229
20;20;2008;Minnesota;Helianthus annuus;sunflower;1;143
21;21;2008;Minnesota;Glycine max;soybean;1;144
23;23;2008;North Dakota;Helianthus annuus;sunflower;1;149
24;24;2008;Nebraska;Phaseolus vulgaris;dry bean;1;154
25;25a;2008;Kansas;Helianthus annuus;sunflower;1;246
25;25a;2008;North Dakota;Phaseolus vulgaris;dry bean;1;121
25;25a;2008;North Dakota;Helianthus annuus;sunflower;1;155
25;25b;2008;North Dakota;Glycine max;soybean;1;241
28;28a;2008;North Dakota;Phaseolus vulgaris;dry bean;1;158
28;28a;2008;North Dakota;Helianthus annuus;sunflower;2;160,161
28;28a;2007;North Dakota;Brassica napus;canola;1;500
28;28b;2008;North Dakota;Helianthus annuus;sunflower;1;166
30;30;2008;North Dakota;Helianthus annuus;sunflower;1;163
31;31;2008;North Dakota;Brassica napus;canola;1;165
33;33;2008;North Dakota;Phaseolus vulgaris;dry bean;1;169
34;34;2008;North Dakota;Brassica napus;canola;1;170
36;36;2008;North Dakota;Phaseolus vulgaris;dry bean;1;172
37;37;2008;North Dakota;Brassica napus;canola;1;173
38;38;2008;North Dakota;Helianthus annuus;sunflower;2;174,182
43;43;2008;North Dakota;Glycine max;soybean;1;179
46;46;2008;North Dakota;Helianthus anuus;sunflower;1;187
49;12a;2008;Nebraska;Phaseolus vulgaris;dry bean;1;195
52;23;2008;Wyoming;Phaseolus vulgaris;dry bean;1;200
53;53;2008;South Dakota;Helianthus annuus;sunflower;1;202
56;12a;1997;Iowa;Glycine max;soybean;1;208
57;57;1998;Illinois;Glycine max;soybean;1;209
60;60;2008;Michigan;Glycine max;soybean;2;218,219
61;61;2002;North Dakota;Brassica napus;canola;1;230
61;61;2000;Wisconsin;Glycine max;soybean;1;222
62;62;2003;Wisconsin;Glycine max;soybean;2;227,228
63;24;2008;Wyoming;Phaseolus vulgaris;dry bean;1;199
64;64;2002;Wisconsin;Glycine max;soybean;1;225
66;66;2002;North Dakota;Brassica napus;canola;1;231
69;69;2005;Montana;Cynoglossum officinale;houndstonque;1;243
71;71;2008;Kansas;Phaseolus vulgaris;dry bean;1;247
73;73;2007;North Dakota;Brassica napus;canola;1;600
77;77;2008;North Dakota;Brassica napus;canola;1;162
78;19;2008;North Dakota;Brassica napus;canola;1;164"

isocol <- readr::cols(
  MCG = col_integer(),
  Haplotype = col_character(),
  Year = col_character(),
  `State/Province` = col_character(),
  Host = col_character(),
  `Host common name` = col_character(),
  `Number of isolates` = col_integer(),
  `Isolate designation` = col_character()
)

isolates <- readr::read_delim(isolates, ";", col_types = isocol)
```

Now that I have the data read in to data frames, I can create a master data
frame by joining them together.

```{r }
ssc <- left_join(isolates, haplotypes, by = "Haplotype") %>%
	mutate(`Isolate designation` = strsplit(`Isolate designation`, ",")) %>%
	tbl_df() %>%
	unnest() %>%
	select(`Isolate designation`, everything())
ssc
```

Now we can read the data into poppr

```{r }
gid <- df2genind(ssc[-c(1:8)], ind.names = ssc$`Isolate designation`,
                 strata = ssc[2:8], ploidy = 1, pop = ssc$`Host common name`) %>%
  as.genclone()
gid
```

# Comparison

First, I need to load our data:

```{r}
load(file.path(PROJHOME, "data/sclerotinia_16_loci.rda"))
```

Now, I need to make sure that the locus names between their data and our data are cromulent:

```{r}
locNames(dat11)
locNames(gid)
```

They are not, so I will have to massage them a bit to make sure that we can match up the loci and then subset them to only the common loci.

```{r}
ours <- dat11
locNames(gid) <- gsub("_", "-", locNames(gid))
locNames(ours) <- gsub("[(FH)]", "", locNames(ours))
locNames(gid)
locNames(ours)

(AW <- gid[loc = locNames(gid) %in% locNames(ours), mlg.reset = TRUE])
(ZK <- ours[loc = locNames(ours) %in% locNames(gid), mlg.reset = TRUE])
```

So, out of the 11 loci in this study, only 7 match up with Aldrich-Wolfe 2015. This reduces us to 140 MLGs. 
One quick way of assessing whether or not there are any shared genotypes in this data set is to look at the
allele sizes.

```{r}
sort_alleles <- . %>% as.integer() %>% sort()
allist <- list(
  AW = map(alleles(AW), { . %>% sort_alleles } ),
  ZK = map(alleles(ZK), { . %>% sort_alleles } )
)
purrr::transpose(allist)

# Do any of the alleles match?
(matching_alleles <- map_lgl(purrr::transpose(allist), ~{ any(.$AW %in% .$ZK) }))
```

Out of this, we only have `r sum(matching_alleles)` matches, which means that no genotypes are shared.

<details>
<summary>Session Information</summary>

```{r, echo = FALSE}
options(width = 100)
devtools::session_info()
```

</details>

### 2017-11-07 doc/manuscript/manuscript.Rmd
add WA field and Aldrich-Wolfe comparison for #15

@@
 
-# Additional Mexico analyses
+# Additional Mexico analyses: Reviewer 1 request
 # 
@@
 mx_allele_string <- glue::glue("alleles {mx_alleles[1]}, {mx_alleles[2]}, and {mx_alleles[3]} in loci {mx_loci[1]}, {mx_loci[2]}, and {mx_loci[3]}")
+
+# Washington aggressiveness: Reviewer 3 request
+#
+# Reviewer 3 wanted to see an assessment of the agressiveness difference between the two fields in
+# WA white mold nurseries.
+
+WA <- bind_cols(strata(dat11), other(dat11)$meta) %>%
+  filter(Region == "WA") %>%
+  filter(Source == "wmn") %>%
+  mutate(Field = ifelse(Year == "2008", "Southeast", "Northeast"))
+
+WAnova  <- lm(Severity ~ Field + Year, data = WA, contrasts = "contr.helmert")
+WAnova  <- anova(WAnova)
+PWA     <- signif(WAnova[1, 5], 3)
+PWA
 ```
@@
 After clone-correction on the hierarchy of Region/Source/Host/Year, a total of `r nInd(dat11) - nInd(dat11cc)` isolates were removed from the data set, resulting in `r nInd(dat11cc)` isolates representing `r nmll(dat11cc)` MLHs that were used in subsequent analyses (Table \@ref(tab:iatab)).
-<!-- 
-The fact that genotypic diversity in poppr uses H, G and lambda should be clarified here.
-It should reflect the results interpretation as well.
-According to Table 2, MI and NE subpopulations exhibit high genotypic diversity when compared to other subpopulations (3.6, 29 and 0.97 for H, G and lambda, respectively for MI and 3.2, 17.8 and 0.94 for NE).
-If this was based on evenness, that should be explained as well.
-It can be confusing stated like this.
-When you shift focus onto diversity (h), the story changes – so please clarify this section to reflect that
--->
 The results showed that, in terms of genotypic diversity (*H*, *G*, and $\lambda$), WA was the most diverse population with both *G* (`r signif(setNames(gdt$G, gdt$Pop)["WA"], 3)`) and $e^H$ (`r signif(setNames(exp(gdt$H), gdt$Pop)["WA"], 3)`) being close to the observed number of MLGs (`r setNames(gdt$MLG, gdt$Pop)["WA"]`).
@@
 Both of these fields were inoculated with sclerotia in 2002, the Northeastern field with sclerotia provided by a commercial bean producer and the Southeastern field with sclerotia from peas (although this was thought to be unsuccessful).
-Despite this information, it is still unclear what has contributed to the differentiation of the 2008 population from WA.
+Despite this information, it is still unclear what has contributed to the differentiation of the 2008 population from WA or why it shares haplotypes with CA and France.
+When we assessed agressiveness between the two fields across years with an ANOVA model, we found that there was a slight effect based on field (P = `r PWA`).
 While the evidence may suggest host as being a factor, previous studies have shown no significant differentiation across host species [@aldrich-wolfe2015genetic].
+It was of interest to compare our data with that of @aldrich-wolfe2015genetic, but we found that, due to differences in data generation, we were unable to confidently perform a comparison (See supplemental file compare-aldrich-wolfe.md[^3] [@kamvar2017data]).
+
+[^3]: Direct link: https://github.com/everhartlab/sclerotinia-366/blob/master/results/compare-aldrich-wolfe.md
 
@@
 This may, in turn increase the risk of introducing new sources of genetic variation through potential outcrossing events.
-<!-- I was also interested in more interpretation/analysis as to why the Mexican population did not have MLH shared with any other regions, the absence of clarity on this is done at a loss. -->
 

### 2017-11-07 doc/manuscript/manuscript.Rmd
removed embarrassing printing and attempt to fix table 2
fix silly mistake

@@
 fa_alleles   <- poptab[, first_axis > q95]
-mx_percent   <- t(fa_alleles[, names(near_private)])/colSums(fa_alleles)
+mx_percent   <- t(fa_alleles[, names(near_private)])/colSums(fa_alleles)[names(near_private)]
 mx_percent   <- prettify(mx_percent[, "Mexico", drop = TRUE], TRUE)
@@
 PWA     <- signif(WAnova[1, 5], 3)
-PWA
 ```
@@
 alleles = `r signif(Amean, 3)`
+
+\newpage
 

### 2017-11-08 doc/manuscript/manuscript.Rmd
acknowledge contributors and add link to interactive vis for #15

@@
 The authors would like to thank Rebecca Higgens for technical support in generating the data for the MCG assessment, aggressiveness ratings, and genotyping; and for providing valuable insights into the historical context of the data collection and curation.
+We would also like to thank Denita Hadziabdic and two anonymous reviewers for their valuable comments and insights that improved the quality of the manuscript.
 
@@
 Node color represents the community assignment based on the walktrap algorithm with a maximum of four steps [@pons2006computing]. 
+An interactive version of this network can be recreated using the code in the "Interactive visualizations" section of the mlg-mcg.md file in the supplementary information (Direct Link: https://github.com/everhartlab/sclerotinia-366/blob/master/results/mlg-mcg.md#interactive-visualizations)  [@kamvar2017data].
 ](../../results/figures/publication/full-graph.pdf){#fullgraph width=100%}

### 2017-11-09 doc/manuscript/manuscript.Rmd
add citation for Rocker

@@
 
-All data processing and analyses were performed in `r R.version.string %>% strsplit("\\(") %>% map_chr(1)`[@R] and are openly available and reproducible at https://github.com/everhartlab/sclerotinia-366/.
+All data processing and analyses were performed in a Rocker "verse" project container running `r R.version.string %>% strsplit("\\(") %>% map_chr(1)`[@boettiger2017introduction; @R] and are openly available and reproducible at https://github.com/everhartlab/sclerotinia-366/.
 Of the 16 microsatellite loci genotyped, five included compound repeats, which made it challenging to accurately/confidently bin alleles into fragment sizes expected for each locus based on the described repeat motif.

### 2017-11-09 doc/manuscript/manuscript.tex
update manuscript to current version

@@
 \title{Population structure and phenotypic variation of \emph{Sclerotinia
-sclerotiorum} from dry bean in the United States}
+sclerotiorum} from dry bean (\emph{Phaseolus vulgaris}) in the United
+States}
 
@@
 producer fields and white mold screening nurseries surveyed over 10
-years in 2003--2012 representing 11 states, Mexico, France, and
-Australia at 11 microsatellite loci resulting in 165 MLHs. Populations
-were loosely structured over space and time based on analysis of
-molecular variance and discriminant analysis of principal components,
-but not by cultivar, aggressiveness, or field source. Of all the regions
-tested, only Mexico (n=18) shared no MLHs with any other region. Using a
-bipartite network-based approach, we found no evidence that the MCGs
-accurately represent MLHs. Our study suggests that breeders should
-continue to test dry bean lines in several white mold screening
-nurseries across the US to account for both the phenotypic and genotypic
-variation that exists across regions.
+years in 2003--2012 representing 11 states in the United States of
+America, Australia, France, and Mexico at 11 microsatellite loci
+resulting in 165 MLHs. Populations were loosely structured over space
+and time based on analysis of molecular variance and discriminant
+analysis of principal components, but not by cultivar, aggressiveness,
+or field source. Of all the regions tested, only Mexico (n=18) shared no
+MLHs with any other region. Using a bipartite network-based approach, we
+found no evidence that the MCGs accurately represent MLHs. Our study
+suggests that breeders should continue to test dry bean lines in several
+white mold screening nurseries across the US to account for both the
+phenotypic and genotypic variation that exists across regions.
 % Dummy abstract text.
@@
 \emph{Sclerotinia sclerotiorum} (Lib.) de Bary is an ascomycete plant
-pathogen with a worldwide distribution. This is a necrotrophic pathogen
-that is primarily homothallic (self-fertilization) and has the ability
-to survive for more than five years in soil using melanized survival
-structures called sclerotia (Bolton et al., 2006; Sexton et al., 2006).
-It causes disease on more than 400 plant species belonging to 75
-families (Boland \& Hall, 1994) including crops of major economic
-importance such as sunflower (\emph{Helianthus spp.}), soybean
+pathogen with a worldwide distribution (Bolton et al., 2006). This is a
+necrotrophic pathogen that is primarily homothallic (self-fertilization)
+and has the ability to survive for more than five years in soil using
+melanized survival structures called sclerotia (Bolton et al., 2006;
+Sexton et al., 2006). It causes disease on more than 400 plant species
+belonging to 75 families (Boland \& Hall, 1994) including crops of major
+economic importance such as sunflower (\emph{Helianthus spp.}), soybean
 (\emph{Glycine max} L.), canola (\emph{Brassica napa} L., \emph{Brassica
 campestris} L.), and dry bean (\emph{Phaseolus vulgaris} L.) (Bolton et
-al., 2006). On dry bean, \emph{S. sclerotiorum} is the causal agent of
-white mold, a devastating disease that can be yield-limiting in
-temperate climates (Steadman, 1983).
+al., 2006).
+
+On dry bean, \emph{S. sclerotiorum} is the causal agent of white mold, a
+devastating disease that can be yield-limiting in temperate climates
+(Steadman, 1983). All above-ground tissues (flowers, stems, leaves,
+pods) are susceptible to infection, first appearing as wet lesions with
+white mycelial tufts, and then bleaching as the tissue senesces
+(Steadman, 1983; Bolton et al., 2006). For many years, white mold has
+been the most serious dry bean disease in the Northwestern United States
+(Otto-Hanson et al., 2011; Knodel et al., 2012, 2015, 2016). The impact
+of white mold on the dry bean industry in the Northwestern United States
+alone has been estimated at a loss of 140 kg/ha with just 10\% disease
+incidence (Ramasubramaniam et al., 2008).
 
@@
 bean production areas across the United States with additional sites in
-Mexico, France, and Australia (Steadman et al., 2003, 2004, 2005, 2006;
+Australia, France, and Mexico (Steadman et al., 2003, 2004, 2005, 2006;
 Otto-Hanson \& Steadman, 2007, 2008; McCoy \& Steadman, 2009). These
@@
 against \emph{S. sclerotiorum} and employ standardized protocols for
-screening new cultivars for resistance to white mold. It has previously
-been shown that aggressiveness (the severity of disease symptoms on the
-host) is significantly different across white mold screening nursery
-sites in separate geographic regions (Otto-Hanson et al., 2011). The
-genetic structure and mode of reproduction in these populations,
-however, is currently unknown.
+screening new cultivars for resistance to white mold (Steadman et al.,
+2003; Otto-Hanson et al., 2011). These protocols included three
+established cultivars used for comparison in the trials: Beryl (great
+northern bean, susceptible), Bunsi (aka Ex Rico, navy bean, low
+susceptibility), and G122 (cranberry bean, partial resistance) (Tu \&
+Beversdorf, 1982; Steadman et al., 2005; Otto-Hanson et al., 2011). It
+has previously been shown that aggressiveness (the severity of disease
+symptoms on the host) is significantly different across white mold
+screening nursery sites in separate geographic regions (Otto-Hanson et
+al., 2011). The genetic structure and mode of reproduction in these
+populations, however, is currently unknown.
 
@@
 fungal hyphae from different colonies to appear to grow together without
-forming a barrier of dead cells between them (known as a barrage line)
-(Leslie, 1993; Sirjusingh \& Kohn, 2001). Mycelial compatibility has
-been used as a proxy for vegetative compatibility, a fungal trait
-controlled by several independent genes controlling the ability for two
-hyphae to fuse and grow as a single unit (Leslie, 1993; Schafer \& Kohn,
-2006). Because of the genetic connection to vegetative compatibility,
-two isolates that are mycelially compatible were considered clones
-(Leslie, 1993); but correlation with genetic markers, such as
-microsatellites, have shown mixed results (Ford et al., 1995; Micali \&
-Smith, 2003; Jo et al., 2008; Attanayake et al., 2012; Papaioannou \&
-Typas, 2014; Lehner et al., 2017).
+forming a barrier of dead cells between them (known as a barrage line,
+Fig. \ref{mcg-fig}B) (Leslie, 1993; Sirjusingh \& Kohn, 2001). Mycelial
+compatibility has been used as a proxy for vegetative compatibility, a
+fungal trait controlled by several independent genes controlling the
+ability for two hyphae to fuse and grow as a single unit (Fig.
+\ref{mcg-fig}A) (Leslie, 1993; Schafer \& Kohn, 2006). Because of the
+genetic connection to vegetative compatibility, two isolates that are
+mycelially compatible were considered clones (Leslie, 1993); but
+correlation with genetic markers, such as microsatellites, have shown
+mixed results (Ford et al., 1995; Micali \& Smith, 2003; Jo et al.,
+2008; Attanayake et al., 2012; Papaioannou \& Typas, 2014; Lehner et
+al., 2017). However, a question remains about the relationship between
+mycelial compatibility groups and clonal genotypes.
 
@@
 and 2012 from dry bean cultivars among different geographic locations in
-the United States, France, Mexico, and Australia. We wanted to know if
-the \emph{S. sclerotiorum} populations from white mold screening
+the Australia, France, Mexico, and the United States. We wanted to know
+if the \emph{S. sclerotiorum} populations from white mold screening
 nurseries were representative of the fields within the same region. As
@@
 structure can help direct breeding efforts. By investigating these aims,
-we hope to effectively describe the population structure of \emph{S.
+we will effectively describe the population structure of \emph{S.
 sclerotiorum} in the USA and make available our database of isolates for
@@
 
-Several of the isolates used for this study were collected as reported
-in previous studies using the same methods (Otto-Hanson et al., 2011).
-Broadly, isolates were collected from two sources: white mold screening
-nurseries (wmn) or producer fields. White mold screening nurseries were
-5m x 10m in size and maintained without application of fungicides to
-observe natural incidence of white mold. The early nursery plots were
-incorporated with a basal dressing of N:P:K = 1:3:2 and side dressing of
-0:3:2 during the growing season (Steadman et al., 2003).
+Several (156) of the isolates used for this study were collected as
+reported in previous studies using the same methods (Otto-Hanson et al.,
+2011). Broadly, isolates were collected from two sources: white mold
+screening nurseries (wmn) or producer fields. White mold screening
+nurseries were 5m x 10m in size and maintained without application of
+fungicides to observe natural incidence of white mold. The early nursery
+plots were incorporated with a basal dressing of N:P:K = 1:3:2 and side
+dressing of 0:3:2 during the growing season (Steadman et al., 2003).
 
@@
 Sclerotia of \emph{S. sclerotiorum} were collected over several years
-from grower fields and/or wmn in 11 states of the United States of
-America, Mexico, France, and Australia (Table \ref{tab:isolate-table}).
-After collection, sclerotia were stored in Petri plates lined with
-filter paper, then stored at \(20~^{\circ}\)F or -\(4~^{\circ}\)C.
-Sclerotia were surface-sterilized with 50\% Clorox bleach (at least 6\%
-NaOCL, The Clorox Company, Oakland, CA) solution for 3 min, and double
-rinsed with ddH\(_2\)O for 3 min. The sterilized sclerotia were then
-placed on water agar plates (16g of Bacto agar per liter of ddH\(_2\)O,
-BD Diagnostic Systems, Sparks, MD), with four to five sclerotia of each
-isolate separated on each plate and stored on the counter top at room
+from grower fields and/or wmn in 11 states of the Australia, France,
+Mexico, and the United States (Table \ref{tab:isolate-table}). After
+collection, sclerotia were stored in Petri plates lined with filter
+paper, then stored at \(20~^{\circ}\)F or -\(4~^{\circ}\)C. Sclerotia
+were surface-sterilized with 50\% Clorox bleach (at least 6\% NaOCl, The
+Clorox Company, Oakland, CA) solution for 3 min, and double rinsed with
+ddH\(_2\)O for 3 min. The sterilized sclerotia were then placed on water
+agar plates (16g of Bacto agar per liter of ddH\(_2\)O, BD Diagnostic
+Systems, Sparks, MD), with four to five sclerotia of each isolate
+separated on each plate and stored on the counter top at room
 temperature for 5 to 6 days. An 8-mm plug from a 5- or 6-day-old culture
@@
 MCG was determined as described previously through co-culturing pairs of
-2-day-old isolates 2.5 cm apart on Diana Sermons (DS) Medium (Cubeta et
-al., 2001). Incompatibility of different MCGs resulted in formation of a
-barrage line accompanied by formation of sclerotia on either side of the
-barrage line, indicating the limits of the isolates' growth (Kohn et
-al., 1990; Leslie, 1993; Otto-Hanson et al., 2011). Isolates were
-compared in a pairwise manner for each site and then representatives
-among sites were compared to determine mycelial compatibility groups by
-scoring compatible and incompatible interactions (Otto-Hanson et al.,
-2011). No MCGs were compatible with any other MCG.
+2-day-old isolates 2.5 cm apart on Diana Sermons (DS) Medium (Fig.
+\ref{mcg-fig}) (Cubeta et al., 2001). Incompatibility of different MCGs
+resulted in formation of a barrage line accompanied by formation of
+sclerotia on either side of the barrage line, indicating the limits of
+the isolates' growth (Kohn et al., 1990; Leslie, 1993; Otto-Hanson et
+al., 2011). Isolates were compared in a pairwise manner for each site
+and then representatives among sites were compared to determine mycelial
+compatibility groups by scoring compatible and incompatible interactions
+(Otto-Hanson et al., 2011). No MCGs were compatible with any other MCG.
 
@@
 
-All data processing and analyses were performed in R version 3.4.1 (R
-Core Team, 2017) and are openly available and reproducible at
+All data processing and analyses were performed in a Rocker ``verse''
+project container running R version 3.4.2 (Boettiger \& Eddelbuettel,
+2017; R Core Team, 2017) and are openly available and reproducible at
 \url{https://github.com/everhartlab/sclerotinia-366/}. Of the 16
@@
 
-We used ANOVA to assess if aggressiveness was significantly different
-with respect to Region, MCG, or multilocus haplotype (MLH). To minimize
-complications due to small sample sizes, we chose the top 10 MCGs,
-representing 56.5\% of the isolates collected, the 10 most abundant MLHs
-representing 26.7\% of the isolates, and populations with more than five
-isolates. If ANOVA results were significantly different at \(\alpha\) =
-0.05, pairwise differences were assessed using Tukey's HSD test
-(\(\alpha\) = 0.05) using the \texttt{HSD.test()} function in the
-package \emph{agricolae} version 1.2.8 (Mendiburu \& Simon, 2015).
+We used ANOVA to assess if aggressiveness (determined via straw test on
+a scale of 1--9) was significantly different with respect to Region,
+MCG, or multilocus haplotype (MLH). To minimize complications due to
+small sample sizes, we chose the top 10 MCGs, representing 56.5\% of the
+isolates collected, the 10 most abundant MLHs representing 26.7\% of the
+isolates, and populations with more than five isolates. If ANOVA results
+were significantly different at \(\alpha\) = 0.05, pairwise differences
+were assessed using Tukey's HSD test (\(\alpha\) = 0.05) using the
+\texttt{HSD.test()} function in the package \emph{agricolae} version
+1.2.8 (Mendiburu \& Simon, 2015).
 
@@
 2006), \emph{dplyr} version 0.7.4 (Wickham et al., 2017), and
-\emph{purrr} version 0.2.3 (Henry \& Wickham, 2017). Each node was
+\emph{purrr} version 0.2.4 (Henry \& Wickham, 2017). Each node was
 weighted by the fraction of shared MLHs. Each edge represented a single
@@
 A total of 366 isolates were collected from 2003 to 2012 (except 2006
-and 2011) from diseased dry bean plants in Australia, Mexico, France,
-and 11 states in the USA (Table \ref{tab:isolate-table}). We observed a
-total of 165 MLHs (215 with 16 loci). Over 50\% of the isolates came
-from four states, MI (62), ND (60), WA (59), NE (47). Four regions had
-fewer than 10 isolates, Australia (6), WI (2), NY (1), ID (1). We
-observed 87 MCGs, the most abundant of which (`MCG 5') was represented
-by 73 isolates over 37 MLHs (Fig \ref{biggraph}A,C).
+and 2011) from diseased dry bean plants in 11 states in the United
+States as well as Australia, France, and Mexico (Table
+\ref{tab:isolate-table}). With the 11 loci used in the analyses (Table
+\ref{tab:locus-stats}), we observed a total of 165 MLHs (215 with 16
+loci). These 11 loci represented 7 chromosomes in the \emph{S.
+sclerotiorum} genome with a minimum distance of 55Kbp between two loci
+on the same chromosome. Over 50\% of the isolates came from four states,
+MI (62), ND (60), WA (59), NE (47). Four regions had fewer than 10
+isolates, Australia (6), WI (2), NY (1), ID (1). We observed 87 MCGs,
+the most abundant of which (`MCG 5') was represented by 73 isolates over
+37 MLHs (Fig \ref{biggraph}A,C).
 
@@
 present. Analysis of the haplotype accumulation curve showed no clear
-plateau for 11 or 16 loci (Supplemental Information), indicating that we
-would likely obtain more multilocus haplotypes if we were to genotype
+plateau for 11 or 16 loci (See section on `Loading Data and Setting
+Strata' in the MLG-distribution.md\footnote{Direct link:
+  \url{https://github.com/everhartlab/sclerotinia-366/blob/master/results/MLG-distribution.md\#loading-data-and-setting-strata}}
+file in the supplemental files (Kamvar et al., 2017)), indicating that
+we would likely obtain more multilocus haplotypes if we were to genotype
 more loci.
 
-After clone-correction on the hierarchy of Region/Source/Host/Year, we
-were left with 318 isolates representing 165 MLHs (Table
-\ref{tab:iatab}). The results show that most populations exhibited
-relatively high amounts of genotypic diversity, with the exception of MI
-and NE, which exhibited lower-than average \(E_5\) values. Mexico had
-the lowest average value of gene diversity, \emph{h} and the lowest
-\emph{eMLH}, suggesting low overall genetic diversity. Nearly all
-populations showed evidence for linkage (Table \ref{tab:iatab}), which
-serves as evidence for clonal reproduction or other forms of non-random
-mating. The only exceptions are CA (\emph{P} = 0.043) and Australia
-(\emph{P} = 0.052). Both of these populations showed only moderate
-significance with \(\bar{r}_d\) values of 0.03 and 0.12, respectively.
-
-\begin{longtable}[]{@{}rrllrl@{}}
+After clone-correction on the hierarchy of Region/Source/Host/Year, a
+total of 48 isolates were removed from the data set, resulting in 318
+isolates representing 165 MLHs that were used in subsequent analyses
+(Table \ref{tab:iatab}). The results showed that, in terms of genotypic
+diversity (\emph{H}, \emph{G}, and \(\lambda\)), WA was the most diverse
+population with both \emph{G} (54.3) and \(e^H\) (55.3) being close to
+the observed number of MLGs (56). This indicated that there are few
+duplicated genotypes in WA (Table \ref{tab:iatab}). A more useful metric
+to compare populations, however, is \(E_5\), which scales from 0 to 1
+where 1 indicates all unique genotypes (Grünwald et al., 2003).
+Evaluating by \(E_5\) shows that both MI and NE exhibit lower than
+average values, indicating that there are over-represented genotypes in
+the popualtions (table \ref{tab:iatab}). When we look at Mexico, we can
+observed that it had relatively high values of \(E_5\) and genotypic
+diversity, but low richness, as measured by \emph{eMLG}. Moreover,
+Mexico had the lowest value for \emph{h}, which is a measure of allelic
+diversity. Nearly all populations showed evidence for linkage (Table
+\ref{tab:iatab}), which serves as evidence for clonal reproduction or
+other forms of non-random mating. The only exceptions are CA (\emph{P} =
+0.043) and Australia (\emph{P} = 0.052). Both of these populations
+showed only moderate significance with \(\bar{r}_d\) values of 0.03 and
+0.12, respectively.
+
+\begin{longtable}[]{@{}rrrlrl@{}}
 \caption{\label{tab:locus-stats} Allelic diversity on full data set at loci
@@
 Range\strut
-\end{minipage} & \begin{minipage}[b]{0.15\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[b]{0.29\columnwidth}\raggedleft\strut
 Repeat Motif\strut
@@
 318--324\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright\strut
-di-\strut
+\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
+(GT)\strut
 \end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
@@
 483--495\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright\strut
-hexa-\strut
+\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
+(TTTTTC)(TTTTTG)(TTTTTC)\strut
 \end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
@@
 158--174\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright\strut
-di-\strut
+\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
+(GA)\strut
 \end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
@@
 244--270\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright\strut
-di-\strut
+\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
+(CA)\strut
 \end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
@@
 360--382\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright\strut
-di-\strut
+\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
+(CA)(CT)\strut
 \end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
@@
 214--222\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright\strut
-di-\strut
+\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
+(CA)\strut
 \end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
@@
 342--363\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright\strut
-tri-\strut
+\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
+(TTA)\strut
 \end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
@@
 280--282\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright\strut
-di-\strut
+\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
+(GT)GG(GT)\strut
 \end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
@@
 153--216\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright\strut
-tetra-\strut
+\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
+(TACA)\strut
 \end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
@@
 370--386\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright\strut
-tetra-\strut
+\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
+(TATG)\strut
 \end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
@@
 339--416\strut
-\end{minipage} & \begin{minipage}[t]{0.15\columnwidth}\raggedright\strut
-tetra-\strut
+\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
+(TAGA)\strut
 \end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
@@
 \end{longtable}
+
+\newpage
 
@@
 \includegraphics{../../results/figures/publication/aggressiveness.pdf}
-\caption{Strip Plot of aggressiveness by population arranged in
+\caption{Strip plot of aggressiveness by population arranged in
 descending order of mean aggressiveness for all populations with N
@@
 with an adjusted R\textsuperscript{2} of 0.0675 (\emph{P} = 0.001).
-Assessment of the marginal effects showed that all varaibles
+Assessment of the marginal effects showed that all varibles
 significantly explained genetic variation (\(P \leq 0.007\)). We found
@@
 and median of 4.85. The group mean averages were 4.88, 5.13, and 5.19
-for Region, MCG, and MLH, respectively. Our assessment of aggressiveness
-in association with Region showed a significant effect (\emph{P}
-\textless{} 1.00e\textsuperscript{-4}), with means that ranged from 5.8
-(MN) to 4.0 (CA) (Fig. \ref{aggressiveness-boxplot}, Table
+for Region, MCG, and MLH, respectively. A strip plot showing the
+distribution of severity across these three variables simultaneously can
+be seen in figure \ref{mcg-mlh-region-aggressiveness}. Our assessment of
+aggressiveness in association with Region showed a significant effect
+(\emph{P} \textless{} 1.00e\textsuperscript{-4}), with means that ranged
+from 5.8 (MN) to 4.0 (CA) (Fig. \ref{aggressiveness-boxplot}, Table
 \ref{tab:region-aggressiveness}). MCGs also showed a significant effect
-(\emph{P} \textless{} 1.00e\textsuperscript{-4}), with means that ranged
-from 6.0 (`MCG 44') to 4.6 (`MCG 49'; Table
-\ref{tab:MCG-aggressiveness}). We additionally found a significant
-effect for MLHs (\emph{P} = 7.44e\textsuperscript{-4}), with means that
-ranged from 6.0 (`MLH 78') to 4.3 (`MLH 140') (Table
+(\emph{P} \textless{} 0.001), with means that ranged from 6.0 (`MCG 44')
+to 4.6 (`MCG 49'; Table \ref{tab:MCG-aggressiveness}). We additionally
+found a significant effect for MLHs (\emph{P} \textless{} 0.001), with
+means that ranged from 6.0 (`MLH 78') to 4.3 (`MLH 140') (Table
 \ref{tab:MLG-aggressiveness}).
@@
 with a modularity of 0.17: A coastal community (CA, OR, WA, and NY), a
-midwest community (CO, NE, ND, MI), and an international community (MN,
-France, Australia). Although analysis with 16 loci resulted in the
+midwest community (CO, ND, NE, MI), and an international community
+(Australia, France, MN). Although analysis with 16 loci resulted in the
 removal of the NY node because it no longer shared a haplotype with OR,
@@
 when we compared the three cultivars, Beryl, Bunsi, and G122, we found
-no significant differentiation (Supplementary Information).
+no significant differentiation (See section on `Host Differentiation' in
+the wmn-differentiation.md\footnote{Direct link:
+  \url{https://github.com/everhartlab/sclerotinia-366/blob/master/results/wmn-differentiation.md\#host-differentiation}}
+file in the supplemental files (Kamvar et al., 2017)).
 
@@
 \end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering\strut
-0.029\strut
+0.033\strut
 \end{minipage}\tabularnewline
@@
 \end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering\strut
-0.486\strut
+0.448\strut
 \end{minipage}\tabularnewline
@@
 
-Discriminant Analysis of Principal Components (DAPC) was performed
-grouping by Region with the first 21 principal components, representing
+Discriminant Analysis of Principal Components (DAPC) was performed by
+grouping Region with the first 21 principal components, representing
 88.1\% of the total variance. The first discriminant axis (representing
-36.1\% of the discriminatory power) separated the centroid for the
+63.9\% of the discriminatory power) separated the centroid for the
 Mexico isolates from the rest of the data, indicating strong
 differentiation (Fig. \ref{DAPC}b). The second discriminant axis,
-representing 25.3\% of the discriminatory power, separated the centroid
+representing 10.8\% of the discriminatory power, separated the centroid
 for the CA isolates. The mean population assignment probabilities for
@@
 representing 89\% of the total variance. The first two discriminant axes
-(representing 38.5\% of the discriminatory power) failed to separate any
+(representing 100\% of the discriminatory power) failed to separate any
 of the cultivars where the mean posterior assignment probabilities were
@@
 unclear what has contributed to the differentiation of the 2008
-population from WA. While the evidence may suggest host as being a
-factor, previous studies have shown no significant differentiation
-across host species (Aldrich-Wolfe et al., 2015).
+population from WA or why it shares haplotypes with CA and France. When
+we assessed agressiveness between the two fields across years with an
+ANOVA model, we found that there was a slight effect based on field (P =
+0.0127). While the evidence may suggest host as being a factor, previous
+studies have shown no significant differentiation across host species
+(Aldrich-Wolfe et al., 2015). It was of interest to compare our data
+with that of Aldrich-Wolfe et al. (2015), but we found that, due to
+differences in data generation, we were unable to confidently perform a
+comparison (See supplemental file compare-aldrich-wolfe.md\footnote{Direct
+  link:
+  \url{https://github.com/everhartlab/sclerotinia-366/blob/master/results/compare-aldrich-wolfe.md}}
+(Kamvar et al., 2017)).
 
@@
 overlapping distributions in the DAPC (i.e.~NE and MI, Fig.
-\ref{DAPC-RY}). DAPC is based on the principal components of Euclidean
-distance between genotypes (Jombart et al., 2010). Unlike Bruvo's
-distance, Euclidean distance does not take into account the magnitude of
-the difference between alleles, which could inflate the distance measure
-in the presence of private alleles (Bruvo et al., 2004). We found no
-evidence of private alleles in the Mexico and CA isolates, suggesting
-that this pattern separating these populations from the others was not
-an artifact. We believe that the differences we observe from Mexico may
-reflect differences in climate.
+\ref{DAPC-RY}). DAPC is based on the principal components of allele
+counts (Jombart et al., 2010). Unlike Bruvo's distance, this does not
+take into account the magnitude of the difference between alleles, which
+could inflate the distance measure in the presence of private alleles
+(Bruvo et al., 2004). While we found no evidence of private alleles in
+the Mexico and CA isolates, we did find that the alleles driving the
+first axis in figure \ref{DAPC}A (alleles 174, 256, and 372 in loci 7-2,
+8-3, and 9-2, respectively) were overrepresented in Mexico (where
+\textgreater{}75\% of the alleles came from the region). However, all
+three of these alleles i) conform to the expected stepwise mutation
+model (Bruvo et al., 2004) and ii) are at or near the extremes of the
+total range (except for allele 372 at locus 9-2). Moreover, the fact
+that we find three alleles at three independent loci segregating for
+Mexico suggests that the pattern separating these populations from the
+others was not an artifact. We believe that the differences we observe
+from Mexico may reflect differences in climate.
 
@@
 isolates from sub-tropical and tropical climates are differentiated or
-more variable than populations from temperate climates {[}Carbone \&
-Kohn (2001); Attanayake et al. (2013); Lehner \& Mizubuti (2017);{]}.
-This has been attributed to the notion that the fungus has the chance to
-undergo more reproductive cycles in the warmer climate (Carbone \& Kohn,
-2001; Attanayake et al., 2013). The strongest evidence to date
-supporting this hypothesis is from Attanayake et al. (2013), showing
-that populations in sub-tropical regions of China have been found to be
-more variable, sexually reproducing, and unrelated to populations in
-temperate regions of the USA. This result however, may be driven more by
-geography and agricultural practice as opposed to climate.
+more variable than populations from temperate climates (Carbone \& Kohn,
+2001; Attanayake et al., 2013; Lehner \& Mizubuti, 2017). This has been
+attributed to the notion that the fungus has the chance to undergo more
+reproductive cycles in the warmer climate (Carbone \& Kohn, 2001;
+Attanayake et al., 2013). The strongest evidence to date supporting this
+hypothesis is from Attanayake et al. (2013), showing that populations in
+sub-tropical regions of China have been found to be more variable,
+sexually reproducing, and unrelated to populations in temperate regions
+of the USA. This result however, may be driven more by geography and
+agricultural practice as opposed to climate.
 
@@
 sclerotia. This could explain the fact that we see shared genotypes with
-low \(P_{sex}\) values shared between France, Australia, and the US.
-While we speculate that these transmission events are rare due to the
-genetic structuring by Region, these results suggest that seedborne
-infections may indeed reflect a source of inoculum. This may, in turn
-increase the risk of introducing new sources of genetic variation
-through potential outcrossing events.
+low \(P_{sex}\) values shared between Australia, France, and the United
+States. While we speculate that these transmission events are rare due
+to the genetic structuring by Region, these results suggest that
+seedborne infections may indeed reflect a source of inoculum. This may,
+in turn increase the risk of introducing new sources of genetic
+variation through potential outcrossing events.
 
 When we tested for sexual reproduction, we were unable to find evidence
-for it in any region except for CA and Australia. While the Australia
+for it in any region except for Australia and CA. While the Australia
 population had a non-significant value of \(\bar{r}_d\)---which would
@@
 sample may yet be representative of the genetic pool present in other
-crops and weedy species.
+crops and weedy species. Additionally, while we found no signficant
+association between genotype and aggressiveness, it is important to note
+that the straw test is only one measure of aggressiveness. Additional
+phenotypes for aggressiveness should be evaluated for future studieds.
 
@@
 reproducible and available at The Open Science Framework
-\url{https://osf.io/ejb5y}.
+\url{https://osf.io/ejb5y} (Kamvar et al., 2017).
 
@@
 genotyping; and for providing valuable insights into the historical
-context of the data collection and curation.
+context of the data collection and curation. We would also like to thank
+Denita Hadziabdic and two anonymous reviewers for their valuable
+comments and insights that improved the quality of the manuscript.
 
@@
 \centering
+\includegraphics[width=1.00000\textwidth]{images/mcg-trimmed.png}
+\caption{Example of MCG test plates showing (A) a compatible reaction
+with mycelia from two strains overgrowing each other and (B) an
+incompatible reaction with a barrage line of dead tissue forming between
+the two strains. Photo Credit: Rebecca Higgens.}\label{mcg-fig}
+\end{figure}
+
+\begin{figure}
+\centering
+\includegraphics[width=1.00000\textwidth]{../../results/figures/publication/RMM-aggressiveness.pdf}
+\caption{Strip plot of aggressiveness for the eight most abundant MCGs
+partitioned by region. Filled circles indicate one of the five most
+abundant MLHs and open circles indicate a MLH of lesser
+abundance.}\label{mcg-mlh-region-aggressiveness}
+\end{figure}
+
+\begin{figure}
+\centering
 \includegraphics[width=1.00000\textwidth]{../../results/figures/publication/full-graph.pdf}
@@
 color represents the community assignment based on the walktrap
-algorithm with a maximum of four steps (Pons \& Latapy,
-2006).}\label{fullgraph}
+algorithm with a maximum of four steps (Pons \& Latapy, 2006). An
+interactive version of this network can be recreated using the code in
+the ``Interactive visualizations'' section of the mlg-mcg.md file in the
+supplementary information (Direct Link:
+\url{https://github.com/everhartlab/sclerotinia-366/blob/master/results/mlg-mcg.md\#interactive-visualizations})
+(Kamvar et al., 2017).}\label{fullgraph}
 \end{figure}
@@
 
+\hypertarget{ref-boettiger2017introduction}{}
+Boettiger, C., \& Eddelbuettel, D. (2017). An introduction to rocker:
+Docker containers for R. \emph{CoRR}, \emph{abs/1710.03675}. Retrieved
+from \url{http://arxiv.org/abs/1710.03675}
+
 \hypertarget{ref-boland1994index}{}
@@
 
+\hypertarget{ref-kamvar2017data}{}
+Kamvar, Z. N., Amaradasa, B. S., Jhala, R., McCoy, S., Steadman, J., \&
+Everhart, S. E. (2017, November). Data and analysis for population
+structure and phenotypic variation of \emph{Sclerotinia sclerotiorum}
+from dry bean (\emph{Phaseolus vulgaris}) in the United States. Open
+Science Framework. \url{https://doi.org/10.17605/OSF.IO/EJB5Y}
+
 \hypertarget{ref-kamvar2015novel}{}
@@
 \url{https://doi.org/10.7717/peerj.281}
+
+\hypertarget{ref-knodel2012dry}{}
+Knodel, J., Beauzay, P., Franzen, D., Kandel, H., Markell, S., Osorno,
+J., Pasche, J., \& Zollinger, R. (2012). 2012 dry bean grower survey of
+production, pest problems and pesticide use in Minnesota and North
+Dakota. \emph{North Dakota State University Extension}.
+
+\hypertarget{ref-knodel2015dry}{}
+Knodel, J., Beauzay, P., Franzen, D., Kandel, H., Markell, S., Osorno,
+J., Pasche, J., \& Zollinger, R. (2015). 2015 dry bean grower survey of
+production, pest problems and pesticide use in Minnesota and North
+Dakota. \emph{North Dakota State University Extension}.
+
+\hypertarget{ref-knodel2016dry}{}
+Knodel, J., Beauzay, P., Franzen, D., Kandel, H., Markell, S., Osorno,
+J., Pasche, J., \& Zollinger, R. (2016). 2016 dry bean grower survey of
+production, pest problems and pesticide use in Minnesota and North
+Dakota. \emph{North Dakota State University Extension}.
 
@@
 
+\hypertarget{ref-ramasubramaniam2008estimates}{}
+Ramasubramaniam, H., Río Mendoza, L. E. del, \& Bradley, C. A. (2008).
+Estimates of yield and economic losses associated with white mold of
+rain-fed dry bean in north dakota. \emph{Agronomy Journal},
+\emph{100}(2), 315. \url{https://doi.org/10.2134/agronj2007.0127}
+
 \hypertarget{ref-sambrook1989molecular}{}
@@
 
+\hypertarget{ref-tu1982tolerance}{}
+Tu, J. C., \& Beversdorf, W. D. (1982). Tolerance to white mold
+(\emph{Sclerotinia sclerotiorum} (lib.) de bary) in Ex Rico 23, a
+cultivar of white bean (\emph{Phaseolus vulgaris} l.). \emph{Canadian
+Journal of Plant Science}, \emph{62}(1), 65--69.
+\url{https://doi.org/10.4141/cjps82-010}
+
 \hypertarget{ref-ggplot2}{}

### 2017-11-10 doc/RMD/by-year.Rmd
add seeds to RMD

@@
 (wmn.amova <- poppr.amova(wmncc, ~Region/Year, dist = bdwmn))
+set.seed(2017-11-10)
 (wmn.amova.test <- randtest(wmn.amova, nrepet = 999))
@@
 (full.amova.2 <- poppr.amova(dat11cc, ~Region/Year, dist = bdfull))
+set.seed(2017-11-10)
 (full.amova.2.test <- randtest(full.amova.2, nrepet = 999)) 
 (full.amova.3 <- poppr.amova(dat11cc, ~Region/SourceType/Year, dist = bdfull))
+set.seed(2017-11-10)
 (full.amova.3.test <- randtest(full.amova.3, nrepet = 999))

### 2017-11-10 tests.sh
add tests for circle

@@ -1,0 +1,17 @@
+#!/bin/bash
+
+# Test if we have all the results. Note, this only tests if the exist. It does
+# not test if they have the same results as the previous files. Some of these
+# files will change due to randomness of the analyses, but the overall results
+# should not change. 
+resfiles=$(grep -E results[^%]+md /analysis/Makefile \
+| sed -r "s_^.{14}_/analysis/_" \
+| sed 's_\\__')
+
+resfiles+=" /analysis/doc/manuscript/manuscript.pdf"
+resfiles+=" /analysis/doc/manuscript/manuscript.tex"
+
+for i in ${resfiles}; 
+do 
+	md5sum $i
+done

### 2017-11-10 doc/RMD/pop-diff.Rmd
add colon after source

@@
   xlab("Population Assignment") +
-  ylab("Mean Probability")
+  ylab("Mean Probability") +
+  labs(list(fill = "Source:"))
 REGION_PLOT

### 2017-11-10 doc/manuscript/manuscript.Rmd
update manuscript with SE comments

@@
 sclerotiorum* and employ standardized protocols for screening new cultivars for resistance to white mold [@steadman2003identification; @otto-hanson2011variation].
-These protocols included three established cultivars used for comparison in the trials: Beryl (great northern bean, susceptible), Bunsi (aka Ex Rico, navy bean, low susceptibility), and G122 (cranberry bean, partial resistance) [@tu1982tolerance; @steadman2005identification; @otto-hanson2011variation].
-It has previously been shown that aggressiveness (the severity of disease symptoms on the host) is significantly different across white mold screening nursery sites in separate geographic regions [@otto-hanson2011variation].
+These protocols included three established cultivars used for comparison in the trials: Beryl (great northern bean, susceptible), Bunsi (a.k.a.
+Ex Rico, navy bean, low susceptibility), and G122 (cranberry bean, partial resistance) [@tu1982tolerance; @steadman2005identification; @otto-hanson2011variation].
+It was previously shown that aggressiveness (the severity of disease symptoms on the host) is significantly different across white mold screening nursery sites in separate geographic regions [@otto-hanson2011variation].
 The genetic structure and mode of reproduction in these populations, however, is currently unknown.
@@
 Because of the genetic connection to vegetative compatibility, two isolates that are mycelially compatible were considered clones [@leslie1993fungal]; but correlation with genetic markers, such as microsatellites, have shown mixed results [@ford1995heterokaryon; @micali2003independence; @jo2008reassessment; @attanayake2012genetic; @papaioannou2014barrage; @lehner2017independently].
-However, a question remains about the relationship between mycelial compatibility groups and clonal genotypes.
+Thus, the relationship between mycelial compatibility groups and clonal genotypes remains unclear.
 
@@
 
-We used ANOVA to assess if aggressiveness (determined via straw test on a scale of 1--9) was significantly different with respect to Region, MCG, or multilocus haplotype (MLH). 
+We used ANOVA to assess if aggressiveness (determined via straw test on a scale of 1--9 as described above) was significantly different with respect to Region, MCG, or multilocus haplotype (MLH). 
 To minimize complications due to small sample sizes, we chose the top 10 MCGs, representing 56.5% of the isolates collected, the 10 most abundant MLHs representing 26.7% of the isolates, and populations with more than five isolates.
@@
 With the 11 loci used in the analyses (Table \@ref(tab:locus-stats)), we observed a total of 165 MLHs (215 with 16 loci).
-These 11 loci represented 7 chromosomes in the *S. sclerotiorum* genome with a minimum distance of 55Kbp between two loci on the same chromosome.
+These 11 loci are found on 7 chromosomes in the *S. sclerotiorum* genome with a minimum distance of 55Kbp between two loci on the same chromosome.
 Over 50% of the isolates came from four states, MI (`r pop_size[["MI"]]`), ND (`r pop_size[["ND"]]`), WA (`r pop_size[["WA"]]`), NE (`r pop_size[["NE"]]`). 
@@
 This indicated that there are few duplicated genotypes in WA (Table \@ref(tab:iatab)).
-A more useful metric to compare populations, however, is $E_5$, which scales from 0 to 1 where 1 indicates all unique genotypes [@grunwald2003analysis]. 
+A more useful metric to compare populations, however, is $E_5$, which scales from 0 to 1, where 1 indicates all unique genotypes [@grunwald2003analysis]. 
 Evaluating by $E_5$ shows that both MI and NE exhibit lower than average values, indicating that there are over-represented genotypes in the popualtions (table \@ref(tab:iatab)). 
-When we look at Mexico, we can observed that it had relatively high values of $E_5$ and genotypic diversity, but low richness, as measured by *eMLG*.
+When we look at Mexico, we observed that it had relatively high values of $E_5$ and genotypic diversity, but low richness, as measured by *eMLG*.
 Moreover, Mexico had the lowest value for *h*, which is a measure of allelic diversity.
-<!-- The results show that most populations exhibited relatively high amounts of genotypic diversity, with the exception of MI and NE, which have which exhibited lower-than average $E_5$ values. -->
-<!-- Mexico had the lowest average value of gene diversity, *h* and the lowest *eMLH*, suggesting low overall genetic diversity.  -->
 Nearly all populations showed evidence for linkage (Table \@ref(tab:iatab)), which serves as evidence for clonal reproduction or other forms of non-random mating. 
@@
 While we found no evidence of private alleles in the Mexico and CA isolates, we did find that the alleles driving the first axis in figure \@ref(DAPC)A (`r mx_allele_string`, respectively) were overrepresented in Mexico (where >75% of the alleles came from the region).
-However, all three of these alleles i) conform to the expected stepwise mutation model [@bruvo2004simple] and ii) are at or near the extremes of the total range (except for `r glue::glue("allele {mx_alleles[3]} at locus {mx_loci[3]}")`).
+However, all three of these alleles, i) conform to the expected stepwise mutation model [@bruvo2004simple] and ii) are at or near the extremes of the total range (except for `r glue::glue("allele {mx_alleles[3]} at locus {mx_loci[3]}")`).
 Moreover, the fact that we find three alleles at three independent loci segregating for Mexico suggests that the pattern separating these populations from the others was not an artifact. 
-We believe that the differences we observe from Mexico may reflect differences in climate.
+We believe that the differences in populations observed from Mexico may be due to differences in climate that allow greater diversification via sexual outcrossing.
 
@@
 Additionally, while we found no signficant association between genotype and aggressiveness, it is important to note that the straw test is only one measure of aggressiveness. 
-Additional phenotypes for aggressiveness should be evaluated for future studieds.
+Additional phenotypes for aggressiveness should be evaluated for future studies.
 
@@
 The authors would like to thank Rebecca Higgens for technical support in generating the data for the MCG assessment, aggressiveness ratings, and genotyping; and for providing valuable insights into the historical context of the data collection and curation.
-We would also like to thank Denita Hadziabdic and two anonymous reviewers for their valuable comments and insights that improved the quality of the manuscript.
+We would also like to thank Denita Hadziabdic and two other anonymous reviewers for their valuable comments and insights that improved the quality of the manuscript.
 

### 2017-11-10 doc/manuscript/manuscript.tex
add current results
update manuscript with SE comments

@@
 established cultivars used for comparison in the trials: Beryl (great
-northern bean, susceptible), Bunsi (aka Ex Rico, navy bean, low
+northern bean, susceptible), Bunsi (a.k.a. Ex Rico, navy bean, low
 susceptibility), and G122 (cranberry bean, partial resistance) (Tu \&
 Beversdorf, 1982; Steadman et al., 2005; Otto-Hanson et al., 2011). It
-has previously been shown that aggressiveness (the severity of disease
+was previously shown that aggressiveness (the severity of disease
 symptoms on the host) is significantly different across white mold
@@
 2008; Attanayake et al., 2012; Papaioannou \& Typas, 2014; Lehner et
-al., 2017). However, a question remains about the relationship between
-mycelial compatibility groups and clonal genotypes.
+al., 2017). Thus, the relationship between mycelial compatibility groups
+and clonal genotypes remains unclear.
 
@@
 We used ANOVA to assess if aggressiveness (determined via straw test on
-a scale of 1--9) was significantly different with respect to Region,
-MCG, or multilocus haplotype (MLH). To minimize complications due to
-small sample sizes, we chose the top 10 MCGs, representing 56.5\% of the
-isolates collected, the 10 most abundant MLHs representing 26.7\% of the
-isolates, and populations with more than five isolates. If ANOVA results
-were significantly different at \(\alpha\) = 0.05, pairwise differences
-were assessed using Tukey's HSD test (\(\alpha\) = 0.05) using the
-\texttt{HSD.test()} function in the package \emph{agricolae} version
-1.2.8 (Mendiburu \& Simon, 2015).
+a scale of 1--9 as described above) was significantly different with
+respect to Region, MCG, or multilocus haplotype (MLH). To minimize
+complications due to small sample sizes, we chose the top 10 MCGs,
+representing 56.5\% of the isolates collected, the 10 most abundant MLHs
+representing 26.7\% of the isolates, and populations with more than five
+isolates. If ANOVA results were significantly different at \(\alpha\) =
+0.05, pairwise differences were assessed using Tukey's HSD test
+(\(\alpha\) = 0.05) using the \texttt{HSD.test()} function in the
+package \emph{agricolae} version 1.2.8 (Mendiburu \& Simon, 2015).
 
@@
 \ref{tab:locus-stats}), we observed a total of 165 MLHs (215 with 16
-loci). These 11 loci represented 7 chromosomes in the \emph{S.
+loci). These 11 loci are found on 7 chromosomes in the \emph{S.
 sclerotiorum} genome with a minimum distance of 55Kbp between two loci
@@
 duplicated genotypes in WA (Table \ref{tab:iatab}). A more useful metric
-to compare populations, however, is \(E_5\), which scales from 0 to 1
+to compare populations, however, is \(E_5\), which scales from 0 to 1,
 where 1 indicates all unique genotypes (Grünwald et al., 2003).
@@
 average values, indicating that there are over-represented genotypes in
-the popualtions (table \ref{tab:iatab}). When we look at Mexico, we can
+the popualtions (table \ref{tab:iatab}). When we look at Mexico, we
 observed that it had relatively high values of \(E_5\) and genotypic
@@
 \end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering\strut
-0.033\strut
+0.031\strut
 \end{minipage}\tabularnewline
@@
 \end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering\strut
-0.448\strut
+0.497\strut
 \end{minipage}\tabularnewline
@@
 \textgreater{}75\% of the alleles came from the region). However, all
-three of these alleles i) conform to the expected stepwise mutation
+three of these alleles, i) conform to the expected stepwise mutation
 model (Bruvo et al., 2004) and ii) are at or near the extremes of the
@@
 Mexico suggests that the pattern separating these populations from the
-others was not an artifact. We believe that the differences we observe
-from Mexico may reflect differences in climate.
+others was not an artifact. We believe that the differences in
+populations observed from Mexico may be due to differences in climate
+that allow greater diversification via sexual outcrossing.
 
@@
 that the straw test is only one measure of aggressiveness. Additional
-phenotypes for aggressiveness should be evaluated for future studieds.
+phenotypes for aggressiveness should be evaluated for future studies.
 
@@
 context of the data collection and curation. We would also like to thank
-Denita Hadziabdic and two anonymous reviewers for their valuable
+Denita Hadziabdic and two other anonymous reviewers for their valuable
 comments and insights that improved the quality of the manuscript.
@@
 production, pest problems and pesticide use in Minnesota and North
-Dakota. \emph{North Dakota State University Extension}.
+Dakota. \emph{North Dakota State University Extension}, \emph{E1640}.
 
@@
 production, pest problems and pesticide use in Minnesota and North
-Dakota. \emph{North Dakota State University Extension}.
+Dakota. \emph{North Dakota State University Extension}, \emph{E1802}.
 
@@
 production, pest problems and pesticide use in Minnesota and North
-Dakota. \emph{North Dakota State University Extension}.
+Dakota. \emph{North Dakota State University Extension}, \emph{E1841}.
 
@@
 \hypertarget{ref-ramasubramaniam2008estimates}{}
-Ramasubramaniam, H., Río Mendoza, L. E. del, \& Bradley, C. A. (2008).
+Ramasubramaniam, H., del Río Mendoza, L. E., \& Bradley, C. A. (2008).
 Estimates of yield and economic losses associated with white mold of
-rain-fed dry bean in north dakota. \emph{Agronomy Journal},
+rain-fed dry bean in North Dakota. \emph{Agronomy Journal},
 \emph{100}(2), 315. \url{https://doi.org/10.2134/agronj2007.0127}
@@
 Strausbaugh, C., \& Forster, R. (2003). Management of white mold of
-beans. \emph{Pacific Northwest Extension}. Retrieved from
+beans. \emph{Pacific Northwest Extension}, \emph{PNW568}. Retrieved from
 \url{http://http://www.extension.uidaho.edu/publishing/pdf/PNW/PNW0568.pdf}
@@
 Tu, J. C., \& Beversdorf, W. D. (1982). Tolerance to white mold
-(\emph{Sclerotinia sclerotiorum} (lib.) de bary) in Ex Rico 23, a
-cultivar of white bean (\emph{Phaseolus vulgaris} l.). \emph{Canadian
+(\emph{Sclerotinia sclerotiorum} (Lib.) De Bary) in Ex Rico 23, a
+cultivar of white bean (\emph{Phaseolus vulgaris} L.). \emph{Canadian
 Journal of Plant Science}, \emph{62}(1), 65--69.

### 2017-11-10 doc/manuscript/manuscript.tex
minor formatting for labels [skip ci]

@@
 the most abundant of which (`MCG 5') was represented by 73 isolates over
-37 MLHs (Fig \ref{biggraph}A,C).
+37 MLHs (Fig. \ref{biggraph}A,C).
 
@@
 distribution of severity across these three variables simultaneously can
-be seen in figure \ref{mcg-mlh-region-aggressiveness}. Our assessment of
+be seen in Fig. \ref{mcg-mlh-region-aggressiveness}. Our assessment of
 aggressiveness in association with Region showed a significant effect
@@
 isolates, `MLH 75' had 57 neighbors that shared the same mycelial
-compatibility group (Fig \ref{biggraph}, \ref{fullgraph}). Overall,
+compatibility group (Fig. \ref{biggraph}, \ref{fullgraph}). Overall,
 there was no clear pattern to the association between MLH and MCGs.
@@
 the Mexico and CA isolates, we did find that the alleles driving the
-first axis in figure \ref{DAPC}A (alleles 174, 256, and 372 in loci 7-2,
+first axis in Fig. \ref{DAPC}A (alleles 174, 256, and 372 in loci 7-2,
 8-3, and 9-2, respectively) were overrepresented in Mexico (where
@@
 1996; Agapow \& Burt, 2001). The low value of \(\bar{r}_d\) in the CA
-population may represent sexual reproduction, but we can see in Figure
+population may represent sexual reproduction, but we can see in Fig.
 \ref{DAPC-RY} that there is differentiation by year. Thus, this could
@@
 genotyping; and for providing valuable insights into the historical
-context of the data collection and curation. We would also like to thank
-Denita Hadziabdic and two other anonymous reviewers for their valuable
-comments and insights that improved the quality of the manuscript.
-
-\subsection*{Conflicts of Interest}\label{conflicts-of-interest}
-\addcontentsline{toc}{subsection}{Conflicts of Interest}
+context of the data collection and curation.
+
+We would also like to thank Denita Hadziabdic and two other anonymous
+reviewers for their valuable comments and insights that improved the
+quality of the manuscript. \#\# Conflicts of Interest \{-\}
 

### 2017-11-10 doc/manuscript/manuscript.Rmd
minor formatting for labels [skip ci]

@@
 Four regions had fewer than 10 isolates, Australia (`r pop_size[["Australia"]]`), WI (`r pop_size[["WI"]]`), NY (`r pop_size[["NY"]]`), ID (`r pop_size[["ID"]]`).
-We observed 87 MCGs, the most abundant of which ('MCG 5') was represented by 73 isolates over 37 MLHs (Fig \@ref(biggraph)A,C).
+We observed 87 MCGs, the most abundant of which ('MCG 5') was represented by 73 isolates over 37 MLHs (Fig. \@ref(biggraph)A,C).
 
@@
 The group mean averages were `r sprintf("%0.2f, %0.2f, and %0.2f", Raggmean, Maggmean, Haggmean)` for Region, MCG, and MLH, respectively.
-A strip plot showing the distribution of severity across these three variables simultaneously can be seen in figure \@ref(mcg-mlh-region-aggressiveness).
+A strip plot showing the distribution of severity across these three variables simultaneously can be seen in Fig. \@ref(mcg-mlh-region-aggressiveness).
 Our assessment of aggressiveness in association with Region showed a significant effect (*P* < 1.00e^-4^), with means that ranged from 5.8 (MN) to 4.0 (CA) (Fig.
@@
 When we assessed the number of times two different MLHs that are in the same MCG, considering these as potential heterothallic pairings that could result in sexual recombination, we found an average of `r signif(mean(in_reach$N), 3)` potential heterothallic parings per MLH.
-Representing just four isolates, 'MLH 75' had `r big_haplotype$N` neighbors that shared the same mycelial compatibility group (Fig \@ref(biggraph), \@ref(fullgraph)). 
+Representing just four isolates, 'MLH 75' had `r big_haplotype$N` neighbors that shared the same mycelial compatibility group (Fig. \@ref(biggraph), \@ref(fullgraph)). 
 Overall, there was no clear pattern to the association between MLH and MCGs.
@@
 Unlike Bruvo's distance, this does not take into account the magnitude of the difference between alleles, which could inflate the distance measure in the presence of private alleles [@bruvo2004simple].
-While we found no evidence of private alleles in the Mexico and CA isolates, we did find that the alleles driving the first axis in figure \@ref(DAPC)A (`r mx_allele_string`, respectively) were overrepresented in Mexico (where >75% of the alleles came from the region).
+While we found no evidence of private alleles in the Mexico and CA isolates, we did find that the alleles driving the first axis in Fig.
+\@ref(DAPC)A (`r mx_allele_string`, respectively) were overrepresented in Mexico (where >75% of the alleles came from the region).
 However, all three of these alleles, i) conform to the expected stepwise mutation model [@bruvo2004simple] and ii) are at or near the extremes of the total range (except for `r glue::glue("allele {mx_alleles[3]} at locus {mx_loci[3]}")`).
@@
 While the Australia population had a non-significant value of $\bar{r}_d$---which would suggest that we cannot reject the null hypothesis of random mating---the sample size was insufficient from which to draw conclusions [@milgroom1996recombination; @agapow2001indices].
-The low value of $\bar{r}_d$ in the CA population may represent sexual reproduction, but we can see in Figure \@ref(DAPC-RY) that there is differentiation by year. 
+The low value of $\bar{r}_d$ in the CA population may represent sexual reproduction, but we can see in Fig. \@ref(DAPC-RY) that there is differentiation by year. 
 Thus, this could also be an artifact of sampling two different populations, which is known to reduce the value of $\bar{r}_d$ [@prugnolle2010apparent].
@@
 The authors would like to thank Rebecca Higgens for technical support in generating the data for the MCG assessment, aggressiveness ratings, and genotyping; and for providing valuable insights into the historical context of the data collection and curation.
+
 We would also like to thank Denita Hadziabdic and two other anonymous reviewers for their valuable comments and insights that improved the quality of the manuscript.
-
 ## Conflicts of Interest {-}

### 2017-11-13 doc/manuscript/manuscript.Rmd
fix typo that clobbered COI
fix embarassing mis-spelling

@@
 
-The authors would like to thank Rebecca Higgens for technical support in generating the data for the MCG assessment, aggressiveness ratings, and genotyping; and for providing valuable insights into the historical context of the data collection and curation.
+The authors would like to thank Rebecca Higgins for technical support in generating the data for the MCG assessment, aggressiveness ratings, and genotyping; and for providing valuable insights into the historical context of the data collection and curation.
 
 We would also like to thank Denita Hadziabdic and two other anonymous reviewers for their valuable comments and insights that improved the quality of the manuscript.
+
 ## Conflicts of Interest {-}
@@
 
- - **James Steadman** Conceived and designed experiments, organized network of white mold screening nurseries, provided *S. sclerotiorum* isolates, edited and reviewed drafts of the paper.
+ - **James R. Steadman** Conceived and designed experiments, organized network of white mold screening nurseries, provided *S. sclerotiorum* isolates, edited and reviewed drafts of the paper.
 
@@
 Example of MCG test plates showing (A) a compatible reaction with mycelia from two strains overgrowing each other and (B) an incompatible reaction with a barrage line of dead tissue forming between the two strains.
-Photo Credit: Rebecca Higgens.
+Photo Credit: Rebecca Higgins.
 ](images/mcg-trimmed.png){#mcg-fig width=100%}

### 2017-11-13 doc/manuscript/manuscript.tex
fix typo that clobbered COI
fix embarassing mis-spelling

@@
 
-The authors would like to thank Rebecca Higgens for technical support in
+The authors would like to thank Rebecca Higgins for technical support in
 generating the data for the MCG assessment, aggressiveness ratings, and
@@
 reviewers for their valuable comments and insights that improved the
-quality of the manuscript. \#\# Conflicts of Interest \{-\}
+quality of the manuscript.
+
+\subsection*{Conflicts of Interest}\label{conflicts-of-interest}
+\addcontentsline{toc}{subsection}{Conflicts of Interest}
 
@@
 \item
-  \textbf{James Steadman} Conceived and designed experiments, organized
-  network of white mold screening nurseries, provided \emph{S.
+  \textbf{James R. Steadman} Conceived and designed experiments,
+  organized network of white mold screening nurseries, provided \emph{S.
   sclerotiorum} isolates, edited and reviewed drafts of the paper.
@@
 incompatible reaction with a barrage line of dead tissue forming between
-the two strains. Photo Credit: Rebecca Higgens.}\label{mcg-fig}
+the two strains. Photo Credit: Rebecca Higgins.}\label{mcg-fig}
 \end{figure}

### 2017-11-13 doc/manuscript/manuscript.Rmd
add R to Jim's name in author list

@@
     affiliation: 1
-  - name: James Steadman
+  - name: James R. Steadman
     email: jsteadman1@unl.edu

### 2017-11-14 README.Rmd
update README.
update instructions

@@
 
-[![Last-changedate](https://img.shields.io/badge/last%20change-`r gsub('-', '--', Sys.Date())`-brightgreen.svg)](https://github.com/everhartlab/sclerotinia-366/commits/master) [![minimal R version](https://img.shields.io/badge/R%3E%3D-3.4.1-brightgreen.svg)](https://cran.r-project.org/) [![Licence](https://img.shields.io/github/license/mashape/apistatus.svg)](http://choosealicense.com/licenses/mit/)  [![Circle CI](https://circleci.com/gh/everhartlab/sclerotinia-366.svg?style=shield&circle-token=:circle-token)](https://circleci.com/gh/everhartlab/sclerotinia-366)
+[![Last-changedate](https://img.shields.io/badge/last%20change-`r gsub('-', '--', Sys.Date())`-2F4096.svg)](https://github.com/everhartlab/sclerotinia-366/commits/master) 
+[![Licence](https://img.shields.io/badge/license-MIT%20License-2F4096.svg)](http://choosealicense.com/licenses/mit/)  
+[![minimal R version](https://img.shields.io/badge/R%3E%3D-3.4.2-2F4096.svg)](https://cran.r-project.org/) 
+[![Circle CI](https://circleci.com/gh/everhartlab/sclerotinia-366.svg?style=shield&circle-token=:circle-token)](https://circleci.com/gh/everhartlab/sclerotinia-366)
 
@@
 
-# Citation
+# Citations
 
-## Manuscript
+## Preprint
 
@@
 N., Amaradasa, B.
-S., Jhala, R., McCoy, S., Steadman, J., & Everhart, S.
+S., Jhala, R., McCoy, S., Steadman, J.
+R., & Everhart, S.
 E.
-(2017, October 3).
-Population structure and phenotypic variation of *Sclerotinia sclerotiorum* from dry bean in the United States.
+(2017, November).
+Data and analysis for population structure and phenotypic variation of *Sclerotinia sclerotiorum* from dry bean (*Phaseolus vulgaris*) in the United States.
 <http://doi.org/10.17605/OSF.IO/EJB5Y>
@@
  1. [by-year.md]
+ 1. [compare-aldrich-wolfe.md]
 
-# Reproducing the analysis
+# Analysis
 
-## Locally
+## Background
 
-This project is controlled via a [Makefile] which means that everything (analyses, tables, figures, the paper itself) is controlled via one command:
+The analysis is controlled via two docker containers:
 
-```
-make
-```
+ 1.
+[sclerotinia-366-dependencies](https://hub.docker.com/r/zkamvar/sclerotinia-366-dependencies/): defines the complete software environment used, built on top of the [rocker/verse:3.4.2](https://hub.docker.com/r/rocker/verse) container.
+**(See the [Dockerfile](https://osf.io/mwv6d/))**
+ 2.
+[sclerotinia-366](https://hub.docker.com/r/zkamvar/sclerotinia-366/) is built on top of the above container and contains the results of the analysis.
+**(See the [Dockerfile](https://osf.io/tvfju/))**
 
-This will bootstrap the installation (warning: it will update packages),
-process the data, perform the analyses, and compile the paper.
+The `sclerotinia-366-dependencies` container is regularly rebuilt on docker hub whenever `rocker/verse:3.4.2` updates and the `sclerotinia-366` container is rebuilt on [CircleCI](https://circleci.com/gh/everhartlab/sclerotinia-366) whenever the git repository is updated.
 
-Required software:
+As of this writing, the containers are up to date with R version 3.4.2 and packages downloaded from the [MRAN snapshot on 2017-10-31](https://mran.microsoft.com/snapshot/2017-10-31/).
 
- - GNU Make (If you're on Windows, you can use MinGW: http://www.mingw.org/)
- - [R (version 3.4.1 or greater)](https://r-project.org)
- - [LaTeX](https://www.latex-project.org/get)
- - [pandoc](http://pandoc.org/) (Note: pandoc ships with Rstudio)
- - [devtools](https://github.com/hadley/devtools#readme)
+Jump to [Reproduction via Docker](#reproduction-via-docker) or [Reproduction Locally](#locally).
 
-## Docker
+------
+
+## Reproduction via Docker
 
@@
 designed to capture the complete development environment of the analysis so
-that it can be accurately reproduced. The image is ~2.71Gb, so be sure that you
+that it can be accurately reproduced. The image is ~3.21Gb, so be sure that you
 have enough memory on your computer to run it.
@@
 
-> Note: both ways assume that you are in the analysis directory
-
 ### RStudio Server
 
-```
-docker run --rm -dp 8787:8787 zkamvar/sclerotinia-366:latest
+To run the docker container, make sure you have Docker running, open your terminal and type:
+
+```bash
+docker run --name ssc --rm -dp 8787:8787 -e ROOT=TRUE zkamvar/sclerotinia-366:latest
 ```
 
-This will spin up the Docker container on your machine and expose it to port 8787.
+This will first check to make sure you have the container on your machine.
+If you don't, Docker will automatically download it for you.
+It will spin up the Docker container on your machine, give it the name "ssc", and expose it to port 8787.
 You can open your browser and type `localhost:8787`, and an instance of Rstudio server will appear.
-Sign in with the username: rstudio and password: rstudio.
+Sign in with the following credentials:
+
+ - username: rstudio 
+ - password: rstudio.
 
@@
 From here you can re-run the analyses to your heart's content. 
+**Don't forget to stop the container when you are finished:**
+
+```bash
+docker stop ssc
+```
 
@@
 
-If you don't want to pull from docker hub, you can build the container.
+If you don't want to pull from docker hub, you can build the container locally.
+First, download the repository
 
-```sh
+```bash
+git clone https://github.com/everhartlab/sclerotinia-366.git
+cd sclerotinia-366/
 docker build -t sclerotinia-366 .
@@
 
+## Locally
+
+This project is controlled via a [Makefile] which means that everything (analyses, tables, figures, the paper itself) is controlled via one command:
+
+```
+make
+```
+
+This will bootstrap the installation (warning: it will update packages),
+process the data, perform the analyses, and compile the paper.
+
+> Note: This analysis is only guaranteed to work with [the stated software environment](#packages-used).
+
+Required software:
+
+ - GNU Make (If you're on Windows, you can use MinGW: http://www.mingw.org/)
+ - [R (version 3.4.1 or greater)](https://r-project.org)
+ - [LaTeX](https://www.latex-project.org/get)
+ - [pandoc](http://pandoc.org/) (Note: pandoc ships with Rstudio)
+ - [devtools](https://github.com/hadley/devtools#readme)
+
 [Makefile]: Makefile
@@
 [by-year.md]: results/by-year.md
+[compare-aldrich-wolfe.md]: results/compare-aldrich-wolfe.md
 

### 2017-11-14 doc/manuscript/manuscript.Rmd
Final minor minor changes

@@
 \@ref(mcg-fig)A) [@leslie1993fungal; @schafer2006optimized].
-Because of the genetic connection to vegetative compatibility, two isolates that are mycelially compatible were considered clones [@leslie1993fungal]; but correlation with genetic markers, such as microsatellites, have shown mixed results [@ford1995heterokaryon; @micali2003independence; @jo2008reassessment; @attanayake2012genetic; @papaioannou2014barrage; @lehner2017independently].
+Because of the genetic connection to vegetative compatibility, two isolates that are mycelially compatible were considered clones [@leslie1993fungal]; but correlation with genetic markers, such as microsatellites, have shown inconsistent results [@ford1995heterokaryon; @micali2003independence; @jo2008reassessment; @attanayake2012genetic; @papaioannou2014barrage; @lehner2017independently].
 Thus, the relationship between mycelial compatibility groups and clonal genotypes remains unclear.
 
-In this study, we analyze and characterize the genetic and phenotypic diversity of 366 *S.
+In the present study, we analyze and characterize the genetic and phenotypic diversity of 366 *S.
 sclerotiorum* isolates collected between 2003 and 2012 from dry bean cultivars among different geographic locations in the Australia, France, Mexico, and the United States.
-We wanted to know if the *S. sclerotiorum* populations from white mold screening nurseries were representative of the fields within the same region. 
+We wanted to know if the *S. sclerotiorum* populations from white mold screening nurseries were representative of the producer fields within the same region. 
 As these nurseries were not treated with any chemical or cultural control of white mold, we hypothesized that these nurseries would represent the natural population of *S. sclerotiorum*.
@@
 
-## Mycelial Compatibility {-}
+## Mycelial compatibility {-}
 
 MCG was determined as described previously through co-culturing pairs of 2-day-old isolates 2.5 cm apart on Diana Sermons (DS) Medium (Fig. \@ref(mcg-fig)) [@cubeta2001mycelial]. 
-Incompatibility of different MCGs resulted in formation of a barrage line accompanied by formation of sclerotia on either side of the barrage line, indicating the limits of the isolates' growth [@otto-hanson2011variation; @kohn1990mycelial; @leslie1993fungal].
+Incompatibility of different MCGs resulted in formation of a barrage line accompanied by formation of sclerotia on either side of the barrage line, indicating the limits of mycelial growth [@otto-hanson2011variation; @kohn1990mycelial; @leslie1993fungal].
 Isolates were compared in a pairwise manner for each site and then representatives among sites were compared to determine mycelial compatibility groups by scoring compatible and incompatible interactions [@otto-hanson2011variation].
@@
 Briefly, the straw test uses 21-day-old G122 plants as the host in a greenhouse setting. 
-Clear drinking straws cut to 2.5 cm and sealed were used to place two mycelial plugs of inoculum on the host plant after removing plant growth beyond 2.5 cm above the fourth node.
+Clear drinking straws cut to 2.5 cm and heat sealed were used to place two mycelial plugs of inoculum on the host plant after removing plant growth beyond 2.5 cm above the fourth node.
 Measurements of the necrotic lesion were taken 8 days later using the Modified Petzoldt and Dickson scale of 1--9, where 1 is no disease and 9 is plant death [@petzoldt1996straw; @teran2006modified].
@@
 
-## Data processing and Analysis {-}
+## Data processing and analysis {-}
 
@@
 
-## Assessing Importance of Variables  {-}
-
-### Distance-based Redundancy Analysis {-}
+## Assessing importance of variables  {-}
+
+### Distance-based redundancy analysis {-}
 
@@
 
-### Aggressiveness Assessment {-}
+### Aggressiveness assessment {-}
 
@@
 
-### Correlating Multilocus Haplotypes with Mycelial Compatibility Groups {-}
+### Correlating multilocus haplotypes with mycelial compatibility groups {-}
 
@@
 To identify clusters of MLHs and MCGs within the network, we used the cluster walktrap community detection algorithm as implemented in the `cluster_walktrap()` function in *igraph* version `r packageVersion("igraph")` [@csardi2006igraph; @pons2006computing].
-This algorithm attempts to define clusters of nodes by starting at a random nodes and performing short, random "walks" along the edges between nodes, assuming that these walks would stay within clusters.
+This algorithm attempts to define clusters of nodes by starting at a random node and performing short, random "walks" along the edges between nodes, assuming that these walks would stay within clusters.
 For this analysis, we set the number of steps within a walk to four and allowed the algorithm to use the edge weights in determining the path.
@@
 
-## Genetic Diversity {-}
-
-### Population Differentiation {-}
+## Genetic diversity {-}
+
+### Population differentiation {-}
 
@@
 
-### Analysis of Shared Multilocus Haplotypes {-}
+### Analysis of shared multilocus haplotypes {-}
 
@@
 With the 11 loci used in the analyses (Table \@ref(tab:locus-stats)), we observed a total of 165 MLHs (215 with 16 loci).
-These 11 loci are found on 7 chromosomes in the *S. sclerotiorum* genome with a minimum distance of 55Kbp between two loci on the same chromosome.
+These 11 loci are located on 7 chromosomes in the *S. sclerotiorum* genome with a minimum distance of 55Kbp between two loci on the same chromosome.
 Over 50% of the isolates came from four states, MI (`r pop_size[["MI"]]`), ND (`r pop_size[["ND"]]`), WA (`r pop_size[["WA"]]`), NE (`r pop_size[["NE"]]`). 
@@
 After clone-correction on the hierarchy of Region/Source/Host/Year, a total of `r nInd(dat11) - nInd(dat11cc)` isolates were removed from the data set, resulting in `r nInd(dat11cc)` isolates representing `r nmll(dat11cc)` MLHs that were used in subsequent analyses (Table \@ref(tab:iatab)).
-The results showed that, in terms of genotypic diversity (*H*, *G*, and $\lambda$), WA was the most diverse population with both *G* (`r signif(setNames(gdt$G, gdt$Pop)["WA"], 3)`) and $e^H$ (`r signif(setNames(exp(gdt$H), gdt$Pop)["WA"], 3)`) being close to the observed number of MLGs (`r setNames(gdt$MLG, gdt$Pop)["WA"]`).
+The results showed that, in terms of genotypic diversity (*H*, *G*, and $\lambda$), WA was the most diverse population with both *G* (`r signif(setNames(gdt$G, gdt$Pop)["WA"], 3)`) and $e^H$ (`r signif(setNames(exp(gdt$H), gdt$Pop)["WA"], 3)`) being close to the observed number of MLHs (`r setNames(gdt$MLG, gdt$Pop)["WA"]`).
 This indicated that there are few duplicated genotypes in WA (Table \@ref(tab:iatab)).
@@
 Moreover, Mexico had the lowest value for *h*, which is a measure of allelic diversity.
-Nearly all populations showed evidence for linkage (Table \@ref(tab:iatab)), which serves as evidence for clonal reproduction or other forms of non-random mating. 
-The only exceptions are CA (*P* = `r gdtrd[["CA"]]`) and Australia (*P* = `r gdtrd[["Australia"]]`).
+Nearly all populations showed evidence of linkage (Table \@ref(tab:iatab)), which serves as evidence for clonal reproduction or other forms of non-random mating. 
+The only exceptions were CA (*P* = `r gdtrd[["CA"]]`) and Australia (*P* = `r gdtrd[["Australia"]]`).
 Both of these populations showed only moderate significance with $\bar{r}_d$ values of 0.03 and 0.12, respectively.
@@
 
-## Variable Assessment {-}
-
-### Variable Contributions {-}
+## Variable assessment {-}
+
+### Variable contributions {-}
 
@@
 
-Agressiveness of the isolates ranged from `r paste(range(other(dat11)$meta$Severity), collapse = " to ")` with a mean of `r signif(mean(other(dat11)$meta$Severity), 3)` and median of `r signif(median(other(dat11)$meta$Severity), 3)`.
+Aggressiveness of the isolates ranged from `r paste(range(other(dat11)$meta$Severity), collapse = " to ")` with a mean of `r signif(mean(other(dat11)$meta$Severity), 3)` and median of `r signif(median(other(dat11)$meta$Severity), 3)`.
 The group mean averages were `r sprintf("%0.2f, %0.2f, and %0.2f", Raggmean, Maggmean, Haggmean)` for Region, MCG, and MLH, respectively.
@@
 
-### Correlation of Mulitlocus Haplotypes and Mycelial Compatibility Groups {-}
+### Correlation of mulitlocus haplotypes and mycelial compatibility groups {-}
 
@@
 When we assessed the number of times two different MLHs that are in the same MCG, considering these as potential heterothallic pairings that could result in sexual recombination, we found an average of `r signif(mean(in_reach$N), 3)` potential heterothallic parings per MLH.
-Representing just four isolates, 'MLH 75' had `r big_haplotype$N` neighbors that shared the same mycelial compatibility group (Fig. \@ref(biggraph), \@ref(fullgraph)). 
+Representing just four isolates, 'MLH 75' had `r big_haplotype$N` neighbors that shared the same MCG (Fig. \@ref(biggraph), \@ref(fullgraph)). 
 Overall, there was no clear pattern to the association between MLH and MCGs.
@@
 
-## Structure of Shared Multilocus Haplotypes {-}
+## Structure of shared multilocus haplotypes {-}
 
@@
 We had performed an analysis on a network where the connections represented shared MLHs across populations, weighted by $1 - P_{sex}$ (Fig. \@ref(community-graph), Table \@ref(tab:mlg-table)). 
-Community analysis of the MLHs shared between populations revealed 4 communities with a modularity of 0.17: A coastal community (CA, OR, WA, and NY), a midwest community (CO, ND, NE, MI), and an international community (Australia, France, MN).
+Community analysis of the MLHs shared between populations revealed 4 communities with a modularity of 0.17: A coastal community (CA, OR, WA, and NY), a Midwest community (CO, ND, NE, MI), and an international community (Australia, France, MN).
 Although analysis with 16 loci resulted in the removal of the NY node because it no longer shared a haplotype with OR, the same overall community structure was present with a modularity of 0.2 (Fig.
@@
 
-## Population Differentiation {-}
-
-### Analysis of Molecular Variance {-}
-
-The analysis of molecular variance (AMOVA) for clone-corrected samples over the hierarchy of Region, Source, and Year showed significant variation between Regions and Years, but no significant variation between wmn and producer fields (Table \@ref(tab:amova)).
+## Population differentiation {-}
+
+### Analysis of molecular variance {-}
+
+The AMOVA for clone-corrected samples over the hierarchy of Region, Source, and Year showed significant variation between Regions and Years, but no significant variation between wmn and producer fields (Table \@ref(tab:amova)).
 In contrast, when we compared the three cultivars, Beryl, Bunsi, and G122, we found no significant differentiation (See section on 'Host Differentiation' in the wmn-differentiation.md[^2] file in the supplemental files [@kamvar2017data]).
@@
 
-Table: (\#tab:amova) Comparison of populations in the white mold screening nurseries (wmn) and producer fields using an analysis of molecular variance on Bruvo's genetic distance showing no apparent differentiation between wmn and other sources.
+Table: (\#tab:amova) Comparison of populations in the white mold screening nurseries (wmn) and producer fields using an analysis of molecular variance (AMOVA) on Bruvo's genetic distance showing no apparent differentiation between wmn and other sources.
 The hierarchy was constructed as Source/Region where source is defined as belonging to a wmn or producer field.
@@
 
-### Discriminant Analysis of Principal Components {-}
-
-Discriminant Analysis of Principal Components (DAPC) was performed by grouping Region with the first `r dapc_region$n.pca` principal components, representing `r prettify(dapc_region$var)`% of the total variance.
+### Discriminant analysis of principal components {-}
+
+DAPC was performed by grouping Region with the first `r dapc_region$n.pca` principal components, representing `r prettify(dapc_region$var)`% of the total variance. 
 The first discriminant axis (representing `r paxis(dapc_region)`% of the discriminatory power) separated the centroid for the Mexico isolates from the rest of the data, indicating strong differentiation (Fig.
@@
 ![
-Discriminant Analysis of Principal Components on regions showing that Mexico is differentiated from other populations. 
+Discriminant Analysis of Principal Components (DAPC) on regions showing that Mexico is differentiated from other populations. 
 **A)** Scatter plot of first two components from DAPC.
@@
 ![
-Scatter plot of Discriminant Analysis of Principal Components on Regions and Years showing non-differentiated temporal variation NE and MI and temporal variation in WA and CA. 
+Scatter plot of Discriminant Analysis of Principal Components (DAPC) on Regions and Years showing non-differentiated temporal variation NE and MI and temporal variation in WA and CA. 
 Points (text labels) represent observed individuals connected to the population centroids with ellipses representing a 66% confidence interval for a normal distribution. 
@@
 However, all three of these alleles, i) conform to the expected stepwise mutation model [@bruvo2004simple] and ii) are at or near the extremes of the total range (except for `r glue::glue("allele {mx_alleles[3]} at locus {mx_loci[3]}")`).
-Moreover, the fact that we find three alleles at three independent loci segregating for Mexico suggests that the pattern separating these populations from the others was not an artifact. 
+Moreover, the fact that we find three alleles at three independent loci segregating the Mexican genotypes suggests that the pattern separating these populations from the others was not an artifact. 
 We believe that the differences in populations observed from Mexico may be due to differences in climate that allow greater diversification via sexual outcrossing.
@@
 
-The results from our shared haplotype analysis show several populations with at least one haplotype between them, except for Mexico and two states that had fewer than three samples each (Fig.
+The results from our shared haplotype analysis showed several populations with at least one haplotype between them, except for Mexico and two states that had fewer than three samples each (Fig.
 \@ref(community-graph)).
 Our network-based approach by treating the haplotypes as edges and weighting each edge with the inverse of $P_{sex}$ treated the edges as springs connecting the populations with the strength proportional to the probability of obtaining the same haplotype as a clone.
-This allowed us to use a graph walking algorithm to see how close the populations were simply based off of the proportion of clones they shared.
+This allowed us to use a graph walking algorithm to see how close the populations were, simply based off of the proportion of clones they shared.
 The most abundant haplotype was shared across four populations, but its high value of $P_{sex}$ meant that it did not contribute significantly to the overall structure. 
@@
 Thus, we hypothesize that shared haplotypes between populations may arise due to transmission events of seed or sclerotia.
-This could explain the fact that we see shared genotypes with low $P_{sex}$ values shared between Australia, France, and the United States.
+This could explain the fact that we see shared haplotypes with low $P_{sex}$ values shared between Australia, France, and the United States.
 While we speculate that these transmission events are rare due to the genetic structuring by Region, these results suggest that seedborne infections may indeed reflect a source of inoculum.
@@
 While our protocol for assessing MCGs utilized Diana Sermons Medium [@cubeta2001mycelial] as compared to Patterson's Medium or Potato Dextrose Agar [@schafer2006optimized] for the MCG reactions, the patterns we observe are not dissimilar from what have previously been reported in the literature.
-It has been demonstrated in several Ascomycetes---including *Neurospora crassa* [@micali2003independence], *Sclerotinia homoecarpa* [@jo2008reassessment], *Verticillium dahliae* [@papaioannou2014barrage], and *S.
+It has been demonstrated in several Ascomycetes---including *Neurospora crassa* [@micali2003independence], *Sclerotinia homoeocarpa* [@jo2008reassessment], *Verticillium dahliae* [@papaioannou2014barrage], and *S.
 sclerotiorum* [@ford1995heterokaryon]---that barrage reactions are independent from stable anastomosis.
@@
 
- - **Bimal Sajeewa Amaradasa** analyzed the data, contributed analysis tools, wrote drafts of the paper, edited and reviewed drafts of the paper.
+ - **B. Sajeewa Amaradasa** analyzed the data, contributed analysis tools, wrote drafts of the paper, edited and reviewed drafts of the paper.
 

### 2017-11-15 doc/manuscript/manuscript.tex
add R to Jim's name in author list
render manuscript for merging

@@
 
-\author[1]{James Steadman}
+\author[1]{James R. Steadman}
 
@@
 correlation with genetic markers, such as microsatellites, have shown
-mixed results (Ford et al., 1995; Micali \& Smith, 2003; Jo et al.,
-2008; Attanayake et al., 2012; Papaioannou \& Typas, 2014; Lehner et
-al., 2017). Thus, the relationship between mycelial compatibility groups
-and clonal genotypes remains unclear.
-
-In this study, we analyze and characterize the genetic and phenotypic
-diversity of 366 \emph{S. sclerotiorum} isolates collected between 2003
-and 2012 from dry bean cultivars among different geographic locations in
-the Australia, France, Mexico, and the United States. We wanted to know
-if the \emph{S. sclerotiorum} populations from white mold screening
-nurseries were representative of the fields within the same region. As
-these nurseries were not treated with any chemical or cultural control
-of white mold, we hypothesized that these nurseries would represent the
-natural population of \emph{S. sclerotiorum}. Furthermore, we wanted to
-investigate the potential effect of cultivar on genetic diversity of the
-pathogen by assessing three dry bean cultivars with different levels of
-resistance, Beryl (great northern bean, susceptible), Bunsi (navy bean,
-low susceptibility), and G122 (cranberry bean, partial resistance)
-(Otto-Hanson et al., 2011). We additionally wanted to determine
-categorical or phenotypic variables that best predicted genetic
-structure and if there was correlation between multilocus haplotype and
-mycelial compatibility group. Knowing what variables predict genetic
-structure can help direct breeding efforts. By investigating these aims,
-we will effectively describe the population structure of \emph{S.
-sclerotiorum} in the USA and make available our database of isolates for
-use in future dry bean breeding efforts.
+inconsistent results (Ford et al., 1995; Micali \& Smith, 2003; Jo et
+al., 2008; Attanayake et al., 2012; Papaioannou \& Typas, 2014; Lehner
+et al., 2017). Thus, the relationship between mycelial compatibility
+groups and clonal genotypes remains unclear.
+
+In the present study, we analyze and characterize the genetic and
+phenotypic diversity of 366 \emph{S. sclerotiorum} isolates collected
+between 2003 and 2012 from dry bean cultivars among different geographic
+locations in the Australia, France, Mexico, and the United States. We
+wanted to know if the \emph{S. sclerotiorum} populations from white mold
+screening nurseries were representative of the producer fields within
+the same region. As these nurseries were not treated with any chemical
+or cultural control of white mold, we hypothesized that these nurseries
+would represent the natural population of \emph{S. sclerotiorum}.
+Furthermore, we wanted to investigate the potential effect of cultivar
+on genetic diversity of the pathogen by assessing three dry bean
+cultivars with different levels of resistance, Beryl (great northern
+bean, susceptible), Bunsi (navy bean, low susceptibility), and G122
+(cranberry bean, partial resistance) (Otto-Hanson et al., 2011). We
+additionally wanted to determine categorical or phenotypic variables
+that best predicted genetic structure and if there was correlation
+between multilocus haplotype and mycelial compatibility group. Knowing
+what variables predict genetic structure can help direct breeding
+efforts. By investigating these aims, we will effectively describe the
+population structure of \emph{S. sclerotiorum} in the USA and make
+available our database of isolates for use in future dry bean breeding
+efforts.
 
@@
 
-\subsection*{Mycelial Compatibility}\label{mycelial-compatibility}
-\addcontentsline{toc}{subsection}{Mycelial Compatibility}
+\subsection*{Mycelial compatibility}\label{mycelial-compatibility}
+\addcontentsline{toc}{subsection}{Mycelial compatibility}
 
@@
 sclerotia on either side of the barrage line, indicating the limits of
-the isolates' growth (Kohn et al., 1990; Leslie, 1993; Otto-Hanson et
-al., 2011). Isolates were compared in a pairwise manner for each site
-and then representatives among sites were compared to determine mycelial
+mycelial growth (Kohn et al., 1990; Leslie, 1993; Otto-Hanson et al.,
+2011). Isolates were compared in a pairwise manner for each site and
+then representatives among sites were compared to determine mycelial
 compatibility groups by scoring compatible and incompatible interactions
@@
 uses 21-day-old G122 plants as the host in a greenhouse setting. Clear
-drinking straws cut to 2.5 cm and sealed were used to place two mycelial
-plugs of inoculum on the host plant after removing plant growth beyond
-2.5 cm above the fourth node. Measurements of the necrotic lesion were
-taken 8 days later using the Modified Petzoldt and Dickson scale of
+drinking straws cut to 2.5 cm and heat sealed were used to place two
+mycelial plugs of inoculum on the host plant after removing plant growth
+beyond 2.5 cm above the fourth node. Measurements of the necrotic lesion
+were taken 8 days later using the Modified Petzoldt and Dickson scale of
 1--9, where 1 is no disease and 9 is plant death (Petzoldt \& Dickson,
@@
 \subsection*{Data processing and
-Analysis}\label{data-processing-and-analysis}
-\addcontentsline{toc}{subsection}{Data processing and Analysis}
+analysis}\label{data-processing-and-analysis}
+\addcontentsline{toc}{subsection}{Data processing and analysis}
 
@@
 
-\subsection*{Assessing Importance of
-Variables}\label{assessing-importance-of-variables}
-\addcontentsline{toc}{subsection}{Assessing Importance of Variables}
-
-\subsubsection*{Distance-based Redundancy
-Analysis}\label{distance-based-redundancy-analysis}
-\addcontentsline{toc}{subsubsection}{Distance-based Redundancy Analysis}
+\subsection*{Assessing importance of
+variables}\label{assessing-importance-of-variables}
+\addcontentsline{toc}{subsection}{Assessing importance of variables}
+
+\subsubsection*{Distance-based redundancy
+analysis}\label{distance-based-redundancy-analysis}
+\addcontentsline{toc}{subsubsection}{Distance-based redundancy analysis}
 
@@
 \subsubsection*{Aggressiveness
-Assessment}\label{aggressiveness-assessment}
-\addcontentsline{toc}{subsubsection}{Aggressiveness Assessment}
+assessment}\label{aggressiveness-assessment}
+\addcontentsline{toc}{subsubsection}{Aggressiveness assessment}
 
@@
 
-\subsubsection*{Correlating Multilocus Haplotypes with Mycelial
-Compatibility
-Groups}\label{correlating-multilocus-haplotypes-with-mycelial-compatibility-groups}
-\addcontentsline{toc}{subsubsection}{Correlating Multilocus Haplotypes
-with Mycelial Compatibility Groups}
+\subsubsection*{Correlating multilocus haplotypes with mycelial
+compatibility
+groups}\label{correlating-multilocus-haplotypes-with-mycelial-compatibility-groups}
+\addcontentsline{toc}{subsubsection}{Correlating multilocus haplotypes
+with mycelial compatibility groups}
 
@@
 (Csardi \& Nepusz, 2006; Pons \& Latapy, 2006). This algorithm attempts
-to define clusters of nodes by starting at a random nodes and performing
+to define clusters of nodes by starting at a random node and performing
 short, random ``walks'' along the edges between nodes, assuming that
@@
 
-\subsection*{Genetic Diversity}\label{genetic-diversity}
-\addcontentsline{toc}{subsection}{Genetic Diversity}
+\subsection*{Genetic diversity}\label{genetic-diversity}
+\addcontentsline{toc}{subsection}{Genetic diversity}
 
 \subsubsection*{Population
-Differentiation}\label{population-differentiation}
-\addcontentsline{toc}{subsubsection}{Population Differentiation}
+differentiation}\label{population-differentiation}
+\addcontentsline{toc}{subsubsection}{Population differentiation}
 
@@
 
-\subsubsection*{Analysis of Shared Multilocus
-Haplotypes}\label{analysis-of-shared-multilocus-haplotypes}
-\addcontentsline{toc}{subsubsection}{Analysis of Shared Multilocus
-Haplotypes}
+\subsubsection*{Analysis of shared multilocus
+haplotypes}\label{analysis-of-shared-multilocus-haplotypes}
+\addcontentsline{toc}{subsubsection}{Analysis of shared multilocus
+haplotypes}
 
@@
 \ref{tab:locus-stats}), we observed a total of 165 MLHs (215 with 16
-loci). These 11 loci are found on 7 chromosomes in the \emph{S.
+loci). These 11 loci are located on 7 chromosomes in the \emph{S.
 sclerotiorum} genome with a minimum distance of 55Kbp between two loci
@@
 population with both \emph{G} (54.3) and \(e^H\) (55.3) being close to
-the observed number of MLGs (56). This indicated that there are few
+the observed number of MLHs (56). This indicated that there are few
 duplicated genotypes in WA (Table \ref{tab:iatab}). A more useful metric
@@
 Mexico had the lowest value for \emph{h}, which is a measure of allelic
-diversity. Nearly all populations showed evidence for linkage (Table
+diversity. Nearly all populations showed evidence of linkage (Table
 \ref{tab:iatab}), which serves as evidence for clonal reproduction or
-other forms of non-random mating. The only exceptions are CA (\emph{P} =
-0.043) and Australia (\emph{P} = 0.052). Both of these populations
+other forms of non-random mating. The only exceptions were CA (\emph{P}
+= 0.043) and Australia (\emph{P} = 0.052). Both of these populations
 showed only moderate significance with \(\bar{r}_d\) values of 0.03 and
@@
 
-\subsection*{Variable Assessment}\label{variable-assessment}
-\addcontentsline{toc}{subsection}{Variable Assessment}
-
-\subsubsection*{Variable Contributions}\label{variable-contributions}
-\addcontentsline{toc}{subsubsection}{Variable Contributions}
+\subsection*{Variable assessment}\label{variable-assessment}
+\addcontentsline{toc}{subsection}{Variable assessment}
+
+\subsubsection*{Variable contributions}\label{variable-contributions}
+\addcontentsline{toc}{subsubsection}{Variable contributions}
 
@@
 
-Agressiveness of the isolates ranged from 1.4 to 7.9 with a mean of 5.02
-and median of 4.85. The group mean averages were 4.88, 5.13, and 5.19
-for Region, MCG, and MLH, respectively. A strip plot showing the
+Aggressiveness of the isolates ranged from 1.4 to 7.9 with a mean of
+5.02 and median of 4.85. The group mean averages were 4.88, 5.13, and
+5.19 for Region, MCG, and MLH, respectively. A strip plot showing the
 distribution of severity across these three variables simultaneously can
@@
 
-\subsubsection*{Correlation of Mulitlocus Haplotypes and Mycelial
-Compatibility
-Groups}\label{correlation-of-mulitlocus-haplotypes-and-mycelial-compatibility-groups}
-\addcontentsline{toc}{subsubsection}{Correlation of Mulitlocus
-Haplotypes and Mycelial Compatibility Groups}
+\subsubsection*{Correlation of mulitlocus haplotypes and mycelial
+compatibility
+groups}\label{correlation-of-mulitlocus-haplotypes-and-mycelial-compatibility-groups}
+\addcontentsline{toc}{subsubsection}{Correlation of mulitlocus
+haplotypes and mycelial compatibility groups}
 
@@
 of 14.3 potential heterothallic parings per MLH. Representing just four
-isolates, `MLH 75' had 57 neighbors that shared the same mycelial
-compatibility group (Fig. \ref{biggraph}, \ref{fullgraph}). Overall,
-there was no clear pattern to the association between MLH and MCGs.
+isolates, `MLH 75' had 57 neighbors that shared the same MCG (Fig.
+\ref{biggraph}, \ref{fullgraph}). Overall, there was no clear pattern to
+the association between MLH and MCGs.
 
@@
 
-\subsection*{Structure of Shared Multilocus
-Haplotypes}\label{structure-of-shared-multilocus-haplotypes}
-\addcontentsline{toc}{subsection}{Structure of Shared Multilocus
-Haplotypes}
+\subsection*{Structure of shared multilocus
+haplotypes}\label{structure-of-shared-multilocus-haplotypes}
+\addcontentsline{toc}{subsection}{Structure of shared multilocus
+haplotypes}
 
@@
 with a modularity of 0.17: A coastal community (CA, OR, WA, and NY), a
-midwest community (CO, ND, NE, MI), and an international community
+Midwest community (CO, ND, NE, MI), and an international community
 (Australia, France, MN). Although analysis with 16 loci resulted in the
@@
 \subsection*{Population
-Differentiation}\label{population-differentiation-1}
-\addcontentsline{toc}{subsection}{Population Differentiation}
-
-\subsubsection*{Analysis of Molecular
-Variance}\label{analysis-of-molecular-variance}
-\addcontentsline{toc}{subsubsection}{Analysis of Molecular Variance}
-
-The analysis of molecular variance (AMOVA) for clone-corrected samples
-over the hierarchy of Region, Source, and Year showed significant
-variation between Regions and Years, but no significant variation
-between wmn and producer fields (Table \ref{tab:amova}). In contrast,
-when we compared the three cultivars, Beryl, Bunsi, and G122, we found
-no significant differentiation (See section on `Host Differentiation' in
-the wmn-differentiation.md\footnote{Direct link:
+differentiation}\label{population-differentiation-1}
+\addcontentsline{toc}{subsection}{Population differentiation}
+
+\subsubsection*{Analysis of molecular
+variance}\label{analysis-of-molecular-variance}
+\addcontentsline{toc}{subsubsection}{Analysis of molecular variance}
+
+The AMOVA for clone-corrected samples over the hierarchy of Region,
+Source, and Year showed significant variation between Regions and Years,
+but no significant variation between wmn and producer fields (Table
+\ref{tab:amova}). In contrast, when we compared the three cultivars,
+Beryl, Bunsi, and G122, we found no significant differentiation (See
+section on `Host Differentiation' in the
+wmn-differentiation.md\footnote{Direct link:
   \url{https://github.com/everhartlab/sclerotinia-366/blob/master/results/wmn-differentiation.md\#host-differentiation}}
@@
 screening nurseries (wmn) and producer fields using an analysis of
-molecular variance on Bruvo's genetic distance showing no apparent
-differentiation between wmn and other sources. The hierarchy was
-constructed as Source/Region where source is defined as belonging to a
-wmn or producer field. Bold \(\Phi\) values indicate significant
+molecular variance (AMOVA) on Bruvo's genetic distance showing no
+apparent differentiation between wmn and other sources. The hierarchy
+was constructed as Source/Region where source is defined as belonging to
+a wmn or producer field. Bold \(\Phi\) values indicate significant
 difference (\emph{P} \textless{} 0.05). S.S. = Sum of Squares, d.f. =
@@
 
-\subsubsection*{Discriminant Analysis of Principal
-Components}\label{discriminant-analysis-of-principal-components}
-\addcontentsline{toc}{subsubsection}{Discriminant Analysis of Principal
-Components}
-
-Discriminant Analysis of Principal Components (DAPC) was performed by
-grouping Region with the first 21 principal components, representing
-88.1\% of the total variance. The first discriminant axis (representing
-63.9\% of the discriminatory power) separated the centroid for the
-Mexico isolates from the rest of the data, indicating strong
-differentiation (Fig. \ref{DAPC}b). The second discriminant axis,
-representing 10.8\% of the discriminatory power, separated the centroid
-for the CA isolates. The mean population assignment probabilities for
-all populations with n \textgreater{} 10 showed that only isolates from
-Mexico, CA, and France had \textgreater{} 50\% probabilities of being
-reassigned to their source populations (Fig. \ref{DAPC}a).
+\subsubsection*{Discriminant analysis of principal
+components}\label{discriminant-analysis-of-principal-components}
+\addcontentsline{toc}{subsubsection}{Discriminant analysis of principal
+components}
+
+DAPC was performed by grouping Region with the first 21 principal
+components, representing 88.1\% of the total variance. The first
+discriminant axis (representing 63.9\% of the discriminatory power)
+separated the centroid for the Mexico isolates from the rest of the
+data, indicating strong differentiation (Fig. \ref{DAPC}b). The second
+discriminant axis, representing 10.8\% of the discriminatory power,
+separated the centroid for the CA isolates. The mean population
+assignment probabilities for all populations with n \textgreater{} 10
+showed that only isolates from Mexico, CA, and France had \textgreater{}
+50\% probabilities of being reassigned to their source populations (Fig.
+\ref{DAPC}a).
 
@@
 \includegraphics[width=1.00000\textwidth]{../../results/figures/publication/DAPC.pdf}
-\caption{Discriminant Analysis of Principal Components on regions
+\caption{Discriminant Analysis of Principal Components (DAPC) on regions
 showing that Mexico is differentiated from other populations.
@@
 \caption{Scatter plot of Discriminant Analysis of Principal Components
-on Regions and Years showing non-differentiated temporal variation NE
-and MI and temporal variation in WA and CA. Points (text labels)
-represent observed individuals connected to the population centroids
-with ellipses representing a 66\% confidence interval for a normal
-distribution. The center of each component is represented as black grid
-lines. A more detailed view is shown in Fig.
+(DAPC) on Regions and Years showing non-differentiated temporal
+variation NE and MI and temporal variation in WA and CA. Points (text
+labels) represent observed individuals connected to the population
+centroids with ellipses representing a 66\% confidence interval for a
+normal distribution. The center of each component is represented as
+black grid lines. A more detailed view is shown in Fig.
 \ref{DAPC-RY-FULL}.}\label{DAPC-RY}
@@
 total range (except for allele 372 at locus 9-2). Moreover, the fact
-that we find three alleles at three independent loci segregating for
-Mexico suggests that the pattern separating these populations from the
-others was not an artifact. We believe that the differences in
+that we find three alleles at three independent loci segregating the
+Mexican genotypes suggests that the pattern separating these populations
+from the others was not an artifact. We believe that the differences in
 populations observed from Mexico may be due to differences in climate
@@
 
-The results from our shared haplotype analysis show several populations
-with at least one haplotype between them, except for Mexico and two
-states that had fewer than three samples each (Fig.
+The results from our shared haplotype analysis showed several
+populations with at least one haplotype between them, except for Mexico
+and two states that had fewer than three samples each (Fig.
 \ref{community-graph}). Our network-based approach by treating the
@@
 haplotype as a clone. This allowed us to use a graph walking algorithm
-to see how close the populations were simply based off of the proportion
-of clones they shared. The most abundant haplotype was shared across
-four populations, but its high value of \(P_{sex}\) meant that it did
-not contribute significantly to the overall structure. The graph walking
-algorithm was able to divide the network into three groups, but had a
-modularity of 0.17, which indicates that the groups are only weakly
-differentiated.
+to see how close the populations were, simply based off of the
+proportion of clones they shared. The most abundant haplotype was shared
+across four populations, but its high value of \(P_{sex}\) meant that it
+did not contribute significantly to the overall structure. The graph
+walking algorithm was able to divide the network into three groups, but
+had a modularity of 0.17, which indicates that the groups are only
+weakly differentiated.
 
@@
 between populations may arise due to transmission events of seed or
-sclerotia. This could explain the fact that we see shared genotypes with
-low \(P_{sex}\) values shared between Australia, France, and the United
-States. While we speculate that these transmission events are rare due
-to the genetic structuring by Region, these results suggest that
-seedborne infections may indeed reflect a source of inoculum. This may,
-in turn increase the risk of introducing new sources of genetic
+sclerotia. This could explain the fact that we see shared haplotypes
+with low \(P_{sex}\) values shared between Australia, France, and the
+United States. While we speculate that these transmission events are
+rare due to the genetic structuring by Region, these results suggest
+that seedborne infections may indeed reflect a source of inoculum. This
+may, in turn increase the risk of introducing new sources of genetic
 variation through potential outcrossing events.
@@
 Ascomycetes---including \emph{Neurospora crassa} (Micali \& Smith,
-2003), \emph{Sclerotinia homoecarpa} (Jo et al., 2008),
+2003), \emph{Sclerotinia homoeocarpa} (Jo et al., 2008),
 \emph{Verticillium dahliae} (Papaioannou \& Typas, 2014), and \emph{S.
@@
 \item
-  \textbf{Bimal Sajeewa Amaradasa} analyzed the data, contributed
-  analysis tools, wrote drafts of the paper, edited and reviewed drafts
-  of the paper.
+  \textbf{B. Sajeewa Amaradasa} analyzed the data, contributed analysis
+  tools, wrote drafts of the paper, edited and reviewed drafts of the
+  paper.
 \item

### 2017-11-15 README.Rmd
update (partially) to everhartlab dockerhub

@@
  1.
-[sclerotinia-366-dependencies](https://hub.docker.com/r/zkamvar/sclerotinia-366-dependencies/): defines the complete software environment used, built on top of the [rocker/verse:3.4.2](https://hub.docker.com/r/rocker/verse) container.
+[sclerotinia-366-dependencies](https://hub.docker.com/r/everhartlab/sclerotinia-366-dependencies/): defines the complete software environment used, built on top of the [rocker/verse:3.4.2](https://hub.docker.com/r/rocker/verse) container.
 **(See the [Dockerfile](https://osf.io/mwv6d/))**
  2.
-[sclerotinia-366](https://hub.docker.com/r/zkamvar/sclerotinia-366/) is built on top of the above container and contains the results of the analysis.
+[sclerotinia-366](https://hub.docker.com/r/everhartlab/sclerotinia-366/) is built on top of the above container and contains the results of the analysis.
 **(See the [Dockerfile](https://osf.io/tvfju/))**
@@
 Pulling is by far the quickest way to do this.
-The docker container is located at https://hub.docker.com/r/zkamvar/sclerotinia-366/
+The docker container is located at https://hub.docker.com/r/everhartlab/sclerotinia-366/
 
@@
 ```bash
-docker run --name ssc --rm -dp 8787:8787 -e ROOT=TRUE zkamvar/sclerotinia-366:latest
+docker run --name ssc --rm -dp 8787:8787 -e ROOT=TRUE everhartlab/sclerotinia-366:latest
 ```

### 2017-11-17 doc/manuscript/manuscript.Rmd
remove funding, COI, and contrib to address #20

@@
 
-## Funding {-}
-
-Funding for this research was provided by the North Central Soybean Research Program (#639K623) to SEE, a Layman Award (#2446) to SEE, USDA-ARS National Sclerotinia Initiative (#58-5442-2-209) to JRS/SEE, and start-up funds from the University of Nebraska-Lincoln to SEE.
-The funders had no role in study design, data collection and analysis, decision to publish, or preparation of the manuscript.
-
 ## Acknowledgements {-}
@@
 We would also like to thank Denita Hadziabdic and two other anonymous reviewers for their valuable comments and insights that improved the quality of the manuscript.
-
-## Conflicts of Interest {-}
-
-The authors declare no conflict of interest.
-
-## Author Contributions {-}
-
- - **Zhian N. Kamvar** analyzed the data, contributed analysis tools, wrote the paper, prepared figures and tables, edited and reviewed drafts of the paper.
-
- - **B. Sajeewa Amaradasa** analyzed the data, contributed analysis tools, wrote drafts of the paper, edited and reviewed drafts of the paper.
-
- - **Rachana Jhala** Carried out experiments (MCG assessment, aggressiveness ratings, genotyping), edited and reviewed drafts of the paper.
-
- - **Serena McCoy** Carried out experiments (MCG assessment, aggressiveness ratings, genotyping), edited and reviewed drafts of the paper.
-
- - **James R. Steadman** Conceived and designed experiments, organized network of white mold screening nurseries, provided *S. sclerotiorum* isolates, edited and reviewed drafts of the paper.
-
- - **Sydney E. Everhart** supervised data analysis, analyzed the data, contributed analysis tools, wrote the paper, edited and reviewed drafts of the paper.
 

### 2017-11-17 peerj-nitpick.sh
begin script to update figures and tex file

@@ -1,0 +1,35 @@
+#!/bin/bash
+#
+# 2017-11-17
+#
+# This script is an attempt at addressing some of the issues brought up by
+# PeerJ to address their workflow. Specifically we need to:
+#
+# 1. Separate out the supplementary figures/tables from the manuscript
+# 2. Rename figures as Fig1.pdf ... Fig6.pdf for uploading.
+#
+# This script will perform all steps, including rendering to PDF.
+TEXFILE="doc/manuscript/manuscript.tex"
+
+# step 1: figure out where the supplementary information begins and ends.
+supplement_line=$(grep -n 'section\*{Supplementary' ${TEXFILE} | awk -F: '{print $1}')
+references_line=$(grep -n 'section\*{References' ${TEXFILE} | awk -F: '{print $1}')
+
+echo "Supplementary info starts on line ${supplement_line} and ends on $((references_line - 1))"
+
+# step 2: find the figures
+#    - Use () to create an array from the numbers
+figures=($(grep -n 'includegraphics' ${TEXFILE} | awk -F: '{print $1}'))
+#    - Get six elements of the array, starting at index 0: https://stackoverflow.com/a/1336245/2752888
+figures=(${figures[@]:0:6})
+
+echo "Figures are located on lines $(printf '%s ' ${figures[*]})"
+echo "I am updating their names now ..."
+counter=1;
+for i in ${figures[@]};
+do
+  fig=$(sed "${i}q;d" ${TEXFILE} | sed -r "s/^.+\{(.+)\}/\1/")
+  new=$(sed -r "s@^(.+)/([^/]+\.pdf)@\1/Fig${counter}.pdf@" <<< ${fig})
+  echo "Figure ${fig} will become ${new}"
+  counter=$((counter + 1))
+done

### 2017-11-17 peerj-nitpick.sh
script to change figure names complete

@@ -28,8 +28,13 @@ echo "I am updating their names now ..."
 counter=1;
 for i in ${figures[@]};
 do
-  fig=$(sed "${i}q;d" ${TEXFILE} | sed -r "s/^.+\{(.+)\}/\1/")
+  fig=$(sed "${i}q;d" ${TEXFILE} | sed -r "s/^.+\{(.+)\}/\1/" | sed 's@../../@@')
   new=$(sed -r "s@^(.+)/([^/]+\.pdf)@\1/Fig${counter}.pdf@" <<< ${fig})
-  echo "Figure ${fig} will become ${new}"
+  printf "sed -i \"s@${fig}@${new}@\" ${TEXFILE}\n"
+  sed -i "s@${fig}@${new}@" ${TEXFILE}
+  printf "cp ${fig} ${new}\n"
+  cp ${fig} ${new}
+  now=$(sed "${i}q;d" ${TEXFILE} | sed -r "s/^.+\{(.+)\}/\1/" | sed 's@../../@@')
+  printf "Figure ${fig} is now ${now}\n\n"
   counter=$((counter + 1))
 done

### 2017-11-17 peerj-nitpick.sh
add splitting of manuscript and supplementary

@@ -10,12 +10,15 @@
 #
 # This script will perform all steps, including rendering to PDF.
 TEXFILE="doc/manuscript/manuscript.tex"
+SUPFILE="doc/manuscript/supplementary.tex"
+PAPER="doc/manuscript/paper.tex"
+TEXLEN=$(wc -l ${TEXFILE} | awk '{print $1}')
 
 # step 1: figure out where the supplementary information begins and ends.
 supplement_line=$(grep -n 'section\*{Supplementary' ${TEXFILE} | awk -F: '{print $1}')
 references_line=$(grep -n 'section\*{References' ${TEXFILE} | awk -F: '{print $1}')
 
-echo "Supplementary info starts on line ${supplement_line} and ends on $((references_line - 1))"
+printf "Supplementary info starts on line ${supplement_line} and ends on $((references_line - 1))\n"
 
 # step 2: find the figures
 #    - Use () to create an array from the numbers
@@ -23,18 +26,31 @@ figures=($(grep -n 'includegraphics' ${TEXFILE} | awk -F: '{print $1}'))
 #    - Get six elements of the array, starting at index 0: https://stackoverflow.com/a/1336245/2752888
 figures=(${figures[@]:0:6})
 
-echo "Figures are located on lines $(printf '%s ' ${figures[*]})"
-echo "I am updating their names now ..."
-counter=1;
-for i in ${figures[@]};
-do
-  fig=$(sed "${i}q;d" ${TEXFILE} | sed -r "s/^.+\{(.+)\}/\1/" | sed 's@../../@@')
-  new=$(sed -r "s@^(.+)/([^/]+\.pdf)@\1/Fig${counter}.pdf@" <<< ${fig})
-  printf "sed -i \"s@${fig}@${new}@\" ${TEXFILE}\n"
-  sed -i "s@${fig}@${new}@" ${TEXFILE}
-  printf "cp ${fig} ${new}\n"
-  cp ${fig} ${new}
-  now=$(sed "${i}q;d" ${TEXFILE} | sed -r "s/^.+\{(.+)\}/\1/" | sed 's@../../@@')
-  printf "Figure ${fig} is now ${now}\n\n"
-  counter=$((counter + 1))
-done
+printf "\nFigures are located on lines $(printf '%s ' ${figures[*]})\n"
[…]

### 2017-11-17 peerj-nitpick.sh
add code to render the manuscript in the right order

@@ -12,11 +12,13 @@
 TEXFILE="doc/manuscript/manuscript.tex"
 SUPFILE="doc/manuscript/supplementary.tex"
 PAPER="doc/manuscript/paper.tex"
-TEXLEN=$(wc -l ${TEXFILE} | awk '{print $1}')
+MANUSCRIPT_SHELL="doc/manuscript/manuscript_shell.tex"
 
 # step 1: figure out where the supplementary information begins and ends.
 supplement_line=$(grep -n 'section\*{Supplementary' ${TEXFILE} | awk -F: '{print $1}')
 references_line=$(grep -n 'section\*{References' ${TEXFILE} | awk -F: '{print $1}')
+intro_line=$(grep -n 'section\*{Introduction' ${TEXFILE} | awk -F: '{print $1}')
+end_line=$(grep -n '\end{document' ${TEXFILE} | awk -F: '{print $1}')
 
 printf "Supplementary info starts on line ${supplement_line} and ends on $((references_line - 1))\n"
 
@@ -48,9 +50,38 @@ fi;
 
 # step 3: split the tex documents into two files to follow https://tex.stackexchange.com/a/87011/77699
 printf "\nSplitting the document...\n"
-printf "\tsed -n ${supplement_line},$((references_line-1))p ${TEXFILE} > ${SUPFILE}\n"
-sed -n ${supplement_line},$((references_line-1))p ${TEXFILE} > ${SUPFILE}
-printf "\tsed -n 1,$((supplement_line-1))p ${TEXFILE} > ${PAPER}\n"
-sed -n 1,$((supplement_line-1))p ${TEXFILE} > ${PAPER}
-printf "\tsed -n ${references_line},${texlen}p ${TEXFILE} >> ${PAPER}\n"
-sed -n ${references_line},${TEXLEN}p ${TEXFILE} >> ${PAPER}
+
+split_sup="sed -n ${supplement_line},$((references_line-1))p ${TEXFILE} > ${SUPFILE}"
+printf "\t${split_sup}\n"
+eval ${split_sup}
+
+split_body="sed -n ${intro_line},$((supplement_line-1))p ${TEXFILE} > ${PAPER}"
+printf "\t${split_body}\n"
+eval ${split_body}
+
+split_ref="sed -n ${references_line},$((end_line-1))p ${TEXFILE} >> ${PAPER}"
[…]

### 2017-11-17 peerj-nitpick.sh
found solution, but this involves a manual step

@@ -13,6 +13,8 @@ TEXFILE="doc/manuscript/manuscript.tex"
 SUPFILE="doc/manuscript/supplementary.tex"
 PAPER="doc/manuscript/paper.tex"
 MANUSCRIPT_SHELL="doc/manuscript/manuscript_shell.tex"
+PAPER_ONLY="doc/manuscript/paper-only.tex"
+SUPPLEMENTARY_ONLY="doc/manuscript/supplementary-only.tex"
 
 # step 1: figure out where the supplementary information begins and ends.
 supplement_line=$(grep -n 'section\*{Supplementary' ${TEXFILE} | awk -F: '{print $1}')
@@ -69,15 +71,16 @@ split_head="sed -n 1,$((intro_line-1))p ${TEXFILE} > ${MANUSCRIPT_SHELL}"
 printf "\t${split_head}\n"
 eval ${split_head}
 
-printf "\\\\include{paper}" >> ${MANUSCRIPT_SHELL}
-printf "\n" >> ${MANUSCRIPT_SHELL}
-printf "\\\\include{supplementary}" >> ${MANUSCRIPT_SHELL}
+printf "\\\\input{paper}" >> ${MANUSCRIPT_SHELL}
+printf "\n\\\\newpage\n" >> ${MANUSCRIPT_SHELL}
+printf "\\\\input{supplementary}" >> ${MANUSCRIPT_SHELL}
 printf "\n" >> ${MANUSCRIPT_SHELL}
 
-split_foot="printf '\\\\end{document}' >> ${MANUSCRIPT_SHELL}"
-printf "\t${split_foot}\n"
-eval ${split_foot}
+split_foot="printf '\\\\end{document}' >> "
+printf "\t${split_foot}${MANUSCRIPT_SHELL}\n"
+eval "${split_foot}${MANUSCRIPT_SHELL}"
 printf "\n" >> ${MANUSCRIPT_SHELL}
+
 
 printf "\nCreating manuscript...\n"
 cd doc/manuscript
@@ -85,3 +88,6 @@ pdflatex manuscript_shell
 pdflatex manuscript_shell
 pdflatex manuscript_shell
 cd -
+
+printf "\ndone\n"
+printf "\n\nNow open manuscript_shell.pdf and split it into the supplementary info and the manuscript manually\n"

### 2017-11-17 doc/manuscript/manuscript_shell.tex
run peerj-nitpick.sh

--- 2017-11-15 doc/manuscript/manuscript.tex
@@
 
-\section*{Introduction}\label{introduction}
-\addcontentsline{toc}{section}{Introduction}
-
-\emph{Sclerotinia sclerotiorum} (Lib.) de Bary is an ascomycete plant
-pathogen with a worldwide distribution (Bolton et al., 2006). This is a
-necrotrophic pathogen that is primarily homothallic (self-fertilization)
-and has the ability to survive for more than five years in soil using
-melanized survival structures called sclerotia (Bolton et al., 2006;
-Sexton et al., 2006). It causes disease on more than 400 plant species
-belonging to 75 families (Boland \& Hall, 1994) including crops of major
-economic importance such as sunflower (\emph{Helianthus spp.}), soybean
-(\emph{Glycine max} L.), canola (\emph{Brassica napa} L., \emph{Brassica
-campestris} L.), and dry bean (\emph{Phaseolus vulgaris} L.) (Bolton et
-al., 2006).
-
-On dry bean, \emph{S. sclerotiorum} is the causal agent of white mold, a
-devastating disease that can be yield-limiting in temperate climates
-(Steadman, 1983). All above-ground tissues (flowers, stems, leaves,
-pods) are susceptible to infection, first appearing as wet lesions with
-white mycelial tufts, and then bleaching as the tissue senesces
-(Steadman, 1983; Bolton et al., 2006). For many years, white mold has
-been the most serious dry bean disease in the Northwestern United States
-(Otto-Hanson et al., 2011; Knodel et al., 2012, 2015, 2016). The impact
-of white mold on the dry bean industry in the Northwestern United States
-alone has been estimated at a loss of 140 kg/ha with just 10\% disease
-incidence (Ramasubramaniam et al., 2008).
-
-Currently, there are no commercially available resistant cultivars of
-dry bean (Otto-Hanson et al., 2011). Organized breeding efforts have
-used a common-garden approach with white mold screening nurseries in dry
-bean production areas across the United States with additional sites in
-Australia, France, and Mexico (Steadman et al., 2003, 2004, 2005, 2006;
-Otto-Hanson \& Steadman, 2007, 2008; McCoy \& Steadman, 2009). These
-white mold screening nurseries use no chemical or cultural treatments
-against \emph{S. sclerotiorum} and employ standardized protocols for
-screening new cultivars for resistance to white mold (Steadman et al.,
-2003; Otto-Hanson et al., 2011). These protocols included three
-established cultivars used for comparison in the trials: Beryl (great
-northern bean, susceptible), Bunsi (a.k.a. Ex Rico, navy bean, low
-susceptibility), and G122 (cranberry bean, partial resistance) (Tu \&
-Beversdorf, 1982; Steadman et al., 2005; Otto-Hanson et al., 2011). It
-was previously shown that aggressiveness (the severity of disease
-symptoms on the host) is significantly different across white mold
-screening nursery sites in separate geographic regions (Otto-Hanson et
-al., 2011). The genetic structure and mode of reproduction in these
-populations, however, is currently unknown.
-
-Understanding genetic relationships and reproduction behavior of
-\emph{S. sclerotiorum} populations is beneficial for breeders seeking to
-develop new resistant cultivars for worldwide deployment (Milgroom,
-1996; McDonald \& Linde, 2002). In particular, genetically diverse
-populations with high rates of sexual reproduction are more likely to
-overcome host resistance. Most populations of \emph{S. sclerotiorum} are
-predominantly clonal with low genetic diversity and have a large degree
-of population fragmentation (Kohli et al., 1995; Cubeta et al., 1997;
-Kohli \& Kohn, 1998; Carbone \& Kohn, 2001; Ekins et al., 2011;
-Attanayake et al., 2012). Some studies, however have found populations
-that show signatures of sexual reproduction (Atallah et al., 2004;
-Sexton \& Howlett, 2004; Attanayake et al., 2013; Aldrich-Wolfe et al.,
-2015).
-
-Nearly all population genetic studies of \emph{S. sclerotiorum} employ a
-macroscopic assay to determine mycelial compatibility, the ability for
-fungal hyphae from different colonies to appear to grow together without
-forming a barrier of dead cells between them (known as a barrage line,
-Fig. \ref{mcg-fig}B) (Leslie, 1993; Sirjusingh \& Kohn, 2001). Mycelial
-compatibility has been used as a proxy for vegetative compatibility, a
-fungal trait controlled by several independent genes controlling the
-ability for two hyphae to fuse and grow as a single unit (Fig.
-\ref{mcg-fig}A) (Leslie, 1993; Schafer \& Kohn, 2006). Because of the
-genetic connection to vegetative compatibility, two isolates that are
-mycelially compatible were considered clones (Leslie, 1993); but
-correlation with genetic markers, such as microsatellites, have shown
-inconsistent results (Ford et al., 1995; Micali \& Smith, 2003; Jo et
-al., 2008; Attanayake et al., 2012; Papaioannou \& Typas, 2014; Lehner
-et al., 2017). Thus, the relationship between mycelial compatibility
-groups and clonal genotypes remains unclear.
-
-In the present study, we analyze and characterize the genetic and
-phenotypic diversity of 366 \emph{S. sclerotiorum} isolates collected
-between 2003 and 2012 from dry bean cultivars among different geographic
-locations in the Australia, France, Mexico, and the United States. We
-wanted to know if the \emph{S. sclerotiorum} populations from white mold
-screening nurseries were representative of the producer fields within
-the same region. As these nurseries were not treated with any chemical
-or cultural control of white mold, we hypothesized that these nurseries
-would represent the natural population of \emph{S. sclerotiorum}.
-Furthermore, we wanted to investigate the potential effect of cultivar
-on genetic diversity of the pathogen by assessing three dry bean
-cultivars with different levels of resistance, Beryl (great northern
-bean, susceptible), Bunsi (navy bean, low susceptibility), and G122
-(cranberry bean, partial resistance) (Otto-Hanson et al., 2011). We
-additionally wanted to determine categorical or phenotypic variables
-that best predicted genetic structure and if there was correlation
-between multilocus haplotype and mycelial compatibility group. Knowing
-what variables predict genetic structure can help direct breeding
-efforts. By investigating these aims, we will effectively describe the
-population structure of \emph{S. sclerotiorum} in the USA and make
-available our database of isolates for use in future dry bean breeding
-efforts.
-
-\section*{Materials and Methods}\label{materials-and-methods}
-\addcontentsline{toc}{section}{Materials and Methods}
-
-\subsection*{Isolate collection}\label{isolate-collection}
-\addcontentsline{toc}{subsection}{Isolate collection}
-
-Several (156) of the isolates used for this study were collected as
-reported in previous studies using the same methods (Otto-Hanson et al.,
-2011). Broadly, isolates were collected from two sources: white mold
-screening nurseries (wmn) or producer fields. White mold screening
-nurseries were 5m x 10m in size and maintained without application of
-fungicides to observe natural incidence of white mold. The early nursery
-plots were incorporated with a basal dressing of N:P:K = 1:3:2 and side
-dressing of 0:3:2 during the growing season (Steadman et al., 2003).
-
-Sampling was carried out by collecting sclerotia from diseased tissue in
-zig-zag transects across field plots. Because sampling depended on
-disease incidence, the number of samples isolated varied from year to
-year. Although the nursery locations were the same over sampling years,
-sampling plots within a location varied for sampling years.
-
-Sclerotia of \emph{S. sclerotiorum} were collected over several years
-from grower fields and/or wmn in 11 states of the Australia, France,
-Mexico, and the United States (Table \ref{tab:isolate-table}). After
-collection, sclerotia were stored in Petri plates lined with filter
-paper, then stored at \(20~^{\circ}\)F or -\(4~^{\circ}\)C. Sclerotia
-were surface-sterilized with 50\% Clorox bleach (at least 6\% NaOCl, The
-Clorox Company, Oakland, CA) solution for 3 min, and double rinsed with
-ddH\(_2\)O for 3 min. The sterilized sclerotia were then placed on water
-agar plates (16g of Bacto agar per liter of ddH\(_2\)O, BD Diagnostic
-Systems, Sparks, MD), with four to five sclerotia of each isolate
-separated on each plate and stored on the counter top at room
-temperature for 5 to 6 days. An 8-mm plug from a 5- or 6-day-old culture
-was transferred from the advancing margin of the mycelia onto a plate of
-Difco potato dextrose agar (PDA at 39 g/liter of ddH\(_2\)O)
-(Otto-Hanson et al., 2011). In combination with the 156 isolates
-described previously, we collected 210 isolates for a total of 366
-isolates (Otto-Hanson et al., 2011).
-
-\subsection*{Mycelial compatibility}\label{mycelial-compatibility}
-\addcontentsline{toc}{subsection}{Mycelial compatibility}
-
-MCG was determined as described previously through co-culturing pairs of
-2-day-old isolates 2.5 cm apart on Diana Sermons (DS) Medium (Fig.
-\ref{mcg-fig}) (Cubeta et al., 2001). Incompatibility of different MCGs
-resulted in formation of a barrage line accompanied by formation of
-sclerotia on either side of the barrage line, indicating the limits of
-mycelial growth (Kohn et al., 1990; Leslie, 1993; Otto-Hanson et al.,
-2011). Isolates were compared in a pairwise manner for each site and
-then representatives among sites were compared to determine mycelial
-compatibility groups by scoring compatible and incompatible interactions
-(Otto-Hanson et al., 2011). No MCGs were compatible with any other MCG.
-
-\subsection*{Aggressiveness}\label{aggressiveness}
-\addcontentsline{toc}{subsection}{Aggressiveness}
-
-Aggressiveness of each isolate was assessed using a straw test as
-described in Otto-Hanson et al. (2011) that rated necrotic lesion size
-(Petzoldt \& Dickson, 1996; Teran et al., 2006). Briefly, the straw test
-uses 21-day-old G122 plants as the host in a greenhouse setting. Clear
-drinking straws cut to 2.5 cm and heat sealed were used to place two
-mycelial plugs of inoculum on the host plant after removing plant growth
-beyond 2.5 cm above the fourth node. Measurements of the necrotic lesion
-were taken 8 days later using the Modified Petzoldt and Dickson scale of
-1--9, where 1 is no disease and 9 is plant death (Petzoldt \& Dickson,
-1996; Teran et al., 2006).
-
-\subsection*{Microsatellite genotyping}\label{microsatellite-genotyping}
-\addcontentsline{toc}{subsection}{Microsatellite genotyping}
-
-Prior to DNA extraction, isolates were grown on PDA and plugs were
-subsequently transferred to Potato Dextrose Broth (PDB) where they were
-grown until there was significant mycelial growth, but before the
-mycelial mat became solidified (4--5 days). Each mycelial mat was
-collected in a filtered Büchner funnel, agar plugs removed, lyophilized
-and pulverized manually in Whirl-pak® HDPE sampling bags (Sigma-Aldrich,
-St.~Louis, MO). Lyophilized mycelia was then stored in microcentrifuge
-tubes at -\(20~^{\circ}\)C until needed for DNA extraction. DNA from
-25mg of pulverized mycelia was purified using a phenol-chloroform
-extraction method followed by alcohol precipitation and evaporation,
-suspending the DNA in 200\(\mu\)l TE (Sambrook et al., 1989). Suspended
-DNA was stored at \(4~^{\circ}\)C until genotyping.
-
-We genotyped each \emph{S. sclerotiorum} isolate using 16 microsatellite
-primer pairs developed previously (Sirjusingh \& Kohn, 2001). PCR was
-carried out as described previously, using primers labeled with FAM
-fluorophore. Resulting amplicons were first resolved in a 1.5\% agarose
-gel stained with ethidium bromide to ensure product was within the
-expected size range prior to capillary electrophoresis. Capillary
-electrophoresis (fragment analysis) of amplicons, with size standard
-GeneScan™ 500 LIZ®, was performed using an ABI 3730 genetic analyzer
-(Life Technologies Corporation, Carlsbad, CA) at the Michigan State
-University Genomic Sequencing Center (East Lansing, MI). Alleles were
-scored using PeakScanner version 1.0 (Life Technologies Corporation,
-Carlsbad, CA) and recorded manually in a spreadsheet.
-
-\subsection*{Data processing and
-analysis}\label{data-processing-and-analysis}
-\addcontentsline{toc}{subsection}{Data processing and analysis}
-
-All data processing and analyses were performed in a Rocker ``verse''
-project container running R version 3.4.2 (Boettiger \& Eddelbuettel,
-2017; R Core Team, 2017) and are openly available and reproducible at
-\url{https://github.com/everhartlab/sclerotinia-366/}. Of the 16
-microsatellite loci genotyped, five included compound repeats, which
-made it challenging to accurately/confidently bin alleles into fragment
-sizes expected for each locus based on the described repeat motif. Loci
-with compound repeats were removed for the reported statistics. To
-ensure the integrity of the results we additionally processed these loci
-and included them in concurrent analyses. We assessed the power of our
-11 markers by generating a genotype accumulation curve in the R package
-\emph{poppr} version 2.5.0, looking for evidence of saturation, which
-would indicate that loci were sufficiently sampled to adequately
-represent the full set of haplotypes (Arnaud-Hanod et al., 2007; Kamvar
-et al., 2015). To avoid including isolates potentially collected from
-the same plant (which increases the probability of collecting sclerotia
-from the same point of infection more than once), data were
-clone-corrected on a hierarchy of Region/Source/Host/Year---meaning that
-duplicated genotypes were reduced to a single observation when they were
-collected in the same year from the same host cultivar located in the
-same source field (wmn or producer)---for subsequent analysis. We
-assessed haplotype diversity by calculating Stoddart and Taylor's index
-(\emph{G}) (Stoddart \& Taylor, 1988), Shannon's index (\emph{H})
-(Shannon, 1948), Simpson's index (\(\lambda\)) (Simpson, 1949), evenness
-(\(E_5\)), and the expected number of multilocus haplotypes
-(\emph{eMLH}) (Hurlbert, 1971; Heck et al., 1975; Pielou, 1975; Grünwald
-et al., 2003). To assess the potential for random mating, we tested for
-linkage disequilibrium with the index of association, \(I_A\) and its
-standardized version, \(\bar{r}_d\) using 999 permutations (Brown et
-al., 1980; Smith et al., 1993; Agapow \& Burt, 2001). Both haplotype
-diversity and linkage disequilibrium were calculated in \emph{poppr}
-(Kamvar et al., 2014).
-
-\subsection*{Assessing importance of
-variables}\label{assessing-importance-of-variables}
-\addcontentsline{toc}{subsection}{Assessing importance of variables}
-
-\subsubsection*{Distance-based redundancy
-analysis}\label{distance-based-redundancy-analysis}
-\addcontentsline{toc}{subsubsection}{Distance-based redundancy analysis}
-
-A distance-based redundancy analysis (dbRDA) (Legendre \& Anderson,
-1999) was performed with the function \texttt{capscale()} in the
-\emph{vegan} package version 2.4.4 (Oksanen et al., 2017). This method
-uses constrained ordinations on a distance matrix representing the
-response variable to delineate relative contribution of any number of
-independent explanatory variables. We used this method to delineate the
-phenotypic (Aggressiveness, Mycelial Compatibility Group (MCG)),
-geographic (Region, Host, Location), and temporal (Year) components in
-predicting genetic composition of the populations. The distance matrix
-we used as the response variable was generated using Bruvo's genetic
-distance from clone-corrected data (procedure described above) as
-implemented in \emph{poppr}, which employed a stepwise mutation model
-for microsatellite data (Bruvo et al., 2004; Kamvar et al., 2014).
-Because aggressiveness measures differed between isolates that were
-reduced to a single observation during clone-correction, aggressiveness
-was first averaged across clone-corrected isolates. To identify
-explanatory variable(s) correlated with genetic variation, a
-forward-backward selection process was applied with the vegan function
-\texttt{ordistep()}. An analysis of variance (ANOVA) was then performed
-to test for significance of the reduced model and marginal effects using
-999 permutations. The \texttt{varpart()} function of vegan was used to
-determine variation partitioning of explanatory variables.
-
-\subsubsection*{Aggressiveness
-assessment}\label{aggressiveness-assessment}
-\addcontentsline{toc}{subsubsection}{Aggressiveness assessment}
-
-We used ANOVA to assess if aggressiveness (determined via straw test on
-a scale of 1--9 as described above) was significantly different with
-respect to Region, MCG, or multilocus haplotype (MLH). To minimize
-complications due to small sample sizes, we chose the top 10 MCGs,
-representing 56.5\% of the isolates collected, the 10 most abundant MLHs
-representing 26.7\% of the isolates, and populations with more than five
-isolates. If ANOVA results were significantly different at \(\alpha\) =
-0.05, pairwise differences were assessed using Tukey's HSD test
-(\(\alpha\) = 0.05) using the \texttt{HSD.test()} function in the
-package \emph{agricolae} version 1.2.8 (Mendiburu \& Simon, 2015).
-
-\subsubsection*{Correlating multilocus haplotypes with mycelial
-compatibility
-groups}\label{correlating-multilocus-haplotypes-with-mycelial-compatibility-groups}
-\addcontentsline{toc}{subsubsection}{Correlating multilocus haplotypes
-with mycelial compatibility groups}
-
-We wanted to assess if there was correlation between MLHs and MCGs. This
-was performed using a network-based approach where both MLHs and MCGs
-were considered nodes and the number of isolates in which they were
-found together was the strength of the connection between an MLH and and
-MCG node. The network-based approach allowed us to assess the
-associations between MLHs and MCGs. To construct the network, a
-contingency table was created with MLHs and MCGs and converted to a
-directed and weighted edgelist where each edge represented a connection
-from an MCG to an MLH, weighted by the number of samples shared in the
-connection. This was then converted to a bipartite graph where top nodes
-represented MLHs and bottom nodes represented MCGs. To identify clusters
-of MLHs and MCGs within the network, we used the cluster walktrap
-community detection algorithm as implemented in the
-\texttt{cluster\_walktrap()} function in \emph{igraph} version 1.1.2
-(Csardi \& Nepusz, 2006; Pons \& Latapy, 2006). This algorithm attempts
-to define clusters of nodes by starting at a random node and performing
-short, random ``walks'' along the edges between nodes, assuming that
-these walks would stay within clusters. For this analysis, we set the
-number of steps within a walk to four and allowed the algorithm to use
-the edge weights in determining the path. All of the resulting
-communities that had fewer than 10 members were then consolidated into
-one. Community definitions were used to assess the average genetic
-distance (as defined by Bruvo's distance) within members of the
-community (Bruvo et al., 2004).
-
-\subsection*{Genetic diversity}\label{genetic-diversity}
-\addcontentsline{toc}{subsection}{Genetic diversity}
-
-\subsubsection*{Population
-differentiation}\label{population-differentiation}
-\addcontentsline{toc}{subsubsection}{Population differentiation}
-
-We used analysis of molecular variance (AMOVA) with Bruvo's genetic
-distance in \emph{poppr} to test for differentiation between populations
-in wmn and producer fields from the same region and collected in the
-same year (Excoffier et al., 1992; Bruvo et al., 2004; Kamvar et al.,
-2014). To identify Regions with greater differentiation, we used
-discriminant analysis of principal components (DAPC) as implemented in
-\emph{adegenet} version 2.1.0, assessing the per-sample posterior group
-assignment probability (Jombart, 2008). This method decomposes the
-genetic data into principal components, and then uses a subset of these
-as the inputs for discriminant analysis, which attempts to minimize
-within-group variation and maximize among-group variation (Jombart et
-al., 2010). To avoid over-fitting data, the optimal number of principal
-components was selected by using the \emph{adegenet} function
-\texttt{xvalDapc()}. This function implements a cross-validation
-procedure to iterate over an increasing number of principal components
-on a subset (90\%) of the data, trying to find the minimum number of
-principal components that maximizes the rate of successful group
-reassignment. To assess if cultivar had an influence on genetic
-diversity between wmn, we first subset the clone-corrected data to
-contain only samples from wmn and from the cultivars Beryl, Bunsi, and
-G122 and tested differentiation using AMOVA and DAPC as described above.
-We additionally assessed population stability over time by calculating
-DAPC over the combined groups of Region and Year as described above.
-
-\subsubsection*{Analysis of shared multilocus
-haplotypes}\label{analysis-of-shared-multilocus-haplotypes}
-\addcontentsline{toc}{subsubsection}{Analysis of shared multilocus
-haplotypes}
-
-We wanted to evaluate patterns of connectivity between shared multilocus
-haplotypes across geographic regions. We first tabulated the multilocus
-haplotypes shared between at least two populations (defined as states or
-countries) with the \emph{poppr} function \texttt{mlg.crosspop()}
-(Kamvar et al., 2014). From these data, we constructed a graph with
-populations as nodes and shared haplotypes as edges (connections)
-between nodes using the R packages \emph{igraph} (Csardi \& Nepusz,
-2006), \emph{dplyr} version 0.7.4 (Wickham et al., 2017), and
-\emph{purrr} version 0.2.4 (Henry \& Wickham, 2017). Each node was
-weighted by the fraction of shared MLHs. Each edge represented a single
-MLH, but because a single MLH could be present in more than one
-population, that MLH would have a number of edges equivalent to the
-total number of possible connections, calculated as
-(\emph{n}*(\emph{n}-1))/2 edges where n represents the number of
-populations crossed. Edges were weighted by \(1-P_{sex}\), where
-\(P_{sex}\) is the probability of encountering the same haplotype via
-two independent meiotic events (Parks \& Werth, 1993; Arnaud-Hanod et
-al., 2007). This weighting scheme would thus strengthen the connection
-of edges that represented genotypes with a low probability of being
-produced via sexual reproduction. We then identified communities (among
-the Regions) in the graph using the \texttt{cluster\_optimal()} function
-from \emph{igraph} (Csardi \& Nepusz, 2006). The graph was plotted using
-the R packages \emph{ggplot2} version 2.2.1 (Wickham, 2009) and
-\emph{ggraph} 1.0.0 (Pedersen, 2017). To ensure that we captured the
-same community signal, we additionally performed this analysis including
-the five polymorphic markers described above.
-
-\section*{Results}\label{results}
-\addcontentsline{toc}{section}{Results}
-
-A total of 366 isolates were collected from 2003 to 2012 (except 2006
-and 2011) from diseased dry bean plants in 11 states in the United
-States as well as Australia, France, and Mexico (Table
-\ref{tab:isolate-table}). With the 11 loci used in the analyses (Table
-\ref{tab:locus-stats}), we observed a total of 165 MLHs (215 with 16
-loci). These 11 loci are located on 7 chromosomes in the \emph{S.
-sclerotiorum} genome with a minimum distance of 55Kbp between two loci
-on the same chromosome. Over 50\% of the isolates came from four states,
-MI (62), ND (60), WA (59), NE (47). Four regions had fewer than 10
-isolates, Australia (6), WI (2), NY (1), ID (1). We observed 87 MCGs,
-the most abundant of which (`MCG 5') was represented by 73 isolates over
-37 MLHs (Fig. \ref{biggraph}A,C).
-
-The number of observed alleles per locus ranged from two to 10 with an
-average of 6.27 (Table \ref{tab:locus-stats}). Locus 20-3, which
-contained only 2 alleles, showed low values of both \emph{h} (0.0533)
-and evenness (0.42), indicating that there was one dominant allele
-present. Analysis of the haplotype accumulation curve showed no clear
-plateau for 11 or 16 loci (See section on `Loading Data and Setting
-Strata' in the MLG-distribution.md\footnote{Direct link:
-  \url{https://github.com/everhartlab/sclerotinia-366/blob/master/results/MLG-distribution.md\#loading-data-and-setting-strata}}
-file in the supplemental files (Kamvar et al., 2017)), indicating that
-we would likely obtain more multilocus haplotypes if we were to genotype
-more loci.
-
-After clone-correction on the hierarchy of Region/Source/Host/Year, a
-total of 48 isolates were removed from the data set, resulting in 318
-isolates representing 165 MLHs that were used in subsequent analyses
-(Table \ref{tab:iatab}). The results showed that, in terms of genotypic
-diversity (\emph{H}, \emph{G}, and \(\lambda\)), WA was the most diverse
-population with both \emph{G} (54.3) and \(e^H\) (55.3) being close to
-the observed number of MLHs (56). This indicated that there are few
-duplicated genotypes in WA (Table \ref{tab:iatab}). A more useful metric
-to compare populations, however, is \(E_5\), which scales from 0 to 1,
-where 1 indicates all unique genotypes (Grünwald et al., 2003).
-Evaluating by \(E_5\) shows that both MI and NE exhibit lower than
-average values, indicating that there are over-represented genotypes in
-the popualtions (table \ref{tab:iatab}). When we look at Mexico, we
-observed that it had relatively high values of \(E_5\) and genotypic
-diversity, but low richness, as measured by \emph{eMLG}. Moreover,
-Mexico had the lowest value for \emph{h}, which is a measure of allelic
-diversity. Nearly all populations showed evidence of linkage (Table
-\ref{tab:iatab}), which serves as evidence for clonal reproduction or
-other forms of non-random mating. The only exceptions were CA (\emph{P}
-= 0.043) and Australia (\emph{P} = 0.052). Both of these populations
-showed only moderate significance with \(\bar{r}_d\) values of 0.03 and
-0.12, respectively.
-
-\begin{longtable}[]{@{}rrrlrl@{}}
-\caption{\label{tab:locus-stats} Allelic diversity on full data set at loci
-used in this study. \emph{h} = Nei's Gene Diversity (Nei, 1978). Average
-\emph{h} = 0.583, average Evenness = 0.693, average no. alleles =
-6.27}\tabularnewline
-\toprule
-\begin{minipage}[b]{0.08\columnwidth}\raggedleft\strut
-Locus\strut
-\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedleft\strut
-Range\strut
-\end{minipage} & \begin{minipage}[b]{0.29\columnwidth}\raggedleft\strut
-Repeat Motif\strut
-\end{minipage} & \begin{minipage}[b]{0.14\columnwidth}\raggedright\strut
-No. alleles\strut
-\end{minipage} & \begin{minipage}[b]{0.06\columnwidth}\raggedleft\strut
-\emph{h}\strut
-\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
-Evenness\strut
-\end{minipage}\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-\begin{minipage}[b]{0.08\columnwidth}\raggedleft\strut
-Locus\strut
-\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedleft\strut
-Range\strut
-\end{minipage} & \begin{minipage}[b]{0.29\columnwidth}\raggedleft\strut
-Repeat Motif\strut
-\end{minipage} & \begin{minipage}[b]{0.14\columnwidth}\raggedright\strut
-No. alleles\strut
-\end{minipage} & \begin{minipage}[b]{0.06\columnwidth}\raggedleft\strut
-\emph{h}\strut
-\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
-Evenness\strut
-\end{minipage}\tabularnewline
-\midrule
-\endhead
-\begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
-5-2\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
-318--324\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
-(GT)\strut
-\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
-4\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
-0.45\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-0.62\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
-6-2\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
-483--495\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
-(TTTTTC)(TTTTTG)(TTTTTC)\strut
-\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
-3\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
-0.64\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-0.95\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
-7-2\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
-158--174\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
-(GA)\strut
-\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
-7\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
-0.73\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-0.76\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
-8-3\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
-244--270\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
-(CA)\strut
-\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
-7\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
-0.74\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-0.79\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
-9-2\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
-360--382\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
-(CA)(CT)\strut
-\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
-9\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
-0.35\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-0.41\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
-12-2\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
-214--222\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
-(CA)\strut
-\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
-5\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
-0.58\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-0.78\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
-17-3\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
-342--363\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
-(TTA)\strut
-\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
-7\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
-0.55\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-0.53\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
-20-3\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
-280--282\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
-(GT)GG(GT)\strut
-\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
-2\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
-0.05\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-0.42\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
-55-4\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
-153--216\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
-(TACA)\strut
-\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
-10\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
-0.72\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-0.66\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
-110-4\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
-370--386\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
-(TATG)\strut
-\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
-5\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
-0.76\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-0.91\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
-114-4\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedleft\strut
-339--416\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedleft\strut
-(TAGA)\strut
-\end{minipage} & \begin{minipage}[t]{0.14\columnwidth}\raggedright\strut
-10\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
-0.83\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
-0.80\strut
-\end{minipage}\tabularnewline
-\bottomrule
-\end{longtable}
-
+\input{paper}
 \newpage
-
-\begin{longtable}[]{@{}crrrrllrl@{}}
-\caption{\label{tab:iatab} Genotypic diversity and Linkage Disequilibrium
-summary for geographic populations arranged by abundance after
-clone-correction by a hierarchy of Region/Source/Host/Year. Pop =
-Population, N = number of individuals (number of MLH in parentheses),
-eMLH = expected number of MLHs based on rarefaction at 10 individuals
-(standard error in parentheses), H = Shannon-Weiner Index, G = Stoddardt
-and Taylor's Index, \(\lambda\) = Simpson's Index, \emph{h} = Nei's 1978
-gene diversity, \(E_5\) = Evenness, \(\bar{r}_d\) = standardized index
-of association. An asterix indicates a significant value of
-\(\bar{r}_d\) after 999 permutations, \(P \leq 0.001\).}\tabularnewline
-\toprule
-\begin{minipage}[b]{0.12\columnwidth}\centering\strut
-Pop\strut
-\end{minipage} & \begin{minipage}[b]{0.08\columnwidth}\raggedleft\strut
-N\strut
-\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\raggedleft\strut
-eMLH\strut
-\end{minipage} & \begin{minipage}[b]{0.04\columnwidth}\raggedleft\strut
-H\strut
-\end{minipage} & \begin{minipage}[b]{0.05\columnwidth}\raggedleft\strut
-G\strut
-\end{minipage} & \begin{minipage}[b]{0.11\columnwidth}\raggedright\strut
-\(\lambda\)\strut
-\end{minipage} & \begin{minipage}[b]{0.06\columnwidth}\raggedright\strut
-\(E_5\)\strut
-\end{minipage} & \begin{minipage}[b]{0.06\columnwidth}\raggedleft\strut
-\emph{h}\strut
-\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\raggedright\strut
-\(\bar{ r}_d\)\strut
-\end{minipage}\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-\begin{minipage}[b]{0.12\columnwidth}\centering\strut
-Pop\strut
-\end{minipage} & \begin{minipage}[b]{0.08\columnwidth}\raggedleft\strut
-N\strut
-\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\raggedleft\strut
-eMLH\strut
-\end{minipage} & \begin{minipage}[b]{0.04\columnwidth}\raggedleft\strut
-H\strut
-\end{minipage} & \begin{minipage}[b]{0.05\columnwidth}\raggedleft\strut
-G\strut
-\end{minipage} & \begin{minipage}[b]{0.11\columnwidth}\raggedright\strut
-\(\lambda\)\strut
-\end{minipage} & \begin{minipage}[b]{0.06\columnwidth}\raggedright\strut
-\(E_5\)\strut
-\end{minipage} & \begin{minipage}[b]{0.06\columnwidth}\raggedleft\strut
-\emph{h}\strut
-\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\raggedright\strut
-\(\bar{ r}_d\)\strut
-\end{minipage}\tabularnewline
-\midrule
-\endhead
-\begin{minipage}[t]{0.12\columnwidth}\centering\strut
-WA\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
-58 (56)\strut
-\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedleft\strut
-9.95 (0.23)\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-4.0\strut
-\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
-54.3\strut
-\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-0.98\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
-0.98\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
-0.60\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
-0.07*\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\centering\strut
-MI\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
-58 (43)\strut
-\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedleft\strut
-9.3 (0.79)\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-3.6\strut
-\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
-29.0\strut
-\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-0.97\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
-0.78\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
-0.54\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
-0.14*\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\centering\strut
-ND\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
-41 (35)\strut
-\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedleft\strut
-9.44 (0.73)\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-3.5\strut
-\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
-25.9\strut
-\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-0.96\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
-0.82\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
-0.54\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
-0.1*\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\centering\strut
-NE\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
-37 (28)\strut
-\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedleft\strut
-8.93 (0.94)\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-3.2\strut
-\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
-17.8\strut
-\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-0.94\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
-0.75\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
-0.55\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
-0.25*\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\centering\strut
-CO\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
-34 (28)\strut
-\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedleft\strut
-9.46 (0.67)\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-3.3\strut
-\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
-24.1\strut
-\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-0.96\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
-0.92\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
-0.56\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
-0.27*\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\centering\strut
-France\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
-21 (14)\strut
-\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedleft\strut
-8.5 (0.85)\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-2.6\strut
-\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
-12.6\strut
-\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-0.92\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
-0.95\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
-0.48\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
-0.11*\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\centering\strut
-CA\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
-18 (15)\strut
-\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedleft\strut
-9.12 (0.72)\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-2.7\strut
-\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
-13.5\strut
-\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-0.93\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
-0.94\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
-0.51\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
-0.03\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\centering\strut
-OR\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
-17 (13)\strut
-\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedleft\strut
-8.52 (0.85)\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-2.5\strut
-\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
-10.7\strut
-\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-0.91\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
-0.89\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
-0.47\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
-0.1*\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\centering\strut
-Mexico\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
-15 (9)\strut
-\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedleft\strut
-7.1 (0.85)\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-2.1\strut
-\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
-7.3\strut
-\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-0.86\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
-0.89\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
-0.28\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
-0.37*\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\centering\strut
-MN\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
-9 (7)\strut
-\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedleft\strut
-7 (0)\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-1.9\strut
-\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
-6.2\strut
-\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-0.84\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
-0.93\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
-0.47\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
-0.19*\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\centering\strut
-Australia\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
-6 (6)\strut
-\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedleft\strut
-6 (0)\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-1.8\strut
-\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
-6.0\strut
-\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-0.83\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
-1.00\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
-0.48\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
-0.12\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\centering\strut
-WI\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
-2 (2)\strut
-\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedleft\strut
-2 (0)\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-0.7\strut
-\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
-2.0\strut
-\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-0.50\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
-1.00\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
-0.27\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
--\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\centering\strut
-NY\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
-1 (1)\strut
-\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedleft\strut
-1 (0)\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-0.0\strut
-\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
-1.0\strut
-\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-0.00\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
-NaN\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
-NaN\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
--\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.12\columnwidth}\centering\strut
-ID\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedleft\strut
-1 (1)\strut
-\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedleft\strut
-1 (0)\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-0.0\strut
-\end{minipage} & \begin{minipage}[t]{0.05\columnwidth}\raggedleft\strut
-1.0\strut
-\end{minipage} & \begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-0.00\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
-NaN\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedleft\strut
-NaN\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
--\strut
-\end{minipage}\tabularnewline
-\bottomrule
-\end{longtable}
-
-\begin{figure}
-\centering
-\includegraphics[width=1.00000\textwidth]{../../results/figures/publication/mcg-subgraph-with-context.pdf}
-\caption{Associations between Mycelial Compatibility Groups and
-Multilocus Haplotypes. \textbf{A)} Barplot of Mycelial Compatibility
-Group (MCG) abundance in descending order. Singletons (46) were
-truncated, leaving 41 MCGs. White bars represent sample counts and grey
-bars represent counts of unique multilocus haplotypes (MLH). The
-transparency of the bars represent the evenness of the distribution of
-the MLHs within a given MCG. A dashed box surrounds the eight most
-common MCGs representing \textgreater{}~51\% of the data. \textbf{B)}
-Full graph-representation of the relationship between MCGs (open
-circles) and MLHs (filled circles). Details in Fig. \ref{fullgraph}.
-\textbf{C)} A subset of \textbf{B} representing the 8 most common MCGs
-and their associated MLHs (dashed box in \textbf{A}). Filled nodes
-(circles) represent MLHs and open nodes represent MCGs. Node area scaled
-to the number of samples represented (range: 1--73). Numbers inside
-nodes are the MLH/MCG label (if n \textgreater{} 1). Edges (arrows)
-point from MLH to MCG where the weight (thickness) of the edge
-represents the number of shared isolates (range: 1--19). Edges extending
-from MLHs displayed to other MCGs are not shown.}\label{biggraph}
-\end{figure}
-
-\subsection*{Variable assessment}\label{variable-assessment}
-\addcontentsline{toc}{subsection}{Variable assessment}
-
-\subsubsection*{Variable contributions}\label{variable-contributions}
-\addcontentsline{toc}{subsubsection}{Variable contributions}
-
-\begin{figure}
-\centering
-\includegraphics[width=0.61000\textwidth]{../../results/figures/publication/Figure7Z.pdf}
-\caption{Biplot showing five most influential explanatory variables
-(arrows) overlayed on the first two eigenvectors of distance based
-redundancy analysis of \emph{Sclerotinia sclerotiorum} isolates. The
-length of the arrows are directly proportional to the strength of the
-correlation between explanatory and molecular variables. Open circles
-represent the 318 clone-corrected haplotypes in ordination
-space.}\label{rda-biplot}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{../../results/figures/publication/aggressiveness.pdf}
-\caption{Strip plot of aggressiveness by population arranged in
-descending order of mean aggressiveness for all populations with N
-\textgreater{} 5. White bars represent mean value. Circles represent
-individual isolates where filled circles are isolates from white mold
-screening nurseries (wmn) and open circles are isolates from producer
-fields.}\label{aggressiveness-boxplot}
-\end{figure}
-
-The forward-backward selection process of the dbRDA models on
-clone-corrected data revealed Year, Region, Host, and MCG to be the
-optimal variables for the reduced model, accounting for 45\% of the
-total variation. ANOVA showed that the reduced model was significant
-with an adjusted R\textsuperscript{2} of 0.0675 (\emph{P} = 0.001).
-Assessment of the marginal effects showed that all varibles
-significantly explained genetic variation (\(P \leq 0.007\)). We found
-that there was multicollinearity when MCG was combined with any other
-variable, so repeated the analysis, dropping MCG from the list of
-potential predictors. From these results, Year, Region, Host, and
-Aggressiveness were found to be optimal, accounting for 17.6\% of the
-total variation. ANOVA revealed significant effects with an adjusted
-R\textsuperscript{2} of 0.0325 (\emph{P} = 0.001). While the marginal
-effect assessment revealed that Year, Region, and Host significantly
-explained variation at \emph{P} = 0.001, and Aggressiveness
-significantly explained variation at \emph{P} = 0.039. Much of the
-variation appeared to be driven by isolates from Mexico and 2005 (Fig.
-\ref{rda-biplot}). Variance partitioning of the independent variables
-without MCG indicated aggressiveness to be the least influential factor
-with 0.1\% contributing to explaining the variation of molecular data,
-whereas the combination of variables accounted for 3.3\%.
-
-\subsubsection*{Aggressiveness}\label{aggressiveness-1}
-\addcontentsline{toc}{subsubsection}{Aggressiveness}
-
-Aggressiveness of the isolates ranged from 1.4 to 7.9 with a mean of
-5.02 and median of 4.85. The group mean averages were 4.88, 5.13, and
-5.19 for Region, MCG, and MLH, respectively. A strip plot showing the
-distribution of severity across these three variables simultaneously can
-be seen in Fig. \ref{mcg-mlh-region-aggressiveness}. Our assessment of
-aggressiveness in association with Region showed a significant effect
-(\emph{P} \textless{} 1.00e\textsuperscript{-4}), with means that ranged
-from 5.8 (MN) to 4.0 (CA) (Fig. \ref{aggressiveness-boxplot}, Table
-\ref{tab:region-aggressiveness}). MCGs also showed a significant effect
-(\emph{P} \textless{} 0.001), with means that ranged from 6.0 (`MCG 44')
-to 4.6 (`MCG 49'; Table \ref{tab:MCG-aggressiveness}). We additionally
-found a significant effect for MLHs (\emph{P} \textless{} 0.001), with
-means that ranged from 6.0 (`MLH 78') to 4.3 (`MLH 140') (Table
-\ref{tab:MLG-aggressiveness}).
-
-\subsubsection*{Correlation of mulitlocus haplotypes and mycelial
-compatibility
-groups}\label{correlation-of-mulitlocus-haplotypes-and-mycelial-compatibility-groups}
-\addcontentsline{toc}{subsubsection}{Correlation of mulitlocus
-haplotypes and mycelial compatibility groups}
-
-In our analysis, we found 165 MLHs with 70 singletons and 87 MCGs with
-43 singletons (Fig. \ref{biggraph}A,B) where the eight most abundant
-MCGs represented \textgreater{} 51\% of the data over 11 Regions, and
-all years except for 2012. Our network-based approach to correlating
-MLHs with MCGs revealed a large and complex network (Fig.
-\ref{biggraph}, Table \ref{tab:mlg-table}). Community analysis showed 51
-communities, 15 of which consisted of a single MLH unconnected with any
-other community indicating that just 9.09\% of the 165 MLHs are unable
-to cross with any other MLH in this data set (Fig. \ref{fullgraph}). The
-three communities with the most members contained eight of the 10 most
-abundant MCGs. Comparing these communities with Bruvo's genetic distance
-showed an average distance of 0.451 among communities and an average
-distance of 0.437 within communities, which were not significantly
-different. When we assessed the number of times two different MLHs that
-are in the same MCG, considering these as potential heterothallic
-pairings that could result in sexual recombination, we found an average
-of 14.3 potential heterothallic parings per MLH. Representing just four
-isolates, `MLH 75' had 57 neighbors that shared the same MCG (Fig.
-\ref{biggraph}, \ref{fullgraph}). Overall, there was no clear pattern to
-the association between MLH and MCGs.
-
-\begin{longtable}[]{@{}llll@{}}
-\caption{\label{tab:mlg-table} The five most abundant Multilocus Haplotypes
-(MLH) with the probability of second encounter (\(P_{sex}\)), Mycelial
-Compatibility Groups (MCG), and Regions with sample sizes in
-parentheses.}\tabularnewline
-\toprule
-MLH & \(P_{sex}\) & MCG & Region\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-MLH & \(P_{sex}\) & MCG & Region\tabularnewline
-\midrule
-\endhead
-25 & 0.016824 & 5 & ND (15), CO (2), MI (2)\tabularnewline
-& & 13 & ND (3)\tabularnewline
-& & 60 & ND (2), WA (1)\tabularnewline
-& & 1 & NE (1)\tabularnewline
-& & 4 & MI (1)\tabularnewline
-163 & 0.049932 & 45 & CO (5), ND (2), NE (1)\tabularnewline
-& & 5 & MI (7)\tabularnewline
-65 & 0.000071 & 44 & NE (10)\tabularnewline
-& & 5 & MI (1)\tabularnewline
-140 & 0.000155 & 8 & CO (5)\tabularnewline
-& & 5 & MI (3)\tabularnewline
-& & 20 & MI (2)\tabularnewline
-66 & 0.000016 & 9 & NE (4), CO (2), MI (2)\tabularnewline
-\bottomrule
-\end{longtable}
-
-\subsection*{Structure of shared multilocus
-haplotypes}\label{structure-of-shared-multilocus-haplotypes}
-\addcontentsline{toc}{subsection}{Structure of shared multilocus
-haplotypes}
-
-The most abundant MLH was represented by 27 isolates (Table
-\ref{tab:mlg-table}) from five Regions (NE, MI, WA, CO, and ND). Within
-Regions, haplotypes were relatively evenly distributed with moderate to
-high diversity (Table \ref{tab:iatab}). Of the 165 MLHs, 76 (46\%) were
-found in at least two Regions, except those found in WI (2), ID (1), and
-Mexico (18) (Fig. \ref{community-graph}).
-
-We had performed an analysis on a network where the connections
-represented shared MLHs across populations, weighted by \(1 - P_{sex}\)
-(Fig. \ref{community-graph}, Table \ref{tab:mlg-table}). Community
-analysis of the MLHs shared between populations revealed 4 communities
-with a modularity of 0.17: A coastal community (CA, OR, WA, and NY), a
-Midwest community (CO, ND, NE, MI), and an international community
-(Australia, France, MN). Although analysis with 16 loci resulted in the
-removal of the NY node because it no longer shared a haplotype with OR,
-the same overall community structure was present with a modularity of
-0.2 (Fig. \ref{community-graph16}). Relative to the US, the
-international community appears to be driven by MLH 4, which is shared
-between all three populations and has a \(P_{sex}\) value of
-2.87e\textsuperscript{-5}, in contrast to the abundant MLH 25, which has
-a \(P_{sex}\) value of 0.0168.
-
-\begin{figure}
-\centering
-\includegraphics[width=0.61000\textwidth]{../../results/figures/publication/mlg-11.pdf}
-\caption{Network of populations (nodes/circles) and their shared
-multilocus haplotypes (MLH) (edges/lines) genotyped over 11 loci. Each
-node is labeled with \textbf{name (number of MLHs shared/number of MLHs
-total).} The shade and area of the nodes are proportional to the number
-of unique MLHs within the node and the inner nodes are proportional to
-the number of private MLHs to the region (bottom legend). Each edge
-represents a single MLH where its thickness represents the number of
-populations that share the MLH and the shade represents the value of
-\(P_{sex}\), or the probability of encountering that MLH from two
-independent meiotic events.}\label{community-graph}
-\end{figure}
-
-\subsection*{Population
-differentiation}\label{population-differentiation-1}
-\addcontentsline{toc}{subsection}{Population differentiation}
-
-\subsubsection*{Analysis of molecular
-variance}\label{analysis-of-molecular-variance}
-\addcontentsline{toc}{subsubsection}{Analysis of molecular variance}
-
-The AMOVA for clone-corrected samples over the hierarchy of Region,
-Source, and Year showed significant variation between Regions and Years,
-but no significant variation between wmn and producer fields (Table
-\ref{tab:amova}). In contrast, when we compared the three cultivars,
-Beryl, Bunsi, and G122, we found no significant differentiation (See
-section on `Host Differentiation' in the
-wmn-differentiation.md\footnote{Direct link:
-  \url{https://github.com/everhartlab/sclerotinia-366/blob/master/results/wmn-differentiation.md\#host-differentiation}}
-file in the supplemental files (Kamvar et al., 2017)).
-
-\begin{longtable}[]{@{}clrllc@{}}
-\caption{\label{tab:amova} Comparison of populations in the white mold
-screening nurseries (wmn) and producer fields using an analysis of
-molecular variance (AMOVA) on Bruvo's genetic distance showing no
-apparent differentiation between wmn and other sources. The hierarchy
-was constructed as Source/Region where source is defined as belonging to
-a wmn or producer field. Bold \(\Phi\) values indicate significant
-difference (\emph{P} \textless{} 0.05). S.S. = Sum of Squares, d.f. =
-degrees of freedom.}\tabularnewline
-\toprule
-\begin{minipage}[b]{0.32\columnwidth}\centering\strut
-Hierarchy\strut
-\end{minipage} & \begin{minipage}[b]{0.06\columnwidth}\raggedright\strut
-d.f.\strut
-\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\raggedleft\strut
-S.S.\strut
-\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\raggedright\strut
-\% variation\strut
-\end{minipage} & \begin{minipage}[b]{0.19\columnwidth}\raggedright\strut
-\(\Phi\) statistic\strut
-\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\centering\strut
-\emph{P}\strut
-\end{minipage}\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-\begin{minipage}[b]{0.32\columnwidth}\centering\strut
-Hierarchy\strut
-\end{minipage} & \begin{minipage}[b]{0.06\columnwidth}\raggedright\strut
-d.f.\strut
-\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\raggedleft\strut
-S.S.\strut
-\end{minipage} & \begin{minipage}[b]{0.13\columnwidth}\raggedright\strut
-\% variation\strut
-\end{minipage} & \begin{minipage}[b]{0.19\columnwidth}\raggedright\strut
-\(\Phi\) statistic\strut
-\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\centering\strut
-\emph{P}\strut
-\end{minipage}\tabularnewline
-\midrule
-\endhead
-\begin{minipage}[t]{0.32\columnwidth}\centering\strut
-Between Region\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
-13\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedleft\strut
-10.19\strut
-\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
-8.45\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
-\textbf{0.0845}\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering\strut
-0.031\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.32\columnwidth}\centering\strut
-Between Source within Region\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
-8\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedleft\strut
-2.74\strut
-\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
--2.29\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
--0.0250\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering\strut
-0.497\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.32\columnwidth}\centering\strut
-Between Year within Source\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
-22\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedleft\strut
-9.37\strut
-\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
-16.28\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
-\textbf{0.173}\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering\strut
-0.001\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.32\columnwidth}\centering\strut
-Within Year\strut
-\end{minipage} & \begin{minipage}[t]{0.06\columnwidth}\raggedright\strut
-274\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedleft\strut
-47.30\strut
-\end{minipage} & \begin{minipage}[t]{0.13\columnwidth}\raggedright\strut
-77.56\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedright\strut
-\textbf{0.224}\strut
-\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\centering\strut
-0.001\strut
-\end{minipage}\tabularnewline
-\bottomrule
-\end{longtable}
-
-\newpage
-
-\subsubsection*{Discriminant analysis of principal
-components}\label{discriminant-analysis-of-principal-components}
-\addcontentsline{toc}{subsubsection}{Discriminant analysis of principal
-components}
-
-DAPC was performed by grouping Region with the first 21 principal
-components, representing 88.1\% of the total variance. The first
-discriminant axis (representing 63.9\% of the discriminatory power)
-separated the centroid for the Mexico isolates from the rest of the
-data, indicating strong differentiation (Fig. \ref{DAPC}b). The second
-discriminant axis, representing 10.8\% of the discriminatory power,
-separated the centroid for the CA isolates. The mean population
-assignment probabilities for all populations with n \textgreater{} 10
-showed that only isolates from Mexico, CA, and France had \textgreater{}
-50\% probabilities of being reassigned to their source populations (Fig.
-\ref{DAPC}a).
-
-DAPC grouping by cultivar used the first 20 principal components,
-representing 89\% of the total variance. The first two discriminant axes
-(representing 100\% of the discriminatory power) failed to separate any
-of the cultivars where the mean posterior assignment probabilities were
-34\% (G122), 35.9\% (Beryl), and 30.1\% (Bunsi). DAPC grouping by Region
-and Year used the first 15 principal components, representing 80.3\% of
-the total variance. The North Central USA populations (NE, MI, CO, ND)
-did not appear to have any variation across time in contrast to WA,
-which showed a shift in population structure in the last year of
-sampling, 2008 (Fig. \ref{DAPC-RY}). Further analysis of this population
-revealed that all 12 isolates in WA circa 2008 originated in a wmn; nine
-haplotypes were shared with CA, and three were shared with France (Fig.
-\ref{community-graph}, \ref{community-graph16}).
-
-\begin{figure}
-\centering
-\includegraphics[width=1.00000\textwidth]{../../results/figures/publication/DAPC.pdf}
-\caption{Discriminant Analysis of Principal Components (DAPC) on regions
-showing that Mexico is differentiated from other populations.
-\textbf{A)} Scatter plot of first two components from DAPC. Points
-represent observed individuals connected to the population centroids
-with ellipses representing a 66\% confidence interval for a normal
-distribution. The center of each component is represented as black grid
-lines. \textbf{B)} Mean population assignment probability from the DAPC
-for all populations with N \textgreater{} 10 (facets). Populations
-represented along the horizontal axis and probability of assignment on
-the vertical. Numbers next to source populations indicate population
-size. All values sum to one.}\label{DAPC}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=0.61000\textwidth]{../../results/figures/publication/dapc_region_year_micanewa.pdf}
-\caption{Scatter plot of Discriminant Analysis of Principal Components
-(DAPC) on Regions and Years showing non-differentiated temporal
-variation NE and MI and temporal variation in WA and CA. Points (text
-labels) represent observed individuals connected to the population
-centroids with ellipses representing a 66\% confidence interval for a
-normal distribution. The center of each component is represented as
-black grid lines. A more detailed view is shown in Fig.
-\ref{DAPC-RY-FULL}.}\label{DAPC-RY}
-\end{figure}
-
-\section*{Discussion}\label{discussion}
-\addcontentsline{toc}{section}{Discussion}
-
-In this study, we characterized the diversity of \emph{Sclerotinia
-sclerotiorum} from dry bean fields across the United States. Our results
-suggest that, broadly, populations from white mold screening nurseries
-reflect the populations of the surrounding regions, indicating that
-resistance screening may be successful within regions. We found
-significant population differentiation by geographic region and year,
-mainly differentiated into three broad North American groups based on
-shared haplotypes and posterior groupings, a Coastal Region, Midwestern
-Region, and Mexico. To date, with 366 isolates, this is the largest
-single population genetic study of \emph{S. sclerotiorum} assessing
-population structure within managed and unmanaged agricultural
-environments. These findings indicate that the white mold screening
-nurseries can be effective at screening for potential resistant lines
-within growing regions.
-
-We found that the best predictors of genetic structure are Region and
-Year, supporting the hypothesis that \emph{S. sclerotiorum} populations
-are spatially structured (Carbone \& Kohn, 2001). Borrowing a technique
-often used in the ecological literature, we used dbRDA to elucidate the
-effect of all variables (MCG, Region, Source, Year, Host, and
-Aggressiveness) (Legendre \& Anderson, 1999). From the initial results,
-it appeared that the most important factors for predicting genetic
-structure were MCG, region, and year. When we inspected the biplot of
-the initial results, we saw that the most important predictors were `MCG
-44', `MCG 5', and `MCG 9'. We believe that this was driven by the fact
-that these particular MCGs have uneven MLH distributions, meaning that
-they are heavily associated with one particular MLH (Fig.
-\ref{biggraph}). We note these results with caution because of the
-apparent multicolinearity between MCG and Region, which is a violation
-of the analysis (Legendre \& Anderson, 1999). While the results
-indicated that Mexico and the year 2005 were the two most important
-variables, it's worth noting that all Mexico isolates were collected in
-2005 (Fig. \ref{rda-biplot}). The results also show that the Vista
-cultivar explains some of the variance, but this represents six isolates
-in MI, and thus we cannot draw broad conclusions from this axis.
-Aggressiveness and source field had little to no effect on prediction of
-genetic diversity. These results are in agreement with studies that
-examined differentiation based on Host (Aldrich-Wolfe et al., 2015) and
-Aggressiveness (Atallah et al., 2004; Attanayake et al., 2012, 2013)
-reporting little or no correlation of genetic diversity to these
-variables. This indicates that a) breeders should keep in mind regional
-differences when assessing resistance and b) it is possible that we have
-not yet measured biologically relevant variables that can predict
-genetic differentiation, which could include variables such as soil
-community composition.
-
-While aggressiveness was not shown to predict genetic structure, it is
-an important factor in breeding efforts, and we observed significant
-differences in aggressiveness based on Region (Fig.
-\ref{aggressiveness-boxplot}, Table \ref{tab:region-aggressiveness}).
-These results show a similar pattern to what was found previously in
-Otto-Hanson et al. (2011) with the exception of North Dakota, which
-increased in mean aggressiveness from 5 to 5.77. This increase was due
-in part to new data from producer field isolates collected after the
-previous study. These straw tests were performed by a different person
-for these later isolates, which could suggest a more lenient or strict
-scoring system. However, when we examined the within-region differences,
-we found no significant effect by individual. Many of the ND isolates
-fell within the 6--7 range, which denotes a physical boundary (disease
-symptoms around the second node) between intermediate and susceptible
-(Otto-Hanson et al., 2011). Thus, we observed a shift in aggressiveness
-without a significant shift in genotypic structure, which may indicate
-that aggressiveness may be controlled by environmental factors as
-opposed to genetic profile.
-
-The primary interest of this study was to assess if isolates sampled
-from white mold screening nurseries represent isolates from producer
-fields within the region (Steadman et al., 2003; Otto-Hanson et al.,
-2011). According to our AMOVA results, we have evidence for
-differentiation at the Region and Year, but little to no differentiation
-between wmn isolates and production field isolates (Table
-\ref{tab:amova}). This lack of differentiation, however, may reflect the
-breeder practice of inoculating screening plots with sclerotia collected
-from sources within the region. When we analyze the AMOVA results in
-light of the DAPC results (Fig. \ref{DAPC}), it becomes clear that the
-regional patterns of differentiation are largely driven by isolates from
-Mexico and CA. Isolates from these Regions had a higher posterior
-probability (\textgreater{} 0.75) of being reassigned to their own
-populations than any other (Fig. \ref{DAPC}A). All other populations in
-comparison (except France) has reassignment probabilities of \textless{}
-0.5, which is reflected in the failure of the first two discriminant
-functions to separate these populations (Fig. \ref{DAPC}B).
-
-Despite the evidence that Mexico and CA contributed to much of the
-population differentiation, Regions like WA still had a large amount of
-internal variation. The two distinct clusters for the WA Region showed
-that the 2008 population appeared differentiated and, under further
-investigation, we found that all the haplotypes from this year were
-shared between CA and France (Fig. \ref{community-graph}, \ref{DAPC-RY},
-\ref{DAPC-RY-FULL}). All of the isolates from WA in 2003--2005, and 2008
-came from the same wmn; within the wmn, those in 2003--2005 came a
-Northeastern field location cropped with dry bean since 2002, and those
-in 2008 from a Southeastern field that was previously cropped with
-brassica, sundgrass, peas, beans, and potatoes (Miklas, Phil Pers.
-comm.). Both of these fields were inoculated with sclerotia in 2002, the
-Northeastern field with sclerotia provided by a commercial bean producer
-and the Southeastern field with sclerotia from peas (although this was
-thought to be unsuccessful). Despite this information, it is still
-unclear what has contributed to the differentiation of the 2008
-population from WA or why it shares haplotypes with CA and France. When
-we assessed agressiveness between the two fields across years with an
-ANOVA model, we found that there was a slight effect based on field (P =
-0.0127). While the evidence may suggest host as being a factor, previous
-studies have shown no significant differentiation across host species
-(Aldrich-Wolfe et al., 2015). It was of interest to compare our data
-with that of Aldrich-Wolfe et al. (2015), but we found that, due to
-differences in data generation, we were unable to confidently perform a
-comparison (See supplemental file compare-aldrich-wolfe.md\footnote{Direct
-  link:
-  \url{https://github.com/everhartlab/sclerotinia-366/blob/master/results/compare-aldrich-wolfe.md}}
-(Kamvar et al., 2017)).
-
-With the exception of the WA Region, populations that were sampled
-across several years appeared to be relatively stable over time with
-overlapping distributions in the DAPC (i.e.~NE and MI, Fig.
-\ref{DAPC-RY}). DAPC is based on the principal components of allele
-counts (Jombart et al., 2010). Unlike Bruvo's distance, this does not
-take into account the magnitude of the difference between alleles, which
-could inflate the distance measure in the presence of private alleles
-(Bruvo et al., 2004). While we found no evidence of private alleles in
-the Mexico and CA isolates, we did find that the alleles driving the
-first axis in Fig. \ref{DAPC}A (alleles 174, 256, and 372 in loci 7-2,
-8-3, and 9-2, respectively) were overrepresented in Mexico (where
-\textgreater{}75\% of the alleles came from the region). However, all
-three of these alleles, i) conform to the expected stepwise mutation
-model (Bruvo et al., 2004) and ii) are at or near the extremes of the
-total range (except for allele 372 at locus 9-2). Moreover, the fact
-that we find three alleles at three independent loci segregating the
-Mexican genotypes suggests that the pattern separating these populations
-from the others was not an artifact. We believe that the differences in
-populations observed from Mexico may be due to differences in climate
-that allow greater diversification via sexual outcrossing.
-
-Many of the isolates in our study were from temperate climates and the
-only isolates representing a sub-tropical climate were from Mexico. It
-has been proposed within the \emph{S. sclerotiorum} literature that
-isolates from sub-tropical and tropical climates are differentiated or
-more variable than populations from temperate climates (Carbone \& Kohn,
-2001; Attanayake et al., 2013; Lehner \& Mizubuti, 2017). This has been
-attributed to the notion that the fungus has the chance to undergo more
-reproductive cycles in the warmer climate (Carbone \& Kohn, 2001;
-Attanayake et al., 2013). The strongest evidence to date supporting this
-hypothesis is from Attanayake et al. (2013), showing that populations in
-sub-tropical regions of China have been found to be more variable,
-sexually reproducing, and unrelated to populations in temperate regions
-of the USA. This result however, may be driven more by geography and
-agricultural practice as opposed to climate.
-
-The results from our shared haplotype analysis showed several
-populations with at least one haplotype between them, except for Mexico
-and two states that had fewer than three samples each (Fig.
-\ref{community-graph}). Our network-based approach by treating the
-haplotypes as edges and weighting each edge with the inverse of
-\(P_{sex}\) treated the edges as springs connecting the populations with
-the strength proportional to the probability of obtaining the same
-haplotype as a clone. This allowed us to use a graph walking algorithm
-to see how close the populations were, simply based off of the
-proportion of clones they shared. The most abundant haplotype was shared
-across four populations, but its high value of \(P_{sex}\) meant that it
-did not contribute significantly to the overall structure. The graph
-walking algorithm was able to divide the network into three groups, but
-had a modularity of 0.17, which indicates that the groups are only
-weakly differentiated.
-
-The widespread nature of multilocus haplotypes in both wmn and
-production fields with relatively small values of \(P_{sex}\) may
-indicate the spread of inoculum between regions. While seedborne
-transmission is thought to be of insignificant epidemiological
-importance (Strausbaugh \& Forster, 2003), it has since been shown that
-\emph{S. sclerotiorum} infections can be transmitted through seed
-(Botelho et al., 2013). Thus, we hypothesize that shared haplotypes
-between populations may arise due to transmission events of seed or
-sclerotia. This could explain the fact that we see shared haplotypes
-with low \(P_{sex}\) values shared between Australia, France, and the
-United States. While we speculate that these transmission events are
-rare due to the genetic structuring by Region, these results suggest
-that seedborne infections may indeed reflect a source of inoculum. This
-may, in turn increase the risk of introducing new sources of genetic
-variation through potential outcrossing events.
-
-When we tested for sexual reproduction, we were unable to find evidence
-for it in any region except for Australia and CA. While the Australia
-population had a non-significant value of \(\bar{r}_d\)---which would
-suggest that we cannot reject the null hypothesis of random mating---the
-sample size was insufficient from which to draw conclusions (Milgroom,
-1996; Agapow \& Burt, 2001). The low value of \(\bar{r}_d\) in the CA
-population may represent sexual reproduction, but we can see in Fig.
-\ref{DAPC-RY} that there is differentiation by year. Thus, this could
-also be an artifact of sampling two different populations, which is
-known to reduce the value of \(\bar{r}_d\) (Prugnolle \& de Meeûs,
-2010).
-
-The previous study of the white mold screening nursery populations used
-MCGs to assess genotypic diversity (Otto-Hanson et al., 2011).
-Historically, MCGs have been used as a proxy for clonal lineages, and
-thus, of interest in this study was testing the association between
-multilocus haplotypes (MLHs) and mycelial compatibility groups (MCGs)
-(Kohn et al., 1990; Leslie, 1993; Kohn, 1995; Carbone et al., 1999;
-Schafer \& Kohn, 2006; Otto-Hanson et al., 2011). Our results, however,
-do not support this assumption. It can be seen in Fig. \ref{biggraph}A
-that the most abundant MCG contains several MLHs, but the diversity of
-those MLHs are low as indicated by the evenness (transparency), which
-indicates that there is one dominant MLH (`MLH 25'). What is not shown
-in Fig. \ref{biggraph}A is the MLHs that are shared between MCGs. This
-is illustrated in both Table \ref{tab:mlg-table} and Fig.
-\ref{biggraph}B,C. It could be argued, however that `MLH 25', with its
-high value of \(P_{sex}\) represents different true MLHs across the five
-MCGs it occupies, but this does not account for the overall structure of
-Fig. \ref{fullgraph} where, for example, `MLH 75' (\(P_{sex}\) =
-1.81e\textsuperscript{-4}) is compatible with 57 other haplotypes
-through three MCG when the population structure of \emph{S.
-sclerotiorum} is known to be clonal.
-
-Over the past few years, researchers have noticed inconsistencies among
-the relationship between MCGs and MLHs (Carbone et al., 1999; Attanayake
-et al., 2012; Aldrich-Wolfe et al., 2015; Lehner et al., 2015). Either
-several MCGs belong to one MLH, which could be explained by insufficient
-sampling of loci; several MLHs belong to one MCG, which could be
-explained by clonal expansion; or a mixture of both. Some studies have
-shown a correlation between MCG and MLH (Carbone et al., 1999;
-Aldrich-Wolfe et al., 2015; Lehner et al., 2015), whereas other studies
-have shown no apparent correlation, even on small spatial scales
-(Atallah et al., 2004; Attanayake et al., 2012, 2013).
-
-One long-held assumption was that MCGs (as determined via barrage
-reaction) represent vegetative compatibility groups (VCGs) (Kohn et al.,
-1990; Schafer \& Kohn, 2006; Lehner et al., 2015), which are known to
-have a genetic component (Saupe, 2000; Hall et al., 2010; Strom \&
-Bushley, 2016). While our protocol for assessing MCGs utilized Diana
-Sermons Medium (Cubeta et al., 2001) as compared to Patterson's Medium
-or Potato Dextrose Agar (Schafer \& Kohn, 2006) for the MCG reactions,
-the patterns we observe are not dissimilar from what have previously
-been reported in the literature. It has been demonstrated in several
-Ascomycetes---including \emph{Neurospora crassa} (Micali \& Smith,
-2003), \emph{Sclerotinia homoeocarpa} (Jo et al., 2008),
-\emph{Verticillium dahliae} (Papaioannou \& Typas, 2014), and \emph{S.
-sclerotiorum} (Ford et al., 1995)---that barrage reactions are
-independent from stable anastomosis. Thus, the inconsistencies in this
-study and other studies indicate that researchers studying \emph{S.
-sclerotiorum} should not rely on MCG data derived from barrage reactions
-as an indicator for genetic diversity.
-
-\subsection*{Limitations}\label{limitations}
-\addcontentsline{toc}{subsection}{Limitations}
-
-One of the main limitations of this study is the focus on \emph{P.
-vulgaris} as a host. It has been shown that \emph{S. sclerotiorum} in
-the midwestern United States does not have a particular preference for
-host (Aldrich-Wolfe et al., 2015). If the distribution of \emph{S.
-sclerotiorum} is even across agricultural hosts in the USA, then our
-sample may yet be representative of the genetic pool present in other
-crops and weedy species. Additionally, while we found no signficant
-association between genotype and aggressiveness, it is important to note
-that the straw test is only one measure of aggressiveness. Additional
-phenotypes for aggressiveness should be evaluated for future studies.
-
-Another limitation was the microsatellite markers used for this
-particular study (Sirjusingh \& Kohn, 2001). The haplotype accumulation
-curve showed no indication of a plateau, indicating that if we had
-sampled more loci, we would have resolved more multilocus haplotypes.
-While 16 loci showed us similar results and began to show a plateau for
-the haplotype accumulation curve, we were unable to use these results
-due to our uncertainty in the allele calls for these five extra loci.
-With the availability of an optically-mapped genome (Derbyshire et al.,
-2017), future studies describing the genetic diversity of \emph{S.
-sclerotiorum} should employ techniques such as Genotyping-By-Sequencing
-(Davey et al., 2011), Sequence Capture (Grover et al., 2012), or Whole
-Genome Sequencing.
-
-\subsection*{Conclusions}\label{conclusions}
-\addcontentsline{toc}{subsection}{Conclusions}
-
-This study represents the largest genetic analysis of \emph{S.
-sclerotiorum} from the USA to date, giving us a unique insight to
-continent-wide population structure and relationships between phenotypic
-and genotypic variables. Populations in wmn appear to show no
-significant differentiation when compared to their production field
-counterparts, suggesting that the wmn populations of \emph{S.
-sclerotiorum} may be considered representative of the surrounding
-regions. While we found no direct relationship between haplotype and
-severity, it is evident that there is a gradient of severity by region,
-further supporting the need for screening in multiple locations. Based
-on our analysis of the relationships between MCG and MLH, we found no
-clear evidence that the two are directly related, suggesting that MCG
-does not necessarily represent vegetative compatibility groups and thus
-should not be used as a proxy for identifying clones.
-
-\subsection*{Data Availability}\label{data-availability}
-\addcontentsline{toc}{subsection}{Data Availability}
-
-All scripts, data, and resources used to generate the results presented
-in this publication (including Supplementary Information) are fully
-reproducible and available at The Open Science Framework
-\url{https://osf.io/ejb5y} (Kamvar et al., 2017).
-
-\subsection*{Funding}\label{funding}
-\addcontentsline{toc}{subsection}{Funding}
-
-Funding for this research was provided by the North Central Soybean
-Research Program (\#639K623) to SEE, a Layman Award (\#2446) to SEE,
-USDA-ARS National Sclerotinia Initiative (\#58-5442-2-209) to JRS/SEE,
-and start-up funds from the University of Nebraska-Lincoln to SEE. The
-funders had no role in study design, data collection and analysis,
-decision to publish, or preparation of the manuscript.
-
-\subsection*{Acknowledgements}\label{acknowledgements}
-\addcontentsline{toc}{subsection}{Acknowledgements}
-
-The authors would like to thank Rebecca Higgins for technical support in
-generating the data for the MCG assessment, aggressiveness ratings, and
-genotyping; and for providing valuable insights into the historical
-context of the data collection and curation.
-
-We would also like to thank Denita Hadziabdic and two other anonymous
-reviewers for their valuable comments and insights that improved the
-quality of the manuscript.
-
-\subsection*{Conflicts of Interest}\label{conflicts-of-interest}
-\addcontentsline{toc}{subsection}{Conflicts of Interest}
-
-The authors declare no conflict of interest.
-
-\subsection*{Author Contributions}\label{author-contributions}
-\addcontentsline{toc}{subsection}{Author Contributions}
-
-\begin{itemize}
-\item
-  \textbf{Zhian N. Kamvar} analyzed the data, contributed analysis
-  tools, wrote the paper, prepared figures and tables, edited and
-  reviewed drafts of the paper.
-\item
-  \textbf{B. Sajeewa Amaradasa} analyzed the data, contributed analysis
-  tools, wrote drafts of the paper, edited and reviewed drafts of the
-  paper.
-\item
-  \textbf{Rachana Jhala} Carried out experiments (MCG assessment,
-  aggressiveness ratings, genotyping), edited and reviewed drafts of the
-  paper.
-\item
-  \textbf{Serena McCoy} Carried out experiments (MCG assessment,
-  aggressiveness ratings, genotyping), edited and reviewed drafts of the
-  paper.
-\item
-  \textbf{James R. Steadman} Conceived and designed experiments,
-  organized network of white mold screening nurseries, provided \emph{S.
-  sclerotiorum} isolates, edited and reviewed drafts of the paper.
-\item
-  \textbf{Sydney E. Everhart} supervised data analysis, analyzed the
-  data, contributed analysis tools, wrote the paper, edited and reviewed
-  drafts of the paper.
-\end{itemize}
-
-\newpage
-
-\section*{Supplementary Information}\label{supplementary-information}
-\addcontentsline{toc}{section}{Supplementary Information}
-
-\setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}}
-\setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}
-
-\begin{longtable}[]{@{}lllrlr@{}}
-\caption{\label{tab:isolate-table} Description of \emph{Sclerotinia
-sclerotiorum} isolates used in this study. N = Number of Isolates. Key
-abbreviations: wmn = white mold screening nursery, producer = producer
-field, unk = unknown cultivar.}\tabularnewline
-\toprule
-\begin{minipage}[b]{0.11\columnwidth}\raggedright\strut
-Country\strut
-\end{minipage} & \begin{minipage}[b]{0.08\columnwidth}\raggedright\strut
-State\strut
-\end{minipage} & \begin{minipage}[b]{0.12\columnwidth}\raggedright\strut
-Field Code\strut
-\end{minipage} & \begin{minipage}[b]{0.19\columnwidth}\raggedleft\strut
-Year\strut
-\end{minipage} & \begin{minipage}[b]{0.29\columnwidth}\raggedright\strut
-Host\strut
-\end{minipage} & \begin{minipage}[b]{0.04\columnwidth}\raggedleft\strut
-N\strut
-\end{minipage}\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-\begin{minipage}[b]{0.11\columnwidth}\raggedright\strut
-Country\strut
-\end{minipage} & \begin{minipage}[b]{0.08\columnwidth}\raggedright\strut
-State\strut
-\end{minipage} & \begin{minipage}[b]{0.12\columnwidth}\raggedright\strut
-Field Code\strut
-\end{minipage} & \begin{minipage}[b]{0.19\columnwidth}\raggedleft\strut
-Year\strut
-\end{minipage} & \begin{minipage}[b]{0.29\columnwidth}\raggedright\strut
-Host\strut
-\end{minipage} & \begin{minipage}[b]{0.04\columnwidth}\raggedleft\strut
-N\strut
-\end{minipage}\tabularnewline
-\midrule
-\endhead
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-USA\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-CA\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-wmn\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2004, 2005\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-Beryl, Bunsi, G122\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-18\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-USA\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-CO\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-producer\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2007, 2010\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-Pinto, Yellow\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-41\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-wmn\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2003\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-GH\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-1\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-USA\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-ID\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-producer\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2003\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-GH\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-1\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-USA\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-MI\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-wmn\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2003, 2004, 2005, 2008, 2009\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-11A, 37, 38, B07104, Beryl, Bunsi, Cornell, G122, Orion, PO7863,
-WM31\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-43\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-producer\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2003, 2008, 2009\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-BL, Black, Fuji, GH, Merlot, SR06233, unk, Vista, Zorro\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-19\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-USA\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-MN\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-wmn\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2003, 2004\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-Beryl, Bunsi, G122\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-11\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-USA\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-ND\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-producer\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2007, 2010\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-unk\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-53\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-wmn\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2005\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-Beryl, Bunsi, G122\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-7\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-USA\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-NE\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-wmn\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2004, 2005, 2008, 2010\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-Beryl, Bunsi, G122, PO7683, unk\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-27\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-producer\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2003, 2007, 2009, 2010\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-Beryl, Emerson, GH, Orion, Pinto, Weihing\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-20\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-USA\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-NY\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-producer\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2003\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-GH\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-1\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-USA\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-OR\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-wmn\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2003, 2004\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-Beryl, Bunsi, G122\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-15\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-producer\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2003\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-G122, GH\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-2\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-USA\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-WA\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-wmn\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2003, 2004, 2005, 2008\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-11A, 37, 38, Beryl, Bunsi, Cornell, G122, Orion, PO7 104, PO7863,
-WM31\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-36\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-producer\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2003, 2007\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-GH, Merlot, Pinto, Redkid\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-23\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-USA\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-WI\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-producer\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2003\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-GH\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-2\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-Mexico\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
--\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-wmn\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2005\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-Beryl, Bunsi, G122\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-18\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-France\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
--\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-wmn\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2004, 2005\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-Beryl, Bunsi, G122\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-18\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-producer\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2012\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-unk\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-4\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-Australia\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
--\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-wmn\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2004\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-Beryl, Bunsi, G122\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-4\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-producer\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2004\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-Beryl\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-2\strut
-\end{minipage}\tabularnewline
-\bottomrule
-\end{longtable}
-
-\begin{longtable}[]{@{}cll@{}}
-\caption{\label{tab:region-aggressiveness} Mean aggressiveness ratings for
-Regions with more than five samples; groupings according to 95\%
-family-wise confidence interval.}\tabularnewline
-\toprule
-\begin{minipage}[b]{0.14\columnwidth}\centering\strut
-Region\strut
-\end{minipage} & \begin{minipage}[b]{0.25\columnwidth}\raggedright\strut
-Mean Aggressiveness\strut
-\end{minipage} & \begin{minipage}[b]{0.08\columnwidth}\raggedright\strut
-Group\strut
-\end{minipage}\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-\begin{minipage}[b]{0.14\columnwidth}\centering\strut
-Region\strut
-\end{minipage} & \begin{minipage}[b]{0.25\columnwidth}\raggedright\strut
-Mean Aggressiveness\strut
-\end{minipage} & \begin{minipage}[b]{0.08\columnwidth}\raggedright\strut
-Group\strut
-\end{minipage}\tabularnewline
-\midrule
-\endhead
-\begin{minipage}[t]{0.14\columnwidth}\centering\strut
-MN\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-5.84\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-a\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\centering\strut
-ND\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-5.77\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-a\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\centering\strut
-NE\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-5.29\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-ab\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\centering\strut
-MI\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-5.13\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-abc\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\centering\strut
-OR\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-4.84\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-abcd\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\centering\strut
-CO\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-4.72\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-bcd\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\centering\strut
-WA\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-4.67\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-cd\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\centering\strut
-France\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-4.66\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-cd\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\centering\strut
-Mexico\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-4.58\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-cd\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\centering\strut
-Australia\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-4.12\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-cd\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\centering\strut
-CA\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-4.01\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-d\strut
-\end{minipage}\tabularnewline
-\bottomrule
-\end{longtable}
-
-\begin{longtable}[]{@{}cll@{}}
-\caption{\label{tab:MCG-aggressiveness} Mean aggressiveness ratings for the
-10 most abundant MCG; groupings according to 95\% family-wise confidence
-interval.}\tabularnewline
-\toprule
-\begin{minipage}[b]{0.10\columnwidth}\centering\strut
-MCG\strut
-\end{minipage} & \begin{minipage}[b]{0.25\columnwidth}\raggedright\strut
-Mean Aggressiveness\strut
-\end{minipage} & \begin{minipage}[b]{0.08\columnwidth}\raggedright\strut
-Group\strut
-\end{minipage}\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-\begin{minipage}[b]{0.10\columnwidth}\centering\strut
-MCG\strut
-\end{minipage} & \begin{minipage}[b]{0.25\columnwidth}\raggedright\strut
-Mean Aggressiveness\strut
-\end{minipage} & \begin{minipage}[b]{0.08\columnwidth}\raggedright\strut
-Group\strut
-\end{minipage}\tabularnewline
-\midrule
-\endhead
-\begin{minipage}[t]{0.10\columnwidth}\centering\strut
-44\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-6.03\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-a\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\centering\strut
-3\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-5.50\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-ab\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\centering\strut
-5\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-5.40\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-b\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\centering\strut
-2\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-5.25\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-b\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\centering\strut
-9\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-5.11\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-b\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\centering\strut
-1\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-4.95\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-b\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\centering\strut
-45\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-4.88\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-b\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\centering\strut
-4\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-4.87\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-b\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\centering\strut
-53\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-4.69\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-b\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\centering\strut
-49\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-4.60\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-b\strut
-\end{minipage}\tabularnewline
-\bottomrule
-\end{longtable}
-
-\begin{longtable}[]{@{}cll@{}}
-\caption{\label{tab:MLG-aggressiveness} Mean aggressiveness ratings for the
-10 MLH most abundant; groupings according to 95\% family-wise confidence
-interval.}\tabularnewline
-\toprule
-\begin{minipage}[b]{0.10\columnwidth}\centering\strut
-MLH\strut
-\end{minipage} & \begin{minipage}[b]{0.25\columnwidth}\raggedright\strut
-Mean Aggressiveness\strut
-\end{minipage} & \begin{minipage}[b]{0.08\columnwidth}\raggedright\strut
-Group\strut
-\end{minipage}\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-\begin{minipage}[b]{0.10\columnwidth}\centering\strut
-MLH\strut
-\end{minipage} & \begin{minipage}[b]{0.25\columnwidth}\raggedright\strut
-Mean Aggressiveness\strut
-\end{minipage} & \begin{minipage}[b]{0.08\columnwidth}\raggedright\strut
-Group\strut
-\end{minipage}\tabularnewline
-\midrule
-\endhead
-\begin{minipage}[t]{0.10\columnwidth}\centering\strut
-78\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-6.07\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-a\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\centering\strut
-65\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-5.94\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-a\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\centering\strut
-9\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-5.67\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-ab\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\centering\strut
-25\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-5.41\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-ab\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\centering\strut
-66\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-5.30\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-ab\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\centering\strut
-104\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-5.22\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-ab\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\centering\strut
-160\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-4.80\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-ab\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\centering\strut
-163\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-4.80\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-ab\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\centering\strut
-165\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-4.34\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-b\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\centering\strut
-140\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-4.31\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-b\strut
-\end{minipage}\tabularnewline
-\bottomrule
-\end{longtable}
-
-\begin{figure}
-\centering
-\includegraphics[width=1.00000\textwidth]{images/mcg-trimmed.png}
-\caption{Example of MCG test plates showing (A) a compatible reaction
-with mycelia from two strains overgrowing each other and (B) an
-incompatible reaction with a barrage line of dead tissue forming between
-the two strains. Photo Credit: Rebecca Higgins.}\label{mcg-fig}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=1.00000\textwidth]{../../results/figures/publication/RMM-aggressiveness.pdf}
-\caption{Strip plot of aggressiveness for the eight most abundant MCGs
-partitioned by region. Filled circles indicate one of the five most
-abundant MLHs and open circles indicate a MLH of lesser
-abundance.}\label{mcg-mlh-region-aggressiveness}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=1.00000\textwidth]{../../results/figures/publication/full-graph.pdf}
-\caption{Graph showing complex associations between Mycelial
-Compatibility Groups (MCG) (dotted nodes) and Multilocus Haplotypes
-(MLH) (full nodes) where the number in each node represents the MLH/MCG
-assignment. Node size reflect the number of samples represented by each
-node (circle). Edges (arrows) point from MLH to MCG where the weight
-(thickness) of the edge represents the number of samples shared. Node
-color represents the community assignment based on the walktrap
-algorithm with a maximum of four steps (Pons \& Latapy, 2006). An
-interactive version of this network can be recreated using the code in
-the ``Interactive visualizations'' section of the mlg-mcg.md file in the
-supplementary information (Direct Link:
-\url{https://github.com/everhartlab/sclerotinia-366/blob/master/results/mlg-mcg.md\#interactive-visualizations})
-(Kamvar et al., 2017).}\label{fullgraph}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=1.00000\textwidth]{../../results/figures/publication/mlg-16.pdf}
-\caption{Network of populations (nodes/circles) and their shared
-multilocus haplotypes (MLH) (edges/lines) haplotyped over 16 loci. Each
-node is labeled with \textbf{name (number of MLHs shared/number of MLHs
-total).} The shade and area of the nodes are proportional to the number
-of unique MLHs within the node and the inner nodes are proportional to
-the number of private MLHs to the region (bottom legend). Each edge
-represents a single MLH where its thickness represents the number of
-populations that share the MLH and the shade represents the value of
-\(P_{sex}\), or the probability of encountering that MLH from two
-independent meiotic events.}\label{community-graph16}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=1.00000\textwidth]{../../results/figures/publication/dapc_region_year.pdf}
-\caption{Scatter plot of Discriminant Analysis of Principal Components
-on Regions and Years showing temporal variation across all Regions.
-Points (text labels) represent observed individuals connected to the
-population centroids with ellipses representing a 66\% confidence
-interval for a normal distribution. The center of each component is
-represented as black grid lines.}\label{DAPC-RY-FULL}
-\end{figure}
-
-\newpage
-
-\section*{References}\label{references}
-\addcontentsline{toc}{section}{References}
-
-\hypertarget{refs}{}
-\hypertarget{ref-agapow2001indices}{}
-Agapow, P., \& Burt, A. (2001). Indices of multilocus linkage
-disequilibrium. \emph{Molecular Ecology Notes}, \emph{1}, 101--102.
-
-\hypertarget{ref-aldrich-wolfe2015genetic}{}
-Aldrich-Wolfe, L., Travers, S., \& Nelson, B. D. (2015). Genetic
-variation of \emph{Sclerotinia sclerotiorum} from multiple crops in the
-north central United States. \emph{PLOS ONE}, \emph{10}(9), e0139188.
-\url{https://doi.org/10.1371/journal.pone.0139188}
-
-\hypertarget{ref-arnaud2007standardizing}{}
-Arnaud-Hanod, S., Duarte, C. M., Alberto, F., \& Serrão, E. A. (2007).
-Standardizing methods to address clonality in population studies.
-\emph{Molecular Ecology}, \emph{16}(24), 5115--5139.
-\url{https://doi.org/10.1111/j.1365-294X.2007.03535.x}
-
-\hypertarget{ref-atallah2004high}{}
-Atallah, Z. K., Larget, B., Chen, X., \& Johnson, D. A. (2004). High
-genetic diversity, phenotypic uniformity, and evidence of outcrossing in
-\emph{Sclerotinia sclerotiorum} in the Columbia Basin of Washington
-State. \emph{Phytopathology}, \emph{94}, 737--742.
-
-\hypertarget{ref-attanayake2013sclerotinia}{}
-Attanayake, R. N., Carter, P. A., Jiang, D., del Río-Mendoza, L., \&
-Chen, W. (2013). \emph{Sclerotinia sclerotiorum} populations infecting
-canola from China and the United States are genetically and
-phenotypically distinct. \emph{Phytopathology}, \emph{103}(7), 750--761.
-\url{https://doi.org/10.1094/phyto-07-12-0159-r}
-
-\hypertarget{ref-attanayake2012genetic}{}
-Attanayake, R., Porter, L., Johnson, D., \& Chen, W. (2012). Genetic and
-phenotypic diversity and random association of DNA markers of isolates
-of the fungal plant pathogen \emph{Sclerotinia sclerotiorum} from soil
-on a fine geographic scale. \emph{Soil Biology and Biochemistry},
-\emph{55}, 28--36. \url{https://doi.org/10.1016/j.soilbio.2012.06.002}
-
-\hypertarget{ref-boettiger2017introduction}{}
-Boettiger, C., \& Eddelbuettel, D. (2017). An introduction to rocker:
-Docker containers for R. \emph{CoRR}, \emph{abs/1710.03675}. Retrieved
-from \url{http://arxiv.org/abs/1710.03675}
-
-\hypertarget{ref-boland1994index}{}
-Boland, G., \& Hall, R. (1994). Index of plant hosts of
-\emph{Sclerotinia sclerotiorum}. \emph{Canadian Journal of Plant
-Pathology}, \emph{16}(2), 93--108.
-\url{https://doi.org/10.1080/07060669409500766}
-
-\hypertarget{ref-bolton2006sclerotinia}{}
-Bolton, M. D., Thomma, B. P. H. J., \& Nelson, B. D. (2006).
-\emph{Sclerotinia sclerotiorum} (Lib.) de Bary: Biology and molecular
-traits of a cosmopolitan pathogen. \emph{Molecular Plant Pathology},
-\emph{7}(1), 1--16.
-\url{https://doi.org/10.1111/j.1364-3703.2005.00316.x}
-
-\hypertarget{ref-botelho2013performance}{}
-Botelho, L. d., Zancan, W. L. A., Cruz Machado, J. da, \& Barrocas, E.
-N. (2013). Performance of common bean seeds infected by the fungus
-\emph{Sclerotinia sclerotiorum}. \emph{Journal of Seed Science},
-\emph{35}(2), 153--160.
-\url{https://doi.org/10.1590/s2317-15372013000200003}
-
-\hypertarget{ref-brown1980multilocus}{}
-Brown, A. H. D., Feldman, M. W., \& Nevo, E. (1980). Multilocus
-structure of natural populations of \emph{Hordeum Spontaneum}.
-\emph{Genetics}, \emph{96}(2), 523--536. Retrieved from
-\url{http://www.genetics.org/content/96/2/523}
-
-\hypertarget{ref-bruvo2004simple}{}
-Bruvo, R., Michiels, N. K., D'Souza, T. G., \& Schulenburg, H. (2004). A
-simple method for the calculation of microsatellite genotype distances
-irrespective of ploidy level. \emph{Molecular Ecology}, \emph{13}(7),
-2101--2106.
-
-\hypertarget{ref-carbone2001multilocus}{}
-Carbone, I., \& Kohn, L. M. (2001). Multilocus nested haplotype networks
-extended with DNA fingerprints show common origin and fine-scale,
-ongoing genetic divergence in a wild microbial metapopulation.
-\emph{Molecular Ecology}, \emph{10}(10), 2409--2422.
-\url{https://doi.org/10.1046/j.0962-1083.2001.01380.x}
-
-\hypertarget{ref-carbone1999patterns}{}
-Carbone, I., Anderson, J. B., \& Kohn, L. M. (1999). Patterns of descent
-in clonal lineages and their multilocus fingerprints are resolved with
-combined gene genealogies. \emph{Evolution}, \emph{53}(1), 11--21.
-\url{https://doi.org/10.1111/j.1558-5646.1999.tb05329.x}
-
-\hypertarget{ref-csardi2006igraph}{}
-Csardi, G., \& Nepusz, T. (2006). The igraph software package for
-complex network research. \emph{InterJournal}, \emph{Complex Systems},
-1695. Retrieved from \url{http://igraph.org}
-
-\hypertarget{ref-cubeta1997clonality}{}
-Cubeta, M. A., Cody, B. R., Kohli, Y., \& Kohn, L. M. (1997). Clonality
-in \emph{Sclerotinia sclerotiorum} on infected cabbage in eastern North
-Carolina. \emph{Phytopathology}, \emph{87}, 1000--1004.
-
-\hypertarget{ref-cubeta2001mycelial}{}
-Cubeta, M., Sermons, D., \& Cody, B. (2001). Mycelial interactions of
-\emph{Sclerotinia minor}. \emph{Phytopathology}, \emph{91}(6S), S19.
-\url{https://doi.org/10.1094/phyto.2001.91.6.s1}
-
-\hypertarget{ref-davey2011genome}{}
-Davey, J. W., Hohenlohe, P. A., Etter, P. D., Boone, J. Q., Catchen, J.
-M., \& Blaxter, M. L. (2011). Genome-wide genetic marker discovery and
-genotyping using next-generation sequencing. \emph{Nature Reviews
-Genetics}, \emph{12}(7), 499--510. \url{https://doi.org/10.1038/nrg3012}
-
-\hypertarget{ref-derbyshire2017complete}{}
-Derbyshire, M., Denton-Giles, M., Hegedus, D., Seifbarghy, S., Rollins,
-J., van Kan, J., Seidl, M. F., Faino, L., Mbengue, M., Navaud, O.,
-Raffaele, S., Hammond-Kosack, K., Heard, S., \& Oliver, R. (2017). The
-complete genome sequence of the phytopathogenic fungus \emph{Sclerotinia
-sclerotiorum} reveals insights into the genome architecture of broad
-host range pathogens. \emph{Genome Biology and Evolution}, \emph{9}(3),
-593--618. \url{https://doi.org/10.1093/gbe/evx030}
-
-\hypertarget{ref-ekins2011population}{}
-Ekins, M. G., Hayden, H. L., Aitken, E. A. B., \& Goulter, K. C. (2011).
-Population structure of \emph{Sclerotinia sclerotiorum} on sunflower in
-Australia. \emph{Australasian Plant Pathology}, \emph{40}, 99--108.
-
-\hypertarget{ref-excoffier1992analysis}{}
-Excoffier, L., Smouse, P. E., \& Quattro, J. M. (1992). Analysis of
-molecular variance inferred from metric distances among DNA haplotypes:
-Application to human mitochondrial DNA restriction data.
-\emph{Genetics}, \emph{131}(2), 479--91.
-
-\hypertarget{ref-ford1995heterokaryon}{}
-Ford, E., Miller, R., Gray, H., \& Sherwood, J. (1995). Heterokaryon
-formation and vegetative compatibility in \emph{Sclerotinia
-sclerotiorum}. \emph{Mycological Research}, \emph{99}(2), 241--247.
-
-\hypertarget{ref-grover2012targeted}{}
-Grover, C. E., Salmon, A., \& Wendel, J. F. (2012). Targeted sequence
-capture as a powerful tool for evolutionary analysis. \emph{American
-Journal of Botany}, \emph{99}(2), 312--319.
-\url{https://doi.org/10.3732/ajb.1100323}
-
-\hypertarget{ref-grunwald2003analysis}{}
-Grünwald, N. J., Goodwin, S. B., Milgroom, M. G., \& Fry, W. E. (2003).
-Analysis of genotypic diversity data for populations of microorganisms.
-\emph{Phytopathology}, \emph{93}(6), 738--746.
-\url{https://doi.org/10.1094/phyto.2003.93.6.738}
-
-\hypertarget{ref-hall2010evolution}{}
-Hall, C., Welch, J., Kowbel, D. J., \& Glass, N. L. (2010). Evolution
-and diversity of a fungal self/nonself recognition locus. \emph{PLoS
-ONE}, \emph{5}(11), e14055.
-\url{https://doi.org/10.1371/journal.pone.0014055}
-
-\hypertarget{ref-heck1975explicit}{}
-Heck, K. L., van Belle, G., \& Simberloff, D. (1975). Explicit
-calculation of the rarefaction diversity measurement and the
-determination of sufficient sample size. \emph{Ecology}, \emph{56}(6),
-1459--1461. \url{https://doi.org/10.2307/1934716}
-
-\hypertarget{ref-purrr}{}
-Henry, L., \& Wickham, H. (2017). \emph{purrr: Functional programming
-tools}. Retrieved from \url{https://CRAN.R-project.org/package=purrr}
-
-\hypertarget{ref-hurlbert1971nonconcept}{}
-Hurlbert, S. H. (1971). The nonconcept of species diversity: A critique
-and alternative parameters. \emph{Ecology}, \emph{52}(4), 577--586.
-\url{https://doi.org/10.2307/1934145}
-
-\hypertarget{ref-jo2008reassessment}{}
-Jo, Y.-K., Chang, S. W., Rees, J., \& Jung, G. (2008). Reassessment of
-vegetative compatibility of \emph{Sclerotinia homoeocarpa} using
-nitrate-nonutilizing mutants. \emph{Phytopathology}, \emph{98}(1),
-108--114. \url{https://doi.org/10.1094/phyto-98-1-0108}
-
-\hypertarget{ref-jombart2008adegenet}{}
-Jombart, T. (2008). adegenet: A R package for the multivariate analysis
-of genetic markers. \emph{Bioinformatics}, \emph{24}(11), 1403--1405.
-\url{https://doi.org/10.1093/bioinformatics/btn129}
-
-\hypertarget{ref-jombart2010discriminant}{}
-Jombart, T., Devillard, S., \& Balloux, F. (2010). Discriminant analysis
-of principal components: A new method for the analysis of genetically
-structured populations. \emph{BMC Genetics}, \emph{11:94}.
-\url{https://doi.org/10.1186/1471--2156--11--94}
-
-\hypertarget{ref-kamvar2017data}{}
-Kamvar, Z. N., Amaradasa, B. S., Jhala, R., McCoy, S., Steadman, J., \&
-Everhart, S. E. (2017, November). Data and analysis for population
-structure and phenotypic variation of \emph{Sclerotinia sclerotiorum}
-from dry bean (\emph{Phaseolus vulgaris}) in the United States. Open
-Science Framework. \url{https://doi.org/10.17605/OSF.IO/EJB5Y}
-
-\hypertarget{ref-kamvar2015novel}{}
-Kamvar, Z. N., Brooks, J. C., \& Grünwald, N. J. (2015). Novel R tools
-for analysis of genome-wide population genetic data with emphasis on
-clonality. \emph{Frontiers in Genetics}, \emph{6}.
-\url{https://doi.org/10.3389/fgene.2015.00208}
-
-\hypertarget{ref-kamvar2014poppr}{}
-Kamvar, Z. N., Tabima, J. F., \& Grünwald, N. J. (2014). Poppr: An R
-package for genetic analysis of populations with clonal, partially
-clonal, and/or sexual reproduction. \emph{PeerJ}, \emph{2}, e281.
-\url{https://doi.org/10.7717/peerj.281}
-
-\hypertarget{ref-knodel2012dry}{}
-Knodel, J., Beauzay, P., Franzen, D., Kandel, H., Markell, S., Osorno,
-J., Pasche, J., \& Zollinger, R. (2012). 2012 dry bean grower survey of
-production, pest problems and pesticide use in Minnesota and North
-Dakota. \emph{North Dakota State University Extension}, \emph{E1640}.
-
-\hypertarget{ref-knodel2015dry}{}
-Knodel, J., Beauzay, P., Franzen, D., Kandel, H., Markell, S., Osorno,
-J., Pasche, J., \& Zollinger, R. (2015). 2015 dry bean grower survey of
-production, pest problems and pesticide use in Minnesota and North
-Dakota. \emph{North Dakota State University Extension}, \emph{E1802}.
-
-\hypertarget{ref-knodel2016dry}{}
-Knodel, J., Beauzay, P., Franzen, D., Kandel, H., Markell, S., Osorno,
-J., Pasche, J., \& Zollinger, R. (2016). 2016 dry bean grower survey of
-production, pest problems and pesticide use in Minnesota and North
-Dakota. \emph{North Dakota State University Extension}, \emph{E1841}.
-
-\hypertarget{ref-kohli1998random}{}
-Kohli, Y., \& Kohn, L. M. (1998). Random association among alleles in
-clonal populations of \emph{Sclerotinia sclerotiorum}. \emph{Fungal
-Genetics and Biology}, \emph{23}, 139--149.
-
-\hypertarget{ref-kohli1995clonal}{}
-Kohli, Y., Brunner, L. J., Yoell, H., Milgroom, M. G., Anderson, J. B.,
-Morrall, R. A. A., \& Kohn, L. M. (1995). Clonal dispersal and spatial
-mixing in populations of the plant pathogenic fungus, \emph{Sclerotinia
-sclerotiorum}. \emph{Molecular Ecology}, \emph{4}, 69--77.
-
-\hypertarget{ref-kohn1995clonal}{}
-Kohn, L. M. (1995). The clonal dynamic in wild and agricultural
-plant-pathogen populations. \emph{Canadian Journal of Botany},
-\emph{73}(S1), 1231--1240. \url{https://doi.org/10.1139/b95-383}
-
-\hypertarget{ref-kohn1990mycelial}{}
-Kohn, L. M., Carbone, I., \& Anderson, J. B. (1990). Mycelial
-interactions in \emph{Sclerotinia sclerotiorum}. \emph{Experimental
-Mycology}, \emph{14}, 255--267.
-
-\hypertarget{ref-legendre1999distance}{}
-Legendre, P., \& Anderson, M. J. (1999). Distance-based redundancy
-analysis: Testing multispecies responses in multifactorial ecological
-experiments. \emph{Ecological Monographs}, \emph{69}, 1--24.
-
-\hypertarget{ref-lehner2017sclerotinia}{}
-Lehner, M. S., \& Mizubuti, E. S. G. (2017). Are \emph{Sclerotinia
-sclerotiorum} populations from the tropics more variable than those from
-subtropical and temperate zones? \emph{Tropical Plant Pathology},
-\emph{42}(2), 61--69. \url{https://doi.org/10.1007/s40858-016-0125-1}
-
-\hypertarget{ref-lehner2015genetic}{}
-Lehner, M. S., Júnior, T. J. P., Júnior, B. T. H., Teixeira, H., Vieira,
-R. F., Carneiro, J. E. S., \& Mizubuti, E. S. G. (2015). Low genetic
-variability in \emph{Sclerotinia sclerotiorum} populations from common
-bean fields in Minas Gerais State, Brazil, at regional, local and
-micro-scales. \emph{Plant Pathology}, \emph{64}(4), 921--931.
-\url{https://doi.org/10.1111/ppa.12322}
-
-\hypertarget{ref-lehner2017independently}{}
-Lehner, M. S., Paula Júnior, T. J. de, Del Ponte, E. M., Mizubuti, E.
-S., \& Pethybridge, S. J. (2017). Independently founded populations of
-\emph{Sclerotinia sclerotiorum} from a tropical and a temperate region
-have similar genetic structure. \emph{PloS One}, \emph{12}(3), e0173915.
-\url{https://doi.org/10.1371/journal.pone.0173915}
-
-\hypertarget{ref-leslie1993fungal}{}
-Leslie, J. (1993). Fungal vegetative compatibility. \emph{Annual Review
-of Phytopathology}, \emph{31}, 127--150. Review.
-\url{https://doi.org/10.1146/annurev.py.31.090193.001015}
-
-\hypertarget{ref-mccoy2009use}{}
-McCoy, S., \& Steadman, J. R. (2009). Use of multi-site screening to
-identify partial resistance to white mold in common bean in 2008.
-\emph{Bean Improvement Cooperative Annual Report}, 86--87. Retrieved
-from \url{https://naldc.nal.usda.gov/naldc/catalog.xhtml?id=IND44207142}
-
-\hypertarget{ref-mcdonald2002pathogen}{}
-McDonald, B. A., \& Linde, C. (2002). Pathogen population genetics,
-evolutionary potential, and durable resistance. \emph{Annual Review of
-Phytopathology}, \emph{40}(1), 349--379.
-\url{https://doi.org/10.1146/annurev.phyto.40.120501.101443}
-
-\hypertarget{ref-mendiburu2015agricolae}{}
-Mendiburu, F. D., \& Simon, R. (2015). Agricolae - ten years of an open
-source statistical tool for experiments in breeding, agriculture and
-biology. \url{https://doi.org/10.7287/peerj.preprints.1404v1}
-
-\hypertarget{ref-micali2003independence}{}
-Micali, C. O., \& Smith, M. L. (2003). On the independence of barrage
-formation and heterokaryon incompatibility in \emph{Neurospora crassa}.
-\emph{Fungal Genetics and Biology}, \emph{38}(2), 209--219.
-\url{https://doi.org/10.1016/s1087-1845(02)00533-9}
-
-\hypertarget{ref-milgroom1996recombination}{}
-Milgroom, M. G. (1996). Recombination and the multilocus structure of
-fungal populations. \emph{Annual Review of Phytopathology},
-\emph{34}(1), 457--477.
-
-\hypertarget{ref-nei1978estimation}{}
-Nei, M. (1978). Estimation of average heterozygosity and genetic
-distance from a small number of individuals. \emph{Genetics}, \emph{89},
-583--590.
-
-\hypertarget{ref-vegan}{}
-Oksanen, J., Blanchet, F. G., Friendly, M., Kindt, R., Legendre, P.,
-McGlinn, D., Minchin, P. R., O'Hara, R. B., Simpson, G. L., Solymos, P.,
-Stevens, M. H. H., Szoecs, E., \& Wagner, H. (2017). \emph{Vegan:
-Community ecology package}. Retrieved from
-\url{https://CRAN.R-project.org/package=vegan}
-
-\hypertarget{ref-otto2007identification}{}
-Otto-Hanson, L., \& Steadman, J. R. (2007). Identification of partial
-resistance to \emph{Sclerotinia sclerotiorum} in common bean at multiple
-locations in 2006. \emph{Bean Improvement Cooperative Annual Report},
-133--134. Retrieved from
-\url{https://naldc.nal.usda.gov/naldc/catalog.xhtml?id=IND43940892}
-
-\hypertarget{ref-otto2008identification}{}
-Otto-Hanson, L., \& Steadman, J. R. (2008). Identification of partial
-resistance to \emph{Sclerotinia sclerotiorum} in common bean at multiple
-locations in 2007. \emph{Bean Improvement Cooperative Annual Report},
-214--215. Retrieved from
-\url{https://naldc.nal.usda.gov/naldc/catalog.xhtml?id=IND44063230}
-
-\hypertarget{ref-otto-hanson2011variation}{}
-Otto-Hanson, L., Steadman, J. R., Higgins, R., \& Eskridge, K. M.
-(2011). Variation in \emph{Sclerotinia sclerotiorum} bean isolates from
-multisite resistance screening locations. \emph{Plant Disease},
-\emph{95}(11), 1370--1377. \url{https://doi.org/10.1094/pdis-11-10-0865}
-
-\hypertarget{ref-papaioannou2014barrage}{}
-Papaioannou, I. A., \& Typas, M. A. (2014). Barrage formation is
-independent from heterokaryon incompatibility in \emph{Verticillium
-dahliae}. \emph{European Journal of Plant Pathology}, \emph{141}(1),
-71--82. \url{https://doi.org/10.1007/s10658-014-0525-3}
-
-\hypertarget{ref-parks1993study}{}
-Parks, J. C., \& Werth, C. R. (1993). A study of spatial features of
-clones in a population of bracken fern, \emph{Pteridium aquilinum}
-(Dennstaedtiaceae). \emph{American Journal of Botany}, \emph{80}(5),
-537. \url{https://doi.org/10.2307/2445369}
-
-\hypertarget{ref-ggraph}{}
-Pedersen, T. L. (2017). \emph{ggraph: An implementation of grammar of
-graphics for graphs and networks}. Retrieved from
-\url{https://CRAN.R-project.org/package=ggraph}
-
-\hypertarget{ref-petzoldt1996straw}{}
-Petzoldt, R., \& Dickson, M. H. (1996). Straw test for resistance to
-white mold in beans. \emph{Bean Improvement Cooperative Annual Report},
-\emph{39}, 142--143. Retrieved from
-\url{https://naldc.nal.usda.gov/naldc/catalog.xhtml?id=IND20562675}
-
-\hypertarget{ref-pielou1975ecological}{}
-Pielou, E. (1975). \emph{Ecological Diversity}. New York: Wiley \& Sons.
-
-\hypertarget{ref-pons2006computing}{}
-Pons, P., \& Latapy, M. (2006). Computing communities in large networks
-using random walks. \emph{Journal of Graph Algorithms and Applications},
-\emph{10}(2), 191--218. \url{https://doi.org/10.7155/jgaa.00124}
-
-\hypertarget{ref-prugnolle2010apparent}{}
-Prugnolle, F., \& de Meeûs, T. (2010). Apparent high recombination rates
-in clonal parasitic organisms due to inappropriate sampling design.
-\emph{Heredity}, \emph{104}(2), 135--140.
-\url{https://doi.org/10.1038/hdy.2009.128}
-
-\hypertarget{ref-R}{}
-R Core Team. (2017). \emph{R: A language and environment for statistical
-computing}. Vienna, Austria: R Foundation for Statistical Computing.
-Retrieved from \url{https://www.R-project.org/}
-
-\hypertarget{ref-ramasubramaniam2008estimates}{}
-Ramasubramaniam, H., del Río Mendoza, L. E., \& Bradley, C. A. (2008).
-Estimates of yield and economic losses associated with white mold of
-rain-fed dry bean in North Dakota. \emph{Agronomy Journal},
-\emph{100}(2), 315. \url{https://doi.org/10.2134/agronj2007.0127}
-
-\hypertarget{ref-sambrook1989molecular}{}
-Sambrook, J., Fritsch, E. F., Maniatis, T., \& others. (1989).
-\emph{Molecular Cloning: A Laboratory Manual.} Cold spring harbor
-laboratory press.
-
-\hypertarget{ref-saupe2000molecular}{}
-Saupe, S. J. (2000). Molecular genetics of heterokaryon incompatibility
-in filamentous ascomycetes. \emph{Microbiology and Molecular Biology
-Reviews}, \emph{64}(3), 489--502.
-\url{https://doi.org/10.1128/mmbr.64.3.489-502.2000}
-
-\hypertarget{ref-schafer2006optimized}{}
-Schafer, M. R., \& Kohn, L. M. (2006). An optimized method for mycelial
-compatibility testing in \emph{Sclerotinia sclerotiorum}.
-\emph{Mycologia}, \emph{98}(4), 593--597.
-\url{https://doi.org/10.1080/15572536.2006.11832662}
-
-\hypertarget{ref-sexton2004microsatellite}{}
-Sexton, A. C., \& Howlett, B. J. (2004). Microsatellite markers reveal
-genetic differentiation among populations of \emph{Sclerotinia
-sclerotiorum} from Australian canola fields. \emph{Current Genetics},
-\emph{46}(6), 357--365. \url{https://doi.org/10.1007/s00294-004-0543-3}
-
-\hypertarget{ref-sexton2006population}{}
-Sexton, A. C., Whitten, A. R., \& Howlett, B. J. (2006). Population
-structure of \emph{Sclerotinia sclerotiorum} in an Australian canola
-field at flowering and stem-infection stages of the disease cycle.
-\emph{Genome}, \emph{49}(11), 1408--1415.
-\url{https://doi.org/10.1139/g06-101}
-
-\hypertarget{ref-shannon2001mathematical}{}
-Shannon, C. E. (1948). A mathematical theory of communication. \emph{ACM
-SIGMOBILE Mobile Computing and Communications Review}, \emph{5}(1),
-3--55.
-
-\hypertarget{ref-simpson1949measurement}{}
-Simpson, E. H. (1949). Measurement of diversity. \emph{Nature},
-\emph{163}(4148), 688--688. \url{https://doi.org/10.1038/163688a0}
-
-\hypertarget{ref-sirjusingh2001characterisation}{}
-Sirjusingh, C., \& Kohn, L. M. (2001). Characterisation of
-microsatellites in the fungal plant pathogen, \emph{Sclerotinia
-sclerotiorum}. \emph{Molecular Ecology Notes}, \emph{1}(4), 267--269.
-\url{https://doi.org/10.1046/j.1471-8278.2001.00102.x}
-
-\hypertarget{ref-smith1993how}{}
-Smith, J. M., Smith, N. H., O'Rourke, M., \& Spratt, B. G. (1993). How
-clonal are bacteria? \emph{Proceedings of the National Academy of
-Sciences}, \emph{90}(10), 4384--4388.
-\url{https://doi.org/10.1073/pnas.90.10.4384}
-
-\hypertarget{ref-steadman1983white}{}
-Steadman, J. R. (1983). White mold - a serious yield-limiting disease of
-bean. \emph{Plant Disease}, \emph{67}, 346--350.
-
-\hypertarget{ref-steadman2003identification}{}
-Steadman, J. R., Eskridge, K., \& Powers, K. (2003). Identification of
-partial resistance to \emph{Sclerotinia sclerotiorum} in common bean at
-multiple locations. \emph{Bean Improvement Cooperative Annual Report},
-225--226. Retrieved from
-\url{https://naldc.nal.usda.gov/naldc/catalog.xhtml?id=IND43757287}
-
-\hypertarget{ref-steadman2006identification}{}
-Steadman, J. R., Otto-Hanson, L., \& Breathnach, J. (2006).
-Identification of partial resistance to \emph{Sclerotinia sclerotiorum}
-in common bean at multiple locations in 2005. \emph{Bean Improvement
-Cooperative Annual Report}, 223--224. Retrieved from
-\url{https://naldc.nal.usda.gov/naldc/catalog.xhtml?id=IND43805570}
-
-\hypertarget{ref-steadman2004identification}{}
-Steadman, J. R., Otto-Hanson, L., \& Powers, K. (2004). Identification
-of partial resistance to \emph{Sclerotinia sclerotiorum} in common bean
-at multiple locations. \emph{Bean Improvement Cooperative Annual
-Report}, \emph{47}, 281--282. Retrieved from
-\url{https://naldc.nal.usda.gov/naldc/catalog.xhtml?id=IND43758354}
-
-\hypertarget{ref-steadman2005identification}{}
-Steadman, J. R., Otto-Hanson, L., \& Powers, K. (2005). Identification
-of partial resistance to \emph{Sclerotinia sclerotiorum} in common bean
-at multiple locations in 2004. \emph{Bean Improvement Cooperative Annual
-Report}, 124--125. Retrieved from
-\url{https://naldc.nal.usda.gov/naldc/catalog.xhtml?id=IND43759243}
-
-\hypertarget{ref-stoddart1988genotypic}{}
-Stoddart, J. A., \& Taylor, J. F. (1988). Genotypic diversity:
-Estimation and prediction in samples. \emph{Genetics}, \emph{118}(4),
-705--11.
-
-\hypertarget{ref-strausbaugh2003management}{}
-Strausbaugh, C., \& Forster, R. (2003). Management of white mold of
-beans. \emph{Pacific Northwest Extension}, \emph{PNW568}. Retrieved from
-\url{http://http://www.extension.uidaho.edu/publishing/pdf/PNW/PNW0568.pdf}
-
-\hypertarget{ref-strom2016genomes}{}
-Strom, N. B., \& Bushley, K. E. (2016). Two genomes are better than one:
-History, genetics, and biotechnological applications of fungal
-heterokaryons. \emph{Fungal Biology and Biotechnology}, \emph{3}(1).
-\url{https://doi.org/10.1186/s40694-016-0022-x}
-
-\hypertarget{ref-teran2006modified}{}
-Teran, H., Lema, M., Schwartz, H. F., Duncan, R., Gilbeitson, R., \&
-Singh, S. P. (2006). Modified Petzoldt and Dickson scale for white mold
-rating of common bean. \emph{Bean Improvement Cooperative Annual
-Report}, \emph{49}, 115--116. Retrieved from
-\url{https://naldc.nal.usda.gov/naldc/catalog.xhtml?id=IND43805401}
-
-\hypertarget{ref-tu1982tolerance}{}
-Tu, J. C., \& Beversdorf, W. D. (1982). Tolerance to white mold
-(\emph{Sclerotinia sclerotiorum} (Lib.) De Bary) in Ex Rico 23, a
-cultivar of white bean (\emph{Phaseolus vulgaris} L.). \emph{Canadian
-Journal of Plant Science}, \emph{62}(1), 65--69.
-\url{https://doi.org/10.4141/cjps82-010}
-
-\hypertarget{ref-ggplot2}{}
-Wickham, H. (2009). \emph{ggplot2: Elegant graphics for data analysis}.
-Springer-Verlag New York. Retrieved from \url{http://ggplot2.org}
-
-\hypertarget{ref-dplyr}{}
-Wickham, H., Francois, R., Henry, L., \& Müller, K. (2017). \emph{dplyr:
-A grammar of data manipulation}. Retrieved from
-\url{https://CRAN.R-project.org/package=dplyr}
-
+\input{supplementary}
 \end{document}

### 2017-11-17 doc/manuscript/paper.tex
run peerj-nitpick.sh

--- 2017-11-15 doc/manuscript/manuscript.tex
@@
-%% Submissions for peer-review must enable line-numbering
-%% using the lineno option in the \documentclass command.
-%%
-%% Preprints and camera-ready submissions do not need
-%% line numbers, and should have this option removed.
-%%
-%% Please note that the line numbering option requires
-%% version 1.1 or newer of the wlpeerj.cls file, and
-%% the corresponding author info requires v1.2
-
-\documentclass[fleqn,10pt,lineno]{wlpeerj} % for journal submissions
-
-% ZNK -- Adding headers for pandoc
-
-\setlength{\emergencystretch}{3em}
-\providecommand{\tightlist}{
-\setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-\usepackage{lipsum}
-\usepackage[unicode=true]{hyperref}
-\usepackage{longtable}
-
-\usepackage{indentfirst}
-
-\title{Population structure and phenotypic variation of \emph{Sclerotinia
-sclerotiorum} from dry bean (\emph{Phaseolus vulgaris}) in the United
-States}
-
-\author[1]{Zhian N. Kamvar}
-
-\author[1]{Bimal Sajeewa Amaradasa}
-
-\author[1]{Rachana Jhala}
-
-\author[1]{Serena McCoy}
-
-\author[1]{James R. Steadman}
-
-\author[1]{Sydney E. Everhart}
-
-\corrauthor[1]{Sydney E. Everhart}{\href{mailto:everhart@unl.edu}{\nolinkurl{everhart@unl.edu}}}
-
-\affil[1]{Department of Plant Pathology, University of Nebraska-Lincoln, Lincoln,
-NE 68583}
-
-%
-% \author[1]{First Author}
-% \author[2]{Second Author}
-% \affil[1]{Address of first author}
-% \affil[2]{Address of second author}
-% \corrauthor[1]{First Author}{f.author@email.com}
-
-%
-
-\begin{abstract}
-The ascomycete pathogen \emph{Sclerotinia sclerotiorum} is a
-necrotrophic pathogen on over 400 known host plants, and is the causal
-agent of white mold on dry bean. Currently, there are no known cultivars
-of dry bean with complete resistance to white mold. For more than 20
-years, bean breeders have been using white mold screening nurseries with
-natural populations of \emph{S. sclerotiorum} to screen new cultivars
-for resistance. It is thus important to know if the genetic diversity in
-populations of \emph{S. sclerotiorum} within these nurseries a) reflect
-the genetic diversity of the populations in the surrounding region and
-b) are stable over time. Furthermore, previous studies have investigated
-the correlation between mycelial compatibility groups (MCG) and
-multilocus haplotypes (MLH), but none have formally tested these
-patterns. We genotyped 366 isolates of \emph{S. sclerotiorum} from
-producer fields and white mold screening nurseries surveyed over 10
-years in 2003--2012 representing 11 states in the United States of
-America, Australia, France, and Mexico at 11 microsatellite loci
-resulting in 165 MLHs. Populations were loosely structured over space
-and time based on analysis of molecular variance and discriminant
-analysis of principal components, but not by cultivar, aggressiveness,
-or field source. Of all the regions tested, only Mexico (n=18) shared no
-MLHs with any other region. Using a bipartite network-based approach, we
-found no evidence that the MCGs accurately represent MLHs. Our study
-suggests that breeders should continue to test dry bean lines in several
-white mold screening nurseries across the US to account for both the
-phenotypic and genotypic variation that exists across regions.
-% Dummy abstract text.
-Dummy abstract text.
-Dummy abstract text.
-Dummy abstract text.
-Dummy abstract text.
-Dummy abstract text.
-Dummy abstract text.
-Dummy abstract text.
-Dummy abstract text.
-Dummy abstract text.
-Dummy abstract text.
-\end{abstract}
-
-\usepackage{amsthm}
-\newtheorem{theorem}{Theorem}[section]
-\newtheorem{lemma}{Lemma}[section]
-\theoremstyle{definition}
-\newtheorem{definition}{Definition}[section]
-\newtheorem{corollary}{Corollary}[section]
-\newtheorem{proposition}{Proposition}[section]
-\theoremstyle{definition}
-\newtheorem{example}{Example}[section]
-\theoremstyle{definition}
-\newtheorem{exercise}{Exercise}[section]
-\theoremstyle{remark}
-\newtheorem*{remark}{Remark}
-\newtheorem*{solution}{Solution}
-\begin{document}
-
-\flushbottom
-\maketitle
-\thispagestyle{empty}
-
 \section*{Introduction}\label{introduction}
@@
 \centering
-\includegraphics[width=1.00000\textwidth]{../../results/figures/publication/mcg-subgraph-with-context.pdf}
+\includegraphics[width=1.00000\textwidth]{../../results/figures/publication/Fig1.pdf}
 \caption{Associations between Mycelial Compatibility Groups and
@@
 \centering
-\includegraphics[width=0.61000\textwidth]{../../results/figures/publication/Figure7Z.pdf}
+\includegraphics[width=0.61000\textwidth]{../../results/figures/publication/Fig2.pdf}
 \caption{Biplot showing five most influential explanatory variables
@@
 \centering
-\includegraphics{../../results/figures/publication/aggressiveness.pdf}
+\includegraphics{../../results/figures/publication/Fig3.pdf}
 \caption{Strip plot of aggressiveness by population arranged in
@@
 \centering
-\includegraphics[width=0.61000\textwidth]{../../results/figures/publication/mlg-11.pdf}
+\includegraphics[width=0.61000\textwidth]{../../results/figures/publication/Fig4.pdf}
 \caption{Network of populations (nodes/circles) and their shared
@@
 \centering
-\includegraphics[width=1.00000\textwidth]{../../results/figures/publication/DAPC.pdf}
+\includegraphics[width=1.00000\textwidth]{../../results/figures/publication/Fig5.pdf}
 \caption{Discriminant Analysis of Principal Components (DAPC) on regions
@@
 \centering
-\includegraphics[width=0.61000\textwidth]{../../results/figures/publication/dapc_region_year_micanewa.pdf}
+\includegraphics[width=0.61000\textwidth]{../../results/figures/publication/Fig6.pdf}
 \caption{Scatter plot of Discriminant Analysis of Principal Components
@@
 
-\subsection*{Funding}\label{funding}
-\addcontentsline{toc}{subsection}{Funding}
-
-Funding for this research was provided by the North Central Soybean
-Research Program (\#639K623) to SEE, a Layman Award (\#2446) to SEE,
-USDA-ARS National Sclerotinia Initiative (\#58-5442-2-209) to JRS/SEE,
-and start-up funds from the University of Nebraska-Lincoln to SEE. The
-funders had no role in study design, data collection and analysis,
-decision to publish, or preparation of the manuscript.
-
 \subsection*{Acknowledgements}\label{acknowledgements}
@@
 quality of the manuscript.
-
-\subsection*{Conflicts of Interest}\label{conflicts-of-interest}
-\addcontentsline{toc}{subsection}{Conflicts of Interest}
-
-The authors declare no conflict of interest.
-
-\subsection*{Author Contributions}\label{author-contributions}
-\addcontentsline{toc}{subsection}{Author Contributions}
-
-\begin{itemize}
-\item
-  \textbf{Zhian N. Kamvar} analyzed the data, contributed analysis
-  tools, wrote the paper, prepared figures and tables, edited and
-  reviewed drafts of the paper.
-\item
-  \textbf{B. Sajeewa Amaradasa} analyzed the data, contributed analysis
-  tools, wrote drafts of the paper, edited and reviewed drafts of the
-  paper.
-\item
-  \textbf{Rachana Jhala} Carried out experiments (MCG assessment,
-  aggressiveness ratings, genotyping), edited and reviewed drafts of the
-  paper.
-\item
-  \textbf{Serena McCoy} Carried out experiments (MCG assessment,
-  aggressiveness ratings, genotyping), edited and reviewed drafts of the
-  paper.
-\item
-  \textbf{James R. Steadman} Conceived and designed experiments,
-  organized network of white mold screening nurseries, provided \emph{S.
-  sclerotiorum} isolates, edited and reviewed drafts of the paper.
-\item
-  \textbf{Sydney E. Everhart} supervised data analysis, analyzed the
-  data, contributed analysis tools, wrote the paper, edited and reviewed
-  drafts of the paper.
-\end{itemize}
-
-\newpage
-
-\section*{Supplementary Information}\label{supplementary-information}
-\addcontentsline{toc}{section}{Supplementary Information}
-
-\setcounter{table}{0} \renewcommand{\thetable}{S\arabic{table}}
-\setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}
-
-\begin{longtable}[]{@{}lllrlr@{}}
-\caption{\label{tab:isolate-table} Description of \emph{Sclerotinia
-sclerotiorum} isolates used in this study. N = Number of Isolates. Key
-abbreviations: wmn = white mold screening nursery, producer = producer
-field, unk = unknown cultivar.}\tabularnewline
-\toprule
-\begin{minipage}[b]{0.11\columnwidth}\raggedright\strut
-Country\strut
-\end{minipage} & \begin{minipage}[b]{0.08\columnwidth}\raggedright\strut
-State\strut
-\end{minipage} & \begin{minipage}[b]{0.12\columnwidth}\raggedright\strut
-Field Code\strut
-\end{minipage} & \begin{minipage}[b]{0.19\columnwidth}\raggedleft\strut
-Year\strut
-\end{minipage} & \begin{minipage}[b]{0.29\columnwidth}\raggedright\strut
-Host\strut
-\end{minipage} & \begin{minipage}[b]{0.04\columnwidth}\raggedleft\strut
-N\strut
-\end{minipage}\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-\begin{minipage}[b]{0.11\columnwidth}\raggedright\strut
-Country\strut
-\end{minipage} & \begin{minipage}[b]{0.08\columnwidth}\raggedright\strut
-State\strut
-\end{minipage} & \begin{minipage}[b]{0.12\columnwidth}\raggedright\strut
-Field Code\strut
-\end{minipage} & \begin{minipage}[b]{0.19\columnwidth}\raggedleft\strut
-Year\strut
-\end{minipage} & \begin{minipage}[b]{0.29\columnwidth}\raggedright\strut
-Host\strut
-\end{minipage} & \begin{minipage}[b]{0.04\columnwidth}\raggedleft\strut
-N\strut
-\end{minipage}\tabularnewline
-\midrule
-\endhead
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-USA\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-CA\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-wmn\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2004, 2005\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-Beryl, Bunsi, G122\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-18\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-USA\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-CO\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-producer\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2007, 2010\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-Pinto, Yellow\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-41\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-wmn\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2003\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-GH\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-1\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-USA\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-ID\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-producer\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2003\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-GH\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-1\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-USA\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-MI\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-wmn\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2003, 2004, 2005, 2008, 2009\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-11A, 37, 38, B07104, Beryl, Bunsi, Cornell, G122, Orion, PO7863,
-WM31\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-43\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-producer\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2003, 2008, 2009\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-BL, Black, Fuji, GH, Merlot, SR06233, unk, Vista, Zorro\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-19\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-USA\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-MN\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-wmn\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2003, 2004\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-Beryl, Bunsi, G122\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-11\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-USA\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-ND\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-producer\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2007, 2010\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-unk\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-53\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-wmn\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2005\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-Beryl, Bunsi, G122\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-7\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-USA\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-NE\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-wmn\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2004, 2005, 2008, 2010\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-Beryl, Bunsi, G122, PO7683, unk\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-27\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-producer\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2003, 2007, 2009, 2010\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-Beryl, Emerson, GH, Orion, Pinto, Weihing\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-20\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-USA\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-NY\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-producer\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2003\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-GH\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-1\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-USA\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-OR\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-wmn\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2003, 2004\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-Beryl, Bunsi, G122\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-15\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-producer\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2003\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-G122, GH\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-2\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-USA\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-WA\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-wmn\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2003, 2004, 2005, 2008\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-11A, 37, 38, Beryl, Bunsi, Cornell, G122, Orion, PO7 104, PO7863,
-WM31\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-36\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-producer\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2003, 2007\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-GH, Merlot, Pinto, Redkid\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-23\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-USA\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-WI\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-producer\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2003\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-GH\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-2\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-Mexico\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
--\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-wmn\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2005\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-Beryl, Bunsi, G122\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-18\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-France\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
--\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-wmn\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2004, 2005\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-Beryl, Bunsi, G122\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-18\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-producer\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2012\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-unk\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-4\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-Australia\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
--\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-wmn\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2004\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-Beryl, Bunsi, G122\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-4\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.11\columnwidth}\raggedright\strut
-\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-\strut
-\end{minipage} & \begin{minipage}[t]{0.12\columnwidth}\raggedright\strut
-producer\strut
-\end{minipage} & \begin{minipage}[t]{0.19\columnwidth}\raggedleft\strut
-2004\strut
-\end{minipage} & \begin{minipage}[t]{0.29\columnwidth}\raggedright\strut
-Beryl\strut
-\end{minipage} & \begin{minipage}[t]{0.04\columnwidth}\raggedleft\strut
-2\strut
-\end{minipage}\tabularnewline
-\bottomrule
-\end{longtable}
-
-\begin{longtable}[]{@{}cll@{}}
-\caption{\label{tab:region-aggressiveness} Mean aggressiveness ratings for
-Regions with more than five samples; groupings according to 95\%
-family-wise confidence interval.}\tabularnewline
-\toprule
-\begin{minipage}[b]{0.14\columnwidth}\centering\strut
-Region\strut
-\end{minipage} & \begin{minipage}[b]{0.25\columnwidth}\raggedright\strut
-Mean Aggressiveness\strut
-\end{minipage} & \begin{minipage}[b]{0.08\columnwidth}\raggedright\strut
-Group\strut
-\end{minipage}\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-\begin{minipage}[b]{0.14\columnwidth}\centering\strut
-Region\strut
-\end{minipage} & \begin{minipage}[b]{0.25\columnwidth}\raggedright\strut
-Mean Aggressiveness\strut
-\end{minipage} & \begin{minipage}[b]{0.08\columnwidth}\raggedright\strut
-Group\strut
-\end{minipage}\tabularnewline
-\midrule
-\endhead
-\begin{minipage}[t]{0.14\columnwidth}\centering\strut
-MN\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-5.84\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-a\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\centering\strut
-ND\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-5.77\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-a\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\centering\strut
-NE\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-5.29\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-ab\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\centering\strut
-MI\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-5.13\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-abc\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\centering\strut
-OR\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-4.84\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-abcd\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\centering\strut
-CO\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-4.72\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-bcd\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\centering\strut
-WA\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-4.67\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-cd\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\centering\strut
-France\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-4.66\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-cd\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\centering\strut
-Mexico\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-4.58\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-cd\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\centering\strut
-Australia\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-4.12\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-cd\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.14\columnwidth}\centering\strut
-CA\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-4.01\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-d\strut
-\end{minipage}\tabularnewline
-\bottomrule
-\end{longtable}
-
-\begin{longtable}[]{@{}cll@{}}
-\caption{\label{tab:MCG-aggressiveness} Mean aggressiveness ratings for the
-10 most abundant MCG; groupings according to 95\% family-wise confidence
-interval.}\tabularnewline
-\toprule
-\begin{minipage}[b]{0.10\columnwidth}\centering\strut
-MCG\strut
-\end{minipage} & \begin{minipage}[b]{0.25\columnwidth}\raggedright\strut
-Mean Aggressiveness\strut
-\end{minipage} & \begin{minipage}[b]{0.08\columnwidth}\raggedright\strut
-Group\strut
-\end{minipage}\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-\begin{minipage}[b]{0.10\columnwidth}\centering\strut
-MCG\strut
-\end{minipage} & \begin{minipage}[b]{0.25\columnwidth}\raggedright\strut
-Mean Aggressiveness\strut
-\end{minipage} & \begin{minipage}[b]{0.08\columnwidth}\raggedright\strut
-Group\strut
-\end{minipage}\tabularnewline
-\midrule
-\endhead
-\begin{minipage}[t]{0.10\columnwidth}\centering\strut
-44\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-6.03\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-a\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\centering\strut
-3\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-5.50\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-ab\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\centering\strut
-5\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-5.40\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-b\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\centering\strut
-2\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-5.25\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-b\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\centering\strut
-9\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-5.11\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-b\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\centering\strut
-1\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-4.95\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-b\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\centering\strut
-45\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-4.88\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-b\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\centering\strut
-4\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-4.87\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-b\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\centering\strut
-53\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-4.69\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-b\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\centering\strut
-49\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-4.60\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-b\strut
-\end{minipage}\tabularnewline
-\bottomrule
-\end{longtable}
-
-\begin{longtable}[]{@{}cll@{}}
-\caption{\label{tab:MLG-aggressiveness} Mean aggressiveness ratings for the
-10 MLH most abundant; groupings according to 95\% family-wise confidence
-interval.}\tabularnewline
-\toprule
-\begin{minipage}[b]{0.10\columnwidth}\centering\strut
-MLH\strut
-\end{minipage} & \begin{minipage}[b]{0.25\columnwidth}\raggedright\strut
-Mean Aggressiveness\strut
-\end{minipage} & \begin{minipage}[b]{0.08\columnwidth}\raggedright\strut
-Group\strut
-\end{minipage}\tabularnewline
-\midrule
-\endfirsthead
-\toprule
-\begin{minipage}[b]{0.10\columnwidth}\centering\strut
-MLH\strut
-\end{minipage} & \begin{minipage}[b]{0.25\columnwidth}\raggedright\strut
-Mean Aggressiveness\strut
-\end{minipage} & \begin{minipage}[b]{0.08\columnwidth}\raggedright\strut
-Group\strut
-\end{minipage}\tabularnewline
-\midrule
-\endhead
-\begin{minipage}[t]{0.10\columnwidth}\centering\strut
-78\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-6.07\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-a\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\centering\strut
-65\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-5.94\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-a\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\centering\strut
-9\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-5.67\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-ab\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\centering\strut
-25\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-5.41\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-ab\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\centering\strut
-66\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-5.30\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-ab\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\centering\strut
-104\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-5.22\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-ab\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\centering\strut
-160\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-4.80\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-ab\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\centering\strut
-163\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-4.80\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-ab\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\centering\strut
-165\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-4.34\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-b\strut
-\end{minipage}\tabularnewline
-\begin{minipage}[t]{0.10\columnwidth}\centering\strut
-140\strut
-\end{minipage} & \begin{minipage}[t]{0.25\columnwidth}\raggedright\strut
-4.31\strut
-\end{minipage} & \begin{minipage}[t]{0.08\columnwidth}\raggedright\strut
-b\strut
-\end{minipage}\tabularnewline
-\bottomrule
-\end{longtable}
-
-\begin{figure}
-\centering
-\includegraphics[width=1.00000\textwidth]{images/mcg-trimmed.png}
-\caption{Example of MCG test plates showing (A) a compatible reaction
-with mycelia from two strains overgrowing each other and (B) an
-incompatible reaction with a barrage line of dead tissue forming between
-the two strains. Photo Credit: Rebecca Higgins.}\label{mcg-fig}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=1.00000\textwidth]{../../results/figures/publication/RMM-aggressiveness.pdf}
-\caption{Strip plot of aggressiveness for the eight most abundant MCGs
-partitioned by region. Filled circles indicate one of the five most
-abundant MLHs and open circles indicate a MLH of lesser
-abundance.}\label{mcg-mlh-region-aggressiveness}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=1.00000\textwidth]{../../results/figures/publication/full-graph.pdf}
-\caption{Graph showing complex associations between Mycelial
-Compatibility Groups (MCG) (dotted nodes) and Multilocus Haplotypes
-(MLH) (full nodes) where the number in each node represents the MLH/MCG
-assignment. Node size reflect the number of samples represented by each
-node (circle). Edges (arrows) point from MLH to MCG where the weight
-(thickness) of the edge represents the number of samples shared. Node
-color represents the community assignment based on the walktrap
-algorithm with a maximum of four steps (Pons \& Latapy, 2006). An
-interactive version of this network can be recreated using the code in
-the ``Interactive visualizations'' section of the mlg-mcg.md file in the
-supplementary information (Direct Link:
-\url{https://github.com/everhartlab/sclerotinia-366/blob/master/results/mlg-mcg.md\#interactive-visualizations})
-(Kamvar et al., 2017).}\label{fullgraph}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=1.00000\textwidth]{../../results/figures/publication/mlg-16.pdf}
-\caption{Network of populations (nodes/circles) and their shared
-multilocus haplotypes (MLH) (edges/lines) haplotyped over 16 loci. Each
-node is labeled with \textbf{name (number of MLHs shared/number of MLHs
-total).} The shade and area of the nodes are proportional to the number
-of unique MLHs within the node and the inner nodes are proportional to
-the number of private MLHs to the region (bottom legend). Each edge
-represents a single MLH where its thickness represents the number of
-populations that share the MLH and the shade represents the value of
-\(P_{sex}\), or the probability of encountering that MLH from two
-independent meiotic events.}\label{community-graph16}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=1.00000\textwidth]{../../results/figures/publication/dapc_region_year.pdf}
-\caption{Scatter plot of Discriminant Analysis of Principal Components
-on Regions and Years showing temporal variation across all Regions.
-Points (text labels) represent observed individuals connected to the
-population centroids with ellipses representing a 66\% confidence
-interval for a normal distribution. The center of each component is
-represented as black grid lines.}\label{DAPC-RY-FULL}
-\end{figure}
 
@@
 \hypertarget{ref-kamvar2017data}{}
-Kamvar, Z. N., Amaradasa, B. S., Jhala, R., McCoy, S., Steadman, J., \&
-Everhart, S. E. (2017, November). Data and analysis for population
+Kamvar, Z. N., Amaradasa, B. S., Jhala, R., McCoy, S., Steadman, J. R.,
+\& Everhart, S. E. (2017, November). Data and analysis for population
 structure and phenotypic variation of \emph{Sclerotinia sclerotiorum}
 from dry bean (\emph{Phaseolus vulgaris}) in the United States. Open
-Science Framework. \url{https://doi.org/10.17605/OSF.IO/EJB5Y}
+Science Framework. \url{https://doi.org/10.17605/OSF.IO/K8WTM}
 
@@
 \url{https://CRAN.R-project.org/package=dplyr}
-
-\end{document}

### 2017-11-18 doc/RMD/by-year.Rmd
update Figure 6 to PeerJ's standards

@@
   theme(legend.key = element_rect(fill = "grey95"))
-gg_region_year_micanewa
+# 2017-11-18
+# PeerJ has asked for me to label this figure with A-D for silly reasons
+gg_region_year_mincanewa_lab <- gg_region_year_micanewa %+%
+  mutate(gg_region_year_micanewa$data, 
+         Region = case_when(
+           Region == "NE" ~ "A (NE)",
+           Region == "MI" ~ "B (MI)",
+           Region == "WA" ~ "C (WA)",
+           Region == "CA" ~ "D (CA)"
+         ))
+gg_region_year_mincanewa_lab
 if (!interactive()) {
     ggsave(filename = file.path(PROJHOME, "results/figures/publication/dapc_region_year_micanewa.pdf"),
-       plot = gg_region_year_micanewa,
+       plot = gg_region_year_mincanewa_lab,
        width = 88,

### 2017-11-18 README.Rmd
update README

@@
 > Kamvar ZN, Amaradasa BS, Jhala R, McCoy S, Steadman JR, Everhart SE.
-(2017) Population structure and phenotypic variation of *Sclerotinia sclerotiorum* from dry bean in the United States.
-*PeerJ Preprints* 5:e3311v1 <https://doi.org/10.7287/peerj.preprints.3311v1>
+(2017) Population structure and phenotypic variation of *Sclerotinia sclerotiorum* from dry bean (*Phaseolus vulgaris*) in the United States.
+*PeerJ Preprints* 5:e3311v2 <https://doi.org/10.7287/peerj.preprints.3311v2>
 
@@
 Data and analysis for population structure and phenotypic variation of *Sclerotinia sclerotiorum* from dry bean (*Phaseolus vulgaris*) in the United States.
-<http://doi.org/10.17605/OSF.IO/EJB5Y>
+<https://doi.org/10.17605/OSF.IO/K8WTM>
 

### 2017-11-28 doc/manuscript/paper.tex
add citations

@@
 project container running R version 3.4.2 (Boettiger \& Eddelbuettel,
-2017; R Core Team, 2017) and are openly available and reproducible at
-\url{https://github.com/everhartlab/sclerotinia-366/}. Of the 16
-microsatellite loci genotyped, five included compound repeats, which
-made it challenging to accurately/confidently bin alleles into fragment
-sizes expected for each locus based on the described repeat motif. Loci
-with compound repeats were removed for the reported statistics. To
-ensure the integrity of the results we additionally processed these loci
-and included them in concurrent analyses. We assessed the power of our
-11 markers by generating a genotype accumulation curve in the R package
-\emph{poppr} version 2.5.0, looking for evidence of saturation, which
-would indicate that loci were sufficiently sampled to adequately
-represent the full set of haplotypes (Arnaud-Hanod et al., 2007; Kamvar
-et al., 2015). To avoid including isolates potentially collected from
-the same plant (which increases the probability of collecting sclerotia
-from the same point of infection more than once), data were
-clone-corrected on a hierarchy of Region/Source/Host/Year---meaning that
-duplicated genotypes were reduced to a single observation when they were
-collected in the same year from the same host cultivar located in the
-same source field (wmn or producer)---for subsequent analysis. We
-assessed haplotype diversity by calculating Stoddart and Taylor's index
-(\emph{G}) (Stoddart \& Taylor, 1988), Shannon's index (\emph{H})
-(Shannon, 1948), Simpson's index (\(\lambda\)) (Simpson, 1949), evenness
-(\(E_5\)), and the expected number of multilocus haplotypes
-(\emph{eMLH}) (Hurlbert, 1971; Heck et al., 1975; Pielou, 1975; Grünwald
-et al., 2003). To assess the potential for random mating, we tested for
-linkage disequilibrium with the index of association, \(I_A\) and its
-standardized version, \(\bar{r}_d\) using 999 permutations (Brown et
-al., 1980; Smith et al., 1993; Agapow \& Burt, 2001). Both haplotype
-diversity and linkage disequilibrium were calculated in \emph{poppr}
-(Kamvar et al., 2014).
+2017; R Core Team, 2017). These analyses were rendered as dynamic
+documents with the R packages \emph{knitr} (version 1.17) and
+\emph{ezknitr} (version 0.6) and are openly available and reproducible
+at \url{https://github.com/everhartlab/sclerotinia-366/} (Attali, 2016;
+Xie, 2017). Of the 16 microsatellite loci genotyped, five included
+compound repeats, which made it challenging to accurately/confidently
+bin alleles into fragment sizes expected for each locus based on the
+described repeat motif. Loci with compound repeats were removed for the
+reported statistics. To ensure the integrity of the results we
+additionally processed these loci and included them in concurrent
+analyses. We assessed the power of our 11 markers by generating a
+genotype accumulation curve in the R package \emph{poppr} version 2.5.0,
+looking for evidence of saturation, which would indicate that loci were
+sufficiently sampled to adequately represent the full set of haplotypes
+(Arnaud-Hanod et al., 2007; Kamvar et al., 2015). To avoid including
+isolates potentially collected from the same plant (which increases the
+probability of collecting sclerotia from the same point of infection
+more than once), data were clone-corrected on a hierarchy of
+Region/Source/Host/Year---meaning that duplicated genotypes were reduced
+to a single observation when they were collected in the same year from
+the same host cultivar located in the same source field (wmn or
+producer)---for subsequent analysis. We assessed haplotype diversity by
+calculating Stoddart and Taylor's index (\emph{G}) (Stoddart \& Taylor,
+1988), Shannon's index (\emph{H}) (Shannon, 1948), Simpson's index
+(\(\lambda\)) (Simpson, 1949), evenness (\(E_5\)), and the expected
+number of multilocus haplotypes (\emph{eMLH}) (Hurlbert, 1971; Heck et
+al., 1975; Pielou, 1975; Grünwald et al., 2003). To assess the potential
+for random mating, we tested for linkage disequilibrium with the index
+of association, \(I_A\) and its standardized version, \(\bar{r}_d\)
+using 999 permutations (Brown et al., 1980; Smith et al., 1993; Agapow
+\& Burt, 2001). Both haplotype diversity and linkage disequilibrium were
+calculated in \emph{poppr} (Kamvar et al., 2014).
 
@@
 \centering
-\includegraphics[width=1.00000\textwidth]{../../results/figures/publication/Fig1.pdf}
+\includegraphics[width=1.00000\textwidth]{../../results/figures/publication/mcg-subgraph-with-context.pdf}
 \caption{Associations between Mycelial Compatibility Groups and
@@
 \centering
-\includegraphics[width=0.61000\textwidth]{../../results/figures/publication/Fig2.pdf}
+\includegraphics[width=0.61000\textwidth]{../../results/figures/publication/Figure7Z.pdf}
 \caption{Biplot showing five most influential explanatory variables
@@
 \centering
-\includegraphics{../../results/figures/publication/Fig3.pdf}
+\includegraphics{../../results/figures/publication/aggressiveness.pdf}
 \caption{Strip plot of aggressiveness by population arranged in
@@
 \centering
-\includegraphics[width=0.61000\textwidth]{../../results/figures/publication/Fig4.pdf}
+\includegraphics[width=0.61000\textwidth]{../../results/figures/publication/mlg-11.pdf}
 \caption{Network of populations (nodes/circles) and their shared
@@
 \centering
-\includegraphics[width=1.00000\textwidth]{../../results/figures/publication/Fig5.pdf}
+\includegraphics[width=1.00000\textwidth]{../../results/figures/publication/DAPC.pdf}
 \caption{Discriminant Analysis of Principal Components (DAPC) on regions
@@
 \centering
-\includegraphics[width=0.61000\textwidth]{../../results/figures/publication/Fig6.pdf}
+\includegraphics[width=0.61000\textwidth]{../../results/figures/publication/dapc_region_year_micanewa.pdf}
 \caption{Scatter plot of Discriminant Analysis of Principal Components
@@
 
+\hypertarget{ref-attali2016ezknitr}{}
+Attali, D. (2016). ezknitr: Avoid the typical working directory pain
+when using ``knitr''. \emph{The Journal of Open Source Software},
+\emph{1}(5), 75. \url{https://doi.org/10.21105/joss.00075}
+
 \hypertarget{ref-attanayake2013sclerotinia}{}
@@
 \url{https://CRAN.R-project.org/package=dplyr}
+
+\hypertarget{ref-xie2017knitr}{}
+Xie, Y. (2017). \emph{knitr: A general-purpose package for dynamic
+report generation in R}. Retrieved from \url{https://yihui.name/knitr/}

### 2017-11-28 doc/manuscript/manuscript.Rmd
Update citations for publication (#21)
* add citations

* update citations

@@
 
-All data processing and analyses were performed in a Rocker "verse" project container running `r R.version.string %>% strsplit("\\(") %>% map_chr(1)`[@boettiger2017introduction; @R] and are openly available and reproducible at https://github.com/everhartlab/sclerotinia-366/.
+All data processing and analyses were performed in a Rocker "verse" project container running `r R.version.string %>% strsplit("\\(") %>% map_chr(1)`[@boettiger2017introduction; @R]. 
+These analyses were rendered as dynamic documents with the R packages *knitr* (version `r packageVersion("knitr")`) and *ezknitr* (version `r packageVersion("ezknitr")`) and are openly available and reproducible at https://github.com/everhartlab/sclerotinia-366/ [@attali2016ezknitr; @xie2017knitr].
 Of the 16 microsatellite loci genotyped, five included compound repeats, which made it challenging to accurately/confidently bin alleles into fragment sizes expected for each locus based on the described repeat motif.

### 2017-11-29 doc/manuscript/manuscript.Rmd
add Sydney's comments

@@
 We assessed the power of our 11 markers by generating a genotype accumulation curve in the R package *poppr* version `r packageVersion("poppr")`, looking for evidence of saturation, which would indicate that loci were sufficiently sampled to adequately represent the full set of haplotypes [@arnaud2007standardizing; @kamvar2015novel].
+We additionally assessed within-locus allelic diversity by measuring Nei's gene diversity (*h*) [@nei1978estimation] and allelic evenness ($E_5$) [@pielou1975ecological; @grunwald2003analysis].
 To avoid including isolates potentially collected from the same plant (which increases the probability of collecting sclerotia from the same point of infection more than once), data were clone-corrected on a hierarchy of Region/Source/Host/Year---meaning that duplicated genotypes were reduced to a single observation when they were collected in the same year from the same host cultivar located in the same source field (wmn or producer)---for subsequent analysis.
-We assessed haplotype diversity by calculating Stoddart and Taylor's index (*G*) [@stoddart1988genotypic], Shannon's index (*H*) [@shannon2001mathematical], Simpson's index ($\lambda$) [@simpson1949measurement], evenness ($E_5$), and the expected number of multilocus haplotypes (*eMLH*) [@heck1975explicit; @hurlbert1971nonconcept; @pielou1975ecological; @grunwald2003analysis].
+We assessed haplotype diversity by calculating Stoddart and Taylor's index (*G*) [@stoddart1988genotypic], Shannon's index (*H*) [@shannon2001mathematical], Simpson's index ($\lambda$) [@simpson1949measurement], evenness ($E_5$), and the expected number of multilocus haplotypes via rarefaction (*eMLH*) with 10 samples [@heck1975explicit; @hurlbert1971nonconcept; @pielou1975ecological; @grunwald2003analysis].
+If all haplotypes in are equally abundant, both *G* and $e^H$ (the exponentiation of *H*) are expected to be equal to the number of haplotypes, $\lambda$ and $E_5$ are expected to equal one, and *eMLH* is expected to be at its maximum value (in this case, 10) [@grunwald2003analysis].
 To assess the potential for random mating, we tested for linkage disequilibrium with the index of association, $I_A$ and its standardized version, $\bar{r}_d$ using 999 permutations [@agapow2001indices; @smith1993how; @brown1980multilocus].
@@
 The number of observed alleles per locus ranged from two to 10 with an average of `r signif(Amean, 3)` (Table \@ref(tab:locus-stats)).
-Locus `r loc_table$Locus[loc_table$allele==2]`, which contained only 2 alleles, showed low values of both *h* (`r signif(loc_table$Hexp[loc_table$allele==2], 3)`) and evenness (`r signif(loc_table$Evenness[loc_table$allele==2], 3)`), indicating that there was one dominant allele present.
+Locus `r loc_table$Locus[loc_table$allele==2]`, which contained only 2 alleles, showed low values of both *h* (`r signif(loc_table$Hexp[loc_table$allele==2], 3)`) and $E_5$ (`r signif(loc_table$Evenness[loc_table$allele==2], 3)`), indicating that there was one dominant allele present.
 Analysis of the haplotype accumulation curve showed no clear plateau for 11 or 16 loci (See section on 'Loading Data and Setting Strata' in the MLG-distribution.md[^1] file in the supplemental files [@kamvar2017data]), indicating that we would likely obtain more multilocus haplotypes if we were to genotype more loci.
@@
 rename(`*h*` = Hexp) %>%
+rename(`$E_5$` = Evenness) %>%
 rename(`No. alleles` = allele) %>%
@@
 Table: (\#tab:locus-stats) Allelic diversity on full data set at loci used in this study.
-*h* = Nei's Gene Diversity [@nei1978estimation].
-Average *h* = `r signif(Hmean, 3)`, average Evenness = `r signif(Emean, 3)`, average no.
+*h* = Nei's 1978 gene diversity, $E_5$ = Evenness.
+Average *h* = `r signif(Hmean, 3)`, average $E_5$ = `r signif(Emean, 3)`, average no.
 alleles = `r signif(Amean, 3)`

### 2017-11-29 doc/manuscript/manuscript.tex
update OSF citation
rerun make

@@
 project container running R version 3.4.2 (Boettiger \& Eddelbuettel,
-2017; R Core Team, 2017) and are openly available and reproducible at
-\url{https://github.com/everhartlab/sclerotinia-366/}. Of the 16
-microsatellite loci genotyped, five included compound repeats, which
-made it challenging to accurately/confidently bin alleles into fragment
-sizes expected for each locus based on the described repeat motif. Loci
-with compound repeats were removed for the reported statistics. To
-ensure the integrity of the results we additionally processed these loci
-and included them in concurrent analyses. We assessed the power of our
-11 markers by generating a genotype accumulation curve in the R package
-\emph{poppr} version 2.5.0, looking for evidence of saturation, which
-would indicate that loci were sufficiently sampled to adequately
-represent the full set of haplotypes (Arnaud-Hanod et al., 2007; Kamvar
-et al., 2015). To avoid including isolates potentially collected from
-the same plant (which increases the probability of collecting sclerotia
-from the same point of infection more than once), data were
-clone-corrected on a hierarchy of Region/Source/Host/Year---meaning that
-duplicated genotypes were reduced to a single observation when they were
-collected in the same year from the same host cultivar located in the
-same source field (wmn or producer)---for subsequent analysis. We
-assessed haplotype diversity by calculating Stoddart and Taylor's index
-(\emph{G}) (Stoddart \& Taylor, 1988), Shannon's index (\emph{H})
-(Shannon, 1948), Simpson's index (\(\lambda\)) (Simpson, 1949), evenness
-(\(E_5\)), and the expected number of multilocus haplotypes
-(\emph{eMLH}) (Hurlbert, 1971; Heck et al., 1975; Pielou, 1975; Grünwald
-et al., 2003). To assess the potential for random mating, we tested for
-linkage disequilibrium with the index of association, \(I_A\) and its
-standardized version, \(\bar{r}_d\) using 999 permutations (Brown et
-al., 1980; Smith et al., 1993; Agapow \& Burt, 2001). Both haplotype
-diversity and linkage disequilibrium were calculated in \emph{poppr}
-(Kamvar et al., 2014).
+2017; R Core Team, 2017). These analyses were rendered as dynamic
+documents with the R packages \emph{knitr} (version 1.17) and
+\emph{ezknitr} (version 0.6) and are openly available and reproducible
+at \url{https://github.com/everhartlab/sclerotinia-366/} (Attali, 2016;
+Xie, 2017). Of the 16 microsatellite loci genotyped, five included
+compound repeats, which made it challenging to accurately/confidently
+bin alleles into fragment sizes expected for each locus based on the
+described repeat motif. Loci with compound repeats were removed for the
+reported statistics. To ensure the integrity of the results we
+additionally processed these loci and included them in concurrent
+analyses. We assessed the power of our 11 markers by generating a
+genotype accumulation curve in the R package \emph{poppr} version 2.5.0,
+looking for evidence of saturation, which would indicate that loci were
+sufficiently sampled to adequately represent the full set of haplotypes
+(Arnaud-Hanod et al., 2007; Kamvar et al., 2015). We additionally
+assessed within-locus allelic diversity by measuring Nei's gene
+diversity (\emph{h}) (Nei, 1978) and allelic evenness (\(E_5\)) (Pielou,
+1975; Grünwald et al., 2003). To avoid including isolates potentially
+collected from the same plant (which increases the probability of
+collecting sclerotia from the same point of infection more than once),
+data were clone-corrected on a hierarchy of
+Region/Source/Host/Year---meaning that duplicated genotypes were reduced
+to a single observation when they were collected in the same year from
+the same host cultivar located in the same source field (wmn or
+producer)---for subsequent analysis. We assessed haplotype diversity by
+calculating Stoddart and Taylor's index (\emph{G}) (Stoddart \& Taylor,
+1988), Shannon's index (\emph{H}) (Shannon, 1948), Simpson's index
+(\(\lambda\)) (Simpson, 1949), evenness (\(E_5\)), and the expected
+number of multilocus haplotypes via rarefaction (\emph{eMLH}) with 10
+samples (Hurlbert, 1971; Heck et al., 1975; Pielou, 1975; Grünwald et
+al., 2003). If all haplotypes in are equally abundant, both \emph{G} and
+\(e^H\) (the exponentiation of \emph{H}) are expected to be equal to the
+number of haplotypes, \(\lambda\) and \(E_5\) are expected to equal one,
+and \emph{eMLH} is expected to be at its maximum value (in this case,
+10) (Grünwald et al., 2003). To assess the potential for random mating,
+we tested for linkage disequilibrium with the index of association,
+\(I_A\) and its standardized version, \(\bar{r}_d\) using 999
+permutations (Brown et al., 1980; Smith et al., 1993; Agapow \& Burt,
+2001). Both haplotype diversity and linkage disequilibrium were
+calculated in \emph{poppr} (Kamvar et al., 2014).
 
@@
 contained only 2 alleles, showed low values of both \emph{h} (0.0533)
-and evenness (0.42), indicating that there was one dominant allele
+and \(E_5\) (0.42), indicating that there was one dominant allele
 present. Analysis of the haplotype accumulation curve showed no clear
@@
 \caption{\label{tab:locus-stats} Allelic diversity on full data set at loci
-used in this study. \emph{h} = Nei's Gene Diversity (Nei, 1978). Average
-\emph{h} = 0.583, average Evenness = 0.693, average no. alleles =
-6.27}\tabularnewline
+used in this study. \emph{h} = Nei's 1978 gene diversity, \(E_5\) =
+Evenness. Average \emph{h} = 0.583, average \(E_5\) = 0.693, average no.
+alleles = 6.27}\tabularnewline
 \toprule
@@
 \emph{h}\strut
-\end{minipage} & \begin{minipage}[b]{0.10\columnwidth}\raggedright\strut
-Evenness\strut
+\end{minipage} & \begin{minipage}[b]{0.07\columnwidth}\raggedright\strut
+\(E_5\)\strut
 \end{minipage}\tabularnewline
@@
 0.45\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
 0.62\strut
@@
 0.64\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
 0.95\strut
@@
 0.73\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
 0.76\strut
@@
 0.74\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
 0.79\strut
@@
 0.35\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
 0.41\strut
@@
 0.58\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
 0.78\strut
@@
 0.55\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
 0.53\strut
@@
 0.05\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
 0.42\strut
@@
 0.72\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
 0.66\strut
@@
 0.76\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
 0.91\strut
@@
 0.83\strut
-\end{minipage} & \begin{minipage}[t]{0.10\columnwidth}\raggedright\strut
+\end{minipage} & \begin{minipage}[t]{0.07\columnwidth}\raggedright\strut
 0.80\strut
@@
 quality of the manuscript.
-
-\subsection*{Conflicts of Interest}\label{conflicts-of-interest}
-\addcontentsline{toc}{subsection}{Conflicts of Interest}
-
-The authors declare no conflict of interest.
-
-\subsection*{Author Contributions}\label{author-contributions}
-\addcontentsline{toc}{subsection}{Author Contributions}
-
-\begin{itemize}
-\item
-  \textbf{Zhian N. Kamvar} analyzed the data, contributed analysis
-  tools, wrote the paper, prepared figures and tables, edited and
-  reviewed drafts of the paper.
-\item
-  \textbf{B. Sajeewa Amaradasa} analyzed the data, contributed analysis
-  tools, wrote drafts of the paper, edited and reviewed drafts of the
-  paper.
-\item
-  \textbf{Rachana Jhala} Carried out experiments (MCG assessment,
-  aggressiveness ratings, genotyping), edited and reviewed drafts of the
-  paper.
-\item
-  \textbf{Serena McCoy} Carried out experiments (MCG assessment,
-  aggressiveness ratings, genotyping), edited and reviewed drafts of the
-  paper.
-\item
-  \textbf{James R. Steadman} Conceived and designed experiments,
-  organized network of white mold screening nurseries, provided \emph{S.
-  sclerotiorum} isolates, edited and reviewed drafts of the paper.
-\item
-  \textbf{Sydney E. Everhart} supervised data analysis, analyzed the
-  data, contributed analysis tools, wrote the paper, edited and reviewed
-  drafts of the paper.
-\end{itemize}
 
@@
 for analysis of genome-wide population genetic data with emphasis on
-clonality. \emph{Frontiers in Genetics}, \emph{6}.
+clonality. \emph{Frontiers in Genetics}, \emph{6}, 208.
 \url{https://doi.org/10.3389/fgene.2015.00208}
@@
 identify partial resistance to white mold in common bean in 2008.
-\emph{Bean Improvement Cooperative Annual Report}, 86--87. Retrieved
-from \url{https://naldc.nal.usda.gov/naldc/catalog.xhtml?id=IND44207142}
+\emph{Bean Improvement Cooperative Annual Report}, \emph{52}, 86--87.
+Retrieved from
+\url{https://naldc.nal.usda.gov/naldc/catalog.xhtml?id=IND44207142}
 
@@
 source statistical tool for experiments in breeding, agriculture and
-biology. \url{https://doi.org/10.7287/peerj.preprints.1404v1}
+biology. \emph{PeerJ PrePrints}, \emph{3}, e1404v1.
+\url{https://doi.org/10.7287/peerj.preprints.1404v1}
 
@@
 locations in 2006. \emph{Bean Improvement Cooperative Annual Report},
-133--134. Retrieved from
+\emph{50}, 133--134. Retrieved from
 \url{https://naldc.nal.usda.gov/naldc/catalog.xhtml?id=IND43940892}
@@
 locations in 2007. \emph{Bean Improvement Cooperative Annual Report},
-214--215. Retrieved from
+\emph{51}, 214--215. Retrieved from
 \url{https://naldc.nal.usda.gov/naldc/catalog.xhtml?id=IND44063230}
@@
 multiple locations. \emph{Bean Improvement Cooperative Annual Report},
-225--226. Retrieved from
+\emph{46}, 225--226. Retrieved from
 \url{https://naldc.nal.usda.gov/naldc/catalog.xhtml?id=IND43757287}
@@
 in common bean at multiple locations in 2005. \emph{Bean Improvement
-Cooperative Annual Report}, 223--224. Retrieved from
+Cooperative Annual Report}, \emph{49}, 223--224. Retrieved from
 \url{https://naldc.nal.usda.gov/naldc/catalog.xhtml?id=IND43805570}
@@
 at multiple locations in 2004. \emph{Bean Improvement Cooperative Annual
-Report}, 124--125. Retrieved from
+Report}, \emph{48}, 124--125. Retrieved from
 \url{https://naldc.nal.usda.gov/naldc/catalog.xhtml?id=IND43759243}
@@
 
+\hypertarget{ref-xie2017knitr}{}
+Xie, Y. (2017). \emph{knitr: A general-purpose package for dynamic
+report generation in R}. Retrieved from \url{https://yihui.name/knitr/}
+
 \end{document}

### 2017-11-29 doc/manuscript/paper.tex
update citations
rerun make

@@
 sufficiently sampled to adequately represent the full set of haplotypes
-(Arnaud-Hanod et al., 2007; Kamvar et al., 2015). To avoid including
-isolates potentially collected from the same plant (which increases the
-probability of collecting sclerotia from the same point of infection
-more than once), data were clone-corrected on a hierarchy of
+(Arnaud-Hanod et al., 2007; Kamvar et al., 2015). We additionally
+assessed within-locus allelic diversity by measuring Nei's gene
+diversity (\emph{h}) (Nei, 1978) and allelic evenness (\(E_5\)) (Pielou,
+1975; Grünwald et al., 2003). To avoid including isolates potentially
+collected from the same plant (which increases the probability of
+collecting sclerotia from the same point of infection more than once),
+data were clone-corrected on a hierarchy of
 Region/Source/Host/Year---meaning that duplicated genotypes were reduced
@@
 (\(\lambda\)) (Simpson, 1949), evenness (\(E_5\)), and the expected
-number of multilocus haplotypes (\emph{eMLH}) (Hurlbert, 1971; Heck et
-al., 1975; Pielou, 1975; Grünwald et al., 2003). To assess the potential
-for random mating, we tested for linkage disequilibrium with the index
-of association, \(I_A\) and its standardized version, \(\bar{r}_d\)
-using 999 permutations (Brown et al., 1980; Smith et al., 1993; Agapow
-\& Burt, 2001). Both haplotype diversity and linkage disequilibrium were
+number of multilocus haplotypes via rarefaction (\emph{eMLH}) with 10
+samples (Hurlbert, 1971; Heck et al., 1975; Pielou, 1975; Grünwald et
+al., 2003). If all haplotypes in are equally abundant, both \emph{G} and
+\(e^H\) (the exponentiation of \emph{H}) are expected to be equal to the
+number of haplotypes, \(\lambda\) and \(E_5\) are expected to equal one,
+and \emph{eMLH} is expected to be at its maximum value (in this case,
+10) (Grünwald et al., 2003). To assess the potential for random mating,
+we tested for linkage disequilibrium with the index of association,
+\(I_A\) and its standardized version, \(\bar{r}_d\) using 999
+permutations (Brown et al., 1980; Smith et al., 1993; Agapow \& Burt,
+2001). Both haplotype diversity and linkage disequilibrium were
 calculated in \emph{poppr} (Kamvar et al., 2014).

### 2018-01-22 README.Rmd
update README with publication information
add note to README; update with latest container
add poster to README.Rmd

@@
 
-## Preprint
+## Publication
 
@@
 (2017) Population structure and phenotypic variation of *Sclerotinia sclerotiorum* from dry bean (*Phaseolus vulgaris*) in the United States.
-*PeerJ Preprints* 5:e3311v2 <https://doi.org/10.7287/peerj.preprints.3311v2>
+*PeerJ* 5:e4152 <https://doi.org/10.7717/peerj.4152>
 
-## Data and Code
+> Poster presented at the 2018 National Sclerotinia Initiative: <https://doi.org/10.6084/m9.figshare.5791713>
+
+Data and Code
+-------------
 
@@
 <https://doi.org/10.17605/OSF.IO/K8WTM>
+
+Note: This citation represents the data and code that was in this repository at the time of re-submission (2017-11-15 20:42 UTC), For the docker container, this represents tag v1.1.
+You can find the live code at <https://osf.io/ejb5y>
 
